### 2019-07-31 Analyses/long_ts.Rmd
First_commit

---
title: "long_ts"
author: "Amieroh Abrahams"
date: "31 July 2019"
output: html_document
---

# Detectig frequency and intensity of upwelling events - Use MHW algorithm 
  Long time series of 30years
  Use OISST 
  Match this up with MCS occuring in the region
  Paper 1 (Local paper)
  Upwelling: 30th percentiles within EBUS

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE 
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(data.table)
library(heatwaveR)
library(viridis)
library(gridExtra)
library(circular)
library(fossil)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ggpubr)
library(ggpubr); theme_set(theme_pubr())# library(devtools)

```

# Loading the data and finding a 30year time series

```{r}
load("~/Documents/Masters_2019/Upwelling_MCS/Data/SACTN_daily_v4.2.RData")
load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_v4.2.RData")

site_list_sub <- site_list %>%
  filter(coast == "wc") %>%
  filter(length > 10950)

SACTN_US <- SACTN_daily_v4.2 %>%
  left_join(site_list[,c(4,13)], by = "index") %>%
  filter(index %in% site_list_sub$index) %>%
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) 
```

# Long time series in_situ temperature
  First creating datasets for each region
  Now I determine the frequency and intensity of MCS using the MHW algorithm

```{r}
SACTN_PN <- SACTN_US %>%
  filter(site == "Port Nolloth") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_LB <- SACTN_US %>%
  filter(site == "Lamberts Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SB <- SACTN_US %>%
  filter(site == "Saldanha Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SP <- SACTN_US %>%
  filter(site == "Sea Point") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_HB <- SACTN_US %>%
  filter(site == "Hout Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

  The rate of onset of MCS can be determined here

```{r}
#' # Or if one wants to calculate MCSs
res_clim <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 10)
PN_MCS <- detect_event(res_clim, coldSpells = TRUE)

res_clim <- ts2clm(SACTN_LB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 10)
LB_MCS <- detect_event(res_clim, coldSpells = TRUE)

res_MCS <- ts2clm(SACTN_SB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 10)
SB_MCS <- detect_event(res_clim, coldSpells = TRUE)

res_clim <- ts2clm(SACTN_SP, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 10)
SP_MCS <- detect_event(res_clim, coldSpells = TRUE)

res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"),
                   pctile = 10)
HB_MCS <- detect_event(res_clim, coldSpells = TRUE)
```

## Upwelling events - 30th percentile used as a threshold to detect upwellling events

```{r}
res_clim <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 30)
PN_upwelling <- detect_event(res_clim)

res_clim <- ts2clm(SACTN_LB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 30)
LB_MCS <- detect_event(res_clim)

res_MCS <- ts2clm(SACTN_SB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 30)
SB_MCS <- detect_event(res_clim)

res_clim <- ts2clm(SACTN_SP, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 30)
SP_MCS <- detect_event(res_clim)

res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"),
                   pctile = 30)
HB_MCS <- detect_event(res_clim)
```

# Will have to be repeated for the remotely sensed sites

### 2019-08-01 Write_up/write_up.Rmd
updated

---
title: "Thesis - Upwelling and MHW/Cold spells analyses"
author: "Amieroh Abrahams"
date: "30 January 2019"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: yes
  html_document:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
  pdf_document:
    toc: yes
    toc_depth: 4
    highlight: default
    keep_tex: yes
    latex_engine: xelatex
fontsize: 10pt
geometry: margin=1in
language: Australian
mainfont: Minion Pro
monofont: Courier
papersize: A4
sansfont: Myriad Pro
tables: yes
bibliography: ../bib_master.bib
---

# Chapter 1: Introduction

## Eastern boundary upwelling systems

Globally, there are four major eastern boundary upwelling systems (EBUS). Upwelling systems include the Benguela Current off south-western Africa, the Canary Current off north-western Africa together with its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly]. Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans. EBUS cover only 1% of the world’s ocean surface, but despite this these are among the most productive regions of the world [@pauly]. This is due to the  systems exposure to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer]. As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas due to the abundance of nutrients and favourable temperatures.

Upwelling is primarily caused by alongshore, equator ward winds. These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods. Bakun proposed that if there was an increase in greenhouse gases, it would cause more heating during the day and more cooling during the night causing the temperature gradient to increase resulting in a strong pressure gradient, Ultimately effecting wind patterns and therefore upwelling. Several people have tried to model this using different datasets of SST and wind pattern data at various resolutions and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling. These models however, were based on coarse data and when higher resolutions were used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].

### Ecology and primary production within upwelling zones

The duration and intensity of coastal upwelling are known to play a critical role in the phenology of key marine ecosystem processes. Similarly, changes in the characteristics of upwelling are shown to cause large disturbances to ecosystems at several trophic levels [@wang]. Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies]. In these upwelling systems, the high productivity supports vast and diverse marine populations. Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean. Most of the world’s major fishing industries are dependent on upwelling, as this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis]. Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well. There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES). Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast. Inshore regions and embayments that are protected from wind action are likely to be subjected to intense episodes of primary production [@graham]. Vander Wouder et al. (2006) found that these sites represent the highest levels of chlorophyll concentration. Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions [@capone].

For several decades, regime shifts have become an important factor of marine research. Regime shifts are often defined as a shift from one stable ecosystem state to another. This process affects several species within a short period of time [@cury; @jarre; @blamey]. Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey]. The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components. Over the past  three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al. (2007) associated these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey]. This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region. These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al. 1993A, @pitcher1998]. Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf. Red tide events are greatest during the latter part of the upwelling season during transitions in the synoptic weather patterns which result in diminished upwelling activity and increased thermal stratification (Pitcher et al. 1993a).

### Atmosphere

Throughout regions surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high-pressure cells with quasi-stationary positions. The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van]. Solar heating during summer may result in the development of low-pressure cells, known as heat lows, which are absent during the winter [@tyson]. As the anticyclones shift north during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van]. The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].

## Sea surface temperature and data collection

Water temperature affects the survival, abundance, distribution and growth of marine species (REFERENCE). As such, a clear understanding of the relationship between marine ecosystems and their thermal tolerance is necessary for conservation and management planning. Future predictions in biogeographical patterns and ecosystem functioning may also be estimated based on various climatic behaviours. Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets. Unfortunately, not all data are created equally and under the same conditions and as such, not all datasets are appropriate for a particular task.

### *In situ* data

Having coastal, * in situ* seawater temperature data is a rare privilege to researchers. Most biological and ecological research requires coastal temperature data. The South African Coastal Temperature Network (SACTN). The availability of the *in situ* collected coastal temperature data products for several decades provides a reliable source of accurate coastal seawater temperature data [@smit]. SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments. At present, this dataset contains temperatures of 129 sites along the three distinct coasts of the South African coastline. Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneven per site. Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily. During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.

The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors. For instance, as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data. One problematic outcome of this sampling methodology was that, due to specific work of the South African Weather Service (SAWS) and Kwa-Zulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and recorded at a data precision of 0.500 ◦ C, however, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud]. Due of these inconsistencies, this dataset required verification from an independent source before its use in research. Smit et al. (2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results. Using these *in situ* collected temperature data along with remotely sensed SST data requires a clear understanding of the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long-term measurements are for use in climate change studies.

It should also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar]. Specifically, uncertainty lies within the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period. There is certainty that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs in this study.  Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.

For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset. Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers. These data points were then changed to NA (not available) so as to not interfere with analysis. All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/). The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MastersProject.

## Satellite data

Reliable datasets are often difficult to maintain and obtain. In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography. These datasets however, may be inaccurate along the coast and local *in situ* collected data should be used when possible (Smit et al., 2013). Remotely sensed data products are obtained by several different organisations worldwide. Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST. Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset. This is demonstrated as the size of the pixels one sees in the final product.

Satellite SST products undergo a range of quality control and validation processes (e.g. Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful. As previously mentioned, satellites may show a warm or cold bias along the coastal region but are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault). Remotely-sensed SST products are obtained at different resolutions, and as a result do not represent the same level of bias against *in situ* collected seawater temperature. Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.

This study made use of various satellite-derived Sea Surface Temperature (SST) datasets. One of datasets used was the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST). This dataset is an analysis of seawater temperature collected by different platforms (satellites, ships and buoys) on a global grid. The gridded daily SST product assimilates both remotely sensed and **in situ** sources of data to create a level-4 gap free product (Banzon et al., 2016). A spatially complete sea surface temperature map is produced by interpolation to fill in gaps within the dataset. The method used to fill these gaps include the adjustment of satellite and ship observations which compensate for bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst). There are two types of daily OISST, named after the relevant satellite sensors, these are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E). Currently the AVHRR has the longest record starting from 1981 to present day. Infrared instruments, like AVHRR, can make observations at relatively high resolutions. Microwave instruments, such as AMSR-E, can measure SSTs in most weather conditions but not adjacent to land. The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region. The AVHRR datasets have provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].

The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, used as the second dataset, was produced using satellite instruments with datasets spanning from1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/). MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals. This dataset is currently among the highest resolution SST dataset available. The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible. The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way in order to produce 1km global maps representing sea surface temperature.

The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates. MODIS was launched in 1991 and 2002 to study global dynamics of the earth’s atmosphere, ice and ocean. MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong]. MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands (NASA.GOV, http://oceancolor.gsfc.nasa.gov).

The CMC dataset comprises the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC). The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with *in situ* observations of SST from ships and buoys from the ICOADS program.

The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and *in situ* data from drifting and moored buoys [@dash]. This dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).

### Climate change

Climate change as a result of anthropogenic activities is a global concern of both scientific and political importance. As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems. Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney]. Direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour of marine organisms and in turn, these changes lead to a change in species integration and trophic pathways.

Significant research studies accept that climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh]. The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of the increase of greenhouse gas concentrations [@solomon]. A large amount of this heat will be absorbed by the world’s ocean. In current years, as a result of changes in air temperature, the heat content of the ocean has increased significantly. Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades. Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007). Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than the global average [@levitus; @santos]. In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].

As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface. This warming slowly propagates down through the water column and in turn increases thermal stratification [@gruber]. Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016]. The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therefore strongly linked to biological productivity. Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker]. Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses. As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted with an increase in nutrient levels [@rykaczewski].  Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq]. It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which may result in unusual trends.

As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs. As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010]. In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad]. Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@keeling]. Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].

Climate change has severe ecological consequences on both plants and animals (Bradley et al. 1999; Walther et al. 2002). In the past two decades, sea surface temperature has risen in many regions and this seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990). The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993). Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process. In a bid to further this understanding, Rockström et al. (2009) introduced the concept of planetary boundaries. According to this concept, within each of these boundaries exists a level of safety that should not be exceeded. According to this concept, the planetary boundary for biodiversity loss, has been surpassed [@steffen]. Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed. The idea of planetary boundaries is essential in that it helps in understanding the threats humans cause via the continued destruction of the natural environment. Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea. Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom]. Climate change has been proven to decrease primary production thus affecting trophic levels and leading to trophic turbulence and the destruction of marine fauna [@lewandowska; @dell, @mcCauley].

Climate change affects coastal species such as oysters, corals and, sea grass. Lord (2017) also found evidence that suggested that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer]. Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity has been observed to be affected. Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis. Here the increase in global temperatures is one of the main focus points along with chlorophyll concentration. By making use of these two measurements, an analysis of how chlorophyll concentration varies with temperature changes could be conducted.

## Benguela

The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundary upwelling system. The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes. These currents are forced locally by the wind stress field off southwest Africa [@nelson ; @fennel]. This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole]. Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms]. The main upwelling period in the Benguela, north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year. Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August). In the central Benguela and also at Cape Frio, upwelling is continuous throughout the year [@shannon1985]. The cold Benguela waters are bound in the south by the warm Agulhas retroflexion region and in the north by the southward flowing Angola current. The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange]. The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].

Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not. This results in a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.  The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon]. There have been a number of warm and cool periods documented in the northern and southern Benguela [e.g. @stander; @walker; @shannon1984; @boyd; @brainard]. Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water. This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water. Another warming event occurred during the summer and autumn months of 1984 [@shannon1986]. The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela. Shannon et al. (1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore, west of Luderitz. The absence of upwelling between Walvis Bay and north of Luderitz is evident, however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal. Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions. The seasonal temperature pattern within the northern Benguela was somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season. During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.

Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson]. Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the downwelling process southward. Southward, the coastal upwelling system is bounded by the Agulhas Bank. South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current [@duncombe]. The region surrounding the Benguela experiences a persistent along shore wind. This wind is associated with the St. Helena high pressure system. Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds. These favourable upwelling equatorial winds represent a maximum at 25°S and decrease towards the northern and southern boundaries of the Benguela upwelling system. Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast. Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel]. Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher]. Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind, with wind increasing somewhat away from the coast. This facilitated the idea of a wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).

Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986]. These warm events profoundly impact fisheries and the climate of this region. They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer]. Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures. Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region [@sutton; @zebiak]. Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod]. Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics. Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline. It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños. ***Need to make changes to this paragraph, random but feels important***

### Extreme events

Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g. @gunter; @lirman; @boucek] or heatwaves [e.g. @gordon; @stott; @perkins]. These events occur rapidly, and unexpectedly and have significant impacts on species diversity, distributions, growth and survival where they take place. These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago. Jentsch et al. (2007) were the first to discuss this mode of climate change and brought it to greater focus. Steps have since been taken towards understanding the drivers, impacts and consequences of these events. Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming [@mccabe; @harley; @schoepf].

Little is known about extreme events occurring within the South African region. Few South African researchers have specifically focussed on investigating this topic, and published literature is sparse. During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, *Donax serra Röding*. Further studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish. *Acropora cervicornis*, was found to be particularly susceptible to low temperatures and experienced a high mortality rate [@shinn; @davis]. While these phenomena are not as common as warming events, they have a dramatic and long-lasting effect on coral communities. Studies in January 2010 revealed that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman]. Species population distribution are often set by cold distribution boundaries.  More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves. It is now becoming accepted that extreme events are a threat to several ecosystems [@thompson]. In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease. This thesis aims to address that assumption.

### Event detection

As specified earlier, the extreme cold events occurring along the west coast of South Africa, which is central to this thesis will be detected and correlated to upwelling events with the algorithm developed by Hobday et al. (2016). The events that recorded *in situ* data from SACTN would be compared against those detected in satellite derived SST. For this analyses, the R packages ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events. In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred. With these annual values it is possible to calculate long-term trends in extreme events.

## Analyses

The analyses conducted in this thesis varied according to the investigations performed. The one common thread uniting all of the analyses was that they were conducted in R-programming language (R Core Team, 2017). This provided reproducibility for all of the analyses and visualisations used.

## Problem statement:

With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future. This increase will undoubtedly have an impact on the world’s oceans and the marine life within it,  both directly and indirectly. The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions. Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing. To assess changes in upwelling events, the aim of this investigation is to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event. Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015). Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS. We also aim to determine whether or not there is a change in the rate of onset of marine heatwaves and marine coldspells.

## Aims and approach

The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley]. This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet]. Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region. This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region. Its aims to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis. In this research project, various satellite derived observations of sea surface temperature (SST) and wind stress will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.

## Study 1

The primary aim of the first study was to make use of various remotely sensed SST data, in-situ collected coastal seawater temperature data and chlorophyll-a data. It was also useful to examine how the intensity and duration of upwelling varies per dataset as these datasets are collected at different resolutions, with the assumption that high resolution data may reveal a more accurate detection of the duration and intensity of upwelling events. Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling. Chlorophyll a data were then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites. It was hypothesised that an increase in the intensity and duration of upwelling will show an increase in the abundance of chlorophyll a concentration. The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets. Thereafter analyses rate of onset of MHW and MCS was done.

Thereafter the aim of the second study was to determine the occurrence of MCSs along the west coast of Southern Africa between the nearshore (<400m from the shore) and offshore. The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al. 2013). With the knowledge obtained from study 1, it was now possible to investigate whether a more intense upwelling event may lead to an increase in the intensity and duration of MCSs. The SACTN dataset and a total of five SST datasets could be used to investigate the occurrence of marine cold-spells (MCSs) and marine heatwaves with the methodology introduced by Hobday et al. 2016. This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends.

## Study 2

The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine wether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves. It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises.

## Research assumptions

The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared. Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets. An additional assumption within this research focused on the effect of upwelling on MCSs and MHWs. Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events. For the purpose of this study the influence of upwelling on MCSs was the primary focus.

# Paper 1
## Chapter 2 Upwelling
### Introduction

Upwelling events are known to have a large influence on the abundance, diversity and distribution of marine organisms and critically affect marine ecosystem processes. The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang]. Globally, there are four major eastern boundary upwelling systems (EBUS). Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly]. Eastern boundary upwelling regions only cover 1% of the worlds ocean surface, but are among the most productive regions of the world. This is because these regions are influenced by upwelling events. Upwelling is defined as a process whereby cold, nutrient rich waters are pushed to the surface as a result of alongshore equatorial winds. The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast an d the Benguela current, an eastern boundary upwelling system, along the west coast. These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al. 2004; Mead et al. 2013; Schlegel and Smit 2016). The Benguela Current is driven by the anticyclone high pressures systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al. 2009). The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings]. Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al. 2009; Schlegel et al. 2017).

Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems. The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney]. According to the Bakun hypothesis, an increase in greenhouse gases, will result in an increase in day time warming and night time cooling, this ultimately causes an increase in temperature gradients and this further results in a strong pressure gradient. Pressure gradients are known to affect wind patterns and wind patterns subsequently control the intensity and duration of upwelling events [@hsieh; @mote; @lima].

Several research studies provide significant evidence suggesting an increase in ocean surface temperatures. These temperatures are agreed to have increased as a result of climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification. This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. This process is thus strongly linked to primary productivity. Climate change impacting ocean processes has major ecological consequences (Bradley et al. 1999; Walther et al. 2002). For example, several regions show an increase in sea surface temperatures. This rise in temperatures largely influence species survival, reproduction and abundance. These changes are known to yield several economic consequences (Murawski 1993), thus, understanding the impact and cause of climate change is a challenging phenomenon. Similarly, climate change also largely influences coastal species such as corals, muscles and oysters. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.

The Benguela system is bounded by the equatorward and poleward ends by warm water regimes. These currents are shown to be forced by local wind stress fields off southwest Africa [@nelson; @fennel]. This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole]. Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms]. The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August. The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole]. Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard]. Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events. During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region. In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events. Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds. Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]

Global warming is increasing at an extremely high rate. This will result in an increase in greenhouse gases which in turn will impact the ocean and terrestrial life. The process of upwelling is one such factor that is predicted to change with an increase in global warming. As previously mentioned, changes in temperature results in changes in pressure gradients, these pressure gradients control wind patterns which ultimately controls the intensity and duration of upwelling. To assess the changes within the upwelling events, various datasets would be used. These are satellite sea surface temperatures (broad scale), in situ collected seawater temperatures (local scale), wind data as well as satellite obtained chlorophyll-a data. Having assessed the eastern boundry upwelling system within a context that comprised of broad and local-scale features permitted us to assess how upwelling varied between sites located closer to the upwelling centres and furthur away from the upwelling centre. Specifically, we aim to assess whether or not it is possible to identify the upwelling signal throughout the various datasets and here we examine how far offshore the upwelling signal is present in the higher resolution data and how far offshore the SACTN signal extend. We hypothesized that (i) wind action may influence the intensity and duration of upwelling; (ii) Remotely sensed SST data will represent a similar event frequency and duration; (iii) Sites located closer to upwelling centres displays distinct variation in the intensity and duration of upwelling events compared to sites furthur away from the upwelling centeres.

## Methods

In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature, chlorophyll-a and wind were analysed and accessed.

## Temperature data sources

This research study lies within the borders of South Africa along the west coast in particular. The locations of each of the points used within this study can be seen in Fig.1. The South African Coastal Temperature Network was the primary dataset used to specify the various sites along the west coast of South Africa. This dataset consists of coastal seawater temperatures obtained from 129 sites along the entire South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al. 2009). These currents are known to largely influence the seawater temperature and marine life along the South African coastline. The SACTN has collected SST data along this unique coastline from as early as 1972 till present date, with contributions made from several organisations and govermental departments. As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs). The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C. The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.  During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s. Recordings are still ongoing at many of these sites.

This study also made use of a range of remotely sensed temperature datasets. Satellites provide a system of large spatial and temporal data retrieval. Remotely sensed SST products are produced by a number of organisations across the globe. The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region. The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006). OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016). The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/). MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available. The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC). The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.

An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well. Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.

Inorder to create a time series of each of the remotely sensed SST data representing and also comparing the broad-scale temperatures against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa. Temperature values from the various satellite datasets were extracted at multiple distances from the coast. These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Figure 1). The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.

## Wind data

Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa. Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions. Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).

## Chlorophyll-a data

In order to observe the patterns in the variation of the chlorophyll-a concentration with the intensity and duration of upwelling events,  MODIS-Aqua Chlorophyll a concentration, OCI Algorithm, 8-day composite were used. The National Aeronautics and Space Administration’s (NASA) Moderate Resolution imaging Spectroradiometer (MODIS) Aqua satellite is an example of a satellite currently collecting the reflected wavelength energy from the earth. The MODIS Aqua satellite records wavelength data down to the 1% photosynthetically available depth (0.1 PAR depth) of the water column, a vital component in the estimation of chlorophyll (NASA.GOV, http://oceancolor.gsfc.nasa.gov). Because chlorophyll-a concentration retrieval is only possible over cloud-free areas the daily data offered may have large data gaps due to cloud cover. The quality of the chlorophyll-a concentration  estimates depends strongly on the quality of the cloud mask. The NASA Aqua satellite contains the MODIS sensor that is used for detecting chlorophyll-a concentrations (ocean color) in the world's oceans, among other applications (NASA.GOV, http://oceancolor.gsfc.nasa.gov). The Level 3 standard mapped image (SMI) chlorophyll-a dataset has a monthly temporal resolution and 4.6 km spatial resolution. The SMI dataset is an image representation of binned MODIS data.

## Site selection

At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead]. The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings]. Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine. In order to examine upwelling patterns between sites, the appropriate sites along the west coast of South Africa (i.e. sites influenced by the Benguela current) were selected. Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites that could be adequately compared. Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 1992 – 2015.  This yield a total of six sites along the west coast of South Africa. Hereafter, each of the remotely sensed SST datasets were matched to the SACTN dataset. This was done for two reasons, to find the same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline. The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("../Figures/map_fixed.png")
```

### Detecting the intensity and duration of upwelling

In South Africa, the changes in wind patterns, which occurred in the early 1980s, have resulted in increased upwelling variability and mean upwelling intensity along the west coast in particular, since the early 1990s (Blamey et al. 2012). This is driven by the shift to stronger south-easterly winds which are the drivers of upwelling in this system (Blamey et al. 2012, Hutchings et al. 2012). Additionally, Rouault et al. (2010) have described a negative trend (0.5°C per decade) in seawater temperature along the west coast which is a suggested result of the intensification of upwelling- favourable wind patterns (Blamey et al. 2015). This is contrary to the global trend of warming ocean temperatures as a result of climate change. *In order to determine wether upwelling had an influence on MCS, First we had to determine the various upwelling events and then using the same dataset we had to determine when marine cold spells occured or if there was a change in upwelling and

## Determining the upwelling events 
Known upwelling events and how they link up with Cold spells

- Step 1 Find upwelling events
- Step 2 when have upwelling see if this match the MCS with that

### MCS

## Results

### Wind patterns during upwelling events

```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagrams representing the the most predominant wind direction for each of the sites for the years 2010 - 2014. Each spoke is divided by color into wind speed ranges. The radial length of each spoke around the circle is the amount of times that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}

 knitr::include_graphics("../Figures/wind.png")
```

## Chlorophyll a content during upwelling and cold spells

### 2019-08-01 site_map.Rmd
updated

---
title: "site_map"
author: "Amieroh Abrahams"
date: "01 August 2019"
output: html_document
---

# Loading Libraries

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE 
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
library(ncdf4)
library(reshape2)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ggpubr)
library(ggpubr); theme_set(theme_pubr())# library(devtools)
# install_github("marchtaylor/sinkr")

## Functions
source("Functions/earthdist.R")
source("Functions/wind.rose.R")
source("Functions/theme.R")
source("Functions/scale.bar.func.R")
```

```{r}
load("Data/site_list_sub.Rdata")
load("Data/africa_coast.RData")
xtable(site_list_sub, auto = TRUE)
west <- site_list_sub
west$coast <- "west" # Chnages wc to west

load("Data/sa_bathy.RData")
## Downloading the bathy data from NOAA
# Download mid-res bathymetry data
# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
# colnames(sa_bathy) <- c("lon", "lat", "depth")
# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")

# Loading in the newly downloaded bathymetry data
load("Data/sa_bathy.RData")
load("Data/africa_coast.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
shore.normal.transect <- function(site, width = 2){
  # Find the site on the coastline and it's nearest neighbour points
  coords <- data.frame(lon = site$lon, lat = site$lat)
  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
  coords3 <- coords3[2:1,1:3]
  # Define the shore normal transect bearing
  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
  if(heading >= 360){
    heading <- heading-360
  } else {
    heading <- heading
  }
  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
  return(heading2)
}

# Creating the transects
site_transects <- data.frame()
for(i in 1:length(west$site)){
 site <- west[i,]
 site_transect <- shore.normal.transect(site, 2)
 site_transects <- rbind(site_transects, site_transect)
}

# Manually correcting Sea Point and Kommetjie
site_transects$heading[4] <- 290 
# save(site_transects, file = "Data/site_transects.RData")
load("Data/site_transects.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath

transect.pixel <- function(site, distances){
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  # deep <- 999
  # distance_multiplier <- 1
  # while(deep > isobath){
  for(i in 1:length(distances)){
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, 
                         distance = distances[i])
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
  }
  if(nrow(pixels) < 1){
    pixels <- data.frame(site, depth = NA)
  }else{
    pixels <- pixels
  }
  return(pixels)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
  site_pixels <- rbind(site_pixels, site_pixel)
}

# save(site_pixels, file = "Data/site_pixels.RData")
load("Data/site_pixels.RData")

# Bounding box
  # Only one is made in order to know how large the the geom_point() squares should be made to match
bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])

# Determining the temperature at the various distances from the coast
```

```{r}
load("Data/south_africa_coast.RData")
load("Data/africa_coast.RData")
load("Data/sa_provinces_new.RData")
load("Data/site_list_sub.Rdata")
load("Data/site_pixels.RData")
load("Data/MUR.RData")
names(south_africa_coast)[1] <- "lon"

# Manually divide up coastline
wc <- south_africa_coast[291:410,]

# Setting up the theme
theme_set(theme_bw())
limits <- c(12,28) # for colour bar
#breaks <- seq(6, 30, 2) # Create breaks to be used for colour bar

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 25)

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  xlab("") + ylab("") +
  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
  #            size = 4, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size =4, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 4, angle = 302, colour = "black") +
    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
    #          size = 5, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5),
        axis.text = element_text(size = 6)) +
    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.direction = "horizontal",
        legend.justification = c(1,0),
        legend.position = c(0.65, 0.80),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
  theme_bw() + 
  coord_equal() + 
  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
  theme(panel.border = element_rect(colour = "black", size = 0.4), 
        plot.background = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
africa_map

# More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html

pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
print(site_map, vp = vp2)
print(africa_map, vp = vp1)
dev.off()

# Site map
# The insetting needs to be placed at appropriate location
```

### 2019-08-01 wind_analyses.Rmd
data_commit

---
title: "wind_diagrams"
author: "Amieroh Abrahams"
date: "01 August 2019"
output: html_document
---

### 2019-08-01 wind_analyses.Rmd
wind_analyses

@@
 ---
+
+Loading the Wind data.
+The wind dataset was supplied by the SAWS in .text format.
+Wind data was provided for the sites closes to the requested site as wind data may not be available for the 6 study sites in this research.
+The wind direction was collected every 2 hours and by using the circular function we calcuated the daily wind direction and mean speed.
+There is a numerically large gap between 360 and 2 where as in degrees its not as large.
+The circular mean function returns the mean direction of a vector of circular data.
+https://cran.r-project.org/web/packages/circular/circular.pdf (Pg 130).
+The circular function creates circular objects around the wind direction.
+
+```{r}
+wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "", 
+                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
+
+wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+# Slecting the important columns for each of the datasets
+wind_fix <- function(df){
+wind <- df %>% 
+  select(station_name, date, hour, dir, speed) %>%  #column names
+  mutate(date = as.Date(as.character(date)),
+         hour = as.numeric(as.character(hour)), 
+         dir = as.numeric(as.character(dir)),
+         speed = as.numeric(as.character(speed)))
+}
+# RWS: We can see when we force the values to be numeric that there are some non-numeric values in the base data
+wind_1 <- wind_fix(df = wind_1)
+wind_2 <- wind_fix(df = wind_2)
+wind_3 <- wind_fix(df = wind_3)
+
+## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+
+renaming_sites_1 <- function(df) {
+  sites <- df %>%
+      # RWS: An alternative way to replace values without having to use multiple ifelse() statements
+      mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
+                                    station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
+                                    station_name  == "PORT NOLLOTH" ~"Port Nolloth",
+                                    station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
+                                    station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
+                                    station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
+  return(sites)
+}
+
+wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
+wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
+wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
+wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+save(wind_data, file = "Data/wind_data.RData")
+
+# # RWS: You should never be removing values by specific call like this
+#   # Rather you should be able to use some sort of conditional to screen out unwanted values
+# wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
+# renaming_sites_2 <- function(df) {
+#   sites <- df %>%
+#     mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
+#                            ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
+#                                 ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
+#                                        ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
+#   return(sites)
+# }
+# 
+# wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
+# wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ] 
+# renaming_sites_3 <- function(df) {
+#   sites <- df %>%
+#     mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
+#                            ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
+#                            ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
+#   return(sites)
+# }
+# 
+# wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
+# wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ] 
+### CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
+# wind_3_HoutBay <- wind_3 %>% 
+#   filter(station_name == "CAPE TOWN SLANGKOP") %>% 
+#   mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
+# wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+# wind_data <- wind_data %>% 
+#   na.omit()
+# 
+# wind_data <- wind_data %>% 
+#   # mutate(date = as.Date(date)) %>%  
+#   group_by(date, hour, temp_sites) 
+# 
+# # write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
+# # save(wind_data, file = "Data_P1/wind_data.RData")
+
+# load("Data_P1/insitu_wind.RData")
+# 
+# test1 <- insitu_wind %>% 
+#   mutate(lat = case_when(site == "Port Nolloth" ~ "-29.25241",
+#                                     site == "Saldanha Bay" ~ "-33.01054",
+#                                     site  == "Yzerfontein" ~ "-33.36068",
+#                                     site  == "Kommetjie" ~ "-34.13667",
+#                                     site  == "Lamberts Bay" ~ "-32.09185",
+#                                     site  == "Sea Point" ~ "-33.91847"))
+# 
+# test <- test1 %>% 
+#   mutate(lon = case_when(site == "Port Nolloth" ~ "16.86710",
+#                                     site == "Saldanha Bay" ~ "17.95063",
+#                                     site  == "Yzerfontein" ~ "18.15731",
+#                                     site  == "Kommetjie" ~ "18.32674",
+#                                     site  == "Lamberts Bay" ~ "18.30262",
+#                                     site  == "Sea Point" ~ "18.38217"))
+# 
+# test_final <- test %>% 
+#   select(date,lat,lon,dir_circ, mean_speed) %>% 
+#   dplyr::rename(speed = mean_speed) %>% 
+#   dplyr::rename(dir = dir_circ)
+
+# Loading the wind data
+library(tidyverse)
+library(circular)
+
+# wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number())) # Dataset too large for Github but better data at line 638
+
+# wind_data <- read_csv("Data_P1/wind_data.csv")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+
+wind_daily <- wind_data %>%
+  # ungroup() %>%
+  # mutate(date = as.Date(date))%>%
+  dplyr::group_by(temp_sites, date) %>%
+  filter(temp_sites %in% selected_sites) %>% 
+  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
+                   mean_speed = round(mean(speed),2))
+
+# save(wind_daily, file = "Data/wind_daily.RData")
+# # Loading in the daily wind data
+
+# Loading the subsetted wind data created from the text files above 
+load("Data/wind_daily.RData") 
+```

### 2019-08-01 Analyses/long_ts.Rmd
2.0
update

@@
   filter(length > 10950) 
+# save(site_list_sub, file = "site_list_sub.Rdata")
 
@@
   separate(index, into = c("site", "src"), sep = "/", remove = FALSE)
+
+save(SACTN_US, file = "Data/SACTN_US.RData")
 ```

### 2019-08-01 wind_analyses.Rmd
update

@@
 # wind_data <- read_csv("Data_P1/wind_data.csv")
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
 
@@
 ```
+
+# Matching
+  Matching the wind data with the temperature data using the match function
+
+```{r}
+match_func <- function(df){
+  match <- SACTN_US  %>%  
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+wind_daily <- wind_daily %>% 
+  dplyr::rename(site = temp_sites)
+
+wind_match <- match_func(df = wind_daily)
+
+library(openair)
+print(
+with(wind_match,
+     windRose(data_frame(ws=mean_speed, wd=dir_circ, 
+                         date=date, station=factor(site)),
+
+              paddle=FALSE, type="station", width=2))
+)
+# png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
+?windRose
+
+# insitu_wind_2 <- insitu_wind %>% 
+#   filter(site == "Port Nolloth")
+# 
+# insitu_wind_3 <- insitu_wind %>% 
+#   filter(site == "Lamberts Bay")
+# 
+# plotPN <- plot(as.factor(format(insitu_wind_2$date, "%Y-%m")), insitu_wind_2$insitu_temp, col = "lightpink")
+# plotLB <- plot(as.factor(format(insitu_wind_3$date, "%Y-%m")), insitu_wind_3$insitu_temp, col = "lightpink")
+#        
+# ?plot
+```

### 2019-08-02 wind_analyses.Rmd
Wind_rose (Progress)

@@
 wind_match <- match_func(df = wind_daily) 
+# save(wind_match, file = "Data/wind_match.RData")
 
@@
 ?windRose
-
-# insitu_wind_2 <- insitu_wind %>% 
-#   filter(site == "Port Nolloth")
-# 
-# insitu_wind_3 <- insitu_wind %>% 
-#   filter(site == "Lamberts Bay")
-# 
-# plotPN <- plot(as.factor(format(insitu_wind_2$date, "%Y-%m")), insitu_wind_2$insitu_temp, col = "lightpink")
-# plotLB <- plot(as.factor(format(insitu_wind_3$date, "%Y-%m")), insitu_wind_3$insitu_temp, col = "lightpink")
-#        
 # ?plot
 ```
+
+## Wind rose plots
+
+```{r}
+# Loading in the matched wind data
+load("Data/wind_match.RData")
+
+wind_match <- wind_match %>% 
+  drop_na()
+
+wave_daily_renamed <- wind_match %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ) 
+source("Functions/wind.rose.R")
+p.wr2 <- plot.windrose(data = wave_daily_renamed,
+              spd = "spd",
+              dir = "dir")
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = "") +
+  theme(strip.text.x = element_text(size = 25))
+p.wr3  # Change colours and remove the NA values
+```

### 2019-08-02 Chlorophyll.Rmd
chloro

---
title: "Chlorophylla"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

## Script info
The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV. The chlorophylla data were obtained from MODIS Aqua. The aim of working with these chlorophyll a data is to examine how the chlorophyll a concentration varies with upwelling.

```{r}
library(ncdf4)
library(data.table)
library(tidyverse)
library(reshape2)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(lubridate)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(ggpubr)
library(fasttime)
library(FNN)
library(heatwaveR)
```

This function convert chlorophyll netCDF datasets to a single CSV datset

```{r}
MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"

region <- "BC" # Benguela Current
coords <- bbox[, region]
coords <- c(-35, -20, 10, 20) # this is the BC

ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
nc.init <- nc_open(ncList[1])
LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
nc_close(nc.init)

ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
nc <- nc_open(nc.file)
  instrument <- ncatt_get(nc, 0, "instrument")$value
  platform <- ncatt_get(nc, 0, "platform")$value
  product_name <- ncatt_get(nc, 0, "product_name")$value
  fNameStem <- substr(product_name, 17, 38)
  timeStamp <- substr(product_name, 2, 8)
  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
  chl <- round(ncvar_get(nc,
                   varid = "chlor_a",
                   start = c(LonIdx[1], LatIdx[1]),
                   count = c(length(LonIdx), length(LatIdx))),
               3)
  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat =  nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  chl <-
    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
    mutate(t = date) %>%
    na.omit()
  fwrite(chl,
         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(chl)
}

llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)

# MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
#             col.names = c("lon", "lat", "chloro", "date"))

# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
```

## Setting up the chlorophylla dataset

```{r}

# Loading the data
load("Data/MODIS_Chloro.RData") # Extracted chlorophyll data
load("Data/SACTN_US.RData") # Site list of all the sites along the west coast

chloro_plot <- MODIS_Chloro %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

unique_pixel <- MODIS_Chloro %>% 
  select(lon, lat) %>% 
  unique()

match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(SACTN_US[,5:6]), k = 1)

# Matching the sites within the chlorophyll data to the sites along the west coast
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = SACTN_US$site)

chloro_match <- MODIS_Chloro %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)

chloro_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

combined_chloro_sites <- chloro_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::mutate(date = as.Date(date))

# Save the complete dataset
# save(combined_chloro_sites, file = "Data/combined_chloro_sites.RData")

Chloro_plot_complete <- ggplot(combined_chloro_sites, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_complete

# Plotting only years 2014-2015
daily_chloro_data_tester <- combined_chloro_sites %>% 
  filter(year(date) %in% seq(2014, 2015))

Chloro_ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_filtered

# # With the chloro pixels filter, it is now faster to create the daily data
# daily_chloro_data <- daily_chloro_match %>%
#   dplyr::mutate(date = as.Date(date)) %>%
#   dplyr::group_by(combi, lon, lat, date) %>%
#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
#   dplyr::group_by(combi, lon, lat) %>%
#   tidyr::nest() %>%
#   dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
#                      climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
#                      maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
#   dplyr::select(-data) %>%
#   tidyr::unnest() %>%
#   select(-doy, -seas, -thresh)

## Plotting daily data
# ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
#   geom_line() +
#   facet_wrap(~combi, ncol = 1)

# daily_chloro_match_sites <- daily_chloro_data %>%
# left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year"))
```

### 2019-08-02 Chlorophyll.Rmd
chlorophyll

@@
 load("Data/MODIS_Chloro.RData") # Extracted chlorophyll data
-load("Data/SACTN_US.RData") # Site list of all the sites along the west coast
+load("Data/site_list_sub.RData") # Site list of all the sites along the west coast
 
@@
 match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-                          query = as.matrix(SACTN_US[,5:6]), k = 1)
+                          query = as.matrix(site_list_sub[,5:6]), k = 1)
 
@@
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  mutate(site = SACTN_US$site)
+  mutate(site = site_list_sub$site)
 

### 2019-08-02 Chlorophyll.Rmd
chloro

@@
 # # With the chloro pixels filter, it is now faster to create the daily data
-# daily_chloro_data <- daily_chloro_match %>%
-#   dplyr::mutate(date = as.Date(date)) %>%
-#   dplyr::group_by(combi, lon, lat, date) %>%
-#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
-#   dplyr::group_by(combi, lon, lat) %>%
-#   tidyr::nest() %>%
-#   dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
-#                      climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
-#                      maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
-#   dplyr::select(-data) %>%
-#   tidyr::unnest() %>%
-#   select(-doy, -seas, -thresh)
+daily_chloro_data <-  %>%
+  dplyr::mutate(date = as.Date(date)) %>%
+  dplyr::group_by(combi, lon, lat, date) %>%
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
+  dplyr::group_by(combi, lon, lat) %>%
+  tidyr::nest() %>%
+  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+                     climatologyPeriod = c("2002-07-13", "2018-12-04"), 
+                     maxPadLength = 8)) %>% 
+  dplyr::select(-data) %>%
+  tidyr::unnest() %>%
+  select(-doy, -seas, -thresh)
 

### 2019-08-02 Chlorophyll.Rmd
update

@@
 
-chloro_match %>%
-  filter(date == max(date)) %>%
-  ggplot(aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = chloro))
+# chloro_match %>%
+#   filter(date == max(date)) %>%
+#   ggplot(aes(x = lon, y = lat)) +
+#   geom_raster(aes(fill = chloro))
+# 
+# chloro_sites <- chloro_match %>%
+#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
+#   dplyr::mutate(date = as.Date(date))
+# # save(chloro_sites, file = "Data_P1/chloro_sites.RData")
+# 
+# # Visualisation
+# load("Data_P1/chloro_sites.RData")
+# Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~site, ncol = 1)
+# Chloro_plot_complete
+# 
+# # Study period from 2010 - 2014
+# chloro_data <- chloro_sites %>% 
+#   filter(year(date) %in% seq(2010, 2014))
+# # save(chloro_data, file = "Data_P1/chloro_data.RData")
+# 
+# # Plotting only years 2014-2015
+# load("Data_P1/chloro_data.RData")
+# Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~site, ncol = 1)
+# Chloro_plot_filtered
 
-combined_chloro_sites <- chloro_match %>%
-  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
-  dplyr::mutate(date = as.Date(date))
+# Creating daily temperatures
+load("Data_P1/chloro_data.RData") # Loading in the chlorophyll-a data
 
-# Save the complete dataset
-# save(combined_chloro_sites, file = "Data/combined_chloro_sites.RData")
-
-Chloro_plot_complete <- ggplot(combined_chloro_sites, aes(x = date, y = chloro)) +
-  geom_line() +
-  facet_wrap(~site, ncol = 1)
-Chloro_plot_complete
-
-# Plotting only years 2014-2015
-daily_chloro_data_tester <- combined_chloro_sites %>% 
-  filter(year(date) %in% seq(2014, 2015))
-
-Chloro_ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
-  geom_line() +
-  facet_wrap(~site, ncol = 1)
-Chloro_plot_filtered
-
-# # With the chloro pixels filter, it is now faster to create the daily data
-daily_chloro_data <-  %>%
-  dplyr::mutate(date = as.Date(date)) %>%
+daily_chloro <- chloro_data %>%
+  # dplyr::mutate(date = as.Date(date)) %>% 
   dplyr::group_by(combi, lon, lat, date) %>%
@@
   dplyr::group_by(combi, lon, lat) %>% 
+  drop_na() %>% 
   tidyr::nest() %>%
   dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
-                     climatologyPeriod = c("2002-07-13", "2018-12-04"), 
-                     maxPadLength = 8)) %>% 
+                     climatologyPeriod = c("2010-02-11", "2014-12-20"),
+                     maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
   dplyr::select(-data) %>%
@@
 
-## Plotting daily data
-# ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
-#   geom_line() +
-#   facet_wrap(~combi, ncol = 1)
-
-# daily_chloro_match_sites <- daily_chloro_data %>%
-# left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year"))
+****Error: Evaluation error: sort(): detected NaN.*** 
 ```

### 2019-08-03 SACTN_lts.Rmd
updating

--- 2019-08-01 Analyses/long_ts.Rmd
@@
 
-save(SACTN_US, file = "Data/SACTN_US.RData")
+# save(SACTN_US, file = "Data/SACTN_US.RData")
 ```
@@
 ```{r}
+
 #' # Or if one wants to calculate MCSs
-res_clim <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2015-12-31"),
+
+MCS_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                    pctile = 10)
-PN_MCS <- detect_event(res_clim, coldSpells = TRUE)
+  return(res_clim)
+}
 
-res_clim <- ts2clm(SACTN_LB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
+PN_MCS <- MCS_func(df = SACTN_PN)
+LB_MCS <- MCS_func(df = SACTN_LB)
+SB_MCS <- MCS_func(df = SACTN_SB)
+SP_MCS <- MCS_func(df = SACTN_SP)
+
+res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
                    pctile = 10)
-LB_MCS <- detect_event(res_clim, coldSpells = TRUE)
 
-res_MCS <- ts2clm(SACTN_SB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 10)
-SB_MCS <- detect_event(res_clim, coldSpells = TRUE)
+## Events
+MCS_event <- function(df){
+  detect <- detect_event(df, coldSpells = TRUE)
+  return(detect)
+}
 
-res_clim <- ts2clm(SACTN_SP, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 10)
-SP_MCS <- detect_event(res_clim, coldSpells = TRUE)
+PN_MCS_events <- MCS_func(df = PN_MCS)
+LB_MCS_events <- MCS_func(df = LB_MCS)
+SB_MCS_events <- MCS_func(df = SB_MCS)
+SP_MCS_events <- MCS_func(df = SP_MCS)
 
-res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"),
-                   pctile = 10)
-HB_MCS <- detect_event(res_clim, coldSpells = TRUE)
+HB_MCS_events <- detect_event(res_clim, coldSpells = TRUE) # line 118 -
 ```
 
-## Upwelling events - 30th percentile used as a threshold to detect upwellling events
+## Upwelling events - 30th percentile used as a threshold to detect upwellling events.
 
 ```{r}
-res_clim <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2015-12-31"),
+upwell_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                    pctile = 30)
-PN_upwelling <- detect_event(res_clim)
+  return(res_clim)
+}
 
-res_clim <- ts2clm(SACTN_LB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 30)
-LB_MCS <- detect_event(res_clim)
+PN_upwell <- upwell_func(df = SACTN_PN)
+LB_upwell <- upwell_func(df = SACTN_LB)
+SB_upwell <- upwell_func(df = SACTN_SB)
+SP_upwell <- upwell_func(df = SACTN_SP)
 
-res_MCS <- ts2clm(SACTN_SB, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 30)
-SB_MCS <- detect_event(res_clim)
+res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
+                   pctile = 10)
 
-res_clim <- ts2clm(SACTN_SP, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 30)
-SP_MCS <- detect_event(res_clim)
+## Events
+upwell_event <- function(df){
+  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
+  return(detect)
+}
 
-res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"),
-                   pctile = 30)
-HB_MCS <- detect_event(res_clim)
+PN_upwell_events <- MCS_func(df = PN_upwell)
+LB_upwell_events <- MCS_func(df = LB_upwell)
+SB_upwell_events <- MCS_func(df = SB_upwell)
+SP_upwell_events <- MCS_func(df = SP_upwell)
+
+HB_upwell_events <- detect_event(res_clim, minDuration = 2, coldSpells = TRUE) # line 1
 ```
-
-# Will have to be repeated for the remotely sensed sites

### 2019-08-05 remote_SST.Rmd
SST

---
title: "Remotely SST"
author: "Amieroh Abrahams"
date: "03 August 2019"
output: html_document
---

## MUR dataset

```{r}
MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
MUR_Hout_Bay<- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Lamberts_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Hout_Bay, MUR_Sea_Point) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp_MUR = nearest1) %>% 
  drop_na()

# save(MUR_SST, file = "Data/MUR_SST.RData")

match_func <- function(df){
  match <- SACTN_US  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR <- match_func(df = MUR_SST) 
```

## CMC dataset

```{r}
Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
Sea_Point<- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")

CMC <- rbind(Lamberts_Bay, Hout_Bay, Port_Nolloth, Saldanha_Bay, Sea_Point) %>%
  dplyr::rename(site = station)

CMC$date <- (ymd(CMC$date))
CMC <- CMC %>% 
  drop_na() %>% 
  dplyr::rename(temp_CMC = nearest1)

# save(CMC, file = "Data/CMC_SST.RData")

## Matching the Insitu data with the CMC SST data
insitu_CMC <- match_func(df = CMC) 
CMC_plot <- match_plot(df = insitu_CMC )
CMC_temp<- temp_plot(df = insitu_CMC)
```

## OISST

```{r}
OISSTDir <- "~/Documents/Masters_2019/Data_SST"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

# Visualise the data
# To explore the data I visualise the min temperatures along the South african coastline.

OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

## The nearest SST pixels
Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

```{r}

load("Data_P1/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data

unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# Here I use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                      query = as.matrix(site_list_sub[,5:6]), k = 1)

pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = site_list_sub$site)

# Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)
# save(OISST_match, file = "Data_P1/OISST_match.RData")

OISST_sites <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
  dplyr::rename(temp_OISST =temp)

OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
# save(OISST_sites, file = "Data_P1/OISST_sites.RData")

## Matching the Insitu data with the OISST SST data
insitu_OISST <- match_func(df = OISST_sites)
```

### 2019-08-06 SACTN_lts.Rmd
func.bits

@@
 library(zoo)
-library(lubridate)
 library(data.table)
@@
 library(gridExtra)
-library(circular)
 library(fossil)
@@
 ```{r}
+
+filter_SACTN <- function(df){
   SACTN_PN <- SACTN_US %>%
-  filter(site == "Port Nolloth") %>%
+  filter(site == df) %>%
   select(-index, -site, -src, -length) %>% 
@@
   drop_na()
+  return(SACTN_PN)
+}
 
-SACTN_LB <- SACTN_US %>%
-  filter(site == "Lamberts Bay") %>%
-  select(-index, -site, -src, -length) %>% 
-  dplyr::rename(t = date) %>% 
-  drop_na()
-
-SACTN_SB <- SACTN_US %>%
-  filter(site == "Saldanha Bay") %>%
-  select(-index, -site, -src, -length) %>% 
-  dplyr::rename(t = date) %>% 
-  drop_na()
-
-SACTN_SP <- SACTN_US %>%
-  filter(site == "Sea Point") %>%
-  select(-index, -site, -src, -length) %>% 
-  dplyr::rename(t = date) %>% 
-  drop_na()
-
-SACTN_HB <- SACTN_US %>%
-  filter(site == "Hout Bay") %>%
-  select(-index, -site, -src, -length) %>% 
-  dplyr::rename(t = date) %>% 
-  drop_na()
+SACTN_LB <- filter_SACTN(df = "Lamberts Bay") 
+SACTN_PN <- filter_SACTN(df = "Port Nolloth") 
+SACTN_SB <- filter_SACTN(df = "Saldanha Bay") 
+SACTN_SP <- filter_SACTN(df = "Sea Point") 
+SACTN_HB <- filter_SACTN(df = "Hout Bay") 
 ```

### 2019-08-06 remote_SST.Rmd
SST.update

@@
 
+# Loading libraries
+
+```{r}
+library(tidyverse)
+library(plyr)
+library(lubridate)
+library(tidyverse)
+library(plyr)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(data.table)
+library(heatwaveR)
+library(viridis)
+library(gridExtra)
+library(fossil)
+library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+library(ggpubr)
+library(ggpubr); theme_set(theme_pubr())# library(devtools)
+```
+
 ## MUR dataset
@@
 ```{r}
+load("~/Documents/Masters_2019/Upwelling_MCS/Data/SACTN_US.RData")
+
 MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
@@
 
+# Upwelling and MCS
+
+```{r}
+
+MUR_SST <- insitu_MUR %>% 
+  select(date,temp_MUR, site) %>% 
+  dplyr::rename(t = date) %>% 
+  dplyr::rename(temp = temp_MUR)
+
+filter_func <- function(df){
+  filtered <- MUR_SST  %>%  
+  filter(site == df) %>%
+  select(-site) %>% 
+  drop_na()
+  return(filtered)
+}
+
+MUR_LB <- filter_func(df = "Lamberts Bay") 
+MUR_SB <- filter_func(df = "Saldanha Bay") 
+MUR_HB <- filter_func(df = "Hout Bay") 
+MUR_PN <- filter_func(df = "Port Nolloth") 
+MUR_SP <- filter_func(df = "Sea Point")
+
+# MCS
+
+MCS_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("2002-06-01", "2015-12-30"),
+                   pctile = 10)
+  return(res_clim)
+}
+
+PN_MCS <- MCS_func(df = MUR_PN)
+LB_MCS <- MCS_func(df = MUR_LB)
+SB_MCS <- MCS_func(df = MUR_SB)
+SP_MCS <- MCS_func(df = MUR_SP)
+
+HB_MCS <- ts2clm(MUR_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
+                   pctile = 10)
+
+## Events
+MCS_event <- function(df){
+  detect <- detect_event(df, coldSpells = TRUE)
+  return(detect)
+}
+
+PN_MCS_events <- MCS_func(df = PN_MCS)
+LB_MCS_events <- MCS_func(df = LB_MCS)
+SB_MCS_events <- MCS_func(df = SB_MCS)
+SP_MCS_events <- MCS_func(df = SP_MCS)
+
+HB_MCS_events <- detect_event(HB_MCS, coldSpells = TRUE) 
+```
+
+## Upwelling events - 30th percentile used as a threshold to detect upwellling events.
+
+```{r}
+upwell_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("2002-06-01", "2015-12-30"),
+                   pctile = 30)
+  return(res_clim)
+}
+
+PN_upwell <- upwell_func(df = MUR_PN)
+LB_upwell <- upwell_func(df = MUR_LB)
+SB_upwell <- upwell_func(df = MUR_SB)
+SP_upwell <- upwell_func(df = MUR_SP)
+
+HB_upwell <- ts2clm(MUR_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
+                   pctile = 10)
+
+## Events
+upwell_event <- function(df){
+  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
+  return(detect)
+}
+
+PN_upwell_events <- MCS_func(df = PN_upwell)
+LB_upwell_events <- MCS_func(df = LB_upwell)
+SB_upwell_events <- MCS_func(df = SB_upwell)
+SP_upwell_events <- MCS_func(df = SP_upwell)
+
+HB_upwell_events <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
+```
+
 ## CMC dataset
@@
 insitu_CMC <- match_func(df = CMC)
-CMC_plot <- match_plot(df = insitu_CMC )
-CMC_temp<- temp_plot(df = insitu_CMC)
+
+CMC_SST <- insitu_CMC %>% 
+  select(date,temp_CMC, site) %>% 
+  dplyr::rename(t = date) %>% 
+  dplyr::rename(temp = temp_CMC)
+
+filter_func <- function(df){
+  filtered <- CMC_SST  %>%  
+  filter(site == df) %>%
+  select(-site) %>% 
+  drop_na()
+  return(filtered)
+}
+
+CMC_LB <- filter_func(df = "Lamberts Bay") 
+CMC_SB <- filter_func(df = "Saldanha Bay") 
+CMC_HB <- filter_func(df = "Hout Bay") 
+CMC_PN <- filter_func(df = "Port Nolloth") 
+CMC_SP <- filter_func(df = "Sea Point")
+
+# MCS
+
+MCS_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("1991-09-01", "2015-12-30"),
+                   pctile = 10)
+  return(res_clim)
+}
+
+PN_MCS <- MCS_func(df = CMC_PN)
+LB_MCS <- MCS_func(df = CMC_LB)
+SB_MCS <- MCS_func(df = CMC_SB)
+SP_MCS <- MCS_func(df = CMC_SP)
+
+HB_MCS <- ts2clm(CMC_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
+                   pctile = 10)
+
+## Events
+MCS_event <- function(df){
+  detect <- detect_event(df, coldSpells = TRUE)
+  return(detect)
+}
+
+PN_MCS_events <- MCS_func(df = PN_MCS)
+LB_MCS_events <- MCS_func(df = LB_MCS)
+SB_MCS_events <- MCS_func(df = SB_MCS)
+SP_MCS_events <- MCS_func(df = SP_MCS)
+
+HB_MCS_events <- detect_event(HB_MCS, coldSpells = TRUE) 
 ```

### 2019-08-06 remote_SST.Rmd
updated

@@
 
-PN_MCS_events <- MCS_func(df = PN_MCS)
-LB_MCS_events <- MCS_func(df = LB_MCS)
-SB_MCS_events <- MCS_func(df = SB_MCS)
-SP_MCS_events <- MCS_func(df = SP_MCS)
+PN_MCS_events_MUR <- MCS_func(df = PN_MCS)
+LB_MCS_events_MUR <- MCS_func(df = LB_MCS)
+SB_MCS_events_MUR <- MCS_func(df = SB_MCS)
+SP_MCS_events_MUR <- MCS_func(df = SP_MCS)
 
-HB_MCS_events <- detect_event(HB_MCS, coldSpells = TRUE) 
+HB_MCS_events_MUR <- detect_event(HB_MCS, coldSpells = TRUE) 
 ```
@@
 
-PN_upwell_events <- MCS_func(df = PN_upwell)
-LB_upwell_events <- MCS_func(df = LB_upwell)
-SB_upwell_events <- MCS_func(df = SB_upwell)
-SP_upwell_events <- MCS_func(df = SP_upwell)
+PN_upwell_events_MUR <- MCS_func(df = PN_upwell)
+LB_upwell_events_MUR <- MCS_func(df = LB_upwell)
+SB_upwell_events_MUR <- MCS_func(df = SB_upwell)
+SP_upwell_events_MUR <- MCS_func(df = SP_upwell)
 
-HB_upwell_events <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
+HB_upwell_events_MUR <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
 ```
@@
 
-PN_MCS_events <- MCS_func(df = PN_MCS)
-LB_MCS_events <- MCS_func(df = LB_MCS)
-SB_MCS_events <- MCS_func(df = SB_MCS)
-SP_MCS_events <- MCS_func(df = SP_MCS)
+PN_MCS_events_CMC <- MCS_func(df = PN_MCS)
+LB_MCS_events_CMC <- MCS_func(df = LB_MCS)
+SB_MCS_events_CMC <- MCS_func(df = SB_MCS)
+SP_MCS_events_CMC <- MCS_func(df = SP_MCS)
 
-HB_MCS_events <- detect_event(HB_MCS, coldSpells = TRUE) 
+HB_MCS_events_CMC <- detect_event(HB_MCS, coldSpells = TRUE) 
+```
+
+## Upwelling events - 30th percentile used as a threshold to detect upwellling events.
+
+```{r}
+upwell_func <- function(df){
+  res_clim <- ts2clm(df, climatologyPeriod = c("1991-09-01", "2015-12-30"),
+                   pctile = 30)
+  return(res_clim)
+}
+
+PN_upwell <- upwell_func(df = CMC_PN)
+LB_upwell <- upwell_func(df = CMC_LB)
+SB_upwell <- upwell_func(df = CMC_SB)
+SP_upwell <- upwell_func(df = CMC_SP)
+
+HB_upwell <- ts2clm(CMC_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
+                   pctile = 10)
+
+## Events
+upwell_event <- function(df){
+  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
+  return(detect)
+}
+
+PN_upwell_events_CMC <- MCS_func(df = PN_upwell)
+LB_upwell_events_CMC <- MCS_func(df = LB_upwell)
+SB_upwell_events_CMC <- MCS_func(df = SB_upwell)
+SP_upwell_events_CMC <- MCS_func(df = SP_upwell)
+
+HB_upwell_events_CMC <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
 ```

### 2019-08-06 remote_SST.Rmd
OISST

@@
 library(ggpubr); theme_set(theme_pubr())# library(devtools)
+library(FNN)
+```
+
+# Plots
+
+```{r}
+OISST %>%
+  filter(date == min(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = temp))
 ```
@@
 Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites. 
-
 ```{r}
-
-load("Data_P1/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
+load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_sub.Rdata")
 
@@
   filter(combi %in% pixel_match$combi)
-# save(OISST_match, file = "Data_P1/OISST_match.RData")
+# save(OISST_match, file = "Data/OISST_match.RData")
 
@@
   drop_na()
-# save(OISST_sites, file = "Data_P1/OISST_sites.RData")
+# save(OISST_sites, file = "Data/OISST_sites.RData")
+```
 
+```{r}
+load("~/Documents/Masters_2019/Upwelling_MCS/Data/OISST_sites.RData")
 ## Matching the Insitu data with the OISST SST data

### 2019-08-06 Chlorophyll.Rmd
Chlorophyll

@@
 load("Data/MODIS_Chloro.RData") # Extracted chlorophyll data
-load("Data/site_list_sub.RData") # Site list of all the sites along the west coast
+load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_sub.Rdata") # Site list of all the sites along the west coast
 
@@
 # 
-# chloro_sites <- chloro_match %>%
-#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
-#   dplyr::mutate(date = as.Date(date))
-# # save(chloro_sites, file = "Data_P1/chloro_sites.RData")
-# 
+chloro_sites <- chloro_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
+  dplyr::mutate(date = as.Date(date))
+save(chloro_sites, file = "Data/chloro_sites.RData")
+
 # # Visualisation
-# load("Data_P1/chloro_sites.RData")
+load("Data/chloro_sites.RData")
 # Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
@@
 # Chloro_plot_complete
-# 
-# # Study period from 2010 - 2014
-# chloro_data <- chloro_sites %>% 
-#   filter(year(date) %in% seq(2010, 2014))
-# # save(chloro_data, file = "Data_P1/chloro_data.RData")
-# 
-# # Plotting only years 2014-2015
-# load("Data_P1/chloro_data.RData")
-# Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
-#   geom_line() +
-#   facet_wrap(~site, ncol = 1)
-# Chloro_plot_filtered
 
 # Creating daily temperatures
-load("Data_P1/chloro_data.RData") # Loading in the chlorophyll-a data
-
-daily_chloro <- chloro_data %>%
+daily_chloro <- chloro_sites %>%
   # dplyr::mutate(date = as.Date(date)) %>% 

### 2019-08-07 SACTN_lts.Rmd
SST.update
Added some code on how to perform a couple of additional tasks.

@@
 # Detectig frequency and intensity of upwelling events - Use MHW algorithm 
-  Long time series of 30years
+  Long time series of 30 years
   Use OISST 
@@
 library(tidyverse)
-library(plyr)
 library(lubridate)
-library(ggpubr)
+library(ggpubr); theme_set(theme_pubr())
 library(zoo)
@@
 library(fossil)
-library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
 library(stringr)
@@
 library(xtable)
-library(ggpubr)
-library(ggpubr); theme_set(theme_pubr())# library(devtools)
 
+# Disable scientific notation for numeric values
+# I just find it annoying
+options(scipen = 999)
 ```
@@
 ```{r}
-load("~/Documents/Masters_2019/Upwelling_MCS/Data/SACTN_daily_v4.2.RData")
-load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_v4.2.RData")
+load("Data/SACTN_daily_v4.2.RData")
+load("Data/site_list_v4.2.RData")
 
@@
 
-# Long time series in_situ temperature
-  First creating datasets for each region
-  Now I determine the frequency and intensity of MCS using the MHW algorithm
+# Long _in situ_ temperature time series
+
+Determine the frequency and intensity of MCS using the MHW algorithm
 
 ```{r}
+# Use purrr::nest() to allow the use of purrr::map() to grouped data
+SACTN_MCS <- SACTN_US %>% 
+  dplyr::rename(t = date) %>%
+  group_by(site) %>% 
+  nest() %>% 
+  mutate(clims = purrr::map(data, ts2clm,
+                          # NB: A 30 year period should be used for a climatology
+                          # Conveniently, this period works with all of the time series
+                          climatologyPeriod = c("1974-01-01", "2003-12-31"),  pctile = 10),
+         events = purrr::map(clims, detect_event, coldSpells = TRUE)) %>% 
+  select(-data, -clims)
 
-filter_SACTN <- function(df){
-  SACTN_PN <- SACTN_US %>%
-  filter(site == df) %>%
-  select(-index, -site, -src, -length) %>% 
-  dplyr::rename(t = date) %>% 
-  drop_na()
-  return(SACTN_PN)
-}
+# Unnest the climatology results
+SACTN_MCS_clims <- SACTN_MCS %>% 
+  unnest(events) %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest(events)
 
-SACTN_LB <- filter_SACTN(df = "Lamberts Bay") 
-SACTN_PN <- filter_SACTN(df = "Port Nolloth") 
-SACTN_SB <- filter_SACTN(df = "Saldanha Bay") 
-SACTN_SP <- filter_SACTN(df = "Sea Point") 
-SACTN_HB <- filter_SACTN(df = "Hout Bay") 
-```
-
-  The rate of onset of MCS can be determined here
-
-```{r}
-
-#' # Or if one wants to calculate MCSs
-
-MCS_func <- function(df){
-  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 10)
-  return(res_clim)
-}
-
-PN_MCS <- MCS_func(df = SACTN_PN)
-LB_MCS <- MCS_func(df = SACTN_LB)
-SB_MCS <- MCS_func(df = SACTN_SB)
-SP_MCS <- MCS_func(df = SACTN_SP)
-
-res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
-                   pctile = 10)
-
-## Events
-MCS_event <- function(df){
-  detect <- detect_event(df, coldSpells = TRUE)
-  return(detect)
-}
-
-PN_MCS_events <- MCS_func(df = PN_MCS)
-LB_MCS_events <- MCS_func(df = LB_MCS)
-SB_MCS_events <- MCS_func(df = SB_MCS)
-SP_MCS_events <- MCS_func(df = SP_MCS)
-
-HB_MCS_events <- detect_event(res_clim, coldSpells = TRUE) # line 118 -
+# Unnest the event results
+SACTN_MCS_events <- SACTN_MCS %>% 
+  unnest(events) %>%
+  filter(row_number() %% 2 == 0) %>%
+  unnest(events)
 ```
@@
 ```{r}
-upwell_func <- function(df){
-  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
-                   pctile = 30)
-  return(res_clim)
+SACTN_upwell <- SACTN_US %>% 
+  dplyr::rename(t = date) %>%
+  group_by(site) %>% 
+  nest() %>% 
+  mutate(clims = purrr::map(data, ts2clm,
+                          # NB: A 30 year period should be used for a climatology
+                          # Conveniently, this period works with all of the time series
+                          climatologyPeriod = c("1974-01-01", "2003-12-31"),  pctile = 30),
+         events = purrr::map(clims, detect_event, coldSpells = TRUE, minDuration = 2)) %>% 
+  select(-data, -clims)
+
+# Unnest the climatology results
+SACTN_upwell_clims <- SACTN_upwell %>% 
+  unnest(events) %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest(events)
+
+# Unnest the event results
+SACTN_upwell_events <- SACTN_upwell %>% 
+  unnest(events) %>%
+  filter(row_number() %% 2 == 0) %>%
+  unnest(events)
+```
+
+# Test visuals
+
+The figures below compare the largest event detected with the classic MCS definition vs.
+the upwelling definition used here (i.e.
+30th perc.
+thresh.
+and 2 day minimum).
+This isn't very usfeul...
+but we do see that only 2 of the 5 time series have largest events match up with the different definitions.
+I'm not sure that is useful though...
+
+```{r}
+# Port Nolloth
+ggarrange(event_line(SACTN_MCS$events[[1]]),
+          event_line(SACTN_upwell$events[[1]]), ncol = 1, nrow = 2)
+
+# Lamberts Bay
+ggarrange(event_line(SACTN_MCS$events[[2]]),
+          event_line(SACTN_upwell$events[[2]]), ncol = 1, nrow = 2)
+
+# Saldanha Bay
+ggarrange(event_line(SACTN_MCS$events[[3]]),
+          event_line(SACTN_upwell$events[[3]]), ncol = 1, nrow = 2)
+
+# Sea Point
+ggarrange(event_line(SACTN_MCS$events[[4]]),
+          event_line(SACTN_upwell$events[[4]]), ncol = 1, nrow = 2)
+
+# Hout Bay
+ggarrange(event_line(SACTN_MCS$events[[5]]),
+          event_line(SACTN_upwell$events[[5]]), ncol = 1, nrow = 2)
+```
+
+The following figures show each of the 5 long time series with the MCS days highlighted in blue, and the upwelling days shown with a rug plot.
+We see that the 30th percentile threshold shows near constant upwelling, which is likely not accurate.
+A more selective threshold should probably be used, but this will depned on what the upweliing index results show.
+
+```{r}
+ggplot(data = SACTN_MCS_clims, aes(x = t, y = temp)) +
+  geom_line() +
+  geom_point(data = filter(SACTN_MCS_clims, event == TRUE), colour = "navy") +
+  geom_rug(data = filter(SACTN_upwell_clims, event == TRUE), sides = "b") +
+  scale_x_date(expand = c(0, 0)) +
+  facet_wrap(~site, ncol = 1) +
+  theme_grey() +
+  labs(y = "Temperature (°C)", x = NULL)
+```
+
+# Linear regressions
+
+Below we use `purrr::map()` to apply a linear regression to every variable from our MCS and upwelling results.
+These results may need some thinking.
+I'm not certain that the negative values inherent in the MCS results are being handled correctly here.
+
+```{r}
+# Convenience function for below
+lm_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
 }
 
-PN_upwell <- upwell_func(df = SACTN_PN)
-LB_upwell <- upwell_func(df = SACTN_LB)
-SB_upwell <- upwell_func(df = SACTN_SB)
-SP_upwell <- upwell_func(df = SACTN_SP)
+# Changes in MCS metrics
+SACTN_MCS_lm <- SACTN_MCS_events %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
 
-res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
-                   pctile = 10)
-
-## Events
-upwell_event <- function(df){
-  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
-  return(detect)
-}
-
-PN_upwell_events <- MCS_func(df = PN_upwell)
-LB_upwell_events <- MCS_func(df = LB_upwell)
-SB_upwell_events <- MCS_func(df = SB_upwell)
-SP_upwell_events <- MCS_func(df = SP_upwell)
-
-HB_upwell_events <- detect_event(res_clim, minDuration = 2, coldSpells = TRUE) # line 1
+# Changes in upwelling metrics
+SACTN_upwell_lm <- SACTN_upwell_events %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
 ```

### 2019-08-09 upwell_IDX.Rmd
upwell_IDX

---
title: "upwell_IDX"
author: "Amieroh Abrahams"
date: "08 August 2019"
output: html_document
---

# Upwelling

Upwelling is primarily caused by alongshore, equator ward winds. These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods. Upwelling is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.

# Upwelling indeces
  # Determining upwelling index from wind data (SAWS)
  # Index Equation from Fielding & Davis 1989 paper

$$ UpwellingIndex = μ{(Cosθ − 160)}$$

In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The 160 degrees is used as this refers to the angle of the coastline. This equation is largely dependant on wind speed and direction data in order to determing the intensity of the upwelling event. Wind data were obtained daily from the South African Weather Service (SAWS). This wind data was then matched to the date at which temperature were collected. With this data the upwelling index was determined.

# libraries

```{r}
library(tidyverse)
library(ggplot2)
library(circular)
```

# Steps
  - Use the circular function 
  - Determine daily wind data obtained from 3hr intervals

```{r}
# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
# 
# wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# # Slecting the important columns for each of the datasets
# wind_fix <- function(df){
# wind <- df %>%
#   select(station_name, date, hour, dir, speed) %>%  #column names
#   mutate(date = as.Date(as.character(date)),
#          hour = as.numeric(as.character(hour)),
#          dir = as.numeric(as.character(dir)),
#          speed = as.numeric(as.character(speed)))
# }
# 
# wind_1 <- wind_fix(df = wind_1)
# wind_2 <- wind_fix(df = wind_2)
# wind_3 <- wind_fix(df = wind_3)
# 
# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
# 
# renaming_sites_1 <- function(df) {
#   sites <- df %>%
#       mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
#                                     station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
#                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
#                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
#                                     station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
#   return(sites)
# }
# 
# 
# wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
# wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
# wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
# wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
# # save(wind_data, file = "Data/wind_data.RData")
# 
# load("Data/wind_data.RData")
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
# 
# wind_daily <- wind_data %>%
#   # ungroup() %>%
#   # mutate(date = as.Date(date))%>%
#   dplyr::group_by(temp_sites, date) %>%
#   filter(temp_sites %in% selected_sites) %>%
#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
#                    mean_speed = round(mean(speed),2))
# #save(wind_daily, file = "Data/wind_daily.RData")
```

# Load wind

```{r}
load("Data/wind_daily.RData") 
wind_daily <- wind_daily %>% 
  dplyr::rename(sites = temp_sites)
```

# Determining the upwelling indeces

```{r}

wind_UI <- wind_daily %>% 
  group_by(sites) %>% 
  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # Angle?
  drop_na

index <- wind_UI %>% 
  group_by(sites) %>% 
  select("date", "ui.saws") %>% 
  mutate(saws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))

```

# Visualisiing the data

```{r}
plot1 <- ggplot(data = index, aes(x = date, y = ui.saws)) +
  geom_line() +
  facet_wrap(~sites, ncol = 1) +
  geom_area(aes( fill = saws.condition)) +
  theme_grey()
plot1
```

# MCS and upwelling indeces plot

```{r}
ggplot(data = SACTN_MCS_clims, aes(x = t, y = temp)) +
  geom_line() +
  geom_point(data = filter(SACTN_MCS_clims, event == TRUE), colour = "navy") +
  geom_rug(data = filter(index, ui.saws == TRUE), sides = "b") +
  scale_x_date(expand = c(0, 0)) +
  facet_wrap(~site, ncol = 1) +
  theme_grey() +
  labs(y = "Temperature (°C)", x = NULL)
```

### 2019-08-09 upwell_IDX.Rmd
update

@@
 ```{r}
-ggplot(data = SACTN_MCS_clims, aes(x = t, y = temp)) +
+match_func <- function(df){
+  match <- SACTN_MCS_clims  %>% 
+  left_join(df, by = c("site", "t")) %>% 
+  na.trim()
+  return(match)
+}
+
+index <- index %>% 
+  dplyr::rename(site = sites) %>% 
+  dplyr::rename(t = date)
+
+match <- match_func(df = index)
+
+upwell_MCS <- ggplot(data = match, aes(x = t, y = temp)) +
   geom_line() +
-  geom_point(data = filter(SACTN_MCS_clims, event == TRUE), colour = "navy") +
-  geom_rug(data = filter(index, ui.saws == TRUE), sides = "b") +
+  geom_point(data = filter(match, event == TRUE), colour = "navy") +
+  geom_rug(data = filter(match, ui.saws == TRUE), sides = "b") +
   scale_x_date(expand = c(0, 0)) +
@@
   labs(y = "Temperature (°C)", x = NULL)
+upwell_MCS
 ```

### 2019-08-09 upwell_IDX.Rmd
fixed

@@
 
-upwell_MCS <- ggplot(data = match, aes(x = t, y = temp)) +
+upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
   geom_line() +
   geom_point(data = filter(match, event == TRUE), colour = "navy") +
-  geom_rug(data = filter(match, ui.saws == TRUE), sides = "b") +
+  geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
   scale_x_date(expand = c(0, 0)) +
@@
 upwell_MCS
+
 ```

### 2019-08-09 upwell_IDX.Rmd
UI

@@
 library(circular)
+# devtools::install_github("robwschlegel/coastR")
+library(coastR)
+library(gridExtra)
+library(dplyr)
+library(geosphere)
 ```
@@
 
-# Determining the upwelling indeces
+# Determining the angle perpendicular to the coastline and plotting it
+
+```{r}
+south_africa_away_wide <- transects(site_list_sub, spread = 30)
+
+world_map <- ggplot() + 
+  borders(fill = "grey40", colour = "black")
+
+# Create titles
+titles <- c("Shore-normal", "Islands")
+
+# Plotting function
+plot_sites <- function(site_list, buffer, title_choice, dist){
+
+  # Find the point 200 km from the site manually to pass to ggplot
+  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
+                                              b = site_list$heading, d = dist))
+
+  # Add the new coordinates tot he site list
+  site_list <- site_list %>% 
+    mutate(lon_dest = heading2$lon,
+           lat_dest = heading2$lat)
+
+  # Visualise
+  world_map +
+    geom_segment(data = site_list, colour = "red4", 
+                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
+    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
+    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
+    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
+                             max(site_list$lon + buffer)),
+                    ylim = c(min(site_list$lat - buffer), 
+                             max(site_list$lat + buffer))) +
+    labs(x = "", y = "", colour = "Site\norder") +
+    ggtitle(titles[title_choice])
+}
+
+south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
+```
+
+# Determining the upwelling indeces by changing the angle for each of the sites
 
@@
 
+wind_angle <- wind_daily %>% 
+  dplyr::rename(site = sites) %>% 
+  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
+  dplyr::rename(coast_angle = heading)
+
+##### Do this
 wind_UI <- wind_daily %>% 
-  group_by(sites) %>% 
-  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # Angle?
+  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% 
   drop_na
@@
   select("date", "ui.saws") %>% 
-  mutate(saws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))
+  mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
 

### 2019-08-12 upwell_IDX.Rmd
working process

@@
 
-# Determining the upwelling indeces by changing the angle for each of the sites
+You can then use the exceedence() function from heatwaveR to check for consecutive numbers of days at or above whatever the UI value is meant to be at 
+Sounds good. Good luck!
 
@@
   left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
-  dplyr::rename(coast_angle = heading)
-
-##### Do this
-wind_UI <- wind_daily %>% 
+  dplyr::rename(coast_angle = heading) %>% 
   mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% 
@@
 
-index <- wind_UI %>% 
-  group_by(sites) %>% 
-  select("date", "ui.saws") %>% 
-  mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-
+# index <- wind_UI %>% 
+#   group_by(sites) %>% 
+#   select("date", "ui.saws") %>% 
+#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
 ```

### 2019-08-13 upwell_IDX.Rmd
progress

--- 2019-08-09 upwell_IDX.Rmd
@@
 library(geosphere)
+library(heatwaveR)
 ```
@@
 
-# Determining the upwelling indeces by changing the angle for each of the sites
+```{r}
+UI_angle <- wind_daily %>% 
+  dplyr::rename(site = sites) %>% 
+  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
+  dplyr::rename(coast_angle = heading) %>% 
+  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
+  drop_na
+
+# index <- wind_UI %>%
+#   group_by(site) %>%
+#   select("date", "ui.saws") %>%
+#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
+
+wind_UI <- wind_daily %>% 
+  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
+  drop_na
+
+# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
+```
+
+# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is ment to be
 
 ```{r}
+# Loading the insitu temperature data along the wc
+load("Data/SACTN_US.RData")
 
-wind_angle <- wind_daily %>% 
-  dplyr::rename(site = sites) %>% 
-  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
-  dplyr::rename(coast_angle = heading)
+# exceedance <-
+#   function(data,
+#            x = t,
+#            y = temp,
+#            threshold,
+#            below = FALSE,
+#            minDuration = 5,
+#            joinAcrossGaps = TRUE,
+#            maxGap = 2,
+#            maxPadLength = 3)
 
-##### Do this
-wind_UI <- wind_daily %>% 
-  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% 
-  drop_na
+# Upwelling temperature threshold 12.4 / 30th percentile
 
-index <- wind_UI %>% 
-  group_by(sites) %>% 
-  select("date", "ui.saws") %>% 
-  mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
+SACTN_upwell <- SACTN_US %>% 
+  dplyr::rename(t = date) %>%
+  group_by(site) %>%
+  nest() %>% 
+  mutate(exceed = purrr::map(data, exceedance, threshold = 12.4, below = TRUE))
 
-```
+# SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
 
-# Visualisiing the data
-
-```{r}
-plot1 <- ggplot(data = index, aes(x = date, y = ui.saws)) +
-  geom_line() +
-  facet_wrap(~sites, ncol = 1) +
-  geom_area(aes( fill = saws.condition)) +
-  theme_grey()
-plot1
 ```
@@
 ```{r}
-match_func <- function(df){
-  match <- SACTN_MCS_clims  %>% 
-  left_join(df, by = c("site", "t")) %>% 
-  na.trim()
-  return(match)
-}
-
-index <- index %>% 
-  dplyr::rename(site = sites) %>% 
-  dplyr::rename(t = date)
-
-match <- match_func(df = index)
-
-upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
-  geom_line() +
-  geom_point(data = filter(match, event == TRUE), colour = "navy") +
-  geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
-  scale_x_date(expand = c(0, 0)) +
-  facet_wrap(~site, ncol = 1) +
-  theme_grey() +
-  labs(y = "Temperature (°C)", x = NULL)
-upwell_MCS
+# match_func <- function(df){
+#   match <- SACTN_MCS_clims  %>% 
+#   left_join(df, by = c("site", "t")) %>% 
+#   na.trim()
+#   return(match)
+# }
+# 
+# index <- index %>% 
+#   dplyr::rename(site = sites) %>% 
+#   dplyr::rename(t = date)
+# 
+# match <- match_func(df = index)
+# 
+# upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
+#   geom_line() +
+#   geom_point(data = filter(match, event == TRUE), colour = "navy") +
+#   geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
+#   scale_x_date(expand = c(0, 0)) +
+#   facet_wrap(~site, ncol = 1) +
+#   theme_grey() +
+#   labs(y = "Temperature (°C)", x = NULL)
+# upwell_MCS
 

### 2019-08-13 upwell_IDX.Rmd
Tweaking

@@
 library(tidyverse)
-library(ggplot2)
 library(circular)
@@
 library(gridExtra)
-library(dplyr)
 library(geosphere)
@@
   dplyr::rename(sites = temp_sites)
+load("Data/site_list_sub.Rdata")
 ```
@@
 
-# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is ment to be
+# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be
 
@@
 
-SACTN_upwell <- SACTN_US %>% 
-  dplyr::rename(t = date) %>%
+SACTN_upwell <- UI_angle %>% 
+  dplyr::rename(t = date,
+                temp = ui.saws) %>%
   group_by(site) %>%
+  # mutate(min_t = min(t), 
+  #        max_t = max(t)) %>% 
   nest() %>% 
-  mutate(exceed = purrr::map(data, exceedance, threshold = 12.4, below = TRUE))
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
+         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 3)) %>% 
+  # mutate() %>% 
+  select(-data, -clim) %>% 
+  unnest()
 

### 2019-08-13 upwell_IDX.Rmd
A bit more

@@
   select(-data, -clim) %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
   unnest()
@@
 # SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
+
+# SACTN_upwell_metrics
 

### 2019-08-13 upwell_IDX.Rmd
Worked out a functioning chunk of code for combining the MCS algorithm with the upwelling index

--- 2019-08-13 upwell_IDX.Rmd
@@
 ```{r}
-library(tidyverse)
 library(circular)
 # devtools::install_github("robwschlegel/coastR")
-library(coastR)
 library(gridExtra)
 library(geosphere)
+library(tidyverse)
 library(heatwaveR)
+library(coastR)
 ```
@@
 ```{r}
-south_africa_away_wide <- transects(site_list_sub, spread = 30)
+south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
 
@@
   mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
-  drop_na
+  drop_na %>% 
+  dplyr::rename(t = date)
 
@@
 SACTN_upwell <- UI_angle %>% 
-  dplyr::rename(t = date,
-                temp = ui.saws) %>%
+  dplyr::rename(temp = ui.saws) %>%
   group_by(site) %>%
@@
   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
-         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 3)) %>% 
+         # NB: A threshold of 3 appeared to be far to strict
+         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 1)) %>% 
   # mutate() %>% 
   select(-data, -clim) %>% 
-  unnest()
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% 
+  dplyr::rename(ui.saws = temp) %>% 
+  select(site, t, ui.saws, exceedance)
 
@@
 
+# There are a few ideas on how to go about doing this
+# The first is to calculatea rolling climatology and to check the upwelling phenology agianst that
+# in order to determine changes in upwelling values, but this won't give us the metrics we want
+# Rather what we can do is set a static bottom threshold below which we are interested in looking for upwelling
+# We then use the upwelling index values created above as a second filter to determine the metrics of the upwelling
+
+# Unfortunately the way that the heatwaveR functions work well does with purrr. 
+# I've figured this out once before but can't remember where now.
+# Rather than go through the process I have rather just made a convenience function below that
+# does what I want and can simply be given to purrr::map() without any faffing about.
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
+  return(res)
+}
+
+# Calculate the upwelling event metrics
+SACTN_upwell_base <- SACTN_US %>% 
+  dplyr::rename(t = date) %>% 
+  group_by(site) %>% 
+  nest() %>% 
+  # The climatology base period used here is up for debate...
+  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
+  mutate(clim = purrr::map(data, ts2clm, pctile = 30, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
+  select(-data) %>% 
+  unnest() %>%
+  left_join(SACTN_upwell, by = c("site", "t")) %>%
+  filter(!is.na(exceedance)) %>%
+  group_by(site) %>% 
+  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
+  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
+  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
+  nest() %>% 
+  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
+  select(-data) #%>% 
+  # unnest() %>% 
+  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
+  # filter(row_number() %% 2 == 1) %>% # Select daily values
+  # unnest()
+
+# Unpack the event metric reults
+SACTN_upwell_metrics <- SACTN_upwell_base %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 0) %>%
+  unnest()
+
+# Unpack the daily climatology results
+SACTN_upwell_clims <- SACTN_upwell_base %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>% 
+  unnest()
+
+# The above chunk of code appears to work as expected.
+# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn.
+# What also needs to be decided is if we are interested in any upwelling results throughout the year,
+# or only during the upwelling season, whenever that may be for each site.
+# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
+# is changing over time.
+# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
 ```
@@
 
+# Plot some of the Upwelling/MCS results
+SP_upwelling <- SACTN_upwell_clims %>% 
+  filter(site == "Sea Point", lubridate::year(t) %in% 2016:2018)
+
+ggplot(SP_upwelling, aes(x = t, y = temp)) +
+  geom_flame(aes(y = thresh, y2 = temp), fill = "navy", n = 3) +
+  geom_point(data = filter(SP_upwelling, !is.na(event_no))) +
+  geom_line(aes(y = thresh)) +
+  geom_line(alpha = 0.7) +
+  geom_rug(data = filter(SP_upwelling, event_no > 0), sides = "b")
+
+# We can see in the figure above that the rug plot, which denotes when the wind driven upwelling was detected,
+# occurrs much less frequently than simply using the 30th percentile threshold.
+# The dots on the plot help to show the length of each wind driven upwelling event detected.
+# This supports the use of this methodology.
+
+# Up next is fitting linear models to the upwelling metric results to see what trends shake out.
 ```

### 2019-08-14 upwell_IDX.Rmd
fix merge

--- 2019-08-13 upwell_IDX.Rmd
@@
 ```{r}
-south_africa_away_wide <- transects(site_list_sub, spread = 30)
-
+south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
 world_map <- ggplot() + 
   borders(fill = "grey40", colour = "black")
-
 # Create titles
 titles <- c("Shore-normal", "Islands")
-
 # Plotting function
@@
 }
-
 south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
@@
   mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
-  drop_na
-
+  drop_na %>% 
+  dplyr::rename(t = date)
 # index <- wind_UI %>%
@@
 #   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-
 wind_UI <- wind_daily %>% 
@@
   drop_na
-
 # Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
@@
 load("Data/SACTN_US.RData")
-
 # exceedance <-
@@
 #            maxPadLength = 3)
-
 # Upwelling temperature threshold 12.4 / 30th percentile
-
 SACTN_upwell <- UI_angle %>% 
-  dplyr::rename(t = date,
-                temp = ui.saws) %>%
+  dplyr::rename(temp = ui.saws) %>%
   group_by(site) %>%
@@
   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
-         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 3)) %>% 
+         # NB: A threshold of 3 appeared to be far to strict
+         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 1)) %>% 
   # mutate() %>% 
   select(-data, -clim) %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% 
+  dplyr::rename(ui.saws = temp) %>% 
+  select(site, t, ui.saws, exceedance)
+# SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
+# There are a few ideas on how to go about doing this
+# The first is to calculatea rolling climatology and to check the upwelling phenology agianst that
+# in order to determine changes in upwelling values, but this won't give us the metrics we want
+# Rather what we can do is set a static bottom threshold below which we are interested in looking for upwelling
+# We then use the upwelling index values created above as a second filter to determine the metrics of the upwelling
+# Unfortunately the way that the heatwaveR functions work well does with purrr. 
+# I've figured this out once before but can't remember where now.
+# Rather than go through the process I have rather just made a convenience function below that
+# does what I want and can simply be given to purrr::map() without any faffing about.
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
+  return(res)
+}
+# Calculate the upwelling event metrics
+SACTN_upwell_base <- SACTN_US %>% 
+  dplyr::rename(t = date) %>% 
+  group_by(site) %>% 
+  nest() %>% 
+  # The climatology base period used here is up for debate...
+  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
+  mutate(clim = purrr::map(data, ts2clm, pctile = 30, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
+  select(-data) %>% 
+  unnest() %>%
+  left_join(SACTN_upwell, by = c("site", "t")) %>%
+  filter(!is.na(exceedance)) %>%
+  group_by(site) %>% 
+  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
+  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
+  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
+  nest() %>% 
+  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
+  select(-data) #%>% 
+  # unnest() %>% 
+  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
+  # filter(row_number() %% 2 == 1) %>% # Select daily values
+  # unnest()
+# Unpack the event metric reults
+SACTN_upwell_metrics <- SACTN_upwell_base %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 0) %>%
   unnest()
-
-# SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
-
+# Unpack the daily climatology results
+SACTN_upwell_clims <- SACTN_upwell_base %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>% 
+  unnest()
+# The above chunk of code appears to work as expected.
+# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn.
+# What also needs to be decided is if we are interested in any upwelling results throughout the year,
+# or only during the upwelling season, whenever that may be for each site.
+# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
+# is changing over time.
+# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
 ```
@@
 # upwell_MCS
+# Plot some of the Upwelling/MCS results
+SP_upwelling <- SACTN_upwell_clims %>% 
+  filter(site == "Sea Point", lubridate::year(t) %in% 2016:2018)
+ggplot(SP_upwelling, aes(x = t, y = temp)) +
+  geom_flame(aes(y = thresh, y2 = temp), fill = "navy", n = 3) +
+  geom_point(data = filter(SP_upwelling, !is.na(event_no))) +
+  geom_line(aes(y = thresh)) +
+  geom_line(alpha = 0.7) +
+  geom_rug(data = filter(SP_upwelling, event_no > 0), sides = "b")
 
+# We can see in the figure above that the rug plot, which denotes when the wind driven upwelling was detected,
+# occurrs much less frequently than simply using the 30th percentile threshold.
+# The dots on the plot help to show the length of each wind driven upwelling event detected.
+# This supports the use of this methodology.
+# Up next is fitting linear models to the upwelling metric results to see what trends shake out.
 ```

### 2019-08-14 upwell_IDX.Rmd
ud

@@
          # NB: A threshold of 3 appeared to be far to strict
-         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 1)) %>% 
+         exceed = purrr::map(clim, exceedance, minDuration = 2, threshold = 1)) %>% 
   # mutate() %>% 
@@
 detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 2, coldSpells = T)
   return(res)
@@
   # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
-  mutate(clim = purrr::map(data, ts2clm, pctile = 30, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
+  mutate(clim = purrr::map(data, ts2clm, pctile = 20, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
   select(-data) %>% 

### 2019-08-15 upwell_IDX.Rmd
updating

@@
   dplyr::rename(t = date)
+
+# save(UI_angle, file = "Data/UI_angle.RData")
+
 # index <- wind_UI %>%
@@
 #   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-wind_UI <- wind_daily %>% 
-  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
-  drop_na
+
+# wind_UI <- wind_daily %>% 
+#   mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
+#   drop_na
 # Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
@@
          # NB: A threshold of 3 appeared to be far to strict
-         exceed = purrr::map(clim, exceedance, minDuration = 2, threshold = 1)) %>% 
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>% 
   # mutate() %>% 
@@
   select(site, t, ui.saws, exceedance)
+
 # SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
@@
 # does what I want and can simply be given to purrr::map() without any faffing about.
+
 detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 2, coldSpells = T)
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
   return(res)
@@
   # unnest()
+
 # Unpack the event metric reults
-SACTN_upwell_metrics <- SACTN_upwell_base %>% 
+SACTN_upwell_events <- SACTN_upwell_base %>% 
   unnest() %>%
@@
   unnest()
+
 # The above chunk of code appears to work as expected.
@@
 # Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
+
+# Linear Models
+lm_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
+}
+
+# Changes in upwelling metrics
+SACTN_upwell_lm <- SACTN_upwell_events %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+
 ```
@@
 # Up next is fitting linear models to the upwelling metric results to see what trends shake out.
+
 ```
+
+Increase in upwelling over the past years as a result of intensified winds over time?
+Weakening of th upwelling over time, leads to a decrease in the upwelling indexs
+Strong south easterly winds - Spring and summer: Upwelling is more intense in spring and summer months
+The formula used suggests that if values at + upwelling occurs and if values are - then downwelling occurs. But the intensity of upwelling changes.
+Compare upwelling indeces per season
+
+```{r}
+## Adding a seasonal component to the UI_angle dataset
+
+# UI_angle_season <- UI_angle %>%
+#   mutate(month = month(t, abbr = T, label = T)) %>%
+#   mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+#                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+#                               ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+#                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+
+## save(UI_angle_season, file = "Data/UI_angle_season.RData")
+
+load("Data/UI_angle_season.RData")
+```
+
+# Show how winds intensified over the years
+# Show how upwwelling intensified over the years
+# Look at upwelling seasonally

### 2019-08-18 upwell_IDX.Rmd
updating

@@
          # NB: A threshold of 3 appeared to be far to strict
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>% 
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+Upwelling can last for a few hours.
+Given that we have daily data, upwelling events minimum duration here will be 1day
   # mutate() %>% 
@@
   unnest()
+# save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
+
 # Unpack the daily climatology results
@@
   unnest()
+# save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
 
@@
 # What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn. 
+    # Min duration for an upwelling event - 1day and threshold -1
 # What also needs to be decided is if we are interested in any upwelling results throughout the year,
@@
 # Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
-
-# Linear Models
-lm_coeff <- function(df){
-  res <- lm(formula = val ~ date_peak, data = df)
-  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
-}
-
-# Changes in upwelling metrics
-SACTN_upwell_lm <- SACTN_upwell_events %>% 
-  select(-c(index_start:index_end)) %>% 
-  gather(key = "var", value = "val", -c(site:date_end)) %>% 
-  group_by(site, var) %>% 
-  nest() %>% 
-  mutate(slope = purrr::map(data, lm_coeff)) %>% 
-  select(-data) %>% 
-  unnest() %>% 
-  # convert from daily to decadal values
-  mutate(slope = slope * 365.25*10)
-
 ```
@@
 
+# Linear Models
+lm_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
+}
+# Changes in upwelling metrics
+SACTN_upwell_lm <- SACTN_upwell_events %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
 ```
@@
 
-# UI_angle_season <- UI_angle %>%
-#   mutate(month = month(t, abbr = T, label = T)) %>%
-#   mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
-#                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
-#                               ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-#                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+season_func <- function(df){
+  season <- df %>%
+  mutate(month = month(t, abbr = T, label = T),
+         year = year(t)) %>%
+  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",
+                          ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                              ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                      ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+return(season)
+}
 
-## save(UI_angle_season, file = "Data/UI_angle_season.RData")
+# UI_angle_season <- season_func(df = UI_angle)
+# save(UI_angle_season, file = "UI_angle_season.RData")
 
 load("Data/UI_angle_season.RData")
+
+# Speed of wind increased over time???
+anova_func <- function(df){
+  season_aov <- aov(mean_speed ~ site  * year * season, data = df)
+return(season_aov)
+}
+
+wind_speed_aov <- anova_func(df = UI_angle_season)
+summary(wind_speed_aov)
+
+anova_func <- function(df){
+  season_aov <- aov(ui.saws ~ site  * year * season, data = df)
+return(season_aov)
+}
+ui.saws_aov <- anova_func(df = UI_angle_season)
+summary(ui.saws_aov)
 ```
-
-# Show how winds intensified over the years
-# Show how upwwelling intensified over the years
-# Look at upwelling seasonally

### 2019-08-20 SST_patterns.Rmd
wind

---
title: "SST_patterns"
author: "Amieroh Abrahams"
date: "19 August 2019"
output: html_document
---

# Are the same upwelling signals present at different distances from the coastline and through the different datasets

### 2019-08-20 upwell_IDX.Rmd
wind

@@
 ```
+
+# Wind speed intensifed ovwer the years
+Take first third (33percentile) of the data and the last third (66th percentile) of the data for every season and then do a t-test
+
+```{r}
+load("Data/wind_daily.RData")
+
+wind_daily <- wind_daily %>% 
+  dplyr::rename(t = date)
+
+# wind_daily_season <- season_func(df = wind_daily)
+
+#save(wind_daily_season, file = "Data/wind_daily_season.RData")
+
+load("Data/wind_daily_season.RData")
+
+wind_daily_season_HB <- wind_daily_season %>% 
+  filter(temp_sites == "Hout Bay") %>% 
+  group_by(season) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+wind_daily_season_SP <- wind_daily_season %>% 
+  filter(temp_sites == "Sea Point") %>% 
+  group_by(season) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+wind_daily_season_SB <- wind_daily_season %>% 
+  filter(temp_sites == "Saldanha Bay") %>% 
+  group_by(season) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+wind_daily_season_PN <- wind_daily_season %>% 
+  filter(temp_sites == "Port Nolloth") %>% 
+  group_by(season) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+wind_daily_season_LB <- wind_daily_season %>% 
+  filter(temp_sites == "Lamberts Bay") %>% 
+  group_by(season) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+# T-test
+t.test(wind_daily_season_HB$msp_1st, wind_daily_season_HB$mspeed_perc, paired=TRUE)
+
+# T- test between seasons for all sites
+
+```

### 2019-08-20 upwell_IDX.Rmd
fixed

--- 2019-08-18 upwell_IDX.Rmd
@@
 ```
+
+# Wind speed intensifed ovwer the years
+Take first third (33percentile) of the data and the last third (66th percentile) of the data for every season and then do a t-test
+
+```{r}
+load("Data/wind_daily.RData")
+
+wind_daily <- wind_daily %>% 
+  dplyr::rename(t = date)
+
+# wind_daily_season <- season_func(df = wind_daily)
+
+#save(wind_daily_season, file = "Data/wind_daily_season.RData")
+
+load("Data/wind_daily_season.RData")
+
+wind_daily_season_test <- wind_daily_season %>% 
+  group_by(season, temp_sites) %>% 
+  arrange(mean_speed) %>% 
+  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
+
+# T-test
+
+# T- test between seasons for all sites
+
+```

### 2019-08-21 SST_patterns.Rmd
push

@@
 
-# Are the same upwelling signals present at different distances from the coastline and through the different datasets
+# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
+
+```{r}
+knitr::opts_chunk$set(
+  comment = "R>",
+  warning = FALSE,
+  message = FALSE 
+)
+
+library(tidyverse)
+# library(plyr)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(forecast)
+library(astrochron)
+library(WaveletComp)
+library(data.table)
+library(heatwaveR)
+library(viridis)
+library(ggrepel)
+library(maptools)
+library(sp)
+library(geosphere)
+library(PBSmapping)
+library(scales)
+library(grid)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(ncdf4)
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+```
+
+```{r}
+load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
+
+SACTN_US <- SACTN_US %>% 
+  dplyr::rename(in_situ_temp = temp)
+
+# Visualising the data
+temp_plot <- function(df){
+  plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
+    geom_line(aes(group = site)) +
+    labs(x = "", y = "Temperature (°C)") +
+    theme(axis.text.x = element_text(angle = 45)) +
+    theme(legend.position = "top")
+}
+
+SACTN_plot <- temp_plot(df = SACTN_US)
+SACTN_plot
+```
+
+# Loading MUR data
+
+```{r}
+MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
+MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
+MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
+MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
+MUR_Hout_Bay <- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")
+
+MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Hout_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Sea_Point) %>%
+  dplyr::rename(site = station)
+
+MUR_SST$date <- (ymd(MUR_SST$date))
+MUR_SST <- MUR_SST %>% 
+  dplyr::rename(temp = nearest1) %>% 
+  drop_na()
+
+match_func <- function(df){
+  match <- SACTN_US  %>%  
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+MUR <- match_func(df = MUR_SST) %>%
+  drop_na()
+
+match_plot <- function(df){
+  plot1 <- df %>% 
+ggplot(aes(x = date, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 2) +
+  theme_bw()
+  return(plot1)
+}
+
+MUR_plot <- match_plot(df = MUR)
+
+```
+
+# CMC data
+
+```{r}
+CMC_Sea_Point <- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
+CMC_Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
+CMC_Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
+CMC_Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
+CMC_Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")
+
+CMC_SST <- rbind(CMC_Lamberts_Bay,CMC_Hout_Bay, CMC_Port_Nolloth, CMC_Saldanha_Bay, CMC_Sea_Point) %>%
+  dplyr::rename(site = station)
+
+CMC_SST$date <- (ymd(CMC_SST$date))
+CMC_SST <- CMC_SST %>% 
+  dplyr::rename(temp = nearest1) %>% 
+  drop_na()
+
+match_func <- function(df){
+  match <- SACTN_US  %>%  
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+CMC <- match_func(df = CMC_SST) %>%
+  drop_na()
+CMC_plot <- match_plot(df = CMC)
+CMC_plot
+```
+
+# Loading the OISST
+
+```{r}
+
+```
+
+```{r}
+load("Data/site_list_sub.Rdata")
+xtable(site_list_sub, auto = TRUE)
+west <- site_list_sub
+west$coast <- "west" # Chnages wc to west
+
+load("Data_P1/africa_coast.RData")
+
+## Downloading the bathy data from NOAA
+# Download mid-res bathymetry data
+# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# colnames(sa_bathy) <- c("lon", "lat", "depth")
+# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+
+# Loading in the newly downloaded bathymetry data
+load("Data_P1/bathy/sa_bathy.RData")
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+shore.normal.transect <- function(site, width = 2){
+  # Find the site on the coastline and it's nearest neighbour points
+  coords <- data.frame(lon = site$lon, lat = site$lat)
+  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
+  coords3 <- coords3[2:1,1:3]
+  # Define the shore normal transect bearing
+  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+  if(heading >= 360){
+    heading <- heading-360
+  } else {
+    heading <- heading
+  }
+  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+  return(heading2)
+}
+
+# Creating the transects
+site_transects <- data.frame()
+for(i in 1:length(west$site)){
+ site <- west[i,]
+ site_transect <- shore.normal.transect(site, 2)
+ site_transects <- rbind(site_transects, site_transect)
+}
+
+# Manually correcting Sea Point and Kommetjie
+site_transects$heading[5:6] <- 290 
+# save(site_transects, file = "Data_P1/site_transects.RData")
+load("Data_P1/site_transects.RData")
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+
+transect.pixel <- function(site, distances){
+  # Extract coordinates
+  coords <- data.frame(lon = site$lon, lat = site$lat)
+  # Find lon/ lats every X metres 
+  pixels <- data.frame()
+  # deep <- 999
+  # distance_multiplier <- 1
+  # while(deep > isobath){
+  for(i in 1:length(distances)){
+    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
+    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
+    bathy2 <- sa_bathy[sitesIdx,]
+    bathy2 <- bathy2[complete.cases(bathy2$depth),]
+    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
+                         heading = site$heading, 
+                         distance = distances[i])
+    pixels <- rbind(pixels, bathy3)
+    coords <- coords2
+  }
+  if(nrow(pixels) < 1){
+    pixels <- data.frame(site, depth = NA)
+  }else{
+    pixels <- pixels
+  }
+  return(pixels)
+}
+
+# Pixel points
+site_pixels <- data.frame()
+for(i in 1:length(west$site)){
+  site <- site_transects[i,]
+  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+  site_pixels <- rbind(site_pixels, site_pixel)
+}
+
+# save(site_pixels, file = "Data_P1/site_pixels.RData")
+load("Data_P1/site_pixels.RData")
+
+# Bounding box
+  # Only one is made in order to know how large the the geom_point() squares should be made to match
+bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
+                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
+                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+
+# Determining the temperature at the various distances from the coast
+```

### 2019-08-21 SST_patterns.Rmd
update

@@
 
+# save(MUR, file = "Data/MUR.RData")
+
 match_plot <- function(df){
@@
 MUR_plot <- match_plot(df = MUR)
-
+MUR_plot
 ```
@@
   drop_na()
+
+# save(CMC, file = "Data/CMC.RData")
+
 CMC_plot <- match_plot(df = CMC)
@@
 ```{r}
+OISSTDir <- "~/Documents/Masters_2019/Data_SST"
+OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
+            col.names = c("lon", "lat", "temp", "date"))
 
+# Visualise the data
+# To explore the data I visualise the min temperatures along the South african coastline.
+
+OISST %>%
+  filter(date == min(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = temp))
+
+load("Data/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
+
+unique_pixel <- OISST %>%
+  select(lon, lat) %>%
+  unique()
+
+# Select nearest 1 pixels (k = 1)
+# Here I use knnx to find the closes 1 pixels to the insitu sites
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+                      query = as.matrix(site_list_sub[,5:6]), k = 1)
+
+pixel_match <- unique_pixel[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  mutate(site = site_list_sub$site)
+
+# Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
+OISST_match <- OISST %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+# save(OISST_match, file = "Data/OISST_match.RData")
+
+OISST_sites <- OISST_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat")) #%>%
+  #dplyr::rename(temp_OISST =temp)
+
+OISST_sites <- OISST_sites %>% 
+  dplyr::mutate(date = as.Date(date)) %>% 
+  drop_na()
+# save(OISST_sites, file = "Data/OISST_sites.RData")
+
+## Matching the Insitu data with the OISST SST data
+OISST <- match_func(df = OISST_sites)
+# save(OISST, file = "Data/OISST.RData")
 ```
+
+Different distances from the coastline
 
@@
 load("Data/site_list_sub.Rdata")
-xtable(site_list_sub, auto = TRUE)
+# xtable(site_list_sub, auto = TRUE)
 west <- site_list_sub
@@
 
-load("Data_P1/africa_coast.RData")
+load("Data/africa_coast.RData")
 
@@
 # Loading in the newly downloaded bathymetry data
-load("Data_P1/bathy/sa_bathy.RData")
+load("Data/bathy/sa_bathy.RData")
 
@@
 # Manually correcting Sea Point and Kommetjie
-site_transects$heading[5:6] <- 290 
-# save(site_transects, file = "Data_P1/site_transects.RData")
-load("Data_P1/site_transects.RData")
+site_transects$heading[4:5] <- 290 
+# save(site_transects, file = "Data/site_transects.RData")
+load("Data/site_transects.RData")
 
@@
 
-# save(site_pixels, file = "Data_P1/site_pixels.RData")
-load("Data_P1/site_pixels.RData")
-
 # Bounding box
@@
 # Determining the temperature at the various distances from the coast
+
+save(site_pixels, file = "Data/site_pixels.RData")
+load("Data/site_pixels.RData")
 ```
+
+#

### 2019-08-21 SST_patterns.Rmd
update

--- 2019-08-20 SST_patterns.Rmd
@@
 
-# Are the same upwelling signals present at different distances from the coastline and through the different datasets
+# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
+
+```{r}
+knitr::opts_chunk$set(
+  comment = "R>",
+  warning = FALSE,
+  message = FALSE 
+)
+
+library(tidyverse)
+# library(plyr)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(forecast)
+library(astrochron)
+library(WaveletComp)
+library(data.table)
+library(heatwaveR)
+library(viridis)
+library(ggrepel)
+library(maptools)
+library(sp)
+library(geosphere)
+library(PBSmapping)
+library(scales)
+library(grid)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(ncdf4)
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+```
+
+# Loading in all the data created using the code bellow
+
+```{r}
+load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
+load("Data/site_pixels.RData")
+load("Data/OISST.RData")
+load("Data/CMC.RData")
+load("Data/MUR.RData")
+```
+
+```{r}
+# SACTN_US <- SACTN_US %>% 
+#   dplyr::rename(in_situ_temp = temp)
+# 
+# # Visualising the data
+# temp_plot <- function(df){
+#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
+#     geom_line(aes(group = site)) +
+#     labs(x = "", y = "Temperature (°C)") +
+#     theme(axis.text.x = element_text(angle = 45)) +
+#     theme(legend.position = "top")
+# }
+# 
+# SACTN_plot <- temp_plot(df = SACTN_US)
+# SACTN_plot
+```
+
+# Loading MUR data
+
+```{r}
+# MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
+# MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
+# MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
+# MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
+# MUR_Hout_Bay <- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")
+# 
+# MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Hout_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Sea_Point) %>%
+#   dplyr::rename(site = station)
+# 
+# MUR_SST$date <- (ymd(MUR_SST$date))
+# MUR_SST <- MUR_SST %>% 
+#   dplyr::rename(temp = nearest1) %>% 
+#   drop_na()
+# 
+# match_func <- function(df){
+#   match <- SACTN_US  %>%  
+#   left_join(df, by = c("site", "date")) %>% 
+#   na.trim()
+#   return(match)
+# }
+# 
+# MUR <- match_func(df = MUR_SST) %>%
+#   drop_na()
+# 
+# # save(MUR, file = "Data/MUR.RData")
+# 
+# match_plot <- function(df){
+#   plot1 <- df %>% 
+# ggplot(aes(x = date, y = temp)) +
+#   geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+#   geom_line() +
+#   facet_wrap(~ site, nrow = 2) +
+#   theme_bw()
+#   return(plot1)
+# }
+# 
+# MUR_plot <- match_plot(df = MUR)
+# MUR_plot
+```
+
+# CMC data
+
+```{r}
+# CMC_Sea_Point <- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
+# CMC_Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
+# CMC_Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
+# CMC_Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
+# CMC_Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")
+# 
+# CMC_SST <- rbind(CMC_Lamberts_Bay,CMC_Hout_Bay, CMC_Port_Nolloth, CMC_Saldanha_Bay, CMC_Sea_Point) %>%
+#   dplyr::rename(site = station)
+# 
+# CMC_SST$date <- (ymd(CMC_SST$date))
+# CMC_SST <- CMC_SST %>% 
+#   dplyr::rename(temp = nearest1) %>% 
+#   drop_na()
+# 
+# match_func <- function(df){
+#   match <- SACTN_US  %>%  
+#   left_join(df, by = c("site", "date")) %>% 
+#   na.trim()
+#   return(match)
+# }
+# 
+# CMC <- match_func(df = CMC_SST) %>%
+#   drop_na()
+# 
+# # save(CMC, file = "Data/CMC.RData")
+# 
+# CMC_plot <- match_plot(df = CMC)
+# CMC_plot
+```
+
+# Loading the OISST
+
+```{r}
+# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
+# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
+#             col.names = c("lon", "lat", "temp", "date"))
+# 
+# # Visualise the data
+# # To explore the data I visualise the min temperatures along the South african coastline.
+# 
+# OISST %>%
+#   filter(date == min(date)) %>%
+#   ggplot(aes(x = lon, y = lat)) +
+#   geom_raster(aes(fill = temp))
+# 
+# load("Data/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data 
+# 
+# unique_pixel <- OISST %>%
+#   select(lon, lat) %>%
+#   unique()
+# 
+# # Select nearest 1 pixels (k = 1)
+# # Here I use knnx to find the closes 1 pixels to the insitu sites
+# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+#                       query = as.matrix(site_list_sub[,5:6]), k = 1)
+# 
+# pixel_match <- unique_pixel[match_index,] %>%
+#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+#   mutate(site = site_list_sub$site)
+# 
+# # Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
+# OISST_match <- OISST %>%
+#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+#   filter(combi %in% pixel_match$combi)
+# # save(OISST_match, file = "Data/OISST_match.RData")
+# 
+# OISST_sites <- OISST_match %>%
+#   left_join(pixel_match, by = c("combi", "lon", "lat")) #%>%
+#   #dplyr::rename(temp_OISST =temp)
+# 
+# OISST_sites <- OISST_sites %>% 
+#   dplyr::mutate(date = as.Date(date)) %>% 
+#   drop_na()
+# # save(OISST_sites, file = "Data/OISST_sites.RData")
+# 
+# ## Matching the Insitu data with the OISST SST data
+# OISST <- match_func(df = OISST_sites)
+# # save(OISST, file = "Data/OISST.RData")
+```
+
+Different distances from the coastline
+
+```{r}
+# load("Data/site_list_sub.Rdata")
+# # xtable(site_list_sub, auto = TRUE)
+# west <- site_list_sub
+# west$coast <- "west" # Chnages wc to west
+# 
+# load("Data/africa_coast.RData")
+# 
+# ## Downloading the bathy data from NOAA
+# # Download mid-res bathymetry data
+# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# # colnames(sa_bathy) <- c("lon", "lat", "depth")
+# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+# 
+# # Loading in the newly downloaded bathymetry data
+# load("Data/bathy/sa_bathy.RData")
+# 
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# shore.normal.transect <- function(site, width = 2){
+#   # Find the site on the coastline and it's nearest neighbour points
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
+#   coords3 <- coords3[2:1,1:3]
+#   # Define the shore normal transect bearing
+#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+#   if(heading >= 360){
+#     heading <- heading-360
+#   } else {
+#     heading <- heading
+#   }
+#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+#   return(heading2)
+# }
+# 
+# # Creating the transects
+# site_transects <- data.frame()
+# for(i in 1:length(west$site)){
+#  site <- west[i,]
+#  site_transect <- shore.normal.transect(site, 2)
+#  site_transects <- rbind(site_transects, site_transect)
+# }
+# 
+# # Manually correcting Sea Point and Kommetjie
+# site_transects$heading[4:5] <- 290 
+# # save(site_transects, file = "Data/site_transects.RData")
+# load("Data/site_transects.RData")
+# 
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+# 
+# transect.pixel <- function(site, distances){
+#   # Extract coordinates
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   # Find lon/ lats every X metres 
+#   pixels <- data.frame()
+#   # deep <- 999
+#   # distance_multiplier <- 1
+#   # while(deep > isobath){
+#   for(i in 1:length(distances)){
+#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
+#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
+#     bathy2 <- sa_bathy[sitesIdx,]
+#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
+#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
+#                          heading = site$heading, 
+#                          distance = distances[i])
+#     pixels <- rbind(pixels, bathy3)
+#     coords <- coords2
+#   }
+#   if(nrow(pixels) < 1){
+#     pixels <- data.frame(site, depth = NA)
+#   }else{
+#     pixels <- pixels
+#   }
+#   return(pixels)
+# }
+# 
+# # Pixel points
+# site_pixels <- data.frame()
+# for(i in 1:length(west$site)){
+#   site <- site_transects[i,]
+#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+#   site_pixels <- rbind(site_pixels, site_pixel)
+# }
+# 
+# # Bounding box
+#   # Only one is made in order to know how large the the geom_point() squares should be made to match
+# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
+#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
+#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+# 
+# # Determining the temperature at the various distances from the coast
+# 
+# # save(site_pixels, file = "Data/site_pixels.RData")
+# load("Data/site_pixels.RData")
+```
+
+# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
+# Is this upwelling event seen throughout the different distances from the coastline
+# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected is present in each of the datasets

### 2019-08-21 upwell_IDX.Rmd
update

@@
 
-Increase in upwelling over the past years as a result of intensified winds over time?
-Weakening of th upwelling over time, leads to a decrease in the upwelling indexs
-Strong south easterly winds - Spring and summer: Upwelling is more intense in spring and summer months
-The formula used suggests that if values at + upwelling occurs and if values are - then downwelling occurs. But the intensity of upwelling changes.
-Compare upwelling indeces per season
+##### Upwelling #####
+# Min duration for an upwelling event - 1day and threshold -1. Upwelling can over within a few hours, but given that we have daily data - upwelling has a minimum duration of 1 day
+# Seasonal upwelling. Upwelling seasons are summer and spring
+# Did the wind intensify? This is usually done by taking first third of the dataset and the last thid of the dataset and doing a t-test on this
+# Strong south easterly winds lead to a stronger upwelling event
 
@@
 
-# Wind speed intensifed ovwer the years
+# Wind speed intensifed over the years
 Take first third (33percentile) of the data and the last third (66th percentile) of the data for every season and then do a t-test

### 2019-08-21 SST_patterns.Rmd
updated

@@
 # Is this upwelling event seen throughout the different distances from the coastline
-# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected is present in each of the datasets
+# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets

### 2019-08-21 upwell_IDX.Rmd
updated

@@
 
-##### Upwelling #####
+##################### Upwelling ######################
 # Min duration for an upwelling event - 1day and threshold -1. Upwelling can over within a few hours, but given that we have daily data - upwelling has a minimum duration of 1 day
@@
 # Strong south easterly winds lead to a stronger upwelling event
+##############################################
 

### 2019-08-21 SST_patterns.Rmd
Calculated upwelling index quantiles

@@
 library(tidyverse)
-# library(plyr)
 library(lubridate)
@@
 library(scales)
-library(grid)
+# library(grid)
 library(gridExtra)
@@
 library(mapproj)
-library(ncdf4)
+# library(ncdf4) # This was used to process NetCDF files in an earlier version
 library(stringr)

### 2019-08-21 upwell_IDX.Rmd
Calculated upwelling index quantiles

@@
 
+# Calculate quantiles of upwelling index
+SACTN_upwell_quantiles <- SACTN_upwell %>% 
+  group_by(site) %>% 
+  summarize(quant_10 = quantile(ui.saws, probs = 0.10, na.rm = TRUE),
+            quant_25 = quantile(ui.saws, probs = 0.25, na.rm = TRUE),
+            quant_50 = quantile(ui.saws, probs = 0.50, na.rm = TRUE),
+            quant_75 = quantile(ui.saws, probs = 0.75, na.rm = TRUE),
+            quant_90 = quantile(ui.saws, probs = 0.90, na.rm = TRUE))
+
 # SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
@@
   arrange(mean_speed) %>% 
+  # Add the following line with mutate to filter the data and add a text label to use with an ANOVA
   summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),

### 2019-08-21 upwell_IDX.Rmd
Added a filter

@@
 SACTN_upwell_quantiles <- SACTN_upwell %>%
+  filter(ui.saws >= 0) %>% 
   group_by(site) %>% 

### 2019-08-21 upwell_IDX.Rmd
Rounding the quantiles

--- 2019-08-21 upwell_IDX.Rmd
@@
             quant_75 = quantile(ui.saws, probs = 0.75, na.rm = TRUE),
-            quant_90 = quantile(ui.saws, probs = 0.90, na.rm = TRUE))
+            quant_90 = quantile(ui.saws, probs = 0.90, na.rm = TRUE)) %>% 
+  mutate_if(is.numeric, round, digits = 2)
 

### 2019-08-27 SST_patterns.Rmd
Updated

@@
 # SACTN_plot
+
+# 25th quantile
 ```
@@
 #     pixels <- pixels
-#   }
+#   }+
+
 #   return(pixels)
@@
 # Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
+
+## Function
+Determining the temperatures at the various distances from the coastline
+
+```{r}
+# Splitting the dataset
+
+PN_site <- site_pixels %>% 
+  filter(site == "Port Nolloth")
+
+LB_site <- site_pixels %>% 
+  filter(site == "Lamberts Bay")
+
+SB_site <- site_pixels %>% 
+  filter(site == "Saldanha Bay")
+
+SP_site <- site_pixels %>% 
+  filter(site == "Sea Point")
+
+HB_site <- site_pixels %>% 
+  filter(site == "Hout Bay")
+
+## Functions:
+
+MUR <- MUR %>% 
+  select(site, lon, lat, temp.y)
+
+OISST <- OISST %>% 
+  select(site, lon, lat, temp.y)
+
+unique_pixels <- function(df){
+  unique_pixel <- df %>%
+  select(lon, lat) %>%
+  unique()
+  return(unique_pixel)
+}
+
+MUR_pixels <- unique_pixels(df = MUR)
+# SACTN_pixels <- unique_pixels(df = SACTN)
+OISST_pixels <- unique_pixels(df = OISST)
+CMC_pixels <- unique_pixels(df = CMC)
+
+match_func <- function(df, xy){
+  match_index <- knnx.index(data = as.matrix(df[,1:2]),
+                    query = as.matrix(xy[,2:3]), k = 1)
+  return(match_index)
+}
+
+MUR_PN <- match_func(df = MUR_pixels, xy = PN_site)
+MUR_LB <- match_func(df = MUR_pixels, xy = LB_site)
+MUR_SB <- match_func(df = MUR_pixels, xy = SB_site)
+MUR_SP <- match_func(df = MUR_pixels, xy = SP_site)
+MUR_HB <- match_func(df = MUR_pixels, xy = HB_site)
+
+# SACTN_pixels <- match_func(df = SACTN)
+OISST_PN <- match_func(df = OISST_pixels, xy = PN_site)
+OISST_LB <- match_func(df = OISST_pixels, xy = LB_site)
+OISST_SB <- match_func(df = OISST_pixels, xy = SB_site)
+OISST_SP <- match_func(df = OISST_pixels, xy = SP_site)
+OISST_HB <- match_func(df = OISST_pixels, xy = HB_site)
+
+CMC_PN <- match_func(df = CMC_pixels, xy = PN_site)
+CMC_LB <- match_func(df = CMC_pixels, xy = LB_site)
+CMC_SB <- match_func(df = CMC_pixels, xy = SB_site)
+CMC_SP <- match_func(df = CMC_pixels, xy = SP_site)
+CMC_HB <- match_func(df = CMC_pixels, xy = HB_site)
+
+pixel_match_func <- function(df, xy){ 
+  pixel_match <- df[xy,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  mutate(site = PN_site$site)
+  return(pixel_match)
+}
+
+MUR_match_PN <- pixel_match_func(df = MUR_pixels, xy = MUR_PN)
+MUR_match_LB <- pixel_match_func(df = MUR_pixels, xy = MUR_LB)
+MUR_match_SB <- pixel_match_func(df = MUR_pixels, xy = MUR_SB)
+MUR_match_SP <- pixel_match_func(df = MUR_pixels, xy = MUR_SP)
+MUR_match_HB <- pixel_match_func(df = MUR_pixels, xy = MUR_HB)
+
+CMC_match_PN <- pixel_match_func(df = CMC_pixels, xy = CMC_PN)
+CMC_match_LB <- pixel_match_func(df = CMC_pixels, xy = CMC_LB)
+CMC_match_SB <- pixel_match_func(df = CMC_pixels, xy = CMC_SB)
+CMC_match_SP <- pixel_match_func(df = CMC_pixels, xy = CMC_SP)
+CMC_match_HB <- pixel_match_func(df = CMC_pixels, xy = CMC_HB)
+
+OISST_match_PN <- pixel_match_func(df = OISST_pixels, xy = OISST_PN)
+OISST_match_LB <- pixel_match_func(df = OISST_pixels, xy = OISST_LB)
+OISST_match_SB <- pixel_match_func(df = OISST_pixels, xy = OISST_SB)
+OISST_match_SP <- pixel_match_func(df = OISST_pixels, xy = OISST_SP)
+OISST_match_HB <- pixel_match_func(df = OISST_pixels, xy = OISST_HB)
+
+match_func <- function(df,xy){
+matched_coords <- df %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% xy$combi)
+return(matched_coords)
+}
+
+MUR_temp_PN <- match_func(df = MUR, xy = MUR_match_PN)
+MUR_temp_LB <- match_func(df = MUR, xy = MUR_match_LB)
+MUR_temp_SB <- match_func(df = MUR, xy = MUR_match_SB)
+MUR_temp_SP <- match_func(df = MUR, xy = MUR_match_SP)
+MUR_temp_HB <- match_func(df = MUR, xy = MUR_match_HB)
+
+CMC_temp_PN <- match_func(df = CMC, xy = CMC_match_PN)
+CMC_temp_LB <- match_func(df = CMC, xy = CMC_match_LB)
+CMC_temp_SB <- match_func(df = CMC, xy = CMC_match_SB)
+CMC_temp_SP <- match_func(df = CMC, xy = CMC_match_SP)
+CMC_temp_HB <- match_func(df = CMC, xy = CMC_match_HB)
+```

### 2019-08-27 upwell_IDX.Rmd
Updated

@@
   # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
-  mutate(clim = purrr::map(data, ts2clm, pctile = 20, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
+  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
   select(-data) %>% 
@@
 ```{r}
-load("Data/wind_daily.RData")
-
-wind_daily <- wind_daily %>% 
-  dplyr::rename(t = date)
-
-# wind_daily_season <- season_func(df = wind_daily)
-
-#save(wind_daily_season, file = "Data/wind_daily_season.RData")
-
-load("Data/wind_daily_season.RData")
-
-wind_daily_season_test <- wind_daily_season %>% 
-  group_by(season, temp_sites) %>% 
-  arrange(mean_speed) %>% 
-  # Add the following line with mutate to filter the data and add a text label to use with an ANOVA
-  summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
-            mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
-
-# T-test
-
-# T- test between seasons for all sites
-
+# load("Data/wind_daily.RData") 
+# 
+# wind_daily <- wind_daily %>% 
+#   dplyr::rename(t = date)
+# 
+# # wind_daily_season <- season_func(df = wind_daily)
+# 
+# #save(wind_daily_season, file = "Data/wind_daily_season.RData")
+# 
+# load("Data/wind_daily_season.RData")
+# 
+# wind_daily_season_test <- wind_daily_season %>% 
+#   group_by(season, temp_sites) %>% 
+#   arrange(mean_speed) %>% 
+#   # Add the following line with mutate to filter the data and add a text label to use with an ANOVA
+#   summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
+#             mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
 ```

### 2019-08-28 SST_patterns.Rmd
updated

@@
 
+CMC <- CMC %>% 
+  select(site, lon, lat, temp.y)
+
 unique_pixels <- function(df){
@@
 OISST_pixels <- unique_pixels(df = OISST)
+OISST_pixels <- OISST_pixels[-c(2),]
+
 CMC_pixels <- unique_pixels(df = CMC)
@@
 CMC_temp_HB <- match_func(df = CMC, xy = CMC_match_HB)
+
+OISST_temp_PN <- match_func(df = OISST, xy = OISST_match_PN)
+OISST_temp_LB <- match_func(df = OISST, xy = OISST_match_LB)
+OISST_temp_SB <- match_func(df = OISST, xy = OISST_match_SB)
+OISST_temp_SP <- match_func(df = OISST, xy = OISST_match_SP)
+OISST_temp_HB <- match_func(df = OISST, xy = OISST_match_HB)
 ```

### 2019-08-28 Write_up/write_up.Rmd
wind_analyses
fixed

@@
 Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing.
-To assess changes in upwelling events, the aim of this investigation is to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event.
+To assess changes in upwelling events, the aim of this investigation is to use SST and wind data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event.
 Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
 Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS.
-We also aim to determine whether or not there is a change in the rate of onset of marine heatwaves and marine coldspells.
+We also aim to determine whether or not there is a change in the rate of onset of upwelling and marine coldspells.
 
@@
 
-The primary aim of the first study was to make use of various remotely sensed SST data, in-situ collected coastal seawater temperature data and chlorophyll-a data.
-It was also useful to examine how the intensity and duration of upwelling varies per dataset as these datasets are collected at different resolutions, with the assumption that high resolution data may reveal a more accurate detection of the duration and intensity of upwelling events.
-Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
-Chlorophyll a data were then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites.
-It was hypothesised that an increase in the intensity and duration of upwelling will show an increase in the abundance of chlorophyll a concentration.
-The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets.
-Thereafter analyses rate of onset of MHW and MCS was done.
-
-Thereafter the aim of the second study was to determine the occurrence of MCSs along the west coast of Southern Africa between the nearshore (<400m from the shore) and offshore.
+The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
 The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
 2013).
-With the knowledge obtained from study 1, it was now possible to investigate whether a more intense upwelling event may lead to an increase in the intensity and duration of MCSs.
-The SACTN dataset and a total of five SST datasets could be used to investigate the occurrence of marine cold-spells (MCSs) and marine heatwaves with the methodology introduced by Hobday et al.
-2016.
-This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends.
+It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and whether or not the same patterns exist between the different datasets.
 
@@
 
-The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine wether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
+The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
 It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises.
+Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
+The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets.
+Thereafter analyses rate of onset of MCS was done.This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends
 

### 2019-08-28 SST_patterns.Rmd
Updated the pixel finding pipeline to use purrr

--- 2019-08-27 SST_patterns.Rmd
@@
 ```{r}
-# Splitting the dataset
 
-PN_site <- site_pixels %>% 
-  filter(site == "Port Nolloth")
+# RWS: It appears as though the satellite data you are loading here have already been trimmed down to the one nearest pixel to each SACTN site
+# This won't work for the following steps because you need to be finding pixels from the COMPLETE set of satellite data you downloaded
+# Once you load the complete satellite data lon/lat values you can run the following code and it should give you the coordinates you want
 
-LB_site <- site_pixels %>% 
-  filter(site == "Lamberts Bay")
-
-SB_site <- site_pixels %>% 
-  filter(site == "Saldanha Bay")
-
-SP_site <- site_pixels %>% 
-  filter(site == "Sea Point")
-
-HB_site <- site_pixels %>% 
-  filter(site == "Hout Bay")
-
-## Functions:
-
+# These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
 MUR <- MUR %>% 
-  select(site, lon, lat, temp.y)
+  select(lon, lat) %>% 
+  mutate(product = "MUR")
 
 OISST <- OISST %>% 
-  select(site, lon, lat, temp.y)
+  select(lon, lat) %>% 
+  mutate(product = "OISST")
 
-unique_pixels <- function(df){
-  unique_pixel <- df %>%
+CMC <- CMC %>% 
   select(lon, lat) %>% 
+  mutate(product = "CMC")
+
+sat_data <- rbind(MUR, OISST) %>% 
+  rbind(., CMC) %>% 
+  select(product, lon, lat)
+
+# rm(MUR, OISST, CMC); gc()
+
+sat_pixels <- sat_data %>% 
+  select(product, lon, lat) %>%
   unique()
-  return(unique_pixel)
+
+## For testing the nest/map pipeline
+# df <- sat_pixels %>% 
+#   filter(product == "MUR", site == "Lamberts Bay") %>% 
+#   select(-c(product, site))
+
+match_func <- function(df){
+  df <- na.omit(df) %>% 
+    dplyr::rename(lon_sat = lon, lat_sat = lat)
+  match_index <- knnx.index(data = as.matrix(site_pixels[,2:3]),
+                    query = as.matrix(df[,1:2]), k = 1)
+  res <- cbind(site_pixels[match_index,], df)
+  return(res)
 }
 
-MUR_pixels <- unique_pixels(df = MUR)
-# SACTN_pixels <- unique_pixels(df = SACTN)
-OISST_pixels <- unique_pixels(df = OISST)
-CMC_pixels <- unique_pixels(df = CMC)
+pixel_match <- sat_pixels %>% 
+  group_by(product) %>% 
+  nest() %>% 
+  mutate(match = purrr:::map(data, match_func)) %>% 
+  select(-data) %>%
+  unnest(match)
 
-match_func <- function(df, xy){
-  match_index <- knnx.index(data = as.matrix(df[,1:2]),
-                    query = as.matrix(xy[,2:3]), k = 1)
-  return(match_index)
-}
-
-MUR_PN <- match_func(df = MUR_pixels, xy = PN_site)
-MUR_LB <- match_func(df = MUR_pixels, xy = LB_site)
-MUR_SB <- match_func(df = MUR_pixels, xy = SB_site)
-MUR_SP <- match_func(df = MUR_pixels, xy = SP_site)
-MUR_HB <- match_func(df = MUR_pixels, xy = HB_site)
-
-# SACTN_pixels <- match_func(df = SACTN)
-OISST_PN <- match_func(df = OISST_pixels, xy = PN_site)
-OISST_LB <- match_func(df = OISST_pixels, xy = LB_site)
-OISST_SB <- match_func(df = OISST_pixels, xy = SB_site)
-OISST_SP <- match_func(df = OISST_pixels, xy = SP_site)
-OISST_HB <- match_func(df = OISST_pixels, xy = HB_site)
-
-CMC_PN <- match_func(df = CMC_pixels, xy = PN_site)
-CMC_LB <- match_func(df = CMC_pixels, xy = LB_site)
-CMC_SB <- match_func(df = CMC_pixels, xy = SB_site)
-CMC_SP <- match_func(df = CMC_pixels, xy = SP_site)
-CMC_HB <- match_func(df = CMC_pixels, xy = HB_site)
-
-pixel_match_func <- function(df, xy){ 
-  pixel_match <- df[xy,] %>%
-  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  mutate(site = PN_site$site)
-  return(pixel_match)
-}
-
-MUR_match_PN <- pixel_match_func(df = MUR_pixels, xy = MUR_PN)
-MUR_match_LB <- pixel_match_func(df = MUR_pixels, xy = MUR_LB)
-MUR_match_SB <- pixel_match_func(df = MUR_pixels, xy = MUR_SB)
-MUR_match_SP <- pixel_match_func(df = MUR_pixels, xy = MUR_SP)
-MUR_match_HB <- pixel_match_func(df = MUR_pixels, xy = MUR_HB)
-
-CMC_match_PN <- pixel_match_func(df = CMC_pixels, xy = CMC_PN)
-CMC_match_LB <- pixel_match_func(df = CMC_pixels, xy = CMC_LB)
-CMC_match_SB <- pixel_match_func(df = CMC_pixels, xy = CMC_SB)
-CMC_match_SP <- pixel_match_func(df = CMC_pixels, xy = CMC_SP)
-CMC_match_HB <- pixel_match_func(df = CMC_pixels, xy = CMC_HB)
-
-OISST_match_PN <- pixel_match_func(df = OISST_pixels, xy = OISST_PN)
-OISST_match_LB <- pixel_match_func(df = OISST_pixels, xy = OISST_LB)
-OISST_match_SB <- pixel_match_func(df = OISST_pixels, xy = OISST_SB)
-OISST_match_SP <- pixel_match_func(df = OISST_pixels, xy = OISST_SP)
-OISST_match_HB <- pixel_match_func(df = OISST_pixels, xy = OISST_HB)
-
-match_func <- function(df,xy){
-matched_coords <- df %>%
-  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  filter(combi %in% xy$combi)
-return(matched_coords)
-}
-
-MUR_temp_PN <- match_func(df = MUR, xy = MUR_match_PN)
-MUR_temp_LB <- match_func(df = MUR, xy = MUR_match_LB)
-MUR_temp_SB <- match_func(df = MUR, xy = MUR_match_SB)
-MUR_temp_SP <- match_func(df = MUR, xy = MUR_match_SP)
-MUR_temp_HB <- match_func(df = MUR, xy = MUR_match_HB)
-
-CMC_temp_PN <- match_func(df = CMC, xy = CMC_match_PN)
-CMC_temp_LB <- match_func(df = CMC, xy = CMC_match_LB)
-CMC_temp_SB <- match_func(df = CMC, xy = CMC_match_SB)
-CMC_temp_SP <- match_func(df = CMC, xy = CMC_match_SP)
-CMC_temp_HB <- match_func(df = CMC, xy = CMC_match_HB)
+# You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+# You will then have one time series for each product for each pixel at each distance for each site that you may then
+# run the upwelling event function on
+# You should do this via purrr::nest+map as you can see how much less code it requires
 ```

### 2019-08-31 SST_patterns.Rmd
Extracting data from netCDFS

@@
 library(FNN)
-library(forecast)
-library(astrochron)
-library(WaveletComp)
-library(data.table)
-library(heatwaveR)
-library(viridis)
-library(ggrepel)
-library(maptools)
-library(sp)
-library(geosphere)
-library(PBSmapping)
 library(scales)
-# library(grid)
 library(gridExtra)
@@
 library(xtable)
+library(ncdf4) # library for processing netCDFs
+library(ncdf4)
+library(data.table)
+library(plyr)
 ```
@@
 load("Data/site_pixels.RData")
-load("Data/OISST.RData")
-load("Data/CMC.RData")
-load("Data/MUR.RData")
 ```
@@
 # 25th quantile
-```
-
-# Loading MUR data
-
-```{r}
-# MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
-# MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
-# MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
-# MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
-# MUR_Hout_Bay <- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")
-# 
-# MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Hout_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Sea_Point) %>%
-#   dplyr::rename(site = station)
-# 
-# MUR_SST$date <- (ymd(MUR_SST$date))
-# MUR_SST <- MUR_SST %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   drop_na()
-# 
-# match_func <- function(df){
-#   match <- SACTN_US  %>%  
-#   left_join(df, by = c("site", "date")) %>% 
-#   na.trim()
-#   return(match)
-# }
-# 
-# MUR <- match_func(df = MUR_SST) %>%
-#   drop_na()
-# 
-# # save(MUR, file = "Data/MUR.RData")
-# 
-# match_plot <- function(df){
-#   plot1 <- df %>% 
-# ggplot(aes(x = date, y = temp)) +
-#   geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
-#   geom_line() +
-#   facet_wrap(~ site, nrow = 2) +
-#   theme_bw()
-#   return(plot1)
-# }
-# 
-# MUR_plot <- match_plot(df = MUR)
-# MUR_plot
-```
-
-# CMC data
-
-```{r}
-# CMC_Sea_Point <- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
-# CMC_Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
-# CMC_Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
-# CMC_Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
-# CMC_Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")
-# 
-# CMC_SST <- rbind(CMC_Lamberts_Bay,CMC_Hout_Bay, CMC_Port_Nolloth, CMC_Saldanha_Bay, CMC_Sea_Point) %>%
-#   dplyr::rename(site = station)
-# 
-# CMC_SST$date <- (ymd(CMC_SST$date))
-# CMC_SST <- CMC_SST %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   drop_na()
-# 
-# match_func <- function(df){
-#   match <- SACTN_US  %>%  
-#   left_join(df, by = c("site", "date")) %>% 
-#   na.trim()
-#   return(match)
-# }
-# 
-# CMC <- match_func(df = CMC_SST) %>%
-#   drop_na()
-# 
-# # save(CMC, file = "Data/CMC.RData")
-# 
-# CMC_plot <- match_plot(df = CMC)
-# CMC_plot
-```
-
-# Loading the OISST
-
-```{r}
-# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
-# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
-#             col.names = c("lon", "lat", "temp", "date"))
-# 
-# # Visualise the data
-# # To explore the data I visualise the min temperatures along the South african coastline.
-# 
-# OISST %>%
-#   filter(date == min(date)) %>%
-#   ggplot(aes(x = lon, y = lat)) +
-#   geom_raster(aes(fill = temp))
-# 
-# load("Data/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data 
-# 
-# unique_pixel <- OISST %>%
-#   select(lon, lat) %>%
-#   unique()
-# 
-# # Select nearest 1 pixels (k = 1)
-# # Here I use knnx to find the closes 1 pixels to the insitu sites
-# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-#                       query = as.matrix(site_list_sub[,5:6]), k = 1)
-# 
-# pixel_match <- unique_pixel[match_index,] %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   mutate(site = site_list_sub$site)
-# 
-# # Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
-# OISST_match <- OISST %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   filter(combi %in% pixel_match$combi)
-# # save(OISST_match, file = "Data/OISST_match.RData")
-# 
-# OISST_sites <- OISST_match %>%
-#   left_join(pixel_match, by = c("combi", "lon", "lat")) #%>%
-#   #dplyr::rename(temp_OISST =temp)
-# 
-# OISST_sites <- OISST_sites %>% 
-#   dplyr::mutate(date = as.Date(date)) %>% 
-#   drop_na()
-# # save(OISST_sites, file = "Data/OISST_sites.RData")
-# 
-# ## Matching the Insitu data with the OISST SST data
-# OISST <- match_func(df = OISST_sites)
-# # save(OISST, file = "Data/OISST.RData")
 ```
@@
 
+# Extrcting the MUR data
+
+```{r}
+ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+csvDir <- "/home/amieroh/Documents"
+
+bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
+                    CC = c(25, 35, 340, 355), # Canary Current
+                   CalC = c(35, 45, 225, 240), # California Current
+                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+  coords <- bbox[,region]
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
+    fNameStem <-
+      substr(ncFile, pathLen + 10, pathLen + 38)
+    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <-
+      as.data.frame(melt(sst, value.name = "temp"), row.names = NULL) %>%
+      mutate(t = ymd(fDate)) %>%
+      na.omit()
+    fwrite(sst,
+           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+}
+
+# the list of files
+ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+strtDate <- str_sub(ncList[1], start = 1, end = 8)
+endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)
+
+# apply the function
+system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))
+```
+
+# Extracting the CMC data
+
+```{r}
+ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_global"
+csvDir <- "/home/amieroh/Documents"
+
+bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
+                    CC = c(25, 35, 340, 355), # Canary Current
+                   CalC = c(35, 45, 225, 240), # California Current
+                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+  coords <- bbox[,region]
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
+    fNameStem <-
+      substr(ncFile, pathLen + 10, pathLen + 38)
+    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <-
+      as.data.frame(melt(sst, value.name = "temp"), row.names = NULL) %>%
+      mutate(t = ymd(fDate)) %>%
+      na.omit()
+    fwrite(sst,
+           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+}
+
+# the list of files
+ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+strtDate <- str_sub(ncList[1], start = 1, end = 8)
+endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)
+
+# apply the function
+system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))
+
+ncDir <- "/Volumes/Challenger_Deep/OceanData/CMC0.2deg-CMC-L4-GLOB-v2.0/netCDF"
+CMC.csv.dir <- "/Volumes/Challenger_Deep/spatial/processed/CMC0.2deg-CMC-L4-GLOB-v2.0/WBC/daily"
+
+system.time(read_nc(ncDir, "AC", CMC.csv.dir))
+```
+
+# Extracting the OISST data
+
+```{r}
+bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+                   CC = c(25, 35, 340, 355), # Canary Current
+                   CalC = c(35, 45, 225, 240), # California Current
+                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+
+OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
+OISST.csv.dir <- "/home/amieroh/Documents/OISST_subset"
+
+#          1         2
+# 1234567890123456789012345
+# avhrr-only-v2.19810901.nc
+
+# function to extract the dims and data from OISST netCDFs
+read_nc <- function(ncFile, region = region, csvDir = csvDir) {
+  coords <- bbox[, region]
+  nc <- nc_open(ncFile)
+  pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
+  fNameStem <-
+    substr(ncFile, pathLen + 1, pathLen + 13)
+  fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
+  LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+  LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+  sst <- ncvar_get(nc,
+                   varid = "sst",
+                   start = c(LonIdx[1], LatIdx[1], 1, 1),
+                   count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
+    round(4)
+  dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
+                        lat = nc$dim$lat$vals[LatIdx])
+  nc_close(nc)
+  sst <-
+    as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
+    mutate(t = ymd(fDate)) %>%
+    na.omit()
+  fwrite(sst,
+         file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
+         append = TRUE, col.names = FALSE)
+  rm(sst)
+}
+
+# the list of files
+ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+strtDate <- str_sub(ncList[1], start = 15, end = 22)
+endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
+
+# apply the function
+system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
+
+```
+
 # Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 

### 2019-08-31 upwell_IDX.Rmd
UPDATE

@@
   dplyr::rename(t = date)
-
 # save(UI_angle, file = "Data/UI_angle.RData")
-
 # index <- wind_UI %>%
@@
 #   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-
 # wind_UI <- wind_daily %>% 
@@
   select(site, t, ui.saws, exceedance)
-
 # Calculate quantiles of upwelling index
@@
   mutate_if(is.numeric, round, digits = 2)
-
 # SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
@@
 # does what I want and can simply be given to purrr::map() without any faffing about.
-
 detect_event_custom <- function(df){
@@
   # unnest()
-
 # Unpack the event metric reults
@@
 # save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
-
 # Unpack the daily climatology results
@@
 # save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
-
 # The above chunk of code appears to work as expected.
@@
 # Up next is fitting linear models to the upwelling metric results to see what trends shake out.
-
 # Linear Models
@@
 ## Adding a seasonal component to the UI_angle dataset
-
 season_func <- function(df){
@@
 }
-
 # UI_angle_season <- season_func(df = UI_angle)
 # save(UI_angle_season, file = "UI_angle_season.RData")
-
 load("Data/UI_angle_season.RData")
-
 # Speed of wind increased over time???
@@
 }
-
 wind_speed_aov <- anova_func(df = UI_angle_season)
 summary(wind_speed_aov)
-
 anova_func <- function(df){

### 2019-08-31 SST_patterns.Rmd
updated

@@
 load("Data/site_pixels.RData")
+load("Data/OISST.RData")
 ```
@@
 ```{r}
-bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-                   CC = c(25, 35, 340, 355), # Canary Current
-                   CalC = c(35, 45, 225, 240), # California Current
-                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
-
-OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
-OISST.csv.dir <- "/home/amieroh/Documents/OISST_subset"
-
-#          1         2
-# 1234567890123456789012345
-# avhrr-only-v2.19810901.nc
-
-# function to extract the dims and data from OISST netCDFs
-read_nc <- function(ncFile, region = region, csvDir = csvDir) {
-  coords <- bbox[, region]
-  nc <- nc_open(ncFile)
-  pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
-  fNameStem <-
-    substr(ncFile, pathLen + 1, pathLen + 13)
-  fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
-  LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-  LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-  sst <- ncvar_get(nc,
-                   varid = "sst",
-                   start = c(LonIdx[1], LatIdx[1], 1, 1),
-                   count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
-    round(4)
-  dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
-                        lat = nc$dim$lat$vals[LatIdx])
-  nc_close(nc)
-  sst <-
-    as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
-    mutate(t = ymd(fDate)) %>%
-    na.omit()
-  fwrite(sst,
-         file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
-         append = TRUE, col.names = FALSE)
-  rm(sst)
-}
-
-# the list of files
-ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-strtDate <- str_sub(ncList[1], start = 15, end = 22)
-endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
-
-# apply the function
-system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
-
+# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+#                    CC = c(25, 35, 340, 355), # Canary Current
+#                    CalC = c(35, 45, 225, 240), # California Current
+#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+# 
+# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
+# OISST.csv.dir <- "/home/amieroh/Documents/OISST_subset"
+# 
+# #          1         2
+# # 1234567890123456789012345
+# # avhrr-only-v2.19810901.nc
+# 
+# # function to extract the dims and data from OISST netCDFs
+# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
+#   coords <- bbox[, region]
+#   nc <- nc_open(ncFile)
+#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
+#   fNameStem <-
+#     substr(ncFile, pathLen + 1, pathLen + 13)
+#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
+#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#   sst <- ncvar_get(nc,
+#                    varid = "sst",
+#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
+#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
+#     round(4)
+#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
+#                         lat = nc$dim$lat$vals[LatIdx])
+#   nc_close(nc)
+#   sst <-
+#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
+#     mutate(t = ymd(fDate)) %>%
+#     na.omit()
+#   fwrite(sst,
+#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
+#          append = TRUE, col.names = FALSE)
+#   rm(sst)
+# }
+# 
+# # the list of files
+# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# strtDate <- str_sub(ncList[1], start = 15, end = 22)
+# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
+# 
+# # apply the function
+# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# 
+# # Loading the data
+# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
+# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
+# 
+# # Saving the data
+# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
 ```

### 2019-08-31 SST_patterns.Rmd
progress

--- 2019-08-31 SST_patterns.Rmd
@@
 
-bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
+bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
                   CC = c(25, 35, 340, 355), # Canary Current
@@
 
-bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
+bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
                     CC = c(25, 35, 340, 355), # Canary Current
@@
 CMC.csv.dir <- "/Volumes/Challenger_Deep/spatial/processed/CMC0.2deg-CMC-L4-GLOB-v2.0/WBC/daily"
-
-system.time(read_nc(ncDir, "AC", CMC.csv.dir))
 ```
@@
 ```{r}
-bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-                   CC = c(25, 35, 340, 355), # Canary Current
-                   CalC = c(35, 45, 225, 240), # California Current
-                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
-
-OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
-OISST.csv.dir <- "/home/amieroh/Documents/OISST_subset"
-
-#          1         2
-# 1234567890123456789012345
-# avhrr-only-v2.19810901.nc
-
-# function to extract the dims and data from OISST netCDFs
-read_nc <- function(ncFile, region = region, csvDir = csvDir) {
-  coords <- bbox[, region]
-  nc <- nc_open(ncFile)
-  pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
-  fNameStem <-
-    substr(ncFile, pathLen + 1, pathLen + 13)
-  fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
-  LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-  LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-  sst <- ncvar_get(nc,
-                   varid = "sst",
-                   start = c(LonIdx[1], LatIdx[1], 1, 1),
-                   count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
-    round(4)
-  dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
-                        lat = nc$dim$lat$vals[LatIdx])
-  nc_close(nc)
-  sst <-
-    as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
-    mutate(t = ymd(fDate)) %>%
-    na.omit()
-  fwrite(sst,
-         file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
-         append = TRUE, col.names = FALSE)
-  rm(sst)
-}
-
-# the list of files
-ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-strtDate <- str_sub(ncList[1], start = 15, end = 22)
-endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
-
-# apply the function
-system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
-
+# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+#                    CC = c(25, 35, 340, 355), # Canary Current
+#                    CalC = c(35, 45, 225, 240), # California Current
+#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+# 
+# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
+# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
+# 
+# #          1         2
+# # 1234567890123456789012345
+# # avhrr-only-v2.19810901.nc
+# 
+# # function to extract the dims and data from OISST netCDFs
+# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
+#   coords <- bbox[, region]
+#   nc <- nc_open(ncFile)
+#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
+#   fNameStem <-
+#     substr(ncFile, pathLen + 1, pathLen + 13)
+#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
+#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#   sst <- ncvar_get(nc,
+#                    varid = "sst",
+#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
+#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
+#     round(4)
+#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
+#                         lat = nc$dim$lat$vals[LatIdx])
+#   nc_close(nc)
+#   sst <-
+#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
+#     mutate(t = ymd(fDate)) %>%
+#     na.omit()
+#   fwrite(sst,
+#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
+#          append = TRUE, col.names = FALSE)
+#   rm(sst)
+# }
+# 
+# # the list of files
+# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# strtDate <- str_sub(ncList[1], start = 15, end = 22)
+# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
+# 
+# # apply the function
+# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# 
+# # Loading the data
+# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
+# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
+# 
+# # Saving the data
+# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
 ```
@@
 ```{r}
-
 # RWS: It appears as though the satellite data you are loading here have already been trimmed down to the one nearest pixel to each SACTN site

### 2019-09-03 SST_patterns.Rmd
update

@@
 load("Data/OISST.RData")
+OISST <- BC_avhrr_only_v2_Document_Document 
+rm(BC_avhrr_only_v2_Document_Document )
 ```
@@
 ```{r}
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+# nc_data <- nc_open('/home/amieroh/Documents/Data/Datasets/MUR/daily/20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc')
+# print(nc_data)
+
 ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
@@
 
-read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-  coords <- bbox[,region]
+read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
     nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
     fNameStem <-
-      substr(ncFile, pathLen + 10, pathLen + 38)
-    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
+      substr(basename(ncFile), 10, 38)
+    fDate <- substr(basename(ncFile), 1, 8)
     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
@@
     nc_close(nc)
-    sst <-
-      as.data.frame(melt(sst, value.name = "temp"), row.names = NULL) %>%
-      mutate(t = ymd(fDate)) %>%
-      na.omit()
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
     fwrite(sst,
-           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
            append = TRUE, col.names = FALSE)
@@
 
-# the list of files
-ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-strtDate <- str_sub(ncList[1], start = 1, end = 8)
-endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)
-
-# apply the function
-system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))
+llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
 ```
@@
 # 12345678901234567890123456789012345678901
-# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
 
@@
 ncDir <- "/Volumes/Challenger_Deep/OceanData/CMC0.2deg-CMC-L4-GLOB-v2.0/netCDF"
-CMC.csv.dir <- "/Volumes/Challenger_Deep/spatial/processed/CMC0.2deg-CMC-L4-GLOB-v2.0/WBC/daily"
 ```
@@
 
-pixel_match <- sat_pixels %>% 
+pixel_match <- OISST %>% 
   group_by(product) %>% 

### 2019-09-03 Write_up/write_up.Rmd
update

@@
 
-In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature, chlorophyll-a and wind were analysed and accessed.
+In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature and wind were analysed and accessed.
 
@@
 Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
-
-## Chlorophyll-a data
-
-In order to observe the patterns in the variation of the chlorophyll-a concentration with the intensity and duration of upwelling events,  MODIS-Aqua Chlorophyll a concentration, OCI Algorithm, 8-day composite were used.
-The National Aeronautics and Space Administration’s (NASA) Moderate Resolution imaging Spectroradiometer (MODIS) Aqua satellite is an example of a satellite currently collecting the reflected wavelength energy from the earth.
-The MODIS Aqua satellite records wavelength data down to the 1% photosynthetically available depth (0.1 PAR depth) of the water column, a vital component in the estimation of chlorophyll (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
-Because chlorophyll-a concentration retrieval is only possible over cloud-free areas the daily data offered may have large data gaps due to cloud cover.
-The quality of the chlorophyll-a concentration  estimates depends strongly on the quality of the cloud mask.
-The NASA Aqua satellite contains the MODIS sensor that is used for detecting chlorophyll-a concentrations (ocean color) in the world's oceans, among other applications (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
-The Level 3 standard mapped image (SMI) chlorophyll-a dataset has a monthly temporal resolution and 4.6 km spatial resolution.
-The SMI dataset is an image representation of binned MODIS data.
 

### 2019-09-04 Write_up/write_up.Rmd
last writing update

@@
 Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing.
-To assess changes in upwelling events, the aim of this investigation is to use SST and wind data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event.
+To assess changes in upwelling events, the aim of this investigation is to make use of SST and wind data to monitor the durations and intensities of upwelling events at different distances from the coastline within the Benguela EBUS.
+
 Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
-Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS.
-We also aim to determine whether or not there is a change in the rate of onset of upwelling and marine coldspells.
+TThereafter,Inorder to determine whether upwelling events had resulted in an increase in the onset of marine cold spells, all four easter boundry upwelling systems were considred.
 
@@
 This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet].
-Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region.
-This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region.
-Its aims to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
-In this research project, various satellite derived observations of sea surface temperature (SST) and wind stress will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
+Very few studies have addressed the occurrence and the changes in the intensity or duration of cold temperature spells or cold events within the Benguela region.
+This investigation examines recent trends and variation in the frequency and duration of upwelling events occurring in this region In this research project, various satellite derived observations of sea surface temperature (SST) and wind will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
+Thereafter, satellite data will be used to examine upwelling on a global scale and whether these events may have influence the rate of onset of marine cold spells.
 
@@
 2013).
-It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and whether or not the same patterns exist between the different datasets.
+It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
 
@@
 
-The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
-It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises.
+The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for all of the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
+It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and over the years given the influence of climate change and sea surface temperature rises.
 Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
@@
 Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
-An additional assumption within this research focused on the effect of upwelling on MCSs and MHWs.
+An additional assumption within this research focused on the influence of the changes of upwelling on marine cold spells.
 Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events.

### 2019-09-04 upwell_IDX.Rmd
Updated

--- 2019-08-27 upwell_IDX.Rmd
@@
 #   # ungroup() %>%
-#   # mutate(date = as.Date(date))%>%
+#   # mutate(date = as.Date(date))%>% # Changing the date to date format 
 #   dplyr::group_by(temp_sites, date) %>% 
-#   filter(temp_sites %in% selected_sites) %>%
-#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
-#                    mean_speed = round(mean(speed),2))
+#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
+#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2), # Direction using the circular function
+#                    mean_speed = round(mean(speed),2)) # Calculating the mean 
 # #save(wind_daily, file = "Data/wind_daily.RData")
@@
 ```{r}
+# Coast R- function that calculates the angle of the site perpendicular to the coastline
+# Transect function in coastR package finds the angle away or along the coastline
 south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
+
+# Plotting the map
 world_map <- ggplot() + 
   borders(fill = "grey40", colour = "black")
+
 # Create titles
@@
 
+# Upwelling indeces using the above formula
+
 ```{r}
-UI_angle <- wind_daily %>% 
-  dplyr::rename(site = sites) %>% 
-  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
-  dplyr::rename(coast_angle = heading) %>% 
-  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
-  drop_na %>% 
-  dplyr::rename(t = date)
+UI_angle <- wind_daily %>%  # Making refeerence to the wind daily datta created above
+  dplyr::rename(site = sites) %>%  # Renaming sites to site
+  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to south africa wide data column 2 and 21. by site 
+  dplyr::rename(coast_angle = heading) %>% # Renamingthe column heading to coast_angle
+  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
+  drop_na %>% # removing the na values
+  dplyr::rename(t = date) #renaming date to t
 
 # save(UI_angle, file = "Data/UI_angle.RData")
-
 # index <- wind_UI %>%
@@
 
-# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be
+# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.
 
@@
 # Loading the insitu temperature data along the wc
-load("Data/SACTN_US.RData")
+load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
 # exceedance <-
@@
   #        max_t = max(t)) %>% 
-  nest() %>% 
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")), # creating a column of climaatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
@@
 Given that we have daily data, upwelling events minimum duration here will be 1day
+  # Detect consecutive days in exceedance of a given threshold.
   # mutate() %>% 
@@
   filter(row_number() %% 2 == 1) %>%
-  unnest() %>% 
-  dplyr::rename(ui.saws = temp) %>% 
-  select(site, t, ui.saws, exceedance)
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance) # selecting only these variables
+# Calculate quantiles of upwelling index
 
-# Calculate quantiles of upwelling index
+# Static numbers are often rejected and so we decided to find a percentile value as these are often more likely to be approved
+
 SACTN_upwell_quantiles <- SACTN_upwell %>% 
-  filter(ui.saws >= 0) %>% 
+  filter(ui.saws >= 0) %>% # Upwelling occurs for all values above 0. Values below this is regarded as downwelling
   group_by(site) %>% 
@@
 
+# Detect event: 
 detect_event_custom <- function(df){
@@
   filter(site == "Sea Point", lubridate::year(t) %in% 2016:2018)
-ggplot(SP_upwelling, aes(x = t, y = temp)) +
+plot1 <- ggplot(SP_upwelling, aes(x = t, y = temp)) +
   geom_flame(aes(y = thresh, y2 = temp), fill = "navy", n = 3) +
@@
   geom_rug(data = filter(SP_upwelling, event_no > 0), sides = "b")
+plot1
 
-# We can see in the figure above that the rug plot, which denotes when the wind driven upwelling was detected,
-# occurrs much less frequently than simply using the 30th percentile threshold.
 # The dots on the plot help to show the length of each wind driven upwelling event detected.

### 2019-09-05 SST_patterns.Rmd
updated

@@
 
-# nc_data <- nc_open('/home/amieroh/Documents/Data/Datasets/MUR/daily/20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc')
-# print(nc_data)
+print(nc_data)
 
 ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-csvDir <- "/home/amieroh/Documents"
-
-bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-                   CC = c(25, 35, 340, 355), # Canary Current
-                   CalC = c(35, 45, 225, 240), # California Current
-                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
 
@@
   endDate <- str_sub(ncLast, start = 1, end = 8)
+
+ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
 
@@
 llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+
+MUR <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv03-MUR-20020601-20140727.csv")
+names(MUR)<-c("lon","lat", "temp", "date")
+MUR <- MUR %>% 
+  mutate(temp = temp - 273.15)
+
+JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
 ```
@@
 ```{r}
-ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_global"
+ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
 csvDir <- "/home/amieroh/Documents"
 
-bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-                    CC = c(25, 35, 340, 355), # Canary Current
-                   CalC = c(35, 45, 225, 240), # California Current
-                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+#          1         2         3         4         5         6
+# 123456789012345678901234567890123456789012345678901234567890
+# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
 
-#          1         2         3         4
-# 12345678901234567890123456789012345678901
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
 
-read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-  coords <- bbox[,region]
+  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
     nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
     fNameStem <-
-      substr(ncFile, pathLen + 10, pathLen + 38)
-    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
+      substr(basename(ncFile), 10, 58)
+    fDate <- substr(basename(ncFile), 1, 8)
     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
@@
 
-# the list of files
-ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-strtDate <- str_sub(ncList[1], start = 1, end = 8)
-endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)
-
-# apply the function
-system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))
-
-ncDir <- "/Volumes/Challenger_Deep/OceanData/CMC0.2deg-CMC-L4-GLOB-v2.0/netCDF"
+llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
 ```

### 2019-09-05 upwell_IDX.Rmd
updated

@@
 south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
+south_africa_away
 ```
@@
 
+# Bits of code testing the thresholds against the upwelling literature
 ##################### Upwelling ######################

### 2019-09-05 SST_patterns.Rmd
with CMC new data

@@
 rm(BC_avhrr_only_v2_Document_Document )
+load("Data/CMC.RData")
 ```
@@
 # 
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
 # # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
@@
 ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
-csvDir <- "/home/amieroh/Documents"
+csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
 
@@
 llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+
+CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
+names(CMC)<-c("lon","lat", "temp", "date")
+CMC <- CMC %>% 
+  mutate(temp = temp - 273.15)
+
+# save(CMC, file = "Data/CMC.RData")
 ```
@@
 
-sat_data <- rbind(MUR, OISST) %>% 
+sat_data <- rbind(CMC, OISST) %>% 
   rbind(., CMC) %>% 

### 2019-09-07 SST_patterns.Rmd
Site Map

@@
 
-print(nc_data)
-
 ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
@@
 
-ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
 
@@
 ```{r}
-ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
-csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
-
-#          1         2         3         4         5         6
-# 123456789012345678901234567890123456789012345678901234567890
-# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  strtDate <- str_sub(ncFirst, start = 1, end = 8)
-  endDate <- str_sub(ncLast, start = 1, end = 8)
-
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
-
-  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-    nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/")) + 1
-    fNameStem <-
-      substr(basename(ncFile), 10, 58)
-    fDate <- substr(basename(ncFile), 1, 8)
-    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-      round(4)
-    dimnames(sst) <- list(lon = nc$dim$lon$vals,
-                          lat = nc$dim$lat$vals)
-    nc_close(nc)
-    sst <- as_tibble(melt(sst, value.name = "temp"))
-    sst$t <- ymd(fDate)
-    na.omit(sst)
-    fwrite(sst,
-           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-           append = TRUE, col.names = FALSE)
-    rm(sst)
-  }
-
-llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-
-CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
-names(CMC)<-c("lon","lat", "temp", "date")
-CMC <- CMC %>% 
-  mutate(temp = temp - 273.15)
-
-# save(CMC, file = "Data/CMC.RData")
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
+# 
+# #          1         2         3         4         5         6
+# # 123456789012345678901234567890123456789012345678901234567890
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# 
+# 
+# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
+# names(CMC)<-c("lon","lat", "temp", "date")
+# CMC <- CMC %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# # save(CMC, file = "Data/CMC.RData")
 ```

### 2019-09-07 site_map.Rmd
Site Map

@@
 load("Data/site_pixels.RData")
-load("Data/MUR.RData")
+load("MUR.RData")
 names(south_africa_coast)[1] <- "lon"
@@
         panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.minor = element_line(colour = "NA"),
-        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
-        legend.text = element_text(size = 8),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        legend.text = element_text(size = 12),
         legend.title = element_text(size = 8),
@@
         axis.ticks = element_line(size = 0.5),
-        axis.text = element_text(size = 6)) +
+        axis.text = element_text(size = 10)) +
     guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
@@
 site_map
+
+pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
 

### 2019-09-08 SST_patterns.Rmd
update

@@
 ```{r}
-library(stringr)
-library(tidyverse)
-library(reshape2)
-library(ncdf4) # library for processing netCDFs
-library(plyr)
-library(lubridate)
-library(data.table)
-library(doMC); doMC::registerDoMC(cores = 7)
-
-ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
-
-#          1         2         3         4
-# 12345678901234567890123456789012345678901
-# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-
-read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  strtDate <- str_sub(ncFirst, start = 1, end = 8)
-  endDate <- str_sub(ncLast, start = 1, end = 8)
-
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-
-  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-    nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/")) + 1
-    fNameStem <-
-      substr(basename(ncFile), 10, 38)
-    fDate <- substr(basename(ncFile), 1, 8)
-    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-      round(4)
-    dimnames(sst) <- list(lon = nc$dim$lon$vals,
-                          lat = nc$dim$lat$vals)
-    nc_close(nc)
-    sst <- as_tibble(melt(sst, value.name = "temp"))
-    sst$t <- ymd(fDate)
-    na.omit(sst)
-    fwrite(sst,
-           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-           append = TRUE, col.names = FALSE)
-    rm(sst)
-  }
-
-llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-
-MUR <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv03-MUR-20020601-20140727.csv")
-names(MUR)<-c("lon","lat", "temp", "date")
-MUR <- MUR %>% 
-  mutate(temp = temp - 273.15)
-
-JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# 
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+# 
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# 
+# 
+# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# names(MUR)<-c("lon","lat", "temp", "date")
+# MUR <- MUR %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# 
+# ######################################### SUBSET VIA REGION #################
+# 
+# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+#   
+#   
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)  
+#   
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
 ```
@@
 # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-MUR <- MUR %>% 
-  select(lon, lat) %>% 
-  mutate(product = "MUR")
+# MUR <- MUR %>% 
+#   select(lon, lat) %>% 
+#   mutate(product = "MUR")
 
@@
 
-pixel_match <- OISST %>% 
+pixel_match <- sat_data %>% 
   group_by(product) %>% 
@@
 
+pixel_match <- pixel_match %>%
+  unite(col = combi, lon_sat, lat_sat, sep = "/", remove = F)
+
 # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+
+match_func <- function(df){
+  match <- SACTN_daily_clusters_sub_1  %>% 
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+insitu_MUR <- match_func(df = MUR_data)
+
+# OISST_temp <- OISST %>%
+#   unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+#   filter(combi %in% pixel_match$combi)
+
+# combined_OISST_match <- CMC %>% 
+#   filter(combi %in% pixel_match$combi)
+
 # You will then have one time series for each product for each pixel at each distance for each site that you may then

### 2019-09-08 netCDF2CSV_MUR_CMC_OISST.R

@@ -1,0 +1,237 @@
+# Script which converts netCDF to CSV
+# This script extracts MUR, CMC and OISST data
+
+
+# EXtracting MUR
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+
+  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
+    fNameStem <-
+      substr(basename(ncFile), 10, 38)
+    fDate <- substr(basename(ncFile), 1, 8)
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
+    fwrite(sst,
[…]

### 2019-09-08 SST_patterns.Rmd
fixed_maybe?

@@
 # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-
-match_func <- function(df){
-  match <- SACTN_daily_clusters_sub_1  %>% 
-  left_join(df, by = c("site", "date")) %>% 
-  na.trim()
-  return(match)
+temp_match <- function(df){
+  temp <- df %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
 }
 
-insitu_MUR <- match_func(df = MUR_data)
-
-# OISST_temp <- OISST %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
-#   filter(combi %in% pixel_match$combi)
-
-# combined_OISST_match <- CMC %>% 
-#   filter(combi %in% pixel_match$combi)
+OISST_temp <- temp_match(df = OISST)
+CMC_temp <- temp_match(df = CMC)
 

### 2019-09-08 SST_patterns.Rmd
updated

@@
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  filter(combi %in% pixel_match$combi)
+  filter(combi %in% pixel_match$combi) %>% 
+  select(distance)
 }

### 2019-09-08 upwell_IDX.Rmd
updated

@@
 ##############################################
-
-```{r}
-## Adding a seasonal component to the UI_angle dataset
-season_func <- function(df){
-  season <- df %>%
-  mutate(month = month(t, abbr = T, label = T),
-         year = year(t)) %>%
-  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",
-                          ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
-                              ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-                                      ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
-return(season)
-}
-# UI_angle_season <- season_func(df = UI_angle)
-# save(UI_angle_season, file = "UI_angle_season.RData")
-load("Data/UI_angle_season.RData")
-# Speed of wind increased over time???
-anova_func <- function(df){
-  season_aov <- aov(mean_speed ~ site  * year * season, data = df)
-return(season_aov)
-}
-wind_speed_aov <- anova_func(df = UI_angle_season)
-summary(wind_speed_aov)
-anova_func <- function(df){
-  season_aov <- aov(ui.saws ~ site  * year * season, data = df)
-return(season_aov)
-}
-ui.saws_aov <- anova_func(df = UI_angle_season)
-summary(ui.saws_aov)
-```
-
-# Wind speed intensifed over the years
-Take first third (33percentile) of the data and the last third (66th percentile) of the data for every season and then do a t-test
-
-```{r}
-# load("Data/wind_daily.RData") 
-# 
-# wind_daily <- wind_daily %>% 
-#   dplyr::rename(t = date)
-# 
-# # wind_daily_season <- season_func(df = wind_daily)
-# 
-# #save(wind_daily_season, file = "Data/wind_daily_season.RData")
-# 
-# load("Data/wind_daily_season.RData")
-# 
-# wind_daily_season_test <- wind_daily_season %>% 
-#   group_by(season, temp_sites) %>% 
-#   arrange(mean_speed) %>% 
-#   # Add the following line with mutate to filter the data and add a text label to use with an ANOVA
-#   summarize(msp_1st = quantile(mean_speed, probs=0.33, na.rm=TRUE),
-#             mspeed_2nd = quantile(mean_speed, probs=0.66, na.rm=TRUE))
-```

### 2019-09-08 Write_up/write_up.Rmd
updated

@@
 
-This research study lies within the borders of South Africa along the west coast in particular.
-The locations of each of the points used within this study can be seen in Fig.1.
-The South African Coastal Temperature Network was the primary dataset used to specify the various sites along the west coast of South Africa.
-This dataset consists of coastal seawater temperatures obtained from 129 sites along the entire South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al.
+This research study lies within the Benguela region along the west coast of South Africa.
+Our study sites covered the full range of this coastal region (Fig.1).
+The South African Coastal Temperature Network (SACTN) data serves as the *in situ* temperature dataset used in this study.
+This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al.
 2009).
@@
 
-Inorder to create a time series of each of the remotely sensed SST data representing and also comparing the broad-scale temperatures against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa.
+Inorder to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist at the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa.
 Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
-These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Figure 1).
+These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig.
+1).
 The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
@@
 At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
-The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings].
+The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study.
 Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
+This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].
 In order to examine upwelling patterns between sites, the appropriate sites along the west coast of South Africa (i.e.
 sites influenced by the Benguela current) were selected.
-Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites that could be adequately compared.
-Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 1992 – 2015.
-This yield a total of six sites along the west coast of South Africa.
-Hereafter, each of the remotely sensed SST datasets were matched to the SACTN dataset.
-This was done for two reasons, to find the same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.
+Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
+This yield a total of five sites within the Benguela region.
 
@@
 
-knitr::include_graphics("../Figures/map_fixed.png")
+knitr::include_graphics("../Figures/site_map.png")
 ```
+
+# Defining and calculating the occurance of upwelling
+
+Inorder to determine whether or not upwelling signals are present at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was necessary to define when upwelling is occuring.
+For this research study we aimed to determine a threshold value for which we could identify that upwelling was occuring at a particular time.
+Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider temperature and wind values when determining a threshold for temperature to be experiencing an upwelling event.
+We then decided to make use of the following equation $$ UpwellingIndex = μ{(Cosθ − coastal angle)} $$
+
+In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
+The coastal angle is the angle of each of the sites perpendicular to the coastline.
+Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value is ment to be.
+Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling indeces as 1.
+Inorder to obtain a more approved threshold value, a percentile value was determined.
+To determine changes within the upwelling values, the 'detect event' fuction within the heatwaveR package was used.
+Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
+Values obtained from the upwelling indeces were used as a second threshold.
+An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to further used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
+Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
 

### 2019-09-16 SST_patterns.Rmd
minor edits

--- 2019-09-08 SST_patterns.Rmd
@@
 # load("Data/bathy/sa_bathy.RData")
-# 
+
 # # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
@@
 
-pixel_match <- pixel_match %>%
-  unite(col = combi, lon_sat, lat_sat, sep = "/", remove = F)
-
 # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-
-match_func <- function(df){
-  match <- SACTN_daily_clusters_sub_1  %>% 
-  left_join(df, by = c("site", "date")) %>% 
-  na.trim()
-  return(match)
-}
-
-insitu_MUR <- match_func(df = MUR_data)
-
-# OISST_temp <- OISST %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
-#   filter(combi %in% pixel_match$combi)
-
-# combined_OISST_match <- CMC %>% 
-#   filter(combi %in% pixel_match$combi)
 

### 2019-09-16 Write_up/Write_up_Paper2.Rmd
minor edits

--- 2019-09-03 Write_up/write_up.Rmd
@@
 ---
-title: "Thesis - Upwelling and MHW/Cold spells analyses"
+title: "Paper 2"
 author: "Amieroh Abrahams"
-date: "30 January 2019"
-output:
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
-    highlight: zenburn
-    toc: yes
-  html_document:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  bookdown::html_document2:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  pdf_document:
-    toc: yes
-    toc_depth: 4
-    highlight: default
-    keep_tex: yes
-    latex_engine: xelatex
-fontsize: 10pt
-geometry: margin=1in
-language: Australian
-mainfont: Minion Pro
-monofont: Courier
-papersize: A4
-sansfont: Myriad Pro
-tables: yes
-bibliography: ../bib_master.bib
+date: "16 September 2019"
+output: html_document
 ---
 
-# Chapter 1: Introduction
+## Introduction to paper 2: INcrease in intensity and duration of upwelling events
 
-## Eastern boundary upwelling systems
-
-Globally, there are four major eastern boundary upwelling systems (EBUS).
-Upwelling systems include the Benguela Current off south-western Africa, the Canary Current off north-western Africa together with its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly].
-Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
-EBUS cover only 1% of the world’s ocean surface, but despite this these are among the most productive regions of the world [@pauly].
-This is due to the  systems exposure to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer].
-As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas due to the abundance of nutrients and favourable temperatures.
-
-Upwelling is primarily caused by alongshore, equator ward winds.
-These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods.
-Bakun proposed that if there was an increase in greenhouse gases, it would cause more heating during the day and more cooling during the night causing the temperature gradient to increase resulting in a strong pressure gradient, Ultimately effecting wind patterns and therefore upwelling.
-Several people have tried to model this using different datasets of SST and wind pattern data at various resolutions and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
-These models however, were based on coarse data and when higher resolutions were used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].
-
-### Ecology and primary production within upwelling zones
-
-The duration and intensity of coastal upwelling are known to play a critical role in the phenology of key marine ecosystem processes.
-Similarly, changes in the characteristics of upwelling are shown to cause large disturbances to ecosystems at several trophic levels [@wang].
-Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies].
-In these upwelling systems, the high productivity supports vast and diverse marine populations.
-Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean.
-Most of the world’s major fishing industries are dependent on upwelling, as this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis].
-Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well.
-There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES).
-Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast.
-Inshore regions and embayments that are protected from wind action are likely to be subjected to intense episodes of primary production [@graham].
-Vander Wouder et al.
-(2006) found that these sites represent the highest levels of chlorophyll concentration.
-Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions [@capone].
-
-For several decades, regime shifts have become an important factor of marine research.
-Regime shifts are often defined as a shift from one stable ecosystem state to another.
-This process affects several species within a short period of time [@cury; @jarre; @blamey].
-Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey].
-The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components.
-Over the past  three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
-(2007) associated these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey].
-This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region.
-These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al.
-1993A, @pitcher1998].
-Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf.
-Red tide events are greatest during the latter part of the upwelling season during transitions in the synoptic weather patterns which result in diminished upwelling activity and increased thermal stratification (Pitcher et al.
-1993a).
-
-### Atmosphere
-
-Throughout regions surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high-pressure cells with quasi-stationary positions.
-The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van].
-Solar heating during summer may result in the development of low-pressure cells, known as heat lows, which are absent during the winter [@tyson].
-As the anticyclones shift north during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
-
-## Sea surface temperature and data collection
-
-Water temperature affects the survival, abundance, distribution and growth of marine species (REFERENCE).
-As such, a clear understanding of the relationship between marine ecosystems and their thermal tolerance is necessary for conservation and management planning.
-Future predictions in biogeographical patterns and ecosystem functioning may also be estimated based on various climatic behaviours.
-Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets.
-Unfortunately, not all data are created equally and under the same conditions and as such, not all datasets are appropriate for a particular task.
-
-### *In situ* data
-
-Having coastal, * in situ* seawater temperature data is a rare privilege to researchers.
-Most biological and ecological research requires coastal temperature data.
-The South African Coastal Temperature Network (SACTN).
-The availability of the *in situ* collected coastal temperature data products for several decades provides a reliable source of accurate coastal seawater temperature data [@smit].
-SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments.
-At present, this dataset contains temperatures of 129 sites along the three distinct coasts of the South African coastline.
-Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneven per site.
-Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
-During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
-
-The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors.
-For instance, as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data.
-One problematic outcome of this sampling methodology was that, due to specific work of the South African Weather Service (SAWS) and Kwa-Zulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and recorded at a data precision of 0.500 ◦ C, however, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud].
-Due of these inconsistencies, this dataset required verification from an independent source before its use in research.
-Smit et al.
-(2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results.
-Using these *in situ* collected temperature data along with remotely sensed SST data requires a clear understanding of the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long-term measurements are for use in climate change studies.
-
-It should also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar].
-Specifically, uncertainty lies within the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
-There is certainty that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs in this study.
-Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.
-
-For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset.
-Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers.
-These data points were then changed to NA (not available) so as to not interfere with analysis.
-All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/).
-The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MastersProject.
-
-## Satellite data
-
-Reliable datasets are often difficult to maintain and obtain.
-In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography.
-These datasets however, may be inaccurate along the coast and local *in situ* collected data should be used when possible (Smit et al., 2013).
-Remotely sensed data products are obtained by several different organisations worldwide.
-Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST.
-Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset.
-This is demonstrated as the size of the pixels one sees in the final product.
-
-Satellite SST products undergo a range of quality control and validation processes (e.g.
-Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful.
-As previously mentioned, satellites may show a warm or cold bias along the coastal region but are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault).
-Remotely-sensed SST products are obtained at different resolutions, and as a result do not represent the same level of bias against *in situ* collected seawater temperature.
-Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.
-
-This study made use of various satellite-derived Sea Surface Temperature (SST) datasets.
-One of datasets used was the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST).
-This dataset is an analysis of seawater temperature collected by different platforms (satellites, ships and buoys) on a global grid.
-The gridded daily SST product assimilates both remotely sensed and **in situ** sources of data to create a level-4 gap free product (Banzon et al., 2016).
-A spatially complete sea surface temperature map is produced by interpolation to fill in gaps within the dataset.
-The method used to fill these gaps include the adjustment of satellite and ship observations which compensate for bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst).
-There are two types of daily OISST, named after the relevant satellite sensors, these are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E).
-Currently the AVHRR has the longest record starting from 1981 to present day.
-Infrared instruments, like AVHRR, can make observations at relatively high resolutions.
-Microwave instruments, such as AMSR-E, can measure SSTs in most weather conditions but not adjacent to land.
-The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].
-
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, used as the second dataset, was produced using satellite instruments with datasets spanning from1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals.
-This dataset is currently among the highest resolution SST dataset available.
-The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible.
-The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way in order to produce 1km global maps representing sea surface temperature.
-
-The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates.
-MODIS was launched in 1991 and 2002 to study global dynamics of the earth’s atmosphere, ice and ocean.
-MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong].
-MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
-
-The CMC dataset comprises the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with *in situ* observations of SST from ships and buoys from the ICOADS program.
-
-The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and *in situ* data from drifting and moored buoys [@dash].
-This dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).
-
-### Climate change
-
-Climate change as a result of anthropogenic activities is a global concern of both scientific and political importance.
-As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
-Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney].
-Direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour of marine organisms and in turn, these changes lead to a change in species integration and trophic pathways.
-
-Significant research studies accept that climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh].
-The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of the increase of greenhouse gas concentrations [@solomon].
-A large amount of this heat will be absorbed by the world’s ocean.
-In current years, as a result of changes in air temperature, the heat content of the ocean has increased significantly.
-Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades.
-Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than the global average [@levitus; @santos].
-In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].
-
-As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface.
-This warming slowly propagates down through the water column and in turn increases thermal stratification [@gruber].
-Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016].
-The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therefore strongly linked to biological productivity.
-Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker].
-Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted with an increase in nutrient levels [@rykaczewski].
-Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
-It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which may result in unusual trends.
-
-As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs.
-As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010].
-In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad].
-Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@keeling].
-Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].
-
-Climate change has severe ecological consequences on both plants and animals (Bradley et al.
-1999; Walther et al.
-2002).
-In the past two decades, sea surface temperature has risen in many regions and this seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
-The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993).
-Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process.
-In a bid to further this understanding, Rockström et al.
-(2009) introduced the concept of planetary boundaries.
-According to this concept, within each of these boundaries exists a level of safety that should not be exceeded.
-According to this concept, the planetary boundary for biodiversity loss, has been surpassed [@steffen].
-Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed.
-The idea of planetary boundaries is essential in that it helps in understanding the threats humans cause via the continued destruction of the natural environment.
-Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea.
-Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom].
-Climate change has been proven to decrease primary production thus affecting trophic levels and leading to trophic turbulence and the destruction of marine fauna [@lewandowska; @dell, @mcCauley].
-
-Climate change affects coastal species such as oysters, corals and, sea grass.
-Lord (2017) also found evidence that suggested that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer].
-Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity has been observed to be affected.
-Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis.
-Here the increase in global temperatures is one of the main focus points along with chlorophyll concentration.
-By making use of these two measurements, an analysis of how chlorophyll concentration varies with temperature changes could be conducted.
-
-## Benguela
-
-The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundary upwelling system.
-The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes.
-These currents are forced locally by the wind stress field off southwest Africa [@nelson ; @fennel].
-This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
-Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The main upwelling period in the Benguela, north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year.
-Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August).
-In the central Benguela and also at Cape Frio, upwelling is continuous throughout the year [@shannon1985].
-The cold Benguela waters are bound in the south by the warm Agulhas retroflexion region and in the north by the southward flowing Angola current.
-The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange].
-The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-
-Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not.
-This results in a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
-The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon].
-There have been a number of warm and cool periods documented in the northern and southern Benguela [e.g.
-@stander; @walker; @shannon1984; @boyd; @brainard].
-Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water.
-This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water.
-Another warming event occurred during the summer and autumn months of 1984 [@shannon1986].
-The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela.
-Shannon et al.
-(1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore, west of Luderitz.
-The absence of upwelling between Walvis Bay and north of Luderitz is evident, however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal.
-Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions.
-The seasonal temperature pattern within the northern Benguela was somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
-During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
-
-Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson].
-Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the downwelling process southward.
-Southward, the coastal upwelling system is bounded by the Agulhas Bank.
-South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current [@duncombe].
-The region surrounding the Benguela experiences a persistent along shore wind.
-This wind is associated with the St.
-Helena high pressure system.
-Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.
-These favourable upwelling equatorial winds represent a maximum at 25°S and decrease towards the northern and southern boundaries of the Benguela upwelling system.
-Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast.
-Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel].
-Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher].
-Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind, with wind increasing somewhat away from the coast.
-This facilitated the idea of a wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).
-
-Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986].
-These warm events profoundly impact fisheries and the climate of this region.
-They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer].
-Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures.
-Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region [@sutton; @zebiak].
-Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod].
-Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics.
-Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline.
-It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.
-***Need to make changes to this paragraph, random but feels important***
-
-### Extreme events
-
-Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g.
-@gunter; @lirman; @boucek] or heatwaves [e.g.
-@gordon; @stott; @perkins].
-These events occur rapidly, and unexpectedly and have significant impacts on species diversity, distributions, growth and survival where they take place.
-These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago.
-Jentsch et al.
-(2007) were the first to discuss this mode of climate change and brought it to greater focus.
-Steps have since been taken towards understanding the drivers, impacts and consequences of these events.
-Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming [@mccabe; @harley; @schoepf].
-
-Little is known about extreme events occurring within the South African region.
-Few South African researchers have specifically focussed on investigating this topic, and published literature is sparse.
-During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, *Donax serra Röding*.
-Further studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
-*Acropora cervicornis*, was found to be particularly susceptible to low temperatures and experienced a high mortality rate [@shinn; @davis].
-While these phenomena are not as common as warming events, they have a dramatic and long-lasting effect on coral communities.
-Studies in January 2010 revealed that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman].
-Species population distribution are often set by cold distribution boundaries.
-More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves.
-It is now becoming accepted that extreme events are a threat to several ecosystems [@thompson].
-In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease.
-This thesis aims to address that assumption.
-
-### Event detection
-
-As specified earlier, the extreme cold events occurring along the west coast of South Africa, which is central to this thesis will be detected and correlated to upwelling events with the algorithm developed by Hobday et al.
-(2016).
-The events that recorded *in situ* data from SACTN would be compared against those detected in satellite derived SST.
-For this analyses, the R packages ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events.
-In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred.
-With these annual values it is possible to calculate long-term trends in extreme events.
-
-## Analyses
-
-The analyses conducted in this thesis varied according to the investigations performed.
-The one common thread uniting all of the analyses was that they were conducted in R-programming language (R Core Team, 2017).
-This provided reproducibility for all of the analyses and visualisations used.
-
-## Problem statement:
-
-With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future.
-This increase will undoubtedly have an impact on the world’s oceans and the marine life within it, both directly and indirectly.
-The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
-Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing.
-To assess changes in upwelling events, the aim of this investigation is to use SST and wind data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event.
-Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
-Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS.
-We also aim to determine whether or not there is a change in the rate of onset of upwelling and marine coldspells.
-
-## Aims and approach
-
-The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley].
-This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet].
-Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region.
-This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region.
-Its aims to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
-In this research project, various satellite derived observations of sea surface temperature (SST) and wind stress will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
-
-## Study 1
-
-The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
-2013).
-It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and whether or not the same patterns exist between the different datasets.
-
-## Study 2
-
-The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
-It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises.
-Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
-The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets.
-Thereafter analyses rate of onset of MCS was done.This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends
-
-## Research assumptions
-
-The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared.
-Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
-An additional assumption within this research focused on the effect of upwelling on MCSs and MHWs.
-Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events.
-For the purpose of this study the influence of upwelling on MCSs was the primary focus.
-
-# Paper 1
-## Chapter 2 Upwelling
-### Introduction
-
-Upwelling events are known to have a large influence on the abundance, diversity and distribution of marine organisms and critically affect marine ecosystem processes.
-The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
-Globally, there are four major eastern boundary upwelling systems (EBUS).
-Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
-Eastern boundary upwelling regions only cover 1% of the worlds ocean surface, but are among the most productive regions of the world.
-This is because these regions are influenced by upwelling events.
-Upwelling is defined as a process whereby cold, nutrient rich waters are pushed to the surface as a result of alongshore equatorial winds.
-The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast an d the Benguela current, an eastern boundary upwelling system, along the west coast.
-These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
-2004; Mead et al.
-2013; Schlegel and Smit 2016).
-The Benguela Current is driven by the anticyclone high pressures systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
-2009).
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
-Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al.
-2009; Schlegel et al.
-2017).
-
-Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
-The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases, will result in an increase in day time warming and night time cooling, this ultimately causes an increase in temperature gradients and this further results in a strong pressure gradient.
-Pressure gradients are known to affect wind patterns and wind patterns subsequently control the intensity and duration of upwelling events [@hsieh; @mote; @lima].
-
-Several research studies provide significant evidence suggesting an increase in ocean surface temperatures.
-These temperatures are agreed to have increased as a result of climate change [@gruber].
-Warming is shown to slowly move down into the water column thus increasing thermal stratification.
-This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
-This process is thus strongly linked to primary productivity.
-Climate change impacting ocean processes has major ecological consequences (Bradley et al.
-1999; Walther et al.
-2002).
-For example, several regions show an increase in sea surface temperatures.
-This rise in temperatures largely influence species survival, reproduction and abundance.
-These changes are known to yield several economic consequences (Murawski 1993), thus, understanding the impact and cause of climate change is a challenging phenomenon.
-Similarly, climate change also largely influences coastal species such as corals, muscles and oysters.
-Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-
-The Benguela system is bounded by the equatorward and poleward ends by warm water regimes.
-These currents are shown to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
-Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August.
-The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
-Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events.
-During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
-In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.
-Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]
-
+### Last paragraph in paper 2 introduction ###
 Global warming is increasing at an extremely high rate.
@@
 We hypothesized that (i) wind action may influence the intensity and duration of upwelling; (ii) Remotely sensed SST data will represent a similar event frequency and duration; (iii) Sites located closer to upwelling centres displays distinct variation in the intensity and duration of upwelling events compared to sites furthur away from the upwelling centeres.
-
-## Methods
-
-In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature and wind were analysed and accessed.
-
-## Temperature data sources
-
-This research study lies within the borders of South Africa along the west coast in particular.
-The locations of each of the points used within this study can be seen in Fig.1.
-The South African Coastal Temperature Network was the primary dataset used to specify the various sites along the west coast of South Africa.
-This dataset consists of coastal seawater temperatures obtained from 129 sites along the entire South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al.
-2009).
-These currents are known to largely influence the seawater temperature and marine life along the South African coastline.
-The SACTN has collected SST data along this unique coastline from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
-As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
-The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
-The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.
-During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
-Recordings are still ongoing at many of these sites.
-
-This study also made use of a range of remotely sensed temperature datasets.
-Satellites provide a system of large spatial and temporal data retrieval.
-Remotely sensed SST products are produced by a number of organisations across the globe.
-The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
-OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available.
-The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
-
-An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well.
-Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
-
-Inorder to create a time series of each of the remotely sensed SST data representing and also comparing the broad-scale temperatures against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa.
-Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
-These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Figure 1).
-The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
-
-## Wind data
-
-Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
-Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
-Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
-
-## Site selection
-
-At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
-The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings].
-Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-In order to examine upwelling patterns between sites, the appropriate sites along the west coast of South Africa (i.e.
-sites influenced by the Benguela current) were selected.
-Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites that could be adequately compared.
-Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 1992 – 2015.
-This yield a total of six sites along the west coast of South Africa.
-Hereafter, each of the remotely sensed SST datasets were matched to the SACTN dataset.
-This was done for two reasons, to find the same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.
-
-```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
-The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}
-
-knitr::include_graphics("../Figures/map_fixed.png")
-```
-
-### Detecting the intensity and duration of upwelling
-
-In South Africa, the changes in wind patterns, which occurred in the early 1980s, have resulted in increased upwelling variability and mean upwelling intensity along the west coast in particular, since the early 1990s (Blamey et al.
-2012).
-This is driven by the shift to stronger south-easterly winds which are the drivers of upwelling in this system (Blamey et al.
-2012, Hutchings et al.
-2012).
-Additionally, Rouault et al.
-(2010) have described a negative trend (0.5°C per decade) in seawater temperature along the west coast which is a suggested result of the intensification of upwelling- favourable wind patterns (Blamey et al.
-2015).
-This is contrary to the global trend of warming ocean temperatures as a result of climate change.
-*In order to determine wether upwelling had an influence on MCS, First we had to determine the various upwelling events and then using the same dataset we had to determine when marine cold spells occured or if there was a change in upwelling and
-
-## Determining the upwelling events 
-Known upwelling events and how they link up with Cold spells
-
-- Step 1 Find upwelling events
-- Step 2 when have upwelling see if this match the MCS with that
-
-### MCS
-
-## Results
-
-### Wind patterns during upwelling events
-
-```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagrams representing the the most predominant wind direction for each of the sites for the years 2010 - 2014.
-Each spoke is divided by color into wind speed ranges.
-The radial length of each spoke around the circle is the amount of times that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}
-
- knitr::include_graphics("../Figures/wind.png")
-```
-
-## Chlorophyll a content during upwelling and cold spells

### 2019-09-16 Write_up/write_up.Rmd
minor edits

@@
 ## Chapter 2 Upwelling
+
+## Abstract
+
 ### Introduction
 
-Upwelling events are known to have a large influence on the abundance, diversity and distribution of marine organisms and critically affect marine ecosystem processes.
-The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
-Globally, there are four major eastern boundary upwelling systems (EBUS).
-Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
-Eastern boundary upwelling regions only cover 1% of the worlds ocean surface, but are among the most productive regions of the world.
-This is because these regions are influenced by upwelling events.
-Upwelling is defined as a process whereby cold, nutrient rich waters are pushed to the surface as a result of alongshore equatorial winds.
-The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast an d the Benguela current, an eastern boundary upwelling system, along the west coast.
-These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
+Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of motion caused by alongshore equatorial winds.
+These upwelling events are known to have a large influence on the abundances, diversity, and distributions of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
+Globally, upwelling occurs throughout the world’s oceans but there are four major eastern boundary upwelling systems (EBUS) that offer considerably larger impacts on coastal systems.
+Each of these systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
+Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
+
+The 3000km South African coastline, is bordered by the Agulhas current along the east coast and the EBUS Benguela current, along the west coast.
+These coastal regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
 2004; Mead et al.
 2013; Schlegel and Smit 2016).
-The Benguela Current is driven by the anticyclone high pressures systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
+The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
 2009).
@@
 2017).
+The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
 
@@
 The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases, will result in an increase in day time warming and night time cooling, this ultimately causes an increase in temperature gradients and this further results in a strong pressure gradient.
-Pressure gradients are known to affect wind patterns and wind patterns subsequently control the intensity and duration of upwelling events [@hsieh; @mote; @lima].
+According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling andultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
+Pressure gradients are known to affect wind patterns and thus will subsequently  affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
 
-Several research studies provide significant evidence suggesting an increase in ocean surface temperatures.
-These temperatures are agreed to have increased as a result of climate change [@gruber].
+Several  studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references).
+These temperatures are agreed to have increased as a result of anthropogenically driven climate change [@gruber].
 Warming is shown to slowly move down into the water column thus increasing thermal stratification.
 This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
-This process is thus strongly linked to primary productivity.
-Climate change impacting ocean processes has major ecological consequences (Bradley et al.
+Climate change impacting oceanic processes has major ecological consequences (Bradley et al.
 1999; Walther et al.
@@
 The Benguela system is bounded by the equatorward and poleward ends by warm water regimes.
-These currents are shown to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
+These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
+This EBUS stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
 Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August.
+The upwelling period in the Benguela north of 25°C persist throughout the entire year but peaks during August.
 The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
-Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events.
-During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
-In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.
+Several  studies have shown that extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
+Walter (1937) foundthat warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
+During the periods of warming the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
+In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
+Upwelling occurring along the south coastline of the of Southern Africa are particularly generated by local winds.
 Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]
 
-Global warming is increasing at an extremely high rate.
-This will result in an increase in greenhouse gases which in turn will impact the ocean and terrestrial life.
-The process of upwelling is one such factor that is predicted to change with an increase in global warming.
-As previously mentioned, changes in temperature results in changes in pressure gradients, these pressure gradients control wind patterns which ultimately controls the intensity and duration of upwelling.
-To assess the changes within the upwelling events, various datasets would be used.
-These are satellite sea surface temperatures (broad scale), in situ collected seawater temperatures (local scale), wind data as well as satellite obtained chlorophyll-a data.
-Having assessed the eastern boundry upwelling system within a context that comprised of broad and local-scale features permitted us to assess how upwelling varied between sites located closer to the upwelling centres and furthur away from the upwelling centre.
-Specifically, we aim to assess whether or not it is possible to identify the upwelling signal throughout the various datasets and here we examine how far offshore the upwelling signal is present in the higher resolution data and how far offshore the SACTN signal extend.
-We hypothesized that (i) wind action may influence the intensity and duration of upwelling; (ii) Remotely sensed SST data will represent a similar event frequency and duration; (iii) Sites located closer to upwelling centres displays distinct variation in the intensity and duration of upwelling events compared to sites furthur away from the upwelling centeres.
+The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood..
+Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore (Reference).
+Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
+There is by no means an understanding of how the upwelling process varies at local versus broader scales.
+We hypothesize that upwelling is more dominant within coastal regions instead of broader scale.
+Very few studies have addressed the changes in the intensity or duration of cold temperature spells or cold events at varying distances from the coast within the Benguela region.
+
+The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared.
+Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
+The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
+The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
+2013).
+It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
+This investigation examines recent trends and variation of upwelling events at different distances from the coastline.
 
@@
 
-# Defining and calculating the occurance of upwelling
+# Defining and determining upwelling
 
@@
 Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
+
+##########
 

### 2019-09-16 Write_up/write_up.Rmd
edits

@@
 Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
-
-##########
-
-### Detecting the intensity and duration of upwelling
-
-In South Africa, the changes in wind patterns, which occurred in the early 1980s, have resulted in increased upwelling variability and mean upwelling intensity along the west coast in particular, since the early 1990s (Blamey et al.
-2012).
-This is driven by the shift to stronger south-easterly winds which are the drivers of upwelling in this system (Blamey et al.
-2012, Hutchings et al.
-2012).
-Additionally, Rouault et al.
-(2010) have described a negative trend (0.5°C per decade) in seawater temperature along the west coast which is a suggested result of the intensification of upwelling- favourable wind patterns (Blamey et al.
-2015).
-This is contrary to the global trend of warming ocean temperatures as a result of climate change.
-*In order to determine wether upwelling had an influence on MCS, First we had to determine the various upwelling events and then using the same dataset we had to determine when marine cold spells occured or if there was a change in upwelling and
-
-## Determining the upwelling events 
-Known upwelling events and how they link up with Cold spells
-
-- Step 1 Find upwelling events
-- Step 2 when have upwelling see if this match the MCS with that
-
-### MCS
-
-## Results
-
-### Wind patterns during upwelling events
-
-```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagrams representing the the most predominant wind direction for each of the sites for the years 2010 - 2014.
-Each spoke is divided by color into wind speed ranges.
-The radial length of each spoke around the circle is the amount of times that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}
-
- knitr::include_graphics("../Figures/wind.png")
-```
-
-## Chlorophyll a content during upwelling and cold spells

### 2019-09-16 Write_up/write_up.Rmd
update

@@
 These upwelling events are known to have a large influence on the abundances, diversity, and distributions of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
-Globally, upwelling occurs throughout the world’s oceans but there are four major eastern boundary upwelling systems (EBUS) that offer considerably larger impacts on coastal systems.
-Each of these systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
+Globally, upwelling occurs throughout the world’s oceans, however four major eastern boundary upwelling systems (EBUS) are recognised, each considerably impacting coastal systems.
+These respective systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
 Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
@@
 2017).
-The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
+Additionally,  changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
 
@@
 The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling andultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
+According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
 Pressure gradients are known to affect wind patterns and thus will subsequently  affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
@@
 2002).
-For example, several regions show an increase in sea surface temperatures.
-This rise in temperatures largely influence species survival, reproduction and abundance.
-These changes are known to yield several economic consequences (Murawski 1993), thus, understanding the impact and cause of climate change is a challenging phenomenon.
-Similarly, climate change also largely influences coastal species such as corals, muscles and oysters.
+This is demonstrated in several regions whereby which increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
+Climate change has also been shown tp largely influence coastal species such as corals, muscles and oysters.
 Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
+Furthermore it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993); thus understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in it’s negative affects to local and broader economies as well as human populations dependent on these coastal regions.
 
@@
 These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This EBUS stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
-Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
+This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole].
+With water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
 The upwelling period in the Benguela north of 25°C persist throughout the entire year but peaks during August.
@@
 Several  studies have shown that extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
-Walter (1937) foundthat warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
+Furthermore, Walter (1937) found that warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
 During the periods of warming the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
 In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occurring along the south coastline of the of Southern Africa are particularly generated by local winds.
+Upwelling occurring along the south coastline of the Southern Africa region are particularly generated by local winds.
 Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]

### 2019-09-17 Write_up/write_up.Rmd
update

@@
 These respective systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
+The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which is regarded along with the Peru, California and Humboldt current as Eastern boundry upwelling regions [@andrews].
 Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
+As such there has been a constant interest in upwelling systems.
+Major upwelling programs were carried through off off Peru (e.g.
+WALSH.
+KELLEY, DUGDALE and FROST, 1971) and North-West Africa (e.g.
+BARBER, 1977).
 
@@
 
-The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood..
-Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore (Reference).
+The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood.
+Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore [@andrews].
 Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
@@
 The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
-2013).
+The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit].
 It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
@@
 
-In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature and wind were analysed and accessed.
+## Site selection
+
+At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
+The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study.
+Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the weastern limit near the Namibian border.
+Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
+This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].
+In order to accurately examine upwelling patterns at various distances from the coast several sites were selected as this provides a more accurate representation, the appropriate sites along the west coast of South Africa (i.e.
+sites influenced by the Benguela current) were selected.
+Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
+This yield a total of five sites within the Benguela region (Fig 1).
+
+```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
+The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/site_map.png")
+```
 
@@
 
-This research study lies within the Benguela region along the west coast of South Africa.
-Our study sites covered the full range of this coastal region (Fig.1).
 The South African Coastal Temperature Network (SACTN) data serves as the *in situ* temperature dataset used in this study.
-This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al.
-2009).
-These currents are known to largely influence the seawater temperature and marine life along the South African coastline.
-The SACTN has collected SST data along this unique coastline from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
+This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
+The SACTN has collected SST data along the unique coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
 As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
@@
 
-An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well.
+An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and thus can more accurately explain upwelling within the coastal inshore environment.
 Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
@@
 
-Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
 Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
 Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
+Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
+The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters.
+ERA5 hourly data on pressure levels was used as a second wind source.
+This dataset comprise of u and v variables from 1973 to present.
 
-## Site selection
-
-At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
-The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study.
-Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].
-In order to examine upwelling patterns between sites, the appropriate sites along the west coast of South Africa (i.e.
-sites influenced by the Benguela current) were selected.
-Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
-This yield a total of five sites within the Benguela region.
-
-```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
-The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}
-
-knitr::include_graphics("../Figures/site_map.png")
-```
+With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested.
+To do this, it was important to identify upwelling.
 
@@
 
-Inorder to determine whether or not upwelling signals are present at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was necessary to define when upwelling is occuring.
-For this research study we aimed to determine a threshold value for which we could identify that upwelling was occuring at a particular time.
-Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider temperature and wind values when determining a threshold for temperature to be experiencing an upwelling event.
-We then decided to make use of the following equation $$ UpwellingIndex = μ{(Cosθ − coastal angle)} $$
+Inorder to determine the change in upwelling signals at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was first necessary to define when upwelling was occuring.
+We then aimed to determine a threshold value for which we could identify when upwelling was occuring within this region.
+Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider both temperature and wind values.
+To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989).
+The upwelling index provided an indication of the intensity of the upwelling process taking place using the following equation:
 
-In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
+                  $$ UpwellingIndex = μ{(Cosθ − coastal angle)} $$
+
+In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
 The coastal angle is the angle of each of the sites perpendicular to the coastline.
+The index relies heavily on wind speed and direction data inorder to determine the intensity of upwelling likely to occur.
 Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value is ment to be.
-Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling indeces as 1.
+Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling index as 1.
 Inorder to obtain a more approved threshold value, a percentile value was determined.
@@
 Values obtained from the upwelling indeces were used as a second threshold.
-An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to further used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
+An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
 Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
+For the use of the heatwaveR package it was best to use a time series of 30years.

### 2019-09-17 Write_up/write_up.Rmd
Edits

@@
 
-Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of motion caused by alongshore equatorial winds.
-These upwelling events are known to have a large influence on the abundances, diversity, and distributions of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
-Globally, upwelling occurs throughout the world’s oceans, however four major eastern boundary upwelling systems (EBUS) are recognised, each considerably impacting coastal systems.
-These respective systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
-The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which is regarded along with the Peru, California and Humboldt current as Eastern boundry upwelling regions [@andrews].
+Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of the motion caused by alongshore equatorial winds.
+These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
+Globally, upwelling occurs throughout the world’s oceans, however, four major eastern boundary upwelling systems (EBUS) are recognised, each significantly impacting coastal systems.
+These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly].
+The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews].
 Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
-As such there has been a constant interest in upwelling systems.
-Major upwelling programs were carried through off off Peru (e.g.
+As such, there has been a constant interest in upwelling systems.
+Major upwelling programs have already been carried through off of Peru (e.g.
 WALSH.
@@
 BARBER, 1977).
+
+Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
+The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
+According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
+Pressure gradients are known to affect wind patterns which will subsequently affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
+
+Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references).
+These temperatures have increased as a result of anthropogenically driven climate change [@gruber].
+Warming is shown to slowly move down into the water column thus increasing thermal stratification.
+This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
+Climate change impacting oceanic processes has major ecological consequences (Bradley et al.
+1999; Walther et al.
+2002).
+This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
+Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters.
+Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
+Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993) therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
 
@@
 2017).
-Additionally,  changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
-
-Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
-The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
-Pressure gradients are known to affect wind patterns and thus will subsequently  affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
-
-Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references).
-These temperatures are agreed to have increased as a result of anthropogenically driven climate change [@gruber].
-Warming is shown to slowly move down into the water column thus increasing thermal stratification.
-This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
-Climate change impacting oceanic processes has major ecological consequences (Bradley et al.
-1999; Walther et al.
-2002).
-This is demonstrated in several regions whereby which increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
-Climate change has also been shown tp largely influence coastal species such as corals, muscles and oysters.
-Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-Furthermore it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993); thus understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in it’s negative affects to local and broader economies as well as human populations dependent on these coastal regions.
+Changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
 
@@
 These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole].
-With water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The upwelling period in the Benguela north of 25°C persist throughout the entire year but peaks during August.
+This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole], with water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
+The upwelling period in the Benguela north of 25°C persists throughout the entire year but peaks during August.
 The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-Several  studies have shown that extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
+Several studies have shown that extreme warming and cooling periods occurring in the northern and southern Benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
 Furthermore, Walter (1937) found that warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
-During the periods of warming the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
+During the periods of warming, the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
 In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occurring along the south coastline of the Southern Africa region are particularly generated by local winds.
-Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]
+Upwelling occurring along the south coastline of the Southern Africa region are usually generated by local winds.
+Event-scale upwelling has been demonstrated to have substantial impacts on fisheries [@roy; @van].
+
+The Cape peninsula is influenced by south-east trade winds during summer and westerlies during winter, thus a pronounced seasonal cycle is evident within patterns of wind movement across the coastline.
+During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent.
+During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months.
+During the winter seasons, low pressure cells are driven from the west to the east as a result of the atmospheric circumpolar waves and so north-westerly winds dominate during winter.
+South easterly winds dominate during summer but are still present for short durations during winter [@andrews].
+Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region.
 
@@
 Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
-There is by no means an understanding of how the upwelling process varies at local versus broader scales.
-We hypothesize that upwelling is more dominant within coastal regions instead of broader scale.
+There has not yet been any understanding of how the upwelling process varies at local versus broader scales.
+In this study, it was hypothesize that upwelling is more dominant within coastal regions instead of broader scale.
 Very few studies have addressed the changes in the intensity or duration of cold temperature spells or cold events at varying distances from the coast within the Benguela region.
+
+Strong temperature gradients existing between offshore and inshore oceanic waters may result in different upwelling event readings.
+Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tends to vary based on seasonal wind variation [@andrews].
+Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters[@andrews,@Shannon].
+Due to the temperature gradients that exist inshore and offshore, this study focused on observing upwelling patterns from the coastal region to 50km from the coastline.
 
@@
 It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
-This investigation examines recent trends and variation of upwelling events at different distances from the coastline.
+This investigation examined recent trends and variation of upwelling events at different distances from the coastline.
 

### 2019-09-17 Write_up/write_up.Rmd
updated

@@
 
-At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
-The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study.
-Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the weastern limit near the Namibian border.
-Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].
-In order to accurately examine upwelling patterns at various distances from the coast several sites were selected as this provides a more accurate representation, the appropriate sites along the west coast of South Africa (i.e.
+On a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
+The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings], which provides a natural laboratory for this study.
+Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the Western limit near the Namibian border.
+Seasonal upwelling is controlled by south-easterly trade winds and as a result reflects distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
+This region represents monthly mean temperatures at generally intermediate at cold and warm temperatures [@luning].
+In order to accurately examine upwelling patterns at various distances from the coast, several sites were selected providing a more accurate representation.
+The appropriate sites along the west coast of South Africa (i.e.
 sites influenced by the Benguela current) were selected.
 Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
-This yield a total of five sites within the Benguela region (Fig 1).
+This yielded a total of five sites within the Benguela region (Fig 1).
 
@@
 
-The South African Coastal Temperature Network (SACTN) data serves as the *in situ* temperature dataset used in this study.
+The South African Coastal Temperature Network (SACTN) data served as the *in situ* temperature dataset used in this study.
 This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
-The SACTN has collected SST data along the unique coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
-As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
-The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
+The SACTN has collected SST data along the coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and governmental departments.
+As a result, temperatures within these datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
+The precision of the instruments used vary accordingly, with hand-held thermometers recording temperatures at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
 The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.
 During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
+A further 53 entries were added during the 1980s, 34 during the 1990s, and 18 entries were added during the 2000s.
 Recordings are still ongoing at many of these sites.
@@
 The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
-OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available.
+The AVHRR datasets have been providing global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
+OISST is a global 1/4° gridded, daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
+The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, was produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
+MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and are currently among the highest resolution SST datasets available.
 The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
@@
 
-An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and thus can more accurately explain upwelling within the coastal inshore environment.
-Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
+An advantage of using of *in situ* data over satellite data is that they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and can therefore more accurately explain upwelling within the coastal inshore environment.
+Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm bias reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
 
-Inorder to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist at the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa.
+In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa.
 Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
-These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig.
+These distances were 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig.
 1).
-The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
+The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated in order to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
 
@@
 
-Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
 Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
-Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
+Specific wind characteristics were measured, namely; wind direction (dirw) and wind speed (spw).
 Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
@@
 ERA5 hourly data on pressure levels was used as a second wind source.
-This dataset comprise of u and v variables from 1973 to present.
+This dataset comprised of u and v variables from 1973 to present.
 
@@
 
-Inorder to determine the change in upwelling signals at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was first necessary to define when upwelling was occuring.
-We then aimed to determine a threshold value for which we could identify when upwelling was occuring within this region.
+In order to determine the change in upwelling signals at different distances from the coastline and to test which dataset showed clear, visible representation of these events, it was first necessary to define when upwelling was occurring.
+The next aim was to determine a threshold value to identify when upwelling was occurring within this region.
 Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider both temperature and wind values.
@@
 The coastal angle is the angle of each of the sites perpendicular to the coastline.
-The index relies heavily on wind speed and direction data inorder to determine the intensity of upwelling likely to occur.
-Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value is ment to be.
-Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling index as 1.
-Inorder to obtain a more approved threshold value, a percentile value was determined.
-To determine changes within the upwelling values, the 'detect event' fuction within the heatwaveR package was used.
+The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur.
+Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value was ment to be.
+Here, it was specified that the minimum duration for upwelling was 1 day with a threshold value obtained from the upwelling index as 1.
+In order to obtain a more approved threshold value, a percentile value was determined.
+To determine changes within the upwelling values, the 'detect event' function within the heatwaveR package was used.
 Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
-Values obtained from the upwelling indeces were used as a second threshold.
+Values obtained from the upwelling index were used as a second threshold.
 An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
-Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
+Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months.
 For the use of the heatwaveR package it was best to use a time series of 30years.

### 2019-09-17 Write_up/Paper1_writeup.Rmd
update

--- 2019-09-17 Write_up/write_up.Rmd
@@
 ---
-title: "Thesis - Upwelling and MHW/Cold spells analyses"
+title: "Paper1"
 author: "Amieroh Abrahams"
-date: "30 January 2019"
-output:
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
-    highlight: zenburn
-    toc: yes
-  html_document:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  bookdown::html_document2:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  pdf_document:
-    toc: yes
-    toc_depth: 4
-    highlight: default
-    keep_tex: yes
-    latex_engine: xelatex
-fontsize: 10pt
-geometry: margin=1in
-language: Australian
-mainfont: Minion Pro
-monofont: Courier
-papersize: A4
-sansfont: Myriad Pro
-tables: yes
-bibliography: ../bib_master.bib
+date: "17 September 2019"
+output: html_document
 ---
-
-# Chapter 1: Introduction
-
-## Eastern boundary upwelling systems
-
-Globally, there are four major eastern boundary upwelling systems (EBUS).
-Upwelling systems include the Benguela Current off south-western Africa, the Canary Current off north-western Africa together with its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly].
-Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
-EBUS cover only 1% of the world’s ocean surface, but despite this these are among the most productive regions of the world [@pauly].
-This is due to the  systems exposure to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer].
-As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas due to the abundance of nutrients and favourable temperatures.
-
-Upwelling is primarily caused by alongshore, equator ward winds.
-These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods.
-Bakun proposed that if there was an increase in greenhouse gases, it would cause more heating during the day and more cooling during the night causing the temperature gradient to increase resulting in a strong pressure gradient, Ultimately effecting wind patterns and therefore upwelling.
-Several people have tried to model this using different datasets of SST and wind pattern data at various resolutions and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
-These models however, were based on coarse data and when higher resolutions were used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].
-
-### Ecology and primary production within upwelling zones
-
-The duration and intensity of coastal upwelling are known to play a critical role in the phenology of key marine ecosystem processes.
-Similarly, changes in the characteristics of upwelling are shown to cause large disturbances to ecosystems at several trophic levels [@wang].
-Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies].
-In these upwelling systems, the high productivity supports vast and diverse marine populations.
-Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean.
-Most of the world’s major fishing industries are dependent on upwelling, as this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis].
-Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well.
-There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES).
-Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast.
-Inshore regions and embayments that are protected from wind action are likely to be subjected to intense episodes of primary production [@graham].
-Vander Wouder et al.
-(2006) found that these sites represent the highest levels of chlorophyll concentration.
-Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions [@capone].
-
-For several decades, regime shifts have become an important factor of marine research.
-Regime shifts are often defined as a shift from one stable ecosystem state to another.
-This process affects several species within a short period of time [@cury; @jarre; @blamey].
-Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey].
-The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components.
-Over the past  three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
-(2007) associated these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey].
-This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region.
-These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al.
-1993A, @pitcher1998].
-Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf.
-Red tide events are greatest during the latter part of the upwelling season during transitions in the synoptic weather patterns which result in diminished upwelling activity and increased thermal stratification (Pitcher et al.
-1993a).
-
-### Atmosphere
-
-Throughout regions surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high-pressure cells with quasi-stationary positions.
-The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van].
-Solar heating during summer may result in the development of low-pressure cells, known as heat lows, which are absent during the winter [@tyson].
-As the anticyclones shift north during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
-
-## Sea surface temperature and data collection
-
-Water temperature affects the survival, abundance, distribution and growth of marine species (REFERENCE).
-As such, a clear understanding of the relationship between marine ecosystems and their thermal tolerance is necessary for conservation and management planning.
-Future predictions in biogeographical patterns and ecosystem functioning may also be estimated based on various climatic behaviours.
-Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets.
-Unfortunately, not all data are created equally and under the same conditions and as such, not all datasets are appropriate for a particular task.
-
-### *In situ* data
-
-Having coastal, * in situ* seawater temperature data is a rare privilege to researchers.
-Most biological and ecological research requires coastal temperature data.
-The South African Coastal Temperature Network (SACTN).
-The availability of the *in situ* collected coastal temperature data products for several decades provides a reliable source of accurate coastal seawater temperature data [@smit].
-SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments.
-At present, this dataset contains temperatures of 129 sites along the three distinct coasts of the South African coastline.
-Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneven per site.
-Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
-During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
-
-The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors.
-For instance, as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data.
-One problematic outcome of this sampling methodology was that, due to specific work of the South African Weather Service (SAWS) and Kwa-Zulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and recorded at a data precision of 0.500 ◦ C, however, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud].
-Due of these inconsistencies, this dataset required verification from an independent source before its use in research.
-Smit et al.
-(2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results.
-Using these *in situ* collected temperature data along with remotely sensed SST data requires a clear understanding of the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long-term measurements are for use in climate change studies.
-
-It should also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar].
-Specifically, uncertainty lies within the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
-There is certainty that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs in this study.
-Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.
-
-For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset.
-Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers.
-These data points were then changed to NA (not available) so as to not interfere with analysis.
-All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/).
-The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MastersProject.
-
-## Satellite data
-
-Reliable datasets are often difficult to maintain and obtain.
-In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography.
-These datasets however, may be inaccurate along the coast and local *in situ* collected data should be used when possible (Smit et al., 2013).
-Remotely sensed data products are obtained by several different organisations worldwide.
-Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST.
-Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset.
-This is demonstrated as the size of the pixels one sees in the final product.
-
-Satellite SST products undergo a range of quality control and validation processes (e.g.
-Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful.
-As previously mentioned, satellites may show a warm or cold bias along the coastal region but are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault).
-Remotely-sensed SST products are obtained at different resolutions, and as a result do not represent the same level of bias against *in situ* collected seawater temperature.
-Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.
-
-This study made use of various satellite-derived Sea Surface Temperature (SST) datasets.
-One of datasets used was the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST).
-This dataset is an analysis of seawater temperature collected by different platforms (satellites, ships and buoys) on a global grid.
-The gridded daily SST product assimilates both remotely sensed and **in situ** sources of data to create a level-4 gap free product (Banzon et al., 2016).
-A spatially complete sea surface temperature map is produced by interpolation to fill in gaps within the dataset.
-The method used to fill these gaps include the adjustment of satellite and ship observations which compensate for bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst).
-There are two types of daily OISST, named after the relevant satellite sensors, these are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E).
-Currently the AVHRR has the longest record starting from 1981 to present day.
-Infrared instruments, like AVHRR, can make observations at relatively high resolutions.
-Microwave instruments, such as AMSR-E, can measure SSTs in most weather conditions but not adjacent to land.
-The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].
-
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, used as the second dataset, was produced using satellite instruments with datasets spanning from1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals.
-This dataset is currently among the highest resolution SST dataset available.
-The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible.
-The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way in order to produce 1km global maps representing sea surface temperature.
-
-The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates.
-MODIS was launched in 1991 and 2002 to study global dynamics of the earth’s atmosphere, ice and ocean.
-MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong].
-MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
-
-The CMC dataset comprises the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with *in situ* observations of SST from ships and buoys from the ICOADS program.
-
-The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and *in situ* data from drifting and moored buoys [@dash].
-This dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).
-
-### Climate change
-
-Climate change as a result of anthropogenic activities is a global concern of both scientific and political importance.
-As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
-Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney].
-Direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour of marine organisms and in turn, these changes lead to a change in species integration and trophic pathways.
-
-Significant research studies accept that climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh].
-The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of the increase of greenhouse gas concentrations [@solomon].
-A large amount of this heat will be absorbed by the world’s ocean.
-In current years, as a result of changes in air temperature, the heat content of the ocean has increased significantly.
-Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades.
-Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than the global average [@levitus; @santos].
-In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].
-
-As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface.
-This warming slowly propagates down through the water column and in turn increases thermal stratification [@gruber].
-Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016].
-The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therefore strongly linked to biological productivity.
-Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker].
-Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted with an increase in nutrient levels [@rykaczewski].
-Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
-It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which may result in unusual trends.
-
-As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs.
-As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010].
-In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad].
-Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@keeling].
-Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].
-
-Climate change has severe ecological consequences on both plants and animals (Bradley et al.
-1999; Walther et al.
-2002).
-In the past two decades, sea surface temperature has risen in many regions and this seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
-The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993).
-Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process.
-In a bid to further this understanding, Rockström et al.
-(2009) introduced the concept of planetary boundaries.
-According to this concept, within each of these boundaries exists a level of safety that should not be exceeded.
-According to this concept, the planetary boundary for biodiversity loss, has been surpassed [@steffen].
-Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed.
-The idea of planetary boundaries is essential in that it helps in understanding the threats humans cause via the continued destruction of the natural environment.
-Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea.
-Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom].
-Climate change has been proven to decrease primary production thus affecting trophic levels and leading to trophic turbulence and the destruction of marine fauna [@lewandowska; @dell, @mcCauley].
-
-Climate change affects coastal species such as oysters, corals and, sea grass.
-Lord (2017) also found evidence that suggested that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer].
-Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity has been observed to be affected.
-Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis.
-Here the increase in global temperatures is one of the main focus points along with chlorophyll concentration.
-By making use of these two measurements, an analysis of how chlorophyll concentration varies with temperature changes could be conducted.
-
-## Benguela
-
-The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundary upwelling system.
-The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes.
-These currents are forced locally by the wind stress field off southwest Africa [@nelson ; @fennel].
-This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
-Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The main upwelling period in the Benguela, north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year.
-Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August).
-In the central Benguela and also at Cape Frio, upwelling is continuous throughout the year [@shannon1985].
-The cold Benguela waters are bound in the south by the warm Agulhas retroflexion region and in the north by the southward flowing Angola current.
-The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange].
-The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-
-Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not.
-This results in a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
-The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon].
-There have been a number of warm and cool periods documented in the northern and southern Benguela [e.g.
-@stander; @walker; @shannon1984; @boyd; @brainard].
-Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water.
-This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water.
-Another warming event occurred during the summer and autumn months of 1984 [@shannon1986].
-The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela.
-Shannon et al.
-(1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore, west of Luderitz.
-The absence of upwelling between Walvis Bay and north of Luderitz is evident, however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal.
-Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions.
-The seasonal temperature pattern within the northern Benguela was somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
-During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
-
-Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson].
-Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the downwelling process southward.
-Southward, the coastal upwelling system is bounded by the Agulhas Bank.
-South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current [@duncombe].
-The region surrounding the Benguela experiences a persistent along shore wind.
-This wind is associated with the St.
-Helena high pressure system.
-Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.
-These favourable upwelling equatorial winds represent a maximum at 25°S and decrease towards the northern and southern boundaries of the Benguela upwelling system.
-Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast.
-Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel].
-Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher].
-Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind, with wind increasing somewhat away from the coast.
-This facilitated the idea of a wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).
-
-Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986].
-These warm events profoundly impact fisheries and the climate of this region.
-They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer].
-Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures.
-Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region [@sutton; @zebiak].
-Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod].
-Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics.
-Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline.
-It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.
-***Need to make changes to this paragraph, random but feels important***
-
-### Extreme events
-
-Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g.
-@gunter; @lirman; @boucek] or heatwaves [e.g.
-@gordon; @stott; @perkins].
-These events occur rapidly, and unexpectedly and have significant impacts on species diversity, distributions, growth and survival where they take place.
-These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago.
-Jentsch et al.
-(2007) were the first to discuss this mode of climate change and brought it to greater focus.
-Steps have since been taken towards understanding the drivers, impacts and consequences of these events.
-Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming [@mccabe; @harley; @schoepf].
-
-Little is known about extreme events occurring within the South African region.
-Few South African researchers have specifically focussed on investigating this topic, and published literature is sparse.
-During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, *Donax serra Röding*.
-Further studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
-*Acropora cervicornis*, was found to be particularly susceptible to low temperatures and experienced a high mortality rate [@shinn; @davis].
-While these phenomena are not as common as warming events, they have a dramatic and long-lasting effect on coral communities.
-Studies in January 2010 revealed that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman].
-Species population distribution are often set by cold distribution boundaries.
-More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves.
-It is now becoming accepted that extreme events are a threat to several ecosystems [@thompson].
-In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease.
-This thesis aims to address that assumption.
-
-### Event detection
-
-As specified earlier, the extreme cold events occurring along the west coast of South Africa, which is central to this thesis will be detected and correlated to upwelling events with the algorithm developed by Hobday et al.
-(2016).
-The events that recorded *in situ* data from SACTN would be compared against those detected in satellite derived SST.
-For this analyses, the R packages ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events.
-In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred.
-With these annual values it is possible to calculate long-term trends in extreme events.
-
-## Analyses
-
-The analyses conducted in this thesis varied according to the investigations performed.
-The one common thread uniting all of the analyses was that they were conducted in R-programming language (R Core Team, 2017).
-This provided reproducibility for all of the analyses and visualisations used.
-
-## Problem statement:
-
-With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future.
-This increase will undoubtedly have an impact on the world’s oceans and the marine life within it, both directly and indirectly.
-The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
-Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing.
-To assess changes in upwelling events, the aim of this investigation is to make use of SST and wind data to monitor the durations and intensities of upwelling events at different distances from the coastline within the Benguela EBUS.
-
-Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
-TThereafter,Inorder to determine whether upwelling events had resulted in an increase in the onset of marine cold spells, all four easter boundry upwelling systems were considred.
-
-## Aims and approach
-
-The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley].
-This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet].
-Very few studies have addressed the occurrence and the changes in the intensity or duration of cold temperature spells or cold events within the Benguela region.
-This investigation examines recent trends and variation in the frequency and duration of upwelling events occurring in this region In this research project, various satellite derived observations of sea surface temperature (SST) and wind will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
-Thereafter, satellite data will be used to examine upwelling on a global scale and whether these events may have influence the rate of onset of marine cold spells.
-
-## Study 1
-
-The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
-2013).
-It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
-
-## Study 2
-
-The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for all of the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves.
-It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and over the years given the influence of climate change and sea surface temperature rises.
-Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
-The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets.
-Thereafter analyses rate of onset of MCS was done.This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends
-
-## Research assumptions
-
-The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared.
-Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
-An additional assumption within this research focused on the influence of the changes of upwelling on marine cold spells.
-Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events.
-For the purpose of this study the influence of upwelling on MCSs was the primary focus.
 

### 2019-09-18 Download_MODIS.R
Download Error

@@ -1,0 +1,11 @@
+## Downloading MODIS
+https://oceancolor.gsfc.nasa.gov/forum/oceancolor/topic_show.pl?pid=12520
+
+
+# K10 SST download
+https://podaac.jpl.nasa.gov/dataset/NAVO-L4HR1m-GLOB-K10_SST?ids=Platform:ProcessingLevel&values=Aqua:*4* 
+wget -q --post-data="sensor=aqua&sdate=2009-07-04&edate=2018-12-13&dtype=L4&addurl=1&results_as_file=1&search=*PODAAC-GHK10-41N01*" -O - https://oceandata.sci.gsfc.nasa.gov/api/file_search |wget -i -
+  
+  
+# This wget script will download the MODIS-Aqua Chlorophyll Concentration, OCI Algorithm, 8-day composite, for the period 2002-07-04 to 2018-12-13:
+wget -q --post-data="sensor=aqua&sdate=2002-07-04&edate=2018-12-13&dtype=L3m&addurl=1&results_as_file=1&search=*L3m_8D_CHL_chlor_a_9km*" -O - https://oceandata.sci.gsfc.nasa.gov/api/file_search |wget -i -

### 2019-09-18 Write_up/Paper1_writeup.Rmd
updated

@@
 ---
-title: "Paper1"
+title: "Upwelling variability across the Benguela upwelling systems"
 author: "Amieroh Abrahams"
-date: "17 September 2019"
-output: html_document
+date: "18 September 2019"
+output:
+  word_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
+    highlight: zenburn
+    toc: yes
+  html_document:
+    fig_caption: yes
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: sandstone
+    toc_float: no
+  bookdown::html_document2:
+    fig_caption: yes
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: sandstone
+    toc_float: no
+  pdf_document:
+    toc: yes
+    toc_depth: 4
+    highlight: default
+    keep_tex: yes
+    latex_engine: xelatex
+fontsize: 10pt
+geometry: margin=1in
+language: Australian
+mainfont: Minion Pro
+monofont: Courier
+papersize: A4
+sansfont: Myriad Pro
+tables: yes
+bibliography: ../bib_master.bib
 ---

### 2019-09-18 Write_up/Abstract_edits/Paper1_abstract.Rmd
update

---
title: "Abstract"
author: "Amieroh Abrahams"
date: "18 September 2019"
output: html_document
---

## Abstract

### 2019-09-18 Write_up/Acknowledgements_edits/Paper1_acknowledgements.Rmd
update

---
title: "Acknowledgements"
author: "Amieroh Abrahams"
date: "18 September 2019"
output: html_document
---

## Acknowledgements

We would like to thank SAWS, DAFF, DEA, EKZNW and SAEON for the contribution of all the raw data used in this study. Without the collection of this data this article and the South African Coastal Temperature Network (SACTN) would not be possible. This research is funded by a South African National Research Foundation Grant (SFH180604339745). The data and analyses used in this paper may be found at [Add appendix link here]

### 2019-09-19 SST_patterns.Rmd
update
incomplete commit

@@
 
-OISST <- OISST %>% 
+# Decided to work with OISST and CMC as both has a time series of 30years
+#
+
+OISST_prod <- OISST %>% 
   select(lon, lat) %>% 
@@
 
-CMC <- CMC %>% 
+CMC_prod <- CMC %>% 
   select(lon, lat) %>% 
@@
 
-sat_data <- rbind(CMC, OISST) %>% 
-  rbind(., CMC) %>% 
+sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+  #rbind(., CMC) %>% 
   select(product, lon, lat)
@@
   select(-data) %>%
-  unnest(match)
+  unnest(match) %>% 
+  unite(col = combi, lon, lat, sep = "/", remove = F)
 
 # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+# one time series for each product for each pixel at each distance for each site that you may then
+dataset_filt <- function(df){
+  filtered_products <- OISST %>% 
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+  filter(combi %in% pixel_match$combi)
+}
 
-# You will then have one time series for each product for each pixel at each distance for each site that you may then
-# run the upwelling event function on
-# You should do this via purrr::nest+map as you can see how much less code it requires
+OISST_fil <- dataset_filt(df = OISST)
+CMC_fil <- dataset_filt(df = CMC)
+
+# Repeat this on MODIS K10 and another time series of 10 years
+
 ```

### 2019-09-19 SST_patterns.Rmd
left_join

@@
 load("Data/SACTN_US.RData")
-load("Data/site_pixels.RData")
-load("Data/OISST.RData")
+load("Data/site_pixels.RData") # 5 decimal places
+load("Data/OISST.RData") # 2 decimal places
 OISST <- BC_avhrr_only_v2_Document_Document 
 rm(BC_avhrr_only_v2_Document_Document )
-load("Data/CMC.RData")
+load("Data/CMC.RData") # 1decimal places
 ```
@@
 # load("Data/site_pixels.RData")
+
 ```
@@
 
+site_pixels$lat<-round(site_pixels$lat, 1) #CMC is to 1 decimal place and OISST is to 2
+site_pixels$lon<-round(site_pixels$lon, 1)
+
 # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
@@
 # Decided to work with OISST and CMC as both has a time series of 30years
-#
 
@@
   unique()
-
 ## For testing the nest/map pipeline
@@
     dplyr::rename(lon_sat = lon, lat_sat = lat)
-  match_index <- knnx.index(data = as.matrix(site_pixels[,2:3]),
+  match_index <- knnx.index(data = as.matrix(round(site_pixels[,2:3]),1),
                     query = as.matrix(df[,1:2]), k = 1)
@@
   select(-data) %>%
-  unnest(match) %>% 
-  unite(col = combi, lon, lat, sep = "/", remove = F)
+  unnest(match) #%>% 
+  #unite(col = combi, lon, lat, sep = "/", remove = F)
+
+pixel_match$lat<-round(pixel_match$lat, 1) #CMC is to 1 decimal place and OISST is to 2
+pixel_match$lon<-round(pixel_match$lon, 1)
 
@@
 # one time series for each product for each pixel at each distance for each site that you may then
-dataset_filt <- function(df){
-  filtered_products <- OISST %>% 
-  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
-  filter(combi %in% pixel_match$combi)
+# dataset_filt <- function(df){
+#   filtered_products <- df %>% 
+#   unite(col = combi, lon, lat, sep = "/", remove = F) #%>% 
+#   #filter(combi %in% pixel_match$combi)
+# }
+
+# OISST_fil <- dataset_filt(df = OISST)
+# CMC_fil <- dataset_filt(df = CMC)
+# 
+# 
+# pixel_match$lat
+# pixel_match$lon
+
+filterted_product <- function(df){
+  extracted <- pixel_match %>% 
+  right_join(df, by = c("lon", "lat")) %>% 
+  select(date,temp,lat,lon)
 }
 
-OISST_fil <- dataset_filt(df = OISST)
-CMC_fil <- dataset_filt(df = CMC)
+OISST_fil <- filterted_product(df = OISST)
+CMC_fil <- filterted_product(df = CMC)
 
-# Repeat this on MODIS K10 and another time series of 10 years
-
+# Next to run upwelling index
 ```

### 2019-09-19 SST_patterns.Rmd
Fixed
updated

@@
 library(plyr)
+library(heatwaveR)
 ```
@@
 
-OISST_fil <- filterted_product(df = OISST)
-CMC_fil <- filterted_product(df = CMC)
+OISST_fil <- filterted_product(df = OISST) # Need to add a site column
+CMC_fil <- filterted_product(df = CMC) # Need to add a site column
 
+# OISST_sites <- OISST_fil %>% 
+#   left_join(site_list_sub[,c(2,3)], by = c("lat","lon"))
+
+# All the code below can be viewed in upwell_IDX.Rmd
 # Next to run upwelling index
+# UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+load("Data/UI_angle.RData")
+
+upwelling <- UI_angle %>% 
+  dplyr::rename(temp = ui.saws) %>%
+  group_by(site) %>%
+  # mutate(min_t = min(t), 
+  #        max_t = max(t)) %>% 
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climaatologies. Column will be named clim
+         # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+Upwelling can last for a few hours.
+Given that we have daily data, upwelling events minimum duration here will be 1day
+  # Detect consecutive days in exceedance of a given threshold.
+  # mutate() %>% 
+  select(-data, -clim) %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance)
+
+# Now applying the Upwelling func
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
+  return(res)
+}
+
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+upwell_base <- df %>% 
+  dplyr::rename(t = date) %>% 
+  group_by(lat,lon) %>% 
+  nest() %>% 
+  # The climatology base period used here is up for debate...
+  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
+  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1989-01-01", "2017-12-31"))) %>%
+  select(-data) %>% 
+  unnest() %>%
+  left_join(upwelling, by = c("site", "t")) %>%
+  filter(!is.na(exceedance)) %>%
+  group_by(lat, lon) %>% 
+  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
+  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
+  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
+  nest() %>% 
+  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
+  select(-data)
+}
+
+OISST_upwell_base <- upwelling_detect_event(df = OISST_fil)
+CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
 ```

### 2019-09-20 netCDF2CSV_ERA5.R
Extracting_scripts

@@ -1,0 +1,141 @@
+# Extracting ERA 5 wind u and v variables
+# N/W/S/E coordinate format when downloading
+# -26.00/15.00/-36.00/22.00
+
+# Loading libraries
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+# ERA.dir <- "/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate"
+# ERA.csv.dir <- "/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate/Extracted"
+
+####### u-wind 2008-2019
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # wu_daily08-19.nc
+
+ncFile <- '/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate/wu_daily08-19.nc'  
+nc <- nc_open(ncFile)
+fNameStem <-
+  substr(basename(ncFile), 1, 13)
+u_10 <- ncvar_get(nc, varid = "u10") %>%
+  round(4)
+dimnames(u_10) <- list(lon = nc$dim$lon$vals,
+                         lat = nc$dim$lat$vals,
+                         time = nc$dim$time$vals)
+nc_close(nc)
+u_10 <- as_tibble(melt(u_10, value.name = "u_10"))
+u_10$time <- as.POSIXct(u_10$time, origin = "1970-01-01")
+
+
+
+###### v-wind 2008-2019
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # wu_daily08-19.nc
+
+ncFile <- '/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate/wv_daily08-19.nc'
+nc <- nc_open(ncFile)
+fNameStem <-
+  substr(basename(ncFile), 1, 13)
+v_10 <- ncvar_get(nc, varid = "v10") %>%
+  round(4)
+dimnames(v_10) <- list(lon = nc$dim$lon$vals,
+                       lat = nc$dim$lat$vals,
+                       time = nc$dim$time$vals)
+nc_close(nc)
+v_10 <- as_tibble(melt(v_10, value.name = "v_10"))
+v_10$time <- as.POSIXct(v_10$time, origin = "1970-01-01")
+
[…]

### 2019-09-20 netCDF2CSV_MUR_CMC_OISST.R

@@ -2,7 +2,7 @@
 # This script extracts MUR, CMC and OISST data
 
 
-# EXtracting MUR
+# Laoding Libraries
 library(stringr)
 library(tidyverse)
 library(reshape2)

### 2019-09-20 SST_patterns.Rmd
updated
The pixel filtering by distance now functions as expected

--- 2019-09-19 SST_patterns.Rmd
@@
 library(ncdf4) # library for processing netCDFs
-library(ncdf4)
 library(data.table)
-library(plyr)
+# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
+library(heatwaveR)
 ```
@@
 OISST <- BC_avhrr_only_v2_Document_Document 
-rm(BC_avhrr_only_v2_Document_Document )
+rm(BC_avhrr_only_v2_Document_Document ); gc()
 load("Data/CMC.RData") # 1decimal places
@@
 ```{r}
-# RWS: It appears as though the satellite data you are loading here have already been trimmed down to the one nearest pixel to each SACTN site
-# This won't work for the following steps because you need to be finding pixels from the COMPLETE set of satellite data you downloaded
-# Once you load the complete satellite data lon/lat values you can run the following code and it should give you the coordinates you want
-
-site_pixels$lat<-round(site_pixels$lat, 1) #CMC is to 1 decimal place and OISST is to 2
-site_pixels$lon<-round(site_pixels$lon, 1)
-
 # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
@@
 # Decided to work with OISST and CMC as both has a time series of 30years
-
 OISST_prod <- OISST %>% 
   select(lon, lat) %>% 
+  unique() %>% 
   mutate(product = "OISST")
@@
   select(lon, lat) %>%
+  unique() %>% 
   mutate(product = "CMC")
 
-sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+# sat_data <- rbind(CMC_prod, OISST_prod) %>% 
   #rbind(., CMC) %>% 
-  select(product, lon, lat)
+  # select(product, lon, lat)
 
-# rm(MUR, OISST, CMC); gc()
+# sat_pixels <- sat_data %>% 
+  # select(product, lon, lat) %>%
+  # unique()
 
-sat_pixels <- sat_data %>% 
-  select(product, lon, lat) %>%
-  unique()
 ## For testing the nest/map pipeline
-# df <- sat_pixels %>% 
-#   filter(product == "MUR", site == "Lamberts Bay") %>% 
-#   select(-c(product, site))
+# df <- site_pixels %>%
+#   filter(site == "Lamberts Bay") %>%
+#   select(-site)
 
 match_func <- function(df){
-  df <- na.omit(df) %>% 
-    dplyr::rename(lon_sat = lon, lat_sat = lat)
-  match_index <- knnx.index(data = as.matrix(round(site_pixels[,2:3]),1),
-                    query = as.matrix(df[,1:2]), k = 1)
-  res <- cbind(site_pixels[match_index,], df)
+  df <- df %>%
+    dplyr::rename(lon_site = lon, lat_site = lat)
+  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
+    cbind(., df)
+  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
+                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
+    cbind(., df)
+  res <- rbind(OISST_index, CMC_index)
   return(res)
@@
 
-pixel_match <- sat_data %>% 
-  group_by(product) %>% 
-  nest() %>% 
-  mutate(match = purrr:::map(data, match_func)) %>% 
-  select(-data) %>%
-  unnest(match) #%>% 
-  #unite(col = combi, lon, lat, sep = "/", remove = F)
-
-pixel_match$lat<-round(pixel_match$lat, 1) #CMC is to 1 decimal place and OISST is to 2
-pixel_match$lon<-round(pixel_match$lon, 1)
+# Find the nearest pixels to each spot along the transect
+# NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
+# points to not fall inside of the same 25 KM pixel
+pixel_match <- site_pixels %>% 
+  group_by(site) %>% 
+  group_modify(~match_func(.x))
 
 # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-# one time series for each product for each pixel at each distance for each site that you may then
-# dataset_filt <- function(df){
-#   filtered_products <- df %>% 
-#   unite(col = combi, lon, lat, sep = "/", remove = F) #%>% 
-#   #filter(combi %in% pixel_match$combi)
-# }
+OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
 
-# OISST_fil <- dataset_filt(df = OISST)
-# CMC_fil <- dataset_filt(df = CMC)
-# 
-# 
-# pixel_match$lat
-# pixel_match$lon
+# Clean up some RAM space
+rm(OISST, CMC); gc()
 
-filterted_product <- function(df){
-  extracted <- pixel_match %>% 
-  right_join(df, by = c("lon", "lat")) %>% 
-  select(date,temp,lat,lon)
+# All the code below can be viewed in upwell_IDX.Rmd
+# Next to run upwelling index
+# UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+load("Data/UI_angle.RData")
+
+upwelling <- UI_angle %>% 
+  dplyr::rename(temp = ui.saws) %>%
+  group_by(site) %>%
+  # mutate(min_t = min(t), 
+  #        max_t = max(t)) %>% 
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+         # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+Upwelling can last for a few hours.
+Given that we have daily data, upwelling events minimum duration here will be 1day
+  # Detect consecutive days in exceedance of a given threshold.
+  # mutate() %>% 
+  select(-data, -clim) %>% 
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance)
+
+# Now applying the Upwelling func
+# NB: This only pulls out the event results and not the climatology results
+# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
+# To create a list output one would use group_map(), 
+# but this then loses the labels of which sites etc. the results belong to
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
+  return(res)
 }
 
-OISST_fil <- filterted_product(df = OISST)
-CMC_fil <- filterted_product(df = CMC)
+ts2clm_custom <- function(df){
+  # The climatology base period used here is up for debate...
+  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  return(res)
+}
 
-# Next to run upwelling index
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+  upwell_base <- df %>% 
+    dplyr::rename(t = date) %>% 
+    group_by(site, product, heading, distance, lon, lat) %>% 
+    group_modify(~ts2clm_custom(.x)) %>% 
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, product, heading, distance, lon, lat) %>% 
+    group_modify(~detect_event_custom(.x))
+  }
+
+OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
+save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
+save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
 ```

### 2019-09-22 SACTN_lts.Rmd
some_changes

@@
   filter(length > 10950) 
-# save(site_list_sub, file = "site_list_sub.Rdata")
+#site_list_sub <- site_list_sub[-5,]
+save(site_list_sub, file = "Data/site_list_sub.Rdata")
 
@@
 
-# save(SACTN_US, file = "Data/SACTN_US.RData")
+save(SACTN_US, file = "Data/SACTN_US.RData")
 ```

### 2019-09-22 SST_patterns.Rmd
some_changes

@@
 CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+SACTN_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
 
@@
 rm(OISST, CMC); gc()
+
+# sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
+# # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
+# # Hout Bay will be ignored
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+
+OISST_fill <- OISST_fill %>%
+  filter(site %in% selected_sites)
+CMC_fill <- OISST_fill %>%
+  filter(site %in% selected_sites)
 
@@
 save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
+
+# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
+SACTN_US <- SACTN_US %>%
+  filter(site %in% selected_sites)
+
+SACTN_upwell_base <- SACTN_US %>%
+  left_join(site_list[,c(4, 5, 6)], by = "index")%>% 
+    dplyr::rename(t = date) %>% 
+    group_by(site, lon, lat) %>% 
+    group_modify(~ts2clm_custom(.x)) %>% 
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, lon, lat) %>% 
+    group_modify(~detect_event_custom(.x))
+
+save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
 ```

### 2019-09-22 upwell_IDX.Rmd
some_changes

@@
 ```{r}
-library(circular)
-# devtools::install_github("robwschlegel/coastR")
-library(gridExtra)
-library(geosphere)
-library(tidyverse)
-library(heatwaveR)
-library(coastR)
-```
-
-# Steps
-  - Use the circular function 
-  - Determine daily wind data obtained from 3hr intervals
-
-```{r}
+# library(circular)
+# # devtools::install_github("robwschlegel/coastR")
+# library(gridExtra)
+# library(geosphere)
+# library(tidyverse)
+# library(heatwaveR)
+# library(coastR)
+# ```
+# 
+# # Steps
+#   - Use the circular function 
+#   - Determine daily wind data obtained from 3hr intervals
+# 
+# ```{r}
 # wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
@@
 #                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
+#                                     #station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
 #                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
@@
 # wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-# # save(wind_data, file = "Data/wind_data.RData")
+# save(wind_data, file = "Data/wind_data.RData")
 # 
@@
 #                    mean_speed = round(mean(speed),2)) # Calculating the mean
-# #save(wind_daily, file = "Data/wind_daily.RData")
+# save(wind_daily, file = "Data/wind_daily.RData")
 ```

### 2019-09-22 site_map.Rmd
Map update

@@
 xtable(site_list_sub, auto = TRUE)
-west <- site_list_sub
+west <- site_list_sub[-5,]
 west$coast <- "west" # Chnages wc to west
+
+site_list_sub <- site_list_sub[-5,]
 
@@
 
+site_list_sub <- site_list_sub[-5,]
+
 # Setting up the theme

### 2019-09-22 SST_patterns.Rmd
update

--- 2019-09-20 SST_patterns.Rmd
@@
 ```{r}
-# These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# MUR <- MUR %>% 
+# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
+# # MUR <- MUR %>% 
+# #   select(lon, lat) %>% 
+# #   mutate(product = "MUR")
+# 
+# # Decided to work with OISST and CMC as both has a time series of 30years
+# OISST_prod <- OISST %>% 
 #   select(lon, lat) %>% 
-#   mutate(product = "MUR")
-
-# Decided to work with OISST and CMC as both has a time series of 30years
-OISST_prod <- OISST %>% 
-  select(lon, lat) %>% 
-  unique() %>% 
-  mutate(product = "OISST")
-
-CMC_prod <- CMC %>% 
-  select(lon, lat) %>%
-  unique() %>% 
-  mutate(product = "CMC")
-
-# sat_data <- rbind(CMC_prod, OISST_prod) %>% 
-  #rbind(., CMC) %>% 
-  # select(product, lon, lat)
-
-# sat_pixels <- sat_data %>% 
-  # select(product, lon, lat) %>%
-  # unique()
-
-## For testing the nest/map pipeline
-# df <- site_pixels %>%
-#   filter(site == "Lamberts Bay") %>%
-#   select(-site)
-
-match_func <- function(df){
-  df <- df %>%
-    dplyr::rename(lon_site = lon, lat_site = lat)
-  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
-                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
-    cbind(., df)
-  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
-                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
-    cbind(., df)
-  res <- rbind(OISST_index, CMC_index)
-  return(res)
-}
-
-# Find the nearest pixels to each spot along the transect
-# NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
-# points to not fall inside of the same 25 KM pixel
-pixel_match <- site_pixels %>% 
-  group_by(site) %>% 
-  group_modify(~match_func(.x))
-
-# You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
-
-# Clean up some RAM space
-rm(OISST, CMC); gc()
-
-# All the code below can be viewed in upwell_IDX.Rmd
-# Next to run upwelling index
-# UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
-load("Data/UI_angle.RData")
-
-upwelling <- UI_angle %>% 
-  dplyr::rename(temp = ui.saws) %>%
-  group_by(site) %>%
-  # mutate(min_t = min(t), 
-  #        max_t = max(t)) %>% 
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+#   unique() %>% 
+#   mutate(product = "OISST")
+# 
+# CMC_prod <- CMC %>% 
+#   select(lon, lat) %>%
+#   unique() %>% 
+#   mutate(product = "CMC")
+# 
+# # sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+#   #rbind(., CMC) %>% 
+#   # select(product, lon, lat)
+# 
+# # sat_pixels <- sat_data %>% 
+#   # select(product, lon, lat) %>%
+#   # unique()
+# 
+# ## For testing the nest/map pipeline
+# # df <- site_pixels %>%
+# #   filter(site == "Lamberts Bay") %>%
+# #   select(-site)
+# 
+# match_func <- function(df){
+#   df <- df %>%
+#     dplyr::rename(lon_site = lon, lat_site = lat)
+#   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+#                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
+#     cbind(., df)
+#   CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
+#                                              as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
+#     cbind(., df)
+#   res <- rbind(OISST_index, CMC_index)
+#   return(res)
+# }
+# 
+# # Find the nearest pixels to each spot along the transect
+# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
+# # points to not fall inside of the same 25 KM pixel
+# pixel_match <- site_pixels %>% 
+#   group_by(site) %>% 
+#   group_modify(~match_func(.x))
+# 
+# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+# OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+# CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
+# 
+# # Clean up some RAM space
+# rm(OISST, CMC); gc()
+# 
+# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
+# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
+# # # Hout Bay will be ignored
+# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+# 
+# OISST_fill <- OISST_fill %>%
+#   filter(site %in% selected_sites)
+# 
+# CMC_fill <- CMC_fill %>%
+#   filter(site %in% selected_sites)
+# 
+# 
+# # All the code below can be viewed in upwell_IDX.Rmd
+# # Next to run upwelling index
+# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+# load("Data/UI_angle.RData")
+# 
+# upwelling <- UI_angle %>% 
+#   dplyr::rename(temp = ui.saws) %>%
+#   group_by(site) %>%
+#   # mutate(min_t = min(t), 
+#   #        max_t = max(t)) %>% 
+#   nest() %>% # apply the following functions to all of the variables in te dataset
+#   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+#          # NB: A threshold of 3 appeared to be far to strict
+#          # purr::map - apllies a function to each element of a vector
+#          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
 Upwelling can last for a few hours.
 Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>% 
-  select(-data, -clim) %>% 
-  unnest() %>%
-  filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance)
+#   # Detect consecutive days in exceedance of a given threshold.
+#   # mutate() %>% 
+#   select(-data, -clim) %>% 
+#   unnest() %>%
+#   filter(row_number() %% 2 == 1) %>%
+#   unnest() %>% # creates a column for each variables
+#   dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+#   select(site, t, ui.saws, exceedance) 
+# 
+# # Now applying the Upwelling func
+# # NB: This only pulls out the event results and not the climatology results
+# # This is done to keep the output tidy because group_modify() may only create data.frame type outputs
+# # To create a list output one would use group_map(), 
+# # but this then loses the labels of which sites etc. the results belong to
+# detect_event_custom <- function(df){
+#   res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
+#   return(res)
+# }
+# 
+# ts2clm_custom <- function(df){
+#   # The climatology base period used here is up for debate...
+#   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
+#   res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+#   return(res)
+# }
+# 
+# # Calculate the upwelling event metrics
+# upwelling_detect_event <- function(df){
+#   upwell_base <- df %>% 
+#     dplyr::rename(t = date) %>% 
+#     group_by(site, product, heading, distance, lon, lat) %>% 
+#     group_modify(~ts2clm_custom(.x)) %>% 
+#     left_join(upwelling, by = c("site", "t")) %>%
+#     filter(!is.na(exceedance)) %>%
+#     group_by(site, product, heading, distance, lon, lat) %>% 
+#     group_modify(~detect_event_custom(.x))
+#   }
+# 
+# OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
+# save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+# CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
+# save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
+# 
+# # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
+# SACTN_US <- SACTN_US %>%
+#   filter(site %in% selected_sites)
+# 
+# load("Data/site_list_v4.2.RData")
+# 
+# SACTN_upwell_base <- SACTN_US %>%
+#   left_join(site_list[,c(4, 5, 6)], by = "index")%>% 
+#     dplyr::rename(t = date) %>% 
+#     group_by(site, lon, lat) %>% 
+#     group_modify(~ts2clm_custom(.x)) %>% 
+#     left_join(upwelling, by = c("site", "t")) %>%
+#     filter(!is.na(exceedance)) %>%
+#     group_by(site, lon, lat) %>% 
+#     group_modify(~detect_event_custom(.x))
+# 
+# save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+```
 
-# Now applying the Upwelling func
-# NB: This only pulls out the event results and not the climatology results
-# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
-# To create a list output one would use group_map(), 
-# but this then loses the labels of which sites etc. the results belong to
-detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
-  return(res)
-}
+# PLotting events
 
-ts2clm_custom <- function(df){
-  # The climatology base period used here is up for debate...
-  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
-  return(res)
-}
+Loading the data
+```{r}
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+```
 
-# Calculate the upwelling event metrics
-upwelling_detect_event <- function(df){
-  upwell_base <- df %>% 
-    dplyr::rename(t = date) %>% 
-    group_by(site, product, heading, distance, lon, lat) %>% 
-    group_modify(~ts2clm_custom(.x)) %>% 
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, product, heading, distance, lon, lat) %>% 
-    group_modify(~detect_event_custom(.x))
-  }
+To accurately compare 
+Start date CMC start date 1991-09-02 end date 2017-03-15
+Start date OISST start date 1981-09- 01  and end date 2018-09-28
+Start date SACTN start date 1973 and end date 2017-12-16
 
-OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-```
+As a common date we will use a start date from 1992-01-31 and end date 17-01-31

### 2019-09-22 SST_patterns.Rmd
update

@@
 
-As a common date we will use a start date from 1992-01-31 and end date 17-01-31
+As a common date we will use a start date from 1992-01-31 and end date 16-12-31
+
+Using the generalised linear hypotheses - Look at the difference in count of upwelling within the different datasets (in situ, CMC, OISST).
+Look at the difference in duration of upwelling between the thre datasets.
+
+To determine wether or not the same upwlling patterns are present at different distances from the coastline
+
+Offshore Upwelling:
+Which pixels showed the highest upwelling counts
+Which site showed the most intense upwelling

### 2019-09-23 Write_up/Paper1_writeup.Rmd
Comments throughout the text thus far. My comments begin with RWS, for ease of searching.

@@
 
-### Introduction
+## Introduction
 
@@
 These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
-Globally, upwelling occurs throughout the world’s oceans, however, four major eastern boundary upwelling systems (EBUS) are recognised, each significantly impacting coastal systems.
+Globally, upwelling occurs throughout the world’s oceans, however, four major eastern boundary upwelling systems (EBUS) are recognised, each significantly impacting coastal systems (RWS: This is a run-on sentence).
 These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly].
-The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews].
+The inshore waters off the west coast of South Africa are dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews] (RWS: This could be made more clear.
+What is the main point you want to communicate in this sentence?).
 Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
 As such, there has been a constant interest in upwelling systems.
-Major upwelling programs have already been carried through off of Peru (e.g.
+Major upwelling programs (RWS: This is too vague.
+What are these programs?) have already been carried out in Peru (e.g.
 WALSH.
 KELLEY, DUGDALE and FROST, 1971) and North-West Africa (e.g.
-BARBER, 1977).
+BARBER, 1977) (RWS: Thes references need to be made in a markdown format (e.g.
+[@walsh])).
 
-Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
-The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
+Over the past three decades, anthropogenic climate change has established global concern as it has been shown to have negatively affected marine and terrestrial ecosystems (RWS: citation needed).
+The consequences of these changes are also shown to have had direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
+According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients [@bakun].
 Pressure gradients are known to affect wind patterns which will subsequently affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
 
-Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references).
+Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this).
 These temperatures have increased as a result of anthropogenically driven climate change [@gruber].
-Warming is shown to slowly move down into the water column thus increasing thermal stratification.
+Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed).
 This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
@@
 1999; Walther et al.
-2002).
+2002 (RWS: Make these markdown citation style)).
 This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
-Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters.
+Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters (RWS: Citations).
 Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993) therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
+Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993 (RWS: Markdown citation)) therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
+(RWS: There appear to be two or three separate main points being made in this paragraph.
+I would recomend you split it up into a couple of paragraphs and think about what the main message of each should be.)
 
-The 3000km South African coastline, is bordered by the Agulhas current along the east coast and the EBUS Benguela current, along the west coast.
-These coastal regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
+The 3000km South African coastline is bordered by the Agulhas current along the east coast and the Benguela current (an EBUS) along the west coast.
+These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents (Bolton et al.
 2004; Mead et al.
-2013; Schlegel and Smit 2016).
+2013; Schlegel and Smit 2016 (RWS: Markdown citations.
+Fix this throughout.
+I will stop making this comment here.)).
 The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
@@
 The upwelling period in the Benguela north of 25°C persists throughout the entire year but peaks during August.
-The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-Several studies have shown that extreme warming and cooling periods occurring in the northern and southern Benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
+The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region in Namibia [@shannon; @cole].
+Several studies have shown that extreme warming and cooling periods occur in the northern and southern Benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
 Furthermore, Walter (1937) found that warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
@@
 In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occurring along the south coastline of the Southern Africa region are usually generated by local winds.
-Event-scale upwelling has been demonstrated to have substantial impacts on fisheries [@roy; @van].
+Upwelling occurring along the southern coastline of southern Africa are usually generated by local winds.
+Event-scale upwelling has been demonstrated to have substantial impacts on fisheries [@roy; @van] (RWS: Why is this last sentence here?
+It has nothing to do with the rest of the paragraph.).
 
 The Cape peninsula is influenced by south-east trade winds during summer and westerlies during winter, thus a pronounced seasonal cycle is evident within patterns of wind movement across the coastline.
-During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent.
-During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months.
+During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent (RWS: Citation).
+During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months (RWS: Citation).
 During the winter seasons, low pressure cells are driven from the west to the east as a result of the atmospheric circumpolar waves and so north-westerly winds dominate during winter.
 South easterly winds dominate during summer but are still present for short durations during winter [@andrews].
-Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region.
+Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region (RWS: Citation).
 
@@
 There has not yet been any understanding of how the upwelling process varies at local versus broader scales.
-In this study, it was hypothesize that upwelling is more dominant within coastal regions instead of broader scale.
-Very few studies have addressed the changes in the intensity or duration of cold temperature spells or cold events at varying distances from the coast within the Benguela region.
+In this study, it was hypothesized that upwelling is more dominant within coastal regions instead of broader scale (RWS: This sentence needs to be re-written for clarity).
+Very few studies have addressed the changes in the intensity or duration of cold-spells or cold events at varying distances from the coast within the Benguela region (RWS: this should be said before you announce what your hypothesis is).
 
 Strong temperature gradients existing between offshore and inshore oceanic waters may result in different upwelling event readings.
-Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tends to vary based on seasonal wind variation [@andrews].
-Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters[@andrews,@Shannon].
+Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tend to vary based on seasonal wind variation [@andrews].
+Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters [@andrews; @Shannon].
 Due to the temperature gradients that exist inshore and offshore, this study focused on observing upwelling patterns from the coastal region to 50km from the coastline.
@@
 Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
-The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
+The primary aim of the first study was to make use of various remotely sensed SST and *in-situ* collected coastal seawater temperature data.
 The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit].
@@
 This investigation examined recent trends and variation of upwelling events at different distances from the coastline.
+(RWS: This is a very thorough introduction section but has potentially too much information.
+I would recomend that you write out a bullet list for this section with the main and sub-points that you want to make with each paragraph and then see that they all flow from one paragraph to the next.
+If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)
 
@@
 
-## Site selection
+### Site selection
 
-On a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
+The South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions [@smit, @mead].
 The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings], which provides a natural laboratory for this study.
-Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the Western limit near the Namibian border.
+Annual mean coastal seawater temperatures within this region have a range of 12.3 ± 1.2°C at the Western limit near the Namibian border.
 Seasonal upwelling is controlled by south-easterly trade winds and as a result reflects distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-This region represents monthly mean temperatures at generally intermediate at cold and warm temperatures [@luning].
-In order to accurately examine upwelling patterns at various distances from the coast, several sites were selected providing a more accurate representation.
-The appropriate sites along the west coast of South Africa (i.e.
+This region represents monthly mean temperatures at generally intermediate at cold and warm temperatures [@luning] (RWS: This sentence is unclear, it would also be good to include a temperature value here, like the previous sentence).
+In order to accurately examine upwelling patterns at various distances from the coast, several sites from the SACTN dataset along the west coast of South Africa (i.e.
 sites influenced by the Benguela current) were selected.
 Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
-This yielded a total of five sites within the Benguela region (Fig 1).
+This yielded a total of five sites within the Benguela region (Fig 1 (RWS: The text is too small on this figure.
+I recommend saving it at a smaller size so that the text appears larger.
+I would also move the legend over SA rather than having it on the side of the figure.)).
+(RWS: This sub-section should come after the temperature and wind data sections)
 
@@
 
-## Temperature data sources
+### Temperature data
 
@@
 This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
-The SACTN has collected SST data along the coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and governmental departments.
-As a result, temperatures within these datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
+The SACTN has collected SST data along the coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and governmental departments (RWS: The SACTN itself does not collect data.
+Be careful with your word choice).
+The temperature data within this dataset were obtained with different methods and instruments, such as hand held thermometers or underwater temperature reorders (UTRs).
 The precision of the instruments used vary accordingly, with hand-held thermometers recording temperatures at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
-The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.
+The duration and extent of recordings per site were uneven with the longest time series in the dataset being that of Gordons Bay, which started on 13 September 1972 and concluded on 26 January 2017.
 During the 1970s, a total of 11 time series began recording.
@@
 Remotely sensed SST products are produced by a number of organisations across the globe.
-The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
+(RWS: Consider combining or re-working these first three sentences) The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
 The AVHRR datasets have been providing global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
-OISST is a global 1/4° gridded, daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, was produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and are currently among the highest resolution SST datasets available.
+OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product (Banzon et al., 2016).
+The second remotely sensed dataset used in this study was the Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis.
+This dataset is produced using satellite instruments from 1 June 2002 to present (PO.DAAC, https://podaac.jpl.nasa.gov/).
+MUR provides SST data at a global grid of 0.01° in longitude latitude and are currently among the highest resolution SST datasets available.
 The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
+The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program (RWS: citation needed).
 
@@
 Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm bias reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
+(RWS: Knowing this, why do you say near the end of your intro section that you are okay with assuming that in situ and satellite data are similar along the coast?)
 
 In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa.
-Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
-These distances were 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig.
+Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast  (Fig.
 1).
-The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated in order to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
+The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce offshore, broad-scale (RWS: What does broad-scale mean?) time series that could accurately be used as a dataset to be compared against the *in situ* time series.
 
-## Wind data
+### Wind data
 
-Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
-Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
-Specific wind characteristics were measured, namely; wind direction (dirw) and wind speed (spw).
+The wind was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, the wind was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind data were obtained from the South African Weather Service (SAWS), and were provided at a three-hour temporal resolution.
+The wind variables used here were wind direction (dirw) and wind speed (spw).
 Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
-The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters.
-ERA5 hourly data on pressure levels was used as a second wind source.
-This dataset comprised of u and v variables from 1973 to present.
+The `circular()` function in R software (RWS: Cite R) was used to create circular objects around the wave data in order to calculate the daily wind parameters.
+ERA5 hourly data on pressure levels were used as a second wind source (RWS: Need ERA5 citation and need a sentence explaining what ERA 5 is).
+This dataset comprised of u and v variables from 1973 to present (RWS: Need to be more specific about what the u and v variables are).
 
@@
 To do this, it was important to identify upwelling.
+(RWS: This needs to be re-worked.)
 
-# Defining and determining upwelling
+### Defining and determining upwelling
 
 In order to determine the change in upwelling signals at different distances from the coastline and to test which dataset showed clear, visible representation of these events, it was first necessary to define when upwelling was occurring.
-The next aim was to determine a threshold value to identify when upwelling was occurring within this region.
-Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider both temperature and wind values.
-To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989).
-The upwelling index provided an indication of the intensity of the upwelling process taking place using the following equation:
+To do this it was necessary to determine a threshold value for identifying when upwelling was occurring.
+Given that upwelling is primarily caused by alongshore, equatorward winds, it was necessary to consider both temperature and wind values.
+To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989):
 
@@
 
-In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
+In this equation, μ represents the wind speed (m/s) and θ represents the wind direction, which is measured in degrees.
 The coastal angle is the angle of each of the sites perpendicular to the coastline.
 The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur.
-Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value was ment to be.
-Here, it was specified that the minimum duration for upwelling was 1 day with a threshold value obtained from the upwelling index as 1.
-In order to obtain a more approved threshold value, a percentile value was determined.
-To determine changes within the upwelling values, the 'detect event' function within the heatwaveR package was used.
-Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
+Thereafter, the `exceedance()` function within the __`heatwaveR`__ package (RWS: Cite heatwaveR) was used to determine the consecutive number of days the upwelling index was at or above 1.
+In order to obtain a more approved threshold value, a percentile value was determined (RWS: Was it though?
+This sentence wouldn't make sense to someone that hadn't also seen the R code).
+To determine changes within the upwelling values, the `detect_event()` function within the __`heatwaveR`__ package was used.
+Thereafter, to determine the upwelling event metrics (frequency and intensity) the detect event function was used as a first filter with a threshold at or below the 25th percentile.
 Values obtained from the upwelling index were used as a second threshold.
-An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
+(RWS: The description of the statistical methodology here is unclear and needs to be re-written) An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
 Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months.
-For the use of the heatwaveR package it was best to use a time series of 30years.
+For the use of the heatwaveR package it was best to use a time series of 30 years (RWS: No.
+The 30 years comes from WMO recommendations and is not a requirement for the heatwaveR package specifically).

### 2019-09-23 Write_up/Paper1_writeup.Rmd
merge

--- 2019-09-18 Write_up/Paper1_writeup.Rmd
@@
 The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters.
-ERA5 hourly data on pressure levels was used as a second wind source.
-This dataset comprised of u and v variables from 1973 to present.
 
-With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested.
-To do this, it was important to identify upwelling.
+With temperature and wind values now corresponding to their respective sites and dates, the hypothesis examining upwelling patterns at different distances from the coastline could now be tested.
+To do this, it was first important to identify upwelling.
 
@@
 To determine changes within the upwelling values, the 'detect event' function within the heatwaveR package was used.
-Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
+Thereafter, to determine the upwelling event metrics (frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
 Values obtained from the upwelling index were used as a second threshold.
@@
 Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months.
-For the use of the heatwaveR package it was best to use a time series of 30years.
+For the application of the heatwaveR package it was best to make use a time series of 30 years.

### 2019-09-23 Write_up/Paper1_writeup.Rmd
updates

--- 2019-09-23 Write_up/Paper1_writeup.Rmd
@@
 These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
-Globally, upwelling occurs throughout the world’s oceans, however, four major eastern boundary upwelling systems (EBUS) are recognised, each significantly impacting coastal systems (RWS: This is a run-on sentence).
+Occurring globally throughout the world's oceans, four major eastern boundary upwelling systems (EBUS) can be recognised; each significantly impacting associated coastal systems.
 These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly].
@@
 Major upwelling programs (RWS: This is too vague.
-What are these programs?) have already been carried out in Peru (e.g.
-WALSH.
-KELLEY, DUGDALE and FROST, 1971) and North-West Africa (e.g.
-BARBER, 1977) (RWS: Thes references need to be made in a markdown format (e.g.
-[@walsh])).
+What are these programs?) have already been carried out in Peru [@walsh] and North-West Africa [@barber].
 
-Over the past three decades, anthropogenic climate change has established global concern as it has been shown to have negatively affected marine and terrestrial ecosystems (RWS: citation needed).
+Over the past three decades, anthropogenic climate change has established global concern as it has been shown to have negatively affected marine and terrestrial ecosystems [@hoegh].
 The consequences of these changes are also shown to have had direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
@@
 This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
-Climate change impacting oceanic processes has major ecological consequences (Bradley et al.
-1999; Walther et al.
-2002 (RWS: Make these markdown citation style)).
+Climate change impacting oceanic processes has major ecological consequences [@walther].
 This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
-Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters (RWS: Citations).
+Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters [@hoegh].
 Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993 (RWS: Markdown citation)) therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
+Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
 (RWS: There appear to be two or three separate main points being made in this paragraph.
@@
 The 3000km South African coastline is bordered by the Agulhas current along the east coast and the Benguela current (an EBUS) along the west coast.
-These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents (Bolton et al.
-2004; Mead et al.
-2013; Schlegel and Smit 2016 (RWS: Markdown citations.
-Fix this throughout.
-I will stop making this comment here.)).
-The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
-2009).
+These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@Mead; @schlegel].
+The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region [@jarred; @hutchings].
 The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
-Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al.
-2009; Schlegel et al.
-2017).
+Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast [@luning1990; @lutjeharms2000; @hutchings; @schlegel].
 Changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
@@
 Upwelling occurring along the southern coastline of southern Africa are usually generated by local winds.
-Event-scale upwelling has been demonstrated to have substantial impacts on fisheries [@roy; @van] (RWS: Why is this last sentence here?
-It has nothing to do with the rest of the paragraph.).
 
@@
 
-The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood.
+The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood, with Very few studies addressing the changes in the intensity of cold events at varying distances from the coast within the Benguela region.
+Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
 Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore [@andrews].
-Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
 There has not yet been any understanding of how the upwelling process varies at local versus broader scales.
-In this study, it was hypothesized that upwelling is more dominant within coastal regions instead of broader scale (RWS: This sentence needs to be re-written for clarity).
-Very few studies have addressed the changes in the intensity or duration of cold-spells or cold events at varying distances from the coast within the Benguela region (RWS: this should be said before you announce what your hypothesis is).
+Within this study it is hypothesized that the upwelling metric may vary spatially across different distances, with more intense upwelling events occurring nearest to the coast.
 
@@
 This investigation examined recent trends and variation of upwelling events at different distances from the coastline.
-(RWS: This is a very thorough introduction section but has potentially too much information.
-I would recomend that you write out a bullet list for this section with the main and sub-points that you want to make with each paragraph and then see that they all flow from one paragraph to the next.
-If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)
 
 ## Methods
+
+### Temperature data
+
+The South African Coastal Temperature Network (SACTN) data served as the *in situ* temperature dataset used in this study.
+This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
+This dataset consisted of coastal seawater temperatures for 129 sites along the coast of South Africa, measured daily from 1972 until 2017, with contributions made from several organisations and governmental departments.
+Of these, 80 were measured using hand-held thermometers and the remaining 45 were measured using UTRs.
+The duration and extent of the recordings per site were uneven, with the longest time series in the dataset being that of Gordons Bay.
+Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
+During the 1970s, a total of 11 time series began recording.
+A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
+Recordings are still ongoing at many of these sites.
+For this analysis, the data were combined and formatted into standardized comma delineated values (CSV) files which allowed for a fixed methodology to be used across the entire dataset.
+Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers.
+These data points were then changed to NA (not available) so as to not interfere with analysis.
+Recordings are still ongoing at many of these sites.
+
+Additionally this study has sought to incorporate remotely sensed temperature datasets; complied by a number of organisations across the globe, each specialising in remotely sensed SST data and information products; utilising satellites to provide a system of large spatial and temporal data retrieval.
+The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
+The AVHRR datasets have been providing global SSTs for more than four decades [@reynolds].
+OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product [@banzon].
+The CMC dataset constitutes the second dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
+The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program (PO.DAAC, https://podaac.jpl.nasa.gov/).
+
+An advantage of using of *in situ* data over satellite data is that they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and can therefore more accurately explain upwelling within the coastal inshore environment.
+Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm bias reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
+
+In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa.
+Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast  (Fig.
+1).
+The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce an offshore time series that could accurately be used as a dataset to be compared against the *in situ* time series.
+
+### Wind data
+
+The wind was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, the wind was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind data were obtained from the South African Weather Service (SAWS), and were provided at a three-hour temporal resolution.
+The wind variables used here were wind direction (dirw) and wind speed (spw).
+Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
+The `circular()` function in R software [@team] was used to create circular objects around the wave data in order to calculate the daily wind parameters.
+ERA5 hourly data on pressure levels were used as a second wind source To visualize a synoptic view of the wind state around the Benguela region during coastal upwelling, we chose to use ERA-5 to provide wind vectors (10m above surface).
+ERA-5 is a fifth generation atmospheric reanalysis of global climate product produced by the European Center for Medium-Range Weather Forecasts (ECMWF, http://www.ecmwf.int/) [@molten].
+ERA-5 is a comprehensive global model that assimilates a wide range of data to create a short term forecast.
+At the time of this writing the chosen variables were available for download from January 1st, 1979 to December 31st, 2016.
+The data used in this study were downloaded at a daily resolution on a 0.25° grid and within the latitude/longitude of the study region (Fig 1).
+With temperature and wind values now corresponding, we tested if upwelling pattersns varied at different distances from the coastline.
 
@@
 Seasonal upwelling is controlled by south-easterly trade winds and as a result reflects distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-This region represents monthly mean temperatures at generally intermediate at cold and warm temperatures [@luning] (RWS: This sentence is unclear, it would also be good to include a temperature value here, like the previous sentence).
 In order to accurately examine upwelling patterns at various distances from the coast, several sites from the SACTN dataset along the west coast of South Africa (i.e.
@@
 Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
-This yielded a total of five sites within the Benguela region (Fig 1 (RWS: The text is too small on this figure.
-I recommend saving it at a smaller size so that the text appears larger.
-I would also move the legend over SA rather than having it on the side of the figure.)).
-(RWS: This sub-section should come after the temperature and wind data sections)
+This yielded a total of four sites within the Benguela region (Fig 1).
 
@@
 ```
-
-### Temperature data
-
-The South African Coastal Temperature Network (SACTN) data served as the *in situ* temperature dataset used in this study.
-This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
-The SACTN has collected SST data along the coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and governmental departments (RWS: The SACTN itself does not collect data.
-Be careful with your word choice).
-The temperature data within this dataset were obtained with different methods and instruments, such as hand held thermometers or underwater temperature reorders (UTRs).
-The precision of the instruments used vary accordingly, with hand-held thermometers recording temperatures at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
-The duration and extent of recordings per site were uneven with the longest time series in the dataset being that of Gordons Bay, which started on 13 September 1972 and concluded on 26 January 2017.
-During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 during the 1990s, and 18 entries were added during the 2000s.
-Recordings are still ongoing at many of these sites.
-
-This study also made use of a range of remotely sensed temperature datasets.
-Satellites provide a system of large spatial and temporal data retrieval.
-Remotely sensed SST products are produced by a number of organisations across the globe.
-(RWS: Consider combining or re-working these first three sentences) The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been providing global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
-OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product (Banzon et al., 2016).
-The second remotely sensed dataset used in this study was the Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis.
-This dataset is produced using satellite instruments from 1 June 2002 to present (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a global grid of 0.01° in longitude latitude and are currently among the highest resolution SST datasets available.
-The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program (RWS: citation needed).
-
-An advantage of using of *in situ* data over satellite data is that they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and can therefore more accurately explain upwelling within the coastal inshore environment.
-Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm bias reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
-(RWS: Knowing this, why do you say near the end of your intro section that you are okay with assuming that in situ and satellite data are similar along the coast?)
-
-In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa.
-Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast  (Fig.
-1).
-The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce offshore, broad-scale (RWS: What does broad-scale mean?) time series that could accurately be used as a dataset to be compared against the *in situ* time series.
-
-### Wind data
-
-The wind was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, the wind was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
-Wind data were obtained from the South African Weather Service (SAWS), and were provided at a three-hour temporal resolution.
-The wind variables used here were wind direction (dirw) and wind speed (spw).
-Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
-The `circular()` function in R software (RWS: Cite R) was used to create circular objects around the wave data in order to calculate the daily wind parameters.
-ERA5 hourly data on pressure levels were used as a second wind source (RWS: Need ERA5 citation and need a sentence explaining what ERA 5 is).
-This dataset comprised of u and v variables from 1973 to present (RWS: Need to be more specific about what the u and v variables are).
-
-With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested.
-To do this, it was important to identify upwelling.
-(RWS: This needs to be re-worked.)
 

### 2019-09-23 SST_patterns.Rmd
progress

--- 2019-09-22 SST_patterns.Rmd
@@
 
-As a common date we will use a start date from 1992-01-31 and end date 17-01-31
+As a common date we will use a start date from 1992-01-31 and end date 16-12-31
+
+```{r}
+# Generalosed linear models
+#Using the generalised linear hypotheses - Look at the difference in metrics (duration and intensity) of in situ upwelling and remotely sensed SST data upwelling
+
+lm_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
+}
+# Changes in upwelling metrics
+lm_func <- function(df){
+  upwell_lm <- df %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var, distance) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+}
+
+OISST_lm <- lm_func(df = OISST_upwell_base)
+CMC_lm <- lm_func(df = CMC_upwell_base)
+
+SACTN_lm <- SACTN_upwell_base %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+```
+
+# From the years 1992 - 2017
+
+```{r}
+# Which pixels showed the highest upwelling counts within the different datasets
+# Which site showed the most intense upwelling within the different datasets
+# Where is the largest upwelling event occuring?
+
+metric_func <- function(df){
+ metrics <-df %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  #mutate(year = year(date_start)) %>% 
+  group_by(distance, site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+   select(site, distance, mean_intensity, sum_events)
+}
+
+OISST_metrics <- metric_func(OISST_upwell_base)
+CMC_metrics <- metric_func(CMC_upwell_base)
+
+# No of u pwelling events found at the different sites
+SACTN_metrics <- SACTN_upwell_base %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  group_by(site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+  select(site, mean_intensity, sum_events)
+
+# Create a table showing these metrics
+
+```
+
+Create visualisation which compares the presence and absence of events
+If a event is found in a particular site at a particulate dat, is the same event present in the other datasets? Show how the events detected varies between site and dataset.
+
+```{r}
+# To determine wether or not the same upwlling patterns are present at different distances from the coastline
+
+```
+
+- Pvalues
+- plotting
+- ERA5
+- Plot wind data

### 2019-09-23 Write_up/Paper1_writeup.Rmd
progress

@@
 These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly].
-The inshore waters off the west coast of South Africa are dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews] (RWS: This could be made more clear.
-What is the main point you want to communicate in this sentence?).
-Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
-As such, there has been a constant interest in upwelling systems.
-Major upwelling programs (RWS: This is too vague.
-What are these programs?) have already been carried out in Peru [@walsh] and North-West Africa [@barber].
+There has been significant interest in major upwelling systems since the later 1960s, with major interest occuring off Peru [@walsh] and North-West Africa [@barber].
+Research done by Copenhagen 1953, reported a number of sites along the west coast of southern Africa where upwelling was particularly intense.
+The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrew]
 
@@
 (RWS: There appear to be two or three separate main points being made in this paragraph.
-I would recomend you split it up into a couple of paragraphs and think about what the main message of each should be.)
+I would recomend you use what you have here and try and work these main points into other paragraphs)
 
@@
 This investigation examined recent trends and variation of upwelling events at different distances from the coastline.
+
+(RWS: This is a very thorough introduction section.
+Impotantly see that they all flow from one paragraph to the next.
+If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)
 

### 2019-09-23 site_map.Rmd
progress

@@
 
-```{r}
+```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
+The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=6, fig.width=10, fig.pos="H"}
+
 load("Data/south_africa_coast.RData")
@@
         panel.grid.minor = element_blank(),
-        legend.text = element_text(size = 12),
-        legend.title = element_text(size = 8),
+        legend.text = element_text(size = 14),
+        legend.title = element_text(size = 10),
         legend.key = element_blank(),
@@
         legend.justification = c(1,0),
-        legend.position = c(0.65, 0.80),
+        legend.position = c(0.85, 0.70),
         legend.text = element_text(size = 8),
@@
 
-pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
 vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
@@
 dev.off()
+pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
 

### 2019-09-24 SST_patterns.Rmd
update

@@
 
-As a common date we will use a start date from 1992-01-31 and end date 16-12-31
-
 ```{r}
-# Generalosed linear models
-#Using the generalised linear hypotheses - Look at the difference in metrics (duration and intensity) of in situ upwelling and remotely sensed SST data upwelling
 
@@
 
+As a common date we will use a start date from 1992-01-31 and end date 16-12-31
+
+```{r}
+# Generalosed linear models
+
+# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
+# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)
+
+trends_duration_func <- function(df){
+  upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
+                     family = poisson(link = "log"))
+  upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
+  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
+                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+}
+
+OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
+CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
+SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
+
+### intensoty 
+trends_intensity_func <- function(df){
+  upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
+                     family = gaussian(link = "identity"))
+  upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
+  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
+                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+}
+
+OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
+CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
+SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
+```
+
 # From the years 1992 - 2017
@@
 # Which site showed the most intense upwelling within the different datasets
-# Where is the largest upwelling event occuring?
 
@@
 
+# Wind plots
+
+```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 20}
+library(ggradar)
+library(dplyr)
+library(scales)
+library(tibble)
+load("Data/UI_angle.RData")
+
+# Detects NA values, no NA values are present
+# Plotting wind variables
+source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function
+
+wind_daily_renamed <- UI_angle %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ)
+
+p.wr2 <- plot.windrose(data = wind_daily_renamed,
+              spd = "spd",
+              dir = "dir")
+
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 4, nrow = 5) +
+  theme(strip.text.x = element_text(size = 25))
+p.wr3
+
+################3
+
+load("Data/UI_angle.RData")
+
+wind_plot_prep <- UI_angle %>% 
+  select(-t, -ui.saws)
+
+wind_plot <- wind_plot_prep %>% 
+  as_tibble(rownames = "site") %>% 
+  group_by(site) %>% 
+  mutate_at(vars(-site), rescale) %>% 
+  tail(4)
+
+ggradar(wind_plot)
+
+??countmax
+
+```
+
+# ANOVA
+- Relationship between duration, year and distance as a function of the mean intensity.
+
+```{r}
+anova_func <- function(df){
+  sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration * distance, data = df)
+return(sites_aov)
+}
+
+OISST_anov <- anova_func(df = OISST_upwell_base)
+OISST_anov
+CMC_anov <- anova_func(df = CMC_upwell_base)
+CMC_anov
+
+SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration, data = SACTN_upwell_base)
+SACTN_anov
+```
+
 Create visualisation which compares the presence and absence of events
@@
 ```
-
-- Pvalues
-- plotting
-- ERA5
-- Plot wind data

### 2019-09-24 Write_up/Paper1_writeup.Rmd
update

@@
 In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa.
-Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast  (Fig.
-1).
+Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast shown as black boxes in Fig.
+1.
 The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce an offshore time series that could accurately be used as a dataset to be compared against the *in situ* time series.
+Offshore transects were used in order to generate a time series in this manner as they accurately represent the mesoscale temperatures we were interested in comparing against coastal upwelling.
+Had we only made use of the nearest remotely sensed pixels to compare against the SACTN time series this would simply only draw a comparison between datasets, and not the mesoscape acitivity within the ocean.
 
@@
 
-In order to determine the change in upwelling signals at different distances from the coastline and to test which dataset showed clear, visible representation of these events, it was first necessary to define when upwelling was occurring.
-To do this it was necessary to determine a threshold value for identifying when upwelling was occurring.
+In order to determine the change in upwelling signals at various different distances from the coastline, and to test which dataset showed a clear, and visible representation of these events, it was first necessary to define when upwelling was occurring; however to accomplish this it was first required to establish a threshold value for identifying when the phenomenon was in taking place.
 Given that upwelling is primarily caused by alongshore, equatorward winds, it was necessary to consider both temperature and wind values.
-To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989):
+These data values are instrumental in determining the specific environmental conditions and parameters required for upwelling to occur; to accomplish this an upwelling index was calculated using the formula presented in the work by Fielding and Davis (1989):
 
@@
 
-In this equation, μ represents the wind speed (m/s) and θ represents the wind direction, which is measured in degrees.
-The coastal angle is the angle of each of the sites perpendicular to the coastline.
-The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur.
-Thereafter, the `exceedance()` function within the __`heatwaveR`__ package (RWS: Cite heatwaveR) was used to determine the consecutive number of days the upwelling index was at or above 1.
-In order to obtain a more approved threshold value, a percentile value was determined (RWS: Was it though?
-This sentence wouldn't make sense to someone that hadn't also seen the R code).
-To determine changes within the upwelling values, the `detect_event()` function within the __`heatwaveR`__ package was used.
-Thereafter, to determine the upwelling event metrics (frequency and intensity) the detect event function was used as a first filter with a threshold at or below the 25th percentile.
-Values obtained from the upwelling index were used as a second threshold.
-(RWS: The description of the statistical methodology here is unclear and needs to be re-written) An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
+In this equation, μ represents the wind speed (m/s) and θ represents the wind direction, which is measured in degrees, and the coastal angle as the angle of each of the sites perpendicular to the coastline.
+The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling and the likelihood of the phenomenon occurring.
+The above equation produces a value called the upwelling index; for this study specifically, upwelling was classified to occur when the index is greater than 1.
+However to further ascertain a threshold for upwelling, the `exceedance()` function within the _`heatwaveR`_ package [@schlegel2017heat] was utilised, specifically to determine the consecutive number of days that the event must exceed for it to be regarded and classified as an upwelling event.
+However it should be noted that upwelling is known to vary on a seasonal basis and also may occur hourly.
+Therefore this study sought to establish the minimum duration for the analyses of upwelling as 1 day; the rationale behind this being that data from the SACTN dataset as well as the Satellite remotely sensed SST data is collected daily.
+Furthermore, it was also determined that a 25th percentile threshold for the upwelling index result lasting for >= 1 days matched with the upwelling index value of 1.
+As a result the 25th percentile threshold was used as a more acceptable value to predict upwelling events.
+In order to determine the upwelling metrics, the `detect_event()` function within the _`heatwaveR`_ package was used.An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
 Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months.
-For the use of the heatwaveR package it was best to use a time series of 30 years (RWS: No.
-The 30 years comes from WMO recommendations and is not a requirement for the heatwaveR package specifically).
+It must also be stated that due to irregular sampling efforts in the SACTN dataset the spatial and temporal distributions of upwelling is not necessarily even.

### 2019-09-24 Functions/wind.rose.R

@@ -7,11 +7,11 @@ plot.windrose <- function(data,
                           dir,
                           spdres = 2,
                           dirres = 30,
-                          spdmin = 2,
-                          spdmax = 22,
+                          spdmin = 0.00,
+                          spdmax = 12,
                           spdseq = NULL,
                           palette = "YlGnBu",
-                          countmax = NA,
+                          countmax = NA, #changed count max to 1000
                           debug = 0){
   
   

### 2019-09-25 SST_patterns.Rmd
Updated

@@
 
-### intensoty 
+### Intensity: Chnaged the family and link because of the negative values within the dataset 
 trends_intensity_func <- function(df){
@@
 OISST_anov <- anova_func(df = OISST_upwell_base)
-OISST_anov
+summary(OISST_anov)
 CMC_anov <- anova_func(df = CMC_upwell_base)
-CMC_anov
+summary(CMC_anov)
 
 SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration, data = SACTN_upwell_base)
-SACTN_anov
+summary(SACTN_anov)
 ```

### 2019-09-27 SST_patterns.Rmd
Corrected between different longitude schemes
Getting rid of NA
Update

@@
 
-```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 20}
+```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
 library(ggradar)
@@
 wind_daily_renamed <- UI_angle %>% 
+  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
   dplyr::rename(spd = mean_speed) %>%
-  dplyr::rename(dir = dir_circ)
+  dplyr::rename(dir = dir_circ) %>% 
+  filter(spd > 0)
 
@@
 
-p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 4, nrow = 5) +
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
   theme(strip.text.x = element_text(size = 25))
@@
 
-load("Data/UI_angle.RData")
-
-wind_plot_prep <- UI_angle %>% 
-  select(-t, -ui.saws)
-
-wind_plot <- wind_plot_prep %>% 
-  as_tibble(rownames = "site") %>% 
-  group_by(site) %>% 
-  mutate_at(vars(-site), rescale) %>% 
-  tail(4)
-
-ggradar(wind_plot)
-
-??countmax
+# load("Data/UI_angle.RData")
+# 
+# wind_plot_prep <- UI_angle %>% 
+#   select(-t, -ui.saws)
+# 
+# wind_plot <- wind_plot_prep %>% 
+#   as_tibble(rownames = "site") %>% 
+#   group_by(site) %>% 
+#   mutate_at(vars(-site), rescale) %>% 
+#   tail(4) 
+# 
+# ggradar(wind_plot)
+# 
+# ??countmax
 

### 2019-10-07 SST_patterns.Rmd
updated

@@
 p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
-  theme(strip.text.x = element_text(size = 25))
+  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
 p.wr3
@@
 anova_func <- function(df){
-  sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration * distance, data = df)
+  sites_aov <- aov(intensity_mean ~ site * duration * distance, data = df)
 return(sites_aov)
@@
 
-SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration, data = SACTN_upwell_base)
+SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
 summary(SACTN_anov)

### 2019-10-07 Write_up/Paper1_writeup.Rmd
updated

@@
 Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
-(RWS: There appear to be two or three separate main points being made in this paragraph.
-I would recomend you use what you have here and try and work these main points into other paragraphs)
 
@@
 This investigation examined recent trends and variation of upwelling events at different distances from the coastline.
-
-(RWS: This is a very thorough introduction section.
-Impotantly see that they all flow from one paragraph to the next.
-If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)
 
@@
 It must also be stated that due to irregular sampling efforts in the SACTN dataset the spatial and temporal distributions of upwelling is not necessarily even.
+
+## Statistical analyses
+
+A series of ANOVA tests were used to compare the main effects of the chosen variables on a continuous variable.
+In this analysis the relationship between sites based on the mean intenisity of upwelling as a function of distance and duration were analysed; these analyses tested if significant
+differences occurred between upwelling at different sites and distances from the coastline within each of the various datasets.
+
+## Results
+
+# Table
+
+# Wind rose diagrams
+
+```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagram representing the most predominant wind direction for each of the sites.
+Each spoke is divided by colour into wind speed ranges.
+The radial length of each spoke around the circle is the time that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/wind_roses.png")
+
+```
+
+Wind diagrams help visualise the patterns present at a particular site.
+Moving outward on the radial scale, the frequency associated with wind and waves coming from a particular direction increases.
+The predominant wind direction at Port Nolloth is 135o and at Sea Point it is 315o.
+Sea Point and Port Nollowth shows the fastest wind speed at a direction of 135o.
+The predominant wind direction for Lamberts Bay and Saldanha Bay occured at 195o, with the highest wind speed also found within this direction.
+
+# Average intensity between sites at the various distances
+
+Sites located within the OISST datasets shows no difference in duration of the events at different distances for each of the sites from the coastline.
+However, there is a difference in the mean intensity of events between sites (F=5.74, SS=5.2, p<0.0006).
+It is also evident that there is a significant difference in the mean intensity of the event at different distances from the coastline (F=260.17, SS=78.6, p<0.001), as well as differences in the duration in events (F=51.39, SS=15.53, p<0.001) and between sites .
+Similarly, the second dataset, the CMC dataset, results obtained from a  three-way ANOVA showed that there  was no a significant difference in the duration of events at different distances from the coastline for each of the sites (F=1.20, SS=0.41, p>0.05).
+However, there is a difference in the mean intensity of events at diferent distances (F=471.63, SS=162.0, p<0.05), as well as a signifcant difference in the duration of events (F=38.36, SS=13.2, p<0.006), and between sites (F=17.64, SS=18.2, p<0.002).
+It is also seen that there is no significant difference between the duration of upwelling between the various sites (F=0.22 SS=0.2, p<0.87).
+When looking at the in situ collected dataset, data collected at the coastline, we find no difference in the duration of events at the different sites (F=2.05, SS=1.84, p>0.11), however, we see a difference in the mean intensity of events at the different sites (F=2.05, SS=0.61, p>0.44).
+
+# Plotting images
+
+####TO do
+- Plots
+- Tables
+
+# Discussion

### 2019-10-07 Write_up/Paper1_writeup.Rmd
updated

@@
 
-# Paper 1
-## Chapter 2 Upwelling
-
+\newpage
 ## Abstract
 
+*Keywords*: Seawater temperature, climate change, coastal regions, code: R, variability, upwelling
+
+\newpage
+## Acknowledgments
+
+I would like to thank SAWS, DAFF, DEA, EKZNW and SAEON for the contribution of all the raw data used in this study.
+Without the collection of this data this article and the South African Coastal Temperature Network (SACTN) would not be possible.
+This research is funded by a South African National Research Foundation Grant (SFH180604339745).
+The data and analyses used in this paper may be found at [Add appendix link here]
+
+\newpage
 ## Introduction
@@
 Research done by Copenhagen 1953, reported a number of sites along the west coast of southern Africa where upwelling was particularly intense.
-The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrew]
+The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrew].
+Whilst upwelling plays a significant role in ecological processes, climate change largely influences this phenomenon.
 
@@
 Pressure gradients are known to affect wind patterns which will subsequently affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
+Over the last few decades, climate change has resulted in drastic pressure and temperature gradients.
 
@@
 South easterly winds dominate during summer but are still present for short durations during winter [@andrews].
-Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region (RWS: Citation).
+The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood, with Very few studies addressing the changes in the intensity of cold events at varying distances from the coast within the Benguela region.
 
-The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood, with Very few studies addressing the changes in the intensity of cold events at varying distances from the coast within the Benguela region.
 Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
@@
 Within this study it is hypothesized that the upwelling metric may vary spatially across different distances, with more intense upwelling events occurring nearest to the coast.
-
 Strong temperature gradients existing between offshore and inshore oceanic waters may result in different upwelling event readings.
@@
 
-# Table
+Visualisations
+
+<!-- \def\tiny{\@setfontsize\tiny{10pt}{11pt}} -->
+\begin{tiny}
+\begin{center}
+\setlength\tabcolsep{3pt}
+\begin{longtable}{|r|r|p{2cm}|}
+\caption{OISST} \\
+
+% This is the header for the first page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endfirsthead
+
+% This is the header for the second page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endhead
+
+% This is the footer for all pages except the last page of the table...
+\midrule
+\multicolumn{5}{l}{{Continued on Next Page\ldots}} \\
+\endfoot
+
+% This is the footer for the last page of the table...
+\bottomrule
+\endlastfoot
+
+% Now the data...
+
+Lamberts Bay & 10000 & 5.47875 \\
+Lamberts Bay & 20000 & 5.47875 \\
+Lamberts Bay & 30000 & 7.67025 \\
+Lamberts Bay & 40000 & 1.09575 \\
+Lamberts Bay & 50000 & -8.76600 \\
+Port Nolloth & 10000 & 0.36525  \\
+Port Nolloth & 20000 & -3.28725 \\
+Port Nolloth & 30000 & -2.92200 \\
+Port Nolloth & 40000 & -3.65250 \\
+Port Nolloth & 50000 & -3.28725 \\
+Saldanha Bay & 10000 & -0.73050 \\
+Saldanha Bay & 20000 & -0.36525 \\
+Saldanha Bay & 30000 & -4.38300 \\
+Saldanha Bay & 40000 & -10.22700 \\
+Saldanha Bay & 50000 & -12.41850 \\
+Sea Point & 10000 & -8.76600  \\
+Sea Point & 20000 & -3.65250 \\
+Sea Point & 30000 & -1.46100 \\
+Sea Point & 40000 & -9.49650 \\
+Sea Point & 50000 & -18.62775 \\
+
+\end{longtable}
+\end{center}
+\end{tiny}
+
+<!-- \def\tiny{\@setfontsize\tiny{10pt}{11pt}} -->
+\begin{tiny}
+\begin{center}
+\setlength\tabcolsep{3pt}
+\begin{longtable}{|r|r|p{2cm}|}
+\caption{CMC} \\
+
+% This is the header for the first page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endfirsthead
+
+% This is the header for the second page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endhead
+
+% This is the footer for all pages except the last page of the table...
+\midrule
+\multicolumn{5}{l}{{Continued on Next Page\ldots}} \\
+\endfoot
+
+% This is the footer for the last page of the table...
+\bottomrule
+\endlastfoot
+
+% Now the data...
+
+Lamberts Bay & 10000 & -26.29800 \\
+Lamberts Bay & 20000 & -25.56750 \\
+Lamberts Bay & 30000 & -23.74125 \\
+Lamberts Bay & 40000 & -31.41150 \\
+Lamberts Bay & 50000 & -31.77675 \\
+Port Nolloth & 10000 & -15.34050 \\
+Port Nolloth & 20000 & -16.43625 \\
+Port Nolloth & 30000 & -15.70575 \\
+Port Nolloth & 40000 & -20.45400 \\
+Port Nolloth & 50000 & -24.83700 \\
+Saldanha Bay & 10000 & -29.58525 \\
+Saldanha Bay & 20000 & -36.52500 \\
+Saldanha Bay & 30000 & -42.73425 \\
+Saldanha Bay & 40000 & -46.38675 \\
+Saldanha Bay & 50000 & -45.29100 \\
+Sea Point & 10000 & -28.48950  \\
+Sea Point & 20000 & -30.31575 \\
+Sea Point & 30000 & -36.89025 \\
+Sea Point & 40000 & -38.35125 \\
+Sea Point & 50000 & -37.25550 \\
+
+\end{longtable}
+\end{center}
+\end{tiny}
+
+<!-- \def\tiny{\@setfontsize\tiny{10pt}{11pt}} -->
+\begin{tiny}
+\begin{center}
+\setlength\tabcolsep{3pt}
+\begin{longtable}{|r|r|p{2cm}|}
+\caption{SACTN} \\
+
+% This is the header for the first page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endfirsthead
+
+% This is the header for the second page of the table...
+\toprule
+Site & distance & slope  \\
+\midrule
+\endhead
+
+% This is the footer for all pages except the last page of the table...
+\midrule
+\multicolumn{5}{l}{{Continued on Next Page\ldots}} \\
+\endfoot
+
+% This is the footer for the last page of the table...
+\bottomrule
+\endlastfoot
+
+% Now the data...
+
+Lamberts Bay & -1.46100 \\
+Port Nolloth & -3.65250 \\
+Saldanha Bay & -2.55675 \\
+Sea Point & -10.95750  \\
+
+\end{longtable}
+\end{center}
+\end{tiny}
 
@@
 
-Sites located within the OISST datasets shows no difference in duration of the events at different distances for each of the sites from the coastline.
-However, there is a difference in the mean intensity of events between sites (F=5.74, SS=5.2, p<0.0006).
-It is also evident that there is a significant difference in the mean intensity of the event at different distances from the coastline (F=260.17, SS=78.6, p<0.001), as well as differences in the duration in events (F=51.39, SS=15.53, p<0.001) and between sites .
-Similarly, the second dataset, the CMC dataset, results obtained from a  three-way ANOVA showed that there  was no a significant difference in the duration of events at different distances from the coastline for each of the sites (F=1.20, SS=0.41, p>0.05).
-However, there is a difference in the mean intensity of events at diferent distances (F=471.63, SS=162.0, p<0.05), as well as a signifcant difference in the duration of events (F=38.36, SS=13.2, p<0.006), and between sites (F=17.64, SS=18.2, p<0.002).
-It is also seen that there is no significant difference between the duration of upwelling between the various sites (F=0.22 SS=0.2, p<0.87).
-When looking at the in situ collected dataset, data collected at the coastline, we find no difference in the duration of events at the different sites (F=2.05, SS=1.84, p>0.11), however, we see a difference in the mean intensity of events at the different sites (F=2.05, SS=0.61, p>0.44).
-
-# Plotting images
-
-####TO do
-- Plots
-- Tables
+Sites compared within the OISST datasets showed no difference in the durations of the events at different distances for each of the sites from the coastline.
+However, there is a difference in the mean intensity of events between sites (*F*=5.74, *SS*=5.2, *p*<0.001).
+Additionally, there were also significant differences between the mean intensities of events at different distances from the coastline (*F*=260.17, *SS*=78.6, *p*<0.001), and between the durations in events (*F*=51.39, *SS*= 15.53, *p*<0.001) and between sites.
+Similar comparisons using the CMC dataset showed that there was no significant difference in the duration of events at different distances from the coastline for each of the sites (Three-Way ANOVA: *F*=1.20, *SS*=0.41, *p*>0.05).
+However, there were significant differences in the mean intensities of events at different distances (*F*=471.63, *SS*=162.0, *p*<0.05), between the durations of events (*F*= 38.36, *SS*=13.2, *p*<0.001), and between sites (*F*=17.64, *SS*=18.2, *p*<0.002).
+I also found that there was no significant difference between the durations of upwelling across all sites (*F*=0.22, *SS*=0.2, *p*<0.87).
+When looking at the in situ collected dataset, data collected at the coastline, we find no difference in the durations of events at across sites (*F*=2.05, *SS*=1.84, *p*=0.11) but there was a difference in the mean intensities of those events (*F*=2.05, *SS*=0.61, *p*=0.44).
 

### 2019-10-07 Write_up/Paper1_writeup.Rmd
update

@@
 Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast [@luning1990; @lutjeharms2000; @hutchings; @schlegel].
-Changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
 
@@
 South easterly winds dominate during summer but are still present for short durations during winter [@andrews].
+Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region (RWS: Citation) and as a result conditions are variable in the nearshore and offshore.
+
 The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood, with Very few studies addressing the changes in the intensity of cold events at varying distances from the coast within the Benguela region.
-
 Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
@@
 Within this study it is hypothesized that the upwelling metric may vary spatially across different distances, with more intense upwelling events occurring nearest to the coast.
+
 Strong temperature gradients existing between offshore and inshore oceanic waters may result in different upwelling event readings.
@@
 ## Results
-
-Visualisations
 
@@
 
+# Plotting images
+
+####TO do
+- Plots
+
 # Discussion

### 2019-10-08 SST_patterns.Rmd
update

@@
 ```{r}
+options(scipen = 999)
 anova_func <- function(df){
@@
 
-Create visualisation which compares the presence and absence of events
-If a event is found in a particular site at a particulate dat, is the same event present in the other datasets? Show how the events detected varies between site and dataset.
+Plot showing upwelling at the different sites for each of the different distances
 
 ```{r}
-# To determine wether or not the same upwlling patterns are present at different distances from the coastline
 
+scatter_func <- function(df){
+  CMC_filtered <- df %>%
+filter(year(date_start) %in% seq(2010, 2016, 1))
+}
+
+CMC_filtered <- scatter_func(df = CMC_upwell_base)
+OISST_filtered <- scatter_func(df = OISST_upwell_base)
+SACTN_filtered <- scatter_func(df = SACTN_upwell_base)
+
+plot_scat_func <- function(df){
+  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
+  geom_point() +
+  scale_x_date(expand = c(0, 0)) +
+  facet_wrap(~distance, ncol = 2) +
+  theme_grey() +
+  labs(y = "Mean Intensity", x = NULL)
+}
+(CMC_plot <- plot_scat_func(df = CMC_filtered))
+(OISST_plot <- plot_scat_func(df = OISST_filtered))
+
+plot_scat_func <- function(df){
+  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
+  geom_point() +
+  scale_x_date(expand = c(0, 0)) +
+  facet_wrap(~site, ncol = 2) +
+  theme_grey() +
+  labs(y = "Mean Intensity", x = NULL)
+}
+(SACTN_plot <- plot_scat_func(df = SACTN_filtered))
 ```

### 2019-10-08 Write_up/Paper1_writeup.Rmd
update

@@
 
+##
+
 <!-- \def\tiny{\@setfontsize\tiny{10pt}{11pt}} -->
@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{OISST} \\
+\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the OISST dataset.
+The site column gives the name of the site.
+The distance column gives the distance from the coastline as seen in Fig 1.
+The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
 
@@
 \toprule
-Site & distance & slope  \\
+Site & distance & regression slope  \\
 \midrule
@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{CMC} \\
+\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the CMC dataset.
+The site column gives the name of the site.
+The distance column gives the distance from the coastline as seen in Fig 1.
+The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
 
@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{SACTN} \\
+\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the SACTN dataset.
+The site column gives the name of the site.
+Tempertures were collected at the coastline and not at different distances from the coastline.
+The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
 
@@
 
+The linear regression analysis showed that the absolute cumulative intensity among all variables within each dataset showed the most variation.
+The cumulative intensity acts as absolute magnitude rather than relative seasonal climatology.
+When looking at the results for the data obtained from the OISST dataset it is evident that R^2 values are highest at greater distances from the coastline for each of the sites except for Port Nolloth.
+For examples, Lamberts Bay had an R^2 value of 5.47  at a distance of 10000km but only an R^2 value of -8.76 at 50000km.
+Similar trends were apparent for Saldanha Bay (R^2 -3.28 at 10000 km and -12.42 at 50000 km) and Sea Point (R^2  = -8.77 at 10000 km)
+
+The CMC remotely sensed sea surface temperature dataset had higher R^2 values compared to the OISST dataset.
+Most of the sites within CMC dataset had greater R^2 values at distances furthest from the coastline.
+Here, Lamberts Bay had R^2 value of -26.30 for 10000km and -31.77 for 50000km respectively.
+Similarly, Port Nolloth had R^2 values of -15.34 at 10000km and -20.45 at 50000km.
+Saldanha Bay however presented its lowest value of -24.84 at 10000km and highest R^2 value of -46.39 at 40000km and also the highest value among all of the sites.
+Additionally, Sea Point displayed similar patterns with the lowest value occurring at a distance of 10000km from the coastline and its highest R^2 value of -38.35 occurring at a distance of 40000km from the coastline.
+
+Lastly, the SACTN in situ collected SST data collected directly at the coastline had variable R^2 values.
+Lamberts Bay had an R^2 value of -1.46.
+Port Nolloth, representing the second highest value of -3.65.
+Saldanha Bay representing an R^2 value of -2.56 and the highest R^2 value of -10.96 occurring at Sea Point.
+
 # Wind rose diagrams
@@
 
-# Plotting images
-
-####TO do
-- Plots
-
-# Discussion
+## Discussion

### 2019-10-08 SST_patterns.Rmd
Plots

@@
 
-trends_duration_func <- function(df){
-  upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
-                     family = poisson(link = "log"))
-  upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
-  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
-                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-}
-
-OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
-CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
-SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
-
-### Intensity: Chnaged the family and link because of the negative values within the dataset 
-trends_intensity_func <- function(df){
-  upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
-                     family = gaussian(link = "identity"))
-  upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
-  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
-                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-}
-
-OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
-CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
-SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
+# trends_duration_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
+#                      family = poisson(link = "log"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
+# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
+# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
+# 
+# 
+# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
+# trends_intensity_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
+#                      family = gaussian(link = "identity"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
+# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
+# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
 ```
@@
 
+plot_loli_func <- function(df){
+  ggplot(df, aes(x = date_peak, y = duration)) + 
+  geom_lolli(aes(colour = intensity_cumulative)) + 
+  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  facet_wrap(~distance, ncol = 2) +
+  xlab("Date") + ylab("Event duration [days]") 
+}
+
+(CMC_plot <- plot_loli_func(df = CMC_filtered))
+(OISST_plot <- plot_loli_func(df = OISST_filtered))
+
 plot_scat_func <- function(df){
@@
 (SACTN_plot <- plot_scat_func(df = SACTN_filtered))
+
+plot_loli_func <- function(df){
+  ggplot(df, aes(x = date_peak, y = duration)) + 
+  geom_lolli(aes(colour = intensity_cumulative)) + 
+  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  facet_wrap(~site, ncol = 2) +
+  xlab("Date") + ylab("Event duration [days]") 
+}
+(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
 ```

### 2019-10-08 SST_patterns.Rmd
Plotting Bits

@@
 
-Plot showing upwelling at the different sites for each of the different distances
+Plot showing upwelling at the different sites for each of the different distances. Scatter plot
 
 ```{r}
-
-scatter_func <- function(df){
-  CMC_filtered <- df %>%
+library(ggpubr)
+year_func <- function(df){
+  Filtered <- df %>%
 filter(year(date_start) %in% seq(2010, 2016, 1)) 
@@
 
-CMC_filtered <- scatter_func(df = CMC_upwell_base)
-OISST_filtered <- scatter_func(df = OISST_upwell_base)
-SACTN_filtered <- scatter_func(df = SACTN_upwell_base)
+CMC_filtered <- year_func(df = CMC_upwell_base)
+OISST_filtered <- year_func(df = OISST_upwell_base)
+SACTN_filtered <- year_func(df = SACTN_upwell_base)
+
+CMC_SEAPOINT <- CMC_filtered %>% 
+  filter(site == "Sea Point")
+CMC_SB <- CMC_filtered %>% 
+  filter(site == "Saldanha Bay")
+CMC_PN <- CMC_filtered %>% 
+  filter(site == "Port Nolloth")
+CMC_LB <- CMC_filtered %>% 
+  filter(site == "Lamberts Bay")
 
@@
 }
-(CMC_plot <- plot_scat_func(df = CMC_filtered))
-(OISST_plot <- plot_scat_func(df = OISST_filtered))
+(CMC_SP_plot <- plot_scat_func(df = CMC_SEAPOINT))
+(CMC_SB_plot <- plot_scat_func(df = CMC_SB))
+(CMC_PN_plot <- plot_scat_func(df = CMC_PN))
+(CMC_LB_plot <- plot_scat_func(df = CMC_LB))
 
-plot_loli_func <- function(df){
-  ggplot(df, aes(x = date_peak, y = duration)) + 
-  geom_lolli(aes(colour = intensity_cumulative)) + 
-  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+combined_CMC <- ggarrange(CMC_SP_plot, CMC_SB_plot, CMC_PN_plot,CMC_LB_plot)
+###################
+## OISST
+OISST_SEAPOINT <- OISST_filtered %>% 
+  filter(site == "Sea Point")
+OISST_SB <- OISST_filtered %>% 
+  filter(site == "Saldanha Bay")
+OISST_PN <- OISST_filtered %>% 
+  filter(site == "Port Nolloth")
+OISST_LB <- OISST_filtered %>% 
+  filter(site == "Lamberts Bay")
+
+plot_scat_func <- function(df){
+  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
+  geom_point() +
+  scale_x_date(expand = c(0, 0)) +
   facet_wrap(~distance, ncol = 2) +
-  xlab("Date") + ylab("Event duration [days]") 
+  theme_grey() +
+  labs(y = "Mean Intensity", x = NULL)
 }
+(OISST_SP_plot <- plot_scat_func(df = OISST_SEAPOINT))
+(OISST_SB_plot <- plot_scat_func(df = OISST_SB))
+(OISST_PN_plot <- plot_scat_func(df = OISST_PN))
+(OISST_LB_plot <- plot_scat_func(df = OISST_LB))
+combined_OISST <- ggarrange(OISST_SP_plot, OISST_SB_plot, OISST_PN_plot,OISST_LB_plot)
 
-(CMC_plot <- plot_loli_func(df = CMC_filtered))
-(OISST_plot <- plot_loli_func(df = OISST_filtered))
+############## 
+# SACTN
 
@@
 
+```
+
+# Lolli plot
+
+```{r}
+plot_loli_func <- function(df){
+  ggplot(df, aes(x = date_peak, y = duration)) + 
+  geom_lolli(aes(colour = intensity_cumulative)) + 
+  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  facet_wrap(~distance, ncol = 2) +
+  xlab("Date") + ylab("Event duration [days]") 
+}
+
+(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
+(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
+(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
+(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
+combined_OISST_pl <- ggarrange(OISST_SP_pl, OISST_SB_pl, OISST_PN_pl,OISST_LB_pl)
+
+#######
+# CMC
+
+(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
+(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
+(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
+(CMC_LB_pl <- plot_loli_func(df = CMC_LB))
+combined_CMC_pl <- ggarrange(CMC_SP_pl, CMC_SB_pl, CMC_PN_pl,CMC_LB_pl)
+
 plot_loli_func <- function(df){

### 2019-10-08 SST_patterns.Rmd
update

@@
 
+CMC_SP_plot <- CMC_SP_plot +
+  ggtitle("Sea Point")
+CMC_SB_plot <- CMC_SB_plot +
+  ggtitle("Saldanha Bay")
+CMC_PN_plot <- CMC_PN_plot +
+  ggtitle("Port Nolloth")
+CMC_LB_plot <- CMC_LB_plot +
+  ggtitle("Lamberts Bay")
+
 combined_CMC <- ggarrange(CMC_SP_plot, CMC_SB_plot, CMC_PN_plot,CMC_LB_plot)
@@
 (OISST_LB_plot <- plot_scat_func(df = OISST_LB))
+
+OISST_SP_plot <- OISST_SP_plot +
+  ggtitle("Sea Point")
+OISST_SB_plot <- OISST_SB_plot +
+  ggtitle("Saldanha Bay")
+OISST_PN_plot <- OISST_PN_plot +
+  ggtitle("Port Nolloth")
+OISST_LB_plot <- OISST_LB_plot +
+  ggtitle("Lamberts Bay")
+
 combined_OISST <- ggarrange(OISST_SP_plot, OISST_SB_plot, OISST_PN_plot,OISST_LB_plot)
@@
 (OISST_LB_pl <- plot_loli_func(df = OISST_LB))
+OISST_SP_pl <- OISST_SP_pl +
+  ggtitle("Sea Point")
+OISST_SB_pl <- OISST_SB_pl +
+  ggtitle("Saldanha Bay")
+OISST_PN_pl <- OISST_PN_pl +
+  ggtitle("Port Nolloth")
+OISST_LB_pl <- OISST_LB_pl +
+  ggtitle("Lamberts Bay")
+
 combined_OISST_pl <- ggarrange(OISST_SP_pl, OISST_SB_pl, OISST_PN_pl,OISST_LB_pl)
@@
 (CMC_LB_pl <- plot_loli_func(df = CMC_LB))
+
+CMC_SP_pl <- CMC_SP_pl +
+  ggtitle("Sea Point")
+CMC_SB_pl <- CMC_SB_pl +
+  ggtitle("Saldanha Bay")
+CMC_PN_pl <- CMC_PN_pl +
+  ggtitle("Port Nolloth")
+CMC_LB_pl <- CMC_LB_pl +
+  ggtitle("Lamberts Bay")
+
 combined_CMC_pl <- ggarrange(CMC_SP_pl, CMC_SB_pl, CMC_PN_pl,CMC_LB_pl)

### 2019-10-08 Write_up/Paper1_writeup.Rmd
update

@@
 
-##
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for the OISST dataset.
+The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/OISST.png")
+```
+
+Sea Point showcased the most intense upwelling event at 10000 km from the coastline during the year 2013, lasting for 21 days (Refer to figure, do this throughout).
+The majority of upwelling events in Sea Pont occurred around 10000 km away from the coastline lasted for less than five days, and had an approximate cumulative intensity ranging between of -10.00deg.
+C x days and 12.00 deg.
+C x days.
+Conversely, at shorter distances like 3000 km from the coastline several upwelling events lasted for longer than 15 days with a cumulative intensity ranging between -30deg.
+C x days and -20deg.
+C x days.
+At a distance of 30000 km both dominant events appear at a distance of 10000 km are present along with several other additional events (This doesn’t make sense).
+At a distance of 40000 km, less intense events occurred.
+At a distance of 50000 km from the coastline, several events appear to have lasted for approximately 2 – 3 days with the most intense events still occurring in 2015.Events within Saldanha Bay typically lasted for approximately three days but varying at different distances.
+At a distance of 10000 km one major event occurred that lasted for approximately 27 days with a cumulative intensity of -61.70deg.
+C x days.
+This event was also present at a distance of 20000 km from the coastline, albeit at a lesser intensity, but was not present at 30000km, 40000km and 50000 km respectively.
+Events occurring within 30000 km and 40000 km from the coastline lasting for approximately 10 days in most cases.
+However, the majority of events were observed at a distance of 10000 km and 20000 km from the coastline.
+Fewer events occurred at a distance of 50000 km, with most of the events occurring at this distance having lasted for less than five days.
+
+Throughout Port Nolloth, most of the events detected lasted less than five days.
+At a distance of 10000 km one obvious major event occurred during 2016 which had a cumulative intensity of -43.39deg.
+C x days.
+This event persisted at 20000 km and 30000 km with cumulative intensities of -21.09deg.
+C x days and -14.33 deg.
+C x days respectively.
+Here, most upwelling events were detected at 10000km and 20000km respectively.
+The least number of events were detected at 50 000 km.
+The least intense upwelling event lasted for only 3 days and had a cumulative intensity of -5.45deg.
+C x days during the year 2010.
+
+Events occurring within Lamberts Bay lasted for approximately less than five days.
+Three major events were detected at 10000 km and 20000 km away from the coastline respectively.
+At 30000 km an event with a longer duration occurs in 2010, that was not present at a distance of 10000 km and 20000 km from the coastline.
+At a distance of 40000 km no major events were detected.
+However, at a distance of 50000 km a major upwelling event is detected lasting for longer than 20 days.
+Fewer events were however detected at a distance of 50000 km from the coastline.
+
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for the SACTN dataset.
+The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/SACTN.png")
+```
+
+Within the SACTN dataset, the majority of upwelling events detected in Lamberts Bay lasted for longer than five days.
+The longest upwelling event was detected in 2012 lasting for approximately 12 days with a cumulative intensity of -12.63deg.
+C x days.
+Several upwelling events were detected at this site.
+Only eight upwelling events were recorded and detected at Port Nolloth.
+Here most of the lasted between three to five days with the most intense event lasting for 11 days with a cumulative intensity of -12.81deg.
+C x days.
+Saldanha Bay upwelling events also lasted between three to five days.
+The event with the longest duration occurred in 2016 with a cumulative intensity of -22.58deg.
+C x days.
+Events observed within Sea Point lasted less than five days, but two events lasted for 21 days.
+Those events represented intensities of -54.54 and -36.81 deg.
+C x days respectively.
+
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for the CMC dataset.
+The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/CMC.png")
+```
+
+CMC observations of upwelling events at Sea Point detected durations of less than four days.
+The most intense event occurred in 2010, and I found that intensifies appeared to increase with distance from the coastline.
+The fewest number of events occurred 50000 km away from the coastline.
+At Saldanha Bay, the duration of upwelling events mostly ranged between three to four days.
+Similar intensity patterns occurred at 10000 km, 20000 km and 30000 km from the coastline respectively.
+At 40000 km an event representing a cumulative intensity of -28.64deg.
+C x days for was detected for 20 days within 2010.
+At Port Nolloth, most of the events are found to occur for duration of three days.
+Two major upwelling events were present at different distances within 2010 with a variation in their respective cumulative intensities.
+Those events both lasted for approximately seven days.
+Similar patterns were found within Lamberts Bay.
+At 30000 km a major event was detected that lasted for approximately 19 days, but most events here lasted for less than five days.
+At a distance of 40000 km and 50000 km from the coastline similar patterns were observed, with most events being detected at 40000 km and 50000 km respectively.
 

### 2019-10-08 Appendices.Rmd
update

---
title: "Appendices"
author: "Amieroh Abrahams"
date: "08 October 2019"
output: html_document
---
```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE 
)

# Disable scientific notation for numeric values
# I just find it annoying
options(scipen = 999)

library(tidyverse)
library(lubridate)
library(ggpubr); theme_set(theme_pubr())
library(zoo)
library(data.table)
library(heatwaveR)
library(viridis)
library(gridExtra)
library(fossil)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(plyr)
library(ggpubr)
library(zoo)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(fossil)
library(mapproj)
library(ncdf4)
library(reshape2)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(circular)
# devtools::install_github("robwschlegel/coastR")
library(gridExtra)
library(geosphere)
library(tidyverse)
library(heatwaveR)
# install_github("marchtaylor/sinkr")

## Functions
source("Functions/earthdist.R")
source("Functions/wind.rose.R")
source("Functions/theme.R")
source("Functions/scale.bar.func.R")
```

# Loading the data and finding a 30year time series

```{r}
load("Data/SACTN_daily_v4.2.RData")
load("Data/site_list_v4.2.RData")

site_list_sub <- site_list %>%
  filter(coast == "wc") %>%
  filter(length > 10950) 
#site_list_sub <- site_list_sub[-5,]
save(site_list_sub, file = "Data/site_list_sub.Rdata")

SACTN_US <- SACTN_daily_v4.2 %>%
  left_join(site_list[,c(4,13)], by = "index") %>%
  filter(index %in% site_list_sub$index) %>%
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE)

save(SACTN_US, file = "Data/SACTN_US.RData")
```

```{r}
load("Data/site_list_sub.Rdata")
load("Data/africa_coast.RData")
xtable(site_list_sub, auto = TRUE)
west <- site_list_sub[-5,]
west$coast <- "west" # Chnages wc to west

site_list_sub <- site_list_sub[-5,]

load("Data/sa_bathy.RData")
## Downloading the bathy data from NOAA
# Download mid-res bathymetry data
# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
# colnames(sa_bathy) <- c("lon", "lat", "depth")
# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")

# Loading in the newly downloaded bathymetry data
load("Data/sa_bathy.RData")
load("Data/africa_coast.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
shore.normal.transect <- function(site, width = 2){
  # Find the site on the coastline and it's nearest neighbour points
  coords <- data.frame(lon = site$lon, lat = site$lat)
  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
  coords3 <- coords3[2:1,1:3]
  # Define the shore normal transect bearing
  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
  if(heading >= 360){
    heading <- heading-360
  } else {
    heading <- heading
  }
  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
  return(heading2)
}

# Creating the transects
site_transects <- data.frame()
for(i in 1:length(west$site)){
 site <- west[i,]
 site_transect <- shore.normal.transect(site, 2)
 site_transects <- rbind(site_transects, site_transect)
}

# Manually correcting Sea Point and Kommetjie
site_transects$heading[4] <- 290 
# save(site_transects, file = "Data/site_transects.RData")
load("Data/site_transects.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath

transect.pixel <- function(site, distances){
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  # deep <- 999
  # distance_multiplier <- 1
  # while(deep > isobath){
  for(i in 1:length(distances)){
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, 
                         distance = distances[i])
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
  }
  if(nrow(pixels) < 1){
    pixels <- data.frame(site, depth = NA)
  }else{
    pixels <- pixels
  }
  return(pixels)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
  site_pixels <- rbind(site_pixels, site_pixel)
}

# save(site_pixels, file = "Data/site_pixels.RData")
load("Data/site_pixels.RData")

# Bounding box
  # Only one is made in order to know how large the the geom_point() squares should be made to match
bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])

# Determining the temperature at the various distances from the coast
```

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline. The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=6, fig.width=10, fig.pos="H"}

load("Data/south_africa_coast.RData")
load("Data/africa_coast.RData")
load("Data/sa_provinces_new.RData")
load("Data/site_list_sub.Rdata")
load("Data/site_pixels.RData")
load("MUR.RData")
names(south_africa_coast)[1] <- "lon"

# Manually divide up coastline
wc <- south_africa_coast[291:410,]

site_list_sub <- site_list_sub[-5,]

# Setting up the theme
theme_set(theme_bw())
limits <- c(12,28) # for colour bar
#breaks <- seq(6, 30, 2) # Create breaks to be used for colour bar

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 25)

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  xlab("") + ylab("") +
  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
  #            size = 4, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size =4, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 4, angle = 302, colour = "black") +
    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
    #          size = 5, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5),
        axis.text = element_text(size = 10)) +
    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.direction = "horizontal",
        legend.justification = c(1,0),
        legend.position = c(0.85, 0.70),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions

africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
  theme_bw() + 
  coord_equal() + 
  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
  theme(panel.border = element_rect(colour = "black", size = 0.4), 
        plot.background = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
africa_map

# More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html

vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
print(site_map, vp = vp2)
print(africa_map, vp = vp1)
dev.off()
pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions

# Site map
# The insetting needs to be placed at appropriate location
```

Loading the Wind data. The wind dataset was supplied by the SAWS in .text format. Wind data was provided for the sites closes to the requested site as wind data may not be available for the 6 study sites in this research.  The wind direction was collected every 2 hours and by using the circular function we calcuated the daily wind direction and mean speed. There is a numerically large gap between 360 and 2 where as in degrees its not as large. The circular mean function returns the mean direction of a vector of circular data. https://cran.r-project.org/web/packages/circular/circular.pdf (Pg 130). The circular function creates circular objects around the wind direction.

```{r}
wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "", 
                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))

wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

# Slecting the important columns for each of the datasets
wind_fix <- function(df){
wind <- df %>% 
  select(station_name, date, hour, dir, speed) %>%  #column names
  mutate(date = as.Date(as.character(date)),
         hour = as.numeric(as.character(hour)), 
         dir = as.numeric(as.character(dir)),
         speed = as.numeric(as.character(speed)))
}
# RWS: We can see when we force the values to be numeric that there are some non-numeric values in the base data
wind_1 <- wind_fix(df = wind_1)
wind_2 <- wind_fix(df = wind_2)
wind_3 <- wind_fix(df = wind_3)

## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected

renaming_sites_1 <- function(df) {
  sites <- df %>%
      # RWS: An alternative way to replace values without having to use multiple ifelse() statements
      mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
                                    station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
                                    station_name  == "PORT NOLLOTH" ~"Port Nolloth",
                                    station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
                                    station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
                                    station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
  return(sites)
}

wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
save(wind_data, file = "Data/wind_data.RData")

# # RWS: You should never be removing values by specific call like this
#   # Rather you should be able to use some sort of conditional to screen out unwanted values
# wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
# renaming_sites_2 <- function(df) {
#   sites <- df %>%
#     mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
#                            ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
#                                 ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
#                                        ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
#   return(sites)
# }
# 
# wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
# wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ] 
# renaming_sites_3 <- function(df) {
#   sites <- df %>%
#     mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
#                            ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
#                            ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
#   return(sites)
# }
# 
# wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
# wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ] 
### CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
# wind_3_HoutBay <- wind_3 %>% 
#   filter(station_name == "CAPE TOWN SLANGKOP") %>% 
#   mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
# wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
# wind_data <- wind_data %>% 
#   na.omit()
# 
# wind_data <- wind_data %>% 
#   # mutate(date = as.Date(date)) %>%  
#   group_by(date, hour, temp_sites) 
# 
# # write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
# # save(wind_data, file = "Data_P1/wind_data.RData")

# load("Data_P1/insitu_wind.RData")
# 
# test1 <- insitu_wind %>% 
#   mutate(lat = case_when(site == "Port Nolloth" ~ "-29.25241",
#                                     site == "Saldanha Bay" ~ "-33.01054",
#                                     site  == "Yzerfontein" ~ "-33.36068",
#                                     site  == "Kommetjie" ~ "-34.13667",
#                                     site  == "Lamberts Bay" ~ "-32.09185",
#                                     site  == "Sea Point" ~ "-33.91847"))
# 
# test <- test1 %>% 
#   mutate(lon = case_when(site == "Port Nolloth" ~ "16.86710",
#                                     site == "Saldanha Bay" ~ "17.95063",
#                                     site  == "Yzerfontein" ~ "18.15731",
#                                     site  == "Kommetjie" ~ "18.32674",
#                                     site  == "Lamberts Bay" ~ "18.30262",
#                                     site  == "Sea Point" ~ "18.38217"))
# 
# test_final <- test %>% 
#   select(date,lat,lon,dir_circ, mean_speed) %>% 
#   dplyr::rename(speed = mean_speed) %>% 
#   dplyr::rename(dir = dir_circ)

# Loading the wind data
library(tidyverse)
library(circular)

# wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number())) # Dataset too large for Github but better data at line 638

# wind_data <- read_csv("Data_P1/wind_data.csv")
selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")

wind_daily <- wind_data %>%
  # ungroup() %>%
  # mutate(date = as.Date(date))%>%
  dplyr::group_by(temp_sites, date) %>%
  filter(temp_sites %in% selected_sites) %>% 
  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
                   mean_speed = round(mean(speed),2))

# save(wind_daily, file = "Data/wind_daily.RData")
# # Loading in the daily wind data

# Loading the subsetted wind data created from the text files above 
load("Data/wind_daily.RData") 
```

# Upwelling

Upwelling is primarily caused by alongshore, equator ward winds. These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods. Upwelling is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.

# Upwelling indeces
  # Determining upwelling index from wind data (SAWS)
  # Index Equation from Fielding & Davis 1989 paper

$$ UpwellingIndex = μ{(Cosθ − 160)}$$

In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The 160 degrees is used as this refers to the angle of the coastline. This equation is largely dependant on wind speed and direction data in order to determing the intensity of the upwelling event. Wind data were obtained daily from the South African Weather Service (SAWS). This wind data was then matched to the date at which temperature were collected. With this data the upwelling index was determined.

# Steps
  - Use the circular function
  - Determine daily wind data obtained from 3hr intervals

```{r}
# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
# 
# wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# # Slecting the important columns for each of the datasets
# wind_fix <- function(df){
# wind <- df %>%
#   select(station_name, date, hour, dir, speed) %>%  #column names
#   mutate(date = as.Date(as.character(date)),
#          hour = as.numeric(as.character(hour)),
#          dir = as.numeric(as.character(dir)),
#          speed = as.numeric(as.character(speed)))
# }
# 
# wind_1 <- wind_fix(df = wind_1)
# wind_2 <- wind_fix(df = wind_2)
# wind_3 <- wind_fix(df = wind_3)
# 
# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
# 
# renaming_sites_1 <- function(df) {
#   sites <- df %>%
#       mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
#                                     station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
#                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
#                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
#                                     station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
#   return(sites)
# }
# 
# 
# wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
# wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
# wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
# wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
# save(wind_data, file = "Data/wind_data.RData")
# 
# load("Data/wind_data.RData")
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
# 
# wind_daily <- wind_data %>%
#   # ungroup() %>%
#   # mutate(date = as.Date(date))%>% # Changing the date to date format
#   dplyr::group_by(temp_sites, date) %>%
#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2), # Direction using the circular function
#                    mean_speed = round(mean(speed),2)) # Calculating the mean
# save(wind_daily, file = "Data/wind_daily.RData")
```

# Load wind

```{r}
load("Data/wind_daily.RData") 
wind_daily <- wind_daily %>% 
  dplyr::rename(sites = temp_sites)
load("Data/site_list_sub.Rdata")
```

# Determining the angle perpendicular to the coastline and plotting it

```{r}
# Coast R- function that calculates the angle of the site perpendicular to the coastline
# Transect function in coastR package finds the angle away or along the coastline
south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)

# Plotting the map
world_map <- ggplot() + 
  borders(fill = "grey40", colour = "black")

# Create titles
titles <- c("Shore-normal", "Islands")
# Plotting function
plot_sites <- function(site_list, buffer, title_choice, dist){

  # Find the point 200 km from the site manually to pass to ggplot
  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
                                              b = site_list$heading, d = dist))

  # Add the new coordinates tot he site list
  site_list <- site_list %>% 
    mutate(lon_dest = heading2$lon,
           lat_dest = heading2$lat)

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder") +
    ggtitle(titles[title_choice])
}
south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
south_africa_away
```

# Upwelling indeces using the above formula

```{r}
UI_angle <- wind_daily %>%  # Making refeerence to the wind daily datta created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to south africa wide data column 2 and 21. by site 
  dplyr::rename(coast_angle = heading) %>% # Renamingthe column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  drop_na %>% # removing the na values
  dplyr::rename(t = date) #renaming date to t

# save(UI_angle, file = "Data/UI_angle.RData")
# index <- wind_UI %>%
#   group_by(site) %>%
#   select("date", "ui.saws") %>%
#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
# wind_UI <- wind_daily %>% 
#   mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
#   drop_na
# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
```

# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)
# Upwelling temperature threshold 12.4 / 30th percentile
SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")), # creating a column of climaatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance) # selecting only these variables
# Calculate quantiles of upwelling index

# Static numbers are often rejected and so we decided to find a percentile value as these are often more likely to be approved

SACTN_upwell_quantiles <- SACTN_upwell %>% 
  filter(ui.saws >= 0) %>% # Upwelling occurs for all values above 0. Values below this is regarded as downwelling
  group_by(site) %>% 
  summarize(quant_10 = quantile(ui.saws, probs = 0.10, na.rm = TRUE),
            quant_25 = quantile(ui.saws, probs = 0.25, na.rm = TRUE),
            quant_50 = quantile(ui.saws, probs = 0.50, na.rm = TRUE),
            quant_75 = quantile(ui.saws, probs = 0.75, na.rm = TRUE),
            quant_90 = quantile(ui.saws, probs = 0.90, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, digits = 2)

# SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
# There are a few ideas on how to go about doing this
# The first is to calculatea rolling climatology and to check the upwelling phenology agianst that
# in order to determine changes in upwelling values, but this won't give us the metrics we want
# Rather what we can do is set a static bottom threshold below which we are interested in looking for upwelling
# We then use the upwelling index values created above as a second filter to determine the metrics of the upwelling
# Unfortunately the way that the heatwaveR functions work well does with purrr. 
# I've figured this out once before but can't remember where now.
# Rather than go through the process I have rather just made a convenience function below that
# does what I want and can simply be given to purrr::map() without any faffing about.

# Detect event: 
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
  return(res)
}
# Calculate the upwelling event metrics
SACTN_upwell_base <- SACTN_US %>% 
  dplyr::rename(t = date) %>% 
  group_by(site) %>% 
  nest() %>% 
  # The climatology base period used here is up for debate...
  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
  select(-data) %>% 
  unnest() %>%
  left_join(SACTN_upwell, by = c("site", "t")) %>%
  filter(!is.na(exceedance)) %>%
  group_by(site) %>% 
  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
  nest() %>% 
  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
  select(-data) #%>% 
  # unnest() %>% 
  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
  # filter(row_number() %% 2 == 1) %>% # Select daily values
  # unnest()
# Unpack the event metric reults
SACTN_upwell_events <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 0) %>%
  unnest()
# save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
# Unpack the daily climatology results
SACTN_upwell_clims <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>% 
  unnest()
# save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
# The above chunk of code appears to work as expected.
# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn. 
    # Min duration for an upwelling event - 1day and threshold -1
# What also needs to be decided is if we are interested in any upwelling results throughout the year,
# or only during the upwelling season, whenever that may be for each site.
# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
# is changing over time.
# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
```

# Loading in all the data created using the code bellow

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData") # 5 decimal places
load("Data/OISST.RData") # 2 decimal places
OISST <- BC_avhrr_only_v2_Document_Document 
rm(BC_avhrr_only_v2_Document_Document ); gc()
load("Data/CMC.RData") # 1decimal places
```

 Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
# Is this upwelling event seen throughout the different distances from the coastline
# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets

## Function
Determining the temperatures at the various distances from the coastline

```{r}
# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
# # MUR <- MUR %>% 
# #   select(lon, lat) %>% 
# #   mutate(product = "MUR")
# 
# # Decided to work with OISST and CMC as both has a time series of 30years
# OISST_prod <- OISST %>% 
#   select(lon, lat) %>% 
#   unique() %>% 
#   mutate(product = "OISST")
# 
# CMC_prod <- CMC %>% 
#   select(lon, lat) %>%
#   unique() %>% 
#   mutate(product = "CMC")
# 
# # sat_data <- rbind(CMC_prod, OISST_prod) %>% 
#   #rbind(., CMC) %>% 
#   # select(product, lon, lat)
# 
# # sat_pixels <- sat_data %>% 
#   # select(product, lon, lat) %>%
#   # unique()
# 
# ## For testing the nest/map pipeline
# # df <- site_pixels %>%
# #   filter(site == "Lamberts Bay") %>%
# #   select(-site)
# 
# match_func <- function(df){
#   df <- df %>%
#     dplyr::rename(lon_site = lon, lat_site = lat)
#   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
#                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
#     cbind(., df)
#   CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
#                                              as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
#     cbind(., df)
#   res <- rbind(OISST_index, CMC_index)
#   return(res)
# }
# 
# # Find the nearest pixels to each spot along the transect
# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
# # points to not fall inside of the same 25 KM pixel
# pixel_match <- site_pixels %>% 
#   group_by(site) %>% 
#   group_modify(~match_func(.x))
# 
# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
# OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
# CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
# 
# # Clean up some RAM space
# rm(OISST, CMC); gc()
# 
# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
# # # Hout Bay will be ignored
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
# 
# OISST_fill <- OISST_fill %>%
#   filter(site %in% selected_sites)
# 
# CMC_fill <- CMC_fill %>%
#   filter(site %in% selected_sites)
# 
# 
# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
# load("Data/UI_angle.RData")
# 
# upwelling <- UI_angle %>% 
#   dplyr::rename(temp = ui.saws) %>%
#   group_by(site) %>%
#   # mutate(min_t = min(t), 
#   #        max_t = max(t)) %>% 
#   nest() %>% # apply the following functions to all of the variables in te dataset
#   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
#          # NB: A threshold of 3 appeared to be far to strict
#          # purr::map - apllies a function to each element of a vector
#          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
#   # Detect consecutive days in exceedance of a given threshold.
#   # mutate() %>% 
#   select(-data, -clim) %>% 
#   unnest() %>%
#   filter(row_number() %% 2 == 1) %>%
#   unnest() %>% # creates a column for each variables
#   dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
#   select(site, t, ui.saws, exceedance) 
# 
# # Now applying the Upwelling func
# # NB: This only pulls out the event results and not the climatology results
# # This is done to keep the output tidy because group_modify() may only create data.frame type outputs
# # To create a list output one would use group_map(), 
# # but this then loses the labels of which sites etc. the results belong to
# detect_event_custom <- function(df){
#   res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
#   return(res)
# }
# 
# ts2clm_custom <- function(df){
#   # The climatology base period used here is up for debate...
#   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
#   res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
#   return(res)
# }
# 
# # Calculate the upwelling event metrics
# upwelling_detect_event <- function(df){
#   upwell_base <- df %>% 
#     dplyr::rename(t = date) %>% 
#     group_by(site, product, heading, distance, lon, lat) %>% 
#     group_modify(~ts2clm_custom(.x)) %>% 
#     left_join(upwelling, by = c("site", "t")) %>%
#     filter(!is.na(exceedance)) %>%
#     group_by(site, product, heading, distance, lon, lat) %>% 
#     group_modify(~detect_event_custom(.x))
#   }
# 
# OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
# save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
# CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
# save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
# 
# # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
# SACTN_US <- SACTN_US %>%
#   filter(site %in% selected_sites)
# 
# load("Data/site_list_v4.2.RData")
# 
# SACTN_upwell_base <- SACTN_US %>%
#   left_join(site_list[,c(4, 5, 6)], by = "index")%>% 
#     dplyr::rename(t = date) %>% 
#     group_by(site, lon, lat) %>% 
#     group_modify(~ts2clm_custom(.x)) %>% 
#     left_join(upwelling, by = c("site", "t")) %>%
#     filter(!is.na(exceedance)) %>%
#     group_by(site, lon, lat) %>% 
#     group_modify(~detect_event_custom(.x))
# 
# save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
```

# PLotting events

Loading the data
```{r}
load("Data/OISST_upwell_base.RData")
load("Data/CMC_upwell_base.RData")
load("Data/SACTN_upwell_base.RData")
```

To accurately compare 
  Start date CMC start date 1991-09-02 end date 2017-03-15
  Start date OISST start date 1981-09- 01  and end date 2018-09-28
  Start date SACTN start date 1973 and end date 2017-12-16

```{r}

lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```

# ANOVA
- Relationship between duration, year and distance as a function of the mean intensity. 
```{r}
options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(intensity_mean ~ site * duration * distance, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```

# Wind plots

```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
library(ggradar)
library(dplyr)
library(scales)
library(tibble)
load("Data/UI_angle.RData")

# Detects NA values, no NA values are present
# Plotting wind variables
source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function

wind_daily_renamed <- UI_angle %>% 
  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
  dplyr::rename(spd = mean_speed) %>%
  dplyr::rename(dir = dir_circ) %>% 
  filter(spd > 0)

p.wr2 <- plot.windrose(data = wind_daily_renamed,
              spd = "spd",
              dir = "dir")

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
p.wr3
```

# Lolli plot

```{r}
plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(aes(colour = intensity_cumulative)) + 
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~distance, ncol = 2) +
  xlab("Date") + ylab("Event duration [days]") 
}

(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
OISST_SP_pl <- OISST_SP_pl +
  ggtitle("Sea Point")
OISST_SB_pl <- OISST_SB_pl +
  ggtitle("Saldanha Bay")
OISST_PN_pl <- OISST_PN_pl +
  ggtitle("Port Nolloth")
OISST_LB_pl <- OISST_LB_pl +
  ggtitle("Lamberts Bay")

combined_OISST_pl <- ggarrange(OISST_SP_pl, OISST_SB_pl, OISST_PN_pl,OISST_LB_pl)

#######
# CMC

(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
(CMC_LB_pl <- plot_loli_func(df = CMC_LB))

CMC_SP_pl <- CMC_SP_pl +
  ggtitle("Sea Point")
CMC_SB_pl <- CMC_SB_pl +
  ggtitle("Saldanha Bay")
CMC_PN_pl <- CMC_PN_pl +
  ggtitle("Port Nolloth")
CMC_LB_pl <- CMC_LB_pl +
  ggtitle("Lamberts Bay")

combined_CMC_pl <- ggarrange(CMC_SP_pl, CMC_SB_pl, CMC_PN_pl,CMC_LB_pl)

plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(aes(colour = intensity_cumulative)) + 
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~site, ncol = 2) +
  xlab("Date") + ylab("Event duration [days]") 
}
(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
```

### 2019-10-08 Write_up/Paper1_writeup.Rmd
update

@@
 ---
-title: "Upwelling variability across the Benguela upwelling systems"
+title: "Variability in upwelling patterns across the Benguela upwelling systems"
 author: "Amieroh Abrahams"

### 2019-10-08 Write_up/Appendices.Rmd
updated

--- 2019-10-08 Appendices.Rmd
@@
 ---
-title: "Appendices"
-author: "Amieroh Abrahams"
-date: "08 October 2019"
-output: html_document
+title: Quantifying the impact of wind and wave action on seawater temperature along the South African coastline (Appendices)
+author:
+- affiliation: University of the Western Cape
+  name: Amieroh Abrahams
+date: "19 November 2018"
+output:
+  pdf_document:
+    fig_caption: yes
+    fig_height: 6.5
+    fig_width: 6.5
+    highlight: default
+    keep_tex: yes
+    latex_engine: xelatex
+    template: ../LaTeX/template-article_ajs.tex
+  html_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: cerulean
+    toc_float: no
+  md_document:
+    variant: markdown_github
+  word_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
+    highlight: zenburn
+    toc: no
+fontsize: 10pt
+geometry: margin=1in
+language: Australian
+csl: ../LaTeX/frontiers.csl
+papersize: A4
+tables: yes
+bibliography: ../LaTeX/appendices.bib
 ---
+
+```{r global_options, include = FALSE}
+knitr::opts_chunk$set(fig.path = 'compliled_figures/', include = TRUE, 
+                      echo = TRUE, warning = FALSE, message = FALSE, 
+                      tidy = FALSE, width.cutoff = 80, cache = FALSE, 
+                      results = "hide", size = "small")
+```
+
+This file was generated in R using Rmarkdown, with a bit of \LaTeX\ thrown in:
+```{r include = TRUE, results='markup'}
+sessionInfo()
+```
+
+\newpage
+
+\section*{\large{Appendix A}}
+
+\subsection*{Background and code}
+
+The intention of this section is to show the approach and **R** scripts used to analyse variation in upwelling events at different distances from the coastline within the Benguela upwelling system.
+Specifically, I wish to determine if the same upwelling signals are present throughout the different distances along the coastline and if the same signals are present within the different datasets.
+
+\subsubsection*{The data}
+The SACTN dataset is the primary source of temperature data used in this study.
+This dataset consisted of *in situ* collected coastal seawater temperatures for 129 sites along the coast of South Africa, measured daily from 1972 until 2017.
+This study also make use of AVHRR OISST and CMC satellite-derived SST datasets as well as wind action data that were obtained from the South African Weather Service (SAWS), and were provided at three hour resolutions.
+
+\subsubsection*{Setting up the analysis}
+This is **R**, so first I need to find, install and load various packages. Some of the packages will be available on CRAN and can be accessed and installed in the usual way.
+
 ```{r prelim_opts, echo=FALSE}
@@
 library(reshape2)
-library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
-library(circular)
-# devtools::install_github("robwschlegel/coastR")
 library(gridExtra)
 library(geosphere)
-library(tidyverse)
-library(heatwaveR)
+library(circular)
+#library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
+# devtools::install_github("robwschlegel/coastR")
+
 # install_github("marchtaylor/sinkr")
@@
 ## Functions
-source("Functions/earthdist.R")
-source("Functions/wind.rose.R")
-source("Functions/theme.R")
-source("Functions/scale.bar.func.R")
+source("../Functions/earthdist.R")
+source("../Functions/wind.rose.R")
+source("../Functions/theme.R")
+source("../Functions/scale.bar.func.R")
 ```
 
-# Loading the data and finding a 30year time series
+\newpage
 
-```{r}
-load("Data/SACTN_daily_v4.2.RData")
-load("Data/site_list_v4.2.RData")
+\section*{\large{Appendix B}}
 
+Now to get to the data.
+Here I load the site list dataset.
+This dataset comprise of the statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures.
+These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents.
+At the broad scale, this region exhibits a large variation in seawater temperatures along its coastline [@Mead2013; @Smit2013] and is divided into four bioregions, each with contrasting temperatures.
+These bioregions are the Benguela Marrine Province (BMP), Benguela-Agulhas Transition Zone (B-ATZ), the Agulhas Marine Province (AMP) and the East Coast Transition Zone (ECTZ) [@Smit2017].
+The Benguela Marine Province (BMP) is located to the west of Cape Point and is characterized by the movement of cold water from the Southern Ocean moving north.
+The cold temperate west coast mentioned above is greatly affected by the upwelling caused by offshore winds, generated by the cold Benguela current.
+Thus the seawater temperatures found around South Africa’s coastline exhibits a large variational range.
+My plotting functions partition the data into the clusters and colour code each point accordingly so that I am able to see patterns that exist and distictly group the different sites with similar temperatures along the three distinct coasts.
+
+```{r load_files1, include=TRUE}
+load("../data/site_list_v4.2.RData")
+```
+
+\newpage
+
+The next step involves SACTN temperature dataset.
+First I compute the `filter` function on a data matrix, which only selects the data along the west coast of South Africa (within the Benguela upwelling region) and select all the sites with a time series of at least 30 years of data.
+
+```{r filtered-sites}
 site_list_sub <- site_list %>%
@@
 
-save(SACTN_US, file = "Data/SACTN_US.RData")
+# save(SACTN_US, file = "Data/SACTN_US.RData") #Saving the file
 ```
 
-```{r}
-load("Data/site_list_sub.Rdata")
-load("Data/africa_coast.RData")
+This thermal data contain various variables, in this study I only make use of some of them.
+This resulted in a total of 5 sites.
+Later in the analyses I however selected only 4 of the five sites as they represented better overlapping years.
+
+<!-- \def\tiny{\@setfontsize\tiny{10pt}{11pt}} -->
+\begin{tiny}
+\begin{center}
+\setlength\tabcolsep{3pt}
+\begin{longtable}{|r|r|r|l|p{3cm}|}
+\caption{The 4 sites located within the Benguela, with approximate GPS coordinates and source} \\
+
+% This is the header for the first page of the table...
+\toprule
+ Site & SRC & Lon & Lat  \\
+\midrule
+\endfirsthead
+
+% This is the header for the second page of the table...
+\toprule
+Site & SRC & Lon & Lat  \\
+\midrule
+\endhead
+
+% This is the footer for all pages except the last page of the table...
+\midrule
+\multicolumn{5}{l}{{Continued on Next Page\ldots}} \\
+\endfoot
+
+% This is the footer for the last page of the table...
+\bottomrule
+\endlastfoot
+
+% Now the data...
+Saldanha Bay & SAWS & 17.95 & -33.01 \\
+Port Nolloth & SAWS & 16.87 & -29.25 \\
+Lamberts Bay & SAWS & 18.30 & -32.09 \\
+Sea Point & SAWS & 18.38 & -33.92 \\
+
+\end{longtable}
+\end{center}
+\end{tiny}
+
+```{r load_files2, include=TRUE}
+load("../Data/site_list_sub.Rdata")
+load("../Data/africa_coast.RData")
 xtable(site_list_sub, auto = TRUE)
@@
 
-site_list_sub <- site_list_sub[-5,]
+site_list_sub <- site_list_sub[-5,] # Removing Hout Bay from the site list
 
-load("Data/sa_bathy.RData")
 ## Downloading the bathy data from NOAA
@@
 # Loading in the newly downloaded bathymetry data
-load("Data/sa_bathy.RData")
-load("Data/africa_coast.RData")
+load("../Data/sa_bathy.RData")
+load("../Data/africa_coast.RData")
+```
 
+Thereafter, inorder to determine the variation of upwelling events at different distances from the coastline it was important to first obtain the coordinates for the different distances.
+The function produces a dataframe for coordinates at the different distances from the coastline for each of the sites.
+The coordinates were obtained at 10000km, 20000km, 30000km, 40000km and 50000km respectively.
+
+```{r shorenormal-func}
 # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
@@
 # save(site_transects, file = "Data/site_transects.RData")
-load("Data/site_transects.RData")
+load("../Data/site_transects.RData")
 
@@
 # save(site_pixels, file = "Data/site_pixels.RData")
-load("Data/site_pixels.RData")
+load("../Data/site_pixels.RData")
 
@@
 
-```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
-The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=6, fig.width=10, fig.pos="H"}
-
-load("Data/south_africa_coast.RData")
-load("Data/africa_coast.RData")
-load("Data/sa_provinces_new.RData")
-load("Data/site_list_sub.Rdata")
-load("Data/site_pixels.RData")
-load("MUR.RData")
-names(south_africa_coast)[1] <- "lon"
-
-# Manually divide up coastline
-wc <- south_africa_coast[291:410,]
-
-site_list_sub <- site_list_sub[-5,]
-
-# Setting up the theme
-theme_set(theme_bw())
-limits <- c(12,28) # for colour bar
-#breaks <- seq(6, 30, 2) # Create breaks to be used for colour bar
-
-# Define plotting parameters
-sa_lats <- c(-37, -27); sa_lons <- c(14, 25)
-
-site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
-  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
-   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
-                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
-  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
-  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
-  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
-  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
-  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
-  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
-  scale_fill_viridis(
-    option = "magma",
-    name = "Temperature (°C)",
-    alpha = 0.8) +
-  xlab("") + ylab("") +
-  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
-  #            size = 4, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
-             size =4, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
-             size = 4, angle = 302, colour = "black") +
-    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
-    #          size = 5, angle = 50, colour = "black") +
-  guides(fill = guide_colourbar()) +
-  scale_x_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°E", sep = "")) +
-    scale_y_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°S", sep = "")) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        legend.text = element_text(size = 14),
-        legend.title = element_text(size = 10),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5),
-        axis.text = element_text(size = 10)) +
-    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.minor = element_line(colour = "NA"),
-        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
-        legend.direction = "horizontal",
-        legend.justification = c(1,0),
-        legend.position = c(0.85, 0.70),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5))
-site_map
-
-pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
-
-africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
-  theme_bw() + 
-  coord_equal() + 
-  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
-  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
-  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
-  theme(panel.border = element_rect(colour = "black", size = 0.4), 
-        plot.background = element_blank(), 
-        axis.ticks = element_blank(), 
-        axis.text = element_blank(), 
-        axis.title = element_blank(), 
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank()) +
-  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
-africa_map
-
-# More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html
-
-vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
-vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
-print(site_map, vp = vp2)
-print(africa_map, vp = vp1)
-dev.off()
-pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
-
-# Site map
-# The insetting needs to be placed at appropriate location
-```
-
-Loading the Wind data.
-The wind dataset was supplied by the SAWS in .text format.
-Wind data was provided for the sites closes to the requested site as wind data may not be available for the 6 study sites in this research.
+Here I created a chunck of code to load in the wind data provided by SAWS.
+This data is presented in single text files and as a result I will not be adding the code here.
+As mentioned earlier, this code is present in my Github repository.
 The wind direction was collected every 2 hours and by using the circular function we calcuated the daily wind direction and mean speed.
@@
 
-```{r}
-wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "", 
-                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
-
-wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
-                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-
-wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
-                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-
-# Slecting the important columns for each of the datasets
-wind_fix <- function(df){
-wind <- df %>% 
-  select(station_name, date, hour, dir, speed) %>%  #column names
-  mutate(date = as.Date(as.character(date)),
-         hour = as.numeric(as.character(hour)), 
-         dir = as.numeric(as.character(dir)),
-         speed = as.numeric(as.character(speed)))
-}
-# RWS: We can see when we force the values to be numeric that there are some non-numeric values in the base data
-wind_1 <- wind_fix(df = wind_1)
-wind_2 <- wind_fix(df = wind_2)
-wind_3 <- wind_fix(df = wind_3)
-
-## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
-
-renaming_sites_1 <- function(df) {
-  sites <- df %>%
-      # RWS: An alternative way to replace values without having to use multiple ifelse() statements
-      mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
-                                    station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
-                                    station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-                                    station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
-                                    station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
-                                    station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
-  return(sites)
-}
-
-wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
-wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
-wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
-wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-save(wind_data, file = "Data/wind_data.RData")
-
-# # RWS: You should never be removing values by specific call like this
-#   # Rather you should be able to use some sort of conditional to screen out unwanted values
-# wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
-# renaming_sites_2 <- function(df) {
-#   sites <- df %>%
-#     mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
-#                            ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
-#                                 ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
-#                                        ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
-#   return(sites)
-# }
-# 
-# wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
-# wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ] 
-# renaming_sites_3 <- function(df) {
-#   sites <- df %>%
-#     mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
-#                            ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
-#                            ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
-#   return(sites)
-# }
-# 
-# wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
-# wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ] 
-### CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
-# wind_3_HoutBay <- wind_3 %>% 
-#   filter(station_name == "CAPE TOWN SLANGKOP") %>% 
-#   mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
-# wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-# wind_data <- wind_data %>% 
-#   na.omit()
-# 
-# wind_data <- wind_data %>% 
-#   # mutate(date = as.Date(date)) %>%  
-#   group_by(date, hour, temp_sites) 
-# 
-# # write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
-# # save(wind_data, file = "Data_P1/wind_data.RData")
-
-# load("Data_P1/insitu_wind.RData")
-# 
-# test1 <- insitu_wind %>% 
-#   mutate(lat = case_when(site == "Port Nolloth" ~ "-29.25241",
-#                                     site == "Saldanha Bay" ~ "-33.01054",
-#                                     site  == "Yzerfontein" ~ "-33.36068",
-#                                     site  == "Kommetjie" ~ "-34.13667",
-#                                     site  == "Lamberts Bay" ~ "-32.09185",
-#                                     site  == "Sea Point" ~ "-33.91847"))
-# 
-# test <- test1 %>% 
-#   mutate(lon = case_when(site == "Port Nolloth" ~ "16.86710",
-#                                     site == "Saldanha Bay" ~ "17.95063",
-#                                     site  == "Yzerfontein" ~ "18.15731",
-#                                     site  == "Kommetjie" ~ "18.32674",
-#                                     site  == "Lamberts Bay" ~ "18.30262",
-#                                     site  == "Sea Point" ~ "18.38217"))
-# 
-# test_final <- test %>% 
-#   select(date,lat,lon,dir_circ, mean_speed) %>% 
-#   dplyr::rename(speed = mean_speed) %>% 
-#   dplyr::rename(dir = dir_circ)
-
-# Loading the wind data
-library(tidyverse)
-library(circular)
-
-# wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number())) # Dataset too large for Github but better data at line 638
-
-# wind_data <- read_csv("Data_P1/wind_data.csv")
+```{r load-winddata}
+wind_data <- read_csv("../Data_P1/wind_data.csv")
 selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
@@
 # Loading the subsetted wind data created from the text files above 
-load("Data/wind_daily.RData") 
+load("../Data/wind_daily.RData") 
 ```
 
-# Upwelling
+\subsubsection*{Upwelling Index}
 
@@
 Upwelling is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.
-
-# Upwelling indeces
-  # Determining upwelling index from wind data (SAWS)
-  # Index Equation from Fielding & Davis 1989 paper
 
@@
 With this data the upwelling index was determined.
+Inorder to use this formula I first determined the angle perpendicular to the coastline using the `coastR` function.
+Thereafter, I determined the upwelling index for each of the sites.
 
-# Steps
-  - Use the circular function
-  - Determine daily wind data obtained from 3hr intervals
+```{r upwellingIndex-application}
 
-```{r}
-# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
-# 
-# wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# # Slecting the important columns for each of the datasets
-# wind_fix <- function(df){
-# wind <- df %>%
-#   select(station_name, date, hour, dir, speed) %>%  #column names
-#   mutate(date = as.Date(as.character(date)),
-#          hour = as.numeric(as.character(hour)),
-#          dir = as.numeric(as.character(dir)),
-#          speed = as.numeric(as.character(speed)))
-# }
-# 
-# wind_1 <- wind_fix(df = wind_1)
-# wind_2 <- wind_fix(df = wind_2)
-# wind_3 <- wind_fix(df = wind_3)
-# 
-# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
-# 
-# renaming_sites_1 <- function(df) {
-#   sites <- df %>%
-#       mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
-#                                     station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
-#                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
-#                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
-#                                     station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
-#   return(sites)
-# }
-# 
-# 
-# wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
-# wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
-# wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
-# wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-# save(wind_data, file = "Data/wind_data.RData")
-# 
-# load("Data/wind_data.RData")
-# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
-# 
-# wind_daily <- wind_data %>%
-#   # ungroup() %>%
-#   # mutate(date = as.Date(date))%>% # Changing the date to date format
-#   dplyr::group_by(temp_sites, date) %>%
-#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2), # Direction using the circular function
-#                    mean_speed = round(mean(speed),2)) # Calculating the mean
-# save(wind_daily, file = "Data/wind_daily.RData")
-```
-
-# Load wind
-
-```{r}
-load("Data/wind_daily.RData") 
-wind_daily <- wind_daily %>% 
-  dplyr::rename(sites = temp_sites)
-load("Data/site_list_sub.Rdata")
-```
-
-# Determining the angle perpendicular to the coastline and plotting it
-
-```{r}
-# Coast R- function that calculates the angle of the site perpendicular to the coastline
-# Transect function in coastR package finds the angle away or along the coastline
 south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
 
-# Plotting the map
-world_map <- ggplot() + 
-  borders(fill = "grey40", colour = "black")
-
-# Create titles
-titles <- c("Shore-normal", "Islands")
-# Plotting function
-plot_sites <- function(site_list, buffer, title_choice, dist){
-
-  # Find the point 200 km from the site manually to pass to ggplot
-  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
-                                              b = site_list$heading, d = dist))
-
-  # Add the new coordinates tot he site list
-  site_list <- site_list %>% 
-    mutate(lon_dest = heading2$lon,
-           lat_dest = heading2$lat)
-
-  # Visualise
-  world_map +
-    geom_segment(data = site_list, colour = "red4", 
-                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
-    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
-    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
-    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
-                             max(site_list$lon + buffer)),
-                    ylim = c(min(site_list$lat - buffer), 
-                             max(site_list$lat + buffer))) +
-    labs(x = "", y = "", colour = "Site\norder") +
-    ggtitle(titles[title_choice])
-}
-south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
-south_africa_away
-```
-
-# Upwelling indeces using the above formula
-
-```{r}
+# # Plotting the map
+# world_map <- ggplot() + 
+#   borders(fill = "grey40", colour = "black")
+# 
+# # Create titles
+# titles <- c("Shore-normal", "Islands")
+# # Plotting function
+# plot_sites <- function(site_list, buffer, title_choice, dist){
+#   
+#   # Find the point 200 km from the site manually to pass to ggplot
+#   heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
+#                                               b = site_list$heading, d = dist))
+#   
+#   # Add the new coordinates to the site list
+#   site_list <- site_list %>% 
+#     mutate(lon_dest = heading2$lon,
+#            lat_dest = heading2$lat)
+# 
 UI_angle <- wind_daily %>%  # Making refeerence to the wind daily datta created above
@@
   dplyr::rename(t = date) #renaming date to t
-
-# save(UI_angle, file = "Data/UI_angle.RData")
-# index <- wind_UI %>%
-#   group_by(site) %>%
-#   select("date", "ui.saws") %>%
-#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-# wind_UI <- wind_daily %>% 
-#   mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
-#   drop_na
-# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
 ```
 
-# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.
+Here I make use of the `exceedence` function within the heatwaveR package.
+This was primarily done to determine the consecutive number of days at or above what the UI value is meant to be.
+The scipt however may be found in my [GitHub](https://github.com/AmierohAbrahams/Upwelling_MCS) repository, and the full code can be run in its entiety.
 
-```{r}
+```{r upwelling_thresholds}
 # Loading the insitu temperature data along the wc
@@
 
-# Loading in all the data created using the code bellow
+Thereafter, I downloaded and then create a function which allows me to exctracted remotely sensed sea surface temperatures.Consequently we have a dataframe with a date, temperature, lat and lon columns as seen below.
 
-```{r}
-load("Data/site_list_sub.Rdata")
-load("Data/SACTN_US.RData")
-load("Data/site_pixels.RData") # 5 decimal places
-load("Data/OISST.RData") # 2 decimal places
+```{r load_data2, include=TRUE}
+load("../Data/site_list_sub.Rdata")
+load(".../Data/SACTN_US.RData")
+load("../Data/site_pixels.RData") # 5 decimal places
+load("../Data/OISST.RData") # 2 decimal places
 OISST <- BC_avhrr_only_v2_Document_Document 
 rm(BC_avhrr_only_v2_Document_Document ); gc()
-load("Data/CMC.RData") # 1decimal places
+load("../Data/CMC.RData") # 1decimal places
 ```
 
- Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
-# Is this upwelling event seen throughout the different distances from the coastline
-# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
+The code below uses the extracted SST data and matches this to the lat and lon coordinates obtained at the earlier calculation in order to determine the temperature at a particular distance from the coastline.
 
-## Function
-Determining the temperatures at the various distances from the coastline
-
-```{r}
+```{r matched_distance_coastline}
 # # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
@@
 
-# PLotting events
-
-Loading the data
-```{r}
+```{r loadfiles3, include=TRUE}
 load("Data/OISST_upwell_base.RData")
@@
 
-To accurately compare 
-  Start date CMC start date 1991-09-02 end date 2017-03-15
-  Start date OISST start date 1981-09- 01  and end date 2018-09-28
-  Start date SACTN start date 1973 and end date 2017-12-16
+\subsubsection*{Analyses}
 
@@
 
-# ANOVA
-- Relationship between duration, year and distance as a function of the mean intensity. 
-```{r}
+Here I do a three way ANOVA analayses.
+This allows me to compare one variable in two or more groups taking into account the variability of other variables.
+This analysis of covariance is used to analyse the relationship between the mean intensity  as a function of each site, year and duration.
+
+```{r ANOVA analyses}
 options(scipen = 999)
@@
 
-# Wind plots
+Since the analysis are largely dependant on wind data, I now continue with the analyses to show the various wind patterns that existed during this period.
+Here I created a function to plot the wind data by plotting wind rose diagrams.
 
-```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
+```{r Waverose diagram, fig.keep = "none"}
 library(ggradar)
@@
 library(tibble)
-load("Data/UI_angle.RData")
+load("../Data/UI_angle.RData")
 
-# Detects NA values, no NA values are present
-# Plotting wind variables
-source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function
+source("../Functions/wind.rose.R")
 
@@
 
-# Lolli plot
+The code below produces plots for each of the different datasets representing the cumulative intensity, peak date of event and duration of the detected event.
 
-```{r}
+```{r loliplots, fig.keep="none"}
 plot_loli_func <- function(df){
@@
 ```
+
+\subsection*{References}

### 2019-10-08 Write_up/Paper1_writeup.Rmd
updated

@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the OISST dataset.
+\caption{The cumulative intensity (°C x days) for the sites located within the Benguela upwelling system using the OISST dataset.
 The site column gives the name of the site.
-The distance column gives the distance from the coastline as seen in Fig 1.
-The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
+The distance column gives the distance from the coastline as seen in Fig 1.} \\
 
@@
 
-Lamberts Bay & 10000 & 5.47875 \\
-Lamberts Bay & 20000 & 5.47875 \\
-Lamberts Bay & 30000 & 7.67025 \\
-Lamberts Bay & 40000 & 1.09575 \\
-Lamberts Bay & 50000 & -8.76600 \\
-Port Nolloth & 10000 & 0.36525  \\
-Port Nolloth & 20000 & -3.28725 \\
-Port Nolloth & 30000 & -2.92200 \\
-Port Nolloth & 40000 & -3.65250 \\
-Port Nolloth & 50000 & -3.28725 \\
-Saldanha Bay & 10000 & -0.73050 \\
-Saldanha Bay & 20000 & -0.36525 \\
-Saldanha Bay & 30000 & -4.38300 \\
-Saldanha Bay & 40000 & -10.22700 \\
-Saldanha Bay & 50000 & -12.41850 \\
-Sea Point & 10000 & -8.76600  \\
-Sea Point & 20000 & -3.65250 \\
-Sea Point & 30000 & -1.46100 \\
-Sea Point & 40000 & -9.49650 \\
-Sea Point & 50000 & -18.62775 \\
+Lamberts Bay & 10000 & -1.83 \\
+Lamberts Bay & 20000 & -1.83 \\
+Lamberts Bay & 30000 & -1.83 \\
+Lamberts Bay & 40000 & -0.73 \\
+Lamberts Bay & 50000 & 0.73 \\
+Port Nolloth & 10000 & -1.10 \\
+Port Nolloth & 20000 & -0.37 \\
+Port Nolloth & 30000 & 0.00 \\
+Port Nolloth & 40000 & 0.37 \\
+Port Nolloth & 50000 & 0.37 \\
+Saldanha Bay & 10000 & -2.19 \\
+Saldanha Bay & 20000 & -1.83 \\
+Saldanha Bay & 30000 & -0.73 \\
+Saldanha Bay & 40000 & 0.73 \\
+Saldanha Bay & 50000 & 1.10\\
+Sea Point & 10000 & -2.56 \\
+Sea Point & 20000 & -3.29 \\
+Sea Point & 30000 & -2.56 \\
+Sea Point & 40000 & 0.00\\
+Sea Point & 50000 & 1.46 \\
 
@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the CMC dataset.
+\caption{The cumulative intensity (°C xday) for the sites located within the Benguela upwelling system using the CMC dataset.
 The site column gives the name of the site.
-The distance column gives the distance from the coastline as seen in Fig 1.
-The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
+The distance column gives the distance from the coastline as seen in Fig 1.} \\
 
@@
 
-Lamberts Bay & 10000 & -26.29800 \\
-Lamberts Bay & 20000 & -25.56750 \\
-Lamberts Bay & 30000 & -23.74125 \\
-Lamberts Bay & 40000 & -31.41150 \\
-Lamberts Bay & 50000 & -31.77675 \\
-Port Nolloth & 10000 & -15.34050 \\
-Port Nolloth & 20000 & -16.43625 \\
-Port Nolloth & 30000 & -15.70575 \\
-Port Nolloth & 40000 & -20.45400 \\
-Port Nolloth & 50000 & -24.83700 \\
-Saldanha Bay & 10000 & -29.58525 \\
-Saldanha Bay & 20000 & -36.52500 \\
-Saldanha Bay & 30000 & -42.73425 \\
-Saldanha Bay & 40000 & -46.38675 \\
-Saldanha Bay & 50000 & -45.29100 \\
-Sea Point & 10000 & -28.48950 \\
-Sea Point & 20000 & -30.31575 \\
-Sea Point & 30000 & -36.89025 \\
-Sea Point & 40000 & -38.35125 \\
-Sea Point & 50000 & -37.25550 \\
+Lamberts Bay & 10000 & 3.28725 \\
+Lamberts Bay & 20000 & 2.92200 \\
+Lamberts Bay & 30000 & 2.92200 \\
+Lamberts Bay & 40000 & 2.55675 \\
+Lamberts Bay & 50000 & 1.82625 \\
+Port Nolloth & 10000 & 2.55675 \\
+Port Nolloth & 20000 & 2.55675 \\
+Port Nolloth & 30000 & 1.46100 \\
+Port Nolloth & 40000 & 1.09575 \\
+Port Nolloth & 50000 & 1.09575 \\
+Saldanha Bay & 10000 & 3.28725 \\
+Saldanha Bay & 20000 & 4.38300 \\
+Saldanha Bay & 30000 & 4.01775 \\
+Saldanha Bay & 40000 & 2.92200 \\
+Saldanha Bay & 50000 & 2.92200 \\
+Sea Point & 10000 & 3.28725 \\
+Sea Point & 20000 & 3.65250 \\
+Sea Point & 30000 & 4.01775 \\
+Sea Point & 40000 & 3.28725 \\
+Sea Point & 50000 & 2.19150 \\
 
@@
 
-Lamberts Bay & -1.46100 \\
-Port Nolloth & -3.65250 \\
-Saldanha Bay & -2.55675 \\
-Sea Point & -10.95750 \\
+Lamberts Bay & 1.09575 \\
+Port Nolloth & 1.09575 \\
+Saldanha Bay & 0.36525 \\
+Sea Point & 1.09575 \\
 

### 2019-10-08 Write_up/Paper1_writeup.Rmd
updated

@@
 
-Lamberts Bay & 10000 & 3.28725 \\
-Lamberts Bay & 20000 & 2.92200 \\
-Lamberts Bay & 30000 & 2.92200 \\
-Lamberts Bay & 40000 & 2.55675 \\
-Lamberts Bay & 50000 & 1.82625 \\
-Port Nolloth & 10000 & 2.55675 \\
-Port Nolloth & 20000 & 2.55675 \\
-Port Nolloth & 30000 & 1.46100 \\
-Port Nolloth & 40000 & 1.09575 \\
-Port Nolloth & 50000 & 1.09575 \\
-Saldanha Bay & 10000 & 3.28725 \\
-Saldanha Bay & 20000 & 4.38300 \\
-Saldanha Bay & 30000 & 4.01775 \\
-Saldanha Bay & 40000 & 2.92200 \\
-Saldanha Bay & 50000 & 2.92200 \\
-Sea Point & 10000 & 3.28725 \\
-Sea Point & 20000 & 3.65250 \\
-Sea Point & 30000 & 4.01775 \\
-Sea Point & 40000 & 3.28725 \\
-Sea Point & 50000 & 2.19150 \\
+Lamberts Bay & 10000 & 3.29 \\
+Lamberts Bay & 20000 & 2.92 \\
+Lamberts Bay & 30000 & 2.92 \\
+Lamberts Bay & 40000 & 2.56 \\
+Lamberts Bay & 50000 & 1.83 \\
+Port Nolloth & 10000 & 2.56 \\
+Port Nolloth & 20000 & 2.56 \\
+Port Nolloth & 30000 & 1.46 \\
+Port Nolloth & 40000 & 1.10 \\
+Port Nolloth & 50000 & 1.10 \\
+Saldanha Bay & 10000 & 3.29 \\
+Saldanha Bay & 20000 & 4.38 \\
+Saldanha Bay & 30000 & 4.02 \\
+Saldanha Bay & 40000 & 2.92 \\
+Saldanha Bay & 50000 & 2.92\\
+Sea Point & 10000 & 3.29 \\
+Sea Point & 20000 & 3.65 \\
+Sea Point & 30000 & 4.02 \\
+Sea Point & 40000 & 3.29 \\
+Sea Point & 50000 & 2.19 \\
 
@@
 \begin{longtable}{|r|r|p{2cm}|}
-\caption{The absolute cumulative intensity (°C.day) for the sites located within the Benguela upwelling system using the SACTN dataset.
+\caption{The cumulative intensity (°C x day) for the sites located within the Benguela upwelling system using the SACTN dataset.
 The site column gives the name of the site.
-Tempertures were collected at the coastline and not at different distances from the coastline.
-The cumulative intensity expects as absolute magnitudes rather than relative seasonal climatology} \\
+Tempertures were collected at the coastline and not at different distances from the coastline.} \\
 
@@
 
-Lamberts Bay & 1.09575 \\
-Port Nolloth & 1.09575 \\
-Saldanha Bay & 0.36525 \\
-Sea Point & 1.09575 \\
+Lamberts Bay & 1.10 \\
+Port Nolloth & 1.10 \\
+Saldanha Bay & 0.37 \\
+Sea Point & 1.10 \\
 

### 2019-10-09 Write_up/Paper1_writeup.Rmd
updated

@@
 
-Sea Point showcased the most intense upwelling event at 10000 km from the coastline during the year 2013, lasting for 21 days (Refer to figure, do this throughout).
-The majority of upwelling events in Sea Pont occurred around 10000 km away from the coastline lasted for less than five days, and had an approximate cumulative intensity ranging between of -10.00deg.
+Sea Point showcased the most intense upwelling event at 10 km from the coastline during the year 2013, lasting for 21 days (Refer to figure, do this throughout).
+The majority of upwelling events in Sea Pont occurred around 10 km away from the coastline lasted for less than five days, and had an approximate cumulative intensity ranging between of -10.00deg.
 C x days and 12.00 deg.
 C x days.
-Conversely, at shorter distances like 3000 km from the coastline several upwelling events lasted for longer than 15 days with a cumulative intensity ranging between -30deg.
+Conversely, at shorter distances like 30 km from the coastline several upwelling events lasted for longer than 15 days with a cumulative intensity ranging between -30deg.
 C x days and -20deg.
 C x days.
-At a distance of 30000 km both dominant events appear at a distance of 10000 km are present along with several other additional events (This doesn’t make sense).
-At a distance of 40000 km, less intense events occurred.
+At a distance of 30 km both dominant events appear at a distance of 10 km are present along with several other additional events (This doesn’t make sense).
+At a distance of 40 km, less intense events occurred.
 At a distance of 50000 km from the coastline, several events appear to have lasted for approximately 2 – 3 days with the most intense events still occurring in 2015.Events within Saldanha Bay typically lasted for approximately three days but varying at different distances.
-At a distance of 10000 km one major event occurred that lasted for approximately 27 days with a cumulative intensity of -61.70deg.
+At a distance of 10 km one major event occurred that lasted for approximately 27 days with a cumulative intensity of -61.70deg.
 C x days.
-This event was also present at a distance of 20000 km from the coastline, albeit at a lesser intensity, but was not present at 30000km, 40000km and 50000 km respectively.
-Events occurring within 30000 km and 40000 km from the coastline lasting for approximately 10 days in most cases.
-However, the majority of events were observed at a distance of 10000 km and 20000 km from the coastline.
-Fewer events occurred at a distance of 50000 km, with most of the events occurring at this distance having lasted for less than five days.
+This event was also present at a distance of 20 km from the coastline, albeit at a lesser intensity, but was not present at 30km, 40km and 50km respectively.
+Events occurring within 30 km and 40 km from the coastline lasting for approximately 10 days in most cases.
+However, the majority of events were observed at a distance of 10 km and 20 km from the coastline.
+Fewer events occurred at a distance of 50 km, with most of the events occurring at this distance having lasted for less than five days.
 
@@
 C x days.
-This event persisted at 20000 km and 30000 km with cumulative intensities of -21.09deg.
+This event persisted at 20 km and 30000 km with cumulative intensities of -21.09deg.
 C x days and -14.33 deg.
 C x days respectively.
-Here, most upwelling events were detected at 10000km and 20000km respectively.
-The least number of events were detected at 50 000 km.
+Here, most upwelling events were detected at 10km and 20km respectively.
+The least number of events were detected at 50 km.
 The least intense upwelling event lasted for only 3 days and had a cumulative intensity of -5.45deg.
@@
 Events occurring within Lamberts Bay lasted for approximately less than five days.
-Three major events were detected at 10000 km and 20000 km away from the coastline respectively.
-At 30000 km an event with a longer duration occurs in 2010, that was not present at a distance of 10000 km and 20000 km from the coastline.
-At a distance of 40000 km no major events were detected.
-However, at a distance of 50000 km a major upwelling event is detected lasting for longer than 20 days.
-Fewer events were however detected at a distance of 50000 km from the coastline.
+Three major events were detected at 10 km and 20 km away from the coastline respectively.
+At 30 km an event with a longer duration occurs in 2010, that was not present at a distance of 10 km and 20 km from the coastline.
+At a distance of 40 km no major events were detected.
+However, at a distance of 50 km a major upwelling event is detected lasting for longer than 20 days.
+Fewer events were however detected at a distance of 50 km from the coastline.
 
@@
 The most intense event occurred in 2010, and I found that intensifies appeared to increase with distance from the coastline.
-The fewest number of events occurred 50000 km away from the coastline.
+The fewest number of events occurred 50 km away from the coastline.
 At Saldanha Bay, the duration of upwelling events mostly ranged between three to four days.
-Similar intensity patterns occurred at 10000 km, 20000 km and 30000 km from the coastline respectively.
-At 40000 km an event representing a cumulative intensity of -28.64deg.
+Similar intensity patterns occurred at 10 km, 20 km and 30km from the coastline respectively.
+At 40 km an event representing a cumulative intensity of -28.64deg.
 C x days for was detected for 20 days within 2010.
@@
 Similar patterns were found within Lamberts Bay.
-At 30000 km a major event was detected that lasted for approximately 19 days, but most events here lasted for less than five days.
-At a distance of 40000 km and 50000 km from the coastline similar patterns were observed, with most events being detected at 40000 km and 50000 km respectively.
+At 30 km a major event was detected that lasted for approximately 19 days, but most events here lasted for less than five days.
+At a distance of 40 km and 50 km from the coastline similar patterns were observed, with most events being detected at 40 km and 50 km respectively.
 
@@
 \toprule
-Site & distance & regression slope  \\
+Site & distance (km) & regression slope  \\
 \midrule
@@
 
-Lamberts Bay & 10000 & -1.83 \\
-Lamberts Bay & 20000 & -1.83 \\
-Lamberts Bay & 30000 & -1.83 \\
-Lamberts Bay & 40000 & -0.73 \\
-Lamberts Bay & 50000 & 0.73 \\
-Port Nolloth & 10000 & -1.10 \\
-Port Nolloth & 20000 & -0.37 \\
-Port Nolloth & 30000 & 0.00 \\
-Port Nolloth & 40000 & 0.37 \\
-Port Nolloth & 50000 & 0.37 \\
-Saldanha Bay & 10000 & -2.19 \\
-Saldanha Bay & 20000 & -1.83 \\
-Saldanha Bay & 30000 & -0.73 \\
-Saldanha Bay & 40000 & 0.73 \\
-Saldanha Bay & 50000 & 1.10\\
-Sea Point & 10000 & -2.56 \\
-Sea Point & 20000 & -3.29 \\
-Sea Point & 30000 & -2.56 \\
-Sea Point & 40000 & 0.00\\
-Sea Point & 50000 & 1.46 \\
+Lamberts Bay & 10 & -1.83 \\
+Lamberts Bay & 20 & -1.83 \\
+Lamberts Bay & 30 & -1.83 \\
+Lamberts Bay & 40 & -0.73 \\
+Lamberts Bay & 50 & 0.73 \\
+Port Nolloth & 10 & -1.10 \\
+Port Nolloth & 20 & -0.37 \\
+Port Nolloth & 30 & 0.00 \\
+Port Nolloth & 40 & 0.37 \\
+Port Nolloth & 50 & 0.37 \\
+Saldanha Bay & 10 & -2.19 \\
+Saldanha Bay & 20 & -1.83 \\
+Saldanha Bay & 30 & -0.73 \\
+Saldanha Bay & 40 & 0.73 \\
+Saldanha Bay & 50 & 1.10\\
+Sea Point & 10 & -2.56 \\
+Sea Point & 20 & -3.29 \\
+Sea Point & 30 & -2.56 \\
+Sea Point & 40 & 0.00\\
+Sea Point & 50 & 1.46 \\
 
@@
 
-Lamberts Bay & 10000 & 3.29 \\
-Lamberts Bay & 20000 & 2.92 \\
-Lamberts Bay & 30000 & 2.92 \\
-Lamberts Bay & 40000 & 2.56 \\
-Lamberts Bay & 50000 & 1.83 \\
-Port Nolloth & 10000 & 2.56 \\
-Port Nolloth & 20000 & 2.56 \\
-Port Nolloth & 30000 & 1.46 \\
-Port Nolloth & 40000 & 1.10 \\
-Port Nolloth & 50000 & 1.10 \\
-Saldanha Bay & 10000 & 3.29 \\
-Saldanha Bay & 20000 & 4.38 \\
-Saldanha Bay & 30000 & 4.02 \\
-Saldanha Bay & 40000 & 2.92 \\
-Saldanha Bay & 50000 & 2.92\\
-Sea Point & 10000 & 3.29 \\
-Sea Point & 20000 & 3.65 \\
-Sea Point & 30000 & 4.02 \\
-Sea Point & 40000 & 3.29 \\
-Sea Point & 50000 & 2.19 \\
+Lamberts Bay & 10 & 3.29 \\
+Lamberts Bay & 20 & 2.92 \\
+Lamberts Bay & 30 & 2.92 \\
+Lamberts Bay & 40 & 2.56 \\
+Lamberts Bay & 50 & 1.83 \\
+Port Nolloth & 10 & 2.56 \\
+Port Nolloth & 20 & 2.56 \\
+Port Nolloth & 30 & 1.46 \\
+Port Nolloth & 40 & 1.10 \\
+Port Nolloth & 50 & 1.10 \\
+Saldanha Bay & 10 & 3.29 \\
+Saldanha Bay & 20 & 4.38 \\
+Saldanha Bay & 30 & 4.02 \\
+Saldanha Bay & 40 & 2.92 \\
+Saldanha Bay & 50 & 2.92\\
+Sea Point & 10 & 3.29 \\
+Sea Point & 20 & 3.65 \\
+Sea Point & 30 & 4.02 \\
+Sea Point & 40 & 3.29 \\
+Sea Point & 50 & 2.19 \\
 

### 2019-10-09 Write_up/Abstract_edits/Paper1_abstract.Rmd
Abstract

@@
 ## Abstract
+Abstract
+
+The South African coastline is comprised of three distinct coastal regions, each varying in average temperaturedue to the presence of two distinct ocean currents; namely the Agulhas and Benguela currents, an eastern boundry upwelling system (EBUS).
+This study was designed to see if the same upwelling patterns are visible at different distances from the coastline and in diffeent datasets.
+Analyses of temperature time series data from 4 sites within the west coast of South Africa were conducted to examine variation in intensity occuring at different upwelling events at different distances from the coastline.
+This study futher made use of wind data to determine the upwelling index for each of the sites.
+By using the ANOVA analyses, it was possible to determine the relationship between intensity for each of the sites at variable distnces.
+Results showed high intensity of upwelling events in the OISST dataset but less high intensities in the CMC dataset.
+Similar upwelling patterns are present in the in situ and OISST dataset at a distace of 10km from the coastline.
+At a distance of 50km from the coastline less upwelling events are detected in both remotely sensed datasets.

### 2019-10-09 SST_patterns.Rmd
update

@@
   facet_wrap(~distance, ncol = 2) +
-  xlab("Date") + ylab("Event duration [days]") 
+  scale_y_continuous(breaks = c(5,10,15,20,25,30),
+                     limits = c(1,30)) +
+  xlab("Date (years)") + ylab("Event duration [days]") 
 }
@@
   ggtitle("Sea Point")
+
 OISST_SB_pl <- OISST_SB_pl +
@@
   facet_wrap(~site, ncol = 2) +
-  xlab("Date") + ylab("Event duration [days]") 
+  scale_y_continuous(breaks = c(5,10,15,20,25,30),
+                     limits = c(1,30)) +
+  xlab("Date (years)") + ylab("Event duration [days]") 
 }

### 2019-10-09 Write_up/Paper1_writeup.Rmd
updated push
update

@@
 ## Abstract
+
+The South African coastline is comprised of three distinct coastal regions, each varying in average temperature due to the presence of two distinct ocean currents; namely the Agulhas and Benguela currents, the latter an eastern boundary upwelling system (EBUS).
+This study was designed to examine whether same upwelling patterns were discernible at varying distances from the coastline and if this was apparent from different datasets.
+Analyses of temperature time series data from 4 sites within the west coast of South Africa were conducted to examine the variation in intensity occurring at different upwelling events at different distances from the coastline.
+This study further made use of wind data to determine the upwelling index for each of the sites.
+By using the ANOVA analyses, it was possible to determine the relationship between intensity for each of the sites at variable distances.
+Results showed high intensity of upwelling events in the OISST dataset but less high intensities in the CMC dataset.
+Similar upwelling patterns are present in the in situ and OISST dataset at a distance of 10km from the coastline.
+However, it should be notes that at distances of 50km or more from the coastline, fewer upwelling events are detected in both remotely sensed datasets.
 
@@
 
-The linear regression analysis showed that the absolute cumulative intensity among all variables within each dataset showed the most variation.
+The linear regression analysis showed that the cumulative intensity among all variables within each dataset showed the most variation.
 The cumulative intensity acts as absolute magnitude rather than relative seasonal climatology.

### 2019-10-09 Write_up/Paper1_writeup.Rmd
updated

@@
 Sites compared within the OISST datasets showed no difference in the durations of the events at different distances for each of the sites from the coastline.
-However, there is a difference in the mean intensity of events between sites (*F*=5.74, *SS*=5.2, *p*<0.001).
-Additionally, there were also significant differences between the mean intensities of events at different distances from the coastline (*F*=260.17, *SS*=78.6, *p*<0.001), and between the durations in events (*F*=51.39, *SS*= 15.53, *p*<0.001) and between sites.
+However, there is a difference in the mean intensity of events between sites (*F*=5.74, *SS*=5.20, *p*<0.001).
+Additionally, there were also significant differences between the mean intensities of events at different distances from the coastline (*F*=260.17, *SS*=78.60, *p*<0.001), and between the durations in events (*F*=51.39, *SS*= 15.53, *p*<0.001) and between sites.
 Similar comparisons using the CMC dataset showed that there was no significant difference in the duration of events at different distances from the coastline for each of the sites (Three-Way ANOVA: *F*=1.20, *SS*=0.41, *p*>0.05).
-However, there were significant differences in the mean intensities of events at different distances (*F*=471.63, *SS*=162.0, *p*<0.05), between the durations of events (*F*= 38.36, *SS*=13.2, *p*<0.001), and between sites (*F*=17.64, *SS*=18.2, *p*<0.002).
-I also found that there was no significant difference between the durations of upwelling across all sites (*F*=0.22, *SS*=0.2, *p*<0.87).
+However, there were significant differences in the mean intensities of events at different distances (*F*=471.63, *SS*=162.00, *p*<0.05), between the durations of events (*F*= 38.36, *SS*=13.20, *p*<0.001), and between sites (*F*=17.64, *SS*=18.20, *p*<0.002).
+I also found that there was no significant difference between the durations of upwelling across all sites (*F*=0.22, *SS*=0.20, *p*<0.87).
 When looking at the in situ collected dataset, data collected at the coastline, we find no difference in the durations of events at across sites (*F*=2.05, *SS*=1.84, *p*=0.11) but there was a difference in the mean intensities of those events (*F*=2.05, *SS*=0.61, *p*=0.44).

### 2019-10-09 Write_up/Paper1_writeup.Rmd
Updated

@@
 ## Discussion
+
+This study aimed to investigate patterns occurring within upwelling events at various different spatial intervals from the coastline, as well as to test whether the same patterns are discernible in the various datasets.
+As seawater temperature is known to have a large influence on species distribution and productivity (Bolton 2010; Smit et al.
+2013), thus it is important to understand not only how temperatures vary at different distances from the coastline but what contributing factors to ecosystem maintenance the intensity of these events have across these distance intervals from the coastline.
+Of particular interest to this study; the Benguela upwelling system, which is distinct in that it is one of four most productive regions in the world.
+
+In this way, the proximal causes of environmental change might be understood in terms of its relevance in affecting these important regions.
+Sinnett and Feddesen (2014) expanded on this concept and illustrated that various environmental factors such as solar radiation, air temperature, humidity and wave energy are responsible for temperature variation within coastal regions.
+Here, however, we examine if the same upwelling patterns exist between the local and broader-scale SST fields given that various factors influence temperatures at different scales.
+
+Results gathered during the course of this study clearly illustrate that upwelling events detected in different ocean sea surface temperature datasets showed variations with regards to duration and intensities across different time intervals.
+Noticeably, some patterns were dominant and persistent throughout the datasets; however, it was also observed that fewer upwelling events occurred at distances furthest away from the coastline.
+This was correlated in satellite datasets, where, often at distances of 10km and 20km from the coastline, similar upwelling events were apparent.
+This difference appears to be related to nature and the variability of physical oceanographic processes at both the broader and local scales; however, the potential involvement of other processes cannot be ruled out.
+These findings illustrate how dis-similar satellite data is from in situ collected datasets.
+
+Often the differences noted between the in situ and satellite derived data were striking.
+Initially, it was anticipated a large degree of coupling within the datasets across the local and broader scales would be observed; instead the OISST satellite derived data yielded a more intense representation of events as compared to the SACTN collected data.
+The CMC dataset showed that events were moderate throughout the various sites and matched more closely with the in situ data observed.
+Information of events obtained from the OISST satellite data illustrated that these phenomena occurred more frequently and for longer durations.
+
+Furthermore, it is now believed that the larger intensities witnessed within the in situ dataset may be in actuality an artefact of the inability of remotely sensed data to record the maximum and minimum temperatures that in situ instruments are capable of detecting (Smale and Wernberg, 2009).
+The in situ collected data shows very few events detected at Port Nolloth, similarly, the CMC and OISST dataset also shows fewer upwelling events recorded at a distance of 10km within this region.
+Upwelling events recorded in Sea Point match closely with the events detected in the OISST dataset at a distance of 10km.
+
+Notably the CMC dataset showed large degrees of variance within detected events, however no major events were recorded.
+It is observed that the SACTN in situ collected temperature data shows similar patterns to the OISST dataset at a distance of 10km; with Saldanha Bay showing the highest frequency in upwelling events detected throughout the datasets.
+However, the in situ and OISST dataset at a distance of 10km once again showing similar detections as those recorded by the CMC dataset at the same distance from the coastline, and thus does not accurately represent these detected events.
+At a distance of 30km – 50km the in situ detected events differed greatly as those compared to what was observed within the satellite derived temperature data.
+
+It should be noted that previous, similar studies have suggested that a longer, sampling time series, comprising of more data would have resulted in greater accuracy with regards to detecting subtle changes in temperature differences obtained from variable coastal regions such as the South African coastline.
+Schlegel and Smit (2016) suggested that having a time series of greater than 30 years for sites with a low variance results in increased ability to more accurately detect changes.
+They showed that high quality time series datasets have frequent measurements, with minimal missing (NA) values present.
+Furthermore, low quality datasets (i.e.
+those with more than 5% of NA values) have a higher chance of detecting variation where none exists.
+
+We would be amiss to point out some weaknesses inherent in the SACTN dataset. Temperature
+measurement along the South African coastline started in the 1970s, and has been inconsistent in
+terms of the instruments used, the modifications to styles and methods, calibration, as well as site locations themselves (Smit et al.
+2013; Schlegel and Smit 2016).
+Additionally, older records within some datasets, such as the SACTN dataset, may have been lost or are unreliable since the metadata for these are no longer available.
+These metadata are essential as their absence prevents the understanding of the influence regarding the utilised instruments on temperature recordings.
+It can thus be speculated that some measurements may not accurately represent well-validated temperature data due to the instrumentation used; however, we mitigated these effects by selecting only the most reliable and longest duration data, which other current and on-going studies, such as that by Williamson et al.
+have suggested are most suitable for the problem at hand.
+
+Data in the SACTN are compiled from multiple sources, in addition to being acquired by various means, thus creating considerable heterogeneity; however, efforts are
+currently underway to homogenise these datasets (Williamson et al., in progress). The underwater
+temperature recorders (UTRs) used to collect data in the SACTN dataset expressed a lower number of NA values as compared to the data collected with hand thermometers.
+As such, this may have influenced the overall time series dataset (Schlegel and Smit, 2016).
+The level of precision at which data were collected also influenced the length of the time series needed.
+Time series in which temperature is collected at a precision of 0.5°C may require another 24 months of recordings to precisely detect long term variation (Schlegel and Smit, 2016).
+The average length of the thermometer time series component of the SACTN dataset was 346 months whereas the average length of UTR time series was less than half of that.
+With the extent of these differences in length being so severe, even once correcting for potential negative effects on the measurement precision of the thermometer collected time series, it was clear that thermometer data were more useful than that of UTRs.
+These influences may be expected to affect trend estimates of change derived from the data, however it is not anticipated that they have affected estimates of the variability in time and between sites in the way that they were used in this current study.
+
+Irrespective, satellite-acquired SST records are useful to modern marine scientists.
+These data are often used to model and predict a wide range of oceanic and biological processes within open ocean ecosystems and habitats; however, these have only recently been used to study temperature variations influencing benthic organisms (Pearce, Faskel, and Hyndes, 2006).
+SST data acquired by satellites deviate from coastal in situ collected seawater temperatures for a variety of reasons (Smit et al.
+2013).
+The SST data presented the broader-scale situation along the coast, therefore utilised here as a forcing boundary hypothesised here as influencing coastal temperatures.
+Analysis shows that the offshore SSTs (various sources) at a distance between 30km and 50km often do not match the detected events present within the in situ dataset along the west coast of South African.
+
+EBUS, such at the Benguela upwelling system are known to be highly variable across a wide spectrum of spatial and temporal scales, in both the forcing and response of the system.
+In some cases, a climate event can force large, although temporary changes in the overall productivity and/or community composition; these effects were seen during the large 1997-98 El Niño event and the delayed upwelling in 2005 in the California Current.
+EBUS are therefore be characterized as both resilient (quick recovery from disturbances) and robust (maintenance of ecosystem function and relatively high productivity) to natural climate variability, maintaining an abundance of species of high commercial and conservation value.
+However, the scale of future responses to anthropogenic climate change may be beyond the historical scales of variability, thus making forecasting the response of EBUS to climate change difficult.
+
+The importance of understanding the contributions of variation of upwelling on a local and broad scale relate to how productive these regions are but also indicated that upwelling found at a distance of 50km does not imply that the same upwelling events are present within the nearshore.
+Thus in order to detect coastal upwelling it is better to make use of coastal data.
+Within marine environments, coastal temperature variation allows for a varied spectrum in the spatial arrangements of marine biodiversity.
+Whilst wind and wave action may not be directly affecting ocean temperatures, Blamey and Branch (2009) have found that wave action has a profound influence on species distributions along the coastline.
+The presence or absence of marine species are determined by a variety of factors and whilst those factors may not be influencing each other as was the case here, they do however, collectively play an important role in affecting the marine life of the South African coastline; therefore, identifying those roles can aid in improving our understanding of nearshore dynamics, thereby providing greater knowledge base to be used for conservation efforts.

### 2019-10-09 Write_up/Paper1_writeup.Rmd
updated

@@
 The 3000km South African coastline is bordered by the Agulhas current along the east coast and the Benguela current (an EBUS) along the west coast.
-These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@Mead; @schlegel].
+These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@Mead; @schlegel; @schlegel2016].
 The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region [@jarred; @hutchings].
@@
 
-knitr::include_graphics("../Figures/OISST.png")
+knitr::include_graphics("../Figures/OISST.pdf")
 ```
@@
 
-knitr::include_graphics("../Figures/SACTN.png")
+knitr::include_graphics("../Figures/SACTN.pdf")
 ```
@@
 
-knitr::include_graphics("../Figures/CMC.png")
+knitr::include_graphics("../Figures/CMC.pdf")
 ```
@@
 This study aimed to investigate patterns occurring within upwelling events at various different spatial intervals from the coastline, as well as to test whether the same patterns are discernible in the various datasets.
-As seawater temperature is known to have a large influence on species distribution and productivity (Bolton 2010; Smit et al.
-2013), thus it is important to understand not only how temperatures vary at different distances from the coastline but what contributing factors to ecosystem maintenance the intensity of these events have across these distance intervals from the coastline.
+As seawater temperature is known to have a large influence on species distribution and productivity [@Bolton2010; @smit2013], thus it is important to understand not only how temperatures vary at different distances from the coastline but what contributing factors to ecosystem maintenance the intensity of these events have across these distance intervals from the coastline.
 Of particular interest to this study; the Benguela upwelling system, which is distinct in that it is one of four most productive regions in the world.
@@
 
-We would be amiss to point out some weaknesses inherent in the SACTN dataset. Temperature
-measurement along the South African coastline started in the 1970s, and has been inconsistent in
-terms of the instruments used, the modifications to styles and methods, calibration, as well as site locations themselves (Smit et al.
-2013; Schlegel and Smit 2016).
+We would be amiss to point out some weaknesses inherent in the SACTN dataset.
+Temperature measurement along the South African coastline started in the 1970s, and has been inconsistent in terms of the instruments used, the modifications to styles and methods, calibration, as well as site locations themselves [@Smit2013; @schlegel2016].
 Additionally, older records within some datasets, such as the SACTN dataset, may have been lost or are unreliable since the metadata for these are no longer available.
@@
 
-Data in the SACTN are compiled from multiple sources, in addition to being acquired by various means, thus creating considerable heterogeneity; however, efforts are
-currently underway to homogenise these datasets (Williamson et al., in progress). The underwater
-temperature recorders (UTRs) used to collect data in the SACTN dataset expressed a lower number of NA values as compared to the data collected with hand thermometers.
-As such, this may have influenced the overall time series dataset (Schlegel and Smit, 2016).
+Data in the SACTN are compiled from multiple sources, in addition to being acquired by various means, thus creating considerable heterogeneity; however, efforts are currently underway to homogenise these datasets (Williamson et al., in progress).
+The underwater temperature recorders (UTRs) used to collect data in the SACTN dataset expressed a lower number of NA values as compared to the data collected with hand thermometers.
+As such, this may have influenced the overall time series dataset [@schlegel2016].
 The level of precision at which data were collected also influenced the length of the time series needed.
-Time series in which temperature is collected at a precision of 0.5°C may require another 24 months of recordings to precisely detect long term variation (Schlegel and Smit, 2016).
+Time series in which temperature is collected at a precision of 0.5°C may require another 24 months of recordings to precisely detect long term variation [@schlegel2016].
 The average length of the thermometer time series component of the SACTN dataset was 346 months whereas the average length of UTR time series was less than half of that.
@@
 These data are often used to model and predict a wide range of oceanic and biological processes within open ocean ecosystems and habitats; however, these have only recently been used to study temperature variations influencing benthic organisms (Pearce, Faskel, and Hyndes, 2006).
-SST data acquired by satellites deviate from coastal in situ collected seawater temperatures for a variety of reasons (Smit et al.
-2013).
+SST data acquired by satellites deviate from coastal in situ collected seawater temperatures for a variety of reasons [@smit2013].
 The SST data presented the broader-scale situation along the coast, therefore utilised here as a forcing boundary hypothesised here as influencing coastal temperatures.

### 2019-10-09 Write_up/Paper1_writeup.Rmd
fix
update

@@
 ---
-title: "Variability in upwelling patterns across the Benguela upwelling systems"
+title: "Variability in upwelling patterns across the Benguela region"
 author: "Amieroh Abrahams"
@@
 
-knitr::include_graphics("../Figures/wind_roses.png")
+knitr::include_graphics("../Figures/wind_roses.pdf")
 

### 2019-10-09 Write_up/Paper1_writeup.Rmd
draft_writeup

@@
 
-The linear regression analysis showed that the cumulative intensity among all variables within each dataset showed the most variation.
-The cumulative intensity acts as absolute magnitude rather than relative seasonal climatology.
-When looking at the results for the data obtained from the OISST dataset it is evident that R^2 values are highest at greater distances from the coastline for each of the sites except for Port Nolloth.
-For examples, Lamberts Bay had an R^2 value of 5.47  at a distance of 10000km but only an R^2 value of -8.76 at 50000km.
-Similar trends were apparent for Saldanha Bay (R^2 -3.28 at 10000 km and -12.42 at 50000 km) and Sea Point (R^2  = -8.77 at 10000 km)
-
-The CMC remotely sensed sea surface temperature dataset had higher R^2 values compared to the OISST dataset.
-Most of the sites within CMC dataset had greater R^2 values at distances furthest from the coastline.
-Here, Lamberts Bay had R^2 value of -26.30 for 10000km and -31.77 for 50000km respectively.
-Similarly, Port Nolloth had R^2 values of -15.34 at 10000km and -20.45 at 50000km.
-Saldanha Bay however presented its lowest value of -24.84 at 10000km and highest R^2 value of -46.39 at 40000km and also the highest value among all of the sites.
-Additionally, Sea Point displayed similar patterns with the lowest value occurring at a distance of 10000km from the coastline and its highest R^2 value of -38.35 occurring at a distance of 40000km from the coastline.
-
-Lastly, the SACTN in situ collected SST data collected directly at the coastline had variable R^2 values.
-Lamberts Bay had an R^2 value of -1.46.
-Port Nolloth, representing the second highest value of -3.65.
-Saldanha Bay representing an R^2 value of -2.56 and the highest R^2 value of -10.96 occurring at Sea Point.
+The regression slope shows the slope of the trend for each of the datasets used within this study.
+When looking at the OISST dataset, it is eveident that there is no contstant trend among sites.
+It is evident however that the strongest trends exist at 10km from the coastline, with the weakest trends occuring at a distance of 30 – 40km from the coastline.
+In the CMC dataset events occuring at a distance of 10km from the coastline shows the strongest trend compared to trends observed at a distance of 50km from the coasltine.
+In this dataset we see that at a distance of 10km, Lamberts Bay shows that every decade upwelling lasts for approximatly 3days longer.
+Similarly, Port Nolloth shows that upwelling lasts for approximately 2 days longer.
+Both Saldanha Bay and Sea Point shows a trend of upwelling lasting for approximately 3 days more.
+When looking at the in situ collected temperatur data the trend represents an increase of 1 additional day per decade for each of the sites expect Saldanha Bay.
+Here Saldanha Bay represent a trend value of 0.37 which largely varies with the trend values obtained from the OISST and CMC dataset.
 

### 2019-10-12 Write_up/Appendices.Rmd
update

@@
 ---
-title: Quantifying the impact of wind and wave action on seawater temperature along the South African coastline (Appendices)
+title: Variability in upwelling patterns across the Benguela
 author:
@@
   name: Amieroh Abrahams
-date: "19 November 2018"
+date: "18 September 2019"
 output:

### 2019-10-12 Write_up/Paper1_writeup.Rmd
update

@@
 ---
-title: "Variability in upwelling patterns across the Benguela region"
-author: "Amieroh Abrahams"
+title: Variability in upwelling patterns across the Benguela
+author:
+- affiliation: University of the Western Cape
+  name: Amieroh Abrahams
 date: "18 September 2019"
 output:
+  pdf_document:
+    fig_caption: yes
+    fig_height: 6.5
+    fig_width: 6.5
+    highlight: default
+    keep_tex: yes
+    latex_engine: xelatex
+    template: ../LaTeX/template-article_ajs.tex
+  html_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: cerulean
+    toc_float: no
+  md_document:
+    variant: markdown_github
   word_document:
@@
     highlight: zenburn
-    toc: yes
-  html_document:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  bookdown::html_document2:
-    fig_caption: yes
-    fig_retina: 2
-    fig_width: 7
-    highlight: espresso
-    theme: sandstone
-    toc_float: no
-  pdf_document:
-    toc: yes
-    toc_depth: 4
-    highlight: default
-    keep_tex: yes
-    latex_engine: xelatex
+    toc: no
 fontsize: 10pt
@@
 language: Australian
-mainfont: Minion Pro
-monofont: Courier
+csl: ../LaTeX/frontiers.csl
 papersize: A4
-sansfont: Myriad Pro
 tables: yes
-bibliography: ../bib_master.bib
+bibliography: ../LaTeX/bib_masters.bib
 ---
@@
 This study was designed to examine whether same upwelling patterns were discernible at varying distances from the coastline and if this was apparent from different datasets.
-Analyses of temperature time series data from 4 sites within the west coast of South Africa were conducted to examine the variation in intensity occurring at different upwelling events at different distances from the coastline.
+Analyses of temperature time series data from 4 sites within the west coast of South Africa were conducted to examine the variation in intensity occurring in different upwelling events at different distances from the coastline.
 This study further made use of wind data to determine the upwelling index for each of the sites.
@@
 
-The regression slope shows the slope of the trend for each of the datasets used within this study.
-When looking at the OISST dataset, it is eveident that there is no contstant trend among sites.
-It is evident however that the strongest trends exist at 10km from the coastline, with the weakest trends occuring at a distance of 30 – 40km from the coastline.
-In the CMC dataset events occuring at a distance of 10km from the coastline shows the strongest trend compared to trends observed at a distance of 50km from the coasltine.
-In this dataset we see that at a distance of 10km, Lamberts Bay shows that every decade upwelling lasts for approximatly 3days longer.
-Similarly, Port Nolloth shows that upwelling lasts for approximately 2 days longer.
-Both Saldanha Bay and Sea Point shows a trend of upwelling lasting for approximately 3 days more.
-When looking at the in situ collected temperatur data the trend represents an increase of 1 additional day per decade for each of the sites expect Saldanha Bay.
-Here Saldanha Bay represent a trend value of 0.37 which largely varies with the trend values obtained from the OISST and CMC dataset.
-
 # Wind rose diagrams

### 2019-10-12 Write_up/Paper1_writeup.Rmd
fixing
updated_fixed_refs

@@
 Research done by Copenhagen 1953, reported a number of sites along the west coast of southern Africa where upwelling was particularly intense.
-The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrew].
+The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrews].
 Whilst upwelling plays a significant role in ecological processes, climate change largely influences this phenomenon.
@@
 Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed).
-This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
+This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jacox2016].
 Climate change impacting oceanic processes has major ecological consequences [@walther].
@@
 The 3000km South African coastline is bordered by the Agulhas current along the east coast and the Benguela current (an EBUS) along the west coast.
-These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@Mead; @schlegel; @schlegel2016].
-The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region [@jarred; @hutchings].
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
+These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@mead; @schlegel; @schlegel2016].
+The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region [@jarre; @hutchings].
+The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@lutjeharms1987; @lutjeharms2003; @roberts; @hutchings].
 Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast [@luning1990; @lutjeharms2000; @hutchings; @schlegel].
@@
 These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole], with water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
+This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole], with water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@lutjeharms2000].
 The upwelling period in the Benguela north of 25°C persists throughout the entire year but peaks during August.
@@
 Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tend to vary based on seasonal wind variation [@andrews].
-Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters [@andrews; @Shannon].
+Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters [@andrews; @shannon].
 Due to the temperature gradients that exist inshore and offshore, this study focused on observing upwelling patterns from the coastal region to 50km from the coastline.
@@
 The primary aim of the first study was to make use of various remotely sensed SST and *in-situ* collected coastal seawater temperature data.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit].
+The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit2013].
 It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
@@
 
-The South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions [@smit, @mead].
+The South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions [@smit2013, @mead].
 The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings], which provides a natural laboratory for this study.
@@
 This study aimed to investigate patterns occurring within upwelling events at various different spatial intervals from the coastline, as well as to test whether the same patterns are discernible in the various datasets.
-As seawater temperature is known to have a large influence on species distribution and productivity [@Bolton2010; @smit2013], thus it is important to understand not only how temperatures vary at different distances from the coastline but what contributing factors to ecosystem maintenance the intensity of these events have across these distance intervals from the coastline.
+As seawater temperature is known to have a large influence on species distribution and productivity [@bolton2010; @smit2013], thus it is important to understand not only how temperatures vary at different distances from the coastline but what contributing factors to ecosystem maintenance the intensity of these events have across these distance intervals from the coastline.
 Of particular interest to this study; the Benguela upwelling system, which is distinct in that it is one of four most productive regions in the world.
@@
 We would be amiss to point out some weaknesses inherent in the SACTN dataset.
-Temperature measurement along the South African coastline started in the 1970s, and has been inconsistent in terms of the instruments used, the modifications to styles and methods, calibration, as well as site locations themselves [@Smit2013; @schlegel2016].
+Temperature measurement along the South African coastline started in the 1970s, and has been inconsistent in terms of the instruments used, the modifications to styles and methods, calibration, as well as site locations themselves [@smit2013; @schlegel2016].
 Additionally, older records within some datasets, such as the SACTN dataset, may have been lost or are unreliable since the metadata for these are no longer available.

### 2019-10-13 Write_up/Paper1_writeup.Rmd
updated

@@
 Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of the motion caused by alongshore equatorial winds.
-These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
+These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes.
 Occurring globally throughout the world's oceans, four major eastern boundary upwelling systems (EBUS) can be recognised; each significantly impacting associated coastal systems.
-These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly].
+These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun1990; @pauly; @bakun2015].
 There has been significant interest in major upwelling systems since the later 1960s, with major interest occuring off Peru [@walsh] and North-West Africa [@barber].
@@
 The consequences of these changes are also shown to have had direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients [@bakun].
+According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients [@bakun1990].
 Pressure gradients are known to affect wind patterns which will subsequently affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
@@
 
-Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this).
+Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades [@vose2012noaa].
 These temperatures have increased as a result of anthropogenically driven climate change [@gruber].
-Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed).
+Warming is shown to slowly move down into the water column thus increasing thermal stratification [@chhak; @jacox2011].
 This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jacox2016].
@@
 Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
+Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski1993; @murawski2000] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.
 
 The 3000km South African coastline is bordered by the Agulhas current along the east coast and the Benguela current (an EBUS) along the west coast.
-These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@mead; @schlegel; @schlegel2016].
+These coastal regions display noticeable differences in seawater temperatures primarily due to the influences of the neighbouring ocean currents [@mead; @schlegel2016; @schlegel].
 The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region [@jarre; @hutchings].
@@
 The upwelling period in the Benguela north of 25°C persists throughout the entire year but peaks during August.
-The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region in Namibia [@shannon; @cole].
+The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region in Namibia [@shannon1984; @shannon1986; @cole].
 Several studies have shown that extreme warming and cooling periods occur in the northern and southern Benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
@@
 The Cape peninsula is influenced by south-east trade winds during summer and westerlies during winter, thus a pronounced seasonal cycle is evident within patterns of wind movement across the coastline.
-During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent (RWS: Citation).
-During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months (RWS: Citation).
+During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent [@preston1977towards; @bakun1990].
+During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months [@bakun1990; bakun2015].
 During the winter seasons, low pressure cells are driven from the west to the east as a result of the atmospheric circumpolar waves and so north-westerly winds dominate during winter.
 South easterly winds dominate during summer but are still present for short durations during winter [@andrews].
-Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region (RWS: Citation) and as a result conditions are variable in the nearshore and offshore.
+Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region [@andrews] and as a result conditions are variable in the nearshore and offshore.
 
@@
 Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tend to vary based on seasonal wind variation [@andrews].
-Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters [@andrews; @shannon].
+Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters [@andrews; @shannon1984].
 Due to the temperature gradients that exist inshore and offshore, this study focused on observing upwelling patterns from the coastal region to 50km from the coastline.

### 2019-10-15 Write_up/Paper1_writeup.Rmd
updated

@@
 The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling and the likelihood of the phenomenon occurring.
-The above equation produces a value called the upwelling index; for this study specifically, upwelling was classified to occur when the index is greater than 1.
+The above equation produces a value called the upwelling index; for this study specifically, upwelling was classified to occur when the index value is greater than 1.
 However to further ascertain a threshold for upwelling, the `exceedance()` function within the _`heatwaveR`_ package [@schlegel2017heat] was utilised, specifically to determine the consecutive number of days that the event must exceed for it to be regarded and classified as an upwelling event.
 However it should be noted that upwelling is known to vary on a seasonal basis and also may occur hourly.
-Therefore this study sought to establish the minimum duration for the analyses of upwelling as 1 day; the rationale behind this being that data from the SACTN dataset as well as the Satellite remotely sensed SST data is collected daily.
-Furthermore, it was also determined that a 25th percentile threshold for the upwelling index result lasting for >= 1 days matched with the upwelling index value of 1.
+Therefore this study sought to establish the minimum duration for the analyses of upwelling as 1 day; the rationale behind this being that data from the SACTN dataset as well as the satellite remotely sensed SST data is collected daily.
+Furthermore, it was also decided to make use of a percentile value rather than a static upwelling index value of >1 as percentile values are often more approved.
+It was then determined that a 25th percentile threshold for the upwelling index result lasting for >= 1 days matched with the upwelling index value of >1.
 As a result the 25th percentile threshold was used as a more acceptable value to predict upwelling events.
@@
 A series of ANOVA tests were used to compare the main effects of the chosen variables on a continuous variable.
-In this analysis the relationship between sites based on the mean intenisity of upwelling as a function of distance and duration were analysed; these analyses tested if significant
-differences occurred between upwelling at different sites and distances from the coastline within each of the various datasets.
+In this analysis the relationship between sites based on the mean intenisity of upwelling as a function of distance and duration were analysed; these analyses tested if significant differences occurred between upwelling at different sites and distances from the coastline within each of the various datasets.
 
@@
 
-Sea Point showcased the most intense upwelling event at 10 km from the coastline during the year 2013, lasting for 21 days (Refer to figure, do this throughout).
+Sea Point showcased the most intense upwelling event at 10 km from the coastline during the year 2013, lasting for 21 days (Fig 2).
 The majority of upwelling events in Sea Pont occurred around 10 km away from the coastline lasted for less than five days, and had an approximate cumulative intensity ranging between of -10.00deg.
@@
 
-Within the SACTN dataset, the majority of upwelling events detected in Lamberts Bay lasted for longer than five days.
+Within the SACTN dataset (Fig 3), the majority of upwelling events detected in Lamberts Bay lasted for longer than five days.
 The longest upwelling event was detected in 2012 lasting for approximately 12 days with a cumulative intensity of -12.63deg.
@@
 
-CMC observations of upwelling events at Sea Point detected durations of less than four days.
+CMC observations of upwelling events (Fig 4) at Sea Point detected durations of less than four days.
 The most intense event occurred in 2010, and I found that intensifies appeared to increase with distance from the coastline.
@@
 
-Results gathered during the course of this study clearly illustrate that upwelling events detected in different ocean sea surface temperature datasets showed variations with regards to duration and intensities across different time intervals.
+Results gathered during the course of this study clearly illustrate that upwelling events detected in different ocean sea surface temperature datasets showed variations with regards to duration and intensities across different distances.
 Noticeably, some patterns were dominant and persistent throughout the datasets; however, it was also observed that fewer upwelling events occurred at distances furthest away from the coastline.

### 2019-10-15 SST_patterns.Rmd
Plot changes

@@
                      limits = c(1,30)) +
+  theme_bw() +
   xlab("Date (years)") + ylab("Event duration [days]") 
@@
 
-combined_OISST_pl <- ggarrange(OISST_SP_pl, OISST_SB_pl, OISST_PN_pl,OISST_LB_pl)
+combined_OISST_pl_1 <- ggarrange(OISST_SP_pl, OISST_SB_pl)
+
+combined_OISST_pl_2 <- ggarrange(OISST_PN_pl,OISST_LB_pl)
 
@@
 
-combined_CMC_pl <- ggarrange(CMC_SP_pl, CMC_SB_pl, CMC_PN_pl,CMC_LB_pl)
+combined_CMC_pl_1 <- ggarrange(CMC_SP_pl, CMC_SB_pl)
+
+combined_CMC_pl_1 <- ggarrange(CMC_PN_pl,CMC_LB_pl)
 

### 2019-10-15 Write_up/Paper1_writeup.Rmd
updated

@@
 ```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
-This graph shows the duration of events on the y-axis and the time variable on the x-axis for the OISST dataset.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for Sea Point and Saldanha Bay using the OISST dataset.
 The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
 
-knitr::include_graphics("../Figures/OISST.pdf")
+knitr::include_graphics("../Figures/OISST_1.pdf")
+```
+
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for Port Nolloth and Lamberts Bay using the OISST dataset.
+The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/OISST_2.pdf")
 ```
@@
 
-Throughout Port Nolloth, most of the events detected lasted less than five days.
+Throughout Port Nolloth (Fig 3), most of the events detected lasted less than five days.
 At a distance of 10000 km one obvious major event occurred during 2016 which had a cumulative intensity of -43.39deg.
@@
 
-Events occurring within Lamberts Bay lasted for approximately less than five days.
+Events occurring within Lamberts Bay (Fig 3) lasted for approximately less than five days.
 Three major events were detected at 10 km and 20 km away from the coastline respectively.
@@
 
-Within the SACTN dataset (Fig 3), the majority of upwelling events detected in Lamberts Bay lasted for longer than five days.
+Within the SACTN dataset (Fig 4), the majority of upwelling events detected in Lamberts Bay lasted for longer than five days.
 The longest upwelling event was detected in 2012 lasting for approximately 12 days with a cumulative intensity of -12.63deg.
@@
 ```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
-This graph shows the duration of events on the y-axis and the time variable on the x-axis for the CMC dataset.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for Sea Point and Saldanha Bay using the CMC dataset.
 The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
 
-knitr::include_graphics("../Figures/CMC.pdf")
+knitr::include_graphics("../Figures/CMC_1.pdf")
 ```
 
-CMC observations of upwelling events (Fig 4) at Sea Point detected durations of less than four days.
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Timeline of several event metrics.
+This graph shows the duration of events on the y-axis and the time variable on the x-axis for Port Nolloth and Lamberts Bay using the CMC dataset.
+The number of top events from the chosen metric may be highlighted in a brighter colour.", fig.height=8, fig.width=12, fig.pos="H"}
+
+knitr::include_graphics("../Figures/CMC_2.pdf")
+```
+
+CMC observations of upwelling events (Fig 5) at Sea Point detected durations of less than four days.
 The most intense event occurred in 2010, and I found that intensifies appeared to increase with distance from the coastline.
@@
 C x days for was detected for 20 days within 2010.
-At Port Nolloth, most of the events are found to occur for duration of three days.
+At Port Nolloth (Fig 6), most of the events are found to occur for duration of three days.
 Two major upwelling events were present at different distances within 2010 with a variation in their respective cumulative intensities.
 Those events both lasted for approximately seven days.
-Similar patterns were found within Lamberts Bay.
+Similar patterns were found within Lamberts Bay (Fig 6).
 At 30 km a major event was detected that lasted for approximately 19 days, but most events here lasted for less than five days.

### 2019-10-17 SST_patterns.Rmd
updated

@@
   theme_bw() +
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=15)) +
+  theme(strip.text.x = element_text(size = 15))+
   xlab("Date (years)") + ylab("Event duration [days]") 
@@
 
-combined_CMC_pl_1 <- ggarrange(CMC_PN_pl,CMC_LB_pl)
+combined_CMC_pl_2 <- ggarrange(CMC_PN_pl,CMC_LB_pl)
 
@@
                      limits = c(1,30)) +
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=15)) +
+  theme(strip.text.x = element_text(size = 15))+
   xlab("Date (years)") + ylab("Event duration [days]") 

### 2019-10-17 netCDF2CSV_MUR_CMC_OISST.R
fix

@@ -1,6 +1,12 @@
 # Script which converts netCDF to CSV
 # This script extracts MUR, CMC and OISST data
 
+# the most important of which is that it is self-describing, meaning that software packages can directly read the data and determine its structure
+# variable names and essential metadata such as the units
+# the information needed to ensure accurate work (reduce the incidence of errors) is available with the data itself
+# NetCDF (network Common Data Form) 
+# is a file format for storing multidimensional scientific data (variables) such as temperature, humidity, pressure, wind speed, and direction. 
+
 
 # Laoding Libraries
 library(stringr)

### 2019-10-17 SST_patterns.Rmd
updated

--- 2019-10-15 SST_patterns.Rmd
@@
   theme_bw() +
-  xlab("Date (years)") + ylab("Event duration [days]") 
+  theme(axis.text=element_text(size=17),
+        axis.title=element_text(size=17)) +
+  theme(strip.text.x = element_text(size = 17))+
+  theme(panel.spacing = unit(2, "lines")) +
+  xlab("Date (years)") + ylab("Event duration (days)") 
 }
@@
                      limits = c(1,30)) +
-  xlab("Date (years)") + ylab("Event duration [days]") 
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=15)) +
+  theme(strip.text.x = element_text(size = 15))+
+  xlab("Date (years)") + ylab("Event duration (days)") 
 }

### 2019-10-18 SST_patterns.Rmd
update

--- 2019-10-15 SST_patterns.Rmd
@@
   theme_bw() +
-  xlab("Date (years)") + ylab("Event duration [days]") 
+  theme(axis.text=element_text(size=17),
+        axis.title=element_text(size=17)) +
+  theme(strip.text.x = element_text(size = 17))+
+  theme(panel.spacing = unit(2, "lines")) +
+  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
+  xlab("Date (years)") + ylab("Event duration (days)") 
 }
@@
                      limits = c(1,30)) +
-  xlab("Date (years)") + ylab("Event duration [days]") 
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=15)) +
+  theme(strip.text.x = element_text(size = 15)) +
+  xlab("Date (years)") + ylab("Event duration (days)") 
 }

### 2019-10-18 Write_up/References/write_up.Rmd
update

--- 2019-09-17 Write_up/write_up.Rmd
@@
 For the purpose of this study the influence of upwelling on MCSs was the primary focus.
-
-# Paper 1
-## Chapter 2 Upwelling
-
-## Abstract
-
-### Introduction
-
-Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of motion caused by alongshore equatorial winds.
-These upwelling events are known to have a large influence on the abundances, diversity, and distributions of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references).
-Globally, upwelling occurs throughout the world’s oceans, however four major eastern boundary upwelling systems (EBUS) are recognised, each considerably impacting coastal systems.
-These respective systems are characterized as being vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
-The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which is regarded along with the Peru, California and Humboldt current as Eastern boundry upwelling regions [@andrews].
-Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world.
-As such there has been a constant interest in upwelling systems.
-Major upwelling programs were carried through off off Peru (e.g.
-WALSH.
-KELLEY, DUGDALE and FROST, 1971) and North-West Africa (e.g.
-BARBER, 1977).
-
-The 3000km South African coastline, is bordered by the Agulhas current along the east coast and the EBUS Benguela current, along the west coast.
-These coastal regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
-2004; Mead et al.
-2013; Schlegel and Smit 2016).
-The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
-2009).
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
-Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al.
-2009; Schlegel et al.
-2017).
-Additionally,  changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
-
-Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
-The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
-According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients.
-Pressure gradients are known to affect wind patterns and thus will subsequently  affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].
-
-Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references).
-These temperatures are agreed to have increased as a result of anthropogenically driven climate change [@gruber].
-Warming is shown to slowly move down into the water column thus increasing thermal stratification.
-This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
-Climate change impacting oceanic processes has major ecological consequences (Bradley et al.
-1999; Walther et al.
-2002).
-This is demonstrated in several regions whereby which increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance.
-Climate change has also been shown tp largely influence coastal species such as corals, muscles and oysters.
-Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
-Furthermore it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993); thus understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in it’s negative affects to local and broader economies as well as human populations dependent on these coastal regions.
-
-The Benguela system is bounded by the equatorward and poleward ends by warm water regimes.
-These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel].
-This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole].
-With water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The upwelling period in the Benguela north of 25°C persist throughout the entire year but peaks during August.
-The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
-Several  studies have shown that extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
-Furthermore, Walter (1937) found that warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events.
-During the periods of warming the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
-In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
-Upwelling occurring along the south coastline of the Southern Africa region are particularly generated by local winds.
-Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]
-
-The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood.
-Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore [@andrews].
-Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011).
-There is by no means an understanding of how the upwelling process varies at local versus broader scales.
-We hypothesize that upwelling is more dominant within coastal regions instead of broader scale.
-Very few studies have addressed the changes in the intensity or duration of cold temperature spells or cold events at varying distances from the coast within the Benguela region.
-
-The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared.
-Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
-The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit].
-It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.
-This investigation examines recent trends and variation of upwelling events at different distances from the coastline.
-
-## Methods
-
-## Site selection
-
-At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
-The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study.
-Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the weastern limit near the Namibian border.
-Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
-This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].
-In order to accurately examine upwelling patterns at various distances from the coast several sites were selected as this provides a more accurate representation, the appropriate sites along the west coast of South Africa (i.e.
-sites influenced by the Benguela current) were selected.
-Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years.
-This yield a total of five sites within the Benguela region (Fig 1).
-
-```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
-The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}
-
-knitr::include_graphics("../Figures/site_map.png")
-```
-
-## Temperature data sources
-
-The South African Coastal Temperature Network (SACTN) data serves as the *in situ* temperature dataset used in this study.
-This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline.
-The SACTN has collected SST data along the unique coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
-As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
-The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
-The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.
-During the 1970s, a total of 11 time series began recording.
-A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
-Recordings are still ongoing at many of these sites.
-
-This study also made use of a range of remotely sensed temperature datasets.
-Satellites provide a system of large spatial and temporal data retrieval.
-Remotely sensed SST products are produced by a number of organisations across the globe.
-The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
-OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
-MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available.
-The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
-The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
-
-An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and thus can more accurately explain upwelling within the coastal inshore environment.
-Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.
-
-Inorder to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist at the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa.
-Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
-These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig.
-1).
-The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.
-
-## Wind data
-
-Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
-Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
-Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
-Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data.
-The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters.
-ERA5 hourly data on pressure levels was used as a second wind source.
-This dataset comprise of u and v variables from 1973 to present.
-
-With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested.
-To do this, it was important to identify upwelling.
-
-# Defining and determining upwelling
-
-Inorder to determine the change in upwelling signals at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was first necessary to define when upwelling was occuring.
-We then aimed to determine a threshold value for which we could identify when upwelling was occuring within this region.
-Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider both temperature and wind values.
-To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989).
-The upwelling index provided an indication of the intensity of the upwelling process taking place using the following equation:
-
-                  $$ UpwellingIndex = μ{(Cosθ − coastal angle)} $$
-
-In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
-The coastal angle is the angle of each of the sites perpendicular to the coastline.
-The index relies heavily on wind speed and direction data inorder to determine the intensity of upwelling likely to occur.
-Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value is ment to be.
-Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling index as 1.
-Inorder to obtain a more approved threshold value, a percentile value was determined.
-To determine changes within the upwelling values, the 'detect event' fuction within the heatwaveR package was used.
-Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile.
-Values obtained from the upwelling indeces were used as a second threshold.
-An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data.
-Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months.
-For the use of the heatwaveR package it was best to use a time series of 30years.

### 2019-10-23 SST_patterns.Rmd
paper1 fixing

@@
   ggplot(df, aes(x = date_peak, y = duration)) + 
-  geom_lolli(aes(colour = intensity_cumulative)) + 
-  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  geom_lolli() + 
+  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
   facet_wrap(~distance, ncol = 2) +

### 2019-10-23 SST_patterns.Rmd
extracting G1SST

@@
 # save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
+```
+
+# Extracting G1SST
+
+```{r}
+bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+csvDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/G1SST_extracted"
+
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+#          1         2         3         4         5         6
+# 1234567890123456789012345678901234567890123456789012345678901
+# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+  ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+    coords <- bbox[, region]
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
+    fNameStem <-
+      substr(basename(ncFile), 10, 58)
+    fDate <- substr(basename(ncFile), 1, 8)
+    LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+    LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
+    fwrite(sst,
+           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+  }
+
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
 ```

### 2019-10-24 SST_patterns.Rmd
update

@@
 ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-csvDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/G1SST_extracted"
+csvDir <- "/media/amieroh/Seagate Expansion Drive/G1SST_extracted"
 
@@
 llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+
 ```
@@
   Filtered <- df %>%
-filter(year(date_start) %in% seq(2010, 2016, 1))
+filter(year(date_start) %in% seq(2010, 2015, 1))
 }
@@
   ggplot(df, aes(x = date_peak, y = duration)) + 
-  geom_lolli() + 
+  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
   #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 

### 2019-10-25 SST_patterns.Rmd
updated

@@
   Filtered <- df %>%
-filter(year(date_start) %in% seq(2010, 2015, 1))
+filter(year(date_start) %in% seq(2010, 2014, 1))
 }
@@
   theme(strip.text.x = element_text(size = 17))+
-  theme(panel.spacing = unit(2, "lines")) +
+  theme(panel.spacing = unit(0.5, "lines")) +
   #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
@@
 (OISST_LB_pl <- plot_loli_func(df = OISST_LB))
+
 OISST_SP_pl <- OISST_SP_pl +
-  ggtitle("Sea Point")
+  ggtitle("D.") #Sea Point
+OISST_SB_pl <- OISST_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+OISST_PN_pl <- OISST_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+OISST_LB_pl <- OISST_LB_pl +
+  ggtitle("A.")#Lamberts Bay
 
-OISST_SB_pl <- OISST_SB_pl +
-  ggtitle("Saldanha Bay")
-OISST_PN_pl <- OISST_PN_pl +
-  ggtitle("Port Nolloth")
-OISST_LB_pl <- OISST_LB_pl +
-  ggtitle("Lamberts Bay")
+combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl)
 
-combined_OISST_pl_1 <- ggarrange(OISST_SP_pl, OISST_SB_pl)
-
-combined_OISST_pl_2 <- ggarrange(OISST_PN_pl,OISST_LB_pl)
+combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)
 
@@
 CMC_SP_pl <- CMC_SP_pl +
-  ggtitle("Sea Point")
+  ggtitle("D.") #Sea Point
 CMC_SB_pl <- CMC_SB_pl +
-  ggtitle("Saldanha Bay")
+  ggtitle("C.") # Saldanha Bay
 CMC_PN_pl <- CMC_PN_pl +
-  ggtitle("Port Nolloth")
+  ggtitle("B.")        # Port Nolloth
 CMC_LB_pl <- CMC_LB_pl +
-  ggtitle("Lamberts Bay")
+  ggtitle("A.")#Lamberts Bay
 
-combined_CMC_pl_1 <- ggarrange(CMC_SP_pl, CMC_SB_pl)
+combined_CMC_pl_2 <- ggarrange(CMC_SB_pl, CMC_SP_pl)
 
-combined_CMC_pl_2 <- ggarrange(CMC_PN_pl,CMC_LB_pl)
+combined_CMC_pl_1 <- ggarrange(CMC_LB_pl, CMC_PN_pl)
 
@@
   ggplot(df, aes(x = date_peak, y = duration)) + 
-  geom_lolli(aes(colour = intensity_cumulative)) + 
-  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
+  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
    facet_wrap(~site, ncol = 2) +
@@
                      limits = c(1,30)) +
-  theme(axis.text=element_text(size=15),
-        axis.title=element_text(size=15)) +
-  theme(strip.text.x = element_text(size = 15)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=17),
+        axis.title=element_text(size=17)) +
+  theme(strip.text.x = element_text(size = 17))+
+  theme(panel.spacing = unit(0.5, "lines")) +
+  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
   xlab("Date (years)") + ylab("Event duration (days)") 
 }
+
 (SACTN_plot <- plot_loli_func(df = SACTN_filtered))

### 2019-10-26 G1SST and MUR extraction.Rmd
UPDATE

---
title: "G1SST and MUR"
author: "Amieroh Abrahams"
date: "26 October 2019"
output: html_document
---

This scripts extracts the MUR and G1sst data

```{r}
bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
library(stringr)
library(tidyverse)
library(reshape2)
library(ncdf4) # library for processing netCDFs
library(plyr)
library(lubridate)
library(data.table)
library(doMC); doMC::registerDoMC(cores = 7)
ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/2010"
csvDir <- "/home/amieroh/Documents/tester"

# #          1         2         3         4
# # 12345678901234567890123456789012345678901
# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
#          1         2         3         4         5         6
# 1234567890123456789012345678901234567890123456789012345678901
# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc

# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
  ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
    coords <- bbox[, region]
    nc <- nc_open(ncFile)
    pathLen <- nchar(paste0(ncDir, "/")) + 1
    fNameStem <-
      substr(basename(ncFile), 10, 58)
    fDate <- substr(basename(ncFile), 1, 8)
    LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
    LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
      round(4)
    dimnames(sst) <- list(lon = nc$dim$lon$vals,
                          lat = nc$dim$lat$vals)
    nc_close(nc)
    sst <- as_tibble(melt(sst, value.name = "temp"))
    sst$t <- ymd(fDate)
    na.omit(sst)
    fwrite(sst,
           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
           append = TRUE, col.names = FALSE)
    rm(sst)
  }

  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
  strtDate <- str_sub(ncFirst, start = 1, end = 8)
  endDate <- str_sub(ncLast, start = 1, end = 8)

llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
```

```{r}
G1SST_prod <- G1SST %>%
  select(lon, lat) %>%
  unique() %>%
  mutate(product = "G1SST")

MUR_prod <- MUR %>%
  select(lon, lat) %>%
  mutate(product = "MUR")

sat_pixels <- sat_data %>%
select(product, lon, lat) %>%
unique()

match_func <- function(df){
  df <- df %>%
    dplyr::rename(lon_site = lon, lat_site = lat)
  G1SST_index <- G1SST_prod[as.vector(knnx.index(as.matrix(G1SST_prod[,c("lon", "lat")]),
                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
    cbind(., df)
  MUR_index <- MUR_prod[as.vector(knnx.index(as.matrix(MUR_prod[,c("lon", "lat")]),
                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
    cbind(., df)
  res <- rbind(G1SST_index, MUR_index)
  return(res)
}

pixel_match <- site_pixels %>%
  group_by(site) %>%
  group_modify(~match_func(.x))

G1SST_fill <- right_join(G1SST, filter(pixel_match, product == "G1SST"), by = c("lon", "lat"))
MUR_fill <- right_join(MUR, filter(pixel_match, product == "MUR"), by = c("lon", "lat"))

# Clean up some RAM space
rm(MUR, G1SST); gc()

selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")

G1SST_fill <- G1SST_fill %>%
  filter(site %in% selected_sites)

MUR_fill <- MUR_fill %>%
  filter(site %in% selected_sites)

load("Data/UI_angle.RData")
upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
  return(res)
}

ts2clm_custom <- function(df){
  # The climatology base period used here is up for debate...
  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
  return(res)
}

# Calculate the upwelling event metrics
upwelling_detect_event <- function(df){
  upwell_base <- df %>%
    dplyr::rename(t = date) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~ts2clm_custom(.x)) %>%
    left_join(upwelling, by = c("site", "t")) %>%
    filter(!is.na(exceedance)) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~detect_event_custom(.x))
  }

G1SST_upwell_base <- upwelling_detect_event(df = G1SST_fill)
save(G1SST_upwell_base, file = "Data/G1SST_upwell_base.RData")
MUR_upwell_base <- upwelling_detect_event(df = MUR_fill)
save(MUR_upwell_base, file = "Data/MUR_upwell_base.RData")
```

### 2019-10-26 SST_patterns.Rmd
UPDATE

@@
 ```{r}
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# 
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
-# 
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-    # fNameStem <-
-    #   substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-# 
-# 
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
+
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+
+  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
+fNameStem <-
+  substr(basename(ncFile), 10, 38)
+    fDate <- substr(basename(ncFile), 1, 8)
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
+    fwrite(sst,
+           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+  }
+
+llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+
 # MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
@@
 ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-csvDir <- "/media/amieroh/Seagate Expansion Drive/G1SST_extracted"
+csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
 
@@
 llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+
+ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+
+  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
+fNameStem <-
+  substr(basename(ncFile), 10, 58)
+    fDate <- substr(basename(ncFile), 1, 8)
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
+    fwrite(sst,
+           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+  }
+
+llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
 

### 2019-10-26 G1SST and MUR extraction.Rmd
update

@@
 
-This scripts extracts the MUR and G1sst data
+Extracting G1SST
 
@@
 library(doMC); doMC::registerDoMC(cores = 7)
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/2010"
+library(heatwaveR)
+ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/2010_06-08"
 csvDir <- "/home/amieroh/Documents/tester"
@@
 
+# Extracting MUR
+
+```{r}
+# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+#          1         2         3         4
+# 12345678901234567890123456789012345678901
+# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+  ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+    coords <- bbox[, region]
+    nc <- nc_open(ncFile)
+    pathLen <- nchar(paste0(ncDir, "/")) + 1
+    fNameStem <-
+      substr(basename(ncFile), 10, 38)
+    fDate <- substr(basename(ncFile), 1, 8)
+    LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+    LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+      round(4)
+    dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                          lat = nc$dim$lat$vals)
+    nc_close(nc)
+    sst <- as_tibble(melt(sst, value.name = "temp"))
+    sst$t <- ymd(fDate)
+    na.omit(sst)
+    fwrite(sst,
+           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+           append = TRUE, col.names = FALSE)
+    rm(sst)
+  }
+
+  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+  strtDate <- str_sub(ncFirst, start = 1, end = 8)
+  endDate <- str_sub(ncLast, start = 1, end = 8)
+
+llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+```
+
+Data analysis
+
 ```{r}
@@
 
+load("Data/site_pixels.RData")
 pixel_match <- site_pixels %>%
@@
 
-load("Data/UI_angle.RData")
-upwelling <- UI_angle %>%
-  dplyr::rename(temp = ui.saws) %>%
-  group_by(site) %>%
-  # mutate(min_t = min(t),
-  #        max_t = max(t)) %>%
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
-Upwelling can last for a few hours.
-Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>%
-  select(-data, -clim) %>%
-  unnest() %>%
-  filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance)
+load("Data/upwelling.RData")
 

### 2019-10-26 SST_patterns.Rmd
update

@@
 ```{r}
-library(stringr)
-library(tidyverse)
-library(reshape2)
-library(ncdf4) # library for processing netCDFs
-library(plyr)
-library(lubridate)
-library(data.table)
-library(doMC); doMC::registerDoMC(cores = 7)
-
-ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
-
-#          1         2         3         4
-# 12345678901234567890123456789012345678901
-# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-
-read_nc <- function(ncDir = ncDir, csvDir = csvDir)
-  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  strtDate <- str_sub(ncFirst, start = 1, end = 8)
-  endDate <- str_sub(ncLast, start = 1, end = 8)
-
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-
-  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-    nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/")) + 1
-fNameStem <-
-  substr(basename(ncFile), 10, 38)
-    fDate <- substr(basename(ncFile), 1, 8)
-    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-      round(4)
-    dimnames(sst) <- list(lon = nc$dim$lon$vals,
-                          lat = nc$dim$lat$vals)
-    nc_close(nc)
-    sst <- as_tibble(melt(sst, value.name = "temp"))
-    sst$t <- ymd(fDate)
-    na.omit(sst)
-    fwrite(sst,
-           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-           append = TRUE, col.names = FALSE)
-    rm(sst)
-  }
-
-llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# 
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
+# 
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+# fNameStem <-
+#   substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
 
@@
 ```{r}
-bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
-library(stringr)
-library(tidyverse)
-library(reshape2)
-library(ncdf4) # library for processing netCDFs
-library(plyr)
-library(lubridate)
-library(data.table)
-library(doMC); doMC::registerDoMC(cores = 7)
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
-
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-#          1         2         3         4         5         6
-# 1234567890123456789012345678901234567890123456789012345678901
-# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-  ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-    coords <- bbox[, region]
-    nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/")) + 1
-    fNameStem <-
-      substr(basename(ncFile), 10, 58)
-    fDate <- substr(basename(ncFile), 1, 8)
-    LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-    LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-      round(4)
-    dimnames(sst) <- list(lon = nc$dim$lon$vals,
-                          lat = nc$dim$lat$vals)
-    nc_close(nc)
-    sst <- as_tibble(melt(sst, value.name = "temp"))
-    sst$t <- ymd(fDate)
-    na.omit(sst)
-    fwrite(sst,
-           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-           append = TRUE, col.names = FALSE)
-    rm(sst)
-  }
-
-  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  strtDate <- str_sub(ncFirst, start = 1, end = 8)
-  endDate <- str_sub(ncLast, start = 1, end = 8)
-
-llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
-
-#          1         2         3         4
-# 12345678901234567890123456789012345678901
-# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-
-read_nc <- function(ncDir = ncDir, csvDir = csvDir)
-  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-  strtDate <- str_sub(ncFirst, start = 1, end = 8)
-  endDate <- str_sub(ncLast, start = 1, end = 8)
-
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-
-  ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-    nc <- nc_open(ncFile)
-    pathLen <- nchar(paste0(ncDir, "/")) + 1
-fNameStem <-
-  substr(basename(ncFile), 10, 58)
-    fDate <- substr(basename(ncFile), 1, 8)
-    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-      round(4)
-    dimnames(sst) <- list(lon = nc$dim$lon$vals,
-                          lat = nc$dim$lat$vals)
-    nc_close(nc)
-    sst <- as_tibble(melt(sst, value.name = "temp"))
-    sst$t <- ymd(fDate)
-    na.omit(sst)
-    fwrite(sst,
-           file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-           append = TRUE, col.names = FALSE)
-    rm(sst)
-  }
-
-llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+# 
+# # #          1         2         3         4
+# # # 12345678901234567890123456789012345678901
+# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# #          1         2         3         4         5         6
+# # 1234567890123456789012345678901234567890123456789012345678901
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# 
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
 

### 2019-10-26 SST_patterns.Rmd
Error CMC

@@
 ```{r}
+
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+
 library(ggpubr)
@@
   Filtered <- df %>%
-filter(year(date_start) %in% seq(2010, 2014, 1))
+filter(year(date_start) %in% seq(2011, 2013, 1))
 }
@@
 
-plot_scat_func <- function(df){
-  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
-  geom_point() +
-  scale_x_date(expand = c(0, 0)) +
-  facet_wrap(~distance, ncol = 2) +
-  theme_grey() +
-  labs(y = "Mean Intensity", x = NULL)
-}
-(CMC_SP_plot <- plot_scat_func(df = CMC_SEAPOINT))
-(CMC_SB_plot <- plot_scat_func(df = CMC_SB))
-(CMC_PN_plot <- plot_scat_func(df = CMC_PN))
-(CMC_LB_plot <- plot_scat_func(df = CMC_LB))
-
-CMC_SP_plot <- CMC_SP_plot +
-  ggtitle("Sea Point") 
-CMC_SB_plot <- CMC_SB_plot +
-  ggtitle("Saldanha Bay")
-CMC_PN_plot <- CMC_PN_plot +
-  ggtitle("Port Nolloth")
-CMC_LB_plot <- CMC_LB_plot +
-  ggtitle("Lamberts Bay")
-
-combined_CMC <- ggarrange(CMC_SP_plot, CMC_SB_plot, CMC_PN_plot,CMC_LB_plot)
 ###################
@@
   filter(site == "Lamberts Bay")
-
-plot_scat_func <- function(df){
-  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
-  geom_point() +
-  scale_x_date(expand = c(0, 0)) +
-  facet_wrap(~distance, ncol = 2) +
-  theme_grey() +
-  labs(y = "Mean Intensity", x = NULL)
-}
-(OISST_SP_plot <- plot_scat_func(df = OISST_SEAPOINT))
-(OISST_SB_plot <- plot_scat_func(df = OISST_SB))
-(OISST_PN_plot <- plot_scat_func(df = OISST_PN))
-(OISST_LB_plot <- plot_scat_func(df = OISST_LB))
-
-OISST_SP_plot <- OISST_SP_plot +
-  ggtitle("Sea Point")
-OISST_SB_plot <- OISST_SB_plot +
-  ggtitle("Saldanha Bay")
-OISST_PN_plot <- OISST_PN_plot +
-  ggtitle("Port Nolloth")
-OISST_LB_plot <- OISST_LB_plot +
-  ggtitle("Lamberts Bay")
-
-combined_OISST <- ggarrange(OISST_SP_plot, OISST_SB_plot, OISST_PN_plot,OISST_LB_plot)
-
-############## 
-# SACTN
-
-plot_scat_func <- function(df){
-  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
-  geom_point() +
-  scale_x_date(expand = c(0, 0)) +
-  facet_wrap(~site, ncol = 2) +
-  theme_grey() +
-  labs(y = "Mean Intensity", x = NULL)
-}
-(SACTN_plot <- plot_scat_func(df = SACTN_filtered))
 
@@
   theme_bw() +
-  theme(axis.text=element_text(size=17),
+  theme(axis.text=element_text(size=15),
         axis.title=element_text(size=17)) +
   theme(strip.text.x = element_text(size = 17))+
-  theme(panel.spacing = unit(0.5, "lines")) +
+  theme(panel.spacing = unit(1, "lines")) +
   #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +

### 2019-10-26 SST_patterns.Rmd
Changing distance to km

@@
 
+km_func <- function(df){
+  upwell_base <-  df%>% 
+      mutate(distance_km = case_when(distance == "10000" ~ "10",
+                                    distance == "20000" ~ "20",
+                                    distance  == "30000" ~"30",
+                                    distance  == "40000" ~ "40",
+                                    distance  == "50000" ~ "50"))
+}
+OISST_upwell_base <- km_func(df = OISST_upwell_base)
+OISST_upwell_base$distance_km <- as.numeric(OISST_upwell_base$distance_km)
+
+CMC_upwell_base <- km_func(df = CMC_upwell_base)
+CMC_upwell_base$distance_km <- as.numeric(CMC_upwell_base$distance_km)
+
 library(ggpubr)
@@
   #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
-  facet_wrap(~distance, ncol = 2) +
+  facet_wrap(~distance_km, ncol = 2) +
   scale_y_continuous(breaks = c(5,10,15,20,25,30),
                      limits = c(1,30)) +
+    scale_x_date(limits = as.Date(c("2011-01-01","2014-01-01"))) +
   theme_bw() +
@@
   theme(panel.spacing = unit(1, "lines")) +
+    #theme(strip.text = c("10km", "20km", "30km", "40km", "50km")) +
   #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +

### 2019-10-27 SST_patterns.Rmd
update

@@
 )
-
+#update.packages() 
 library(tidyverse)

### 2019-10-30 SST_patterns.Rmd
update

@@
 # # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# # MUR <- MUR %>% 
-# #   select(lon, lat) %>% 
-# #   mutate(product = "MUR")
+# MUR <- MUR %>%
+#   select(lon, lat) %>%
+#   mutate(product = "MUR")
 # 
 # # Decided to work with OISST and CMC as both has a time series of 30years
-# OISST_prod <- OISST %>% 
-#   select(lon, lat) %>% 
-#   unique() %>% 
-#   mutate(product = "OISST")
+OISST_prod <- OISST %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "OISST")
 # 
-# CMC_prod <- CMC %>% 
-#   select(lon, lat) %>%
-#   unique() %>% 
-#   mutate(product = "CMC")
+CMC_prod <- CMC %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "CMC")
 # 
-# # sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+sat_data <- rbind(CMC_prod, OISST_prod) %>% 
 #   #rbind(., CMC) %>% 
-#   # select(product, lon, lat)
+select(product, lon, lat)
 # 
-# # sat_pixels <- sat_data %>% 
-#   # select(product, lon, lat) %>%
-#   # unique()
+sat_pixels <- sat_data %>%
+select(product, lon, lat) %>%
+unique()
 # 
@@
 # 
-# match_func <- function(df){
-#   df <- df %>%
-#     dplyr::rename(lon_site = lon, lat_site = lat)
-#   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
-#                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
-#     cbind(., df)
-#   CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
-#                                              as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
-#     cbind(., df)
-#   res <- rbind(OISST_index, CMC_index)
-#   return(res)
-# }
+match_func <- function(df){
+  df <- df %>%
+    dplyr::rename(lon_site = lon, lat_site = lat)
+  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
+                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  res <- rbind(OISST_index, CMC_index)
+  return(res)
+}
 # 
@@
 # # points to not fall inside of the same 25 KM pixel
-# pixel_match <- site_pixels %>% 
-#   group_by(site) %>% 
-#   group_modify(~match_func(.x))
-# 
+pixel_match <- site_pixels %>%
+  group_by(site) %>%
+  group_modify(~match_func(.x))
+
 # # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-# OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-# CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
 # #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
 # 
-# # Clean up some RAM space
-# rm(OISST, CMC); gc()
+# Clean up some RAM space
+rm(OISST, CMC); gc()
 # 
@@
 # # # Hout Bay will be ignored
-# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
 # 
-# OISST_fill <- OISST_fill %>%
-#   filter(site %in% selected_sites)
-# 
-# CMC_fill <- CMC_fill %>%
-#   filter(site %in% selected_sites)
+OISST_fill <- OISST_fill %>%
+  filter(site %in% selected_sites)
+
+CMC_fill <- CMC_fill %>%
+  filter(site %in% selected_sites)
 # 
@@
 # 
-# upwelling <- UI_angle %>% 
-#   dplyr::rename(temp = ui.saws) %>%
-#   group_by(site) %>%
-#   # mutate(min_t = min(t), 
-#   #        max_t = max(t)) %>% 
-#   nest() %>% # apply the following functions to all of the variables in te dataset
-#   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-#          # NB: A threshold of 3 appeared to be far to strict
-#          # purr::map - apllies a function to each element of a vector
-#          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+upwelling <- UI_angle %>%
+  dplyr::rename(temp = ui.saws) %>%
+  group_by(site) %>%
+  # mutate(min_t = min(t),
+  #        max_t = max(t)) %>%
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+         # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
 Upwelling can last for a few hours.
 Given that we have daily data, upwelling events minimum duration here will be 1day
-#   # Detect consecutive days in exceedance of a given threshold.
-#   # mutate() %>% 
-#   select(-data, -clim) %>% 
-#   unnest() %>%
-#   filter(row_number() %% 2 == 1) %>%
-#   unnest() %>% # creates a column for each variables
-#   dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-#   select(site, t, ui.saws, exceedance) 
+  # Detect consecutive days in exceedance of a given threshold.
+  # mutate() %>%
+  select(-data, -clim) %>%
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance)
+
+# Now applying the Upwelling func
+# NB: This only pulls out the event results and not the climatology results
+# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
+# To create a list output one would use group_map(),
+# but this then loses the labels of which sites etc. the results belong to
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
+  return(res)
+}
 # 
-# # Now applying the Upwelling func
-# # NB: This only pulls out the event results and not the climatology results
-# # This is done to keep the output tidy because group_modify() may only create data.frame type outputs
-# # To create a list output one would use group_map(), 
-# # but this then loses the labels of which sites etc. the results belong to
-# detect_event_custom <- function(df){
-#   res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
-#   return(res)
-# }
-# 
-# ts2clm_custom <- function(df){
-#   # The climatology base period used here is up for debate...
-#   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-#   res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
-#   return(res)
-# }
+ts2clm_custom <- function(df){
+  # The climatology base period used here is up for debate...
+  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  return(res)
+}
 # 

### 2019-10-30 MUR.R

@@ -1,0 +1,115 @@
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"
+# 
+# # #          1         2         3         4
+# # # 12345678901234567890123456789012345678901
+# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+# fNameStem <-
+#   substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
[…]

### 2019-10-30 MUR.R
update

@@ -1,11 +1,20 @@
-library(stringr)
 library(tidyverse)
-library(reshape2)
-library(ncdf4) # library for processing netCDFs
-library(plyr)
 library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(scales)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+library(ncdf4) # library for processing netCDFs
 library(data.table)
-library(doMC); doMC::registerDoMC(cores = 7)
+library(heatwaveR)
 
 # ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"
 # csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"

### 2019-10-30 MUR.R
update

@@ -70,6 +70,7 @@ MUR_prod <- MUR_1 %>%
   select(lon, lat) %>%
   unique() %>%
   mutate(product = "MUR")
+
 MUR_prod2 <- MUR_2 %>%
   select(lon, lat) %>%
   unique() %>%

### 2019-10-30 MUR.R
update

@@ -98,7 +98,7 @@ match_func <- function(df){
   return(res)
 }
 
-
+load("Data/site_pixels")
 pixel_match <- site_pixels %>%
   group_by(site) %>%
   group_modify(~match_func(.x))

### 2019-10-30 MUR.R
Removed some NA values

@@ -61,20 +61,30 @@ library(heatwaveR)
 # save(MUR, file = "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9/MUR_9.RData")
 # rm(MUR)
 
+
+# RWS: Both of these files load with the same onbject name
+# So the way you had it you were loading one over the other
+# And then assigning different names to the same object
 load("Data/MUR_1.RData")
+MUR_1 <- na.omit(MUR) # Half of this file is deadspace, why?
+
 load("Data/MUR_2.RData")
-MUR_1 <- MUR
-MUR_2 <- MUR
+MUR_2 <- na.omit(MUR)
+
+# Free up RAM
+rm(MUR); gc()
 
 MUR_prod <- MUR_1 %>%
   select(lon, lat) %>%
   unique() %>%
   mutate(product = "MUR")
+gc()
 
 MUR_prod2 <- MUR_2 %>%
   select(lon, lat) %>%
   unique() %>%
   mutate(product = "MUR")
+gc()
 
 sat_data <- rbind(MUR_prod, MUR_prod2) %>% 
   #   #rbind(., CMC) %>% 

### 2019-11-01 MUR.R
housekeeping

@@ -1,65 +1,73 @@
-library(tidyverse)
-library(lubridate)
-library(ggpubr)
-library(zoo)
-library(FNN)
-library(scales)
-library(gridExtra)
-library(circular)
-library(fossil)
-library(mapproj)
 library(stringr)
-library(doMC); doMC::registerDoMC(cores = 4)
-library(fasttime)
-library(xtable)
+library(tidyverse)
+library(reshape2)
 library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
 library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+library(FNN)
 library(heatwaveR)
 
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily/Extract 9"
-# 
-# # #          1         2         3         4
-# # # 12345678901234567890123456789012345678901
-# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-# fNameStem <-
-#   substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
[…]

### 2019-11-01 SST_patterns.Rmd
fixed

@@
 load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
 ```
@@
 
+MUR_upwell_base <- km_func(df = MUR_upwell_base)
+MUR_upwell_base$distance_km <- as.numeric(MUR_upwell_base$distance_km)
+
 library(ggpubr)
@@
 
+############
+#####MUR
+
+MUR_filtered <- year_func(df = MUR_upwell_base)
+
+MUR_SEAPOINT <- MUR_filtered %>% 
+  filter(site == "Sea Point")
+MUR_SB <- MUR_filtered %>% 
+  filter(site == "Saldanha Bay")
+MUR_PN <- MUR_filtered %>% 
+  filter(site == "Port Nolloth")
+MUR_LB <- MUR_filtered %>% 
+  filter(site == "Lamberts Bay")
 ```
@@
 
+##########
+### MUR
+(MUR_SP_pl <- plot_loli_func(df = MUR_SEAPOINT))
+(MUR_SB_pl <- plot_loli_func(df = MUR_SB))
+(MUR_PN_pl <- plot_loli_func(df = MUR_PN))
+(MUR_LB_pl <- plot_loli_func(df = MUR_LB))
+
+MUR_SP_pl <- MUR_SP_pl +
+  ggtitle("D.") #Sea Point
+MUR_SB_pl <- MUR_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+MUR_PN_pl <- MUR_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+MUR_LB_pl <- MUR_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_MUR_pl_2 <- ggarrange(MUR_SB_pl, MUR_SP_pl)
+
+combined_MUR_pl_1 <- ggarrange(MUR_LB_pl, MUR_PN_pl)
+
+########
+
 plot_loli_func <- function(df){

### 2019-11-01 MUR.R

@@ -168,6 +168,7 @@ upwelling_detect_event <- function(df){
 # 
 MUR_upwell_base <- upwelling_detect_event(df = Final)
 # save(MUR_upwell_base, file = "Data/MUR_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
 km_func <- function(df){
   upwell_base <-  df%>% 
     mutate(distance_km = case_when(distance == "10000" ~ "10",

### 2019-11-02 SST_patterns.Rmd
updated
Update

@@
 # ANOVA
-- Relationship between duration, year and distance as a function of the mean intensity.
+- Relationship between duration, year and site as a function of the distance
 
@@
 anova_func <- function(df){
-  sites_aov <- aov(intensity_mean ~ site * duration * distance, data = df)
+  sites_aov <- aov(distance ~ duration * site * site, data = df)
 return(sites_aov)
@@
 
-SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
+MUR_anov <- anova_func(df = MUR_upwell_base)
+summary(MUR_anov)
+
+SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
 summary(SACTN_anov)
@@
 load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
 

### 2019-11-02 G1SST extraction.R

@@ -1,0 +1,167 @@
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+library(FNN)
+library(heatwaveR)
+
+rm(G1SST)
+ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/Extract 47.2"
+csvDir <- "/media/amieroh/Seagate EXpansion/Extract"
+
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+
+ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+strtDate <- str_sub(ncFirst, start = 1, end = 8)
+endDate <- str_sub(ncLast, start = 1, end = 8)
+
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+
+ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+  nc <- nc_open(ncFile)
+  pathLen <- nchar(paste0(ncDir, "/")) + 1
+  fNameStem <-
+    substr(basename(ncFile), 10, 58)
+  fDate <- substr(basename(ncFile), 1, 8)
+  sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+    round(4)
+  dimnames(sst) <- list(lon = nc$dim$lon$vals,
+                        lat = nc$dim$lat$vals)
+  nc_close(nc)
+  sst <- as_tibble(melt(sst, value.name = "temp"))
+  sst$t <- ymd(fDate)
+  na.omit(sst)
+  fwrite(sst,
+         file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
[…]

### 2019-11-03 G1SST extraction.R
Update

@@ -9,159 +9,154 @@ library(doMC); doMC::registerDoMC(cores = 7)
 library(FNN)
 library(heatwaveR)
 
-rm(G1SST)
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/Extract 47.2"
-csvDir <- "/media/amieroh/Seagate EXpansion/Extract"
+# rm(G1SST)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/Extract 90"
+# csvDir <- "/media/amieroh/Seagate EXpansion/Extract"
+# 
+# # #          1         2         3         4
+# # # 12345678901234567890123456789012345678901
+# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+# ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+# strtDate <- str_sub(ncFirst, start = 1, end = 8)
+# endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+# ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#   nc <- nc_open(ncFile)
+#   pathLen <- nchar(paste0(ncDir, "/")) + 1
+#   fNameStem <-
+#     substr(basename(ncFile), 10, 58)
+#   fDate <- substr(basename(ncFile), 1, 8)
+#   sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#     round(4)
+#   dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                         lat = nc$dim$lat$vals)
+#   nc_close(nc)
+#   sst <- as_tibble(melt(sst, value.name = "temp"))
+#   sst$t <- ymd(fDate)
+#   na.omit(sst)
+#   fwrite(sst,
+#          file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
[…]

### 2019-11-03 MUR.R

@@ -139,6 +139,7 @@ MUR_combined_2 <- rbind(MUR_fill, MUR2_fill)
 Final <- rbind(Final, MUR_combined_2)
 save(Final, file = "Data/Final.RData")
 
+load("Data/Final_RData")
 
 
 detect_event_custom <- function(df){
@@ -168,62 +169,3 @@ upwelling_detect_event <- function(df){
 # 
 MUR_upwell_base <- upwelling_detect_event(df = Final)
 # save(MUR_upwell_base, file = "Data/MUR_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-km_func <- function(df){
-  upwell_base <-  df%>% 
-    mutate(distance_km = case_when(distance == "10000" ~ "10",
-                                   distance == "20000" ~ "20",
-                                   distance  == "30000" ~"30",
-                                   distance  == "40000" ~ "40",
-                                   distance  == "50000" ~ "50"))
-}
-MUR_upwell_base <- km_func(df = MUR_upwell_base)
-MUR_upwell_base$distance_km <- as.numeric(MUR_upwell_base$distance_km)
-
-library(ggpubr)
-year_func <- function(df){
-  Filtered <- df %>%
-    filter(year(date_start) %in% seq(2011, 2013, 1))
-}
-
-MUR_filtered <- year_func(df = MUR_upwell_base)
-
-MUR_SEAPOINT <- MUR_filtered %>% 
-  filter(site == "Sea Point")
-MUR_SB <- MUR_filtered %>% 
-  filter(site == "Saldanha Bay")
-MUR_PN <- MUR_filtered %>% 
-  filter(site == "Port Nolloth")
-MUR_LB <- MUR_filtered %>% 
-  filter(site == "Lamberts Bay")
-
-plot_loli_func <- function(df){
-  ggplot(df, aes(x = date_peak, y = duration)) + 
-    geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
-    #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
-    facet_wrap(~distance_km, ncol = 2) +
-    scale_y_continuous(breaks = c(5,10,15,20,25,30),
-                       limits = c(1,30)) +
-    scale_x_date(limits = as.Date(c("2011-01-01","2014-01-01"))) +
[…]

### 2019-11-03 MUR.R
update

@@ -165,7 +165,7 @@ upwelling_detect_event <- function(df){
     filter(!is.na(exceedance)) %>%
     group_by(site, product, heading, distance, lon, lat) %>%
     group_modify(~detect_event_custom(.x))
-  }
+}
 # 
 MUR_upwell_base <- upwelling_detect_event(df = Final)
 # save(MUR_upwell_base, file = "Data/MUR_upwell_base.RData")

### 2019-11-03 G1SST extraction.R
cleaned

@@ -8,51 +8,51 @@ library(data.table)
 library(doMC); doMC::registerDoMC(cores = 7)
 library(FNN)
 library(heatwaveR)
-
+devtools::install_github("robwschlegel/heatwaveR")
 # rm(G1SST)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/Extract 90"
-# csvDir <- "/media/amieroh/Seagate EXpansion/Extract"
-# 
-# # #          1         2         3         4
-# # # 12345678901234567890123456789012345678901
-# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-# ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-# ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-# strtDate <- str_sub(ncFirst, start = 1, end = 8)
-# endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-# ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#   nc <- nc_open(ncFile)
-#   pathLen <- nchar(paste0(ncDir, "/")) + 1
-#   fNameStem <-
-#     substr(basename(ncFile), 10, 58)
-#   fDate <- substr(basename(ncFile), 1, 8)
-#   sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#     round(4)
-#   dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                         lat = nc$dim$lat$vals)
-#   nc_close(nc)
-#   sst <- as_tibble(melt(sst, value.name = "temp"))
-#   sst$t <- ymd(fDate)
-#   na.omit(sst)
-#   fwrite(sst,
-#          file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#          append = TRUE, col.names = FALSE)
-#   rm(sst)
-# }
-# 
[…]

### 2019-11-03 G1SST extraction.R
A couple of quick edits to get the code running

@@ -1,140 +1,144 @@
-library(stringr)
 library(tidyverse)
-library(reshape2)
+# library(stringr) # RWS: 'stringr' is loaded as part of 'tidyverse'
+# library(reshape2) # RWS: Better not to call this ecplicitly as it is pre-tidyverse and may ause issues
 library(ncdf4) # library for processing netCDFs
-library(plyr)
+# library(plyr) # RWS: Please never load plyr directly. You keep doing this...
 library(lubridate)
 library(data.table)
 library(doMC); doMC::registerDoMC(cores = 7)
 library(FNN)
 library(heatwaveR)
-devtools::install_github("robwschlegel/heatwaveR")
-# rm(G1SST)
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/Extract 90.2"
-csvDir <- "/media/amieroh/Seagate EXpansion/Extract"
-
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-
-ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-strtDate <- str_sub(ncFirst, start = 1, end = 8)
-endDate <- str_sub(ncLast, start = 1, end = 8)
-
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-
-ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-  nc <- nc_open(ncFile)
-  pathLen <- nchar(paste0(ncDir, "/")) + 1
-  fNameStem <-
-    substr(basename(ncFile), 10, 58)
-  fDate <- substr(basename(ncFile), 1, 8)
-  sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-    round(4)
-  dimnames(sst) <- list(lon = nc$dim$lon$vals,
[…]

### 2019-11-05 SST_patterns.Rmd
update

@@
 load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
 ```
@@
 - Relationship between duration, year and site as a function of the distance 
+The number of events per distance and then compare this at each distance
 
@@
 load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
 
@@
 
+G1SST_upwell_base <- km_func(df = G1SST_upwell_base)
+G1SST_upwell_base$distance_km <- as.numeric(G1SST_upwell_base$distance_km)
+
 library(ggpubr)
@@
 SACTN_filtered <- year_func(df = SACTN_upwell_base)
+MUR_filtered <- year_func(df = MUR_upwell_base)
+G1SST_filtered <- year_func(df = G1SST_upwell_base)
 
@@
 
-MUR_filtered <- year_func(df = MUR_upwell_base)
-
 MUR_SEAPOINT <- MUR_filtered %>% 
@@
   filter(site == "Lamberts Bay")
+###########
+####G1SST
+
+G1SST_SEAPOINT <- G1SST_filtered %>% 
+  filter(site == "Sea Point")
+G1SST_SB <- G1SST_filtered %>% 
+  filter(site == "Saldanha Bay")
+G1SST_PN <- G1SST_filtered %>% 
+  filter(site == "Port Nolloth")
+G1SST_LB <- G1SST_filtered %>% 
+  filter(site == "Lamberts Bay")
+
 ```
@@
 combined_MUR_pl_1 <- ggarrange(MUR_LB_pl, MUR_PN_pl)
+#######
+#G1SST
+
+(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
+(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
+(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
+(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
+
+OISST_SP_pl <- OISST_SP_pl +
+  ggtitle("D.") #Sea Point
+OISST_SB_pl <- OISST_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+OISST_PN_pl <- OISST_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+OISST_LB_pl <- OISST_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl)
+
+combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)
 

### 2019-11-05 netCDF2CSV_ERA5.R

@@ -20,118 +20,73 @@ library(doMC); doMC::registerDoMC(cores = 7)
 # # 12345678901234567890123456789012345678901
 # # wu_daily08-19.nc
 
-ncFile <- '/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate/wu_daily08-19.nc'  
-nc <- nc_open(ncFile)
-fNameStem <-
-  substr(basename(ncFile), 1, 13)
-u_10 <- ncvar_get(nc, varid = "u10") %>%
-  round(4)
-dimnames(u_10) <- list(lon = nc$dim$lon$vals,
-                         lat = nc$dim$lat$vals,
-                         time = nc$dim$time$vals)
-nc_close(nc)
-u_10 <- as_tibble(melt(u_10, value.name = "u_10"))
-u_10$time <- as.POSIXct(u_10$time, origin = "1970-01-01")
-
-
-
-###### v-wind 2008-2019
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # wu_daily08-19.nc
-
-ncFile <- '/home/amieroh/Documents/Data/Datasets/ERA5/u_v_seperate/wv_daily08-19.nc'
-nc <- nc_open(ncFile)
-fNameStem <-
-  substr(basename(ncFile), 1, 13)
-v_10 <- ncvar_get(nc, varid = "v10") %>%
-  round(4)
-dimnames(v_10) <- list(lon = nc$dim$lon$vals,
-                       lat = nc$dim$lat$vals,
-                       time = nc$dim$time$vals)
-nc_close(nc)
-v_10 <- as_tibble(melt(v_10, value.name = "v_10"))
-v_10$time <- as.POSIXct(v_10$time, origin = "1970-01-01")
-
-#####################################################################################
-### Extracting 1981 - 2007 u10
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # wind_daily81-99_165.nc
 
 ncFile <- '/home/amieroh/Documents/Data/Datasets/ERA5/wind_daily81-99_165.nc'
 nc <- nc_open(ncFile)
-fNameStem <-
-  substr(basename(ncFile), 1, 19)
+fNameStem <- substr(basename(ncFile), 1, 13)
 u_10 <- ncvar_get(nc, varid = "u10") %>%
   round(4)
[…]

### 2019-11-06 Extracting/G1SST and MUR extraction.Rmd
house keeping

--- 2019-10-26 G1SST and MUR extraction.Rmd
@@
 
-This scripts extracts the MUR and G1sst data
+Extracting MUR
 
 ```{r}
-bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
 library(stringr)
@@
 library(doMC); doMC::registerDoMC(cores = 7)
-ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/2010"
-csvDir <- "/home/amieroh/Documents/tester"
+```
 
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+```{r}
+
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+# 
+# #          1         2         3         4         5         6
+# # 1234567890123456789012345678901234567890123456789012345678901
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# 
+# 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+# fNameStem <-
+#   substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+```
+
+# Extracting G1SST
+
+```{r}
+ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+
 #          1         2         3         4         5         6
@@
 # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+ bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
 
-# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
@@
 
+Data analysis
+
 ```{r}

### 2019-11-07 SST_patterns.Rmd
progress

@@
 ```{r}
+#glm to determine if there is a difference in the number of events recorded at different at every distance
+
+load("Data/Count/OISST_count.RData")
+load("Data/Count/G1SST_count.RData")
+load("Data/Count/MUR_count.RData")
+load("Data/Count/CMC_count.RData")
+
+options(scipen = 999)
+anova_func <- function(df){
+  sites_aov <- aov(distance ~ duration * site * site, data = df)
+return(sites_aov)
+}
+
+OISST_anov <- anova_func(df = OISST_upwell_base)
+summary(OISST_anov)
+CMC_anov <- anova_func(df = CMC_upwell_base)
+summary(CMC_anov)
+
+MUR_anov <- anova_func(df = MUR_upwell_base)
+summary(MUR_anov)
+
+SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
+summary(SACTN_anov)
 
@@
 
-(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
-(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
-(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
-(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
-
-OISST_SP_pl <- OISST_SP_pl +
+(G1SST_SP_pl <- plot_loli_func(df = G1SST_SEAPOINT))
+(G1SST_SB_pl <- plot_loli_func(df = G1SST_SB))
+(G1SST_PN_pl <- plot_loli_func(df = G1SST_PN))
+(G1SST_LB_pl <- plot_loli_func(df = G1SST_LB))
+
+G1SST_SP_pl <- G1SST_SP_pl +
   ggtitle("D.") #Sea Point
-OISST_SB_pl <- OISST_SB_pl +
+G1SST_SB_pl <- G1SST_SB_pl +
   ggtitle("C.") # Saldanha Bay
-OISST_PN_pl <- OISST_PN_pl +
+G1SST_PN_pl <- G1SST_PN_pl +
   ggtitle("B.")        # Port Nolloth
-OISST_LB_pl <- OISST_LB_pl +
+G1SST_LB_pl <- G1SST_LB_pl +
   ggtitle("A.")#Lamberts Bay
 
-combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl)
-
-combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)
+combined_G1SST_pl_2 <- ggarrange(G1SST_SB_pl, G1SST_SP_pl)
+
+combined_G1SST_pl_1 <- ggarrange(G1SST_LB_pl, G1SST_PN_pl)
 

### 2019-11-19 SST_patterns.Rmd
update

@@
 
-# PLotting events
-
 Loading the data
+    This is the data obtained once the upwelling signals were filtered out
+
 ```{r}
@@
 ```
+
+# GLM - Determining if a change occur in the number of signals detected at different distancces from the coastline

### 2019-11-25 SST_patterns.Rmd
update
tips

@@
 ---
+# This script is for all the various datasets
 
@@
 
-# GLM - Determining if a change occur in the number of signals detected at different distancces from the coastline
+# GLM - Determining if a change occur in the number of signals detected at different distances from the coastline
+# Included a table in the paper
+  Table discuss/count the number of upwelling signals detected at the different distancees

### 2019-11-26 site_map.Rmd
new laptop commit

@@
 library(ggpubr); theme_set(theme_pubr())# library(devtools)
-# install_github("marchtaylor/sinkr")
+install_github("marchtaylor/sinkr")
 

### 2019-11-26 upwell_IDX.Rmd
new laptop commit

@@
 These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods.
-Upwelling is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.
+Upwelling is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation
 

### 2019-11-28 upwell_IDX.Rmd
coastR install

@@
 library(circular)
-# devtools::install_github("robwschlegel/coastR")
+devtools::install_github("robwschlegel/coastR")
 library(gridExtra)

### 2019-11-28 upwell_IDX.Rmd
update

@@
 library(circular)
-devtools::install_github("robwschlegel/coastR")
+##devtools::install_github("robwschlegel/coastR")
 library(gridExtra)

### 2019-11-29 upwell_IDX.Rmd
new laptop commit

--- 2019-11-28 upwell_IDX.Rmd
@@
 ```{r}
+# Installling andloading libraries
+
 library(circular)
-devtools::install_github("robwschlegel/coastR")
+## devtools::install_github("robwschlegel/coastR")
 library(gridExtra)

### 2019-11-30 wind_analyses.Rmd
edits

@@
 ---
+
+loading libraries
+
+```{r}
+library(tidyverse)
+library(ggpubr)
+library(zoo)
+library(lubridate)
+library(ggrepel)
+library(FNN)
+library(stringr)
+library(circular)
+library(broom)
+library(ggrepel)
+library(purrr)
+library(stlplus)
+```
 
@@
 ```
+
+## Seasonal wind patterns
+
+```{r}
+wind_months <- wind_daily %>% 
+  mutate(month = month(date, abbr = T, label = T),
+         year = year(date)) %>% 
+    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+
+save(wind_months, file = "Data/wind_months.RData")
+```

### 2019-11-30 wind_analyses.Rmd
update
updated commit

@@
 
-save(wind_months, file = "Data/wind_months.RData")
+# save(wind_months, file = "Data/wind_months.RData")
 ```
+
+# Question:
+  # Are there more SE winds in summer but comes later in January?I.e any changes in SE wind and changes in water temperature
+
+```{r}
+load("Data/SACTN_US.RData")
+load("Data/wind_months.RData")
+
+match_func <- function(df){
+  match <- SACTN_US  %>%  
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+wind_months <- wind_months %>% 
+  dplyr::rename(site = temp_sites)
+
+wind_match <- match_func(df = wind_months) 
+```

### 2019-11-30 wind_analyses.Rmd
update

@@
 library(stlplus)
+library(gridExtra)
 ```
@@
 ```
+
+# Visualisation - This does not show the changes per year but for all years
+
+```{r}
+source("Functions/wind.rose.R")
+
+wind_renamed <- wind_months %>% 
+  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>%
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ) %>% 
+  filter(spd > 0)
+
+p.wr2 <- plot.windrose(data = wind_renamed,
+              spd = "spd",
+              dir = "dir")
+p.wr3 <- p.wr2 + facet_wrap(.~ site) +
+  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
+p.wr3
+```

### 2019-12-03 SST_patterns.Rmd
some points

@@
 # Included a table in the paper
-  Table discuss/count the number of upwelling signals detected at the different distancees
+  Table discuss/count the number of upwelling signals detected at the different distances
+
+  - For some of the sites it is seen that fewer upwelling signals are visible at a further distance from the coastline
+  - When adding up all the signals observed at their respective distance it is seen that there are only slight differences in the amount of signals detected In some cases the signals detected at 10km from the coastline exceed that detected at 50km and visa versa

### 2019-12-04 wind_analyses.Rmd
update
Basic results to answer the questions at hand

@@
 # Question:
-  # Are there more SE winds in summer but comes later in January?I.e any changes in SE wind and changes in water temperature
+  # Are there more SE winds in summer but comes later in January? i.e any changes in SE wind and changes in water temperature
 
@@
 load("Data/wind_months.RData")
+
+wind_months <- wind_months %>% 
+  dplyr::rename(site = temp_sites)
 
@@
 
-wind_months <- wind_months %>% 
-  dplyr::rename(site = temp_sites)
-
 wind_match <- match_func(df = wind_months) 
@@
 
-# Visualisation - This does not show the changes per year but for all years
+# Visualisation - This does not show the changes per year but for all  also does not show any temps
 
@@
 
-wind_renamed <- wind_months %>% 
+wind_renamed <- wind_match %>% 
   mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>%
@@
 ```
+
+Are there more SE winds in summer, but come later in January? i.e any changes in SE wind and changes in water temperature
+To answer this first question we will take total counts of SE winds (180 - 270 degrees) for each year and counts for just summer (Jan - Mar).
+We will also take counts of each Summer month individually to see what the long-term trends are.
+Once we've pinned down any trends we will look at the same temporal agregations for temperatures to see what those trends are.
+With all of that determined we can then see what the relationship between the two is.
+
+```{r}
+# First filter out only the SE data
+SE_renamed <- wind_renamed %>% 
+  filter(dir >= 180, dir <= 270)
+
+# Then create diifferent temporal results
+SE_annual <- SE_renamed %>% 
+  group_by(site, year) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_summer <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_monthly <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season, month) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+# Plots
+## Annual count of SE wind
+ggplot(data = SE_annual, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual count of SE wind in Summer
+### The trends between annual and summer SE wind counts are remarkably similar
+ggplot(data = SE_summer, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Summer month count of SE winds
+ggplot(data = SE_monthly, aes(x = year, y = count)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# March SE winds have been increasing in count in Hout Bay, Lamberts Bay, and Sea Point
+# Feb winds are increasing faster in count in Saldanha Bay
+# All SE summer winds are decreasing in Port Nolloth
+
+## Annual mean SST
+ggplot(data = SE_annual, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual Summer mean SST
+ggplot(data = SE_summer, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+# The temperature trends for annual and summer means are very similar
+# THe annual mean increase at Port Nolloth is steeper than the Summer increase
+## Summer months mean SST
+ggplot(data = SE_monthly, aes(x = year, y = mean_temp)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# The trends in increasing temperatures for each month are similar at Lamberts Bay and Port Nolloth
+# The incrase in March temperatures is greater than the other months at Hout Bay
+# The decrease in temperatures per month at Saldanha Bay are similar per month
+# Sea Point shows a more intersting pattern with January temperatures decreasing, Feb nearly flat, and Mar increasing
+
+# Now having had a look at the data it appears there may be some relationships between SE wind and temperature
+# at Sea Point, and perhaps a negative relationship atPort Nolloth
+
+# This outputs the slope of the relationship and the correlation between
+# annual count of SE winds mean temperature 
+SE_annual %>% 
+  group_by(site) %>% 
+  do(mod = lm(mean_temp ~ count, data = .),
+     mod2 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")) %>%
+  mutate(Slope = summary(mod)$coeff[2],
+         Pearson = mod2[1]) %>%
+  select(-mod, -mod2)
+# Summer stats
+SE_summer %>% 
+  group_by(site, season) %>% 
+  do(mod = lm(mean_temp ~ count, data = .),
+     mod2 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")) %>%
+  mutate(Slope = summary(mod)$coeff[2],
+         Pearson = mod2[1]) %>%
+  select(-mod, -mod2)
+# Monthly summer stats
+SE_monthly %>% 
+  group_by(site, season, month) %>% 
+  do(mod = lm(mean_temp ~ count, data = .),
+     mod2 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")) %>%
+  mutate(Slope = summary(mod)$coeff[2],
+         Pearson = mod2[1]) %>%
+  select(-mod, -mod2)
+```
+
+It looks like there is generally a negative relationship between increasing summer winds and mean SST.

### 2019-12-04 wind_analyses.Rmd
Optimised a bit

--- 2019-11-30 wind_analyses.Rmd
@@
 ```
+
+Are there more SE winds in summer, but come later in January? i.e any changes in SE wind and changes in water temperature
+To answer this first question we will take total counts of SE winds (180 - 270 degrees) for each year and counts for just summer (Jan - Mar).
+We will also take counts of each Summer month individually to see what the long-term trends are.
+Once we've pinned down any trends we will look at the same temporal agregations for temperatures to see what those trends are.
+With all of that determined we can then see what the relationship between the two is.
+
+```{r}
+# First filter out only the SE data
+SE_renamed <- wind_renamed %>% 
+  filter(dir >= 180, dir <= 270)
+
+# Then create diifferent temporal results
+SE_annual <- SE_renamed %>% 
+  group_by(site, year) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_summer <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_monthly <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season, month) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+# Plots
+## Annual count of SE wind
+ggplot(data = SE_annual, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual count of SE wind in Summer
+### The trends between annual and summer SE wind counts are remarkably similar
+ggplot(data = SE_summer, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Summer month count of SE winds
+ggplot(data = SE_monthly, aes(x = year, y = count)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# March SE winds have been increasing in count in Hout Bay, Lamberts Bay, and Sea Point
+# Feb winds are increasing faster in count in Saldanha Bay
+# All SE summer winds are decreasing in Port Nolloth
+```
+
+```{r}
+## Annual mean SST
+ggplot(data = SE_annual, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual Summer mean SST
+ggplot(data = SE_summer, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+# The temperature trends for annual and summer means are very similar
+# THe annual mean increase at Port Nolloth is steeper than the Summer increase
+## Summer months mean SST
+ggplot(data = SE_monthly, aes(x = year, y = mean_temp)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# The trends in increasing temperatures for each month are similar at Lamberts Bay and Port Nolloth
+# The incrase in March temperatures is greater than the other months at Hout Bay
+# The decrease in temperatures per month at Saldanha Bay are similar per month
+# Sea Point shows a more intersting pattern with January temperatures decreasing, Feb nearly flat, and Mar increasing
+```
+
+```{r}
+# Now having had a look at the data it appears there may be some relationships between SE wind and temperature
+# at Sea Point, and perhaps a negative relationship atPort Nolloth
+
+# This outputs the slope of the relationship and the correlation between
+# annual count of SE winds mean temperature 
+slope_calc <- function(df){
+  df %>% 
+    mutate(row_num = 1:n()) %>% 
+    do(mod1 = lm(count ~ row_num, data = .),
+       mod2 = lm(mean_temp ~ row_num, data = .),
+       mod3 = lm(mean_temp ~ count, data = .),
+       mod4 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")[1]) %>% 
+    mutate(wind_slope = summary(mod1)$coeff[2],
+           temp_slope = summary(mod2)$coeff[2],
+           temp_wind_slope = summary(mod3)$coeff[2],
+           temp_wind_r = mod4[1],
+           temp_wind_r2 = glance(mod3)$adj.r.squared) %>%
+    select(-mod1, -mod2, -mod3, -mod4) %>% 
+    mutate_if(is.numeric, round, 2)
+}
+# glance(lm(mean_temp ~ count, data = SE_annual))
+
+SE_annual %>% 
+  group_by(site) %>% 
+  slope_calc()
+# Summer stats
+SE_summer %>% 
+  group_by(site, season) %>% 
+  slope_calc()
+# Monthly summer stats
+SE_monthly %>% 
+  group_by(site, season, month) %>% 
+  slope_calc()
+
+# There's a little something at Hout Bay, not much otherwise
+```

### 2019-12-16 SST_patterns.Rmd
fixed error

@@
 }
-# 
+
 # # Find the nearest pixels to each spot along the transect

### 2019-12-16 wind_analyses.Rmd
to do
progress

@@
 
-# save(wind_months, file = "Data/wind_months.RData")
+# save(wind_months, file = "Data/wind_m onths.RData")
 ```
@@
 match_func <- function(df){
-  match <- SACTN_US  %>%  
-  left_join(df, by = c("site", "date")) %>% 
+  match <- df  %>%  
+  left_join(wind_months, by = c("site", "date")) %>% 
   na.trim()
@@
 }
+wind_months <- site_list_sub %>% 
+  left_join(wind_months)
 
-wind_match <- match_func(df = wind_months) 
+wind_match_SACTN <- match_func(df = SACTN_US)
+
+load("Data/OISST.RData") # 2 decimal places
+OISST <- BC_avhrr_only_v2_Document_Document
+
+match_func <- function(df){
+  match <- df  %>%  
+  left_join(wind_months, by = c("lat", "date", "lon")) %>% 
+  na.trim()
+  return(match)
+}
+wind_match_OISST <- match_func(df = OISST)
+
 ```
@@
 ```
+
+Matching the OISST Data and CMC data to match the SACTN dataset
+
+```{r}
+unique_pixel <- OISST %>%
+  select(lon, lat) %>%
+  unique()
+
+# # Select 1 nearest pixel (k = 1)
+# # here we ue knnx to find the closes pixel to the insitu sites
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+                          query = as.matrix(site_list_sub[,5:6]), k = 1)
+
+# # Select SST pixels nearest to insitu sites
+pixel_match <- unique_pixel[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+  mutate(site = site_list_sub$site)
+
+combined_OISST_match <- OISST %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+rm(combined_OISST)
+
+combined_OISST_sites <- combined_OISST_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat"))
+
+OISST_final <- combined_OISST_sites %>% 
+  select(site, date, temp, lat, lon)
+
+############
+unique_pixel <- CMC %>%
+  select(lon, lat) %>%
+  unique()
+
+# # Select 1 nearest pixel (k = 1)
+# # here we ue knnx to find the closes pixel to the insitu sites
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+                          query = as.matrix(site_list_sub[,5:6]), k = 1)
+
+# # Select SST pixels nearest to insitu sites
+pixel_match <- unique_pixel[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+  mutate(site = site_list_sub$site)
+
+combined_CMC_match <- CMC %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+# rm(combined_OISST)
+
+combined_CMC_sites <- combined_CMC_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat"))
+
+CMC_final <- combined_CMC_sites %>% 
+  select(site, date, temp, lat, lon) 
+```
+
+```{r}
+
+```

### 2019-12-16 wind_analyses.Rmd
progress

@@
 ```{r}
+match_funcs <- function(df){
+  unique_pixel <- df %>%
+    select(lon, lat) %>%
+    unique()
+}
+
+match_OISST <- match_funcs(df = OISST)
+match_CMC <- match_funcs(df = CMC)
+
+# # Select 1 nearest pixel (k = 1)
+# # here we ue knnx to find the closes pixel to the insitu sites
+match_index_func <- function(df){
+  match_index <- knnx.index(data = as.matrix(df[,1:2]),
+                            query = as.matrix(site_list_sub[,5:6]), k = 1)
+}
+
+match_index_OISST <- match_funcs(df = match_OISST)
+match_index_CMC <- match_funcs(df = match_CMC)
+
+# # Select SST pixels nearest to insitu sites
+
+pixel_match_OISST <- match_OISST[match_index_OISST,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+  mutate(site = site_list_sub$site)
+
+pixel_match_CMC <- match_OISST[match_index_CMC,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>% 
+  mutate(site = site_list_sub$site)
+
+combined_OISST_match <- OISST %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+rm(combined_OISST)
+
+combined_OISST_sites <- combined_OISST_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat"))
 

### 2019-12-17 wind_analyses.Rmd
update

@@
 ```
-
-## Wind rose plots
-
-```{r}
-# Loading in the matched wind data
-load("Data/wind_match.RData")
-
-wind_match <- wind_match %>% 
-  drop_na()
-
-wave_daily_renamed <- wind_match %>% 
-  dplyr::rename(spd = mean_speed) %>%
-  dplyr::rename(dir = dir_circ) 
-source("Functions/wind.rose.R")
-p.wr2 <- plot.windrose(data = wave_daily_renamed,
-              spd = "spd",
-              dir = "dir")
-p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = "") +
-  theme(strip.text.x = element_text(size = 25))
-p.wr3  # Change colours and remove the NA values
-```
-
-## Seasonal wind patterns
-
-```{r}
-wind_months <- wind_daily %>% 
-  mutate(month = month(date, abbr = T, label = T),
-         year = year(date)) %>% 
-    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
-                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
-                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
-
-# save(wind_months, file = "Data/wind_m onths.RData")
-```
-
-# Question:
-  # Are there more SE winds in summer but comes later in January? i.e any changes in SE wind and changes in water temperature
-
-```{r}
-load("Data/SACTN_US.RData")
-load("Data/wind_months.RData")
-
-wind_months <- wind_months %>% 
-  dplyr::rename(site = temp_sites)
-
-match_func <- function(df){
-  match <- df  %>%  
-  left_join(wind_months, by = c("site", "date")) %>% 
-  na.trim()
-  return(match)
-}
-wind_months <- site_list_sub %>% 
-  left_join(wind_months)
-
-wind_match_SACTN <- match_func(df = SACTN_US)
-
-load("Data/OISST.RData") # 2 decimal places
-OISST <- BC_avhrr_only_v2_Document_Document
-
-match_func <- function(df){
-  match <- df  %>%  
-  left_join(wind_months, by = c("lat", "date", "lon")) %>% 
-  na.trim()
-  return(match)
-}
-wind_match_OISST <- match_func(df = OISST)
-
-```
-
-# Visualisation - This does not show the changes per year but for all  also does not show any temps
-
-```{r}
-source("Functions/wind.rose.R")
-
-wind_renamed <- wind_match %>% 
-  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>%
-  dplyr::rename(spd = mean_speed) %>%
-  dplyr::rename(dir = dir_circ) %>% 
-  filter(spd > 0)
-
-p.wr2 <- plot.windrose(data = wind_renamed,
-              spd = "spd",
-              dir = "dir")
-p.wr3 <- p.wr2 + facet_wrap(.~ site) +
-  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
-p.wr3
-```
-
-Are there more SE winds in summer, but come later in January? i.e any changes in SE wind and changes in water temperature
-To answer this first question we will take total counts of SE winds (180 - 270 degrees) for each year and counts for just summer (Jan - Mar).
-We will also take counts of each Summer month individually to see what the long-term trends are.
-Once we've pinned down any trends we will look at the same temporal agregations for temperatures to see what those trends are.
-With all of that determined we can then see what the relationship between the two is.
-
-```{r}
-# First filter out only the SE data
-SE_renamed <- wind_renamed %>% 
-  filter(dir >= 180, dir <= 270)
-
-# Then create diifferent temporal results
-SE_annual <- SE_renamed %>% 
-  group_by(site, year) %>% 
-  summarise(count = n(),
-            mean_dir = mean(dir, na.rm = T),
-            mean_temp = mean(temp, na.rm = T))
-SE_summer <- SE_renamed %>% 
-  filter(season == "Summer") %>% 
-  group_by(site, year, season) %>% 
-  summarise(count = n(),
-            mean_dir = mean(dir, na.rm = T),
-            mean_temp = mean(temp, na.rm = T))
-SE_monthly <- SE_renamed %>% 
-  filter(season == "Summer") %>% 
-  group_by(site, year, season, month) %>% 
-  summarise(count = n(),
-            mean_dir = mean(dir, na.rm = T),
-            mean_temp = mean(temp, na.rm = T))
-# Plots
-## Annual count of SE wind
-ggplot(data = SE_annual, aes(x = year, y = count)) +
-  geom_line() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~site)
-## Annual count of SE wind in Summer
-### The trends between annual and summer SE wind counts are remarkably similar
-ggplot(data = SE_summer, aes(x = year, y = count)) +
-  geom_line() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~site)
-## Summer month count of SE winds
-ggplot(data = SE_monthly, aes(x = year, y = count)) +
-  geom_line(aes(colour = month)) +
-  geom_smooth(aes(colour = month), method = "lm") +
-  facet_wrap(~site)
-# March SE winds have been increasing in count in Hout Bay, Lamberts Bay, and Sea Point
-# Feb winds are increasing faster in count in Saldanha Bay
-# All SE summer winds are decreasing in Port Nolloth
-```
-
-```{r}
-## Annual mean SST
-ggplot(data = SE_annual, aes(x = year, y = mean_temp)) +
-  geom_line() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~site)
-## Annual Summer mean SST
-ggplot(data = SE_summer, aes(x = year, y = mean_temp)) +
-  geom_line() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~site)
-# The temperature trends for annual and summer means are very similar
-# THe annual mean increase at Port Nolloth is steeper than the Summer increase
-## Summer months mean SST
-ggplot(data = SE_monthly, aes(x = year, y = mean_temp)) +
-  geom_line(aes(colour = month)) +
-  geom_smooth(aes(colour = month), method = "lm") +
-  facet_wrap(~site)
-# The trends in increasing temperatures for each month are similar at Lamberts Bay and Port Nolloth
-# The incrase in March temperatures is greater than the other months at Hout Bay
-# The decrease in temperatures per month at Saldanha Bay are similar per month
-# Sea Point shows a more intersting pattern with January temperatures decreasing, Feb nearly flat, and Mar increasing
-```
-
-```{r}
-# Now having had a look at the data it appears there may be some relationships between SE wind and temperature
-# at Sea Point, and perhaps a negative relationship atPort Nolloth
-
-# This outputs the slope of the relationship and the correlation between
-# annual count of SE winds mean temperature 
-slope_calc <- function(df){
-  df %>% 
-    mutate(row_num = 1:n()) %>% 
-    do(mod1 = lm(count ~ row_num, data = .),
-       mod2 = lm(mean_temp ~ row_num, data = .),
-       mod3 = lm(mean_temp ~ count, data = .),
-       mod4 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")[1]) %>% 
-    mutate(wind_slope = summary(mod1)$coeff[2],
-           temp_slope = summary(mod2)$coeff[2],
-           temp_wind_slope = summary(mod3)$coeff[2],
-           temp_wind_r = mod4[1],
-           temp_wind_r2 = glance(mod3)$adj.r.squared) %>%
-    select(-mod1, -mod2, -mod3, -mod4) %>% 
-    mutate_if(is.numeric, round, 2)
-}
-# glance(lm(mean_temp ~ count, data = SE_annual))
-
-SE_annual %>% 
-  group_by(site) %>% 
-  slope_calc()
-# Summer stats
-SE_summer %>% 
-  group_by(site, season) %>% 
-  slope_calc()
-# Monthly summer stats
-SE_monthly %>% 
-  group_by(site, season, month) %>% 
-  slope_calc()
-
-# There's a little something at Hout Bay, not much otherwise
-```
-
 Matching the OISST Data and CMC data to match the SACTN dataset
@@
 ```
+
+## Wind rose plots
+
+```{r}
+# Loading in the matched wind data
+load("Data/wind_match.RData")
+
+wind_match <- wind_match %>% 
+  drop_na()
+
+wave_daily_renamed <- wind_match %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ) 
+source("Functions/wind.rose.R")
+p.wr2 <- plot.windrose(data = wave_daily_renamed,
+              spd = "spd",
+              dir = "dir")
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = "") +
+  theme(strip.text.x = element_text(size = 25))
+p.wr3  # Change colours and remove the NA values
+```
+
+## Seasonal wind patterns
+
+```{r}
+wind_months <- wind_daily %>% 
+  mutate(month = month(date, abbr = T, label = T),
+         year = year(date)) %>% 
+    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+
+# save(wind_months, file = "Data/wind_m onths.RData")
+```
+
+# Question:
+  # Are there more SE winds in summer but comes later in January? i.e any changes in SE wind and changes in water temperature
+
+```{r}
+load("Data/SACTN_US.RData")
+load("Data/wind_months.RData")
+
+wind_months <- wind_months %>% 
+  dplyr::rename(site = temp_sites)
+
+match_func <- function(df){
+  match <- df  %>%  
+  left_join(wind_months, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+# wind_months <- site_list_sub %>% 
+#   left_join(wind_months)
+
+wind_match_SACTN <- match_func(df = SACTN_US)
+wind_match_OISST <- match_func(df = OISST_final)
+wind_match_CMC <- match_func(df = CMC_final)
+```
+
+# Visualisation - This does not show the changes per year but for all  also does not show any temps
+
+```{r}
+source("Functions/wind.rose.R")
+
+wind_renamed_func <- function(df){
+  wind_renamed <- df %>% 
+    mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>%
+    dplyr::rename(spd = mean_speed) %>%
+    dplyr::rename(dir = dir_circ) %>% 
+    filter(spd > 0)
+}
+
+wind_renamed_SACTN <- wind_renamed_func(df = wind_match_SACTN)
+wind_renmed_OISST <- wind_renamed_func(df = wind_match_OISST)
+wind_renamed_CMC <- wind_renamed_func(df = wind_match_CMC)
+
+p.wr2 <- plot.windrose(data = wind_renamed,
+              spd = "spd",
+              dir = "dir")
+p.wr3 <- p.wr2 + facet_wrap(.~ site) +
+  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
+p.wr3
+```
+
+```{r}
+# First filter out only the SE data
+SE_renamed <- wind_renamed %>% 
+  filter(dir >= 180, dir <= 270)
+# Then create diifferent temporal results
+SE_annual <- SE_renamed %>% 
+  group_by(site, year) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_summer <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+SE_monthly <- SE_renamed %>% 
+  filter(season == "Summer") %>% 
+  group_by(site, year, season, month) %>% 
+  summarise(count = n(),
+            mean_dir = mean(dir, na.rm = T),
+            mean_temp = mean(temp, na.rm = T))
+# Plots
+## Annual count of SE wind
+ggplot(data = SE_annual, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual count of SE wind in Summer
+### The trends between annual and summer SE wind counts are remarkably similar
+ggplot(data = SE_summer, aes(x = year, y = count)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Summer month count of SE winds
+ggplot(data = SE_monthly, aes(x = year, y = count)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# March SE winds have been increasing in count in Hout Bay, Lamberts Bay, and Sea Point
+# Feb winds are increasing faster in count in Saldanha Bay
+# All SE summer winds are decreasing in Port Nolloth
+```
+
+```{r}
+## Annual mean SST
+ggplot(data = SE_annual, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+## Annual Summer mean SST
+ggplot(data = SE_summer, aes(x = year, y = mean_temp)) +
+  geom_line() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+# The temperature trends for annual and summer means are very similar
+# THe annual mean increase at Port Nolloth is steeper than the Summer increase
+## Summer months mean SST
+ggplot(data = SE_monthly, aes(x = year, y = mean_temp)) +
+  geom_line(aes(colour = month)) +
+  geom_smooth(aes(colour = month), method = "lm") +
+  facet_wrap(~site)
+# The trends in increasing temperatures for each month are similar at Lamberts Bay and Port Nolloth
+# The incrase in March temperatures is greater than the other months at Hout Bay
+# The decrease in temperatures per month at Saldanha Bay are similar per month
+# Sea Point shows a more intersting pattern with January temperatures decreasing, Feb nearly flat, and Mar increasing
+```
+
+```{r}
+# Now having had a look at the data it appears there may be some relationships between SE wind and temperature
+# at Sea Point, and perhaps a negative relationship atPort Nolloth
+# This outputs the slope of the relationship and the correlation between
+# annual count of SE winds mean temperature 
+slope_calc <- function(df){
+  df %>% 
+    mutate(row_num = 1:n()) %>% 
+    do(mod1 = lm(count ~ row_num, data = .),
+       mod2 = lm(mean_temp ~ row_num, data = .),
+       mod3 = lm(mean_temp ~ count, data = .),
+       mod4 = cor(.$mean_temp, .$count, method = "pearson", use = "complete.obs")[1]) %>% 
+    mutate(wind_slope = summary(mod1)$coeff[2],
+           temp_slope = summary(mod2)$coeff[2],
+           temp_wind_slope = summary(mod3)$coeff[2],
+           temp_wind_r = mod4[1],
+           temp_wind_r2 = glance(mod3)$adj.r.squared) %>%
+    select(-mod1, -mod2, -mod3, -mod4) %>% 
+    mutate_if(is.numeric, round, 2)
+}
+# glance(lm(mean_temp ~ count, data = SE_annual))
+SE_annual %>% 
+  group_by(site) %>% 
+  slope_calc()
+# Summer stats
+SE_summer %>% 
+  group_by(site, season) %>% 
+  slope_calc()
+# Monthly summer stats
+SE_monthly %>% 
+  group_by(site, season, month) %>% 
+  slope_calc()
+# There's a little something at Hout Bay, not much otherwise
+```

### 2019-12-18 wind_analyses.Rmd
update

@@
 # First filter out only the SE data
-SE_renamed <- wind_renamed %>% 
+SE_renamed <-wind_renamed_SACTN %>% 
   filter(dir >= 180, dir <= 270)

### 2019-12-18 wind_analyses.Rmd
update

--- 2019-12-17 wind_analyses.Rmd
@@
 # First filter out only the SE data
-SE_renamed <- wind_renamed %>% 
+SE_renamed <-wind_renamed_SACTN %>% # Chnaged the names with the data 
   filter(dir >= 180, dir <= 270)

### 2020-01-02 site_map.Rmd
tips
removed
edits

@@
 pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
-
-# Site map
-# The insetting needs to be placed at appropriate location
 ```

### 2020-01-11 site_map.Rmd
update

@@
 
-site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
-  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
-   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
-                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
-  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
-  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
-  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
-  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
-  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
-  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
-  scale_fill_viridis(
-    option = "magma",
-    name = "Temperature (°C)",
-    alpha = 0.8) +
-  xlab("") + ylab("") +
-  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
-  #            size = 4, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
-             size =4, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
-             size = 4, angle = 302, colour = "black") +
-    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
-    #          size = 5, angle = 50, colour = "black") +
-  guides(fill = guide_colourbar()) +
-  scale_x_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°E", sep = "")) +
-    scale_y_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°S", sep = "")) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        legend.text = element_text(size = 14),
-        legend.title = element_text(size = 10),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5),
-        axis.text = element_text(size = 10)) +
-    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.minor = element_line(colour = "NA"),
-        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
-        legend.direction = "horizontal",
-        legend.justification = c(1,0),
-        legend.position = c(0.85, 0.70),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5))
+site
 site_map

### 2020-01-12 SST_patterns.Rmd
update

@@
 # 
-# # Calculate the upwelling event metrics
-# upwelling_detect_event <- function(df){
-#   upwell_base <- df %>% 
-#     dplyr::rename(t = date) %>% 
-#     group_by(site, product, heading, distance, lon, lat) %>% 
-#     group_modify(~ts2clm_custom(.x)) %>% 
-#     left_join(upwelling, by = c("site", "t")) %>%
-#     filter(!is.na(exceedance)) %>%
-#     group_by(site, product, heading, distance, lon, lat) %>% 
-#     group_modify(~detect_event_custom(.x))
-#   }
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+  upwell_base <- df %>%
+    dplyr::rename(t = date) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+  }
 # 

### 2020-01-12 site_map.Rmd
update

@@
 
-site
+site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
+  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
+   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
+                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
+  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 5) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
+  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
+  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
+  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
+  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
+  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
+  scale_fill_viridis(
+    option = "magma",
+    name = "Temperature (°C)",
+    alpha = 0.8) +
+  xlab("") + ylab("") +
+  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
+  #            size = 4, angle = 0, colour = "black") +
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
+             size =6, angle = 0, colour = "black") +
+    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
+             size = 6, angle = 302, colour = "black") +
+    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
+    #          size = 5, angle = 50, colour = "black") +
+  guides(fill = guide_colourbar()) +
+  scale_x_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°E", sep = "")) +
+    scale_y_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°S", sep = "")) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        legend.text = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5),
+        axis.text = element_text(size = 12)) +
+    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.minor = element_line(colour = "NA"),
+        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+        legend.direction = "horizontal",
+        legend.justification = c(1,0),
+        legend.position = c(0.85, 0.70),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5))
 site_map

### 2020-01-12 SST_patterns.Rmd
update

@@
 # 
-# OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-# save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-# CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-# save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
-# # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
-# SACTN_US <- SACTN_US %>%
-#   filter(site %in% selected_sites)
-# 
-# load("Data/site_list_v4.2.RData")
-# 
-# SACTN_upwell_base <- SACTN_US %>%
-#   left_join(site_list[,c(4, 5, 6)], by = "index")%>% 
-#     dplyr::rename(t = date) %>% 
-#     group_by(site, lon, lat) %>% 
-#     group_modify(~ts2clm_custom(.x)) %>% 
-#     left_join(upwelling, by = c("site", "t")) %>%
-#     filter(!is.na(exceedance)) %>%
-#     group_by(site, lon, lat) %>% 
-#     group_modify(~detect_event_custom(.x))
-# 
-# save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
+save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
+save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
+# 
+# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
+SACTN_US <- SACTN_US %>%
+  filter(site %in% selected_sites)
+
+load("Data/site_list_v4.2.RData")
+
+SACTN_upwell_base <- SACTN_US %>%
+  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
+    dplyr::rename(t = date) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+
+save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
 ```

### 2020-01-17 site_map.Rmd
update

--- 2020-01-11 site_map.Rmd
@@
 library(ggpubr); theme_set(theme_pubr())# library(devtools)
-install_github("marchtaylor/sinkr")
+# install_github("marchtaylor/sinkr")
 
@@
   site <- site_transects[i,]
-  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+  # site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+  site_pixel <- transect.pixel(site, c(10000, 30000, 50000))
   site_pixels <- rbind(site_pixels, site_pixel)
@@
 
+# New site pixels
+save(site_pixels, file = "Data/site_pixels_new.RData")
+load("Data/site_pixels_new.RData")
 # save(site_pixels, file = "Data/site_pixels.RData")
-load("Data/site_pixels.RData")
+# load("Data/site_pixels.RData")
 
@@
 load("Data/site_list_sub.Rdata")
-load("Data/site_pixels.RData")
+load("Data/site_pixels_new.RData")
+# load("Data/site_pixels.RData")
 load("MUR.RData")
@@
 site_list_sub <- site_list_sub[-5,]
+
+site_list_sub_map <- site_list_sub %>% 
+  select(site,lat,lon)
 
@@
 
-site
+site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
+  # geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
+   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
+                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
+  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 5) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_label_repel(data =site_list_sub_map, aes(x = lon, y = lat, label = site), 
+                     size = 4, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
+                     segment.alpha = 0.4, force = 0.1) +
+  coord_equal(xlim = c(12, 23), ylim = c(-36,-26), expand = 0) +
+  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
+  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
+  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
+  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
+  scale_fill_viridis(
+    option = "magma",
+    name = "Temperature (°C)",
+    alpha = 0.8) +
+  xlab("") + ylab("") +
+  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
+  #            size = 4, angle = 0, colour = "black") +
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
+             size =6, angle = 0, colour = "black") +
+    annotate("text", label = "Benguela", x = 14.0, y = -29.5,
+             size = 6, angle = 302, colour = "black") +
+    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
+    #          size = 5, angle = 50, colour = "black") +
+  guides(fill = guide_colourbar()) +
+  scale_x_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°E", sep = "")) +
+    scale_y_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°S", sep = "")) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        legend.text = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5),
+        axis.text = element_text(size = 12)) +
+    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.minor = element_line(colour = "NA"),
+        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+        legend.direction = "horizontal",
+        legend.justification = c(1,0),
+        legend.position = c(0.85, 0.70),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5))
 site_map
 
-pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
+# pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
+# 
+# 
+# africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
+#   theme_bw() + 
+#   coord_equal() + 
+#   geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
+#   geom_polygon(data = sa_provinces_new, (aes(group = group))) +
+#   annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
+#   theme(panel.border = element_rect(colour = "black", size = 0.4), 
+#         plot.background = element_blank(), 
+#         axis.ticks = element_blank(), 
+#         axis.text = element_blank(), 
+#         axis.title = element_blank(), 
+#         panel.grid.major = element_blank(), 
+#         panel.grid.minor = element_blank()) +
+#   coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
+# africa_map
+# 
+# # More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html
+# 
+# vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
+# vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
+# print(site_map, vp = vp2)
+# print(africa_map, vp = vp1)
+# dev.off()
+# pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
+```
 
-africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
+# SIte map 2
+
+```{r}
+
+site_list_sub_map <- site_list_sub %>% 
+  select(site,lat,lon)
+
+mapping_plot <- function(df, xy = "salmon"){
+  mapping <- df %>% 
+    ggplot(aes(x = lon, y = lat)) +
+    geom_polygon(data = south_africa_coast, aes(group = group), fill = "white") +
+    coord_cartesian(xlim = c(12, 36), ylim = c(-32, -20)) +
+    geom_path(data = sa_provinces_new, aes(group = group)) +
+    geom_point(data = df, colour = xy, size = 4)  +
+    geom_label_repel(data =df, aes(x = lon, y = lat, label = site), 
+                     size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
+                     segment.alpha = 0.4, force = 0.1) +
+    coord_equal(xlim = c(15, 35), ylim = c(-37, -27)) +
+    annotate("text", label = "INDIAN\nOCEAN", x = 35.00, y = -34.0, size = 6.0, angle = 0, colour = "black") +
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.00, y = -37.0, size = 6.0, angle = 0, colour = "black") +
+    # geom_segment(aes(x = 17.2, y = -32.6, xend = 15.2, yend = -29.5),
+    #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
+    #   annotate("text", label = "Benguela", x = 14.0, y = -31.8, size = 3.5, angle = 298, colour = "black") +
+    #   geom_segment(aes(x = 33, y = -29.5, xend = 29.8, yend = -33.0),
+    #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
+    annotate("text", label = "Agulhas", x = 31.7, y = -31.7, size = 7.5, angle = 53, colour = "black") +
+    scale_x_continuous(labels = scales::unit_format(unit = "°E", sep = "")) +
+    scale_y_continuous(labels = scales::unit_format(unit = "°S", sep = "")) +
+    labs(x = NULL, y = NULL) +
+    coord_cartesian(xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5), expand = F) +
+    scaleBar(lon = 32.0, lat = -38.7, distanceLon = 200, distanceLat = 50, 
+             distanceLegend = 90, dist.unit = "km", arrow.length = 100,
+             arrow.distance = 130, arrow.North.size = 3) +  
+    coord_equal() +
+    theme(aspect.ratio = 1)  +
     theme_bw() +
-  coord_equal() + 
-  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
-  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
-  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
-  theme(panel.border = element_rect(colour = "black", size = 0.4), 
-        plot.background = element_blank(), 
-        axis.ticks = element_blank(), 
-        axis.text = element_blank(), 
-        axis.title = element_blank(), 
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank()) +
-  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
-africa_map
+    theme(panel.border = element_rect(fill = NA, colour = "black", size = 1),
+          axis.text = element_text(colour = "black", size = 22),
+          axis.title = element_text(colour = "black", size = 22))
+  return(mapping)
+}
 
-# More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html
+Map <- mapping_plot(df = site_list_sub_map, xy = "salmon")
 
-vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
-vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
-print(site_map, vp = vp2)
-print(africa_map, vp = vp1)
-dev.off()
-pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
 ```

### 2020-01-17 site_map.Rmd
update

@@
   site <- site_transects[i,]
-  # site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+  #site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
   site_pixel <- transect.pixel(site, c(10000, 30000, 50000))
@@
 # New site pixels
-save(site_pixels, file = "Data/site_pixels_new.RData")
-load("Data/site_pixels_new.RData")
+# save(site_pixels, file = "Data/site_pixels_new.RData")
+# load("Data/site_pixels_new.RData")
 # save(site_pixels, file = "Data/site_pixels.RData")
-# load("Data/site_pixels.RData")
+ load("Data/site_pixels.RData")
 
@@
 load("Data/site_pixels_new.RData")
-# load("Data/site_pixels.RData")
+
+#load("Data/site_pixels.RData")
 load("MUR.RData")
@@
 
-site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
+site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + 
   # geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
@@
   geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 5) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
   geom_label_repel(data =site_list_sub_map, aes(x = lon, y = lat, label = site), 
-                     size = 4, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
+                     size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2, 
                      segment.alpha = 0.4, force = 0.1) +

### 2020-01-18 site_map.Rmd
update

@@
 
-site_list_sub_map <- site_list_sub %>% 
-  select(site,lat,lon)
-
-mapping_plot <- function(df, xy = "salmon"){
-  mapping <- df %>% 
-    ggplot(aes(x = lon, y = lat)) +
-    geom_polygon(data = south_africa_coast, aes(group = group), fill = "white") +
-    coord_cartesian(xlim = c(12, 36), ylim = c(-32, -20)) +
-    geom_path(data = sa_provinces_new, aes(group = group)) +
-    geom_point(data = df, colour = xy, size = 4)  +
-    geom_label_repel(data =df, aes(x = lon, y = lat, label = site), 
-                     size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
-                     segment.alpha = 0.4, force = 0.1) +
-    coord_equal(xlim = c(15, 35), ylim = c(-37, -27)) +
-    annotate("text", label = "INDIAN\nOCEAN", x = 35.00, y = -34.0, size = 6.0, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.00, y = -37.0, size = 6.0, angle = 0, colour = "black") +
-    # geom_segment(aes(x = 17.2, y = -32.6, xend = 15.2, yend = -29.5),
-    #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
-    #   annotate("text", label = "Benguela", x = 14.0, y = -31.8, size = 3.5, angle = 298, colour = "black") +
-    #   geom_segment(aes(x = 33, y = -29.5, xend = 29.8, yend = -33.0),
-    #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
-    annotate("text", label = "Agulhas", x = 31.7, y = -31.7, size = 7.5, angle = 53, colour = "black") +
-    scale_x_continuous(labels = scales::unit_format(unit = "°E", sep = "")) +
-    scale_y_continuous(labels = scales::unit_format(unit = "°S", sep = "")) +
-    labs(x = NULL, y = NULL) +
-    coord_cartesian(xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5), expand = F) +
-    scaleBar(lon = 32.0, lat = -38.7, distanceLon = 200, distanceLat = 50, 
-             distanceLegend = 90, dist.unit = "km", arrow.length = 100,
-             arrow.distance = 130, arrow.North.size = 3) +  
-    coord_equal() +
-    theme(aspect.ratio = 1)  +
-    theme_bw() +
-    theme(panel.border = element_rect(fill = NA, colour = "black", size = 1),
-          axis.text = element_text(colour = "black", size = 22),
-          axis.title = element_text(colour = "black", size = 22))
-  return(mapping)
-}
-
-Map <- mapping_plot(df = site_list_sub_map, xy = "salmon")
+# site_list_sub_map <- site_list_sub %>% 
+#   select(site,lat,lon)
+# 
+# mapping_plot <- function(df, xy = "salmon"){
+#   mapping <- df %>% 
+#     ggplot(aes(x = lon, y = lat)) +
+#     geom_polygon(data = south_africa_coast, aes(group = group), fill = "white") +
+#     coord_cartesian(xlim = c(12, 36), ylim = c(-32, -20)) +
+#     geom_path(data = sa_provinces_new, aes(group = group)) +
+#     geom_point(data = df, colour = xy, size = 4)  +
+#     geom_label_repel(data =df, aes(x = lon, y = lat, label = site), 
+#                      size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
+#                      segment.alpha = 0.4, force = 0.1) +
+#     coord_equal(xlim = c(15, 35), ylim = c(-37, -27)) +
+#     annotate("text", label = "INDIAN\nOCEAN", x = 35.00, y = -34.0, size = 6.0, angle = 0, colour = "black") +
+#     annotate("text", label = "ATLANTIC\nOCEAN", x = 16.00, y = -37.0, size = 6.0, angle = 0, colour = "black") +
+#     # geom_segment(aes(x = 17.2, y = -32.6, xend = 15.2, yend = -29.5),
+#     #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
+#     #   annotate("text", label = "Benguela", x = 14.0, y = -31.8, size = 3.5, angle = 298, colour = "black") +
+#     #   geom_segment(aes(x = 33, y = -29.5, xend = 29.8, yend = -33.0),
+#     #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
+#     annotate("text", label = "Agulhas", x = 31.7, y = -31.7, size = 7.5, angle = 53, colour = "black") +
+#     scale_x_continuous(labels = scales::unit_format(unit = "°E", sep = "")) +
+#     scale_y_continuous(labels = scales::unit_format(unit = "°S", sep = "")) +
+#     labs(x = NULL, y = NULL) +
+#     coord_cartesian(xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5), expand = F) +
+#     scaleBar(lon = 32.0, lat = -38.7, distanceLon = 200, distanceLat = 50, 
+#              distanceLegend = 90, dist.unit = "km", arrow.length = 100,
+#              arrow.distance = 130, arrow.North.size = 3) +  
+#     coord_equal() +
+#     theme(aspect.ratio = 1)  +
+#     theme_bw() +
+#     theme(panel.border = element_rect(fill = NA, colour = "black", size = 1),
+#           axis.text = element_text(colour = "black", size = 22),
+#           axis.title = element_text(colour = "black", size = 22))
+#   return(mapping)
+# }
+# 
+# Map <- mapping_plot(df = site_list_sub_map, xy = "salmon")
 

### 2020-01-18 SST_patterns.Rmd
UPDATE

@@
 load("Data/G1SST_upwell_base.RData")
+
+# Removing the distance of 20 and 40kms
+library(dplyr)
+removing_distance_func <- function(df){
+  removing_dist_func<- df %>% 
+       filter(!distance %in% c(20000,40000))
+}
+
+OISST_final <- removing_distance_func(df = OISST_upwell_base)
+G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+MUR_final <- removing_distance_func(df = MUR_upwell_base)
+CMC_final <- removing_distance_func(df = CMC_upwell_base)
+
+# save(OISST_final, file = "Data/OISST_final.RData")
+# save(G1SST_final, file = "Data/G1SST_final.RData")
+# save(MUR_final, file = "Data/MUR_final.RData")
+# save(CMC_final, file = "Data/CMC_final.RData")
 ```

### 2020-01-18 site_map.Rmd
UPDATE

@@
                      segment.alpha = 0.4, force = 0.1) +
-  coord_equal(xlim = c(12, 23), ylim = c(-36,-26), expand = 0) +
+  coord_equal(xlim = c(13, 23), ylim = c(-36,-26), expand = 0) +
   # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
@@
   #            size = 4, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 15.00, y = -33.0,
              size =6, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 14.0, y = -29.5,
+    annotate("text", label = "Benguela", x = 14.5, y = -30,
              size = 6, angle = 302, colour = "black") +

### 2020-01-19 SST_patterns.Rmd
update

@@
   - When adding up all the signals observed at their respective distance it is seen that there are only slight differences in the amount of signals detected In some cases the signals detected at 10km from the coastline exceed that detected at 50km and visa versa
+
+## Number of events detected
+
+Using the start date and getting the month and year into a column. The month and year will be on the xaxis
+Then count the number of times per month at different years signals are detected
+
+```{r}
+load("Data/OISST_final.RData")
+load("Data/CMC_final.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/G1SST_final.RData")
+
+month_year_func <- function(df){
+  month_year_func<- df %>% 
+    mutate(month = month(date_start, abbr = T, label = T)) %>% 
+    separate(date_start, into = c("year", "month_no")) %>% 
+    unite(month_year, month,year)
+    # Now need to get the year into a seperate column from the start date. So now date_start is split in the dataset
+}
+
+OISST_prep <- month_year_func(df = OISST_final)
+CMC_prep <- month_year_func(df = CMC_final)
+G1SST_prep <- month_year_func(df = G1SST_final)
+MUR_prep <- month_year_func(df = MUR_final)
+
+# Combined dataset
+# Combined_frequency <- read_excel("Data/Combined_frequency.xlsx")
+# save(Combined_frequency, file = "Data/Combined_frequency.RData")
+```
+
+# Plotting
+
+```{r}
+load("Data/Combined_frequency.RData")
+```

### 2020-01-19 SST_patterns.Rmd
update

@@
 load("Data/Combined_frequency.RData")
-```
+
+# Does not accurately show the months
+# Ned to include distance and product
+
+ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+  geom_line(aes(colour = product)) +
+  geom_smooth(aes(colour = product), method = "lm") +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+  geom_point(aes(colour = distance)) +
+  geom_smooth(aes(colour = distance), method = "lm") +
+  facet_wrap(~site)+
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+####
+
+combined_frequency_split <- Combined_frequency %>% 
+  separate(date, into = c("month", "year"))
+
+# ggplot(data = combined_frequency_split, aes(x = year, y = number_of_signals)) +
+#   geom_line(aes(colour = month)) +
+#   geom_smooth(aes(colour = month), method = "lm") +
+#   facet_wrap(~site)
+```
+
+#Pairwise correlations
+Comparing the relaionship of variables within the same sample
+
+  -Correlate the 1st pixel with the second pixel and the first pixel with the last pixel
+  -Correlation of the number of frequencies at different distances at each site
+
+```{r}
+cor.test(x = Combined_frequency$distance, Combined_frequency$number_of_signals,
+    use = "everything", method = "pearson")
+
+slope_calc <- function(df){
+  df %>% 
+    mutate(row_num = 1:n()) %>% 
+    do(mod1 = lm(number_of_signals ~ row_num, data = .),
+       mod2 = lm(distance ~ row_num, data = .),
+       mod3 = lm(distance ~ number_of_signals, data = .),
+       mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
+    mutate(distance_slope = summary(mod1)$coeff[2],
+           signal_slope = summary(mod2)$coeff[2],
+           signal_distance_slope = summary(mod3)$coeff[2],
+           signal_distance_r = mod4[1],
+           signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
+    select(-mod1, -mod2, -mod3, -mod4) %>% 
+    mutate_if(is.numeric, round, 2)
+}
+
+distance_corr <- Combined_frequency %>% 
+  group_by(site) %>% 
+  slope_calc()
+
+```
+
+#
+
+```{r}
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
+
+    #Why are there no signals in 2011 in the SACTN data?
+    # Why are some signals detected in the SACTN data? Is this because of the intensity of the signals?
+
+# Reasons why upwelling might be detected at 10km rather than 50km
+    # Dome in the isotherm blocking upwelling to reach the surface waters but does occur deeper down
+
+```

### 2020-01-19 SST_patterns.Rmd
update

@@
     # Dome in the isotherm blocking upwelling to reach the surface waters but does occur deeper down
-
-```
+# (Figures folder - ig 2 and figure 10)
+```

### 2020-01-20 SST_patterns.Rmd
update
edit

@@
   -Correlate the 1st pixel with the second pixel and the first pixel with the last pixel
-  -Correlation of the number of frequencies at different distances at each site
+  -Correlation between satellite data and number of signals at the different distances from the coastline
 
@@
   slope_calc()
-
 ```

### 2020-01-20 SST_patterns.Rmd
update
Added date column

@@
 
+Combined_frequency <- Combined_frequency %>% 
+  mutate(t = parse_date(date, format = "%b_%Y"),
+         distance = factor(distance))
+
 # Does not accurately show the months
-# Ned to include distance and product
-
-ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+# Plot need to include month_year, number of signals, distance and product ()
+
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
   geom_line(aes(colour = product)) +
@@
 
-ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
+  geom_line(aes(colour = distance)) +
   geom_point(aes(colour = distance)) +

### 2020-01-20 SST_patterns.Rmd
Smol change

@@
   geom_smooth(aes(colour = product), method = "lm") +
-  facet_wrap(~site) +
-  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+  labs(x = NULL) +
+  facet_wrap(~site) #+
+  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
 
@@
   geom_smooth(aes(colour = distance), method = "lm") +
-  facet_wrap(~site)+
-  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+  labs(x = NULL) +
+  facet_wrap(~site) #+
+  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
 

### 2020-01-20 SST_patterns.Rmd
merge

--- 2020-01-20 SST_patterns.Rmd
@@
   theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
-
 ####

### 2020-01-20 SST_patterns.Rmd
Testing

--- 2020-01-20 SST_patterns.Rmd
@@
 library(mapproj)
+library(broom)
 # library(ncdf4) # This was used to process NetCDF files in an earlier version
@@
 
-Combined_frequency <- Combined_frequency %>% 
-  mutate(t = parse_date(date, format = "%b_%Y"),
-         distance = factor(distance))
+# Combined_frequency <- Combined_frequency %>% 
+  # mutate(t = parse_date(date, format = "%b_%Y"),
+         # distance = factor(distance))
 
@@
 distance_corr <- Combined_frequency %>% 
-  group_by(site) %>% 
+  group_by(site, product) %>% 
   slope_calc()

### 2020-01-20 SST_patterns.Rmd
speading

@@
 
-# Combined_frequency <- Combined_frequency %>% 
-  # mutate(t = parse_date(date, format = "%b_%Y"),
-         # distance = factor(distance))
+Combined_frequency <- Combined_frequency %>%
+mutate(t = parse_date(date, format = "%b_%Y"),
+distance = factor(distance))
 
@@
   geom_smooth(aes(colour = distance), method = "lm") +
-  labs(x = NULL) +
+  # labs(x = NULL) +
   facet_wrap(~site) #+
-  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
 
@@
 ```{r}
-cor.test(x = Combined_frequency$distance, Combined_frequency$number_of_signals,
-    use = "everything", method = "pearson")
+
+# chi squared test
+
+combined_freq_spread <- spread(Combined_frequency, distance, number_of_signals)
 

### 2020-01-20 SST_patterns.Rmd
Added correlation results by site and product comparing different distance pixels and the count of events within them

@@
 
-combined_freq_spread <- spread(Combined_frequency, distance, number_of_signals)
+load("Data/Combined_frequency.RData")
+
+combined_freq_spread <- pivot_wider(Combined_frequency, names_prefix = "dist_",
+                                    names_from = distance, values_from = number_of_signals)
 
@@
   df %>% 
-    mutate(row_num = 1:n()) %>% 
-    do(mod1 = lm(number_of_signals ~ row_num, data = .),
-       mod2 = lm(distance ~ row_num, data = .),
-       mod3 = lm(distance ~ number_of_signals, data = .),
-       mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
-    mutate(distance_slope = summary(mod1)$coeff[2],
-           signal_slope = summary(mod2)$coeff[2],
-           signal_distance_slope = summary(mod3)$coeff[2],
-           signal_distance_r = mod4[1],
-           signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
-    select(-mod1, -mod2, -mod3, -mod4) %>% 
+    # mutate(row_num = 1:n()) %>% 
+    # do(mod1 = lm(number_of_signals ~ row_num, data = .),
+       # mod2 = lm(distance ~ row_num, data = .),
+       # mod3 = lm(distance ~ number_of_signals, data = .),
+       # mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
+        # mutate(distance_slope = summary(mod1)$coeff[2],
+        #    signal_slope = summary(mod2)$coeff[2],
+        #    signal_distance_slope = summary(mod3)$coeff[2],
+        #    signal_distance_r = mod4[1],
+        #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
+        # select(-mod1, -mod2, -mod3, -mod4) %>% 
+    do(mod1 = cor(.$dist_10, .$dist_30, method = "pearson", use = "complete.obs"),
+       mod2 = cor(.$dist_10, .$dist_50, method = "pearson", use = "complete.obs")) %>%
+    mutate(dist_10_vs_30_r = mod1[1],
+           dist_10_vs_50_r = mod2[1]) %>% 
+    # select(dist_10_vs_30_r, dist_10_vs_50_r) %>% 
+    select(-mod1, -mod2) %>% 
     mutate_if(is.numeric, round, 2)
@@
 
-distance_corr <- Combined_frequency %>% 
+# test <- cor(combined_freq_spread$`10`, combined_freq_spread$`30`)
+
+distance_corr <- combined_freq_spread %>% 
   group_by(site, product) %>% 

### 2020-01-20 SST_patterns.Rmd
boxplot

@@
 
+ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+  geom_boxplot(aes(colour = product)) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
+  #geom_smooth(aes(colour = product), method = "lm") +
+  #labs(x = NULL) +
+  facet_wrap(~site)
+
 # Does not accurately show the months
@@
 ```
-
-#
-
-```{r}
-load("Data/OISST_upwell_base.RData")
-load("Data/CMC_upwell_base.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-load("Data/G1SST_upwell_base.RData")
-
-    #Why are there no signals in 2011 in the SACTN data?
-    # Why are some signals detected in the SACTN data? Is this because of the intensity of the signals?
-
-# Reasons why upwelling might be detected at 10km rather than 50km
-    # Dome in the isotherm blocking upwelling to reach the surface waters but does occur deeper down
-# (Figures folder - ig 2 and figure 10)
-```

### 2020-01-21 SST_patterns.Rmd
graphs

@@
 
-ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+ggplot(data = Combined_frequency, aes(x = distance, y = number_of_signals)) +
   geom_boxplot(aes(colour = product)) +
@@
   facet_wrap(~site)
+
+ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals, colour = product)) +
+geom_point(aes(size = distance)) +
+geom_smooth(method = "lm", size = 1.2)
+
+ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals, colour = product)) +
+geom_bubble(aes(size = distance)) +
+geom_smooth(method = "lm", size = 1.2)
 
@@
 ```
+
+```{r}
+library(hrbrthemes)
+library(viridis)
+
+Combined_frequency <- Combined_frequency %>%
+  mutate(t = parse_date(date, format = "%b_%Y"),
+         distance = factor(distance))
+
+ggplot(Combined_frequency, aes(x=t, y=number_of_signals, size=distance, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+### Without time
+
+ggplot(Combined_frequency, aes(x=distance, y= number_of_signals, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+```

### 2020-01-24 SST_patterns.Rmd
Update

@@
 
-ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals, colour = product)) +
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals, colour = product)) +
 geom_point(aes(size = distance)) +
-geom_smooth(method = "lm", size = 1.2)
-
-ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals, colour = product)) +
-geom_bubble(aes(size = distance)) +
-geom_smooth(method = "lm", size = 1.2)
+geom_smooth(method = "lm", size = 1.2)+
+theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
 
@@
 ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
+  geom_line(aes(colour = distance)) +
+  geom_point(aes(colour = distance)) +
+  geom_smooth(aes(colour = distance), method = "lm") +
+  # labs(x = NULL) +
+  facet_wrap(~site) #+
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+  ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
   geom_line(aes(colour = distance)) +
@@
 
+ggplot(Combined_frequency, aes(x=date, y=number_of_signals, size=distance, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
 ### Without time
@@
 ```
+
+# Histogram
+
+```{r}
+
+combined_frequency_season <- Combined_frequency %>% 
+  mutate(month = month(date, abbr = T, label = T),
+         year = year(date)) %>% 
+  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+
+CMC_dur <- read_excel("Data/CMC_dur.xlsx")
+
+ggplot(data = CMC_dur, aes(x = duration)) +
+  geom_histogram(aes(fill = product)) +
+  facet_wrap(~site)
+
+```

### 2020-01-26 SST_patterns.Rmd
graphs

@@
 ggplot(data = CMC_dur, aes(x = duration)) +
-  geom_histogram(aes(fill = product)) +
-  facet_wrap(~site)
-
-```
+  geom_histogram(aes(fill = product),  
+                       position="identity", alpha=0.6) +
+  facet_wrap(~site) +
+theme_bw()
+
+ggplot(CMC_dur, aes(duration, colour = product)) +
+  geom_freqpoly() +
+  facet_wrap(~site) +
+  theme_bw()
+```

### 2020-01-29 SST_patterns.Rmd
update

@@
   geom_histogram(aes(fill = product),  
-                       position="identity", alpha=0.6) +
+                       position = "dodge", alpha=0.6) +
+   labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
   facet_wrap(~site) +

### 2020-01-31 SST_patterns.Rmd
intensity

@@
 load("Data/G1SST_final.RData")
+---
+title: "SST_patterns"
+author: "Amieroh Abrahams"
+date: "19 August 2019"
+output: html_document
+---
+# This script is for all the various datasets
+
+# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
+
+```{r}
+knitr::opts_chunk$set(
+  comment = "R>",
+  warning = FALSE,
+  message = FALSE 
+)
+#update.packages() 
+library(tidyverse)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(scales)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(broom)
+# library(ncdf4) # This was used to process NetCDF files in an earlier version
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+library(ncdf4) # library for processing netCDFs
+library(data.table)
+# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
+library(heatwaveR)
+```
+
+# Loading in all the data created using the code bellow
+
+```{r}
+load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
+load("Data/site_pixels.RData") # 5 decimal places
+load("Data/OISST.RData") # 2 decimal places
+OISST <- BC_avhrr_only_v2_Document_Document 
+rm(BC_avhrr_only_v2_Document_Document ); gc()
+load("Data/CMC.RData") # 1decimal places
+```
+
+```{r}
+# SACTN_US <- SACTN_US %>% 
+#   dplyr::rename(in_situ_temp = temp)
+# 
+# # Visualising the data
+# temp_plot <- function(df){
+#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
+#     geom_line(aes(group = site)) +
+#     labs(x = "", y = "Temperature (°C)") +
+#     theme(axis.text.x = element_text(angle = 45)) +
+#     theme(legend.position = "top")
+# }
+# 
+# SACTN_plot <- temp_plot(df = SACTN_US)
+# SACTN_plot
+
+# 25th quantile
+```
+
+Different distances from the coastline
+
+```{r}
+# load("Data/site_list_sub.Rdata")
+# # xtable(site_list_sub, auto = TRUE)
+# west <- site_list_sub
+# west$coast <- "west" # Chnages wc to west
+# 
+# load("Data/africa_coast.RData")
+# 
+# ## Downloading the bathy data from NOAA
+# # Download mid-res bathymetry data
+# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# # colnames(sa_bathy) <- c("lon", "lat", "depth")
+# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+# 
+# # Loading in the newly downloaded bathymetry data               
+# load("Data/bathy/sa_bathy.RData")
+
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# shore.normal.transect <- function(site, width = 2){
+#   # Find the site on the coastline and it's nearest neighbour points
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
+#   coords3 <- coords3[2:1,1:3]
+#   # Define the shore normal transect bearing
+#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+#   if(heading >= 360){
+#     heading <- heading-360
+#   } else {
+#     heading <- heading
+#   }
+#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+#   return(heading2)
+# }
+# 
+# # Creating the transects
+# site_transects <- data.frame()
+# for(i in 1:length(west$site)){
+#  site <- west[i,]
+#  site_transect <- shore.normal.transect(site, 2)
+#  site_transects <- rbind(site_transects, site_transect)
+# }
+# 
+# # Manually correcting Sea Point and Kommetjie
+# site_transects$heading[4:5] <- 290 
+# # save(site_transects, file = "Data/site_transects.RData")
+# load("Data/site_transects.RData")
+# 
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
+# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+# 
+# transect.pixel <- function(site, distances){
+#   # Extract coordinates
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   # Find lon/ lats every X metres 
+#   pixels <- data.frame()
+#   # deep <- 999
+#   # distance_multiplier <- 1
+#   # while(deep > isobath){
+#   for(i in 1:length(distances)){
+#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
+#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
+#     bathy2 <- sa_bathy[sitesIdx,]
+#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
+#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
+#                          heading = site$heading, 
+#                          distance = distances[i])
+#     pixels <- rbind(pixels, bathy3)
+#     coords <- coords2
+#   }
+#   if(nrow(pixels) < 1){
+#     pixels <- data.frame(site, depth = NA)
+#   }else{
+#     pixels <- pixels
+#   }+
+
+#   return(pixels)
+# }
+# 
+# # Pixel points
+# site_pixels <- data.frame()
+# for(i in 1:length(west$site)){
+#   site <- site_transects[i,]
+#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+#   site_pixels <- rbind(site_pixels, site_pixel)
+# }
+# 
+# # Bounding box
+#   # Only one is made in order to know how large the the geom_point() squares should be made to match
+# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
+#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
+#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+# 
+# # Determining the temperature at the various distances from the coast
+# 
+# # save(site_pixels, file = "Data/site_pixels.RData")
+# load("Data/site_pixels.RData")
+
+```
+
+# Extrcting the MUR data
+
+```{r}
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# 
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
+# 
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+# fNameStem <-
+#   substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+
+# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# names(MUR)<-c("lon","lat", "temp", "date")
+# MUR <- MUR %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# 
+# ######################################### SUBSET VIA REGION #################
+# 
+# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+#   
+#   
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)  
+#   
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+```
+
+# Extracting the CMC data
+
+```{r}
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
+# 
+# #          1         2         3         4         5         6
+# # 123456789012345678901234567890123456789012345678901234567890
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# 
+# 
+# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
+# names(CMC)<-c("lon","lat", "temp", "date")
+# CMC <- CMC %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# # save(CMC, file = "Data/CMC.RData")
+```
+
+# Extracting the OISST data
+
+```{r}
+# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+#                    CC = c(25, 35, 340, 355), # Canary Current
+#                    CalC = c(35, 45, 225, 240), # California Current
+#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+# 
+# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
+# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
+# 
+# #          1         2
+# # 1234567890123456789012345
+# # avhrr-only-v2.19810901.nc
+# 
+# # function to extract the dims and data from OISST netCDFs
+# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
+#   coords <- bbox[, region]
+#   nc <- nc_open(ncFile)
+#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
+#   fNameStem <-
+#     substr(ncFile, pathLen + 1, pathLen + 13)
+#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
+#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#   sst <- ncvar_get(nc,
+#                    varid = "sst",
+#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
+#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
+#     round(4)
+#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
+#                         lat = nc$dim$lat$vals[LatIdx])
+#   nc_close(nc)
+#   sst <-
+#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
+#     mutate(t = ymd(fDate)) %>%
+#     na.omit()
+#   fwrite(sst,
+#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
+#          append = TRUE, col.names = FALSE)
+#   rm(sst)
+# }
+# 
+# # the list of files
+# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# strtDate <- str_sub(ncList[1], start = 15, end = 22)
+# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
+# 
+# # apply the function
+# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# 
+# # Loading the data
+# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
+# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
+# 
+# # Saving the data
+# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
+```
+
+# Extracting G1SST
+
+```{r}
+# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+# 
+# # #          1         2         3         4
+# # # 12345678901234567890123456789012345678901
+# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# #          1         2         3         4         5         6
+# # 1234567890123456789012345678901234567890123456789012345678901
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# 
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+
+```
+
+# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
+# Is this upwelling event seen throughout the different distances from the coastline
+# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
+
+## Function
+Determining the temperatures at the various distances from the coastline
+
+```{r}
+# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
+# MUR <- MUR %>%
+#   select(lon, lat) %>%
+#   mutate(product = "MUR")
+# 
+# # Decided to work with OISST and CMC as both has a time series of 30years
+OISST_prod <- OISST %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "OISST")
+# 
+CMC_prod <- CMC %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "CMC")
+# 
+sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+#   #rbind(., CMC) %>% 
+select(product, lon, lat)
+# 
+sat_pixels <- sat_data %>%
+select(product, lon, lat) %>%
+unique()
+# 
+# ## For testing the nest/map pipeline
+# # df <- site_pixels %>%
+# #   filter(site == "Lamberts Bay") %>%
+# #   select(-site)
+# 
+match_func <- function(df){
+  df <- df %>%
+    dplyr::rename(lon_site = lon, lat_site = lat)
+  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
+                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  res <- rbind(OISST_index, CMC_index)
+  return(res)
+}
+
+# # Find the nearest pixels to each spot along the transect
+# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
+# # points to not fall inside of the same 25 KM pixel
+pixel_match <- site_pixels %>%
+  group_by(site) %>%
+  group_modify(~match_func(.x))
+
+# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
+# 
+# Clean up some RAM space
+rm(OISST, CMC); gc()
+# 
+# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
+# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
+# # # Hout Bay will be ignored
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+# 
+OISST_fill <- OISST_fill %>%
+  filter(site %in% selected_sites)
+
+CMC_fill <- CMC_fill %>%
+  filter(site %in% selected_sites)
+# 
+# 
+# # All the code below can be viewed in upwell_IDX.Rmd
+# # Next to run upwelling index
+# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+# load("Data/UI_angle.RData")
+# 
+upwelling <- UI_angle %>%
+  dplyr::rename(temp = ui.saws) %>%
+  group_by(site) %>%
+  # mutate(min_t = min(t),
+  #        max_t = max(t)) %>%
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+         # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+Upwelling can last for a few hours.
+Given that we have daily data, upwelling events minimum duration here will be 1day
+  # Detect consecutive days in exceedance of a given threshold.
+  # mutate() %>%
+  select(-data, -clim) %>%
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance)
+
+# Now applying the Upwelling func
+# NB: This only pulls out the event results and not the climatology results
+# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
+# To create a list output one would use group_map(),
+# but this then loses the labels of which sites etc. the results belong to
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
+  return(res)
+}
+# 
+ts2clm_custom <- function(df){
+  # The climatology base period used here is up for debate...
+  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  return(res)
+}
+# 
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+  upwell_base <- df %>%
+    dplyr::rename(t = date) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+  }
+# 
+OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
+save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
+save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
+# 
+# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
+SACTN_US <- SACTN_US %>%
+  filter(site %in% selected_sites)
+
+load("Data/site_list_v4.2.RData")
+
+SACTN_upwell_base <- SACTN_US %>%
+  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
+    dplyr::rename(t = date) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+
+save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+```
+
+Loading the data
+    This is the data obtained once the upwelling signals were filtered out
+
+```{r}
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
+
+# Removing the distance of 20 and 40kms
+library(dplyr)
+removing_distance_func <- function(df){
+  removing_dist_func<- df %>% 
+       filter(!distance %in% c(20000,40000))
+}
+
+OISST_final <- removing_distance_func(df = OISST_upwell_base)
+G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+MUR_final <- removing_distance_func(df = MUR_upwell_base)
+CMC_final <- removing_distance_func(df = CMC_upwell_base)
+
+# save(OISST_final, file = "Data/OISST_final.RData")
+# save(G1SST_final, file = "Data/G1SST_final.RData")
+# save(MUR_final, file = "Data/MUR_final.RData")
+# save(CMC_final, file = "Data/CMC_final.RData")
+```
+
+To accurately compare 
+  Start date CMC start date 1991-09-02 end date 2017-03-15
+  Start date OISST start date 1981-09- 01  and end date 2018-09-28
+  Start date SACTN start date 1973 and end date 2017-12-16
+
+```{r}
+#glm to determine if there is a difference in the number of events recorded at different at every distance
+
+load("Data/Count/OISST_count.RData")
+load("Data/Count/G1SST_count.RData")
+load("Data/Count/MUR_count.RData")
+load("Data/Count/CMC_count.RData")
+
+options(scipen = 999)
+anova_func <- function(df){
+  sites_aov <- aov(distance ~ duration * site * site, data = df)
+return(sites_aov)
+}
+
+OISST_anov <- anova_func(df = OISST_upwell_base)
+summary(OISST_anov)
+CMC_anov <- anova_func(df = CMC_upwell_base)
+summary(CMC_anov)
+
+MUR_anov <- anova_func(df = MUR_upwell_base)
+summary(MUR_anov)
+
+SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
+summary(SACTN_anov)
+
+lm_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
+}
+# Changes in upwelling metrics
+lm_func <- function(df){
+  upwell_lm <- df %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var, distance) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+}
+
+OISST_lm <- lm_func(df = OISST_upwell_base)
+CMC_lm <- lm_func(df = CMC_upwell_base)
+
+SACTN_lm <- SACTN_upwell_base %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+```
+
+As a common date we will use a start date from 1992-01-31 and end date 16-12-31
+
+```{r}
+# Generalosed linear models
+
+# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
+# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)
+
+# trends_duration_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
+#                      family = poisson(link = "log"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
+# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
+# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
+# 
+# 
+# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
+# trends_intensity_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
+#                      family = gaussian(link = "identity"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
+# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
+# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
+```
+
+# From the years 1992 - 2017
+
+```{r}
+# Which pixels showed the highest upwelling counts within the different datasets
+# Which site showed the most intense upwelling within the different datasets
+
+metric_func <- function(df){
+ metrics <-df %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  #mutate(year = year(date_start)) %>% 
+  group_by(distance, site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+   select(site, distance, mean_intensity, sum_events)
+}
+
+OISST_metrics <- metric_func(OISST_upwell_base)
+CMC_metrics <- metric_func(CMC_upwell_base)
+
+# No of u pwelling events found at the different sites
+SACTN_metrics <- SACTN_upwell_base %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  group_by(site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+  select(site, mean_intensity, sum_events)
+
+# Create a table showing these metrics
+
+```
+
+# Wind plots
+
+```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
+library(ggradar)
+library(dplyr)
+library(scales)
+library(tibble)
+load("Data/UI_angle.RData")
+
+# Detects NA values, no NA values are present
+# Plotting wind variables
+source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function
+
+wind_daily_renamed <- UI_angle %>% 
+  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ) %>% 
+  filter(spd > 0)
+
+p.wr2 <- plot.windrose(data = wind_daily_renamed,
+              spd = "spd",
+              dir = "dir")
+
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
+  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
+p.wr3
+
+################3
+
+# load("Data/UI_angle.RData")
+# 
+# wind_plot_prep <- UI_angle %>% 
+#   select(-t, -ui.saws)
+# 
+# wind_plot <- wind_plot_prep %>% 
+#   as_tibble(rownames = "site") %>% 
+#   group_by(site) %>% 
+#   mutate_at(vars(-site), rescale) %>% 
+#   tail(4) 
+# 
+# ggradar(wind_plot)
+# 
+# ??countmax
+
+```
+
+# ANOVA
+- Relationship between duration, year and site as a function of the distance 
+The number of events per distance and then compare this at each distance
+
+```{r}
+options(scipen = 999)
+anova_func <- function(df){
+  sites_aov <- aov(distance ~ duration * site * site, data = df)
+return(sites_aov)
+}
+
+OISST_anov <- anova_func(df = OISST_upwell_base)
+summary(OISST_anov)
+CMC_anov <- anova_func(df = CMC_upwell_base)
+summary(CMC_anov)
+
+MUR_anov <- anova_func(df = MUR_upwell_base)
+summary(MUR_anov)
+
+SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
+summary(SACTN_anov)
+```
+
+Plot showing upwelling at the different sites for each of the different distances. Scatter plot
+
+```{r}
+
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
+
+km_func <- function(df){
+  upwell_base <-  df%>% 
+      mutate(distance_km = case_when(distance == "10000" ~ "10",
+                                    distance == "20000" ~ "20",
+                                    distance  == "30000" ~"30",
+                                    distance  == "40000" ~ "40",
+                                    distance  == "50000" ~ "50"))
+}
+OISST_upwell_base <- km_func(df = OISST_upwell_base)
+OISST_upwell_base$distance_km <- as.numeric(OISST_upwell_base$distance_km)
+
+CMC_upwell_base <- km_func(df = CMC_upwell_base)
+CMC_upwell_base$distance_km <- as.numeric(CMC_upwell_base$distance_km)
+
+MUR_upwell_base <- km_func(df = MUR_upwell_base)
+MUR_upwell_base$distance_km <- as.numeric(MUR_upwell_base$distance_km)
+
+G1SST_upwell_base <- km_func(df = G1SST_upwell_base)
+G1SST_upwell_base$distance_km <- as.numeric(G1SST_upwell_base$distance_km)
+
+library(ggpubr)
+year_func <- function(df){
+  Filtered <- df %>%
+filter(year(date_start) %in% seq(2011, 2013, 1))
+}
+
+CMC_filtered <- year_func(df = CMC_upwell_base)
+OISST_filtered <- year_func(df = OISST_upwell_base)
+SACTN_filtered <- year_func(df = SACTN_upwell_base)
+MUR_filtered <- year_func(df = MUR_upwell_base)
+G1SST_filtered <- year_func(df = G1SST_upwell_base)
+
+CMC_SEAPOINT <- CMC_filtered %>% 
+  filter(site == "Sea Point")
+CMC_SB <- CMC_filtered %>% 
+  filter(site == "Saldanha Bay")
+CMC_PN <- CMC_filtered %>% 
+  filter(site == "Port Nolloth")
+CMC_LB <- CMC_filtered %>% 
+  filter(site == "Lamberts Bay")
+
+###################
+## OISST
+OISST_SEAPOINT <- OISST_filtered %>% 
+  filter(site == "Sea Point")
+OISST_SB <- OISST_filtered %>% 
+  filter(site == "Saldanha Bay")
+OISST_PN <- OISST_filtered %>% 
+  filter(site == "Port Nolloth")
+OISST_LB <- OISST_filtered %>% 
+  filter(site == "Lamberts Bay")
+
+############
+#####MUR
+
+MUR_SEAPOINT <- MUR_filtered %>% 
+  filter(site == "Sea Point")
+MUR_SB <- MUR_filtered %>% 
+  filter(site == "Saldanha Bay")
+MUR_PN <- MUR_filtered %>% 
+  filter(site == "Port Nolloth")
+MUR_LB <- MUR_filtered %>% 
+  filter(site == "Lamberts Bay")
+###########
+####G1SST
+
+G1SST_SEAPOINT <- G1SST_filtered %>% 
+  filter(site == "Sea Point")
+G1SST_SB <- G1SST_filtered %>% 
+  filter(site == "Saldanha Bay")
+G1SST_PN <- G1SST_filtered %>% 
+  filter(site == "Port Nolloth")
+G1SST_LB <- G1SST_filtered %>% 
+  filter(site == "Lamberts Bay")
+
+```
+
+# Lolli plot
+
+```{r}
+plot_loli_func <- function(df){
+  ggplot(df, aes(x = date_peak, y = duration)) + 
+  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
+  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+  facet_wrap(~distance_km, ncol = 2) +
+  scale_y_continuous(breaks = c(5,10,15,20,25,30),
+                     limits = c(1,30)) +
+    scale_x_date(limits = as.Date(c("2011-01-01","2014-01-01"))) +
+  theme_bw() +
+  theme(axis.text=element_text(size=15),
+        axis.title=element_text(size=17)) +
+  theme(strip.text.x = element_text(size = 17))+
+  theme(panel.spacing = unit(1, "lines")) +
+    #theme(strip.text = c("10km", "20km", "30km", "40km", "50km")) +
+  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
+  xlab("Date (years)") + ylab("Event duration (days)") 
+}
+
+(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
+(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
+(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
+(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
+
+OISST_SP_pl <- OISST_SP_pl +
+  ggtitle("D.") #Sea Point
+OISST_SB_pl <- OISST_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+OISST_PN_pl <- OISST_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+OISST_LB_pl <- OISST_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl)
+
+combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)
+
+#######
+# CMC
+
+(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
+(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
+(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
+(CMC_LB_pl <- plot_loli_func(df = CMC_LB))
+
+CMC_SP_pl <- CMC_SP_pl +
+  ggtitle("D.") #Sea Point
+CMC_SB_pl <- CMC_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+CMC_PN_pl <- CMC_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+CMC_LB_pl <- CMC_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_CMC_pl_2 <- ggarrange(CMC_SB_pl, CMC_SP_pl)
+
+combined_CMC_pl_1 <- ggarrange(CMC_LB_pl, CMC_PN_pl)
+
+##########
+### MUR
+(MUR_SP_pl <- plot_loli_func(df = MUR_SEAPOINT))
+(MUR_SB_pl <- plot_loli_func(df = MUR_SB))
+(MUR_PN_pl <- plot_loli_func(df = MUR_PN))
+(MUR_LB_pl <- plot_loli_func(df = MUR_LB))
+
+MUR_SP_pl <- MUR_SP_pl +
+  ggtitle("D.") #Sea Point
+MUR_SB_pl <- MUR_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+MUR_PN_pl <- MUR_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+MUR_LB_pl <- MUR_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_MUR_pl_2 <- ggarrange(MUR_SB_pl, MUR_SP_pl)
+
+combined_MUR_pl_1 <- ggarrange(MUR_LB_pl, MUR_PN_pl)
+#######
+#G1SST
+
+(G1SST_SP_pl <- plot_loli_func(df = G1SST_SEAPOINT))
+(G1SST_SB_pl <- plot_loli_func(df = G1SST_SB))
+(G1SST_PN_pl <- plot_loli_func(df = G1SST_PN))
+(G1SST_LB_pl <- plot_loli_func(df = G1SST_LB))
+
+G1SST_SP_pl <- G1SST_SP_pl +
+  ggtitle("D.") #Sea Point
+G1SST_SB_pl <- G1SST_SB_pl +
+  ggtitle("C.") # Saldanha Bay
+G1SST_PN_pl <- G1SST_PN_pl +
+  ggtitle("B.")        # Port Nolloth
+G1SST_LB_pl <- G1SST_LB_pl +
+  ggtitle("A.")#Lamberts Bay
+
+combined_G1SST_pl_2 <- ggarrange(G1SST_SB_pl, G1SST_SP_pl)
+
+combined_G1SST_pl_1 <- ggarrange(G1SST_LB_pl, G1SST_PN_pl)
+
+########
+
+plot_loli_func <- function(df){
+  ggplot(df, aes(x = date_peak, y = duration)) + 
+  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
+  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
+   facet_wrap(~site, ncol = 2) +
+  scale_y_continuous(breaks = c(5,10,15,20,25,30),
+                     limits = c(1,30)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=17),
+        axis.title=element_text(size=17)) +
+  theme(strip.text.x = element_text(size = 17))+
+  theme(panel.spacing = unit(0.5, "lines")) +
+  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
+  xlab("Date (years)") + ylab("Event duration (days)") 
+}
+
+(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
+```
+
+# GLM - Determining if a change occur in the number of signals detected at different distances from the coastline
+# Included a table in the paper
+  Table discuss/count the number of upwelling signals detected at the different distances
+
+  - For some of the sites it is seen that fewer upwelling signals are visible at a further distance from the coastline
+  - When adding up all the signals observed at their respective distance it is seen that there are only slight differences in the amount of signals detected In some cases the signals detected at 10km from the coastline exceed that detected at 50km and visa versa
+
+## Number of events detected
+
+Using the start date and getting the month and year into a column. The month and year will be on the xaxis
+Then count the number of times per month at different years signals are detected
+
+```{r}
+load("Data/OISST_final.RData")
+load("Data/CMC_final.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/G1SST_final.RData")
 
@@
   geom_freqpoly() +
+  facet_wrap(~
+month_year_func <- function(df){
+  month_year_func<- df %>% 
+    mutate(month = month(date_start, abbr = T, label = T)) %>% 
+    separate(date_start, into = c("year", "month_no")) %>% 
+    unite(month_year, month,year)
+    # Now need to get the year into a seperate column from the start date. So now date_start is split in the dataset
+}
+
+OISST_prep <- month_year_func(df = OISST_final)
+CMC_prep <- month_year_func(df = CMC_final)
+G1SST_prep <- month_year_func(df = G1SST_final)
+MUR_prep <- month_year_func(df = MUR_final)
+
+# Combined dataset
+# Combined_frequency <- read_excel("Data/Combined_frequency.xlsx")
+# save(Combined_frequency, file = "Data/Combined_frequency.RData")
+```
+
+# Plotting
+
+```{r}
+load("Data/Combined_frequency.RData")
+
+Combined_frequency <- Combined_frequency %>%
+  mutate(t = parse_date(date, format = "%b_%Y"),
+         distance = factor(distance))
+
+ggplot(data = Combined_frequency, aes(x = distance, y = number_of_signals)) +
+  geom_boxplot(aes(colour = product)) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
+  #geom_smooth(aes(colour = product), method = "lm") +
+  #labs(x = NULL) +
+  facet_wrap(~site)
+
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals, colour = product)) +
+geom_point(aes(size = distance)) +
+geom_smooth(method = "lm", size = 1.2)+
+theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+# Does not accurately show the months
+# Plot need to include month_year, number of signals, distance and product ()
+
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
+  geom_line(aes(colour = product)) +
+  geom_smooth(aes(colour = product), method = "lm") +
+  labs(x = NULL) +
+  facet_wrap(~site) #+
+  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
+  geom_line(aes(colour = distance)) +
+  geom_point(aes(colour = distance)) +
+  geom_smooth(aes(colour = distance), method = "lm") +
+  # labs(x = NULL) +
+  facet_wrap(~site) #+
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+  ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
+  geom_line(aes(colour = distance)) +
+  geom_point(aes(colour = distance)) +
+  geom_smooth(aes(colour = distance), method = "lm") +
+  # labs(x = NULL) +
+  facet_wrap(~site) #+
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+####
+
+combined_frequency_split <- Combined_frequency %>% 
+  separate(date, into = c("month", "year"))
+
+# ggplot(data = combined_frequency_split, aes(x = year, y = number_of_signals)) +
+#   geom_line(aes(colour = month)) +
+#   geom_smooth(aes(colour = month), method = "lm") +
+#   facet_wrap(~site)
+```
+
+#Pairwise correlations
+Comparing the relaionship of variables within the same sample
+
+  -Correlate the 1st pixel with the second pixel and the first pixel with the last pixel
+  -Correlation between satellite data and number of signals at the different distances from the coastline
+
+```{r}
+
+# chi squared test
+
+load("Data/Combined_frequency.RData")
+
+combined_freq_spread <- pivot_wider(Combined_frequency, names_prefix = "dist_",
+                                    names_from = distance, values_from = number_of_signals)
+
+slope_calc <- function(df){
+  df %>% 
+    # mutate(row_num = 1:n()) %>% 
+    # do(mod1 = lm(number_of_signals ~ row_num, data = .),
+       # mod2 = lm(distance ~ row_num, data = .),
+       # mod3 = lm(distance ~ number_of_signals, data = .),
+       # mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
+        # mutate(distance_slope = summary(mod1)$coeff[2],
+        #    signal_slope = summary(mod2)$coeff[2],
+        #    signal_distance_slope = summary(mod3)$coeff[2],
+        #    signal_distance_r = mod4[1],
+        #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
+        # select(-mod1, -mod2, -mod3, -mod4) %>% 
+    do(mod1 = cor(.$dist_10, .$dist_30, method = "pearson", use = "complete.obs"),
+       mod2 = cor(.$dist_10, .$dist_50, method = "pearson", use = "complete.obs")) %>%
+    mutate(dist_10_vs_30_r = mod1[1],
+           dist_10_vs_50_r = mod2[1]) %>% 
+    # select(dist_10_vs_30_r, dist_10_vs_50_r) %>% 
+    select(-mod1, -mod2) %>% 
+    mutate_if(is.numeric, round, 2)
+}
+
+#### test <- cor(combined_freq_spread$`10`, combined_freq_spread$`30`)
+
+distance_corr <- combined_freq_spread %>% 
+  group_by(site, product) %>% 
+  slope_calc()
+```
+
+```{r}
+library(hrbrthemes)
+library(viridis)
+
+Combined_frequency <- Combined_frequency %>%
+  mutate(t = parse_date(date, format = "%b_%Y"),
+         distance = factor(distance))
+
+ggplot(Combined_frequency, aes(x=t, y=number_of_signals, size=distance, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+ggplot(Combined_frequency, aes(x=date, y=number_of_signals, size=distance, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+### Without time
+
+ggplot(Combined_frequency, aes(x=distance, y= number_of_signals, color=product)) +
+  geom_point(alpha=0.5) +
+  facet_wrap(~site) +
+  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
+
+```
+
+# Histogram
+
+```{r}
+
+combined_frequency_season <- Combined_frequency %>% 
+  mutate(month = month(date, abbr = T, label = T),
+         year = year(date)) %>% 
+  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+
+CMC_dur <- read_excel("Data/CMC_dur.xlsx")
+
+ggplot(data = CMC_dur, aes(x = duration)) +
+  geom_histogram(aes(fill = product),  
+                       position = "dodge", alpha=0.6) +
+   labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
   facet_wrap(~site) +
 theme_bw()
-```
+
+ggplot(CMC_dur, aes(duration, colour = product)) +
+  geom_freqpoly() +
+  facet_wrap(~site) +
+  theme_bw()
+```
+
+### Looking at intensity
+
+```{r}
+
+filt_func <- function(df){
+  intens <- df %>% 
+  #mutate(year = year(date_start)) %>% 
+  filter(year(date_start) %in% seq(2011, 2012))
+}
+
+filt_func2 <- function(df){
+  intens1 <- df %>% 
+  #mutate(year = year(date_start)) %>% 
+  filter(year(date_start) %in% seq(2013, 2014))
+}
+
+OISST_filt_func <- filt_func(OISST_final)
+OISST_filt_func2 <- filt_func2(OISST_final)
+OISST_binding <- rbind(OISST_filt_func,OISST_filt_func2) %>% 
+  select(site, product, intensity_mean)
+
+MUR_filt_func <- filt_func(MUR_final)
+MUR_filt_func2 <- filt_func2(MUR_final)
+MUR_binding <- rbind(MUR_filt_func,MUR_filt_func2) %>% 
+  select(site, product, intensity_mean)
+
+G1SST_filt_func <- filt_func(G1SST_final)
+G1SST_filt_func2 <- filt_func2(G1SST_final)
+G1SST_binding <- rbind(G1SST_filt_func,G1SST_filt_func2) %>% 
+  select(site, product, intensity_mean)
+
+CMC_filt_func <- filt_func(CMC_final)
+CMC_filt_func2 <- filt_func2(CMC_final)
+CMC_binding <- rbind(CMC_filt_func,CMC_filt_func2) %>% 
+  select(site, product, intensity_mean)
+
+#write.csv(SACTN_upwell_base ,'SACTN_upwell_base .csv') # Copied this and then added in a product
+rm(SACTN_upwell_base)
+SACTN_mean_intensity <- SACTN_upwell_base_ %>% 
+  select(site, product, intensity_mean)
+
+Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) %>% 
+  select(-heading, - distance, - lat, -lon)
+# write.csv(Complete_mean_intensity ,'Complete_mean_intensity .csv') # Copied this and then removedsome columnss
+
+final_intensity <- rbind(SACTN_mean_intensity, Complete_mean_intensity_)
+
+ggplot(data = final_intensity, aes(x = product, y = intensity_mean)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+
+ggplot(data =final_intensity, aes(x = intensity_mean)) +
+  geom_histogram(aes(fill = product),  
+                       position = "dodge", alpha=0.6) +
+  # labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
+  facet_wrap(~site) +
+theme_bw() 
+```

### 2020-01-31 SST_patterns.Rmd
update

@@
 library(heatwaveR)
+library(grid)
+library(ggthemes)
 ```
@@
 load("Data/CMC.RData") # 1decimal places
+```
+
+# PLotting theme
+
+```{r}
+theme_Publication <- function(base_size=14, base_family="helvetica") {
+      library(grid)
+      library(ggthemes)
+      (theme_foundation(base_size=base_size, base_family=base_family)
+       + theme(plot.title = element_text(face = "bold",
+                                         size = rel(1.2), hjust = 0.5),
+               text = element_text(),
+               panel.background = element_rect(colour = NA),
+               plot.background = element_rect(colour = NA),
+               panel.border = element_rect(colour = NA),
+               axis.title = element_text(face = "bold",size = rel(1)),
+               axis.title.y = element_text(angle=90,vjust =2),
+               axis.title.x = element_text(vjust = -0.2),
+               axis.text = element_text(), 
+               axis.line = element_line(colour="black"),
+               axis.ticks = element_line(),
+               panel.grid.major = element_line(colour="#f0f0f0"),
+               panel.grid.minor = element_blank(),
+               legend.key = element_rect(colour = NA),
+               legend.position = "bottom",
+               legend.direction = "horizontal",
+               legend.key.size= unit(0.2, "cm"),
+               legend.margin = unit(0, "cm"),
+               #legend.title = element_text(face="italic"),
+               plot.margin=unit(c(10,5,5,5),"mm"),
+               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+               strip.text = element_text(face="bold")
+          ))
+
+}
+
+scale_fill_Publication <- function(...){
+      library(scales)
+      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+
+}
+
+scale_colour_Publication <- function(...){
+      library(scales)
+      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+
+}
 ```
@@
 
-CMC_dur <- read_excel("Data/CMC_dur.xlsx")
+CMC_dur <- read_csv("Data/CMC_dur.csv")
 
@@
   facet_wrap(~site) +
-theme_bw()
+  scale_colour_Publication()+
+  theme_Publication()
 
@@
 OISST_filt_func2 <- filt_func2(OISST_final)
-OISST_binding <- rbind(OISST_filt_func,OISST_filt_func2) %>% 
-  select(site, product, intensity_mean)
+OISST_binding <- rbind(OISST_filt_func,OISST_filt_func2)# %>% 
+ # select(site, product, intensity_mean)
 
@@
 MUR_filt_func2 <- filt_func2(MUR_final)
-MUR_binding <- rbind(MUR_filt_func,MUR_filt_func2) %>% 
-  select(site, product, intensity_mean)
+MUR_binding <- rbind(MUR_filt_func,MUR_filt_func2)# %>% 
+ # select(site, product, intensity_mean)
 
@@
 G1SST_filt_func2 <- filt_func2(G1SST_final)
-G1SST_binding <- rbind(G1SST_filt_func,G1SST_filt_func2) %>% 
-  select(site, product, intensity_mean)
+G1SST_binding <- rbind(G1SST_filt_func,G1SST_filt_func2) #%>% 
+  #select(site, product, intensity_mean)
 
@@
 CMC_filt_func2 <- filt_func2(CMC_final)
-CMC_binding <- rbind(CMC_filt_func,CMC_filt_func2) %>% 
-  select(site, product, intensity_mean)
-
-#write.csv(SACTN_upwell_base ,'SACTN_upwell_base .csv') # Copied this and then added in a product
-rm(SACTN_upwell_base)
-SACTN_mean_intensity <- SACTN_upwell_base_ %>% 
-  select(site, product, intensity_mean)
-
-Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) %>% 
-  select(-heading, - distance, - lat, -lon)
-# write.csv(Complete_mean_intensity ,'Complete_mean_intensity .csv') # Copied this and then removedsome columnss
+CMC_binding <- rbind(CMC_filt_func,CMC_filt_func2) #%>% 
+ # select(site, product, intensity_mean)
+
+SACTN_filt_func <- filt_func(SACTN_upwell_base_)
+SACTN_filt_func2 <- filt_func2(SACTN_upwell_base_)
+SACTN_binding <- rbind(SACTN_filt_func,SACTN_filt_func2)# %>% 
+# write.csv(SACTN_binding,'SACTNfinal_data.csv')
+
+Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) #%>% 
+  #select(-heading, - distance, - lat, -lon)
+
+# write.csv(Complete_mean_intensity ,'Complete_mean_intensity.csv') # Copied this and then removedsome columnss
+
+Complete_mean_intensity2 <- Complete_mean_intensity2 %>% 
+  select(-distance_km)
+
+final_data <- rbind(Complete_mean_intensity2, SACTN_binding)
+#write.csv(final_data,'final_data.csv')
 
@@
 
-ggplot(data = final_intensity, aes(x = product, y = intensity_mean)) +
-  geom_point() +
+```
+
+# PLotting
+
+```{r}
+
+signals_average <- read_csv("Data/signals_average.csv")
+
+ggplot(data = signals_average, aes(x = no_of_signals, y = intensity_mean)) +
+  geom_point(aes(colour = product), size = 3) +
   geom_smooth(method = "lm") +
-  facet_wrap(~site)
+  facet_wrap(~site) + 
+  labs(x = "Number of signals", y = "Average intensity") +
+  scale_colour_Publication()+
+  theme_Publication()
 

### 2020-01-31 SST_patterns.Rmd
update

@@
 
+SACTN_upwell_base_ <- read_csv("Data/SACTN_upwell_base .csv") # The SACTN product column was added in
+
 SACTN_filt_func <- filt_func(SACTN_upwell_base_)
@@
 
-Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) #%>% 
-  #select(-heading, - distance, - lat, -lon)
-
-# write.csv(Complete_mean_intensity ,'Complete_mean_intensity.csv') # Copied this and then removedsome columnss
-
-Complete_mean_intensity2 <- Complete_mean_intensity2 %>% 
-  select(-distance_km)
-
-final_data <- rbind(Complete_mean_intensity2, SACTN_binding)
-#write.csv(final_data,'final_data.csv')
-
-final_intensity <- rbind(SACTN_mean_intensity, Complete_mean_intensity_)
+Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) %>% 
+  ungroup() %>% 
+  select(-heading, - distance, -distance_km)
+
+#write.csv(Complete_mean_intensity ,'Complete_mean_intensity.csv') # Copied this and then removedsome columnss
+
+final_data <- rbind(Complete_mean_intensity, SACTN_binding)
+# write.csv(final_data,'final_data.csv')
 

### 2020-01-31 SST_patterns.Rmd
update

@@
   ungroup() %>% 
-  select(-heading, - distance, -distance_km)
+  select(-heading, - distance, -distance_km) # Here I remove distance, Keep if distance is needed
 

### 2020-02-01 SST_patterns.Rmd
update

@@
   facet_wrap(~site) + 
-  labs(x = "Number of signals", y = "Average intensity") +
+  labs(x = "Number of signals ", y = "Average intensity (°C)") +
   scale_colour_Publication()+
@@
 ```
+
+# Linear model
+
+```{r}
+remotes::install_github("moderndive/moderndive")
+
+library(moderndive)
+ Regression <- read_csv("Data/Regression.csv")
+score_model <- lm(intensity_mean ~ no_of_signals, data = Regression)
+final_regression <- get_regression_table(score_model)
+
+  linear_regression <- Regression %>% 
+    gather(key = variable, 
+           value = value,
+           -site) %>%
+    group_by(site, variable) %>%
+    na.omit() %>% 
+    nest() %>%
+    mutate(model = purrr::map(data, ~lm(variable ~ value, data = .))) %>%
+    unnest(model %>% purrr::map(glance)) %>%
+    select(cluster:variable, adj.r.squared, p.value) %>% 
+    arrange(site) %>% 
+    mutate(adj.r.squared = round(adj.r.squared, 2),
+           p.value = round(p.value, 4))
+
+```

### 2020-02-01 site_map.Rmd
update

@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        legend.text = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5),
-        axis.text = element_text(size = 12)) +
-    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.minor = element_line(colour = "NA"),
-        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
-        legend.direction = "horizontal",
-        legend.justification = c(1,0),
-        legend.position = c(0.85, 0.70),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5))
+  scale_colour_Publication()+
+  theme_Publication() # R publication theme
+  # theme(panel.background = element_rect(fill = "ivory", colour = NA),
+  #       panel.border = element_rect(colour = "black", size = 0.5),
+  #       panel.grid.major = element_blank(),
+  #       panel.grid.minor = element_blank(),
+  #       legend.text = element_text(size = 12),
+  #       legend.title = element_text(size = 12),
+  #       legend.key = element_blank(),
+  #       legend.background = element_blank(),
+  #       axis.ticks = element_line(size = 0.5),
+  #       axis.text = element_text(size = 12)) +
+  #   guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
+  # theme(panel.background = element_rect(fill = "ivory", colour = NA),
+  #       panel.border = element_rect(colour = "black", size = 0.5),
+  #       panel.grid.minor = element_line(colour = "NA"),
+  #       panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+  #       legend.direction = "horizontal",
+  #       legend.justification = c(1,0),
+  #       legend.position = c(0.85, 0.70),
+  #       legend.text = element_text(size = 8),
+  #       legend.title = element_text(size = 8),
+  #       legend.key = element_blank(),
+  #       legend.background = element_blank(),
+  #       axis.ticks = element_line(size = 0.5))
 site_map

### 2020-04-02 SST_patterns.Rmd
update

--- 2020-01-31 SST_patterns.Rmd
@@
 detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
   return(res)
@@
 load("Data/G1SST_final.RData")
----
-title: "SST_patterns"
-author: "Amieroh Abrahams"
-date: "19 August 2019"
-output: html_document
----
-# This script is for all the various datasets
-
-# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
-
-```{r}
+
 knitr::opts_chunk$set(
@@
 
+save(combined_freq_spread, )
+
 slope_calc <- function(df){
@@
   slope_calc()
+
 ```
@@
   geom_freqpoly() +
-  facet_wrap(~
+  facet_wrap(~ site)
+
 month_year_func <- function(df){
@@
   facet_wrap(~site) +
-theme_bw() 
-```
+theme_bw() ---
+title: "SST_patterns"
+author: "Amieroh Abrahams"
+date: "19 August 2019"
+output: html_document
+---
+# This script is for all the various datasets
+
+# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
+
+```{r}
+knitr::opts_chunk$set(
+  comment = "R>",
+  warning = FALSE,
+  message = FALSE 
+)
+#update.packages() 
+library(tidyverse)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(scales)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(broom)
+# library(ncdf4) # This was used to process NetCDF files in an earlier version
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+library(ncdf4) # library for processing netCDFs
+library(data.table)
+# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
+library(heatwaveR)
+library(grid)
+library(ggthemes)
+```
+
+# Loading in all the data created using the code bellow
+
+```{r}
+load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
+load("Data/site_pixels.RData") # 5 decimal places
+load("Data/OISST.RData") # 2 decimal places
+OISST <- BC_avhrr_only_v2_Document_Document 
+rm(BC_avhrr_only_v2_Document_Document ); gc()
+load("Data/CMC.RData") # 1decimal places
+```
+
+# PLotting theme
+
+```{r}
+theme_Publication <- function(base_size=14, base_family="helvetica") {
+      library(grid)
+      library(ggthemes)
+      (theme_foundation(base_size=base_size, base_family=base_family)
+       + theme(plot.title = element_text(face = "bold",
+                                         size = rel(1.2), hjust = 0.5),
+               text = element_text(),
+               panel.background = element_rect(colour = NA),
+               plot.background = element_rect(colour = NA),
+               panel.border = element_rect(colour = NA),
+               axis.title = element_text(face = "bold",size = rel(1)),
+               axis.title.y = element_text(angle=90,vjust =2),
+               axis.title.x = element_text(vjust = -0.2),
+               axis.text = element_text(), 
+               axis.line = element_line(colour="black"),
+               axis.ticks = element_line(),
+               panel.grid.major = element_line(colour="#f0f0f0"),
+               panel.grid.minor = element_blank(),
+               legend.key = element_rect(colour = NA),
+               legend.position = "bottom",
+               legend.direction = "horizontal",
+               legend.key.size= unit(0.2, "cm"),
+               legend.margin = unit(0, "cm"),
+               #legend.title = element_text(face="italic"),
+               plot.margin=unit(c(10,5,5,5),"mm"),
+               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+               strip.text = element_text(face="bold")
+          ))
+
+}
+
+scale_fill_Publication <- function(...){
+      library(scales)
+      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+
+}
+
+scale_colour_Publication <- function(...){
+      library(scales)
+      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+
+}
+```
+
+```{r}
+# SACTN_US <- SACTN_US %>% 
+#   dplyr::rename(in_situ_temp = temp)
+# 
+# # Visualising the data
+# temp_plot <- function(df){
+#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
+#     geom_line(aes(group = site)) +
+#     labs(x = "", y = "Temperature (°C)") +
+#     theme(axis.text.x = element_text(angle = 45)) +
+#     theme(legend.position = "top")
+# }
+# 
+# SACTN_plot <- temp_plot(df = SACTN_US)
+# SACTN_plot
+
+# 25th quantile
+```
+
+Different distances from the coastline
+
+```{r}
+# load("Data/site_list_sub.Rdata")
+# # xtable(site_list_sub, auto = TRUE)
+# west <- site_list_sub
+# west$coast <- "west" # Chnages wc to west
+# 
+# load("Data/africa_coast.RData")
+# 
+# ## Downloading the bathy data from NOAA
+# # Download mid-res bathymetry data
+# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# # colnames(sa_bathy) <- c("lon", "lat", "depth")
+# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+# 
+# # Loading in the newly downloaded bathymetry data               
+# load("Data/bathy/sa_bathy.RData")
+
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# shore.normal.transect <- function(site, width = 2){
+#   # Find the site on the coastline and it's nearest neighbour points
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
+#   coords3 <- coords3[2:1,1:3]
+#   # Define the shore normal transect bearing
+#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+#   if(heading >= 360){
+#     heading <- heading-360
+#   } else {
+#     heading <- heading
+#   }
+#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+#   return(heading2)
+# }
+# 
+# # Creating the transects
+# site_transects <- data.frame()
+# for(i in 1:length(west$site)){
+#  site <- west[i,]
+#  site_transect <- shore.normal.transect(site, 2)
+#  site_transects <- rbind(site_transects, site_transect)
+# }
+# 
+# # Manually correcting Sea Point and Kommetjie
+# site_transects$heading[4:5] <- 290 
+# # save(site_transects, file = "Data/site_transects.RData")
+# load("Data/site_transects.RData")
+# 
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
+# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+# 
+# transect.pixel <- function(site, distances){
+#   # Extract coordinates
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   # Find lon/ lats every X metres 
+#   pixels <- data.frame()
+#   # deep <- 999
+#   # distance_multiplier <- 1
+#   # while(deep > isobath){
+#   for(i in 1:length(distances)){
+#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
+#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
+#     bathy2 <- sa_bathy[sitesIdx,]
+#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
+#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
+#                          heading = site$heading, 
+#                          distance = distances[i])
+#     pixels <- rbind(pixels, bathy3)
+#     coords <- coords2
+#   }
+#   if(nrow(pixels) < 1){
+#     pixels <- data.frame(site, depth = NA)
+#   }else{
+#     pixels <- pixels
+#   }+
+
+#   return(pixels)
+# }
+# 
+# # Pixel points
+# site_pixels <- data.frame()
+# for(i in 1:length(west$site)){
+#   site <- site_transects[i,]
+#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
+#   site_pixels <- rbind(site_pixels, site_pixel)
+# }
+# 
+# # Bounding box
+#   # Only one is made in order to know how large the the geom_point() squares should be made to match
+# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
+#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
+#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+# 
+# # Determining the temperature at the various distances from the coast
+# 
+# # save(site_pixels, file = "Data/site_pixels.RData")
+# load("Data/site_pixels.RData")
+
+```
+
+# Extrcting the MUR data
+
+```{r}
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# 
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
+# 
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+# fNameStem <-
+#   substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+
+# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# names(MUR)<-c("lon","lat", "temp", "date")
+# MUR <- MUR %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
+# 
+# ######################################### SUBSET VIA REGION #################
+# 
+# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
+# #          1         2         3         4
+# # 12345678901234567890123456789012345678901
+# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 38)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+#   
+#   
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)  
+#   
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+```
+
+# Extracting the CMC data
+
+```{r}
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
+# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
+# 
+# #          1         2         3         4         5         6
+# # 123456789012345678901234567890123456789012345678901234567890
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
+# 
+#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
+# 
+# 
+# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
+# names(CMC)<-c("lon","lat", "temp", "date")
+# CMC <- CMC %>% 
+#   mutate(temp = temp - 273.15)
+# 
+# # save(CMC, file = "Data/CMC.RData")
+```
+
+# Extracting the OISST data
+
+```{r}
+# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
+#                    CC = c(25, 35, 340, 355), # Canary Current
+#                    CalC = c(35, 45, 225, 240), # California Current
+#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
+#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
+# 
+# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
+# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
+# 
+# #          1         2
+# # 1234567890123456789012345
+# # avhrr-only-v2.19810901.nc
+# 
+# # function to extract the dims and data from OISST netCDFs
+# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
+#   coords <- bbox[, region]
+#   nc <- nc_open(ncFile)
+#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
+#   fNameStem <-
+#     substr(ncFile, pathLen + 1, pathLen + 13)
+#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
+#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#   sst <- ncvar_get(nc,
+#                    varid = "sst",
+#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
+#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
+#     round(4)
+#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
+#                         lat = nc$dim$lat$vals[LatIdx])
+#   nc_close(nc)
+#   sst <-
+#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
+#     mutate(t = ymd(fDate)) %>%
+#     na.omit()
+#   fwrite(sst,
+#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
+#          append = TRUE, col.names = FALSE)
+#   rm(sst)
+# }
+# 
+# # the list of files
+# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# strtDate <- str_sub(ncList[1], start = 15, end = 22)
+# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
+# 
+# # apply the function
+# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
+# 
+# # Loading the data
+# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
+# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
+# 
+# # Saving the data
+# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
+```
+
+# Extracting G1SST
+
+```{r}
+# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
+# library(stringr)
+# library(tidyverse)
+# library(reshape2)
+# library(ncdf4) # library for processing netCDFs
+# library(plyr)
+# library(lubridate)
+# library(data.table)
+# library(doMC); doMC::registerDoMC(cores = 7)
+# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
+# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
+# 
+# # #          1         2         3         4
+# # # 12345678901234567890123456789012345678901
+# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
+# #          1         2         3         4         5         6
+# # 1234567890123456789012345678901234567890123456789012345678901
+# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
+# 
+# 
+# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
+#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
+#     coords <- bbox[, region]
+#     nc <- nc_open(ncFile)
+#     pathLen <- nchar(paste0(ncDir, "/")) + 1
+#     fNameStem <-
+#       substr(basename(ncFile), 10, 58)
+#     fDate <- substr(basename(ncFile), 1, 8)
+#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
+#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
+#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
+#       round(4)
+#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
+#                           lat = nc$dim$lat$vals)
+#     nc_close(nc)
+#     sst <- as_tibble(melt(sst, value.name = "temp"))
+#     sst$t <- ymd(fDate)
+#     na.omit(sst)
+#     fwrite(sst,
+#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
+#            append = TRUE, col.names = FALSE)
+#     rm(sst)
+#   }
+# 
+# 
+#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
+#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
+#   endDate <- str_sub(ncLast, start = 1, end = 8)
+# 
+# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
+
+```
+
+# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
+# Is this upwelling event seen throughout the different distances from the coastline
+# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
+
+## Function
+Determining the temperatures at the various distances from the coastline
+
+```{r}
+# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
+# MUR <- MUR %>%
+#   select(lon, lat) %>%
+#   mutate(product = "MUR")
+# 
+# # Decided to work with OISST and CMC as both has a time series of 30years
+OISST_prod <- OISST %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "OISST")
+# 
+CMC_prod <- CMC %>%
+  select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "CMC")
+# 
+sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+#   #rbind(., CMC) %>% 
+select(product, lon, lat)
+# 
+sat_pixels <- sat_data %>%
+select(product, lon, lat) %>%
+unique()
+# 
+# ## For testing the nest/map pipeline
+# # df <- site_pixels %>%
+# #   filter(site == "Lamberts Bay") %>%
+# #   select(-site)
+# 
+match_func <- function(df){
+  df <- df %>%
+    dplyr::rename(lon_site = lon, lat_site = lat)
+  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
+                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  res <- rbind(OISST_index, CMC_index)
+  return(res)
+}
+
+# # Find the nearest pixels to each spot along the transect
+# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
+# # points to not fall inside of the same 25 KM pixel
+pixel_match <- site_pixels %>%
+  group_by(site) %>%
+  group_modify(~match_func(.x))
+
+# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
+OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
+# 
+# Clean up some RAM space
+rm(OISST, CMC); gc()
+# 
+# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
+# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
+# # # Hout Bay will be ignored
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+# 
+OISST_fill <- OISST_fill %>%
+  filter(site %in% selected_sites)
+
+CMC_fill <- CMC_fill %>%
+  filter(site %in% selected_sites)
+# 
+# 
+# # All the code below can be viewed in upwell_IDX.Rmd
+# # Next to run upwelling index
+# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+# load("Data/UI_angle.RData")
+# 
+upwelling <- UI_angle %>%
+  dplyr::rename(temp = ui.saws) %>%
+  group_by(site) %>%
+  # mutate(min_t = min(t),
+  #        max_t = max(t)) %>%
+  nest() %>% # apply the following functions to all of the variables in te dataset
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+         # NB: A threshold of 3 appeared to be far to strict
+         # purr::map - apllies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+Upwelling can last for a few hours.
+Given that we have daily data, upwelling events minimum duration here will be 1day
+  # Detect consecutive days in exceedance of a given threshold.
+  # mutate() %>%
+  select(-data, -clim) %>%
+  unnest() %>%
+  filter(row_number() %% 2 == 1) %>%
+  unnest() %>% # creates a column for each variables
+  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
+  select(site, t, ui.saws, exceedance)
+
+# Now applying the Upwelling func
+# NB: This only pulls out the event results and not the climatology results
+# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
+# To create a list output one would use group_map(),
+# but this then loses the labels of which sites etc. the results belong to
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
+  return(res)
+}
+# 
+ts2clm_custom <- function(df){
+  # The climatology base period used here is up for debate...
+  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  return(res)
+}
+# 
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+  upwell_base <- df %>%
+    dplyr::rename(t = date) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+  }
+# 
+OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
+save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
+save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
+# 
+# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
+SACTN_US <- SACTN_US %>%
+  filter(site %in% selected_sites)
+
+load("Data/site_list_v4.2.RData")
+
+SACTN_upwell_base <- SACTN_US %>%
+  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
+    dplyr::rename(t = date) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+
+save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+```
+
+Loading the data
+    This is the data obtained once the upwelling signals were filtered out
+
+```{r}
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
+
+# Removing the distance of 20 and 40kms
+library(dplyr)
+removing_distance_func <- function(df){
+  removing_dist_func<- df %>% 
+       filter(!distance %in% c(20000,40000))
+}
+
+OISST_final <- removing_distance_func(df = OISST_upwell_base)
+G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+MUR_final <- removing_distance_func(df = MUR_upwell_base)
+CMC_final <- removing_distance_func(df = CMC_upwell_base)
+
+# save(OISST_final, file = "Data/OISST_final.RData")
+# save(G1SST_final, file = "Data/G1SST_final.RData")
+# save(MUR_final, file = "Data/MUR_final.RData")
+# save(CMC_final, file = "Data/CMC_final.RData")
+```
+
+To accurately compare 
+  Start date CMC start date 1991-09-02 end date 2017-03-15
+  Start date OISST start date 1981-09- 01  and end date 2018-09-28
+  Start date SACTN start date 1973 and end date 2017-12-16
+
+```{r}
+#glm to determine if there is a difference in the number of events recorded at different at every distance
+
+load("Data/Count/OISST_count.RData")
+load("Data/Count/G1SST_count.RData")
+load("Data/Count/MUR_count.RData")
+load("Data/Count/CMC_count.RData")
+G1SST_count <- G1SST_count %>% 
+  mutate(src = "G1SST")
+
+combined <- rbind(CMC_count,G1SST_count,MUR_count,OISST_count)
+
+options(scipen = 999)
+
+anova_func <- function(df){
+  sites_aov <- aov(distance ~ no_observation * src, data = df)
+return(sites_aov)
+}
+
+combined <- anova_func(df = combined)
+summary(combined)
+
+m_coeff <- function(df){
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
+}
+# Changes in upwelling metrics
+lm_func <- function(df){
+  upwell_lm <- df %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var, distance) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+}
+
+OISST_lm <- lm_func(df = OISST_upwell_base)
+CMC_lm <- lm_func(df = CMC_upwell_base)
+
+SACTN_lm <- SACTN_upwell_base %>% 
+  select(-c(index_start:index_end)) %>% 
+  gather(key = "var", value = "val", -c(site:date_end)) %>% 
+  group_by(site, var) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() %>% 
+  # convert from daily to decadal values
+  mutate(slope = slope * 365.25*10)
+```
+
+As a common date we will use a start date from 1992-01-31 and end date 16-12-31
+
+```{r}
+# Generalosed linear models
+
+# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
+# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)
+
+# trends_duration_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
+#                      family = poisson(link = "log"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
+# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
+# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
+# 
+# 
+# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
+# trends_intensity_func <- function(df){
+#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
+#                      family = gaussian(link = "identity"))
+#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
+#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
+#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
+#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
+# }
+# 
+# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
+# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
+# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
+```
+
+# From the years 1992 - 2017
+
+```{r}
+# Which pixels showed the highest upwelling counts within the different datasets
+# Which site showed the most intense upwelling within the different datasets
+
+metric_func <- function(df){
+ metrics <-df %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  #mutate(year = year(date_start)) %>% 
+  group_by(distance, site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+   select(site, distance, mean_intensity, sum_events)
+}
+
+OISST_metrics <- metric_func(OISST_upwell_base)
+CMC_metrics <- metric_func(CMC_upwell_base)
+
+# No of u pwelling events found at the different sites
+SACTN_metrics <- SACTN_upwell_base %>% 
+  filter(year(date_start) %in% 1992:2017) %>% 
+  group_by(site) %>% 
+  summarise(mean_intensity = mean(intensity_mean),
+            sum_events = sum(event_no)) %>% 
+  select(site, mean_intensity, sum_events)
+
+# Create a table showing these metrics
+
+```
+
+## Comparing the relationship between 10 km and 30km from coastline and 10km and 50km from the coastline
+
+```{r}
+
+load("Data/Count/OISST_count.RData")
+load("Data/Count/G1SST_count.RData")
+load("Data/Count/MUR_count.RData")
+load("Data/Count/CMC_count.RData")
+G1SST_count <- G1SST_count %>% 
+  mutate(src = "G1SST")
+
+combined <- rbind(CMC_count,G1SST_count,MUR_count,OISST_count)
+```

### 2020-04-02 upwell_IDX.Rmd
update

@@
     ggtitle(titles[title_choice])
+}south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
+
+# Plotting the map
+world_map <- ggplot() + 
+  borders(fill = "grey40", colour = "black")
+
+# Create titles
+titles <- c("Shore-normal", "Islands")
+# Plotting function
+plot_sites <- function(site_list, buffer, title_choice, dist){
+
+  # Find the point 200 km from the site manually to pass to ggplot
+  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
+                                              b = site_list$heading, d = dist))
+
+  # Add the new coordinates tot he site list
+  site_list <- site_list %>% 
+    mutate(lon_dest = heading2$lon,
+           lat_dest = heading2$lat)
+
+  # Visualise
+  world_map +
+    geom_segment(data = site_list, colour = "red4", 
+                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
+    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
+    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
+    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
+                             max(site_list$lon + buffer)),
+                    ylim = c(min(site_list$lat - buffer), 
+             -                max(site_list$lat + buffer))) +
+    labs(x = "", y = "", colour = "Site\norder") +
+    ggtitle(titles[title_choice])
 }
+south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
+south_africa_away
 south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)

### 2020-04-06 SST_patterns.Rmd
UPDATE
update

@@
 
-## Comparing the relationship between 10 km and 30km from coastline and 10km and 50km from the coastline
-
-```{r}
-
-load("Data/Count/OISST_count.RData")
-load("Data/Count/G1SST_count.RData")
-load("Data/Count/MUR_count.RData")
-load("Data/Count/CMC_count.RData")
-G1SST_count <- G1SST_count %>% 
-  mutate(src = "G1SST")
-
-combined <- rbind(CMC_count,G1SST_count,MUR_count,OISST_count)
-```
+## Using series of general linear hypotheses (Hothorn et al., 2008) to look for differences in upwelling signal metrics between datasets and between sites
+# Is there a significant difference in the count between the datasets at a particular site
+
+```{r}
+load("Data/OISST_upwell_base.RData")
+load("Data/CMC_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_upwell_base.RData")
+load("Data/G1SST_upwell_base.RData")
+```

### 2020-04-06 SST_patterns.Rmd
update

--- 2020-04-02 SST_patterns.Rmd
@@
 output: html_document
----
-# This script is for all the various datasets
-
-# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
-
-```{r}
+
 knitr::opts_chunk$set(
@@
  metrics <-df %>% 
-  filter(year(date_start) %in% 1992:2017) %>% 
+  filter(year(date_start) %in% 2011:2014) %>% 
   #mutate(year = year(date_start)) %>% 
@@
 
-# No of u pwelling events found at the different sites
+# No of upwelling events found at the different sites
 SACTN_metrics <- SACTN_upwell_base %>% 
-  filter(year(date_start) %in% 1992:2017) %>% 
+  filter(year(date_start) %in% 2011:2014) %>% 
   group_by(site) %>% 
@@
 
-## Comparing the relationship between 10 km and 30km from coastline and 10km and 50km from the coastline
-
-```{r}
-
-load("Data/Count/OISST_count.RData")
-load("Data/Count/G1SST_count.RData")
-load("Data/Count/MUR_count.RData")
-load("Data/Count/CMC_count.RData")
-G1SST_count <- G1SST_count %>% 
-  mutate(src = "G1SST")
-
-combined <- rbind(CMC_count,G1SST_count,MUR_count,OISST_count)
-```
+## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites
+# Is there a significant difference in the duration/intensity etc between the datasets at a particular site
+
+```{r}
+load("Data/OISST_final.RData")
+load("Data/G1SST_final.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/CMC_final.RData")
+
+combined_products <- rbind(OISST_final,CMC_final,SACTN_upwell_base,MUR_final,G1SST_final)
+
+metric_4years <- combined_products %>% 
+   filter(year(date_start) %in% 2011:2014)
+
+```

### 2020-04-06 upwell_IDX.Rmd
update

@@
 ```
-
-# Bits of code testing the thresholds against the upwelling literature
-##################### Upwelling ######################
-# Min duration for an upwelling event - 1day and threshold -1. Upwelling can over within a few hours, but given that we have daily data - upwelling has a minimum duration of 1 day
-# Seasonal upwelling. Upwelling seasons are summer and spring
-# Did the wind intensify? This is usually done by taking first third of the dataset and the last thid of the dataset and doing a t-test on this
-# Strong south easterly winds lead to a stronger upwelling event
-##############################################

### 2020-04-06 SST_patterns.Rmd
update

@@
 
-## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites
+## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites (?)
 # Is there a significant difference in the duration/intensity etc between the datasets at a particular site
@@
 ```{r}
+
+library(tidyverse)
+library(multcomp)
+
 load("Data/OISST_final.RData")
 load("Data/G1SST_final.RData")
-load("Data/SACTN_upwell_base.RData")
+#load("Data/SACTN_upwell_base.RData")
 load("Data/MUR_final.RData")
@@
 
-combined_products <- rbind(OISST_final,CMC_final,SACTN_upwell_base,MUR_final,G1SST_final)
+combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
 
@@
 
-```
+metrics <- metric_4years %>% 
+  mutate(year = year(date_start)) %>% 
+  group_by(product, site) %>% 
+  summarise(y = n(),
+            mean_intensity = mean(intensity_mean),
+            mean_dur = mean(duration),
+            mean_cumIn = mean(intensity_cumulative)) %>% 
+  rename(count = y)
+
+lmod <- lm(mean_intensity ~ mean_dur, data = metrics)
+summary(lmod)
+
+glht(lmod, linfct = K)
+
+```

### 2020-04-06 SST_patterns.Rmd
update

@@
 metric_4years <- combined_products %>% 
-   filter(year(date_start) %in% 2011:2014)
+   filter(year(date_start) %in% 2011:2014) # only for the years 2011-2014 so 4 year period
 
@@
 
-lmod <- lm(mean_intensity ~ mean_dur, data = metrics)
-summary(lmod)
-
-glht(lmod, linfct = K)
-
-```
+# lmod <- lm(mean_intensity ~ mean_dur, data = metrics)
+# summary(lmod)
+# 
+# glht(lmod, linfct = K)
+
+```

### 2020-04-07 SST_patterns.Rmd
update
glht

@@
 library(tidyverse)
+library(lubridate)
 library(multcomp)
+options(scipen = 999)
 
@@
 
-# lmod <- lm(mean_intensity ~ mean_dur, data = metrics)
-# summary(lmod)
-# 
-# glht(lmod, linfct = K)
-
-```
+# anova_func <- function(df){
+#   currents_aov <- aov(mean_intensity ~ mean_dur  * site, data = df)
+#   return(currents_aov)
+# }
+# anova_metrics <- anova_func(df = metrics)
+# summary(anova_metrics)
+# 
+# anova <- function(df){
+#   currents_aov <- aov(mean_intensity ~ mean_dur  * product, data = df)
+#   return(currents_aov)
+# }
+# anova_metrics <- anova(df = metrics)
+# summary(anova_metrics)
+
+lm_coeff <- function(df){
+  res <- lm(formula = val ~  , data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 2)
+}
+
+lm_metrics <- metrics %>% 
+  gather(key = "var", value = "val", -c(product:site)) %>% 
+  group_by(site, var, product) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest()
+
+### define coefficients of linear function directly
+K <- diag(length(coef(lm_metrics)))[-1,]
+  rownames(K) <- names(coef(lm_metrics))[-1]
+K
+
+### set up general linear hypothesis
+  glht(lm_metrics, linfct = K)
+
+```

### 2020-04-07 general_linear_hypotheses.R
general linear hypothesis

@@ -1,0 +1,65 @@
+## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites (?)
+## Is there a significant difference in the duration/intensity etc between the datasets at a particular site
+
+library(tidyverse)
+library(lubridate)
+library(multcomp)
+options(scipen = 999)
+
+load("Data/OISST_final.RData")
+load("Data/G1SST_final.RData")
+#load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/CMC_final.RData")
+
+combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+
+metric_4years <- combined_products %>% 
+  filter(year(date_start) %in% 2011:2014) # only for the years 2011-2014 so 4 year period
+
+
+metrics <- metric_4years %>% 
+  mutate(year = year(date_start)) %>% 
+  group_by(product, site) %>% 
+  summarise(y = n(),
+            mean_intensity = mean(intensity_mean),
+            mean_dur = mean(duration),
+            mean_cumIn = mean(intensity_cumulative)) %>% 
+  rename(count = y)
+
+# anova_func <- function(df){
+#   currents_aov <- aov(mean_intensity ~ mean_dur  * site, data = df)
+#   return(currents_aov)
+# }
+# anova_metrics <- anova_func(df = metrics)
+# summary(anova_metrics)
+# 
+# anova <- function(df){
+#   currents_aov <- aov(mean_intensity ~ mean_dur  * product, data = df)
+#   return(currents_aov)
+# }
+# anova_metrics <- anova(df = metrics)
+# summary(anova_metrics)
+
+
+lm_coeff <- function(df){
+  res <- lm(formula = val ~  , data = df)
+  res_coeff <- round(as.numeric(res$coefficient[2]), 2)
+}
+
+lm_metrics <- metrics %>% 
+  gather(key = "var", value = "val", -c(product:site)) %>% 
+  group_by(site, var, product) %>% 
+  nest() %>% 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  select(-data) %>% 
+  unnest() 
+
+### define coefficients of linear function directly
+K <- diag(length(coef(lm_metrics)))[-1,]
+rownames(K) <- names(coef(lm_metrics))[-1]
+K
+
+### set up general linear hypothesis
+glht(lm_metrics, linfct = K)
+

### 2020-04-07 general_linear_hypotheses.R
update

@@ -51,7 +51,8 @@ lm_metrics <- metrics %>%
   gather(key = "var", value = "val", -c(product:site)) %>% 
   group_by(site, var, product) %>% 
   nest() %>% 
-  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  #mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  #mutate(model = purrr::map(data, ~lm( ~ value))) %>% 
   select(-data) %>% 
   unnest() 
 

### 2020-04-08 site_map.Rmd
update

--- 2020-01-18 site_map.Rmd
@@
 site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + 
-  # geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
+  #geom_raster(data = OISST, aes(x = lon, y = lat, fill = temp)) +
    geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                 fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
-  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 5) +
+  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
   geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
-  geom_label_repel(data =site_list_sub_map, aes(x = lon, y = lat, label = site), 
+  geom_label_repel(data =site_list_sub, aes(x = lon, y = lat, label = site), 
                      size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2, 
                      segment.alpha = 0.4, force = 0.1) +
-  coord_equal(xlim = c(13, 23), ylim = c(-36,-26), expand = 0) +
+  coord_equal(xlim = c(14, 21.5), ylim = c(-35,-27), expand = 0) +
   # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
@@
   # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
-  scale_fill_viridis(
-    option = "magma",
-    name = "Temperature (°C)",
-    alpha = 0.8) +
+  # scale_fill_viridis(
+  #   option = "magma",
+  #   name = "Temperature (°C)",
+  #   alpha = 0.8) +
   xlab("") + ylab("") +
@@
     annotate("text", label = "ATLANTIC\nOCEAN", x = 15.00, y = -33.0,
-             size =6, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 14.5, y = -30,
-             size = 6, angle = 302, colour = "black") +
+             size =5, angle = 0, colour = "black") +
+    annotate("text", label = "Benguela", x = 15, y = -30,
+             size = 5, angle = 302, colour = "black") +
     # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.major = element_blank(),
+  theme(axis.text = element_text(size = rel(1.2), colour = "black"),
+        axis.text.x = element_text(vjust = 1),
+        axis.text.y = element_text(hjust = 1),
+        axis.title.x = element_text(vjust = 0),
+        axis.title.y = element_text(angle = 90, vjust = 0.3),
+        axis.ticks = element_line(colour = "black"),
+        axis.ticks.length = unit(0.15, "cm"),
+        panel.background = element_rect(fill = "white", colour = "black", size = 0.5),
+        #panel.border = element_blank(),
+        panel.border = element_rect(colour = "black", fill=NA, size=2),
+        #(colour =" black"),
+        panel.grid.major.x = element_blank(),
+        #panel.grid.major.y = element_line(linetype = "dotted", size = 0.2, colour = "black"),
+        panel.grid.major.y = element_blank(),
         panel.grid.minor = element_blank(),
-        legend.text = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5),
-        axis.text = element_text(size = 12)) +
-    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
-  theme(panel.background = element_rect(fill = "ivory", colour = NA),
-        panel.border = element_rect(colour = "black", size = 0.5),
-        panel.grid.minor = element_line(colour = "NA"),
-        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
-        legend.direction = "horizontal",
-        legend.justification = c(1,0),
-        legend.position = c(0.85, 0.70),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        legend.key = element_blank(),
-        legend.background = element_blank(),
-        axis.ticks = element_line(size = 0.5))
+        #panel.margin = unit(0.25, "lines"),
+        strip.text = element_text(size = rel(1.0)),
+        strip.background = element_rect(fill = "white", colour = NA),
+        strip.text.x = element_text(),
+        strip.text.y = element_text(angle = -90),        
+        plot.background = element_rect(colour = "white"),
+        plot.title = element_text(size = rel(1.2), vjust = 1),
+        plot.margin = unit(c(1, 1, 0.5, 0.5), "lines"))
+
 site_map

### 2020-04-08 Themes.R

@@ -1,0 +1,92 @@
+# Themes
+
+theme_Publication <- function(base_size=14, base_family="helvetica") {
+  library(grid)
+  library(ggthemes)
+  (theme_foundation(base_size=base_size, base_family=base_family)
+    + theme(plot.title = element_text(face = "bold",
+                                      size = rel(1.2), hjust = 0.5),
+            text = element_text(),
+            panel.background = element_rect(colour = NA),
+            plot.background = element_rect(colour = NA),
+            panel.border = element_rect(colour = NA),
+            axis.title = element_text(face = "bold",size = rel(1)),
+            axis.title.y = element_text(angle=90,vjust =2),
+            axis.title.x = element_text(vjust = -0.2),
+            axis.text = element_text(), 
+            axis.line = element_line(colour="black"),
+            axis.ticks = element_line(),
+            panel.grid.major = element_line(colour="#f0f0f0"),
+            panel.grid.minor = element_blank(),
+            legend.key = element_rect(colour = NA),
+            legend.position = "bottom",
+            legend.direction = "horizontal",
+            legend.key.size= unit(0.2, "cm"),
+            legend.margin = unit(0, "cm"),
+            #legend.title = element_text(face="italic"),
+            plot.margin=unit(c(10,5,5,5),"mm"),
+            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+            strip.text = element_text(face="bold")
+    ))
+  
+}
+
+scale_fill_Publication <- function(...){
+  library(scales)
+  discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+  
+}
+
+scale_colour_Publication <- function(...){
+  library(scales)
+  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
+  
+}
+
+library("grid")
+theme_sharp <- function(base_size = 10, base_family = "")
+{
+  theme(line = element_line(colour = "black", size = 0.5, linetype = 1, lineend = "butt"),
+        rect = element_rect(fill = "white", colour = "black", size = 0.5, linetype = 1),
+        text = element_text(family = base_family, face = "plain",
+                            colour = "black", size = base_size,
[…]

### 2020-04-08 site_map.Rmd
update

--- 2020-01-18 site_map.Rmd
@@
     annotate("text", label = "ATLANTIC\nOCEAN", x = 15.00, y = -33.0,
-             size =6, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 14.5, y = -30,
-             size = 6, angle = 302, colour = "black") +
+             size =5, angle = 0, colour = "black") +
+    # annotate("text", label = "Benguela", x = 15, y = -30,
+    #          size = 5, angle = 302, colour = "black") +
     # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,

### 2020-04-08 site_map.Rmd
update

--- 2020-04-08 site_map.Rmd
@@
                      segment.alpha = 0.4, force = 0.1) +
-  coord_equal(xlim = c(14, 21.5), ylim = c(-35,-27), expand = 0) +
+  coord_equal(xlim = c(15, 21.5), ylim = c(-35,-27), expand = 0) +
   # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
@@
   #            size = 4, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 15.00, y = -33.0,
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 16, y = -31.5,
              size =5, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 15, y = -30,
-             size = 5, angle = 302, colour = "black") +
+    # annotate("text", label = "Benguela", x = 15, y = -30,
+    #          size = 5, angle = 302, colour = "black") +
     # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,

### 2020-04-08 general_linear_hypotheses.R
Some thoughts

@@ -1,5 +1,5 @@
-## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites (?)
-## Is there a significant difference in the duration/intensity etc between the datasets at a particular site
+## Using a series of general linear hypotheses to look for differences in upwelling signal metrics between datasets and between sites (?)
+## Is there a significant difference in the duration/intensity etc between the datasets at a particular site?
 
 library(tidyverse)
 library(lubridate)
@@ -7,6 +7,7 @@ library(multcomp)
 options(scipen = 999)
 
 load("Data/OISST_final.RData")
+OISST_final$distance_km <- OISST_final$distance/1000 # This column is missing from this dataframe
 load("Data/G1SST_final.RData")
 #load("Data/SACTN_upwell_base.RData")
 load("Data/MUR_final.RData")
@@ -19,7 +20,7 @@ metric_4years <- combined_products %>%
 
 
 metrics <- metric_4years %>% 
-  mutate(year = year(date_start)) %>% 
+  # mutate(year = year(date_start)) %>% # Why is this here? It is removed in the summarise step.
   group_by(product, site) %>% 
   summarise(y = n(),
             mean_intensity = mean(intensity_mean),
@@ -43,7 +44,7 @@ metrics <- metric_4years %>%
 
 
 lm_coeff <- function(df){
-  res <- lm(formula = val ~  , data = df)
+  res <- lm(formula = val ~  date_peak, data = df)
   res_coeff <- round(as.numeric(res$coefficient[2]), 2)
 }
 
@@ -51,9 +52,9 @@ lm_metrics <- metrics %>%
   gather(key = "var", value = "val", -c(product:site)) %>% 
   group_by(site, var, product) %>% 
   nest() %>% 
-  #mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  # mutate(slope = purrr::map(data, lm_coeff)) %>% # Neither of these lines of code run
   #mutate(model = purrr::map(data, ~lm( ~ value))) %>% 
-  select(-data) %>% 
+  select(-data) %>%
   unnest() 
 
 ### define coefficients of linear function directly
@@ -63,4 +64,12 @@ K
 
 ### set up general linear hypothesis
 glht(lm_metrics, linfct = K)
+
+### Why not just perform a normal ANOVA?
+# Why go to all this extra trouble?
+# I would recommend reading up more on ANOVA
+summary(aov(duration ~ site * product * distance, data = metric_4years))
[…]

### 2020-04-08 general_linear_hypotheses.R
ANOVA analysis on upwelling metrics and linear model coefficients therefrom.

@@ -28,48 +28,59 @@ metrics <- metric_4years %>%
             mean_cumIn = mean(intensity_cumulative)) %>% 
   rename(count = y)
 
-# anova_func <- function(df){
-#   currents_aov <- aov(mean_intensity ~ mean_dur  * site, data = df)
-#   return(currents_aov)
-# }
-# anova_metrics <- anova_func(df = metrics)
-# summary(anova_metrics)
-# 
-# anova <- function(df){
-#   currents_aov <- aov(mean_intensity ~ mean_dur  * product, data = df)
-#   return(currents_aov)
-# }
-# anova_metrics <- anova(df = metrics)
-# summary(anova_metrics)
-
-
+# Function for extracting slope from linear model
 lm_coeff <- function(df){
-  res <- lm(formula = val ~  date_peak, data = df)
-  res_coeff <- round(as.numeric(res$coefficient[2]), 2)
+  res <- lm(formula = val ~ date_peak, data = df)
+  res_coeff <- as.numeric(res$coefficient[2])
 }
 
-lm_metrics <- metrics %>% 
-  gather(key = "var", value = "val", -c(product:site)) %>% 
-  group_by(site, var, product) %>% 
+# Calculate all of the linear models
+lm_metrics <- metric_4years %>% 
+  ungroup() %>% 
+  dplyr::select(-heading, -c(lon:index_end), -distance_km) %>% 
+  pivot_longer(cols = c(duration, intensity_mean:rate_decline), 
+               names_to = "var", values_to = "val") %>% 
+  group_by(site, product, distance, var) %>% 
   nest() %>% 
-  # mutate(slope = purrr::map(data, lm_coeff)) %>% # Neither of these lines of code run
-  #mutate(model = purrr::map(data, ~lm( ~ value))) %>% 
-  select(-data) %>%
-  unnest() 
+  mutate(slope = purrr::map(data, lm_coeff)) %>% 
+  dplyr::select(-data) %>% 
+  unnest(cols = slope) %>% 
+  # convert from daily to decadal values
+  mutate(slope = round((slope*365.25*10), 2)) %>% 
+  ungroup()
+
+# Cast the results wide for easier use with ANOVA
+lm_metrics_wide <- pivot_wider(lm_metrics, 
+                               id_cols = site:distance, 
+                               names_from = var, values_from = slope)
+
+### test example
+lmod <- lm(Fertility ~ ., data = swiss)
 
[…]

### 2020-04-08 general_linear_hypotheses.R
update

@@ -55,8 +55,6 @@ lm_metrics_wide <- pivot_wider(lm_metrics,
                                names_from = var, values_from = slope)
 
 ### test example
-lmod <- lm(Fertility ~ ., data = swiss)
-
 ### define coefficients of linear function directly
 K <- diag(length(coef(lmod)))[-1,]
 rownames(K) <- names(coef(lmod))[-1]

### 2020-04-16 SST_patterns.Rmd
update

--- 2020-04-06 SST_patterns.Rmd
@@
   facet_wrap(~site) +
-theme_bw() ---
-title: "SST_patterns"
-author: "Amieroh Abrahams"
-date: "19 August 2019"
-output: html_document
-
-knitr::opts_chunk$set(
-  comment = "R>",
-  warning = FALSE,
-  message = FALSE 
-)
-#update.packages() 
-library(tidyverse)
-library(lubridate)
-library(ggpubr)
-library(zoo)
-library(FNN)
-library(scales)
-library(gridExtra)
-library(circular)
-library(fossil)
-library(mapproj)
-library(broom)
-# library(ncdf4) # This was used to process NetCDF files in an earlier version
-library(stringr)
-library(doMC); doMC::registerDoMC(cores = 4)
-library(fasttime)
-library(xtable)
-library(ncdf4) # library for processing netCDFs
-library(data.table)
-# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
-library(heatwaveR)
-library(grid)
-library(ggthemes)
-```
-
+```
 # Loading in all the data created using the code bellow
@@
   rename(count = y)
-
-# lmod <- lm(mean_intensity ~ mean_dur, data = metrics)
-# summary(lmod)
-# 
-# glht(lmod, linfct = K)
-
-```
+```

### 2020-04-16 box_plot.R

@@ -1,0 +1,88 @@
+# Loading libraries
+library(tidyverse)
+library(lubridate)
+library(multcomp)
+options(scipen = 999)
+
+load("Data/OISST_final.RData")
+load("Data/G1SST_final.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/CMC_final.RData")
+
+combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+
+seasons_func <- function(df){
+  BC_seaons <- df %>% 
+    mutate(month = month(as.Date(as.character(date_start)), abbr = T, label = T),
+           year = year(date_start)) %>% 
+    mutate(season = ifelse(month %in% c("Dec", "Jan", "Feb"), "Summer",        
+                           ifelse(month %in% c("Mar", "Apr", "May"), "Autumn",
+                                  ifelse(month %in% c("Jun", "Jul", "Aug"), "Winter",
+                                         ifelse(month %in% c("Sep", "Oct", "Nov"), "Spring","Error")))))
+}
+
+combined_products <- seasons_func(df = combined_products)
+SACTN <- seasons_func(df = SACTN_upwell_base)
+
+
+metric_prods <- combined_products %>% 
+  filter(year(date_start) %in% 2011:2014) %>% 
+  filter(season == "Summer")
+
+metric_SACTN <- SACTN %>% 
+  filter(year(date_start) %in% 2011:2014) %>% 
+  mutate(product = "SACTN")
+
+combined <- rbind(metric_prods, metric_SACTN)
+# write.csv(combined,'combined.csv')
+
+metrics_func <- function(df){
+  metrics <- df %>% 
+  mutate(year = year(date_start)) %>% 
+  group_by(product, site) %>% 
+  summarise(y = n(),
+            mean_intensity = mean(intensity_mean),
+            mean_dur = mean(duration),
+            mean_cumIn = mean(intensity_cumulative)) %>% 
+  rename(count = y)
+}
+
+
+metrics <- metrics_func(df = metric_prods)
+metrics_SACTN <- metrics_func(df = metric_SACTN)
+
+final_combined <- rbind(metrics, metric_SACTN)
+final_combined$count <- as.numeric(final_combined$count)
+
+# ggplot(data = final_combined, aes(x = mean_dur, y = count)) +
+#   geom_boxplot(aes(colour = product)) +
+#   facet_wrap(~site) 
+# 
+# 
[…]

### 2020-04-16 site_map.Rmd
update

--- 2020-04-08 site_map.Rmd
@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
+  coord_fixed(ratio = 1, xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5),
+                expand = TRUE) +
   theme(axis.text = element_text(size = rel(1.2), colour = "black"),

### 2020-04-16 box_plot.R

@@ -68,11 +68,12 @@ final_combined$count <- as.numeric(final_combined$count)
 
 counts <- read_csv("Data/counts.csv")
 
-ggplot(data = counts, aes(x = duration, y = count)) +
-  geom_boxplot(aes(colour = product)) +
+ggplot(data = combined, aes(x = product,y = duration)) +
+  geom_boxplot(aes(fill = product)) +
   facet_wrap(~site)  +
-  labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "SST products")+
+  labs(y = "Duration (Days)", x = "SST product")+
   theme_bw()+
+  theme(legend.position="none") +
   theme(axis.text = element_text(size = 12),
         axis.title = element_text(size = 12, face = "bold"),
         legend.text = element_text(size = 12),

### 2020-04-16 site_map.Rmd
update

--- 2020-04-08 site_map.Rmd
@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
+  coord_fixed(ratio = 1, xlim = c(15, 21.5), ylim = c(-35,-27),
+                expand = TRUE) +
   theme(axis.text = element_text(size = rel(1.2), colour = "black"),
@@
         axis.ticks.length = unit(0.15, "cm"),
+        axis.ticks.margin = unit(0.1, "cm"),
         panel.background = element_rect(fill = "white", colour = "black", size = 0.5),

### 2020-04-20 SST_patterns.Rmd
update

@@
 load("Data/CMC.RData") # 1decimal places
+load("Data/MUR.RData")
+
+# ### The temperature products
+# load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
+# load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
+# load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
+# load("~/Documents/Upwelling_MCS/Data/OISST_sites.RData") # OISST
+# load("~/Documents/Upwelling_MCS/Data/CMC_SST.RData") # CMC temperature
 ```
@@
 load("Data/G1SST_final.RData")
-#load("Data/SACTN_upwell_base.RData")
+load("Data/SACTN_upwell_base.RData")
 load("Data/MUR_final.RData")
 load("Data/CMC_final.RData")
+SACTN_upwell_base <- SACTN_upwell_base %>% 
+  mutate(product = "SACTN")
+
+### The temperature products
+load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
+load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
+load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
+load("~/Documents/Upwelling_MCS/Data/OISST_sites.RData") # OISST
+load("~/Documents/Upwelling_MCS/Data/CMC_SST.RData") # CMC temperature
 
@@
    filter(year(date_start) %in% 2011:2014) # only for the years 2011-2014 so 4 year period
+
+SACTN <- SACTN_upwell_base %>% 
+   filter(year(date_start) %in% 2011:2014)
 
@@
   rename(count = y)
-```
+
+## Creating a time series graph with all of the products but showing the temperature drop but with no CMC
+# X-axis: Date, Y axis: Temp
+
+CMC_temp <- CMC %>% 
+  filter(site == "Lamberts Bay",
+         year(date) == 2014) %>% 
+  dplyr::rename(temp = temp_CMC) %>% 
+  mutate(product = "CMC") %>% 
+  dplyr::select(date,site,lat,lon,temp,product)#%>% 
+ # select(date,site,product,lat,lon,temp)
+
+OISST_temp <- OISST_sites %>% 
+  filter(site == "Lamberts Bay",
+         year(date)== 2014)%>% 
+  mutate(product = "OISST")%>% 
+  dplyr::select(date,site,lat,lon,temp,product)
+
+# SACTN_temp <- SACTN_US %>% 
+#   filter(site == "Lamberts Bay",
+#          year(date) == 2014)%>% 
+#   mutate(product = "SACTN") %>% 
+#   dplyr::select(date,site,lat,lon,temp,product)
+
+MUR_temp <- Final %>% 
+  filter(site == "Lamberts Bay",
+         year(date) == 2013)%>% 
+  mutate(product = "MUR")%>% 
+  dplyr::select(date,site,lat,lon,temp,product)
+
+G1SST_temp <- Final_G1SST %>% 
+  filter(site == "Lamberts Bay",
+         year(date) == 2014) %>% 
+  mutate(product = "G1SST")%>% 
+  dplyr::select(date,site,lat,lon,temp,product)
+
+combined_SST <- rbind(CMC_temp,MUR_temp,G1SST_temp,OISST_temp)
+
+ggplot(combined_SST, aes(x = date, y = temp, colour = product)) + 
+  geom_line()
+
+```

### 2020-04-20 site_map.Rmd
update

--- 2020-04-08 site_map.Rmd
@@
                      segment.alpha = 0.4, force = 0.1) +
-  coord_equal(xlim = c(15, 21.5), ylim = c(-35,-27), expand = 0) +
+  coord_equal(xlim = c(15.5, 20), ylim = c(-35,-27), expand = 0) +
   # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),

### 2020-04-20 SST_patterns.Rmd
update

@@
 
-# ### The temperature products
+# # ### The temperature products
 # load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
@@
 # load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-# load("~/Documents/Upwelling_MCS/Data/OISST_sites.RData") # OISST
-# load("~/Documents/Upwelling_MCS/Data/CMC_SST.RData") # CMC temperature
+# load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
+# load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
 ```
@@
 OISST_prod <- OISST %>%
-  select(lon, lat) %>%
+  dplyr::select(lon, lat) %>%
   unique() %>%
@@
 CMC_prod <- CMC %>%
-  select(lon, lat) %>%
+  dplyr::select(lon, lat) %>%
   unique() %>%
@@
 #   #rbind(., CMC) %>% 
-select(product, lon, lat)
+dplyr::select(product, lon, lat)
 # 
 sat_pixels <- sat_data %>%
-select(product, lon, lat) %>%
+dplyr::select(product, lon, lat) %>%
 unique()
@@
 CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
+
 # #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
-# 
 # Clean up some RAM space
@@
   filter(site %in% selected_sites)
-# 
-# 
+
+# save(OISST_fill, file = "Data/OISST_fill.RData")
+# save(CMC_fill, file = "Data/CMC_fill.RData")
+
 # # All the code below can be viewed in upwell_IDX.Rmd
@@
 
+combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+
 ### The temperature products
@@
 load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-load("~/Documents/Upwelling_MCS/Data/OISST_sites.RData") # OISST
-load("~/Documents/Upwelling_MCS/Data/CMC_SST.RData") # CMC temperature
-
-combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
+load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
 
 metric_4years <- combined_products %>% 
-   filter(year(date_start) %in% 2011:2014) # only for the years 2011-2014 so 4 year period
+   filter(year(date_start) %in% 2011:2014)
 
@@
 # X-axis: Date, Y axis: Temp
+combined_SST <- rbind(CMC_fill,MUR_temp,G1SST_temp,OISST_temp)
 

### 2020-04-20 SST_patterns.Rmd
time series

@@
 ```
-
-## Using series of general linear hypothese to look for differences in upwelling signal metrics between datasets and between sites (?)
-# Is there a significant difference in the duration/intensity etc between the datasets at a particular site
-
-```{r}
-
-library(tidyverse)
-library(lubridate)
-library(multcomp)
-options(scipen = 999)
-
-load("Data/OISST_final.RData")
-load("Data/G1SST_final.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_final.RData")
-load("Data/CMC_final.RData")
-SACTN_upwell_base <- SACTN_upwell_base %>% 
-  mutate(product = "SACTN")
-
-combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
-
-### The temperature products
-load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
-load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
-load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
-
-metric_4years <- combined_products %>% 
-   filter(year(date_start) %in% 2011:2014)
-
-SACTN <- SACTN_upwell_base %>% 
-   filter(year(date_start) %in% 2011:2014)
-
-metrics <- metric_4years %>% 
-  mutate(year = year(date_start)) %>% 
-  group_by(product, site) %>% 
-  summarise(y = n(),
-            mean_intensity = mean(intensity_mean),
-            mean_dur = mean(duration),
-            mean_cumIn = mean(intensity_cumulative)) %>% 
-  rename(count = y)
-
-## Creating a time series graph with all of the products but showing the temperature drop but with no CMC
-# X-axis: Date, Y axis: Temp
-combined_SST <- rbind(CMC_fill,MUR_temp,G1SST_temp,OISST_temp)
-
-CMC_temp <- CMC %>% 
-  filter(site == "Lamberts Bay",
-         year(date) == 2014) %>% 
-  dplyr::rename(temp = temp_CMC) %>% 
-  mutate(product = "CMC") %>% 
-  dplyr::select(date,site,lat,lon,temp,product)#%>% 
- # select(date,site,product,lat,lon,temp)
-
-OISST_temp <- OISST_sites %>% 
-  filter(site == "Lamberts Bay",
-         year(date)== 2014)%>% 
-  mutate(product = "OISST")%>% 
-  dplyr::select(date,site,lat,lon,temp,product)
-
-# SACTN_temp <- SACTN_US %>% 
-#   filter(site == "Lamberts Bay",
-#          year(date) == 2014)%>% 
-#   mutate(product = "SACTN") %>% 
-#   dplyr::select(date,site,lat,lon,temp,product)
-
-MUR_temp <- Final %>% 
-  filter(site == "Lamberts Bay",
-         year(date) == 2013)%>% 
-  mutate(product = "MUR")%>% 
-  dplyr::select(date,site,lat,lon,temp,product)
-
-G1SST_temp <- Final_G1SST %>% 
-  filter(site == "Lamberts Bay",
-         year(date) == 2014) %>% 
-  mutate(product = "G1SST")%>% 
-  dplyr::select(date,site,lat,lon,temp,product)
-
-combined_SST <- rbind(CMC_temp,MUR_temp,G1SST_temp,OISST_temp)
-
-ggplot(combined_SST, aes(x = date, y = temp, colour = product)) + 
-  geom_line()
-
-```

### 2020-04-20 time_series.R

@@ -1,0 +1,41 @@
+library(tidyverse)
+library(lubridate)
+library(multcomp)
+options(scipen = 999)
+
+# Loading upwelling metrics
+load("Data/OISST_final.RData")
+load("Data/G1SST_final.RData")
+load("Data/SACTN_upwell_base.RData")
+load("Data/MUR_final.RData")
+load("Data/CMC_final.RData")
+SACTN_upwell_base <- SACTN_upwell_base %>% 
+  mutate(product = "SACTN")
+
+combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+combined_products <- combined_products %>% 
+  filter(site == "Saldanha Bay") %>% 
+    filter(distance == "10000") %>% 
+  filter(year(date_start) %in% 2011:2014)
+
+### Loading the temperature products
+load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
+load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
+load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
+load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
+load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
+
+###
+# In Lamberts Bay in June 2012 signals detected below all at a distance of 10 km from the coastline
+# CMC - 2012-06-12 signal lasted 3 days
+#G1SST - 2012-06-10 signal lsted 36 days
+#MUR - 2012-06-10 signal lsted 36 days
+#OISST - 2012-06-10 signal lsted 18 days
+
+combined_SST <- rbind(OISST_fill,CMC_fill,Final,Final_G1SST)
+combined_SST <- combined_SST %>% 
+  filter(year(date) == 2012) # In 2011 in 
+
+# Plotting a time series
+ggplot(combined_SST, aes(x = date, y = temp, colour = product)) +
+  geom_line()

### 2020-04-20 time_series.R
graphs

@@ -1,6 +1,7 @@
 library(tidyverse)
 library(lubridate)
 library(multcomp)
+library(heatwaveR)
 options(scipen = 999)
 
 # Loading upwelling metrics
@@ -14,9 +15,9 @@ SACTN_upwell_base <- SACTN_upwell_base %>%
 
 combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
 combined_products <- combined_products %>% 
-  filter(site == "Saldanha Bay") %>% 
-    filter(distance == "10000") %>% 
-  filter(year(date_start) %in% 2011:2014)
+  filter(site == "Lamberts Bay") %>% 
+  #  filter(distance == "10000") %>% 
+  filter(year(date_start) == 2012)
 
 ### Loading the temperature products
 load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
@@ -25,7 +26,6 @@ load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
 load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
 load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
 
-###
 # In Lamberts Bay in June 2012 signals detected below all at a distance of 10 km from the coastline
 # CMC - 2012-06-12 signal lasted 3 days
 #G1SST - 2012-06-10 signal lsted 36 days
@@ -33,9 +33,17 @@ load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
 #OISST - 2012-06-10 signal lsted 18 days
 
 combined_SST <- rbind(OISST_fill,CMC_fill,Final,Final_G1SST)
-combined_SST <- combined_SST %>% 
-  filter(year(date) == 2012) # In 2011 in 
+combined_SST <- combined_SST %>%
+  filter(year(date) == 2012) # In 2011 in
 
 # Plotting a time series
 ggplot(combined_SST, aes(x = date, y = temp, colour = product)) +
   geom_line()
+
+
+ggplot(combined_products, aes(x = date_peak, y = duration)) +
+    geom_lolli(aes(colour = product)) +
+    xlab("Date") + ylab("Event duration [days]") +
+  theme_bw()
+
+

### 2020-04-20 site_map.Rmd
update

--- 2020-04-16 site_map.Rmd
@@
 site_list_sub_map <- site_list_sub %>%
-  select(site,lat,lon)
+  dplyr::select(site,lat,lon)
 
@@
 site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + 
-  #geom_raster(data = OISST, aes(x = lon, y = lat, fill = temp)) +
-   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
-                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
+   geom_polygon(aes(group = group),fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
   geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
@@
                      segment.alpha = 0.4, force = 0.1) +
-  coord_equal(xlim = c(15, 21.5), ylim = c(-35,-27), expand = 0) +
-  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
-  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
-  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
-  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
-  # scale_fill_viridis(
-  #   option = "magma",
-  #   name = "Temperature (°C)",
-  #   alpha = 0.8) +
   xlab("") + ylab("") +
-  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
-  #            size = 4, angle = 0, colour = "black") +
     annotate("text", label = "ATLANTIC\nOCEAN", x = 16, y = -31.5,
              size =5, angle = 0, colour = "black") +
-    # annotate("text", label = "Benguela", x = 15, y = -30,
-    #          size = 5, angle = 302, colour = "black") +
-    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
-    #          size = 5, angle = 50, colour = "black") +
   guides(fill = guide_colourbar()) +
@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
-  coord_fixed(ratio = 1, xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5),
+  coord_fixed(ratio = 1, xlim = c(15.5, 20), ylim = c(-35,-28),
                 expand = TRUE) +
@@
 site_map 
-
-# pdf("Figures/site_map.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
-# 
-# 
-# africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
-#   theme_bw() + 
-#   coord_equal() + 
-#   geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
-#   geom_polygon(data = sa_provinces_new, (aes(group = group))) +
-#   annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
-#   theme(panel.border = element_rect(colour = "black", size = 0.4), 
-#         plot.background = element_blank(), 
-#         axis.ticks = element_blank(), 
-#         axis.text = element_blank(), 
-#         axis.title = element_blank(), 
-#         panel.grid.major = element_blank(), 
-#         panel.grid.minor = element_blank()) +
-#   coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
-# africa_map
-# 
-# # More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html
-# 
-# vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
-# vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
-# print(site_map, vp = vp2)
-# print(africa_map, vp = vp1)
-# dev.off()
-# pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
 ```

### 2020-04-20 upwell_IDX.Rmd
Some comments

--- 2019-11-29 upwell_IDX.Rmd
@@
 The 160 degrees is used as this refers to the angle of the coastline.
-This equation is largely dependant on wind speed and direction data in order to determing the intensity of the upwelling event.
+This equation is largely dependant on wind speed and direction data in order to determining the intensity of the upwelling event.
 Wind data were obtained daily from the South African Weather Service (SAWS).
-This wind data was then matched to the date at which temperature were collected.
-With this data the upwelling index was determined.
+This wind data were then matched to the date at which temperatures were collected.
+With these data the upwelling index was determined.
 
@@
 ```{r}
-# Installling andloading libraries
+# Installing and loading libraries
 
@@
 
-```{r}
-# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
-# 
-# wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# # Slecting the important columns for each of the datasets
-# wind_fix <- function(df){
-# wind <- df %>%
-#   select(station_name, date, hour, dir, speed) %>%  #column names
-#   mutate(date = as.Date(as.character(date)),
-#          hour = as.numeric(as.character(hour)),
-#          dir = as.numeric(as.character(dir)),
-#          speed = as.numeric(as.character(speed)))
-# }
-# 
-# wind_1 <- wind_fix(df = wind_1)
-# wind_2 <- wind_fix(df = wind_2)
-# wind_3 <- wind_fix(df = wind_3)
-# 
-# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
-# 
-# renaming_sites_1 <- function(df) {
-#   sites <- df %>%
-#       mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
-#                                     station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
-#                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
-#                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
-#                                     station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
-#   return(sites)
-# }
-# 
-# 
-# wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
-# wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
-# wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
-# wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-# save(wind_data, file = "Data/wind_data.RData")
-# 
-# load("Data/wind_data.RData")
-# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
-# 
-# wind_daily <- wind_data %>%
-#   # ungroup() %>%
-#   # mutate(date = as.Date(date))%>% # Changing the date to date format
-#   dplyr::group_by(temp_sites, date) %>%
-#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2), # Direction using the circular function
-#                    mean_speed = round(mean(speed),2)) # Calculating the mean
-# save(wind_daily, file = "Data/wind_daily.RData")
+```{r, eval=FALSE}
+wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
+
+wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+# Selecting the important columns for each of the datasets
+## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+wind_fix <- function(df){
+wind <- df %>%
+  mutate(date = as.Date(as.character(date)),
+         hour = as.numeric(as.character(hour)),
+         dir = as.numeric(as.character(dir)),
+         speed = as.numeric(as.character(speed)),
+         temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
+                                station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
+                                station_name  == "PORT NOLLOTH" ~"Port Nolloth",
+                                station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
+                                station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
+                                station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
+  select(temp_sites, date, hour, dir, speed) %>% 
+  drop_na()
+}
+
+wind_fix_1 <- wind_fix(df = wind_1)
+wind_fix_2 <- wind_fix(df = wind_2)
+wind_fix_3 <- wind_fix(df = wind_3)
+
+# Combine and save
+wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
+save(wind_data, file = "Data/wind_data.RData")
+
+load("Data/wind_data.RData")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
+
+# RWS: Please figure out why some wind direction values are returned as NA
+# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
+# RWS: Please look through these resources and make corrections accordingly:
+# https://cran.r-project.org/web/packages/rWind/rWind.pdf
+# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
+wind_daily <- wind_data %>%
+  # ungroup() %>%
+  # mutate(date = as.Date(date))%>% # Changing the date to date format
+  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
+  dplyr::group_by(temp_sites, date) %>%
+  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
+                                                           template = 'geographics')),2), # Direction using the circular function
+                   mean_speed = round(mean(speed),2)) # Calculating the mean
+save(wind_daily, file = "Data/wind_daily.RData") 
 ```
@@
 # Transect function in coastR package finds the angle away or along the coastline
-south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
+south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30) %>% 
+  mutate(heading = ifelse(heading < 0, heading+360, heading))
 
@@
 
-# Create titles
-titles <- c("Shore-normal", "Islands")
 # Plotting function
-plot_sites <- function(site_list, buffer, title_choice, dist){
+plot_sites <- function(site_list, buffer, dist){
 
@@
                              max(site_list$lat + buffer))) +
-    labs(x = "", y = "", colour = "Site\norder") +
-    ggtitle(titles[title_choice])
+    labs(x = "", y = "", colour = "Site\norder")
 }
-south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
+south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
 south_africa_away
@@
 ```{r}
-UI_angle <- wind_daily %>%  # Making refeerence to the wind daily datta created above
+UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
   dplyr::rename(site = sites) %>%  # Renaming sites to site
-  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to south africa wide data column 2 and 21. by site 
-  dplyr::rename(coast_angle = heading) %>% # Renamingthe column heading to coast_angle
+  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
+  drop_na() %>% # removing the na values
+  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
   mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
-  drop_na %>% # removing the na values
   dplyr::rename(t = date) #renaming date to t
@@
 
-# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.
+# Exceedence function heatwaveR 
+This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.
 

### 2020-04-21 upwell_IDX.Rmd
update

@@
 ```{r}
-# Installing and loading libraries
-
-library(circular)
-## devtools::install_github("robwschlegel/coastR")
-library(gridExtra)
-library(geosphere)
-library(tidyverse)
-library(heatwaveR)
-library(coastR)
-```
-
-# Steps
-  - Use the circular function
-  - Determine daily wind data obtained from 3hr intervals
-
-```{r, eval=FALSE}
-wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
-                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
-
-wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
-                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-
-wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
-                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-
-# Selecting the important columns for each of the datasets
-## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
-wind_fix <- function(df){
-wind <- df %>%
-  mutate(date = as.Date(as.character(date)),
-         hour = as.numeric(as.character(hour)),
-         dir = as.numeric(as.character(dir)),
-         speed = as.numeric(as.character(speed)),
-         temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
-                                station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
-                                station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-                                station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
-                                station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
-                                station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
-  select(temp_sites, date, hour, dir, speed) %>% 
-  drop_na()
-}
-
-wind_fix_1 <- wind_fix(df = wind_1)
-wind_fix_2 <- wind_fix(df = wind_2)
-wind_fix_3 <- wind_fix(df = wind_3)
-
-# Combine and save
-wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
-save(wind_data, file = "Data/wind_data.RData")
-
-load("Data/wind_data.RData")
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
-
-# RWS: Please figure out why some wind direction values are returned as NA
-# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
-# RWS: Please look through these resources and make corrections accordingly:
-# https://cran.r-project.org/web/packages/rWind/rWind.pdf
-# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
-wind_daily <- wind_data %>%
-  # ungroup() %>%
-  # mutate(date = as.Date(date))%>% # Changing the date to date format
-  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-  dplyr::group_by(temp_sites, date) %>%
-  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
-                                                           template = 'geographics')),2), # Direction using the circular function
-                   mean_speed = round(mean(speed),2)) # Calculating the mean
-save(wind_daily, file = "Data/wind_daily.RData") 
+# # Installing and loading libraries
+# 
+# library(circular)
+# ## devtools::install_github("robwschlegel/coastR")
+# library(gridExtra)
+# library(geosphere)
+# library(tidyverse)
+# library(heatwaveR)
+# library(coastR)
+# ```
+# 
+# # Steps
+#   - Use the circular function
+#   - Determine daily wind data obtained from 3hr intervals
+# 
+# ```{r, eval=FALSE}
+# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
+#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
+# 
+# wind_2 <- read.delim("Documents/Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
+#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+# 
+# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
+#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+# 
+# # Selecting the important columns for each of the datasets
+# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+# wind_fix <- function(df){
+# wind <- df %>%
+#   mutate(date = as.Date(as.character(date)),
+#          hour = as.numeric(as.character(hour)),
+#          dir = as.numeric(as.character(dir)),
+#          speed = as.numeric(as.character(speed)),
+#          temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
+#                                 station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
+#                                 station_name  == "PORT NOLLOTH" ~"Port Nolloth",
+#                                 station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
+#                                 station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
+#                                 station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
+#   dplyr::select(temp_sites, date, hour, dir, speed) %>% 
+#   drop_na()
+# }
+# 
+# wind_fix_1 <- wind_fix(df = wind_1)
+# wind_fix_2 <- wind_fix(df = wind_2)
+# wind_fix_3 <- wind_fix(df = wind_3)
+# 
+# # Combine and save
+# wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
+# # save(wind_data, file = "Data/wind_data.RData")
+# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay") #Removed Hout Bay as this is not one of my sites.
+# 
+# # RWS: Please figure out why some wind direction values are returned as NA
+# # RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
+# # RWS: Please look through these resources and make corrections accordingly:
+# # https://cran.r-project.org/web/packages/rWind/rWind.pdf
+# # https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
+# 
+# # Errors in the wind data for Port Nolloth - Recieved the correct wind data with no missing hours
+# 
+# wind_daily <- wind_data %>%
+#   # ungroup() %>%
+#   # mutate(date = as.Date(date))%>% # Changing the date to date format
+#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
+#   dplyr::group_by(temp_sites, date) %>%
+#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
+#                                                            template = 'geographics')),2)) #, # Direction using the circular function
+#                   # mean_speed = round(mean(speed),2)) # Calculating the mean
+# save(wind_daily, file = "Data/wind_daily.RData") 
 ```
@@
 wind_daily <- wind_daily %>%
-  dplyr::rename(sites = temp_sites)
+  dplyr::rename(sites = temp_sites) %>% 
+  drop_na()
 load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
 ```
@@
 # Plotting function
-plot_sites <- function(site_list, buffer, dist){
+# plot_sites <- function(site_list, buffer, dist){
 
   # Find the point 200 km from the site manually to pass to ggplot
-  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
-                                              b = site_list$heading, d = dist))
-
-  # Add the new coordinates tot he site list
-  site_list <- site_list %>% 
-    mutate(lon_dest = heading2$lon,
-           lat_dest = heading2$lat)
-
-  # Visualise
-  world_map +
-    geom_segment(data = site_list, colour = "red4", 
-                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
-    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
-    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
-    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
-                             max(site_list$lon + buffer)),
-                    ylim = c(min(site_list$lat - buffer), 
-                             max(site_list$lat + buffer))) +
-    labs(x = "", y = "", colour = "Site\norder")
-}
-south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
-south_africa_away
+#   heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
+#                                               b = site_list$heading, d = dist))
+#   
+#   # Add the new coordinates tot he site list
+#   site_list <- site_list %>% 
+#     mutate(lon_dest = heading2$lon,
+#            lat_dest = heading2$lat)
+#   
+#   # Visualise
+#   world_map +
+#     geom_segment(data = site_list, colour = "red4", 
+#                  aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
+#     geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
+#     geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
+#     coord_cartesian(xlim = c(min(site_list$lon - buffer), 
+#                              max(site_list$lon + buffer)),
+#                     ylim = c(min(site_list$lat - buffer), 
+#                              max(site_list$lat + buffer))) +
+#     labs(x = "", y = "", colour = "Site\norder")
+# }
+# south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
+# south_africa_away
 ```
@@
   mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
-  dplyr::rename(t = date) #renaming date to t
+  dplyr::rename(t = date) %>% 
+  drop_na()#renaming date to t
 
@@
   nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")), # creating a column of climaatologies. Column will be named clim
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-01-01", "2016-12-31")), # creating a column of climaatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict

### 2020-04-21 upwell_IDX.Rmd
update

@@
 ```{r}
-# # Installing and loading libraries
-# 
-# library(circular)
-# ## devtools::install_github("robwschlegel/coastR")
-# library(gridExtra)
-# library(geosphere)
-# library(tidyverse)
-# library(heatwaveR)
-# library(coastR)
-# ```
-# 
-# # Steps
-#   - Use the circular function
-#   - Determine daily wind data obtained from 3hr intervals
-# 
-# ```{r, eval=FALSE}
-# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
-# 
-# wind_2 <- read.delim("Documents/Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
-#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
-# 
-# # Selecting the important columns for each of the datasets
-# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
-# wind_fix <- function(df){
-# wind <- df %>%
-#   mutate(date = as.Date(as.character(date)),
-#          hour = as.numeric(as.character(hour)),
-#          dir = as.numeric(as.character(dir)),
-#          speed = as.numeric(as.character(speed)),
-#          temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
-#                                 station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
-#                                 station_name  == "PORT NOLLOTH" ~"Port Nolloth",
-#                                 station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
-#                                 station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
-#                                 station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
-#   dplyr::select(temp_sites, date, hour, dir, speed) %>% 
-#   drop_na()
-# }
-# 
-# wind_fix_1 <- wind_fix(df = wind_1)
-# wind_fix_2 <- wind_fix(df = wind_2)
-# wind_fix_3 <- wind_fix(df = wind_3)
-# 
-# # Combine and save
-# wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
-# # save(wind_data, file = "Data/wind_data.RData")
-# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay") #Removed Hout Bay as this is not one of my sites.
-# 
-# # RWS: Please figure out why some wind direction values are returned as NA
-# # RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
-# # RWS: Please look through these resources and make corrections accordingly:
-# # https://cran.r-project.org/web/packages/rWind/rWind.pdf
-# # https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
-# 
-# # Errors in the wind data for Port Nolloth - Recieved the correct wind data with no missing hours
-# 
-# wind_daily <- wind_data %>%
-#   # ungroup() %>%
-#   # mutate(date = as.Date(date))%>% # Changing the date to date format
-#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-#   dplyr::group_by(temp_sites, date) %>%
-#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
-#                                                            template = 'geographics')),2)) #, # Direction using the circular function
-#                   # mean_speed = round(mean(speed),2)) # Calculating the mean
-# save(wind_daily, file = "Data/wind_daily.RData") 
+# Installing and loading libraries
+
+library(circular)
+## devtools::install_github("robwschlegel/coastR")
+library(gridExtra)
+library(geosphere)
+library(tidyverse)
+library(heatwaveR)
+library(coastR)
+```
+
+# Steps
+  - Use the circular function
+  - Determine daily wind data obtained from 3hr intervals
+
+```{r, eval=FALSE}
+wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
+
+wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
+
+# Selecting the important columns for each of the datasets
+## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+wind_fix <- function(df){
+wind <- df %>%
+  mutate(date = as.Date(as.character(date)),
+         hour = as.numeric(as.character(hour)),
+         dir = as.numeric(as.character(dir)),
+         speed = as.numeric(as.character(speed)),
+         temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
+                                station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
+                                station_name  == "PORT NOLLOTH" ~"Port Nolloth",
+                                station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
+                                station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
+                                station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
+  dplyr::select(temp_sites, date, hour, dir, speed) %>% 
+  drop_na()
+}
+
+wind_fix_1 <- wind_fix(df = wind_1)
+wind_fix_2 <- wind_fix(df = wind_2)
+wind_fix_3 <- wind_fix(df = wind_3)
+
+# Combine and save
+wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
+save(wind_data, file = "Data/wind_data.RData")
+load("Data/wind_data.RData")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
+
+# RWS: Please figure out why some wind direction values are returned as NA
+# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
+# RWS: Please look through these resources and make corrections accordingly:
+# https://cran.r-project.org/web/packages/rWind/rWind.pdf
+# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
+wind_daily <- wind_data %>%
+  # ungroup() %>%
+  # mutate(date = as.Date(date))%>% # Changing the date to date format
+  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
+  dplyr::group_by(temp_sites, date) %>%
+  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
+                                                           template = 'geographics')),2),  # Direction using the circular function
+                  mean_speed = round(mean(speed),2)) # Calculating the mean
+
+save(wind_daily, file = "Data/wind_daily.RData") 
 ```
@@
 # Plotting function
-# plot_sites <- function(site_list, buffer, dist){
+plot_sites <- function(site_list, buffer, dist){
 

### 2020-04-21 upwell_IDX.Rmd
edits

@@
 wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
-save(wind_data, file = "Data/wind_data.RData")
-load("Data/wind_data.RData")
+#save(wind_data, file = "Data/wind_data.RData")
+# load("Data/wind_data.RData")
 selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
@@
 # https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
+
 wind_daily <- wind_data %>% 
-  # ungroup() %>%
-  # mutate(date = as.Date(date))%>% # Changing the date to date format
   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
@@
   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
-                                                           template = 'geographics')),2),  # Direction using the circular function
-                  mean_speed = round(mean(speed),2)) # Calculating the mean
+                                                           template = 'geographics')),2),
+                                    spw_circ = mean.circular(circular(speed, units = "degrees"))) # Direction using the circular function
 
@@
 # Plotting function
-plot_sites <- function(site_list, buffer, dist){
+# plot_sites <- function(site_list, buffer, dist){
 

### 2020-04-21 SST_patterns.Rmd
edits

@@
 
-# # ### The temperature products
+# ### The temperature products
 # load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-# load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
-# load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-# load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
-# load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
+load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
+load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
+load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
+load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
 ```
@@
   nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2010-06-23", "2014-08-15")), # creating a column of climatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict
@@
   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2010-06-23", "2014-08-15"))
   return(res)
@@
 save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
+MUR_upwell_base <- upwelling_detect_event(df = Final)
+
 # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here

### 2020-04-23 upwell_IDX.Rmd
edit
update

@@
 
+library(openair)
+wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
 save(wind_daily, file = "Data/wind_daily.RData")
+
+#Convert Wind Speed and Wind Direction to u/v Wind Components 
+# Then get the daily average and then convert it back?
+# https://rdrr.io/github/retostauffer/Rfoehnix/man/ddff2uv.html
 ```

### 2020-04-23 upwell_IDX.Rmd
edits

--- 2020-04-21 upwell_IDX.Rmd
@@
 library(circular)
+library(foehnix)
 ## devtools::install_github("robwschlegel/coastR")
+## devtools::install_github("retostauffer/Rfoehnix")
 library(gridExtra)
@@
 
-save(wind_daily, file = "Data/wind_daily.RData") 
+library(openair)
+wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
+# save(wind_daily, file = "Data/wind_daily.RData")
+
+#Convert Wind Speed and Wind Direction to u/v Wind Components 
+# Then get the daily average and then convert it back?
+# https://rdrr.io/github/retostauffer/Rfoehnix/man/ddff2uv.html
+# Use with zoo
+library("zoo")
+set.seed(100)
+Sys.setenv("TZ" = "UTC")
+## Calculate dd/ff
+wind_data <- wind_data %>% 
+  rename(dd = dir,
+         ff = speed)
+
+uv <- ddff2uv(wind_data)
 ```

### 2020-04-25 SST_patterns.Rmd
update

--- 2020-04-20 SST_patterns.Rmd
@@
 # # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
+
 # load("Data/UI_angle.RData")
-# 
+load("Data_coast_angle/UI_angle.RData")
+
 upwelling <- UI_angle %>%
@@
   nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict
@@
   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2010-06-23", "2014-06-30"))
   return(res)
@@
 OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+#save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
+save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
 CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
 save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
+save(CMC_upwell_base, file = "Data_coast_angle/CMC_upwell_base.RData")
+MUR_upwell_base <- upwelling_detect_event(df = Final)
+save(MUR_upwell_base, file = "Data_coast_angle/MUR_upwell_base.RData")
+
 # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
@@
 
-save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+save(SACTN_upwell_base, file = "Data_coast_angle/SACTN_upwell_base.RData")
 ```
@@
 ```{r}
-load("Data/OISST_upwell_base.RData")
-load("Data/CMC_upwell_base.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-load("Data/G1SST_upwell_base.RData")
+# load("Data/OISST_upwell_base.RData")
+# load("Data/CMC_upwell_base.RData")
+# load("Data/SACTN_upwell_base.RData")
+# load("Data/MUR_upwell_base.RData")
+# load("Data/G1SST_upwell_base.RData")
+
+load("Data_coast_base/OISST_upwell_base.RData")
+load("Data_coast_base/CMC_upwell_base.RData")
+load("Data_coast_base/SACTN_upwell_base.RData")
+load("Data_coast_base/MUR_upwell_base.RData")
+load("Data_coast_base/G1SST_upwell_base.RData")
 
@@
                                     names_from = distance, values_from = number_of_signals)
-
-save(combined_freq_spread, )
 
@@
 load("Data/CMC.RData") # 1decimal places
-```
-
-# PLotting theme
-
-```{r}
-theme_Publication <- function(base_size=14, base_family="helvetica") {
-      library(grid)
-      library(ggthemes)
-      (theme_foundation(base_size=base_size, base_family=base_family)
-       + theme(plot.title = element_text(face = "bold",
-                                         size = rel(1.2), hjust = 0.5),
-               text = element_text(),
-               panel.background = element_rect(colour = NA),
-               plot.background = element_rect(colour = NA),
-               panel.border = element_rect(colour = NA),
-               axis.title = element_text(face = "bold",size = rel(1)),
-               axis.title.y = element_text(angle=90,vjust =2),
-               axis.title.x = element_text(vjust = -0.2),
-               axis.text = element_text(), 
-               axis.line = element_line(colour="black"),
-               axis.ticks = element_line(),
-               panel.grid.major = element_line(colour="#f0f0f0"),
-               panel.grid.minor = element_blank(),
-               legend.key = element_rect(colour = NA),
-               legend.position = "bottom",
-               legend.direction = "horizontal",
-               legend.key.size= unit(0.2, "cm"),
-               legend.margin = unit(0, "cm"),
-               #legend.title = element_text(face="italic"),
-               plot.margin=unit(c(10,5,5,5),"mm"),
-               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
-               strip.text = element_text(face="bold")
-          ))
-
-}
-
-scale_fill_Publication <- function(...){
-      library(scales)
-      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
-
-}
-
-scale_colour_Publication <- function(...){
-      library(scales)
-      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
-
-}
 ```

### 2020-04-25 upwell_IDX.Rmd
update

--- 2020-04-21 upwell_IDX.Rmd
@@
 # RWS: Please look through these resources and make corrections accordingly:
-# https://cran.r-project.org/web/packages/rWind/rWind.pdf
+# https://cran.r-project.org/web/packages/rWind/rWind.pdfsave(, file = "Data_coastal_angle/wind_daily.RData")
 # https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
+# The data provided by SAWS already reads wind from the direction which it is blwing
 
@@
                                     spw_circ = mean.circular(circular(speed, units = "degrees"))) # Direction using the circular function
+# save(wind_daily, file = "Data_coast_angle/wind_daily.RData")
 
-save(wind_daily, file = "Data/wind_daily.RData") 
+# library(openair)
+# wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
+# # save(wind_daily, file = "Data/wind_daily.RData") 
+# #Convert Wind Speed and Wind Direction to u/v Wind Components 
+# # Then get the daily average and then convert it back?
+# # https://rdrr.io/github/retostauffer/Rfoehnix/man/ddff2uv.html
+# # Use with zoo
+# 
+# library("zoo")
+# set.seed(100)
+# Sys.setenv("TZ" = "UTC")
+# ## Calculate dd/ff
+# wind_data <- wind_data %>% 
+#   rename(dd = dir,
+#          ff = speed)
+# 
+# uv <- ddff2uv(wind_data)
+# #blowing to
 ```
@@
 ```{r}
-load("Data/wind_daily.RData") 
+# load("Data/wind_daily.RData") 
+load("Data_coast_angle/wind_daily.RData") 
 wind_daily <- wind_daily %>%
-  dplyr::rename(sites = temp_sites) %>% 
+  dplyr::rename(sites = temp_sites,
+                mean_speed = spw_circ) %>% 
   drop_na()
@@
   dplyr::rename(site = sites) %>%  # Renaming sites to site
-  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
+  #left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
   drop_na() %>% # removing the na values
-  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
-  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
+  #dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
+  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # This is the rientation of the coastal angle for the west coast
   dplyr::rename(t = date) %>% 
   drop_na()#renaming date to t
+
+# save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
 
@@
   nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-01-01", "2016-12-31")), # creating a column of climaatologies. Column will be named clim
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climaatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict
@@
   # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
-  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
+  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1992-07-09", "2005-10-19"))) %>%
   select(-data) %>% 
@@
 # save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
+# save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")
 # Unpack the daily climatology results

### 2020-04-25 box_plot.R

@@ -4,13 +4,33 @@ library(lubridate)
 library(multcomp)
 options(scipen = 999)
 
-load("Data/OISST_final.RData")
-load("Data/G1SST_final.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_final.RData")
-load("Data/CMC_final.RData")
+# load("Data/OISST_final.RData")
+# load("Data/G1SST_final.RData")
+# load("Data/SACTN_upwell_base.RData")
+# load("Data/MUR_final.RData")
+# load("Data/CMC_final.RData")
+
+load("Data_coast_base/OISST_upwell_base.RData")
+load("Data_coast_base/CMC_upwell_base.RData")
+load("Data_coast_base/SACTN_upwell_base.RData")
+load("Data_coast_base/MUR_upwell_base.RData")
+load("Data_coast_base/G1SST_upwell_base.RData")
+
+# Removing the distance of 20 and 40kms
+library(dplyr)
+removing_distance_func <- function(df){
+  removing_dist_func<- df %>% 
+    filter(!distance %in% c(20000,40000))
+}
+
+OISST_final <- removing_distance_func(df = OISST_upwell_base)
+G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+MUR_final <- removing_distance_func(df = MUR_upwell_base)
+CMC_final <- removing_distance_func(df = CMC_upwell_base)
 
 combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+
+#save(combined_products, file = "Data_coast_angle/combined_products.RData")
 
 seasons_func <- function(df){
   BC_seaons <- df %>% 
@@ -35,16 +55,19 @@ metric_SACTN <- SACTN %>%
   mutate(product = "SACTN")
 
 combined <- rbind(metric_prods, metric_SACTN)
-# write.csv(combined,'combined.csv')
+write.csv(combined,'combined.csv')
+
+metrics <- combined %>% 
+  mutate(year = year(date_start)) %>% 
+  group_by(product, site, duration) %>% 
+  summarise(y = n())%>% 
+  rename(count = y)
 
 metrics_func <- function(df){
   metrics <- df %>% 
   mutate(year = year(date_start)) %>% 
-  group_by(product, site) %>% 
-  summarise(y = n(),
-            mean_intensity = mean(intensity_mean),
-            mean_dur = mean(duration),
-            mean_cumIn = mean(intensity_cumulative)) %>% 
[…]

### 2020-04-25 ANOVA.R
ANOVA

@@ -6,12 +6,30 @@ library(lubridate)
 library(multcomp)
 options(scipen = 999)
 
-load("Data/OISST_final.RData")
-OISST_final$distance_km <- OISST_final$distance/1000 # This column is missing from this dataframe
-load("Data/G1SST_final.RData")
-#load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_final.RData")
-load("Data/CMC_final.RData")
+load("Data_coast_base/OISST_upwell_base.RData")
+load("Data_coast_base/CMC_upwell_base.RData")
+load("Data_coast_base/SACTN_upwell_base.RData")
+load("Data_coast_base/MUR_upwell_base.RData")
+load("Data_coast_base/G1SST_upwell_base.RData")
+
+# Removing the distance of 20 and 40kms
+library(dplyr)
+removing_distance_func <- function(df){
+  removing_dist_func<- df %>% 
+    filter(!distance %in% c(20000,40000))
+}
+
+OISST_final <- removing_distance_func(df = OISST_upwell_base)
+G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+MUR_final <- removing_distance_func(df = MUR_upwell_base)
+CMC_final <- removing_distance_func(df = CMC_upwell_base)
+
+# load("Data/OISST_final.RData")
+# OISST_final$distance_km <- OISST_final$distance/1000 # This column is missing from this dataframe
+# load("Data/G1SST_final.RData")
+# #load("Data/SACTN_upwell_base.RData")
+# load("Data/MUR_final.RData")
+# load("Data/CMC_final.RData")
 
 combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
 

### 2020-04-27 exploring_data.Rmd
update

--- 2020-04-20 SST_patterns.Rmd
@@
 ---
-# This script is for all the various datasets
-
-# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets
-
-```{r}
-knitr::opts_chunk$set(
-  comment = "R>",
-  warning = FALSE,
-  message = FALSE 
-)
-#update.packages() 
-library(tidyverse)
-library(lubridate)
-library(ggpubr)
-library(zoo)
-library(FNN)
-library(scales)
-library(gridExtra)
-library(circular)
-library(fossil)
-library(mapproj)
-library(broom)
-# library(ncdf4) # This was used to process NetCDF files in an earlier version
-library(stringr)
-library(doMC); doMC::registerDoMC(cores = 4)
-library(fasttime)
-library(xtable)
-library(ncdf4) # library for processing netCDFs
-library(data.table)
-# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
-library(heatwaveR)
-library(grid)
-library(ggthemes)
-```
-
-# Loading in all the data created using the code bellow
-
-```{r}
-load("Data/site_list_sub.Rdata")
-load("Data/SACTN_US.RData")
-load("Data/site_pixels.RData") # 5 decimal places
-load("Data/OISST.RData") # 2 decimal places
-OISST <- BC_avhrr_only_v2_Document_Document 
-rm(BC_avhrr_only_v2_Document_Document ); gc()
-load("Data/CMC.RData") # 1decimal places
-load("Data/MUR.RData")
-
-# # ### The temperature products
-# load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-# load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
-# load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-# load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
-# load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
-```
-
-# PLotting theme
-
-```{r}
-theme_Publication <- function(base_size=14, base_family="helvetica") {
-      library(grid)
-      library(ggthemes)
-      (theme_foundation(base_size=base_size, base_family=base_family)
-       + theme(plot.title = element_text(face = "bold",
-                                         size = rel(1.2), hjust = 0.5),
-               text = element_text(),
-               panel.background = element_rect(colour = NA),
-               plot.background = element_rect(colour = NA),
-               panel.border = element_rect(colour = NA),
-               axis.title = element_text(face = "bold",size = rel(1)),
-               axis.title.y = element_text(angle=90,vjust =2),
-               axis.title.x = element_text(vjust = -0.2),
-               axis.text = element_text(), 
-               axis.line = element_line(colour="black"),
-               axis.ticks = element_line(),
-               panel.grid.major = element_line(colour="#f0f0f0"),
-               panel.grid.minor = element_blank(),
-               legend.key = element_rect(colour = NA),
-               legend.position = "bottom",
-               legend.direction = "horizontal",
-               legend.key.size= unit(0.2, "cm"),
-               legend.margin = unit(0, "cm"),
-               #legend.title = element_text(face="italic"),
-               plot.margin=unit(c(10,5,5,5),"mm"),
-               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
-               strip.text = element_text(face="bold")
-          ))
-
-}
-
-scale_fill_Publication <- function(...){
-      library(scales)
-      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
-
-}
-
-scale_colour_Publication <- function(...){
-      library(scales)
-      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
-
-}
-```
-
-```{r}
-# SACTN_US <- SACTN_US %>% 
-#   dplyr::rename(in_situ_temp = temp)
-# 
-# # Visualising the data
-# temp_plot <- function(df){
-#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
-#     geom_line(aes(group = site)) +
-#     labs(x = "", y = "Temperature (°C)") +
-#     theme(axis.text.x = element_text(angle = 45)) +
-#     theme(legend.position = "top")
-# }
-# 
-# SACTN_plot <- temp_plot(df = SACTN_US)
-# SACTN_plot
-
-# 25th quantile
-```
-
-Different distances from the coastline
-
-```{r}
-# load("Data/site_list_sub.Rdata")
-# # xtable(site_list_sub, auto = TRUE)
-# west <- site_list_sub
-# west$coast <- "west" # Chnages wc to west
-# 
-# load("Data/africa_coast.RData")
-# 
-# ## Downloading the bathy data from NOAA
-# # Download mid-res bathymetry data
-# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
-# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
-# # colnames(sa_bathy) <- c("lon", "lat", "depth")
-# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
-# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
-# 
-# # Loading in the newly downloaded bathymetry data               
-# load("Data/bathy/sa_bathy.RData")
-
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
-# shore.normal.transect <- function(site, width = 2){
-#   # Find the site on the coastline and it's nearest neighbour points
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
-#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
-#   coords3 <- coords3[2:1,1:3]
-#   # Define the shore normal transect bearing
-#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
-#   if(heading >= 360){
-#     heading <- heading-360
-#   } else {
-#     heading <- heading
-#   }
-#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
-#   return(heading2)
-# }
-# 
-# # Creating the transects
-# site_transects <- data.frame()
-# for(i in 1:length(west$site)){
-#  site <- west[i,]
-#  site_transect <- shore.normal.transect(site, 2)
-#  site_transects <- rbind(site_transects, site_transect)
-# }
-# 
-# # Manually correcting Sea Point and Kommetjie
-# site_transects$heading[4:5] <- 290 
-# # save(site_transects, file = "Data/site_transects.RData")
-# load("Data/site_transects.RData")
-# 
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
-# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
-# 
-# transect.pixel <- function(site, distances){
-#   # Extract coordinates
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   # Find lon/ lats every X metres 
-#   pixels <- data.frame()
-#   # deep <- 999
-#   # distance_multiplier <- 1
-#   # while(deep > isobath){
-#   for(i in 1:length(distances)){
-#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
-#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
-#     bathy2 <- sa_bathy[sitesIdx,]
-#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
-#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
-#                          heading = site$heading, 
-#                          distance = distances[i])
-#     pixels <- rbind(pixels, bathy3)
-#     coords <- coords2
-#   }
-#   if(nrow(pixels) < 1){
-#     pixels <- data.frame(site, depth = NA)
-#   }else{
-#     pixels <- pixels
-#   }+
-
-#   return(pixels)
-# }
-# 
-# # Pixel points
-# site_pixels <- data.frame()
-# for(i in 1:length(west$site)){
-#   site <- site_transects[i,]
-#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
-#   site_pixels <- rbind(site_pixels, site_pixel)
-# }
-# 
-# # Bounding box
-#   # Only one is made in order to know how large the the geom_point() squares should be made to match
-# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
-#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
-#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
-#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
-# 
-# # Determining the temperature at the various distances from the coast
-# 
-# # save(site_pixels, file = "Data/site_pixels.RData")
-# load("Data/site_pixels.RData")
-
-```
-
-# Extrcting the MUR data
-
-```{r}
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# 
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
-# 
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-# fNameStem <-
-#   substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-
-# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# names(MUR)<-c("lon","lat", "temp", "date")
-# MUR <- MUR %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# 
-# ######################################### SUBSET VIA REGION #################
-# 
-# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-#   
-#   
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)  
-#   
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-```
-
-# Extracting the CMC data
-
-```{r}
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
-# 
-# #          1         2         3         4         5         6
-# # 123456789012345678901234567890123456789012345678901234567890
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-# 
-# 
-# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
-# names(CMC)<-c("lon","lat", "temp", "date")
-# CMC <- CMC %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# # save(CMC, file = "Data/CMC.RData")
-```
-
-# Extracting the OISST data
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-#                    CC = c(25, 35, 340, 355), # Canary Current
-#                    CalC = c(35, 45, 225, 240), # California Current
-#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
-# 
-# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
-# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
-# 
-# #          1         2
-# # 1234567890123456789012345
-# # avhrr-only-v2.19810901.nc
-# 
-# # function to extract the dims and data from OISST netCDFs
-# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
-#   coords <- bbox[, region]
-#   nc <- nc_open(ncFile)
-#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
-#   fNameStem <-
-#     substr(ncFile, pathLen + 1, pathLen + 13)
-#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
-#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#   sst <- ncvar_get(nc,
-#                    varid = "sst",
-#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
-#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
-#     round(4)
-#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
-#                         lat = nc$dim$lat$vals[LatIdx])
-#   nc_close(nc)
-#   sst <-
-#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
-#     mutate(t = ymd(fDate)) %>%
-#     na.omit()
-#   fwrite(sst,
-#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
-#          append = TRUE, col.names = FALSE)
-#   rm(sst)
-# }
-# 
-# # the list of files
-# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-# strtDate <- str_sub(ncList[1], start = 15, end = 22)
-# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
-# 
-# # apply the function
-# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# 
-# # Loading the data
-# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
-# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
-# 
-# # Saving the data
-# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
-```
-
-# Extracting G1SST
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
-# 
-# # #          1         2         3         4
-# # # 12345678901234567890123456789012345678901
-# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# #          1         2         3         4         5         6
-# # 1234567890123456789012345678901234567890123456789012345678901
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# 
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-
-```
-
-# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
-# Is this upwelling event seen throughout the different distances from the coastline
-# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
-
-## Function
-Determining the temperatures at the various distances from the coastline
-
-```{r}
-# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# MUR <- MUR %>%
-#   select(lon, lat) %>%
-#   mutate(product = "MUR")
-# 
-# # Decided to work with OISST and CMC as both has a time series of 30years
-OISST_prod <- OISST %>%
-  dplyr::select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "OISST")
-# 
-CMC_prod <- CMC %>%
-  dplyr::select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "CMC")
-# 
-sat_data <- rbind(CMC_prod, OISST_prod) %>% 
-#   #rbind(., CMC) %>% 
-dplyr::select(product, lon, lat)
-# 
-sat_pixels <- sat_data %>%
-dplyr::select(product, lon, lat) %>%
-unique()
-# 
-# ## For testing the nest/map pipeline
-# # df <- site_pixels %>%
-# #   filter(site == "Lamberts Bay") %>%
-# #   select(-site)
-# 
-match_func <- function(df){
-  df <- df %>%
-    dplyr::rename(lon_site = lon, lat_site = lat)
-  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
-                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
-                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  res <- rbind(OISST_index, CMC_index)
-  return(res)
-}
-
-# # Find the nearest pixels to each spot along the transect
-# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
-# # points to not fall inside of the same 25 KM pixel
-pixel_match <- site_pixels %>%
-  group_by(site) %>%
-  group_modify(~match_func(.x))
-
-# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
-
-# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
-# Clean up some RAM space
-rm(OISST, CMC); gc()
-# 
-# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
-# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
-# # # Hout Bay will be ignored
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
-# 
-OISST_fill <- OISST_fill %>%
-  filter(site %in% selected_sites)
-
-CMC_fill <- CMC_fill %>%
-  filter(site %in% selected_sites)
-
-# save(OISST_fill, file = "Data/OISST_fill.RData")
-# save(CMC_fill, file = "Data/CMC_fill.RData")
-
-# # All the code below can be viewed in upwell_IDX.Rmd
-# # Next to run upwelling index
-# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
-# load("Data/UI_angle.RData")
-# 
-upwelling <- UI_angle %>%
-  dplyr::rename(temp = ui.saws) %>%
-  group_by(site) %>%
-  # mutate(min_t = min(t),
-  #        max_t = max(t)) %>%
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
-Upwelling can last for a few hours.
-Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>%
-  select(-data, -clim) %>%
-  unnest() %>%
-  filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance)
-
-# Now applying the Upwelling func
-# NB: This only pulls out the event results and not the climatology results
-# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
-# To create a list output one would use group_map(),
-# but this then loses the labels of which sites etc. the results belong to
-detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
-  return(res)
-}
-# 
-ts2clm_custom <- function(df){
-  # The climatology base period used here is up for debate...
-  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
-  return(res)
-}
-# 
-# Calculate the upwelling event metrics
-upwelling_detect_event <- function(df){
-  upwell_base <- df %>%
-    dplyr::rename(t = date) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-  }
-# 
-OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
-# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
-SACTN_US <- SACTN_US %>%
-  filter(site %in% selected_sites)
-
-load("Data/site_list_v4.2.RData")
-
-SACTN_upwell_base <- SACTN_US %>%
-  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
-    dplyr::rename(t = date) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-
-save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
-```
-
-Loading the data
-    This is the data obtained once the upwelling signals were filtered out
-
-```{r}
-load("Data/OISST_upwell_base.RData")
-load("Data/CMC_upwell_base.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-load("Data/G1SST_upwell_base.RData")
-
-# Removing the distance of 20 and 40kms
-library(dplyr)
-removing_distance_func <- function(df){
-  removing_dist_func<- df %>% 
-       filter(!distance %in% c(20000,40000))
-}
-
-OISST_final <- removing_distance_func(df = OISST_upwell_base)
-G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
-MUR_final <- removing_distance_func(df = MUR_upwell_base)
-CMC_final <- removing_distance_func(df = CMC_upwell_base)
-
-# save(OISST_final, file = "Data/OISST_final.RData")
-# save(G1SST_final, file = "Data/G1SST_final.RData")
-# save(MUR_final, file = "Data/MUR_final.RData")
-# save(CMC_final, file = "Data/CMC_final.RData")
-```
 
@@
 # Create a table showing these metrics
-
 ```
@@
 # ??countmax
-
 ```
@@
 
-# GLM - Determining if a change occur in the number of signals detected at different distances from the coastline
-# Included a table in the paper
-  Table discuss/count the number of upwelling signals detected at the different distances
-
-  - For some of the sites it is seen that fewer upwelling signals are visible at a further distance from the coastline
-  - When adding up all the signals observed at their respective distance it is seen that there are only slight differences in the amount of signals detected In some cases the signals detected at 10km from the coastline exceed that detected at 50km and visa versa
-
-## Number of events detected
-
-Using the start date and getting the month and year into a column. The month and year will be on the xaxis
-Then count the number of times per month at different years signals are detected
-
-```{r}
-load("Data/OISST_final.RData")
-load("Data/CMC_final.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_final.RData")
-load("Data/G1SST_final.RData")
-
-knitr::opts_chunk$set(
-  comment = "R>",
-  warning = FALSE,
-  message = FALSE 
-)
-#update.packages() 
-library(tidyverse)
-library(lubridate)
-library(ggpubr)
-library(zoo)
-library(FNN)
-library(scales)
-library(gridExtra)
-library(circular)
-library(fossil)
-library(mapproj)
-library(broom)
-# library(ncdf4) # This was used to process NetCDF files in an earlier version
-library(stringr)
-library(doMC); doMC::registerDoMC(cores = 4)
-library(fasttime)
-library(xtable)
-library(ncdf4) # library for processing netCDFs
-library(data.table)
-# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
-library(heatwaveR)
-```
-
-# Loading in all the data created using the code bellow
-
-```{r}
-load("Data/site_list_sub.Rdata")
-load("Data/SACTN_US.RData")
-load("Data/site_pixels.RData") # 5 decimal places
-load("Data/OISST.RData") # 2 decimal places
-OISST <- BC_avhrr_only_v2_Document_Document 
-rm(BC_avhrr_only_v2_Document_Document ); gc()
-load("Data/CMC.RData") # 1decimal places
-```
-
-```{r}
-# SACTN_US <- SACTN_US %>% 
-#   dplyr::rename(in_situ_temp = temp)
-# 
-# # Visualising the data
-# temp_plot <- function(df){
-#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
-#     geom_line(aes(group = site)) +
-#     labs(x = "", y = "Temperature (°C)") +
-#     theme(axis.text.x = element_text(angle = 45)) +
-#     theme(legend.position = "top")
-# }
-# 
-# SACTN_plot <- temp_plot(df = SACTN_US)
-# SACTN_plot
-
-# 25th quantile
-```
-
-Different distances from the coastline
-
-```{r}
-# load("Data/site_list_sub.Rdata")
-# # xtable(site_list_sub, auto = TRUE)
-# west <- site_list_sub
-# west$coast <- "west" # Chnages wc to west
-# 
-# load("Data/africa_coast.RData")
-# 
-# ## Downloading the bathy data from NOAA
-# # Download mid-res bathymetry data
-# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
-# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
-# # colnames(sa_bathy) <- c("lon", "lat", "depth")
-# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
-# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
-# 
-# # Loading in the newly downloaded bathymetry data               
-# load("Data/bathy/sa_bathy.RData")
-
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
-# shore.normal.transect <- function(site, width = 2){
-#   # Find the site on the coastline and it's nearest neighbour points
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
-#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
-#   coords3 <- coords3[2:1,1:3]
-#   # Define the shore normal transect bearing
-#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
-#   if(heading >= 360){
-#     heading <- heading-360
-#   } else {
-#     heading <- heading
-#   }
-#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
-#   return(heading2)
-# }
-# 
-# # Creating the transects
-# site_transects <- data.frame()
-# for(i in 1:length(west$site)){
-#  site <- west[i,]
-#  site_transect <- shore.normal.transect(site, 2)
-#  site_transects <- rbind(site_transects, site_transect)
-# }
-# 
-# # Manually correcting Sea Point and Kommetjie
-# site_transects$heading[4:5] <- 290 
-# # save(site_transects, file = "Data/site_transects.RData")
-# load("Data/site_transects.RData")
-# 
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
-# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
-# 
-# transect.pixel <- function(site, distances){
-#   # Extract coordinates
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   # Find lon/ lats every X metres 
-#   pixels <- data.frame()
-#   # deep <- 999
-#   # distance_multiplier <- 1
-#   # while(deep > isobath){
-#   for(i in 1:length(distances)){
-#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
-#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
-#     bathy2 <- sa_bathy[sitesIdx,]
-#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
-#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
-#                          heading = site$heading, 
-#                          distance = distances[i])
-#     pixels <- rbind(pixels, bathy3)
-#     coords <- coords2
-#   }
-#   if(nrow(pixels) < 1){
-#     pixels <- data.frame(site, depth = NA)
-#   }else{
-#     pixels <- pixels
-#   }+
-
-#   return(pixels)
-# }
-# 
-# # Pixel points
-# site_pixels <- data.frame()
-# for(i in 1:length(west$site)){
-#   site <- site_transects[i,]
-#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
-#   site_pixels <- rbind(site_pixels, site_pixel)
-# }
-# 
-# # Bounding box
-#   # Only one is made in order to know how large the the geom_point() squares should be made to match
-# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
-#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
-#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
-#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
-# 
-# # Determining the temperature at the various distances from the coast
-# 
-# # save(site_pixels, file = "Data/site_pixels.RData")
-# load("Data/site_pixels.RData")
-
-```
-
-# Extrcting the MUR data
-
-```{r}
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# 
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
-# 
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-# fNameStem <-
-#   substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-
-# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# names(MUR)<-c("lon","lat", "temp", "date")
-# MUR <- MUR %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# 
-# ######################################### SUBSET VIA REGION #################
-# 
-# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-#   
-#   
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)  
-#   
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-```
-
-# Extracting the CMC data
-
-```{r}
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
-# 
-# #          1         2         3         4         5         6
-# # 123456789012345678901234567890123456789012345678901234567890
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-# 
-# 
-# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
-# names(CMC)<-c("lon","lat", "temp", "date")
-# CMC <- CMC %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# # save(CMC, file = "Data/CMC.RData")
-```
-
-# Extracting the OISST data
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-#                    CC = c(25, 35, 340, 355), # Canary Current
-#                    CalC = c(35, 45, 225, 240), # California Current
-#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
-# 
-# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
-# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
-# 
-# #          1         2
-# # 1234567890123456789012345
-# # avhrr-only-v2.19810901.nc
-# 
-# # function to extract the dims and data from OISST netCDFs
-# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
-#   coords <- bbox[, region]
-#   nc <- nc_open(ncFile)
-#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
-#   fNameStem <-
-#     substr(ncFile, pathLen + 1, pathLen + 13)
-#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
-#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#   sst <- ncvar_get(nc,
-#                    varid = "sst",
-#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
-#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
-#     round(4)
-#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
-#                         lat = nc$dim$lat$vals[LatIdx])
-#   nc_close(nc)
-#   sst <-
-#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
-#     mutate(t = ymd(fDate)) %>%
-#     na.omit()
-#   fwrite(sst,
-#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
-#          append = TRUE, col.names = FALSE)
-#   rm(sst)
-# }
-# 
-# # the list of files
-# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-# strtDate <- str_sub(ncList[1], start = 15, end = 22)
-# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
-# 
-# # apply the function
-# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# 
-# # Loading the data
-# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
-# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
-# 
-# # Saving the data
-# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
-```
-
-# Extracting G1SST
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
-# 
-# # #          1         2         3         4
-# # # 12345678901234567890123456789012345678901
-# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# #          1         2         3         4         5         6
-# # 1234567890123456789012345678901234567890123456789012345678901
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# 
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-
-```
-
-# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
-# Is this upwelling event seen throughout the different distances from the coastline
-# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
-
-## Function
-Determining the temperatures at the various distances from the coastline
-
-```{r}
-# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# MUR <- MUR %>%
-#   select(lon, lat) %>%
-#   mutate(product = "MUR")
-# 
-# # Decided to work with OISST and CMC as both has a time series of 30years
-OISST_prod <- OISST %>%
-  select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "OISST")
-# 
-CMC_prod <- CMC %>%
-  select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "CMC")
-# 
-sat_data <- rbind(CMC_prod, OISST_prod) %>% 
-#   #rbind(., CMC) %>% 
-select(product, lon, lat)
-# 
-sat_pixels <- sat_data %>%
-select(product, lon, lat) %>%
-unique()
-# 
-# ## For testing the nest/map pipeline
-# # df <- site_pixels %>%
-# #   filter(site == "Lamberts Bay") %>%
-# #   select(-site)
-# 
-match_func <- function(df){
-  df <- df %>%
-    dplyr::rename(lon_site = lon, lat_site = lat)
-  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
-                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
-                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  res <- rbind(OISST_index, CMC_index)
-  return(res)
-}
-
-# # Find the nearest pixels to each spot along the transect
-# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
-# # points to not fall inside of the same 25 KM pixel
-pixel_match <- site_pixels %>%
-  group_by(site) %>%
-  group_modify(~match_func(.x))
-
-# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
-# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
-# 
-# Clean up some RAM space
-rm(OISST, CMC); gc()
-# 
-# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
-# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
-# # # Hout Bay will be ignored
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
-# 
-OISST_fill <- OISST_fill %>%
-  filter(site %in% selected_sites)
-
-CMC_fill <- CMC_fill %>%
-  filter(site %in% selected_sites)
-# 
-# 
-# # All the code below can be viewed in upwell_IDX.Rmd
-# # Next to run upwelling index
-# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
-# load("Data/UI_angle.RData")
-# 
-upwelling <- UI_angle %>%
-  dplyr::rename(temp = ui.saws) %>%
-  group_by(site) %>%
-  # mutate(min_t = min(t),
-  #        max_t = max(t)) %>%
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
-Upwelling can last for a few hours.
-Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>%
-  select(-data, -clim) %>%
-  unnest() %>%
-  filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance)
-
-# Now applying the Upwelling func
-# NB: This only pulls out the event results and not the climatology results
-# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
-# To create a list output one would use group_map(),
-# but this then loses the labels of which sites etc. the results belong to
-detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
-  return(res)
-}
-# 
-ts2clm_custom <- function(df){
-  # The climatology base period used here is up for debate...
-  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
-  return(res)
-}
-# 
-# Calculate the upwelling event metrics
-upwelling_detect_event <- function(df){
-  upwell_base <- df %>%
-    dplyr::rename(t = date) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-  }
-# 
-OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
-# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
-SACTN_US <- SACTN_US %>%
-  filter(site %in% selected_sites)
-
-load("Data/site_list_v4.2.RData")
-
-SACTN_upwell_base <- SACTN_US %>%
-  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
-    dplyr::rename(t = date) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-
-save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
-```
-
 Loading the data
@@
   mutate(slope = slope * 365.25*10)
-```
-
-As a common date we will use a start date from 1992-01-31 and end date 16-12-31
-
-```{r}
-# Generalosed linear models
-
-# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
-# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)
-
-# trends_duration_func <- function(df){
-#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
-#                      family = poisson(link = "log"))
-#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
-#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
-#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-# }
-# 
-# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
-# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
-# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
-# 
-# 
-# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
-# trends_intensity_func <- function(df){
-#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
-#                      family = gaussian(link = "identity"))
-#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
-#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
-#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-# }
-# 
-# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
-# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
-# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
-```
-
-# From the years 1992 - 2017
-
-```{r}
-# Which pixels showed the highest upwelling counts within the different datasets
-# Which site showed the most intense upwelling within the different datasets
-
-metric_func <- function(df){
- metrics <-df %>% 
-  filter(year(date_start) %in% 1992:2017) %>% 
-  #mutate(year = year(date_start)) %>% 
-  group_by(distance, site) %>% 
-  summarise(mean_intensity = mean(intensity_mean),
-            sum_events = sum(event_no)) %>% 
-   select(site, distance, mean_intensity, sum_events)
-}
-
-OISST_metrics <- metric_func(OISST_upwell_base)
-CMC_metrics <- metric_func(CMC_upwell_base)
-
-# No of u pwelling events found at the different sites
-SACTN_metrics <- SACTN_upwell_base %>% 
-  filter(year(date_start) %in% 1992:2017) %>% 
-  group_by(site) %>% 
-  summarise(mean_intensity = mean(intensity_mean),
-            sum_events = sum(event_no)) %>% 
-  select(site, mean_intensity, sum_events)
-
-# Create a table showing these metrics
-
-```
-
-# Wind plots
-
-```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
-library(ggradar)
-library(dplyr)
-library(scales)
-library(tibble)
-load("Data/UI_angle.RData")
-
-# Detects NA values, no NA values are present
-# Plotting wind variables
-source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function
-
-wind_daily_renamed <- UI_angle %>% 
-  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
-  dplyr::rename(spd = mean_speed) %>%
-  dplyr::rename(dir = dir_circ) %>% 
-  filter(spd > 0)
-
-p.wr2 <- plot.windrose(data = wind_daily_renamed,
-              spd = "spd",
-              dir = "dir")
-
-p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
-  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
-p.wr3
-
-################3
-
-# load("Data/UI_angle.RData")
-# 
-# wind_plot_prep <- UI_angle %>% 
-#   select(-t, -ui.saws)
-# 
-# wind_plot <- wind_plot_prep %>% 
-#   as_tibble(rownames = "site") %>% 
-#   group_by(site) %>% 
-#   mutate_at(vars(-site), rescale) %>% 
-#   tail(4) 
-# 
-# ggradar(wind_plot)
-# 
-# ??countmax
-
 ```
@@
 
-#Pairwise correlations
-Comparing the relaionship of variables within the same sample
-
-  -Correlate the 1st pixel with the second pixel and the first pixel with the last pixel
-  -Correlation between satellite data and number of signals at the different distances from the coastline
-
-```{r}
-
-# chi squared test
-
-load("Data/Combined_frequency.RData")
-
-combined_freq_spread <- pivot_wider(Combined_frequency, names_prefix = "dist_",
-                                    names_from = distance, values_from = number_of_signals)
-
-slope_calc <- function(df){
-  df %>% 
-    # mutate(row_num = 1:n()) %>% 
-    # do(mod1 = lm(number_of_signals ~ row_num, data = .),
-       # mod2 = lm(distance ~ row_num, data = .),
-       # mod3 = lm(distance ~ number_of_signals, data = .),
-       # mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
-        # mutate(distance_slope = summary(mod1)$coeff[2],
-        #    signal_slope = summary(mod2)$coeff[2],
-        #    signal_distance_slope = summary(mod3)$coeff[2],
-        #    signal_distance_r = mod4[1],
-        #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
-        # select(-mod1, -mod2, -mod3, -mod4) %>% 
-    do(mod1 = cor(.$dist_10, .$dist_30, method = "pearson", use = "complete.obs"),
-       mod2 = cor(.$dist_10, .$dist_50, method = "pearson", use = "complete.obs")) %>%
-    mutate(dist_10_vs_30_r = mod1[1],
-           dist_10_vs_50_r = mod2[1]) %>% 
-    # select(dist_10_vs_30_r, dist_10_vs_50_r) %>% 
-    select(-mod1, -mod2) %>% 
-    mutate_if(is.numeric, round, 2)
-}
-
-#### test <- cor(combined_freq_spread$`10`, combined_freq_spread$`30`)
-
-distance_corr <- combined_freq_spread %>% 
-  group_by(site, product) %>% 
-  slope_calc()
-
-```
-
-```{r}
-library(hrbrthemes)
-library(viridis)
-
-Combined_frequency <- Combined_frequency %>%
-  mutate(t = parse_date(date, format = "%b_%Y"),
-         distance = factor(distance))
-
-ggplot(Combined_frequency, aes(x=t, y=number_of_signals, size=distance, color=product)) +
-  geom_point(alpha=0.5) +
-  facet_wrap(~site) +
-  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
-
-ggplot(Combined_frequency, aes(x=date, y=number_of_signals, size=distance, color=product)) +
-  geom_point(alpha=0.5) +
-  facet_wrap(~site) +
-  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
-
-### Without time
-
-ggplot(Combined_frequency, aes(x=distance, y= number_of_signals, color=product)) +
-  geom_point(alpha=0.5) +
-  facet_wrap(~site) +
-  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
-
-```
-
-# Histogram
-
-```{r}
-
-combined_frequency_season <- Combined_frequency %>% 
-  mutate(month = month(date, abbr = T, label = T),
-         year = year(date)) %>% 
-  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
-                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
-                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
-
-CMC_dur <- read_csv("Data/CMC_dur.csv")
-
-ggplot(data = CMC_dur, aes(x = duration)) +
-  geom_histogram(aes(fill = product),  
-                       position = "dodge", alpha=0.6) +
-   labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
-  facet_wrap(~site) +
-  scale_colour_Publication()+
-  theme_Publication()
-
-ggplot(CMC_dur, aes(duration, colour = product)) +
-  geom_freqpoly() +
-  facet_wrap(~site) +
-  theme_bw()
-```
-
 ### Looking at intensity
@@
 
-Different distances from the coastline
-
-```{r}
-# load("Data/site_list_sub.Rdata")
-# # xtable(site_list_sub, auto = TRUE)
-# west <- site_list_sub
-# west$coast <- "west" # Chnages wc to west
-# 
-# load("Data/africa_coast.RData")
-# 
-# ## Downloading the bathy data from NOAA
-# # Download mid-res bathymetry data
-# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
-# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
-# # colnames(sa_bathy) <- c("lon", "lat", "depth")
-# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
-# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
-# 
-# # Loading in the newly downloaded bathymetry data               
-# load("Data/bathy/sa_bathy.RData")
-
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
-# shore.normal.transect <- function(site, width = 2){
-#   # Find the site on the coastline and it's nearest neighbour points
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
-#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
-#   coords3 <- coords3[2:1,1:3]
-#   # Define the shore normal transect bearing
-#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
-#   if(heading >= 360){
-#     heading <- heading-360
-#   } else {
-#     heading <- heading
-#   }
-#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
-#   return(heading2)
-# }
-# 
-# # Creating the transects
-# site_transects <- data.frame()
-# for(i in 1:length(west$site)){
-#  site <- west[i,]
-#  site_transect <- shore.normal.transect(site, 2)
-#  site_transects <- rbind(site_transects, site_transect)
-# }
-# 
-# # Manually correcting Sea Point and Kommetjie
-# site_transects$heading[4:5] <- 290 
-# # save(site_transects, file = "Data/site_transects.RData")
-# load("Data/site_transects.RData")
-# 
-# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
-# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
-# 
-# transect.pixel <- function(site, distances){
-#   # Extract coordinates
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
-#   # Find lon/ lats every X metres 
-#   pixels <- data.frame()
-#   # deep <- 999
-#   # distance_multiplier <- 1
-#   # while(deep > isobath){
-#   for(i in 1:length(distances)){
-#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
-#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
-#     bathy2 <- sa_bathy[sitesIdx,]
-#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
-#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
-#                          heading = site$heading, 
-#                          distance = distances[i])
-#     pixels <- rbind(pixels, bathy3)
-#     coords <- coords2
-#   }
-#   if(nrow(pixels) < 1){
-#     pixels <- data.frame(site, depth = NA)
-#   }else{
-#     pixels <- pixels
-#   }+
-
-#   return(pixels)
-# }
-# 
-# # Pixel points
-# site_pixels <- data.frame()
-# for(i in 1:length(west$site)){
-#   site <- site_transects[i,]
-#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
-#   site_pixels <- rbind(site_pixels, site_pixel)
-# }
-# 
-# # Bounding box
-#   # Only one is made in order to know how large the the geom_point() squares should be made to match
-# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
-#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
-#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
-#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
-# 
-# # Determining the temperature at the various distances from the coast
-# 
-# # save(site_pixels, file = "Data/site_pixels.RData")
-# load("Data/site_pixels.RData")
-
-```
-
-# Extrcting the MUR data
-
-```{r}
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# 
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST/MUR_extracted"
-# 
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir)
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-# fNameStem <-
-#   substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-
-# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# names(MUR)<-c("lon","lat", "temp", "date")
-# MUR <- MUR %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
-# 
-# ######################################### SUBSET VIA REGION #################
-# 
-# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
-# #          1         2         3         4
-# # 12345678901234567890123456789012345678901
-# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 38)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-#   
-#   
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)  
-#   
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-```
-
-# Extracting the CMC data
-
-```{r}
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
-# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
-# 
-# #          1         2         3         4         5         6
-# # 123456789012345678901234567890123456789012345678901234567890
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
-# 
-#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
-# 
-# 
-# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
-# names(CMC)<-c("lon","lat", "temp", "date")
-# CMC <- CMC %>% 
-#   mutate(temp = temp - 273.15)
-# 
-# # save(CMC, file = "Data/CMC.RData")
-```
-
-# Extracting the OISST data
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
-#                    CC = c(25, 35, 340, 355), # Canary Current
-#                    CalC = c(35, 45, 225, 240), # California Current
-#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
-#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
-# 
-# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
-# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
-# 
-# #          1         2
-# # 1234567890123456789012345
-# # avhrr-only-v2.19810901.nc
-# 
-# # function to extract the dims and data from OISST netCDFs
-# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
-#   coords <- bbox[, region]
-#   nc <- nc_open(ncFile)
-#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
-#   fNameStem <-
-#     substr(ncFile, pathLen + 1, pathLen + 13)
-#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
-#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#   sst <- ncvar_get(nc,
-#                    varid = "sst",
-#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
-#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
-#     round(4)
-#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
-#                         lat = nc$dim$lat$vals[LatIdx])
-#   nc_close(nc)
-#   sst <-
-#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
-#     mutate(t = ymd(fDate)) %>%
-#     na.omit()
-#   fwrite(sst,
-#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
-#          append = TRUE, col.names = FALSE)
-#   rm(sst)
-# }
-# 
-# # the list of files
-# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-# strtDate <- str_sub(ncList[1], start = 15, end = 22)
-# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
-# 
-# # apply the function
-# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
-# 
-# # Loading the data
-# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
-# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
-# 
-# # Saving the data
-# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
-```
-
-# Extracting G1SST
-
-```{r}
-# bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
-# library(stringr)
-# library(tidyverse)
-# library(reshape2)
-# library(ncdf4) # library for processing netCDFs
-# library(plyr)
-# library(lubridate)
-# library(data.table)
-# library(doMC); doMC::registerDoMC(cores = 7)
-# ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily"
-# csvDir <- "/media/amieroh/Seagate Expansion Drive/Extracted G1SST"
-# 
-# # #          1         2         3         4
-# # # 12345678901234567890123456789012345678901
-# # # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
-# #          1         2         3         4         5         6
-# # 1234567890123456789012345678901234567890123456789012345678901
-# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
-# 
-# 
-# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
-#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
-#     coords <- bbox[, region]
-#     nc <- nc_open(ncFile)
-#     pathLen <- nchar(paste0(ncDir, "/")) + 1
-#     fNameStem <-
-#       substr(basename(ncFile), 10, 58)
-#     fDate <- substr(basename(ncFile), 1, 8)
-#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
-#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
-#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
-#       round(4)
-#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
-#                           lat = nc$dim$lat$vals)
-#     nc_close(nc)
-#     sst <- as_tibble(melt(sst, value.name = "temp"))
-#     sst$t <- ymd(fDate)
-#     na.omit(sst)
-#     fwrite(sst,
-#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
-#            append = TRUE, col.names = FALSE)
-#     rm(sst)
-#   }
-# 
-# 
-#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
-#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
-#   endDate <- str_sub(ncLast, start = 1, end = 8)
-# 
-# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
-
-```
-
 # Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
@@
 # Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets
-
-## Function
-Determining the temperatures at the various distances from the coastline
-
-```{r}
-# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# MUR <- MUR %>%
-#   select(lon, lat) %>%
-#   mutate(product = "MUR")
-# 
-# # Decided to work with OISST and CMC as both has a time series of 30years
-OISST_prod <- OISST %>%
-  select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "OISST")
-# 
-CMC_prod <- CMC %>%
-  select(lon, lat) %>%
-  unique() %>%
-  mutate(product = "CMC")
-# 
-sat_data <- rbind(CMC_prod, OISST_prod) %>% 
-#   #rbind(., CMC) %>% 
-select(product, lon, lat)
-# 
-sat_pixels <- sat_data %>%
-select(product, lon, lat) %>%
-unique()
-# 
-# ## For testing the nest/map pipeline
-# # df <- site_pixels %>%
-# #   filter(site == "Lamberts Bay") %>%
-# #   select(-site)
-# 
-match_func <- function(df){
-  df <- df %>%
-    dplyr::rename(lon_site = lon, lat_site = lat)
-  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
-                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
-                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
-    cbind(., df)
-  res <- rbind(OISST_index, CMC_index)
-  return(res)
-}
-
-# # Find the nearest pixels to each spot along the transect
-# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
-# # points to not fall inside of the same 25 KM pixel
-pixel_match <- site_pixels %>%
-  group_by(site) %>%
-  group_modify(~match_func(.x))
-
-# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
-OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
-# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
-# 
-# Clean up some RAM space
-rm(OISST, CMC); gc()
-# 
-# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
-# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
-# # # Hout Bay will be ignored
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
-# 
-OISST_fill <- OISST_fill %>%
-  filter(site %in% selected_sites)
-
-CMC_fill <- CMC_fill %>%
-  filter(site %in% selected_sites)
-# 
-# 
-# # All the code below can be viewed in upwell_IDX.Rmd
-# # Next to run upwelling index
-# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
-# load("Data/UI_angle.RData")
-# 
-upwelling <- UI_angle %>%
-  dplyr::rename(temp = ui.saws) %>%
-  group_by(site) %>%
-  # mutate(min_t = min(t),
-  #        max_t = max(t)) %>%
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
-Upwelling can last for a few hours.
-Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>%
-  select(-data, -clim) %>%
-  unnest() %>%
-  filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance)
-
-# Now applying the Upwelling func
-# NB: This only pulls out the event results and not the climatology results
-# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
-# To create a list output one would use group_map(),
-# but this then loses the labels of which sites etc. the results belong to
-detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
-  return(res)
-}
-# 
-ts2clm_custom <- function(df){
-  # The climatology base period used here is up for debate...
-  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
-  return(res)
-}
-# 
-# Calculate the upwelling event metrics
-upwelling_detect_event <- function(df){
-  upwell_base <- df %>%
-    dplyr::rename(t = date) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-  }
-# 
-OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
-# 
-# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
-SACTN_US <- SACTN_US %>%
-  filter(site %in% selected_sites)
-
-load("Data/site_list_v4.2.RData")
-
-SACTN_upwell_base <- SACTN_US %>%
-  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
-    dplyr::rename(t = date) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-
-save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
-```
-
-Loading the data
-    This is the data obtained once the upwelling signals were filtered out
-
-```{r}
-load("Data/OISST_upwell_base.RData")
-load("Data/CMC_upwell_base.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-load("Data/G1SST_upwell_base.RData")
-
-# Removing the distance of 20 and 40kms
-library(dplyr)
-removing_distance_func <- function(df){
-  removing_dist_func<- df %>% 
-       filter(!distance %in% c(20000,40000))
-}
-
-OISST_final <- removing_distance_func(df = OISST_upwell_base)
-G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
-MUR_final <- removing_distance_func(df = MUR_upwell_base)
-CMC_final <- removing_distance_func(df = CMC_upwell_base)
-
-# save(OISST_final, file = "Data/OISST_final.RData")
-# save(G1SST_final, file = "Data/G1SST_final.RData")
-# save(MUR_final, file = "Data/MUR_final.RData")
-# save(CMC_final, file = "Data/CMC_final.RData")
-```
-
-To accurately compare 
-  Start date CMC start date 1991-09-02 end date 2017-03-15
-  Start date OISST start date 1981-09- 01  and end date 2018-09-28
-  Start date SACTN start date 1973 and end date 2017-12-16
 
@@
 ```
-
-As a common date we will use a start date from 1992-01-31 and end date 16-12-31
-
-```{r}
-# Generalosed linear models
-
-# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
-# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)
-
-# trends_duration_func <- function(df){
-#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
-#                      family = poisson(link = "log"))
-#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
-#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
-#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-# }
-# 
-# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
-# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
-# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
-# 
-# 
-# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
-# trends_intensity_func <- function(df){
-#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
-#                      family = gaussian(link = "identity"))
-#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
-#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
-#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
-#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
-# }
-# 
-# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
-# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
-# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
-```
-
-# From the years 1992 - 2017
-
-```{r}
-# Which pixels showed the highest upwelling counts within the different datasets
-# Which site showed the most intense upwelling within the different datasets
-
-metric_func <- function(df){
- metrics <-df %>% 
-  filter(year(date_start) %in% 2011:2014) %>% 
-  #mutate(year = year(date_start)) %>% 
-  group_by(distance, site) %>% 
-  summarise(mean_intensity = mean(intensity_mean),
-            sum_events = sum(event_no)) %>% 
-   select(site, distance, mean_intensity, sum_events)
-}
-
-OISST_metrics <- metric_func(OISST_upwell_base)
-CMC_metrics <- metric_func(CMC_upwell_base)
-
-# No of upwelling events found at the different sites
-SACTN_metrics <- SACTN_upwell_base %>% 
-  filter(year(date_start) %in% 2011:2014) %>% 
-  group_by(site) %>% 
-  summarise(mean_intensity = mean(intensity_mean),
-            sum_events = sum(event_no)) %>% 
-  select(site, mean_intensity, sum_events)
-
-# Create a table showing these metrics
-
-```

### 2020-04-27 Extracting_data.R

@@ -1,0 +1,405 @@
+# Extracting the data
+# The purpose of this script is to...
+# The steps taken are:
+# 1: Setup environment
+# 2: Find the different distances from the coastline
+# 3: Extracting the MUR data
+# 4: Extracting the CMC data
+# 5: Extracting the OISST data
+# 6: Extracting the G1SST data
+
+# 1: Setup environment ----------------------------------------------------
+library(stringr)
+library(tidyverse)
+library(reshape2)
+library(ncdf4) # library for processing netCDFs
+library(plyr)
+library(lubridate)
+library(data.table)
+library(doMC); doMC::registerDoMC(cores = 7)
+
+# 2: Coastal distances ----------------------------------------------------
+
+# load("Data/site_list_sub.Rdata")
+# # xtable(site_list_sub, auto = TRUE)
+# west <- site_list_sub
+# west$coast <- "west" # Chnages wc to west
+# 
+# load("Data/africa_coast.RData")
+# 
+# ## Downloading the bathy data from NOAA
+# # Download mid-res bathymetry data
+# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# # colnames(sa_bathy) <- c("lon", "lat", "depth")
+# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+# 
+# # Loading in the newly downloaded bathymetry data               
+# load("Data/bathy/sa_bathy.RData")
+
+# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# shore.normal.transect <- function(site, width = 2){
+#   # Find the site on the coastline and it's nearest neighbour points
+#   coords <- data.frame(lon = site$lon, lat = site$lat)
+#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
[…]

### 2020-04-27 Functions/theme.R

@@ -1,57 +1,42 @@
-library(ggplot2)
-# library(extrafont)
-# loadfonts(device = "pdf", quiet = FALSE)
+theme_Publication <- function(base_size=14, base_family="helvetica") {
+  library(grid)
+  library(ggthemes)
+  (theme_foundation(base_size=base_size, base_family=base_family)
+    + theme(plot.title = element_text(face = "bold",
+                                      size = rel(1.2), hjust = 0.5),
+            text = element_text(),
+            panel.background = element_rect(colour = NA),
+            plot.background = element_rect(colour = NA),
+            panel.border = element_rect(colour = NA),
+            axis.title = element_text(face = "bold",size = rel(1)),
+            axis.title.y = element_text(angle=90,vjust =2),
+            axis.title.x = element_text(vjust = -0.2),
+            axis.text = element_text(), 
+            axis.line = element_line(colour="black"),
+            axis.ticks = element_line(),
+            panel.grid.major = element_line(colour="#f0f0f0"),
+            panel.grid.minor = element_blank(),
+            legend.key = element_rect(colour = NA),
+            legend.position = "bottom",
+            legend.direction = "horizontal",
+            legend.key.size= unit(0.2, "cm"),
+            legend.margin = unit(0, "cm"),
+            #legend.title = element_text(face="italic"),
+            plot.margin=unit(c(10,5,5,5),"mm"),
+            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+            strip.text = element_text(face="bold")
+    ))
+  
+}
 
-# make a custom theme
-theme_set(theme_bw())
-bw_update <- theme_bw() +
-  theme(panel.border = element_rect(size = 1.0),
-        # panel.grid.major = element_line(colour = "black", size = 0.2, linetype = 2),
-        panel.grid.major = element_line(colour = NA),
-        panel.grid.minor = element_line(colour = NA),
-        axis.title = element_text(size = 12, face = "bold"),
-        axis.text = element_text(size = 12, colour = "black"),
-        plot.title = element_text(size = 12, hjust = 0),
-        legend.title = element_text(size = 10),
-        legend.text = element_text(size = 10),
-        legend.key = element_rect(size = 0.8, colour = NA),
-        strip.background = element_rect(colour = NA, fill = NA),
[…]

### 2020-04-27 SST_patterns.R

@@ -1,0 +1,263 @@
+# SST_patterns
+# The purpose of this script is to...
+# The steps taken are:
+# 1: Setup environment
+# 2: Loading all of the data 
+# 3: Determining the temperatures at the various distances from the coastline
+# 4: Loading final data products
+
+# 1: Setup environment ----------------------------------------------------
+#update.packages() 
+library(tidyverse)
+library(lubridate)
+library(ggpubr)
+library(zoo)
+library(FNN)
+library(scales)
+library(gridExtra)
+library(circular)
+library(fossil)
+library(mapproj)
+library(broom)
+# library(ncdf4) # This was used to process NetCDF files in an earlier version
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
+library(xtable)
+library(ncdf4) # library for processing netCDFs
+library(data.table)
+# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
+library(heatwaveR)
+library(grid)
+library(ggthemes)
+
+# 2: Loading data ----------------------------------------------------
+
+load("Data/site_list_sub.Rdata")
+load("Data/SACTN_US.RData")
+load("Data/site_pixels.RData") # 5 decimal places
+load("Data/OISST.RData") # 2 decimal places
+OISST <- BC_avhrr_only_v2_Document_Document 
+rm(BC_avhrr_only_v2_Document_Document ); gc()
+load("Data/CMC.RData") # 1decimal places
+load("Data/MUR.RData")
+
+# ### The temperature products
+# load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
+load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
+load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
+load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
+load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
+
+# SACTN_US <- SACTN_US %>% 
+#   dplyr::rename(in_situ_temp = temp)
+# 
+# # Visualising the data
+# temp_plot <- function(df){
+#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
+#     geom_line(aes(group = site)) +
+#     labs(x = "", y = "Temperature (°C)") +
[…]

### 2020-04-27 time_series.R

@@ -1,49 +1,0 @@
-library(tidyverse)
-library(lubridate)
-library(multcomp)
-library(heatwaveR)
-options(scipen = 999)
-
-# Loading upwelling metrics
-load("Data/OISST_final.RData")
-load("Data/G1SST_final.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_final.RData")
-load("Data/CMC_final.RData")
-SACTN_upwell_base <- SACTN_upwell_base %>% 
-  mutate(product = "SACTN")
-
-combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
-combined_products <- combined_products %>% 
-  filter(site == "Lamberts Bay") %>% 
-  #  filter(distance == "10000") %>% 
-  filter(year(date_start) == 2012)
-
-### Loading the temperature products
-load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
-load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
-load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
-
-# In Lamberts Bay in June 2012 signals detected below all at a distance of 10 km from the coastline
-# CMC - 2012-06-12 signal lasted 3 days
-#G1SST - 2012-06-10 signal lsted 36 days
-#MUR - 2012-06-10 signal lsted 36 days
-#OISST - 2012-06-10 signal lsted 18 days
-
-combined_SST <- rbind(OISST_fill,CMC_fill,Final,Final_G1SST)
-combined_SST <- combined_SST %>%
-  filter(year(date) == 2012) # In 2011 in
-
-# Plotting a time series
-ggplot(combined_SST, aes(x = date, y = temp, colour = product)) +
-  geom_line()
-
-
-ggplot(combined_products, aes(x = date_peak, y = duration)) +
-    geom_lolli(aes(colour = product)) +
-    xlab("Date") + ylab("Event duration [days]") +
-  theme_bw()
-
-

### 2020-04-27 upwell_IDX.Rmd
update

@@
 
-# RWS: Please figure out why some wind direction values are returned as NA
+wind_daily_tested <- wind_data %>% 
+  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
+  dplyr::group_by(temp_sites, date)
+
+#save(wind_daily_tested, file = "Data_coast_angle/wind_daily_tested.RData")
+
 # RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
@@
 # The data provided by SAWS already reads wind from the direction which it is blwing
+
+#AA: NA values appear when applying the circular function...
+load(Data_coast_angle/wind_daily_tested.RData)
 

### 2020-04-27 upwell_IDX.Rmd
update

@@
 
-# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
-# RWS: Please look through these resources and make corrections accordingly:
-# https://cran.r-project.org/web/packages/rWind/rWind.pdfsave(, file = "Data_coastal_angle/wind_daily.RData")
-# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
 # The data provided by SAWS already reads wind from the direction which it is blwing

### 2020-04-27 upwell_IDX.Rmd
update

@@
 
-wind_daily <- wind_data %>% 
-  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-  dplyr::group_by(temp_sites, date) %>%
+wind_daily <- wind_daily_tested%>%
   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
@@
 # wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
-# # save(wind_daily, file = "Data/wind_daily.RData") 
+
 # #Convert Wind Speed and Wind Direction to u/v Wind Components 
@@
 # uv <- ddff2uv(wind_data)
-# #blowing to
 ```

### 2020-04-28 box_plot.R
commit

@@ -118,8 +118,6 @@ total_signals <- combined_products %>%
   summarise(y = n()) %>% 
   rename(count = y)
   
-
-
 # Correlation comparing signals detected at 10km with signals detected at 30km and, signals detected at 10km and 50km
 correlation_signals <- combined_products %>%
   mutate(year = year(date_start)) %>% 

### 2020-04-29 upwell_IDX.Rmd
Worked out the upwelling index issues, but some other questions remain.

@@
 In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
-The 160 degrees is used as this refers to the angle of the coastline.
-This equation is largely dependant on wind speed and direction data in order to determining the intensity of the upwelling event.
+The 160 degrees is used as this refers to the angle of the coastline for the West Coast of South Africa (Fielding & Davis, 1989).
+This equation is dependant on wind speed and direction data in order to determining the intensity of the upwelling event.
 Wind data were obtained daily from the South African Weather Service (SAWS).
-This wind data were then matched to the date at which temperatures were collected.
+These wind data were then matched to the date at which temperatures were collected.
 With these data the upwelling index was determined.
@@
 # Installing and loading libraries
-
+## devtools::install_github("robwschlegel/coastR")
+## devtools::install_github("retostauffer/Rfoehnix")
+library(tidyverse)
 library(circular)
 library(foehnix)
-## devtools::install_github("robwschlegel/coastR")
-## devtools::install_github("retostauffer/Rfoehnix")
 library(gridExtra)
 library(geosphere)
-library(tidyverse)
 library(heatwaveR)
 library(coastR)
+library(openair)
 ```
@@
 wind_daily_tested <- wind_data %>% 
-  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
-  dplyr::group_by(temp_sites, date)
+  filter(temp_sites %in% selected_sites) #%>%  # Matching the sites to wind temp sites
+  # dplyr::group_by(temp_sites, date) 
+# save(wind_daily_tested, file = "Data_coast_angle/wind_daily_tested.RData")
 
-#save(wind_daily_tested, file = "Data_coast_angle/wind_daily_tested.RData")
-
-# The data provided by SAWS already reads wind from the direction which it is blwing
-
-#AA: NA values appear when applying the circular function...
-load(Data_coast_angle/wind_daily_tested.RData)
-
-wind_daily <- wind_daily_tested%>%
-  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
-                                                           template = 'geographics')),2),
-                                    spw_circ = mean.circular(circular(speed, units = "degrees"))) # Direction using the circular function
-# save(wind_daily, file = "Data_coast_angle/wind_daily.RData")
-
-# library(openair)
-# wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
-
-# #Convert Wind Speed and Wind Direction to u/v Wind Components 
-# # Then get the daily average and then convert it back?
-# # https://rdrr.io/github/retostauffer/Rfoehnix/man/ddff2uv.html
-# # Use with zoo
-# 
-# library("zoo")
-# set.seed(100)
-# Sys.setenv("TZ" = "UTC")
-# ## Calculate dd/ff
-# wind_data <- wind_data %>% 
-#   rename(dd = dir,
-#          ff = speed)
-# 
-# uv <- ddff2uv(wind_data)
+# The data provided by SAWS already reads wind from the direction which it is blowing
+# RWS: Great find on this timeAverage function!
+wind_daily <- timeAverage(wind_daily_tested, avg.time = "day", type = "temp_sites", vector.ws = TRUE)
+save(wind_daily, file = "Data_coast_angle/wind_daily.RData")
 ```
@@
   dplyr::rename(sites = temp_sites,
-                mean_speed = spw_circ) %>% 
+                mean_speed = speed) %>% 
   drop_na()
@@
 
+<!-- RWS: This section is no longer necessary as we decided to go with the static value of 160 -->
+
 ```{r}
@@
 # Transect function in coastR package finds the angle away or along the coastline
-south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30) %>% 
+south_africa_along_wide <- coastR::transects(site_list_sub, spread = 30, alongshore = T) 
+south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30, alongshore = F) %>% 
   mutate(heading = ifelse(heading < 0, heading+360, heading))
@@
 # Plotting function
-# plot_sites <- function(site_list, buffer, dist){
+plot_sites <- function(site_list, buffer, dist){
 
   # Find the point 200 km from the site manually to pass to ggplot
-#   heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
-#                                               b = site_list$heading, d = dist))
+  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),
+                                              b = site_list$heading, d = dist))
 #
-#   # Add the new coordinates tot he site list
-#   site_list <- site_list %>% 
-#     mutate(lon_dest = heading2$lon,
-#            lat_dest = heading2$lat)
+  # Add the new coordinates tot he site list
+  site_list <- site_list %>%
+    mutate(lon_dest = heading2$lon,
+           lat_dest = heading2$lat)
 #
-#   # Visualise
-#   world_map +
-#     geom_segment(data = site_list, colour = "red4", 
-#                  aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
-#     geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
-#     geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
-#     coord_cartesian(xlim = c(min(site_list$lon - buffer), 
-#                              max(site_list$lon + buffer)),
-#                     ylim = c(min(site_list$lat - buffer), 
-#                              max(site_list$lat + buffer))) +
-#     labs(x = "", y = "", colour = "Site\norder")
-# }
-# south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
-# south_africa_away
+  # Visualise
+  world_map +
+    geom_segment(data = site_list, colour = "red4",
+                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
+    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
+    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
+    coord_quickmap(xlim = c(min(site_list$lon - buffer),
+                             max(site_list$lon + buffer)),
+                    ylim = c(min(site_list$lat - buffer),
+                             max(site_list$lat + buffer))) +
+    labs(x = "", y = "", colour = "Site\norder")
+}
+# Perpendicular transects
+south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
+south_africa_away
+
+# SHore normal transects
+south_africa_along <- plot_sites(south_africa_along_wide, 1, 100000)
+south_africa_along
 ```
@@
 ```{r}
+# RWS: the cos() function works on radians, not degrees.
 UI_angle <- wind_daily %>% # Making reference to the wind daily data created above
@@
   #dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
-  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # This is the rientation of the coastal angle for the west coast
-  dplyr::rename(t = date) %>% 
-  drop_na()#renaming date to t
+  mutate(ui.saws = mean_speed * (cos(deg_rad(dir - 160)))) %>% # This is the orientation of the coastal angle for the west coast
+  dplyr::rename(t = date) %>% # renaming date to t
+  drop_na()
+# save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
+# save(UI_angle, file = "Data/UI_angle.RData")
 
-# save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
-
-# save(UI_angle, file = "Data/UI_angle.RData")
+# RWS: Can the following code be deleted now?
 # index <- wind_UI %>%
@@
 # Exceedence function heatwaveR
-This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.
+
+This is done to determine the consecutive number of days at or above what the UI value is meant to be  
+<!-- # Detect consecutive days in exceedance of a given threshold. -->
 
@@
   group_by(site) %>%
+  mutate(t = as.Date(t)) %>% 
   # mutate(min_t = min(t), 
   #        max_t = max(t)) %>% 
-  nest() %>% # apply the following functions to all of the variables in te dataset
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climaatologies. Column will be named clim
+  nest() %>% # apply the following functions to all of the variables in the dataset
+  # RWS: Please stop using uneven periods for your climatology periods 12 month periods should be used.
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climatologies. Column will be named clim
          # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - apllies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event.
+         # purr::map - applies a function to each element of a vector
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  # Upwelling cannot be described as an event.
 Upwelling can last for a few hours.
@@
   select(-data, -clim) %>% 
-  unnest() %>%
+  unnest(cols = exceed) %>%
   filter(row_number() %% 2 == 1) %>%
-  unnest() %>% # creates a column for each variables
+  unnest(cols = exceed) %>% # creates a column for each variables
   dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
   select(site, t, ui.saws, exceedance) # selecting only these variables
+
+# RWS: If with the above code your primary objective is to determine when ui.saws is above 1, that may be done with the following code
+SACTN_upwell <- UI_angle %>% 
+  mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
+         t = as.Date(t))
+
 # Calculate quantiles of upwelling index
@@
 # There are a few ideas on how to go about doing this
-# The first is to calculatea rolling climatology and to check the upwelling phenology agianst that
+# The first is to calculate rolling climatology and to check the upwelling phenology against that
 # in order to determine changes in upwelling values, but this won't give us the metrics we want
@@
 detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)
   return(res)
@@
   # The climatology base period used here is up for debate...
-  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
+  # The choice of the 30th percentile threshold also needs to be justified and sensitivity tested
   mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1992-07-09", "2005-10-19"))) %>%
   select(-data) %>% 
-  unnest() %>%
+  unnest(cols = clim) %>%
   left_join(SACTN_upwell, by = c("site", "t")) %>%

### 2020-04-30 upwell_IDX.Rmd
deleted code

@@
 # save(wind_daily_tested, file = "Data_coast_angle/wind_daily_tested.RData")
-
 # The data provided by SAWS already reads wind from the direction which it is blowing
@@
 
-# Determining the angle perpendicular to the coastline and plotting it
-
-<!-- RWS: This section is no longer necessary as we decided to go with the static value of 160 -->
-
-```{r}
-# Coast R- function that calculates the angle of the site perpendicular to the coastline
-# Transect function in coastR package finds the angle away or along the coastline
-south_africa_along_wide <- coastR::transects(site_list_sub, spread = 30, alongshore = T) 
-south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30, alongshore = F) %>% 
-  mutate(heading = ifelse(heading < 0, heading+360, heading))
-
-# Plotting the map
-world_map <- ggplot() + 
-  borders(fill = "grey40", colour = "black")
-
-# Plotting function
-plot_sites <- function(site_list, buffer, dist){
-
-  # Find the point 200 km from the site manually to pass to ggplot
-  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),
-                                              b = site_list$heading, d = dist))
-#
-  # Add the new coordinates tot he site list
-  site_list <- site_list %>%
-    mutate(lon_dest = heading2$lon,
-           lat_dest = heading2$lat)
-#
-  # Visualise
-  world_map +
-    geom_segment(data = site_list, colour = "red4",
-                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
-    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
-    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
-    coord_quickmap(xlim = c(min(site_list$lon - buffer),
-                             max(site_list$lon + buffer)),
-                    ylim = c(min(site_list$lat - buffer),
-                             max(site_list$lat + buffer))) +
-    labs(x = "", y = "", colour = "Site\norder")
-}
-# Perpendicular transects
-south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
-south_africa_away
-
-# SHore normal transects
-south_africa_along <- plot_sites(south_africa_along_wide, 1, 100000)
-south_africa_along
-```
-
 # Upwelling indeces using the above formula
@@
 # save(UI_angle, file = "Data/UI_angle.RData")
-
-# RWS: Can the following code be deleted now?
-# index <- wind_UI %>%
-#   group_by(site) %>%
-#   select("date", "ui.saws") %>%
-#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
-# wind_UI <- wind_daily %>% 
-#   mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
-#   drop_na
-# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
 ```

### 2020-04-30 SST_patterns.R
edits

@@ -44,10 +44,10 @@ load("Data/MUR.RData")
 
 # ### The temperature products
 # load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
-load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
-load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
-load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
+load("Data/Final.RData") # MUR
+load("Data/SACTN_US.RData") # SACTN
+load("Data/OISST_fill.RData") # OISST
+load("Data/CMC_fill.RData") # CMC temperature
 
 # SACTN_US <- SACTN_US %>% 
 #   dplyr::rename(in_situ_temp = temp)

### 2020-04-30 upwell_IDX.Rmd
update

@@
 library(circular)
-library(foehnix)
+#library(foehnix)
 library(gridExtra)
@@
 
-# Selecting the important columns for each of the datasets
-## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
-## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
 wind_fix <- function(df){
@@
 wind_daily_tested <- wind_data %>% 
-  filter(temp_sites %in% selected_sites) #%>%  # Matching the sites to wind temp sites
-  # dplyr::group_by(temp_sites, date) 
+  filter(temp_sites %in% selected_sites) 
 # save(wind_daily_tested, file = "Data_coast_angle/wind_daily_tested.RData")
-# The data provided by SAWS already reads wind from the direction which it is blowing
-# RWS: Great find on this timeAverage function!
+
 wind_daily <- timeAverage(wind_daily_tested, avg.time = "day", type = "temp_sites", vector.ws = TRUE)
@@
 ```{r}
-# load("Data/wind_daily.RData") 
 load("Data_coast_angle/wind_daily.RData") 
@@
 ```{r}
-# RWS: the cos() function works on radians, not degrees.
-UI_angle <- wind_daily %>% # Making reference to the wind daily data created above
-  dplyr::rename(site = sites) %>%  # Renaming sites to site
-  #left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
-  drop_na() %>% # removing the na values
-  #dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
+UI_angle <- wind_daily %>% 
+  dplyr::rename(site = sites) %>%  
+  drop_na() %>% 
   mutate(ui.saws = mean_speed * (cos(deg_rad(dir - 160)))) %>% # This is the orientation of the coastal angle for the west coast
@@
 # Loading the insitu temperature data along the wc
-load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
-# exceedance <-
-#   function(data,
-#            x = t,
-#            y = temp,
-#            threshold,
-#            below = FALSE,
-#            minDuration = 5,
-#            joinAcrossGaps = TRUE,
-#            maxGap = 2,
-#            maxPadLength = 3)
-# Upwelling temperature threshold 12.4 / 30th percentile
+load("Data/SACTN_US.RData")
 SACTN_upwell <- UI_angle %>% 
@@
   #        max_t = max(t)) %>% 
-  nest() %>% # apply the following functions to all of the variables in the dataset
-  # RWS: Please stop using uneven periods for your climatology periods 12 month periods should be used.
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climatologies. Column will be named clim
-         # NB: A threshold of 3 appeared to be far to strict
-         # purr::map - applies a function to each element of a vector
-         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  # Upwelling cannot be described as an event.
-Upwelling can last for a few hours.
-Given that we have daily data, upwelling events minimum duration here will be 1day
-  # Detect consecutive days in exceedance of a given threshold.
-  # mutate() %>% 
+  nest() %>% 
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1993-01-01", "2017-12-21")), 
+         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  
   select(-data, -clim) %>% 
@@
   filter(row_number() %% 2 == 1) %>%
-  unnest(cols = exceed) %>% # creates a column for each variables
-  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
-  select(site, t, ui.saws, exceedance) # selecting only these variables
+  unnest(cols = exceed) %>% 
+  dplyr::rename(ui.saws = temp) %>%
+  select(site, t, ui.saws, exceedance)
 
-# RWS: If with the above code your primary objective is to determine when ui.saws is above 1, that may be done with the following code
+# OR
 SACTN_upwell <- UI_angle %>% 
   mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
-         t = as.Date(t))
-
-# Calculate quantiles of upwelling index
+         t = as.Date(t))x
 
@@
 
-# SACTN_upwell_transform <- as.data.frame(read.zoo(transform( SACTN_upwell, Date = as.POSIXct(t) ), FUN = identity ))
-# There are a few ideas on how to go about doing this
-# The first is to calculate rolling climatology and to check the upwelling phenology against that
-# in order to determine changes in upwelling values, but this won't give us the metrics we want
-# Rather what we can do is set a static bottom threshold below which we are interested in looking for upwelling
-# We then use the upwelling index values created above as a second filter to determine the metrics of the upwelling
-# Unfortunately the way that the heatwaveR functions work well does with purrr. 
-# I've figured this out once before but can't remember where now.
-# Rather than go through the process I have rather just made a convenience function below that
-# does what I want and can simply be given to purrr::map() without any faffing about.
-
 # Detect event: 
@@
   nest() %>% 
-  # The climatology base period used here is up for debate...
-  # The choice of the 30th percentile threshold also needs to be justified and sensitivity tested
-  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1992-07-09", "2005-10-19"))) %>%
+  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1993-01-01", "2005-10-19"))) %>%
   select(-data) %>% 
@@
   group_by(site) %>% 
-  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
-  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
-  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
   nest() %>% 
   mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
-  select(-data) #%>% 
-  # unnest() %>% 
-  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
-  # filter(row_number() %% 2 == 1) %>% # Select daily values
-  # unnest()
-# Unpack the event metric reults
+  select(-data)
+
 SACTN_upwell_events <- SACTN_upwell_base %>% 
@@
 # save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")
-# Unpack the daily climatology results
+
 SACTN_upwell_clims <- SACTN_upwell_base %>% 
@@
 # save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
-# The above chunk of code appears to work as expected.
-# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn. 
-    # Min duration for an upwelling event - 1day and threshold -1
-# What also needs to be decided is if we are interested in any upwelling results throughout the year,
-# or only during the upwelling season, whenever that may be for each site.
-# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
-# is changing over time.
-# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
 ```
@@
 ```{r}
-# match_func <- function(df){
-#   match <- SACTN_MCS_clims  %>% 
-#   left_join(df, by = c("site", "t")) %>% 
-#   na.trim()
-#   return(match)
-# }
-# 
-# index <- index %>% 
-#   dplyr::rename(site = sites) %>% 
-#   dplyr::rename(t = date)
-# 
-# match <- match_func(df = index)
-# 
-# upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
-#   geom_line() +
-#   geom_point(data = filter(match, event == TRUE), colour = "navy") +
-#   geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
-#   scale_x_date(expand = c(0, 0)) +
-#   facet_wrap(~site, ncol = 1) +
-#   theme_grey() +
-#   labs(y = "Temperature (°C)", x = NULL)
-# upwell_MCS
-# Plot some of the Upwelling/MCS results
 SP_upwelling <- SACTN_upwell_clims %>% 
@@
 
-# The dots on the plot help to show the length of each wind driven upwelling event detected.
-# This supports the use of this methodology.
-# Up next is fitting linear models to the upwelling metric results to see what trends shake out.
 # Linear Models

### 2020-04-30 Extracting_data.R

@@ -13,7 +13,7 @@ library(stringr)
 library(tidyverse)
 library(reshape2)
 library(ncdf4) # library for processing netCDFs
-library(plyr)
+# library(plyr)
 library(lubridate)
 library(data.table)
 library(doMC); doMC::registerDoMC(cores = 7)

### 2020-04-30 SST_patterns.R

@@ -49,52 +49,22 @@ load("Data/SACTN_US.RData") # SACTN
 load("Data/OISST_fill.RData") # OISST
 load("Data/CMC_fill.RData") # CMC temperature
 
-# SACTN_US <- SACTN_US %>% 
-#   dplyr::rename(in_situ_temp = temp)
-# 
-# # Visualising the data
-# temp_plot <- function(df){
-#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
-#     geom_line(aes(group = site)) +
-#     labs(x = "", y = "Temperature (°C)") +
-#     theme(axis.text.x = element_text(angle = 45)) +
-#     theme(legend.position = "top")
-# }
-# 
-# SACTN_plot <- temp_plot(df = SACTN_US)
-# SACTN_plot
-
-# 3:  Temperatures at the different distances ----------------------------------------------------
-
-# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
-# MUR <- MUR %>%
-#   select(lon, lat) %>%
-#   mutate(product = "MUR")
-# 
-# # Decided to work with OISST and CMC as both has a time series of 30years
 OISST_prod <- OISST %>%
   dplyr::select(lon, lat) %>%
   unique() %>%
   mutate(product = "OISST")
-# 
+
 CMC_prod <- CMC %>%
   dplyr::select(lon, lat) %>%
   unique() %>%
   mutate(product = "CMC")
-# 
+
 sat_data <- rbind(CMC_prod, OISST_prod) %>% 
-  #   #rbind(., CMC) %>% 
   dplyr::select(product, lon, lat)
-# 
 sat_pixels <- sat_data %>%
   dplyr::select(product, lon, lat) %>%
   unique()
-# 
-# ## For testing the nest/map pipeline
-# # df <- site_pixels %>%
-# #   filter(site == "Lamberts Bay") %>%
-# #   select(-site)
-# 
+
 match_func <- function(df){
   df <- df %>%
     dplyr::rename(lon_site = lon, lat_site = lat)
@@ -108,26 +78,16 @@ match_func <- function(df){
   return(res)
 }
 
-# # Find the nearest pixels to each spot along the transect
-# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
-# # points to not fall inside of the same 25 KM pixel
 pixel_match <- site_pixels %>%
   group_by(site) %>%
   group_modify(~match_func(.x))
 
[…]

### 2020-04-30 Extracting_data.R
fixed distances from coastline

@@ -20,107 +20,106 @@ library(doMC); doMC::registerDoMC(cores = 7)
 
 # 2: Coastal distances ----------------------------------------------------
 
-# load("Data/site_list_sub.Rdata")
-# # xtable(site_list_sub, auto = TRUE)
-# west <- site_list_sub
-# west$coast <- "west" # Chnages wc to west
-# 
-# load("Data/africa_coast.RData")
-# 
-# ## Downloading the bathy data from NOAA
-# # Download mid-res bathymetry data
-# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
-# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
-# # colnames(sa_bathy) <- c("lon", "lat", "depth")
-# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
-# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
-# 
-# # Loading in the newly downloaded bathymetry data               
-# load("Data/bathy/sa_bathy.RData")
+load("Data/site_list_sub.Rdata")
+# xtable(site_list_sub, auto = TRUE)
+west <- site_list_sub
+west$coast <- "west" # Chnages wc to west
+
+load("Data/africa_coast.RData")
+
+# Downloading the bathy data from NOAA
+# Download mid-res bathymetry data
+sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+colnames(sa_bathy) <- c("lon", "lat", "depth")
+sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+
+# # Loading in the newly downloaded bathymetry data
+load("Data/bathy/sa_bathy.RData")
 
 # # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
-# shore.normal.transect <- function(site, width = 2){
-#   # Find the site on the coastline and it's nearest neighbour points
-#   coords <- data.frame(lon = site$lon, lat = site$lat)
[…]

### 2020-05-01 upwell_IDX.Rmd
update

@@
   drop_na()
-# save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
+#save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
 # save(UI_angle, file = "Data/UI_angle.RData")
@@
   mutate(t = as.Date(t)) %>% 
-  # mutate(min_t = min(t), 
-  #        max_t = max(t)) %>% 
   nest() %>% 
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1993-01-01", "2017-12-21")), 
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1973-01-01", "2016-12-31")), 
          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  
@@
   mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
-         t = as.Date(t))x
+         t = as.Date(t))
 
@@
   unnest()
-# save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
+save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")
 # save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")

### 2020-05-01 SST_patterns.R

@@ -35,19 +35,16 @@ library(ggthemes)
 
 load("Data/site_list_sub.Rdata")
 load("Data/SACTN_US.RData")
-load("Data/site_pixels.RData") # 5 decimal places
+#load("Data/site_pixels.RData") # 5 decimal places
 load("Data/OISST.RData") # 2 decimal places
 OISST <- BC_avhrr_only_v2_Document_Document 
 rm(BC_avhrr_only_v2_Document_Document ); gc()
 load("Data/CMC.RData") # 1decimal places
 load("Data/MUR.RData")
+load("Data_coast_angle/site_pixels.RData")
 
 # ### The temperature products
 # load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
-load("Data/Final.RData") # MUR
-load("Data/SACTN_US.RData") # SACTN
-load("Data/OISST_fill.RData") # OISST
-load("Data/CMC_fill.RData") # CMC temperature
 
 OISST_prod <- OISST %>%
   dplyr::select(lon, lat) %>%
@@ -94,13 +91,14 @@ OISST_fill <- OISST_fill %>%
 CMC_fill <- CMC_fill %>%
   filter(site %in% selected_sites)
 
-# save(OISST_fill, file = "Data/OISST_fill.RData")
-# save(CMC_fill, file = "Data/CMC_fill.RData")
+# save(OISST_fill, file = "Data_coast_angle/OISST_fill.RData")
+# save(CMC_fill, file = "Data_coast_angle/CMC_fill.RData")
 
 # load("Data/UI_angle.RData")
 load("Data_coast_angle/UI_angle.RData")
 
 upwelling <- UI_angle %>%
+  mutate(date = as.Date(date)) %>% 
   dplyr::rename(temp = ui.saws) %>%
   group_by(site) %>%
   nest() %>% # apply the following functions to all of the variables in te dataset

### 2020-05-01 SST_patterns.R
update

@@ -13,34 +13,23 @@ library(lubridate)
 library(ggpubr)
 library(zoo)
 library(FNN)
-library(scales)
-library(gridExtra)
 library(circular)
 library(fossil)
-library(mapproj)
 library(broom)
-# library(ncdf4) # This was used to process NetCDF files in an earlier version
-library(stringr)
 library(doMC); doMC::registerDoMC(cores = 4)
 library(fasttime)
-library(xtable)
-library(ncdf4) # library for processing netCDFs
 library(data.table)
-# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
 library(heatwaveR)
-library(grid)
-library(ggthemes)
-
 # 2: Loading data ----------------------------------------------------
 
-load("Data/site_list_sub.Rdata")
+# load("Data/site_list_sub.Rdata")
 load("Data/SACTN_US.RData")
 #load("Data/site_pixels.RData") # 5 decimal places
 load("Data/OISST.RData") # 2 decimal places
 OISST <- BC_avhrr_only_v2_Document_Document 
 rm(BC_avhrr_only_v2_Document_Document ); gc()
 load("Data/CMC.RData") # 1decimal places
-load("Data/MUR.RData")
+load("Data/_coast_angleMUR.RData")
 load("Data_coast_angle/site_pixels.RData")
 
 # ### The temperature products
@@ -56,8 +45,19 @@ CMC_prod <- CMC %>%
   unique() %>%
   mutate(product = "CMC")
 
-sat_data <- rbind(CMC_prod, OISST_prod) %>% 
+MUR_prod <- MUR_new %>%
+  dplyr::select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "MUR")
+
+G1SST_prod <- G1SST %>%
+  dplyr::select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "G1SST")
+
+sat_data <- rbind(MUR_prod, OISST_prod) %>% 
   dplyr::select(product, lon, lat)
+
 sat_pixels <- sat_data %>%
   dplyr::select(product, lon, lat) %>%
   unique()
@@ -68,10 +68,10 @@ match_func <- function(df){
   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
     cbind(., df)
[…]

### 2020-05-01 SST_patterns.R
notes

@@ -110,7 +110,7 @@ detect_event_custom <- function(df){
 }
 
 ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2003-01-01", "2015-12-30"))
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2003-01-01", "2015-12-30")) #Length of MUR time series: Chnage according to length os SST product
   return(res)
 }
 

### 2020-05-01 SST_patterns.R
removed the distance 2000/30000 km code

@@ -133,6 +133,8 @@ CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
 save(CMC_upwell_base, file = "Data_coast_angle/CMC_upwell_base.RData")
 MUR_upwell_base <- upwelling_detect_event(df = MUR_fill)
 save(MUR_upwell_base, file = "Data_coast_angle/MUR_upwell_base.RData")
+G1SST_upwell_base <- upwelling_detect_event(df = G1SST_fill)
+save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
 
 # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
 SACTN_US <- SACTN_US %>% 
@@ -151,30 +153,3 @@ SACTN_upwell_base <- SACTN_US %>%
   group_modify(~detect_event_custom(.x))
 
 save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
-
-# 4: Loading the final data  ----------------------------------------------------
-
-load("Data/OISST_upwell_base.RData")
-load("Data/CMC_upwell_base.RData")
-load("Data/SACTN_upwell_base.RData")
-load("Data/MUR_upwell_base.RData")
-load("Data/G1SST_upwell_base.RData")
-
-# Removing the distance of 20 and 40kms
-library(dplyr)
-removing_distance_func <- function(df){
-  removing_dist_func<- df %>% 
-    filter(!distance %in% c(20000,40000))
-}
-
-OISST_final <- removing_distance_func(df = OISST_upwell_base)
-G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
-MUR_final <- removing_distance_func(df = MUR_upwell_base)
-CMC_final <- removing_distance_func(df = CMC_upwell_base)
-
-# save(OISST_final, file = "Data/OISST_final.RData")
-# save(G1SST_final, file = "Data/G1SST_final.RData")
-# save(MUR_final, file = "Data/MUR_final.RData")
-# save(CMC_final, file = "Data/CMC_final.RData")
-
-

### 2020-05-01 box_plot.R
error on script

@@ -10,25 +10,13 @@ options(scipen = 999)
 # load("Data/MUR_final.RData")
 # load("Data/CMC_final.RData")
 
-load("Data_coast_base/OISST_upwell_base.RData")
-load("Data_coast_base/CMC_upwell_base.RData")
-load("Data_coast_base/SACTN_upwell_base.RData")
-load("Data_coast_base/MUR_upwell_base.RData")
-load("Data_coast_base/G1SST_upwell_base.RData")
+load("Data_coast_angle/OISST_upwell_base.RData")
+load("Data_coast_angle/CMC_upwell_base.RData")
+load("Data_coast_angle/SACTN_upwell_base.RData")
+load("Data_coast_angle/MUR_upwell_base.RData")
+load("Data_coast_angle/G1SST_upwell_base.RData")
 
-# Removing the distance of 20 and 40kms
-library(dplyr)
-removing_distance_func <- function(df){
-  removing_dist_func<- df %>% 
-    filter(!distance %in% c(20000,40000))
-}
-
-OISST_final <- removing_distance_func(df = OISST_upwell_base)
-G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
-MUR_final <- removing_distance_func(df = MUR_upwell_base)
-CMC_final <- removing_distance_func(df = CMC_upwell_base)
-
-combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base)
 
 #save(combined_products, file = "Data_coast_angle/combined_products.RData")
 
@@ -44,7 +32,6 @@ seasons_func <- function(df){
 
 combined_products <- seasons_func(df = combined_products)
 SACTN <- seasons_func(df = SACTN_upwell_base)
-
 
 metric_prods <- combined_products %>% 
   filter(year(date_start) %in% 2011:2014) %>% 

### 2020-05-03 box_plot.R
still error G1SST

@@ -42,7 +42,7 @@ metric_prods <- combined_products %>%
 
 metric_SACTN <- SACTN %>% 
   filter(year(date_start) %in% 2011:2014) %>% 
-  mutate(product = "SACTN")
+  mutate(product = "SACTN") %>% 
   filter(season == "Summer")
 
 metrics_func <- function(df){
@@ -86,8 +86,6 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
 #############################################
 #Correlations
 
-load("~/Documents/Upwelling_MCS/Data_coast_angle/combined_products.RData")
-
 #Calculating the number of signals detected at different distances from the coastline
 total_signals <- combined_products %>%
   mutate(year = year(date_start)) %>% 
@@ -101,8 +99,8 @@ correlation_signals <- combined_products %>%
   group_by(product, distance,site) %>% 
   summarise(y = n())
 
-combined_freq_spread <- pivot_wider(Combined_frequency, names_prefix = "dist_",
-                                    names_from = distance, values_from = number_of_signals)
+combined_freq_spread <- pivot_wider(correlation_signals, names_prefix = "dist_",
+                                    names_from = distance, values_from = y)
 
 slope_calc <- function(df){
   df %>% 
@@ -117,8 +115,8 @@ slope_calc <- function(df){
     #    signal_distance_r = mod4[1],
     #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
     # select(-mod1, -mod2, -mod3, -mod4) %>% 
-  do(mod1 = cor(.$dist_10, .$dist_30, method = "pearson", use = "complete.obs"),
-     mod2 = cor(.$dist_10, .$dist_50, method = "pearson", use = "complete.obs")) %>%
+  do(mod1 = cor(.$dist_0, .$dist_25000, method = "pearson", use = "complete.obs"),
+     mod2 = cor(.$dist_0, .$dist_50000, method = "pearson", use = "complete.obs")) %>%
     mutate(dist_10_vs_30_r = mod1[1],
            dist_10_vs_50_r = mod2[1]) %>% 
     # select(dist_10_vs_30_r, dist_10_vs_50_r) %>% 

### 2020-05-04 G1SST_Rob.R
commit

@@ -1,0 +1,37 @@
+# Loading libraries
+library(tidyverse)
+library(lubridate)
+library(heatwaveR)
+
+
+load("Data_coast_angle/UI_angle.RData") 
+load("G1SST.RData") 
+
+upwelling <- UI_angle %>% 
+  mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
+         t = as.Date(t))
+
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event 
+  return(res)
+}
+
+ts2clm_custom <- function(df){
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) #Length of MUR time series: Chnage according to length os SST product
+  return(res)
+}
+
+# Calculate the upwelling event metrics
+upwelling_detect_event <- function(df){
+  upwell_base <- df %>%
+    dplyr::rename(t = date) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~ts2clm_custom(.x)) %>%
+    left_join(upwelling, by = c("site", "t")) %>%
+    filter(!is.na(exceedance)) %>%
+    group_by(site, product, heading, distance, lon, lat) %>%
+    group_modify(~detect_event_custom(.x))
+}
+
+
+G1SST_upwell_base <- upwelling_detect_event(df = G1SST)

### 2020-05-04 SST_patterns.R

@@ -110,7 +110,7 @@ detect_event_custom <- function(df){
 }
 
 ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "17-12-31")) #Length of MUR time series: Chnage according to length os SST product
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) #Length of MUR time series: Chnage according to length os SST product
   return(res)
 }
 

### 2020-05-04 upwell_IDX.Rmd
update

@@
 library(openair)
+
+load("Data_coast_angle/site_list_v4.2.RData")
+load("Data_coast_angle/SACTN_daily_v4.2.RData")
+load("Data_coast_angle/data_prod_dates")
+
+site_list_sub <- site_list %>%
+  filter(coast == "wc") %>%
+  filter(length > 3650)
+
+site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to
+# 2017 -kommetjie and Yzerfontein was removed as temperatures started in 1990s. 
+# I took up till 2017 because the G1SST data starts frm 2010-present. To create a long time series to run through the heatwaveR algorithm, this was needed.
+# save(site_list_sub, file = "Data_coast_angle/site_list_sub.Rdata")
+
+SACTN_US <- SACTN_daily_v4.2 %>%
+  left_join(site_list[,c(4,13)], by = "index") %>%
+  filter(index %in% site_list_sub$index) %>%
+  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
+  dplyr::rename(insitu_temp = temp)
+```
+
+# Creating a table showing the start and end dates for each of the data products used in this project
+
+```{r}
+
 ```

### 2020-05-04 SST_patterns.R

@@ -29,7 +29,7 @@ load("Data/OISST.RData") # 2 decimal places
 OISST <- BC_avhrr_only_v2_Document_Document 
 rm(BC_avhrr_only_v2_Document_Document ); gc()
 load("Data/CMC.RData") # 1decimal places
-load("Data/_coast_angleMUR.RData")
+load("Data_coast_angle/MUR.RData")
 load("Data_coast_angle/site_pixels.RData")
 
 # ### The temperature products

### 2020-05-04 upwell_IDX.Rmd
update
edits

@@
 load("Data_coast_angle/SACTN_daily_v4.2.RData")
-load("Data_coast_angle/data_prod_dates")
+load("Data_coast_angle/data_prod_dates.Rdata")
 
@@
   dplyr::rename(insitu_temp = temp)
-```
-
-# Creating a table showing the start and end dates for each of the data products used in this project
-
-```{r}
-
 ```

### 2020-05-04 G1SST_Rob.R
Minor comments

@@ -7,6 +7,8 @@ library(heatwaveR)
 load("Data_coast_angle/UI_angle.RData") 
 load("G1SST.RData") 
 
+G1SST <- data.frame(G1SST)
+
 upwelling <- UI_angle %>% 
   mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
          t = as.Date(t))
@@ -17,7 +19,7 @@ detect_event_custom <- function(df){
 }
 
 ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) #Length of MUR time series: Chnage according to length os SST product
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) # Length of MUR time series: Change according to length of SST product
   return(res)
 }
 
@@ -33,5 +35,5 @@ upwelling_detect_event <- function(df){
     group_modify(~detect_event_custom(.x))
 }
 
-
+# RWS: Note that this still doesn't run due to mismatches in the dates of the time series
 G1SST_upwell_base <- upwelling_detect_event(df = G1SST)

### 2020-05-05 Extracting/extract_tidync.R
G1SST data progress

@@ -1,0 +1,62 @@
+#library(tidync)
+
+temp_dat <- tidync("/media/amieroh/Amieroh/Data/Datasets/G1SST_data/20100711-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc") %>% 
+  hyper_tibble() %>% 
+  select(lon, lat, analysed_sst) %>% 
+  dplyr::rename(temp = analysed_sst) %>% 
+  mutate(t ="2010-07-11")
+
+# OISST_prod <- OISST %>%
+#   dplyr::select(lon, lat) %>%
+#   unique() %>%
+#   mutate(product = "OISST")
+
+G1SST_prod <- temp_dat %>%
+  dplyr::select(lon, lat) %>%
+  unique() %>%
+  mutate(product = "G1SST")
+
+sat_data <- rbind(G1SST_prod, OISST_prod) %>% 
+  dplyr::select(product, lon, lat)
+
+sat_pixels <- sat_data %>%
+  dplyr::select(product, lon, lat) %>%
+  unique()
+
+match_func <- function(df){
+  df <- df %>%
+    dplyr::rename(lon_site = lon, lat_site = lat)
+  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  G1SST_index <- G1SST_prod[as.vector(knnx.index(as.matrix(G1SST_prod[,c("lon", "lat")]),
+                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
+    cbind(., df)
+  res <- rbind(OISST_index, G1SST_index)
+  return(res)
+}
+
+pixel_match <- site_pixels %>%
+  group_by(site) %>%
+  group_modify(~match_func(.x))
+
+
+extract <- right_join(temp_dat, filter(pixel_match, product == "G1SST"), by = c("lon", "lat"))
+
+
+G1SST_fill <- extract %>% 
+  mutate(temp = temp - 273.15) %>% 
+  mutate(date = as.Date(t)) %>% 
+  dplyr::select(-t)
+
+# G1SST_official <- Final_G1SST %>% 
+#       filter(!distance %in% c(20000,40000)) %>% 
+#   mutate(distance = case_when(distance == "10000" ~ "0",
[…]

### 2020-05-06 Extracting/extract_tidync.R
update

@@ -1,10 +1,14 @@
-#library(tidync)
+library(tidync)
+# load("Data/OISST.RData") # 2 decimal places
+# OISST <- BC_avhrr_only_v2_Document_Document 
+# rm(BC_avhrr_only_v2_Document_Document ); gc()
+# load("Data_coast_angle/site_pixels.RData")
 
-temp_dat <- tidync("/media/amieroh/Amieroh/Data/Datasets/G1SST_data/20100711-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc") %>% 
+temp_dat <- tidync("/media/amieroh/Amieroh/Data/Datasets/G1SST_data/20141226-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc") %>% 
   hyper_tibble() %>% 
   select(lon, lat, analysed_sst) %>% 
   dplyr::rename(temp = analysed_sst) %>% 
-  mutate(t ="2010-07-11")
+  mutate(t ="2014-12-26")
 
 # OISST_prod <- OISST %>%
 #   dplyr::select(lon, lat) %>%
@@ -56,6 +60,9 @@ G1SST_fill <- extract %>%
 #                               distance == "50000" ~ "50000"))
 
 G1SST_finally <- rbind(G1SST_fill,G1SST_finally)
+# G1SST_finally_sub <- rbind(G1SST_fill,G1SST_fill2)
+# G1SST_finally <- rbind(G1SST_finally_sub,G1SST_finally)
+
 save(G1SST_finally, file = "G1SST_finally.RData")
 write_csv(G1SST_finally, path = "G1SST_finally.csv")
 data3 <- arrange(G1SST_finally, date)

### 2020-05-06 G1SST_Rob.R
housekeeping

@@ -1,39 +1,0 @@
-# Loading libraries
-library(tidyverse)
-library(lubridate)
-library(heatwaveR)
-
-
-load("Data_coast_angle/UI_angle.RData") 
-load("G1SST.RData") 
-
-G1SST <- data.frame(G1SST)
-
-upwelling <- UI_angle %>% 
-  mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
-         t = as.Date(t))
-
-detect_event_custom <- function(df){
-  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event 
-  return(res)
-}
-
-ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) # Length of MUR time series: Change according to length of SST product
-  return(res)
-}
-
-# Calculate the upwelling event metrics
-upwelling_detect_event <- function(df){
-  upwell_base <- df %>%
-    dplyr::rename(t = date) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
-    left_join(upwelling, by = c("site", "t")) %>%
-    filter(!is.na(exceedance)) %>%
-    group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~detect_event_custom(.x))
-}
-
-# RWS: Note that this still doesn't run due to mismatches in the dates of the time series
-G1SST_upwell_base <- upwelling_detect_event(df = G1SST)

### 2020-05-07 SST_patterns.R
housekeeping

@@ -126,16 +126,17 @@ upwelling_detect_event <- function(df){
     group_modify(~detect_event_custom(.x))
 }
 
+G1SST_finally <- G1SST_finally %>% 
+  arrange(date)
 
+G1SST_upwell_base <- upwelling_detect_event(df = G1SST_finally)
+#save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
 OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
-#save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
-save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
+#save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
 CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
-save(CMC_upwell_base, file = "Data_coast_angle/CMC_upwell_base.RData")
+#save(CMC_upwell_base, file = "Data_coast_angle/CMC_upwell_base.RData")
 MUR_upwell_base <- upwelling_detect_event(df = MUR_fill)
-save(MUR_upwell_base, file = "Data_coast_angle/MUR_upwell_base.RData")
-G1SST_upwell_base <- upwelling_detect_event(df = G1SST_fill)
-save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
+#save(MUR_upwell_base, file = "Data_coast_angle/MUR_upwell_base.RData")
 
 # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
 SACTN_US <- SACTN_US %>% 

### 2020-05-07 box_plot.R

@@ -6,17 +6,24 @@ options(scipen = 999)
 
 # load("Data/OISST_final.RData")
 # load("Data/G1SST_final.RData")
-load("Data/SACTN_upwell_base.RData")
+# load("Data/SACTN_upwell_base.RData")
 # load("Data/MUR_final.RData")
 # load("Data/CMC_final.RData")
 
 # Loading data: Created in SST_patterns.R and upwell_IDX.Rmd
-# load("Data_coast_angle/OISST_upwell_base.RData")
-# load("Data_coast_angle/CMC_upwell_base.RData")
-# load("Data_coast_angle/SACTN_upwell_base.RData")
-# load("Data_coast_angle/MUR_upwell_base.RData")
+load("Data_coast_angle/OISST_upwell_base.RData")
+load("Data_coast_angle/CMC_upwell_base.RData")
+load("Data_coast_angle/SACTN_upwell_base.RData")
+load("Data_coast_angle/MUR_upwell_base.RData")
+load("Data_coast_angle/G1SST_upwell_base.RData")
+
 # load("Data/G1SST_upwell_base.RData")
-# 
+
+combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base)
+#save(combined_products, file = "Data_coast_angle/combined_products")
+
+G1SST_upwell_base$distance <- as.numeric(G1SST_upwell_base$distance)
+
 seasons_func <- function(df){
   BC_seaons <- df %>%
     mutate(month = month(as.Date(as.character(date_start)), abbr = T, label = T),
@@ -26,13 +33,8 @@ seasons_func <- function(df){
                                   ifelse(month %in% c("Jun", "Jul", "Aug"), "Winter",
                                          ifelse(month %in% c("Sep", "Oct", "Nov"), "Spring","Error")))))
 }
-# 
-# combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base.RData)
-# #write.csv(combined_products,file = "combined_products.csv")
-
-combined_products <- read_csv("Data_coast_angle/combined_products.csv")
-
-combined_products <- seasons_func(df = combined_products)
+ 
+combined_products <- seasons_func(combined_products)
 SACTN <- seasons_func(df = SACTN_upwell_base)
 
 metric_prods <- combined_products %>% 
@@ -42,8 +44,7 @@ metric_prods <- combined_products %>%
 
 metric_SACTN <- SACTN %>% 
[…]

### 2020-05-08 site_map.Rmd
latest version commit

@@
   #site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
-  site_pixel <- transect.pixel(site, c(10000, 30000, 50000))
+  site_pixel <- transect.pixel(site, c(0, 25000, 50000))
   site_pixels <- rbind(site_pixels, site_pixel)
@@
 load("Data/site_list_sub.Rdata")
-load("Data/site_pixels_new.RData")
+load("Data_coast_angle/site_pixels.RData")
 
 #load("Data/site_pixels.RData")
-load("MUR.RData")
+#load("MUR.RData")
 names(south_africa_coast)[1] <- "lon"

### 2020-05-08 ANOVA.R

@@ -6,23 +6,23 @@ library(lubridate)
 library(multcomp)
 options(scipen = 999)
 
-load("Data_coast_base/OISST_upwell_base.RData")
-load("Data_coast_base/CMC_upwell_base.RData")
-load("Data_coast_base/SACTN_upwell_base.RData")
-load("Data_coast_base/MUR_upwell_base.RData")
-load("Data_coast_base/G1SST_upwell_base.RData")
+load("Data_coast_angle/OISST_upwell_base.RData")
+load("Data_coast_angle/CMC_upwell_base.RData")
+load("Data_coast_angle/SACTN_upwell_base.RData")
+load("Data_coast_angle/MUR_upwell_base.RData")
+load("Data_coast_angle/G1SST_upwell_base.RData")
 
-# Removing the distance of 20 and 40kms
-library(dplyr)
-removing_distance_func <- function(df){
-  removing_dist_func<- df %>% 
-    filter(!distance %in% c(20000,40000))
-}
-
-OISST_final <- removing_distance_func(df = OISST_upwell_base)
-G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
-MUR_final <- removing_distance_func(df = MUR_upwell_base)
-CMC_final <- removing_distance_func(df = CMC_upwell_base)
+# # Removing the distance of 20 and 40kms
+# library(dplyr)
+# removing_distance_func <- function(df){
+#   removing_dist_func<- df %>% 
+#     filter(!distance %in% c(20000,40000))
+# }
+# 
+# OISST_final <- removing_distance_func(df = OISST_upwell_base)
+# G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
+# MUR_final <- removing_distance_func(df = MUR_upwell_base)
+# CMC_final <- removing_distance_func(df = CMC_upwell_base)
 
 # load("Data/OISST_final.RData")
 # OISST_final$distance_km <- OISST_final$distance/1000 # This column is missing from this dataframe
@@ -31,7 +31,7 @@ CMC_final <- removing_distance_func(df = CMC_upwell_base)
 # load("Data/MUR_final.RData")
 # load("Data/CMC_final.RData")
 
-combined_products <- rbind(OISST_final,CMC_final,MUR_final,G1SST_final)
+combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base,G1SST_upwell_base)
 
 metric_4years <- combined_products %>% 
[…]

### 2020-05-08 box_plot.R

@@ -74,7 +74,7 @@ final_combined$count <- as.numeric(final_combined$count)
 ggplot(data = final_combined, aes(x = product,y = duration)) +
   geom_boxplot(aes(fill = product)) +
   facet_wrap(~site)  +
-  labs(y = "Duration (Days)", x = "SST product")+
+  labs(y = "Duration (Days)", x = "SST products")+
   theme_bw()+
   theme(legend.position="none") +
   theme(axis.text = element_text(size = 12),
@@ -82,8 +82,7 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
         legend.text = element_text(size = 12),
         legend.title = element_text(size = 14),
         strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
-        strip.text = element_text(face="bold"))
-
+        strip.text = element_text(face="bold", size = 12))
 #############################################
 #Correlations
 

### 2020-05-09 Extracting/extract_tidync.R
fixed G1SST

@@ -63,7 +63,10 @@ G1SST_finally <- rbind(G1SST_fill,G1SST_finally)
 # G1SST_finally_sub <- rbind(G1SST_fill,G1SST_fill2)
 # G1SST_finally <- rbind(G1SST_finally_sub,G1SST_finally)
 
-save(G1SST_finally, file = "G1SST_finally.RData")
-write_csv(G1SST_finally, path = "G1SST_finally.csv")
-data3 <- arrange(G1SST_finally, date)
-write_csv(data3, path = "data3.csv")
+# save(G1SST_finally, file = "G1SST_finally.RData")
+# write_csv(G1SST_finally, path = "G1SST_finally.csv")
+# data3 <- arrange(G1SST_finally, date)
+# write_csv(data3, path = "data3.csv")
+
+save(G1SST_last, file = "G1SST_last.RData") ##Key distance of 0km
+save(G1SST_last, file = "Data_coast_angle/G1SST_last ")

### 2020-05-09 SST_patterns.R

@@ -99,6 +99,7 @@ MUR_fill <- MUR_fill %>%
 
 # load("Data/UI_angle.RData")
 load("Data_coast_angle/UI_angle.RData") # Created in script 'upwell_IDX.Rmd'
+load("Data_coast_angle/G1SST_last.RData") # Created in Extracting folder : extract_tidync.R script"
 
 upwelling <- UI_angle %>% 
   mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
@@ -126,11 +127,14 @@ upwelling_detect_event <- function(df){
     group_modify(~detect_event_custom(.x))
 }
 
-G1SST_finally <- G1SST_finally %>% 
+# G1SST_finally <- G1SST_finally %>% 
+#   arrange(date)
+
+G1SST_last <- G1SST_last %>% 
   arrange(date)
 
-G1SST_upwell_base <- upwelling_detect_event(df = G1SST_finally)
-#save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
+G1SST_upwell_base <- upwelling_detect_event(df = G1SST_last)
+# save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
 OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
 #save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
 CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)

### 2020-05-09 box_plot.R

@@ -18,11 +18,9 @@ load("Data_coast_angle/MUR_upwell_base.RData")
 load("Data_coast_angle/G1SST_upwell_base.RData")
 
 # load("Data/G1SST_upwell_base.RData")
-
+G1SST_upwell_base$distance <- as.numeric(G1SST_upwell_base$distance)
 combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base)
 #save(combined_products, file = "Data_coast_angle/combined_products")
-
-G1SST_upwell_base$distance <- as.numeric(G1SST_upwell_base$distance)
 
 seasons_func <- function(df){
   BC_seaons <- df %>%
@@ -83,6 +81,7 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
         legend.title = element_text(size = 14),
         strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
         strip.text = element_text(face="bold", size = 12))
+
 #############################################
 #Correlations
 
@@ -134,8 +133,17 @@ distance_corr <- combined_freq_spread %>%
   group_by(site, product) %>% 
   slope_calc()
 
+###########################################################################
+
+load("Data_coast_angle/OISST_fill.RData")
+load("Data_coast_angle/CMC_fill.RData")
+load("Data_coast_angle/MUR_fill.RData")
+load("Data_coast_angle/G1SST_last.RData")
+load("Data/SACTN_US.RData")
 
 
 
-
-
+SST_products <- rbind(OISST_fill,G1SST_finally,CMC_fill,MUR_fill)
+SST_anaomaly <- SST_products %>% 
+  group_by(site,product) %>% 
+  summarise(mean_temp = mean(temp))

### 2020-05-13 box_plot.R
MSc writeup

@@ -349,7 +349,7 @@ SST_anaomaly <- SST_products %>%
   group_by(site,product) %>% 
   mutate(anom = temp - mean(temp, na.rm = TRUE))
 
-SACTN_anaomaly <- SST_products %>% 
+SACTN_anaomaly <- SACTN_US %>% 
   group_by(site) %>% 
   mutate(anom = temp - mean(temp, na.rm = TRUE))
 
@@ -380,12 +380,18 @@ matched_anaomaly_sub <- matched_anaomaly %>%
   filter(year(date) %in% 2011:2014) %>% 
   dplyr::select(product,distance,site,temp)
 
+# slope_calc <- function(df){
+#   df %>%
+#     do(mod1 = cor(.$distance, .$temp, method = "pearson", use = "complete.obs")) %>%
+#     mutate(dist_temp= mod1[1]) %>%
+#     dplyr::select(-mod1) %>%
+#     mutate_if(is.numeric, round, 2)
+# }
 
 slope_calc <- function(df){
   df %>%
-    do(mod1 = cor(.$distance, .$temp, method = "pearson", use = "complete.obs")) %>%
-    mutate(dist_temp= mod1[1]) %>%
-    dplyr::select(-mod1) %>%
+    cor.test(x = .$distance, .$temp,
+             use = "complete.obs", method = "pearson") %>% 
     mutate_if(is.numeric, round, 2)
 }
 
@@ -400,16 +406,12 @@ distance_corr <- matched_anaomaly_sub %>%
 #   geom_line() +
 #   facet_wrap(~site)
 
+summary(aov(temp ~site + product + distance , data = matched_anaomaly_sub))
+# summary(aov(temp ~ anom + site + product + distance , data = matched_anaomaly))
 
-
-
-
-
-
-
-
-
-
+mode <- lm(temp ~ anom + site + product + distance, data = matched_anaomaly)
+ANOVA <- aov(model)
+TUKEY <- TukeyHSD(x=ANOVA, 'matched_anaomaly$temp', conf.level=0.95)
 
 
 

### 2020-05-14 SST_patterns.R
fixed MUR

@@ -20,6 +20,7 @@ library(doMC); doMC::registerDoMC(cores = 4)
 library(fasttime)
 library(data.table)
 library(heatwaveR)
+
 # 2: Loading data ----------------------------------------------------
 
 # load("Data/site_list_sub.Rdata")
@@ -45,7 +46,7 @@ CMC_prod <- CMC %>%
   unique() %>%
   mutate(product = "CMC")
 
-MUR_prod <- MUR_new %>%
+MUR_prod <- MUR %>%
   dplyr::select(lon, lat) %>%
   unique() %>%
   mutate(product = "MUR")
@@ -81,7 +82,7 @@ pixel_match <- site_pixels %>%
 
 
 OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-MUR_fill <- right_join(MUR_new, filter(pixel_match, product == "MUR"), by = c("lon", "lat"))
+MUR_fill <- right_join(MUR, filter(pixel_match, product == "MUR"), by = c("lon", "lat"))
 rm(OISST, MUR); gc()
 
 selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
@@ -134,7 +135,7 @@ G1SST_last <- G1SST_last %>%
   arrange(date)
 
 G1SST_upwell_base <- upwelling_detect_event(df = G1SST_last)
-# save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
+#save(G1SST_upwell_base, file = "Data_coast_angle/G1SST_upwell_base.RData")
 OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
 #save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
 CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)

### 2020-05-14 box_plot.R

@@ -1,3 +1,13 @@
+# Analyses
+# The purpose of this script is to...
+# The steps taken are:
+# 1: Setup environment
+# 2: Loading all of the data 
+# 3: Preparing data to plot the box plot
+# 4: Correlations
+# 5: Testing the diffetrence in the anaomly temperatures compared to actual tenperatures
+
+# 1: Setup environment ----------------------------------------------------
 # Loading libraries
 library(tidyverse)
 library(lubridate)
@@ -6,7 +16,7 @@ options(scipen = 999)
 # devtools::install_github("twitter/AnomalyDetection")
 library(AnomalyDetection)
 
-
+# 2: Loading data ----------------------------------------------------
 # load("Data/OISST_final.RData")
 # load("Data/G1SST_final.RData")
 # load("Data/SACTN_upwell_base.RData")
@@ -20,10 +30,10 @@ load("Data_coast_angle/SACTN_upwell_base.RData")
 load("Data_coast_angle/MUR_upwell_base.RData")
 load("Data_coast_angle/G1SST_upwell_base.RData")
 
-# load("Data/G1SST_upwell_base.RData")
-G1SST_upwell_base$distance <- as.numeric(G1SST_upwell_base$distance)
 combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base)
-#save(combined_products, file = "Data_coast_angle/combined_products.RData")
+# save(combined_products, file = "Data_coast_angle/combined_products.RData")
+
+# 2: Preparing box plot data ----------------------------------------------------
 
 seasons_func <- function(df){
   BC_seaons <- df %>%
@@ -60,17 +70,6 @@ metrics_SACTN <- metrics_func(df = metric_SACTN)
 
 final_combined <- rbind(metrics, metric_SACTN)
 final_combined$count <- as.numeric(final_combined$count)
-
-# ggplot(data = final_combined, aes(x = mean_dur, y = count)) +
-#   geom_boxplot(aes(colour = product)) +
-#   facet_wrap(~site) 
-# 
-# 
-# ggplot(data = final_combined, aes(x = mean_dur)) +
-#   geom_histogram(aes(fill = product),  
-#                  position = "dodge", alpha=0.6) +
-#   labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
-#   facet_wrap(~site) 
 
[…]

### 2020-05-14 box_plot.R
fixed cor

@@ -84,63 +84,13 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
         strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
         strip.text = element_text(face="bold", size = 12))
 
-#############################################
-#Correlations
-
-load("Data_coast_angle/combined_products.RData")
-
-#Calculating the number of signals detected at different distances from the coastline
-total_signals <- combined_products %>%
-  filter(year(date_start) %in% 2011:2014) %>% 
-  mutate(year = year(date_start)) %>% 
-  group_by(product, distance) %>% 
-  summarise(y = n()) %>% 
-  rename(count = y)
-  
-# Correlation comparing signals detected at 10km with signals detected at 30km and, signals detected at 10km and 50km
-correlation_signals <- combined_products %>%
-  filter(year(date_start) %in% 2011:2014) %>% 
-  mutate(year = year(date_start)) %>% 
-  group_by(product, distance,site) %>% 
-  summarise(y = n())
-
-combined_freq_spread <- pivot_wider(correlation_signals, names_prefix = "dist_",
-                                    names_from = distance, values_from = y)
-
-slope_calc <- function(df){
-  df %>% 
-    # mutate(row_num = 1:n()) %>% 
-    # do(mod1 = lm(number_of_signals ~ row_num, data = .),
-    # mod2 = lm(distance ~ row_num, data = .),
-    # mod3 = lm(distance ~ number_of_signals, data = .),
-    # mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
-    # mutate(distance_slope = summary(mod1)$coeff[2],
-    #    signal_slope = summary(mod2)$coeff[2],
-    #    signal_distance_slope = summary(mod3)$coeff[2],
-    #    signal_distance_r = mod4[1],
-    #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
-    # select(-mod1, -mod2, -mod3, -mod4) %>% 
-  do(mod1 = cor(.$dist_0, .$dist_25000, method = "pearson", use = "complete.obs"),
-     mod2 = cor(.$dist_0, .$dist_50000, method = "pearson", use = "complete.obs")) %>%
-    mutate(dist_10_vs_30_r = mod1[1],
-           dist_10_vs_50_r = mod2[1]) %>% 
[…]

### 2020-05-14 upwell_IDX.Rmd
housekeeping
climatology

@@
 # 2017 -kommetjie and Yzerfontein was removed as temperatures started in 1990s. 
-# I took up till 2017 because the G1SST data starts frm 2010-present. To create a long time series to run through the heatwaveR algorithm, this was needed.
+# I took up till 2017 because the SACTN data ends here.. To create a long time series to run through the heatwaveR algorithm, this was needed.
 # save(site_list_sub, file = "Data_coast_angle/site_list_sub.Rdata")
@@
   mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1993-01-01", "2005-10-19"))) %>%
-  select(-data) %>% 
+  dplyr::select(-data) %>% 
   unnest(cols = clim) %>%
@@
   nest() %>% 
-  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
-  select(-data)
+  dplyr::mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
+  dplyr::select(-data)
 

### 2020-05-14 SST_patterns.R

@@ -5,6 +5,7 @@
 # 2: Loading all of the data 
 # 3: Determining the temperatures at the various distances from the coastline
 # 4: Loading final data products
+# 5: Climatology
 
 # 1: Setup environment ----------------------------------------------------
 #update.packages() 
@@ -160,3 +161,41 @@ SACTN_upwell_base <- SACTN_US %>%
   group_modify(~detect_event_custom(.x))
 
 save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
+
+# 5: Calculating the climatology ----------------------------------------------------
+
+detect_event_custom <- function(df){
+  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$climatology 
+  return(res)
+}
+
+ts2clm_custom <- function(df){
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2011-01-02 ", "2014-06-30")) #Length of MUR time series: Chnage according to length os SST product
+  return(res)
+}
+
+G1SST_last <- G1SST_last %>% 
+  arrange(date)
+
+G1SST_upwell_clims <- upwelling_detect_event(df = G1SST_last)
+#save(G1SST_upwell_clims, file = "Data_coast_angle/G1SST_upwell_clims.RData")
+OISST_upwell_clims <- upwelling_detect_event(df = OISST_fill)
+#save(OISST_upwell_clims, file = "Data_coast_angle/OISST_upwell_clims.RData")
+CMC_upwell_clims <- upwelling_detect_event(df = CMC_fill)
+# save(CMC_upwell_clims, file = "Data_coast_angle/CMC_upwell_clims.RData")
+MUR_fill$distance <- as.numeric(MUR_fill$distance)
+MUR_upwell_clims <- upwelling_detect_event(df = MUR_fill)
+#save(MUR_upwell_clims, file = "Data_coast_angle/MUR_upwell_clims.RData")
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-05-14 SST_patterns.R
climatology

@@ -170,7 +170,7 @@ detect_event_custom <- function(df){
 }
 
 ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2011-01-02 ", "2014-06-30")) #Length of MUR time series: Chnage according to length os SST product
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) #Length of MUR time series: Chnage according to length os SST product
   return(res)
 }
 

### 2020-05-14 SST_patterns.R
correlation

@@ -178,6 +178,7 @@ G1SST_last <- G1SST_last %>%
   arrange(date)
 
 G1SST_upwell_clims <- upwelling_detect_event(df = G1SST_last)
+G1SST_upwell_clims$distance <- as.numeric(G1SST_upwell_clims$distance)
 #save(G1SST_upwell_clims, file = "Data_coast_angle/G1SST_upwell_clims.RData")
 OISST_upwell_clims <- upwelling_detect_event(df = OISST_fill)
 #save(OISST_upwell_clims, file = "Data_coast_angle/OISST_upwell_clims.RData")
@@ -185,6 +186,7 @@ CMC_upwell_clims <- upwelling_detect_event(df = CMC_fill)
 # save(CMC_upwell_clims, file = "Data_coast_angle/CMC_upwell_clims.RData")
 MUR_fill$distance <- as.numeric(MUR_fill$distance)
 MUR_upwell_clims <- upwelling_detect_event(df = MUR_fill)
+MUR_upwell_clims$distance <- as.numeric(MUR_upwell_clims$distance)
 #save(MUR_upwell_clims, file = "Data_coast_angle/MUR_upwell_clims.RData")
 
 

### 2020-05-14 box_plot.R

@@ -14,7 +14,9 @@ library(lubridate)
 library(multcomp)
 options(scipen = 999)
 # devtools::install_github("twitter/AnomalyDetection")
-library(AnomalyDetection)
+#library(AnomalyDetection)
+#devtools::install_github("easystats/correlation")
+library(correlation)
 
 # 2: Loading data ----------------------------------------------------
 # load("Data/OISST_final.RData")
@@ -91,83 +93,43 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
 # Run only for the dates where upwelling is occuring 
 
 # Loading the daily temperatures
-load("Data_coast_angle/OISST_fill.RData")
-load("Data_coast_angle/CMC_fill.RData")
-load("Data_coast_angle/MUR_fill.RData")
-load("Data_coast_angle/G1SST_last.RData")
-load("Data/SACTN_US.RData")
+load("Data_coast_angle/OISST_upwell_clims.RData")
+load("Data_coast_angle/CMC_upwell_clims.RData")
+load("Data_coast_angle/MUR_upwell_clims.RData")
+load("Data_coast_angle/G1SST_upwell_clims.RData")
 
-SST_products <- rbind(OISST_fill,G1SST_last,CMC_fill,MUR_fill)
+SST_clims <- rbind(OISST_upwell_clims,G1SST_upwell_clims,CMC_upwell_clims,MUR_upwell_clims)
 
 # Determining the anomaly teperature for SST products and SACTN
-SST_anaomaly <- SST_products %>% 
+SST_anom <- SST_clims %>% 
   group_by(site,product) %>% 
-  mutate(anom = temp - mean(temp, na.rm = TRUE))
+  mutate(anom = temp - mean(seas, na.rm = TRUE)) %>% 
+  filter(year(t) %in% 2011:2014) %>% 
+  dplyr::select(site,distance,temp,anom,product)
 
-SACTN_anaomaly <- SACTN_US %>% 
-  group_by(site) %>% 
-  mutate(anom = temp - mean(temp, na.rm = TRUE))
+SST_corr <- SST_anom %>% 
+  group_by(site,temp,anom,product) %>% 
+  correlation(partial = TRUE, multilevel = TRUE) %>% 
+  summary()
 
-SST_anaomaly <- SST_anaomaly %>% 
-  dplyr::select(temp,anom, product,site,date, distance)
-SST_anaomaly$distance <- as.numeric(SST_anaomaly$distance)
+# anom_spread<- SST_anom %>%
[…]

### 2020-05-14 box_plot.R
fixing

@@ -109,7 +109,7 @@ SST_anom <- SST_clims %>%
 
 SST_corr <- SST_anom %>% 
   group_by(site,temp,anom,product) %>% 
-  correlation(partial = TRUE, multilevel = TRUE) %>% 
+  correlation(method = pearson) %>% 
   summary()
 
 # anom_spread<- SST_anom %>%

### 2020-05-15 box_plot.R
Correlation

@@ -5,7 +5,6 @@
 # 2: Loading all of the data 
 # 3: Preparing data to plot the box plot
 # 4: Correlations
-# 5: Testing the diffetrence in the anaomly temperatures compared to actual tenperatures
 
 # 1: Setup environment ----------------------------------------------------
 # Loading libraries
@@ -107,6 +106,8 @@ SST_anom <- SST_clims %>%
   filter(year(t) %in% 2011:2014) %>% 
   dplyr::select(site,distance,temp,anom,product)
 
+# Doing the correlation using the easystat correlation package
+# https://rdrr.io/github/easystats/correlation/man/correlation.html
 SST_corr <- SST_anom %>% 
   group_by(site,temp,anom,product) %>% 
   correlation(method = pearson) %>% 

### 2020-05-31 wind_analyses.Rmd
Tips
to do
update
cleaning

@@
 # load("Data_P1/insitu_wind.RData")
-# 
-# test1 <- insitu_wind %>% 
-#   mutate(lat = case_when(site == "Port Nolloth" ~ "-29.25241",
-#                                     site == "Saldanha Bay" ~ "-33.01054",
-#                                     site  == "Yzerfontein" ~ "-33.36068",
-#                                     site  == "Kommetjie" ~ "-34.13667",
-#                                     site  == "Lamberts Bay" ~ "-32.09185",
-#                                     site  == "Sea Point" ~ "-33.91847"))
-# 
-# test <- test1 %>% 
-#   mutate(lon = case_when(site == "Port Nolloth" ~ "16.86710",
-#                                     site == "Saldanha Bay" ~ "17.95063",
-#                                     site  == "Yzerfontein" ~ "18.15731",
-#                                     site  == "Kommetjie" ~ "18.32674",
-#                                     site  == "Lamberts Bay" ~ "18.30262",
-#                                     site  == "Sea Point" ~ "18.38217"))
-# 
-# test_final <- test %>% 
-#   select(date,lat,lon,dir_circ, mean_speed) %>% 
-#   dplyr::rename(speed = mean_speed) %>% 
-#   dplyr::rename(dir = dir_circ)
-
 # Loading the wind data
@@
 ```
+
+# Data inspection
+    # Find overlapping time series for each of the data and then run analyses
+    # Length of OISST much longer than CMC etc.
+    # Allows for good comparison

### 2020-07-13 site_map.Rmd
update

@@
   #site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
-  site_pixel <- transect.pixel(site, c(0, 25000, 50000))
+  site_pixel <- transect.pixel(site, c(10000, 25000, 50000))
   site_pixels <- rbind(site_pixels, site_pixel)
@@
 
+write.csv(site_list_sub,file = "site_list_sub.csv")
 # Define plotting parameters

### 2020-07-13 ANOVA.R

@@ -55,7 +55,7 @@ lm_coeff <- function(df){
 # Calculate all of the linear models
 lm_metrics <- metric_4years %>% 
   ungroup() %>% 
-  dplyr::select(-heading, -c(lon:index_end), -distance_km) %>% 
+  dplyr::select(-heading, -c(lon:index_end)) %>% 
   pivot_longer(cols = c(duration, intensity_mean:rate_decline), 
                names_to = "var", values_to = "val") %>% 
   group_by(site, product, distance, var) %>% 

### 2020-07-19 site_map.Rmd
update

@@
 library(tidyverse)
-library(plyr)
+# library(plyr)
 library(lubridate)
-library(ggpubr)
 library(zoo)
-library(lubridate)
 library(FNN)
-library(forecast)
-library(astrochron)
-library(WaveletComp)
 library(data.table)
-library(heatwaveR)
 library(viridis)
@@
 library(mapproj)
-library(ncdf4)
 library(reshape2)
-library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
 library(stringr)
@@
 library(xtable)
-library(ggpubr)
 library(ggpubr); theme_set(theme_pubr())# library(devtools)
@@
 
-write.csv(site_list_sub,file = "site_list_sub.csv")
+# write.csv(site_list_sub,file = "site_list_sub.csv")
 # Define plotting parameters

### 2020-07-19 site_map.Rmd
update

@@
 # Map <- mapping_plot(df = site_list_sub_map, xy = "salmon")
-
 ```

### 2020-07-19 box_plot.R
update

@@ -85,11 +85,12 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
-    axis.title = element_text(size = 12, face = "bold"),
-    axis.text = element_text(size = 12, colour = "black"),
-    plot.title = element_text(size = 12, hjust = 0),
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 10),
+    strip.text = element_text(size=14),
+    axis.title = element_text(size = 18, face = "bold"),
+    axis.text = element_text(size = 18, colour = "black"),
+    plot.title = element_text(size = 18, hjust = 0),
+    legend.title = element_text(size = 18),
+    legend.text = element_text(size = 16),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
 
@@ -163,11 +164,12 @@ ggplot(data = no_upwelling_sigs, aes(x = SST_product, y = Number_of_signals, gro
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
-    axis.title = element_text(size = 12, face = "bold"),
-    axis.text = element_text(size = 12, colour = "black"),
-    plot.title = element_text(size = 12, hjust = 0),
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 10),
+    strip.text = element_text(size=12),
+    axis.title = element_text(size = 18, face = "bold"),
+    axis.text = element_text(size = 18, colour = "black"),
+    plot.title = element_text(size = 18, hjust = 0),
+    legend.title = element_text(size = 18),
+    legend.text = element_text(size = 16),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
 

### 2020-07-23 site_map.Rmd
update

@@
 
-# Determining the temperature at the various distances from the coast
+## Determining the temperature at the various distances from the coast
 ```
@@
 
-# write.csv(site_list_sub,file = "site_list_sub.csv")
 # Define plotting parameters

### 2020-07-24 site_map.Rmd
update

@@
 
-# New site pixels
+# New site pixels [AA,changesd some time recorded the land temp instad]
 # save(site_pixels, file = "Data/site_pixels_new.RData")

### 2020-07-24 site_map.Rmd
update

@@
   guides(fill = guide_colourbar()) +
+# scale_x_continuous(breaks = seq(35, 25, by = 5),
+#                      labels  = c("35°S", "30°S", "25°S")) +
+#   scale_y_continuous(breaks = seq(16, 20, by = 2),
+#                      labels = c("16°E", "18°E", "20°E"))+
   scale_x_continuous(expand = c(0, 0),
@@
         axis.ticks = element_line(colour = "black"),
-        axis.ticks.length = unit(0.15, "cm"),
+        axis.ticks.length = unit(0.4, "cm"),
+        #axis.ticks.length = unit(0.15, "cm"),
         axis.ticks.margin = unit(0.1, "cm"),

### 2020-07-24 box_plot.R

@@ -87,6 +87,7 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
     panel.grid.minor = element_line(colour = NA),
     strip.text = element_text(size=14),
     axis.title = element_text(size = 18, face = "bold"),
+    axis.ticks.length = unit(0.4, "cm"),
     axis.text = element_text(size = 18, colour = "black"),
     plot.title = element_text(size = 18, hjust = 0),
     legend.title = element_text(size = 18),
@@ -165,6 +166,7 @@ ggplot(data = no_upwelling_sigs, aes(x = SST_product, y = Number_of_signals, gro
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
     strip.text = element_text(size=12),
+    axis.ticks.length = unit(0.4, "cm"),
     axis.title = element_text(size = 18, face = "bold"),
     axis.text = element_text(size = 18, colour = "black"),
     plot.title = element_text(size = 18, hjust = 0),

### 2020-07-25 site_map.Rmd
update

--- 2020-07-24 site_map.Rmd
@@
                 expand = TRUE) +
-  theme(axis.text = element_text(size = rel(1.2), colour = "black"),
-        axis.text.x = element_text(vjust = 1),
-        axis.text.y = element_text(hjust = 1),
-        axis.title.x = element_text(vjust = 0),
-        axis.title.y = element_text(angle = 90, vjust = 0.3),
-        axis.ticks = element_line(colour = "black"),
-        axis.ticks.length = unit(0.15, "cm"),
-        axis.ticks.margin = unit(0.1, "cm"),
-        panel.background = element_rect(fill = "white", colour = "black", size = 0.5),
-        #panel.border = element_blank(),
-        panel.border = element_rect(colour = "black", fill=NA, size=2),
-        #(colour =" black"),
-        panel.grid.major.x = element_blank(),
-        #panel.grid.major.y = element_line(linetype = "dotted", size = 0.2, colour = "black"),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor = element_blank(),
-        #panel.margin = unit(0.25, "lines"),
-        strip.text = element_text(size = rel(1.0)),
-        strip.background = element_rect(fill = "white", colour = NA),
-        strip.text.x = element_text(),
-        strip.text.y = element_text(angle = -90),        
-        plot.background = element_rect(colour = "white"),
-        plot.title = element_text(size = rel(1.2), vjust = 1),
-        plot.margin = unit(c(1, 1, 0.5, 0.5), "lines"))
+theme_minimal() +
+    theme(
+      axis.text = element_text(size = 10, family = "Palatino"),
+      axis.title = element_text(size = 10, family = "Palatino"),
+      legend.background = element_rect(colour = "black", size = 0.2),
+      legend.position = c(0.093, 0.70),
+      legend.text = element_text(family = "Palatino"),
+      legend.title = element_text(family = "Palatino"),
+      plot.caption = element_text(family = "Palatino"),
+      plot.title = element_text(family = "Palatino"),
+      plot.subtitle = element_text(family = "Palatino"),
+      panel.background = element_rect(colour = NA, fill = "ivory"),
+      panel.grid.major = element_line("grey70", linetype = "dashed", size = 0.2),
+      panel.grid.minor = element_line("grey70", linetype = "dashed", size = 0.2)
+    )
 

### 2020-07-26 site_map.Rmd
updated

@@
                      size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2,
-                     segment.alpha = 0.4, force = 0.1) +
+                     segment.alpha = 0.4, force = 0.1, font = "Palatino") +
   xlab("") + ylab("") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 16, y = -31.5,
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
              size =5, angle = 0, colour = "black") +
@@
                        labels = scales::unit_format(unit = "°S", sep = "")) +
-  coord_fixed(ratio = 1, xlim = c(15.5, 20), ylim = c(-35,-28),
+  coord_fixed(ratio = 1, xlim = c(15.7, 20), ylim = c(-34.7,-28),
                 expand = TRUE) +
-theme_minimal() +
-    theme(
-      axis.text = element_text(size = 10, family = "Palatino"),
-      axis.title = element_text(size = 10, family = "Palatino"),
-      legend.background = element_rect(colour = "black", size = 0.2),
-      legend.position = c(0.093, 0.70),
-      legend.text = element_text(family = "Palatino"),
-      legend.title = element_text(family = "Palatino"),
-      plot.caption = element_text(family = "Palatino"),
-      plot.title = element_text(family = "Palatino"),
-      plot.subtitle = element_text(family = "Palatino"),
-      panel.background = element_rect(colour = NA, fill = "ivory"),
-      panel.grid.major = element_line("grey70", linetype = "dashed", size = 0.2),
-      panel.grid.minor = element_line("grey70", linetype = "dashed", size = 0.2)
-    )
+  theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Palatino"),
+        axis.text.x = element_text(vjust = 1, family = "Palatino"),
+        axis.text.y = element_text(hjust = 1, family = "Palatino"),
+        axis.title.x = element_text(vjust = 0, family = "Palatino"),
+        axis.title.y = element_text(angle = 90, vjust = 0.3, family = "Palatino"),
+        axis.ticks = element_line(colour = "black"),
+        axis.ticks.length = unit(0.4, "cm"),
+        #axis.ticks.length = unit(0.15, "cm"),
+        axis.ticks.margin = unit(0.1, "cm"),
+        panel.background = element_rect(fill = "white", colour = "black", size = 0.5),
+        #panel.border = element_blank(),
+        panel.border = element_rect(colour = "black", fill=NA, size=2),
+        #(colour =" black"),
+        panel.grid.major.x = element_blank(),
+        #panel.grid.major.y = element_line(linetype = "dotted", size = 0.2, colour = "black"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor = element_blank(),
+        #panel.margin = unit(0.25, "lines"),
+        strip.text = element_text(size = rel(1.0), family = "Palatino"),
+        strip.background = element_rect(fill = "white", colour = NA),
+        strip.text.x = element_text(),
+        strip.text.y = element_text(angle = -90),        
+        plot.background = element_rect(colour = "white"),
+        plot.title = element_text(size = rel(1.2), vjust = 1),
+        plot.margin = unit(c(1, 1, 0.5, 0.5), "lines"))
 

### 2020-07-26 box_plot.R

@@ -85,10 +85,10 @@ ggplot(data = final_combined, aes(x = product,y = duration)) +
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
-    strip.text = element_text(size=14),
-    axis.title = element_text(size = 18, face = "bold"),
+    strip.text = element_text(size=14, family = "Palatino"),
+    axis.title = element_text(size = 18, face = "bold", family = "Palatino"),
     axis.ticks.length = unit(0.4, "cm"),
-    axis.text = element_text(size = 18, colour = "black"),
+    axis.text = element_text(size = 18, colour = "black", family = "Palatino"),
     plot.title = element_text(size = 18, hjust = 0),
     legend.title = element_text(size = 18),
     legend.text = element_text(size = 16),
@@ -165,13 +165,14 @@ ggplot(data = no_upwelling_sigs, aes(x = SST_product, y = Number_of_signals, gro
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
-    strip.text = element_text(size=12),
+    strip.text = element_text(size=12, family = "Palatino"),
     axis.ticks.length = unit(0.4, "cm"),
-    axis.title = element_text(size = 18, face = "bold"),
-    axis.text = element_text(size = 18, colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title = element_text(size = 18, face = "bold", family = "Palatino"),
+    axis.text = element_text(size = 18, colour = "black", family = "Palatino"),
     plot.title = element_text(size = 18, hjust = 0),
-    legend.title = element_text(size = 18),
-    legend.text = element_text(size = 16),
+    legend.title = element_text(size = 18, family = "Palatino"),
+    legend.text = element_text(size = 16, family = "Palatino"),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
 

### 2020-07-30 site_map.Rmd
update

--- 2020-07-25 site_map.Rmd
@@
                      size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2,
-                     segment.alpha = 0.4, force = 0.1) +
+                     segment.alpha = 0.4, force = 0.1, font = "Palatino") +
   xlab("") + ylab("") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 16, y = -31.5,
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
              size =5, angle = 0, colour = "black") +
   guides(fill = guide_colourbar()) +
-# scale_x_continuous(breaks = seq(35, 25, by = 5),
-#                      labels  = c("35°S", "30°S", "25°S")) +
-#   scale_y_continuous(breaks = seq(16, 20, by = 2),
-#                      labels = c("16°E", "18°E", "20°E"))+
-  scale_x_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°E", sep = "")) +
-    scale_y_continuous(expand = c(0, 0),
-                       labels = scales::unit_format(unit = "°S", sep = "")) +
-  coord_fixed(ratio = 1, xlim = c(15.5, 20), ylim = c(-35,-28),
+  scale_y_continuous(breaks = seq(-34, -28, by = 2),
+                     labels  = c("34°S","32°S","30°S","28°S")) +
+  scale_x_continuous(breaks = seq(16, 20, by = 2),
+                      labels = c("16°E", "18°E", "20°E"))+
+#   scale_x_continuous(expand = c(0, 0),
+#                        labels = scales::unit_format(unit = "°E", sep = "")) +
+#     scale_y_continuous(expand = c(0, 0),
+#                        labels = scales::unit_format(unit = "°S", sep = "")) +
+ coord_fixed(ratio = 1, xlim = c(15.4, 20), ylim = c(-34.4,-27),
     expand = TRUE) +
-theme_minimal() +
-    theme(
-      axis.text = element_text(size = 10, family = "Palatino"),
-      axis.title = element_text(size = 10, family = "Palatino"),
-      legend.background = element_rect(colour = "black", size = 0.2),
-      legend.position = c(0.093, 0.70),
-      legend.text = element_text(family = "Palatino"),
-      legend.title = element_text(family = "Palatino"),
-      plot.caption = element_text(family = "Palatino"),
-      plot.title = element_text(family = "Palatino"),
-      plot.subtitle = element_text(family = "Palatino"),
-      panel.background = element_rect(colour = NA, fill = "ivory"),
-      panel.grid.major = element_line("grey70", linetype = "dashed", size = 0.2),
-      panel.grid.minor = element_line("grey70", linetype = "dashed", size = 0.2)
-    )
+  theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Palatino"),
+        axis.text.x = element_text(vjust = 1, family = "Palatino", size = 22),
+        axis.text.y = element_text(hjust = 1, family = "Palatino", size = 22),
+        axis.title.x = element_text(vjust = 0, family = "Palatino"),
+        axis.title.y = element_text(angle = 90, vjust = 0.3, family = "Palatino"),
+        axis.ticks = element_line(colour = "black"),
+        axis.ticks.length = unit(0.4, "cm"),
+        #axis.ticks.length = unit(0.15, "cm"),
+        axis.ticks.margin = unit(0.1, "cm"),
+        panel.background = element_rect(fill = "white", colour = "black", size = 0.5),
+        #panel.border = element_blank(),
+        panel.border = element_rect(colour = "black", fill=NA, size=2),
+        #(colour =" black"),
+        panel.grid.major.x = element_blank(),
+        #panel.grid.major.y = element_line(linetype = "dotted", size = 0.2, colour = "black"),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor = element_blank(),
+        #panel.margin = unit(0.25, "lines"),
+        strip.text = element_text(size = rel(1.0), family = "Palatino"),
+        strip.background = element_rect(fill = "white", colour = NA),
+        strip.text.x = element_text(),
+        strip.text.y = element_text(angle = -90),        
+        plot.background = element_rect(colour = "white"),
+        plot.title = element_text(size = rel(1.2), vjust = 1),
+        plot.margin = unit(c(1, 1, 0.5, 0.5), "lines"))
 

### 2020-07-30 ANOVA.R

@@ -100,3 +100,19 @@ summary(aov(intensity_mean ~ site + product + distance, data = lm_metrics_wide))
 summary(aov(intensity_max ~ site + product + distance, data = lm_metrics_wide))
 summary(aov(intensity_cumulative ~ site + product + distance, data = lm_metrics_wide))
 
+### ANOVA suggested by AJ to compare the differences in number of signals detected between sites
+metric_ANOVA <- combined_products %>% 
+  filter(year(date_start) %in% 2011:2014) %>%  # only for the years 2011-2014 so 4 year period
+  ungroup() %>%
+  group_by(product, site, distance) %>% 
+  summarise(y = n()) %>% 
+  rename(count = y) %>% 
+  mutate(distance = case_when(distance == "25000" ~ "25",
+                                distance == "50000" ~ "50",
+                              distance == "0" ~ "0",))
+
+
+# save(metric_ANOVA, file = "Data/metric_ANOVA.RData")
+
+summary(aov(count ~ site + product + distance, data = metric_ANOVA))
+

### 2020-07-30 box_plot.R

@@ -72,12 +72,15 @@ metrics_SACTN <- metrics_func(df = metric_SACTN)
 final_combined <- rbind(metrics, metric_SACTN)
 final_combined$count <- as.numeric(final_combined$count)
 
+Ordering <- c("OISST", "CMC","G1SST", "MUR", "SACTN")
+
 ggplot(data = final_combined, aes(x = product,y = duration)) +
   geom_point(data = metric_prods,shape = 21, fill = "lightgray",
              color = "black", size = 3) +
   geom_point(data = metric_SACTN,shape = 21, fill = "lightgray",
              color = "black", size = 3) +
   geom_boxplot(aes()) +
+  scale_x_discrete(limits = Ordering) +
   facet_wrap(~site)  +
   labs(y = "Duration (Days)", x = "SST products")+
   theme_set(theme_grey()) +
@@ -151,13 +154,15 @@ distance_corr <- SST_anom_spread %>%
   slope_calc()
 
 ## Bar graph plot
+Ordering <- c("OISST", "CMC","G1SST", "MUR")
 no_upwelling_sigs <- read_csv("Data_coast_angle/number_upwelling_sigs.csv")
 ggplot(data = no_upwelling_sigs, aes(x = SST_product, y = Number_of_signals, group = factor(Distance), fill = factor(Distance))) +
   geom_bar(stat = "identity",position = position_dodge2(width =0.5), width = 0.5) +
   scale_y_continuous(breaks = c(50,100,150,200,250,300,350,400))+
-  scale_x_discrete(limits = c("OISST","CMC", "MUR","G1SST")) +
+  scale_x_discrete(limits = Ordering) +
   scale_fill_grey(start = .1, end = .5, labels = c("0", "25", "50")) +
-  labs(x ="SST product", y = "Number of upwelling signals") +
+  scale_fill_discrete(labels = c("0", "25", "50")) +
+    labs(x ="SST product", y = "Number of upwelling signals") +
   labs(fill = "Distance (km)") + 
   theme_set(theme_grey()) +
   theme_grey() +
@@ -176,4 +181,35 @@ ggplot(data = no_upwelling_sigs, aes(x = SST_product, y = Number_of_signals, gro
     legend.text = element_text(size = 16, family = "Palatino"),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
+
+#########################
+load("Data/metric_ANOVA.RData") # Created in the ANOVA.R script
+
[…]

### 2020-08-03 site_map.Rmd
update

@@
 load("Data_coast_angle/site_pixels.RData")
+site_pixels <- read_csv("site_pixels.csv")
 

### 2020-08-03 box_plot.R

@@ -127,7 +127,7 @@ sst_anom_wide <- SST_anom %>%
   pivot_wider(id_cols = c(site, product, t), names_from = distance, values_from = c(temp, anom))
 
 # Doing the correlation using the easystat correlation package
-# https://rdrr.io/github/easystats/correlation/man/correlation.html
+# https://rdrr.io/github/easystats/correlation/man/   correlation.html
 SST_corr <- sst_anom_wide %>% 
   group_by(site, product) %>% 
   correlation(method = "pearson") #%>% 

### 2020-08-04 ANOVA.R
update

@@ -72,21 +72,18 @@ lm_metrics_wide <- pivot_wider(lm_metrics,
                                id_cols = site:distance, 
                                names_from = var, values_from = slope)
 
-### test example
-### define coefficients of linear function directly
-K <- diag(length(coef(lmod)))[-1,]
-rownames(K) <- names(coef(lmod))[-1]
-K
+# ### test example
+# ### define coefficients of linear function directly
+# K <- diag(length(coef(lmod)))[-1,]
+# rownames(K) <- names(coef(lmod))[-1]
+# K
+# ### set up general linear hypothesis
+# summary(glht(lmod, linfct = K))
 
-### set up general linear hypothesis
-summary(glht(lmod, linfct = K))
-
-### Why not just perform a normal ANOVA?
-# Why go to all this extra trouble?
-# I would recommend reading up more on ANOVA
-summary(aov(duration ~ site + product + distance, data = metric_4years))
+# AJ: 3-way ANOVA tests to compare the upwelling metrics such as distance, cummulative intensity and intensity against the different sites, products and distances
+summary(aov(duration ~ site + product + distance , data = metric_4years))
 summary(aov(intensity_mean ~ site + product + distance, data = metric_4years))
-summary(aov(intensity_max ~ site + product + distance, data = metric_4years))
+# summary(aov(intensity_max ~ site + product + distance, data = metric_4years))
 summary(aov(intensity_cumulative ~ site + product + distance, data = metric_4years))
 
 # Here are ANOVAs on the linear model results

### 2020-08-04 ANOVA.R
update

@@ -86,6 +86,12 @@ summary(aov(intensity_mean ~ site + product + distance, data = metric_4years))
 # summary(aov(intensity_max ~ site + product + distance, data = metric_4years))
 summary(aov(intensity_cumulative ~ site + product + distance, data = metric_4years))
 
+## Interaction?
+summary(aov(duration ~ site * product * distance , data = metric_4years))
+summary(aov(intensity_mean ~ site * product * distance, data = metric_4years))
+# summary(aov(intensity_max ~ site + product + distance, data = metric_4years))
+summary(aov(intensity_cumulative ~ site * product * distance, data = metric_4years))
+
 # Here are ANOVAs on the linear model results
 # NB: This is not advisable as these linear models are based on only 4 years of data.
 # That is not enough to draw any conclusions about rates of change over time

### 2020-08-16 ANOVA.R
neat

@@ -95,7 +95,7 @@ load("Data_coast_angle/G1SST_upwell_base.RData")
 combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base)
 # save(combined_products, file = "Data_coast_angle/combined_products.RData")
 
-# 2: Preparing box plot data ----------------------------------------------------
+# 2: Preparing box plot data ---------------------------------------------------------------------------------------------------------------------------------
 
 seasons_func <- function(df){
   BC_seaons <- df %>%

### 2020-08-17 site_map.Rmd
update

@@
    geom_polygon(aes(group = group),fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
+  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
+               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE)+
   geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
@@
                       labels = c("16°E", "18°E", "20°E"))+
+  scale_fill_gradient(low = "steelblue4", high = "steelblue1", na.value = "steelblue4", 
+                      breaks = c(-1000, -2000, -3000, -4000, -5000),
+                      guide_legend(title = "depth (m)")) +
 #   scale_x_continuous(expand = c(0, 0),
@@
 ```
+
+# Bathy plot
+
+```{r}
+fig1_bathy <- ggplot() + theme_bw() + #coord_equal() + 
+  geom_raster(data = sa_bathy, aes(x = lon, y = lat, fill = depth)) +
+  #geom_raster(data = isobath200, aes(x = lon, y = lat, fill = coast)) +
+  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
+               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE) +
+  geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
+               size = 0.1, colour = "black", fill = "grey80") +
+  #geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
+  # geom_polygon(data = scBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
+  #              colour = "green", fill = "green") +
+  # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
+  #              colour = "blue", fill = "blue") +
+  # geom_polygon(data = ecBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
+  #              colour = "orange", fill = "orange") +
+  geom_point(data = site_pixels, aes(x = lon, y = lat, colour = site), alpha = 0.8, size = 3) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "red", shape = 0, alpha = 0.8, size = 2.1) +
+  #geom_rect(data = test, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
+            #alpha = 1, colour = "red", size = 0.1, linetype = 1) +
+   scale_y_continuous(breaks = seq(-34, -30, by = 2),
+                     labels  = c("34°S","32°S","30°S")) +
+  scale_x_continuous(breaks = seq(16, 18, by = 2),
+                      labels = c("16°E", "18°E"))+
+  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
+                         guide_legend(title = "depth (m)")) +
+  scale_fill_gradient(low = "steelblue4", high = "steelblue1", na.value = "steelblue4", 
+                      breaks = c(-1000, -2000, -3000, -4000, -5000),
+                      guide_legend(title = "depth (m)")) +
+  labs(title = NULL, x = NULL, y = NULL) +
+  guides(colour = guide_legend(override.aes = list(size = 3))) +
+  #guide_legend(ncol = 2) +
+  ### Annotate specific things:
+  annotate("text", label = "Cape \n Agulhas", x = 20.0, y = -35.1, size = 2., colour = "white") +
+  #annotate("text", label = "False Bay", x = 18.66, y = -34.55, size = 2.2) +
+  annotate("text", label = "Algoa \n Bay", x = 26.4, y = -34.0, size = 2., colour = "white") +
+  annotate("text", label = "Cape \n Point", x = 18.0, y = -34.4, size = 2, colour = "white") +
+  annotate("text", label = "Hamburg", x = 28.3, y = -33.28611111, size = 2, colour = "white") +
+  theme(#panel.background = element_rect(fill = "steelblue4", colour = NA),
+        #panel.border = element_blank(),
+        #axis.title = element_blank(),
+        #axis.text = element_blank(),
+        #axis.ticks = element_blank(),
+        #panel.grid.minor = element_blank(),
+        #panel.grid.major = element_blank(),
+        #plot.title = element_blank(),
+        #legend.text = element_text(size = 7, colour = "White"),
+        #legend.title = element_text(size = 7, colour = "White"),
+        legend.key = element_rect(colour = NA, size = 0.2),
+        legend.key.height = unit(0.4, "cm"),
+        legend.background = element_blank(),
+        legend.justification = c(1,0), legend.position = c(0.5, 0.4)) +
+  coord_cartesian(xlim = sa_lons, ylim = sa_lats)
+```

### 2020-08-20 site_map.Rmd
updated

@@
 
+load("~/Documents/Upwelling_MCS/Data/sa_bathy.RData")
+
 #load("Data/site_pixels.RData")
@@
    geom_polygon(aes(group = group),fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
-  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
+  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), 
                colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE)+
   geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "red", shape = 0, alpha = 0.8, size = 2.1) +
   geom_label_repel(data =site_list_sub, aes(x = lon, y = lat, label = site),
@@
   scale_x_continuous(breaks = seq(16, 20, by = 2),
-                      labels = c("16°E", "18°E", "20°E"))+
-  scale_fill_gradient(low = "steelblue4", high = "steelblue1", na.value = "steelblue4", 
-                      breaks = c(-1000, -2000, -3000, -4000, -5000),
-                      guide_legend(title = "depth (m)")) +
-#   scale_x_continuous(expand = c(0, 0),
-#                        labels = scales::unit_format(unit = "°E", sep = "")) +
-#     scale_y_continuous(expand = c(0, 0),
-#                        labels = scales::unit_format(unit = "°S", sep = "")) +
+                      labels = c("16°E", "18°E", "20°E")) +
  coord_fixed(ratio = 1, xlim = c(15.4, 20), ylim = c(-34.4,-27),

### 2020-09-16 site_map.Rmd
update

@@
 site_map
+
+africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
+  theme_bw() + 
+  coord_equal() + 
+  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
+  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
+  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
+  theme(panel.border = element_rect(colour = "black", size = 0.4), 
+        plot.background = element_blank(), 
+        axis.ticks = element_blank(), 
+        axis.text = element_blank(), 
+        axis.title = element_blank(), 
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank()) +
+  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
+africa_map
+
+sa_4 <- site_map +
+annotation_custom(grob = ggplotGrob(africa_map),
+xmin = 20.9, xmax = 26.9,
+ymin = -30, ymax = -24)
+sa_4
 ```

### 2020-09-16 ANOVA.R

@@ -130,10 +130,10 @@ summary(aov(duration ~ site, data = final[final$product == "G1SST", ]))
 summary(aov(duration ~ site, data = final[final$product == "MUR", ]))
 summary(aov(duration ~ site, data = final[final$product == "SACTN", ]))
 
-ggplot(data = final, aes(x = site, y = duration)) +
+plot1 <- ggplot(data = final, aes(x = product, y = duration)) +
   geom_boxplot() +
-  facet_wrap(vars(product), ncol = 2) +
-  xlab("Site") + ylab("Duration (days)") +
+  facet_wrap(vars(site), ncol = 2) +
+  xlab("SST product") + ylab("Duration (days)") +
   theme_set(theme_grey()) +
   theme_grey() +
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
@@ -142,10 +142,10 @@ ggplot(data = final, aes(x = site, y = duration)) +
     strip.text = element_text(size=14, family = "Palatino"),
     axis.title = element_text(size = 18, face = "bold", family = "Palatino"),
     axis.ticks.length = unit(0.4, "cm"),
-    axis.text = element_text(size = 18, colour = "black", family = "Palatino"),
+    axis.text = element_text(size = 10, colour = "black", family = "Palatino"),
     plot.title = element_text(size = 18, hjust = 0),
     legend.title = element_text(size = 18),
-    legend.text = element_text(size = 16),
+    legend.text = element_text(size = 12),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
 
@@ -155,10 +155,10 @@ summary(aov(intensity_mean ~ site, data = final[final$product == "G1SST", ]))
 summary(aov(intensity_mean ~ site, data = final[final$product == "MUR", ]))
 summary(aov(intensity_mean ~ site, data = final[final$product == "SACTN", ]))
 
-ggplot(data = final, aes(x = site, y = intensity_mean)) +
+plot2 <- ggplot(data = final, aes(x = product, y = intensity_mean)) +
   geom_boxplot() +
-  facet_wrap(vars(product), ncol = 2) +
-  xlab("Site") + ylab("Upwelling mean intensity (°C)") +
+  facet_wrap(vars(site), ncol = 2) +
+  xlab("SST product") + ylab("Upwelling mean intensity (°C)") +
   theme_set(theme_grey()) +
   theme_grey() +
[…]

### 2020-09-20 site_map.Rmd
update

--- 2020-08-20 site_map.Rmd
@@
 site_map
+
+africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) +
+  theme_bw() +
+  coord_equal() +
+  geom_polygon(aes(group = group), colour = "black", fill = "grey80") +
+  #geom_polygon(data = sa_provinces_new, (aes(group = group))) +
+  geom_rect(data=africa_coast, aes(xmin=15.0, xmax=21.0, ymin=-35.0, ymax=-28.0),  colour = "red", fill = NA, size = 0.75) +
+  #annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) +
+  theme(panel.border = element_rect(colour = "black", size = 0.4),
+        plot.background = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        axis.title = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank()) +
+  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator")
+africa_map
+
+sa_4 <- site_map +
+annotation_custom(grob = ggplotGrob(africa_map),
+xmin = 17, xmax = 21.4,
+ymin = -28.8, ymax = -26.6)
+sa_4
 ```

### 2020-09-20 ANOVA.R

@@ -85,7 +85,7 @@ lm_metrics_wide <- pivot_wider(lm_metrics,
 # summary(aov(intensity_cumulative ~ site * product * distance, data = metric_4years))
 # ---------------------------------------------------------------------------------------------------------------------------------------------------------
 # ANOVA WITH SACTN
-
+# A) DIFFERENCES BETWEEN SITES, PER PRODUCT -------------------------------
 load("Data_coast_angle/OISST_upwell_base.RData")
 load("Data_coast_angle/CMC_upwell_base.RData")
 load("Data_coast_angle/SACTN_upwell_base.RData")
@@ -132,15 +132,15 @@ summary(aov(duration ~ site, data = final[final$product == "SACTN", ]))
 
 plot1 <- ggplot(data = final, aes(x = product, y = duration)) +
   geom_boxplot() +
-  facet_wrap(vars(site), ncol = 2) +
-  xlab("SST product") + ylab("Duration (days)") +
+  facet_wrap(vars(site), ncol = 4) +
+  xlab("") + ylab("Duration (days)") +
   theme_set(theme_grey()) +
   theme_grey() +
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
     panel.grid.major = element_line(size = 0.2, linetype = 2),
     panel.grid.minor = element_line(colour = NA),
     strip.text = element_text(size=14, family = "Palatino"),
-    axis.title = element_text(size = 18, face = "bold", family = "Palatino"),
+    axis.title = element_text(size = 15, face = "bold", family = "Palatino"),
     axis.ticks.length = unit(0.4, "cm"),
     axis.text = element_text(size = 10, colour = "black", family = "Palatino"),
     plot.title = element_text(size = 18, hjust = 0),
@@ -157,15 +157,15 @@ summary(aov(intensity_mean ~ site, data = final[final$product == "SACTN", ]))
 
 plot2 <- ggplot(data = final, aes(x = product, y = intensity_mean)) +
   geom_boxplot() +
-  facet_wrap(vars(site), ncol = 2) +
-  xlab("SST product") + ylab("Upwelling mean intensity (°C)") +
+  facet_wrap(vars(site), ncol = 4) +
+  xlab("") + ylab("Mean intensity (°C)") +
   theme_set(theme_grey()) +
   theme_grey() +
   theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
[…]

### 2020-09-20 box_plot.R

@@ -180,7 +180,7 @@ ggplot(data = test1, aes(x = product, y = mean_signals, group = factor(Distance)
   scale_x_discrete(limits = Ordering) +
   scale_fill_grey(start = .1, end = .5, labels = c("0", "25", "50")) +
   scale_fill_discrete(labels = c("0", "25", "50")) +
-    labs(x ="SST product", y = "Average number of upwelling signals") +
+    labs(x ="", y = "Upwelling events (count)") +
   labs(fill = "Distance (km)") + 
   facet_wrap(~site) +
   theme_set(theme_grey()) +

### 2020-09-20 site_map.Rmd
update

@@
   xlab("") + ylab("") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
-             size =5, angle = 0, colour = "black") +
+    # annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
+    #          size =5, angle = 0, colour = "black") +
+  annotate("text", label = "Cape Columbine", x = 15.5, y = -30.5,
+             size =5, angle = 45, colour = "black")+
+  annotate("text", label = "Cape Peninsula", x = 16.5, y = -33.9,
+             size =5, angle = 45, colour = "black")+
   guides(fill = guide_colourbar()) +

### 2020-09-21 site_map.Rmd
update

--- 2020-09-16 site_map.Rmd
@@
   xlab("") + ylab("") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
-             size =5, angle = 0, colour = "black") +
+    # annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
+    #          size =5, angle = 0, colour = "black") +
+  annotate("text", label = "Cape Columbine", x = 16.0, y = -30.5,
+             size =5, angle = 45, colour = "black")+
+  annotate("text", label = "Cape Peninsula", x = 16.5, y = -33.9,
+             size =5, angle = 45, colour = "black")+
   guides(fill = guide_colourbar()) +
@@
   geom_polygon(aes(group = group), colour = "black", fill = "grey80") +
+  geom_rect(data=africa_coast, aes(xmin=15.0, xmax=21.0, ymin=-34.4, ymax=-29.5),  colour = "red", fill = NA, size = 0.35) +
   geom_polygon(data = sa_provinces_new, (aes(group = group))) +
-  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
+  #annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) +
   theme(panel.border = element_rect(colour = "black", size = 0.4),
@@
 annotation_custom(grob = ggplotGrob(africa_map),
-xmin = 20.9, xmax = 26.9,
-ymin = -30, ymax = -24)
+xmin = 17, xmax = 21.5,
+ymin = -28.8, ymax = -26.6)
 sa_4

### 2020-09-21 ANOVA.R

@@ -130,9 +130,11 @@ summary(aov(duration ~ site, data = final[final$product == "G1SST", ]))
 summary(aov(duration ~ site, data = final[final$product == "MUR", ]))
 summary(aov(duration ~ site, data = final[final$product == "SACTN", ]))
 
+Ordering <- c("OISST", "CMC","G1SST", "MUR")
 plot1 <- ggplot(data = final, aes(x = product, y = duration)) +
-  geom_boxplot() +
+  geom_boxplot(notch=TRUE) +
   facet_wrap(vars(site), ncol = 4) +
+  scale_x_discrete(limits = Ordering) +
   xlab("") + ylab("Duration (days)") +
   theme_set(theme_grey()) +
   theme_grey() +
@@ -156,7 +158,8 @@ summary(aov(intensity_mean ~ site, data = final[final$product == "MUR", ]))
 summary(aov(intensity_mean ~ site, data = final[final$product == "SACTN", ]))
 
 plot2 <- ggplot(data = final, aes(x = product, y = intensity_mean)) +
-  geom_boxplot() +
+  geom_boxplot(notch=TRUE) +
+  scale_x_discrete(limits = Ordering) +
   facet_wrap(vars(site), ncol = 4) +
   xlab("") + ylab("Mean intensity (°C)") +
   theme_set(theme_grey()) +
@@ -183,8 +186,9 @@ summary(aov(intensity_cumulative ~ site, data = final[final$product == "MUR", ])
 summary(aov(intensity_cumulative ~ site, data = final[final$product == "SACTN", ]))
 
 plot3 <- ggplot(data = final, aes(x = product, y = intensity_cumulative)) +
-  geom_boxplot() +
+  geom_boxplot(notch=TRUE) +
   facet_wrap(vars(site), ncol = 4) +
+  scale_x_discrete(limits = Ordering) +
   xlab("") + ylab("Cumulative intensity
   (°C.days)") +
   theme_set(theme_grey()) +
@@ -349,7 +353,7 @@ summary(aov(duration ~ distance, data = metric_prods[metric_prods$product == "OI
 # use above example and make your own ANOVA table, and include the figure below...
 
 plotA <- ggplot(data = metric_prods, aes(x = as.factor(distance), y = duration)) +
-  geom_boxplot() +
+  geom_boxplot(notch = TRUE) +
   facet_wrap(vars(product), ncol = 4) +
[…]

### 2020-09-21 box_plot.R

@@ -12,6 +12,7 @@ library(tidyverse)
 library(lubridate)
 library(multcomp)
 options(scipen = 999)
+library(lubridate)
 # devtools::install_github("twitter/AnomalyDetection")
 #library(AnomalyDetection)
 #devtools::install_github("easystats/correlation")

### 2020-09-30 ANOVA.R
update

@@ -11,7 +11,11 @@
   load("Data_coast_angle/SACTN_upwell_base.RData")
   load("Data_coast_angle/MUR_upwell_base.RData")
   load("Data_coast_angle/G1SST_upwell_base.RData")
-
+  load("Data_coast_angle/CMC_2015_upwell_base.RData")
+  load("Data_coast_angle/OISST_2015_upwell_base.RData")
+  
+  CMC_upwell_base <- rbind(CMC_2015_upwell_base, CMC_upwell_base)
+  OISST_upwell_base <- rbind(OISST_2015_upwell_base, OISST_upwell_base)
 # # Removing the distance of 20 and 40kms
 # library(dplyr)
 # removing_distance_func <- function(df){
@@ -34,7 +38,7 @@
 combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base,G1SST_upwell_base)
 
 metric_4years <- combined_products %>% 
-  filter(year(date_start) %in% 2011:2014) # only for the years 2011-2014 so 4 year period
+  filter(year(date_start) %in% 2011:2016) # only for the years 2011-2014 so 4 year period
 
 
 metrics <- metric_4years %>% 
@@ -111,15 +115,16 @@ combined_products <- seasons_func(combined_products)
 SACTN <- seasons_func(df = SACTN_upwell_base)
 
 metric_prods <- combined_products %>% 
-  filter(year(date_start) %in% 2011:2014) %>% 
+  filter(year(date_start) %in% 2011:2016) %>% 
   filter(season == "Summer") %>% 
   ungroup() %>% 
   dplyr::select(-heading,-distance)
 metric_prods <- as.data.frame(metric_prods)
 
 metric_SACTN <- SACTN %>% 
-  filter(year(date_start) %in% 2011:2014) %>% 
-  mutate(product = "SACTN") 
+  filter(year(date_start) %in% 2011:2016) %>% 
+  mutate(product = "SACTN") %>% 
+  filter(season == "Summer")
 metric_SACTN <- as.data.frame(metric_SACTN)
 
 final <- rbind(metric_SACTN,metric_prods)
@@ -130,7 +135,7 @@ summary(aov(duration ~ site, data = final[final$product == "G1SST", ]))
 summary(aov(duration ~ site, data = final[final$product == "MUR", ]))
 summary(aov(duration ~ site, data = final[final$product == "SACTN", ]))
[…]

### 2020-09-30 SST_patterns.R
update

@@ -37,10 +37,12 @@ load("Data_coast_angle/site_pixels.RData")
 # ### The temperature products
 # load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
 
-OISST_prod <- OISST %>%
+OISST_prod <- BC2015_2016 %>%
   dplyr::select(lon, lat) %>%
   unique() %>%
   mutate(product = "OISST")
+
+CMC_all_yrs <- rbind(CMC, CMC_2015_2016)
 
 CMC_prod <- CMC %>%
   dplyr::select(lon, lat) %>%
@@ -57,7 +59,7 @@ G1SST_prod <- G1SST %>%
   unique() %>%
   mutate(product = "G1SST")
 
-sat_data <- rbind(MUR_prod, OISST_prod) %>% 
+sat_data <- rbind(CMC_prod, OISST_prod) %>% 
   dplyr::select(product, lon, lat)
 
 sat_pixels <- sat_data %>%
@@ -70,29 +72,33 @@ match_func <- function(df){
   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
     cbind(., df)
-  MUR_index <- MUR_prod[as.vector(knnx.index(as.matrix(MUR_prod[,c("lon", "lat")]),
+  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
                                              as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
     cbind(., df)
-  res <- rbind(OISST_index, MUR_index)
+  res <- rbind(OISST_index, CMC_index)
   return(res)
 }
+
 
 pixel_match <- site_pixels %>%
   group_by(site) %>%
   group_modify(~match_func(.x))
 
 
-OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
-MUR_fill <- right_join(MUR, filter(pixel_match, product == "MUR"), by = c("lon", "lat"))
+OISST_fill <- right_join(BC2015_2016, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
+CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
 rm(OISST, MUR); gc()
 
[…]

### 2020-09-30 site_map.Rmd
update

@@
 load("Data_coast_angle/site_pixels.RData")
+load("~/Documents/HONOURSPROJECT-master/data/africa_coast.RData")
 site_pixels <- read_csv("site_pixels.csv")
@@
   geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "red", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat, fill = "red"), colour = "red", shape = 15, alpha = 0.8, size = 2.1) +
   geom_label_repel(data =site_list_sub, aes(x = lon, y = lat, label = site),

### 2020-09-30 wind_analyses.Rmd
update

@@
 ```{r}
-unique_pixel <- OISST %>%
+unique_pixel <- BC2015_2016 %>%
   select(lon, lat) %>%
@@
 
-combined_OISST_match <- OISST %>%
+combined_OISST_match <- BC2015_2016 %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
@@
 
+OISST_final2015_2016 <- combined_OISST_sites %>% 
+  select(site, date, temp, lat, lon)
 ############
-unique_pixel <- CMC %>%
+unique_pixel <- CMC_2015_2016 %>%
   select(lon, lat) %>%
@@
 
-combined_CMC_match <- CMC %>%
+combined_CMC_match <- CMC_2015_2016 %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%

### 2020-09-30 Extracting/Extracting_data.R
update

@@ -272,9 +272,6 @@ save(site_pixels, file = "Data_coast_angle/site_pixels.RData")
 
 # 5: Extracting OISST dataset ----------------------------------------------------
 
-
-
-
 # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
 #                    CC = c(25, 35, 340, 355), # Canary Current
 #                    CalC = c(35, 45, 225, 240), # California Current

### 2020-10-01 SST_patterns.R
edits all yrs

@@ -187,6 +187,10 @@ ts2clm_custom <- function(df){
   return(res)
 }
 
+CMC_clims <- rbind(CMC_fill,CMC_fill_2015_2016)
+OISST_clims <- rbind(OISST_fill,OISST_fill_2015_2016)
+
+
 G1SST_last <- G1SST_last %>% 
   arrange(date)
 
@@ -202,11 +206,9 @@ MUR_upwell_clims <- upwelling_detect_event(df = MUR_fill)
 MUR_upwell_clims$distance <- as.numeric(MUR_upwell_clims$distance)
 #save(MUR_upwell_clims, file = "Data_coast_angle/MUR_upwell_clims.RData")
 
-
-
-
-
-
+# More years
+CMC_all_yrs <- upwelling_detect_event(df = CMC_clims)
+OISST_all_yrs <- upwelling_detect_event(df = OISST_clims)
 
 
 

### 2020-10-01 box_plot.R

@@ -113,7 +113,7 @@ load("Data_coast_angle/G1SST_upwell_clims.RData")
 
 SST_clims <- rbind(OISST_upwell_clims,G1SST_upwell_clims,CMC_upwell_clims,MUR_upwell_clims)
 
-# Determining the anomaly temperature for SST products and SACTN
+# Determining the anomaly temperature for  SST products and SACTN
 SST_anom <- SST_clims %>% 
   group_by(site, product, distance) %>% 
   mutate(anom = temp - seas) %>% 
@@ -207,7 +207,8 @@ load("Data/metric_ANOVA.RData") # Created in the ANOVA.R script
 
 Ordering <- c("OISST", "CMC","G1SST", "MUR")
 ggplot(data = metric_ANOVA, aes(x = product, y = count, group = factor(distance), fill = factor(distance))) +
-  geom_bar(stat = "identity",position = position_dodge2(width =0.5), width = 0.5) +
+  #geom_bar(stat = "identity",position = position_dodge2(width =0.5), width = 0.5) +
+  geom_boxplot()+
   scale_y_continuous(breaks = c(50,100,150,200,250,300,350,400))+
   scale_x_discrete(limits = Ordering) +
   scale_fill_grey(start = .1, end = .5, labels = c("0", "25", "50")) +
@@ -229,6 +230,24 @@ ggplot(data = metric_ANOVA, aes(x = product, y = count, group = factor(distance)
     plot.title = element_text(size = 18, hjust = 0),
     legend.title = element_text(size = 18, family = "Palatino"),
     legend.text = element_text(size = 16, family = "Palatino"),
+    legend.key = element_rect(size = 0.8, colour = NA),
+    legend.background = element_blank())
+
+plot8 <- ggplot(data = metric_ANOVA, aes(x = product, y = count)) +
+  geom_boxplot() +
+  facet_wrap(vars(site), ncol = 4) +
+    xlab("Data product") + ylab("Number of upwelling signals ")+
+  theme_grey() +
+  theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
+    panel.grid.major = element_line(size = 0.2, linetype = 2),
+    panel.grid.minor = element_line(colour = NA),
+    strip.text = element_text(size=14, family = "Palatino"),
+    axis.title = element_text(size = 18, face = "bold", family = "Palatino"),
+    axis.ticks.length = unit(0.4, "cm"),
[…]

### 2020-10-01 ANOVA.R
temporary update

@@ -599,3 +599,7 @@ summary(aov(count ~ site, data = metric_ANOVA))
 summary(aov(count ~ product + distance/site, data = metric_ANOVA))
 summary(aov(count ~ distance, data = metric_ANOVA))
 
+##
+# Plot the count of SST against temperature and see if changes occured
+
+

### 2020-10-01 SST_patterns.R

@@ -128,7 +128,7 @@ upwelling_detect_event <- function(df){
   upwell_base <- df %>%
     dplyr::rename(t = date) %>%
     group_by(site, product, heading, distance, lon, lat) %>%
-    group_modify(~ts2clm_custom(.x)) %>%
+    #group_modify(~ts2clm_custom(.x)) %>%
     left_join(upwelling, by = c("site", "t")) %>%
     filter(!is.na(exceedance)) %>%
     group_by(site, product, heading, distance, lon, lat) %>%
@@ -164,7 +164,7 @@ SACTN_US <- SACTN_US %>%
 load("Data/site_list_v4.2.RData")
 
 SACTN_upwell_base <- SACTN_US %>%
-  left_join(site_list[,c(4, 5, 6)], by = "index")%>%
+ left_join(site_list[,c(4, 5, 6)], by = "index")%>%
   dplyr::rename(t = date) %>%
   group_by(site, lon, lat) %>%
   group_modify(~ts2clm_custom(.x)) %>%
@@ -183,7 +183,7 @@ detect_event_custom <- function(df){
 }
 
 ts2clm_custom <- function(df){
-  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2017-12-31")) #Length of MUR time series: Chnage according to length os SST product
+  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1982-01-01", "2005-10-19 ")) #Length of MUR time series: Chnage according to length os SST product
   return(res)
 }
 
@@ -206,13 +206,30 @@ MUR_upwell_clims <- upwelling_detect_event(df = MUR_fill)
 MUR_upwell_clims$distance <- as.numeric(MUR_upwell_clims$distance)
 #save(MUR_upwell_clims, file = "Data_coast_angle/MUR_upwell_clims.RData")
 
-# More years
-CMC_all_yrs <- upwelling_detect_event(df = CMC_clims)
-OISST_all_yrs <- upwelling_detect_event(df = OISST_clims)
+#### With wind only as a filter
 
+ts2clm_custom <- function(df){
+  res <- ts2clm(df, pctile = 100, climatologyPeriod = c("2011-01-02", "2014-06-30")) #Length of MUR time series: Chnage according to length os SST product
+  return(res)
+}
 
+G1SST_without_temp <- upwelling_detect_event(df = G1SST_last)
[…]

### 2021-02-10 box_plot.R
update

@@ -360,6 +360,3 @@ ggplot(data = final_combined, aes(x = product,y = intensity_cumulative)) +
     legend.text = element_text(size = 16),
     legend.key = element_rect(size = 0.8, colour = NA),
     legend.background = element_blank())
-
-
-

### 2021-03-24 site_map.Rmd
update
Arial font on plots for publication

@@
                      size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2,
-                     segment.alpha = 0.4, force = 0.1, font = "Palatino") +
+                     segment.alpha = 0.4, force = 0.1, font = "Arial") +
   xlab("") + ylab("") +
@@
     expand = TRUE) +
-  theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Palatino"),
-        axis.text.x = element_text(vjust = 1, family = "Palatino", size = 22),
-        axis.text.y = element_text(hjust = 1, family = "Palatino", size = 22),
-        axis.title.x = element_text(vjust = 0, family = "Palatino"),
-        axis.title.y = element_text(angle = 90, vjust = 0.3, family = "Palatino"),
+  theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Arial"),
+        axis.text.x = element_text(vjust = 1, family = "Arial", size = 22),
+        axis.text.y = element_text(hjust = 1, family = "Arial", size = 22),
+        axis.title.x = element_text(vjust = 0, family = "Arial"),
+        axis.title.y = element_text(angle = 90, vjust = 0.3, family = "Arial"),
         axis.ticks = element_line(colour = "black"),
@@
         #panel.margin = unit(0.25, "lines"),
-        strip.text = element_text(size = rel(1.0), family = "Palatino"),
+        strip.text = element_text(size = rel(1.0), family = "Palatino"), #Font used to be Palatino but changed to Arial for publication purposes
         strip.background = element_rect(fill = "white", colour = NA),
@@
 sa_4
+
+ggsave("sa_4.tiff", units="cm", width=19.05, height=22.23, dpi=300)
 ```

### 2021-03-24 ANOVA.R

@@ -163,18 +163,18 @@ plot1 <- ggplot(data = final, aes(x = site, y = duration)) +
   xlab("") + ylab("Duration (days)") +
   theme_minimal() +
   theme(
-    strip.text = element_text(size = 7, family = "Palatino"),
+    strip.text = element_text(size = 7, family = "Arial"),
     axis.text.x = element_text(angle = 45),
     strip.placement = "outside",
-    axis.text = element_text(size = 6, colour = "black", family = "Palatino"),
-    axis.title = element_text(size = 8, face = "bold", family = "Palatino"),
+    axis.text = element_text(size = 6, colour = "black", family = "Arial"),
+    axis.title = element_text(size = 8, face = "bold", family = "Arial"),
     axis.ticks.length = unit(0.2, "cm"),
     panel.grid.major = element_line("grey70", linetype = "dashed", size = 0.2),
     panel.grid.minor = element_line("grey70", linetype = "dashed", size = 0.2),
     plot.title = element_text(size = 18, hjust = 0),
-    legend.title = element_text(size = 8, family = "Palatino"),
+    legend.title = element_text(size = 8, family = "Arial"),
     legend.position = "right",
-    legend.text = element_text(size = 7, family = "Palatino"),
+    legend.text = element_text(size = 7, family = "Arial"),
     legend.key.size = unit(0.2, "cm"),
     legend.background = element_blank())
   # theme_set(theme_grey()) +
@@ -205,18 +205,18 @@ plot2 <- ggplot(data = final, aes(x = site, y = intensity_mean)) +
   xlab("") + ylab("Mean intensity (°C)") +
   theme_minimal() +
   theme(
-    strip.text = element_text(size = 7, family = "Palatino"),
+    strip.text = element_text(size = 7, family = "Arial"),
     axis.text.x = element_text(angle = 45),
     strip.placement = "outside",
-    axis.text = element_text(size = 6, colour = "black", family = "Palatino"),
-    axis.title = element_text(size = 8, face = "bold", family = "Palatino"),
+    axis.text = element_text(size = 6, colour = "black", family = "Arial"),
+    axis.title = element_text(size = 8, face = "bold", family = "Arial"),
     axis.ticks.length = unit(0.2, "cm"),
     panel.grid.major = element_line("grey70", linetype = "dashed", size = 0.2),
     panel.grid.minor = element_line("grey70", linetype = "dashed", size = 0.2),
     plot.title = element_text(size = 18, hjust = 0),
[…]

### 2021-05-11 upwell_IDX.Rmd
manuscript

@@
   dplyr::rename(t = date) %>% # renaming date to t
-  drop_na()
-#save(UI_angle, file = "Data_coast_angle/UI_angle.RData")
+  drop_na() %>% 
+  filter(mean_speed >5)
+save(UI_angle, file = "UI_angle.RData")
 # save(UI_angle, file = "Data/UI_angle.RData")
@@
   nest() %>% 
-  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1973-01-01", "2016-12-31")), 
+  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-10", "2016-12-31")), 
          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  

### 2021-05-11 ANOVA.R

@@ -18,14 +18,11 @@
   OISST_upwell_base <- rbind(OISST_2015_upwell_base, OISST_upwell_base)
   MUR_upwell_base <- rbind(MUR_upwell_base, MUR_upwell_base_2015)
   
-  MUR_upwell_base <-  MUR_upwell_base %>% 
-    filter(duration < 40)
-  
 # # Removing the distance of 20 and 40kms
 # library(dplyr)
 # removing_distance_func <- function(df){
 #   removing_dist_func<- df %>% 
-#     filter(!distance %in% c(20000,40000))
+#     filter(!distance %in% c(20000,40000))gc
 # }
 # 
 # OISST_final <- removing_distance_func(df = OISST_upwell_base)
@@ -40,7 +37,7 @@
 # load("Data/MUR_final.RData")  
 # load("Data/CMC_final.RData")
 
-combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base,G1SST_upwell_base)
+combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base_2015,G1SST_upwell_base)
 
 metric_4years <- combined_products %>% 
   filter(year(date_start) %in% 2011:2016) # only for the years 2011-2014 so 4 year period
@@ -107,9 +104,18 @@ load("Data_coast_angle/SACTN_upwell_base.RData")
 load("Data_coast_angle/MUR_upwell_base.RData")
 load("Data_coast_angle/G1SST_upwell_base.RData")
 
-combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base, G1SST_upwell_base)
-# save(combined_products, file = "Data_coast_angle/combined_products.RData")
+MUR_upwell_base <- MUR_upwell_base_2015 %>% 
+  mutate(distance = as.numeric(as.character(distance)))
 
+G1SST_upwell_base_test <- G1SST_upwell_base %>% 
+  mutate(distance = as.numeric(as.character(distance))) %>% 
+  arrange(desc(duration))
+
+G1SST_upwell_base_test = G1SST_upwell_base_test[-1,] #press 3 times
+G1SST_upwell_base <- G1SST_upwell_base_test
+
+combined_products <- rbind(OISST_upwell_base,CMC_upwell_base,MUR_upwell_base,G1SST_upwell_base)
+# save(combined_products, file = "Data_coast_angle/combined_products.RData")
[…]

### 2021-05-11 SST_patterns.R

@@ -121,30 +121,25 @@ save(CMC_fill_2015_2016, file = "Data_coast_angle/CMC_fill_2015_2016.RData")
 # Loading the data
 # load("Data/UI_angle.RData")
 load("Data_coast_angle/UI_angle.RData") # Created in script 'upwell_IDX.Rmd'
-# load("Data_coast_angle/G1SST_last.RData") # Created in Extracting folder : extract_tidync.R script"
-# load("Data_coast_angle/CMC_fill_2015_2016.RData")
+load("Data_coast_angle/G1SST_last.RData") # Created in Extracting folder : extract_tidync.R script"
+load("Data_coast_angle/CMC_fill_2015_2016.RData")
 load("Data_coast_angle/MUR_fill.RData")
 load("Data_coast_angle/OISST_fill_2015_2016.RData")
 load("Data_coast_angle/MUR_fill_1.RData")
-# load("Data_coast_angle/CMC_fill.RData")
+load("Data_coast_angle/CMC_fill.RData")
 load("Data_coast_angle/OISST_fill.RData")
 load("Data_coast_angle/MUR_updated.RData")
 
-MUR_updated <- MUR_updated %>% 
-  arrange(temp) 
+MUR <- MUR_fill %>% 
+  filter(year(date) %in% seq(2011, 2016, 1))
 
-MUR_1 <- MUR_updated %>% 
-  slice(1:2315)
+MUR_1 <- MUR_fill_1 %>%
+  mutate(temp = temp - 273.15) %>% 
+  rename(date =Month_Yr)
 
-updated <- MUR_updated %>% 
-  slice(2316:4185) %>% 
-  mutate(temp = temp - 273.15)
-
-MUR_na <- MUR_updated %>% 
-  slice(4186:4355)
-
-MUR_combi <- rbind(MUR_1,MUR_na, updated)
-
+MUR_test <- rbind(MUR,MUR_1) %>% 
+  arrange(date)
+  
 MUR_yrs_complete <- rbind(MUR_fill, MUR_updated) %>% 
   mutate(date = as.Date(date)) %>% 
   arrange(date)
@@ -191,13 +186,29 @@ upwelling <- UI_angle %>%
   mutate(exceedance = ifelse(ui.saws >= 1, TRUE, FALSE),
          t = as.Date(t))
 
+# Subsetting data from 2011-01-01 to 2016-12-31
+OISST <- OISST_fill %>% 
[…]

### 2021-05-11 box_plot.R

@@ -117,7 +117,7 @@ SST_clims <- rbind(OISST_upwell_clims,G1SST_upwell_clims,CMC_upwell_clims,MUR_up
 SST_anom <- SST_clims %>% 
   group_by(site, product, distance) %>% 
   mutate(anom = temp - seas) %>% 
-  filter(year(t) %in% 2011:2014) %>% 
+  #filter(year(t) %in% 2011:2014) %>% 
   filter(event_no > 0) %>% # Activate this line to only use dates during an upwelling event
   dplyr::select(site, t, distance, temp, anom, product) %>% 
   na.omit() %>% 
@@ -169,7 +169,7 @@ test <- combined_products %>%
 test1 <- test %>% 
  group_by(site,product,distance) %>% 
   summarise(mean_signals = mean(count),
-            sd_signals = sd(count)) %>% 
+              sd_signals = sd(count)) %>% 
   mutate(Distance = case_when(distance == "25000" ~ "25",
                             distance == "50000" ~ "50",
                             distance == "0" ~ "0",))

### 2021-05-12 ANOVA.R
update

@@ -698,7 +698,7 @@ summary(aov(intensity_cumulative ~ site + product + distance, data = lm_metrics_
 
 ### ANOVA suggested by AJ to compare the differences in number of signals detected between sites
 metric_ANOVA <- combined_products %>% 
-  filter(year(date_start) %in% 2011:2014) %>%  # only for the years 2011-2014 so 4 year period
+ # filter(year(date_start) %in% 2011:2014) %>%  # only for the years 2011-2014 so 4 year period
   ungroup() %>%
   group_by(product, site, distance) %>% 
   summarise(y = n()) %>% 
@@ -712,7 +712,7 @@ metric_ANOVA <- combined_products %>%
 
 summary(aov(count ~ site, data = metric_ANOVA))
 summary(aov(count ~ product + distance/site, data = metric_ANOVA))
-summary(aov(count ~ distance, data = metric_ANOVA))
+  summary(aov(count ~ distance, data = metric_ANOVA))
 
 ##
 # Plot the count of SST against temperature and see if changes occured

### 2021-05-15 box_plot.R
edits

@@ -111,6 +111,9 @@ load("Data_coast_angle/CMC_upwell_clims.RData")
 load("Data_coast_angle/MUR_upwell_clims.RData")
 load("Data_coast_angle/G1SST_upwell_clims.RData")
 
+MUR_upwell_clims <- MUR_upwell_clims %>% 
+  mutate(distance = as.numeric(as.character(distance)))
+
 SST_clims <- rbind(OISST_upwell_clims,G1SST_upwell_clims,CMC_upwell_clims,MUR_upwell_clims)
 
 # Determining the anomaly temperature for  SST products and SACTN
@@ -126,14 +129,18 @@ SST_anom <- SST_clims %>%
 # Create wide anomaly data.frame
 sst_anom_wide <- SST_anom %>% 
   pivot_wider(id_cols = c(site, product, t), names_from = distance, values_from = c(temp, anom))
+sst_anom_wide <- as.data.frame(sst_anom_wide)
+write_csv(sst_anom_wide, path = "sst_anom_wide.csv")
 
 # Doing the correlation using the easystat correlation package
 # https://rdrr.io/github/easystats/correlation/man/   correlation.html
 SST_corr <- sst_anom_wide %>% 
   group_by(site, product) %>% 
   correlation(method = "pearson") #%>% 
-  # summary()
 
+
+  # summary()
+save(sst_anom_wide, file = "sst_anom_wide.RData")
 # anom_spread<- SST_anom %>%
 #   spread(key = distance, value = temp)
 
@@ -141,18 +148,23 @@ SST_anom_spread <- pivot_wider(SST_anom, names_prefix = "dist_",
                                names_from = distance, values_from = temp)
 
 slope_calc <- function(df){
+  cor(df[5:6,], method="pearson")
+}
+
+slope_calc <- function(df){
   df %>% 
     do(mod1 = cor(.$dist_0, .$dist_25000, method = "pearson", use = "complete.obs"),
        mod2 = cor(.$dist_0, .$dist_50000, method = "pearson", use = "complete.obs")) %>%
     mutate(dist_0_vs_25_r = mod1[1],
-           dist_0_vs_50_r = mod2[1]) %>% 
-    dplyr::select(-mod1, -mod2) %>% 
+           dist_0_vs_50_r = mod2[1]) %>%
+    dplyr::select(-mod1, -mod2) %>%
     mutate_if(is.numeric, round, 2)
 }
 
[…]

### 2021-05-16 upwell_IDX.Rmd
update

@@
 # Calculate the upwelling event metrics
+SACTN_US <- SACTN_US %>% 
+  filter(year(date) %in% seq(2011, 2016, 1))
+
 SACTN_upwell_base <- SACTN_US %>% 
@@
   nest() %>% 
-  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1993-01-01", "2005-10-19"))) %>%
+  mutate(clim = purrr::map(data, ts2clm, pctile = 100, climatologyPeriod = c("2011-01-01", "2016-12-31"))) %>%
   dplyr::select(-data) %>% 
@@
   unnest()
-save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")
+
+number_SACTN <-  SACTN_upwell_events%>% 
+  filter(site == "Saldanha Bay")
+
+save(SACTN_upwell_events, file = "SACTN_upwell_events.RData")
 # save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")

### 2021-05-25 site_map.Rmd
Map Font change

@@
 load("~/Documents/HONOURSPROJECT-master/data/africa_coast.RData")
-site_pixels <- read_csv("site_pixels.csv")
+#site_pixels <- read_csv("site_pixels.csv")
 
@@
   annotate("text", label = "Cape Columbine", x = 16.0, y = -30.5,
-             size =5, angle = 45, colour = "black")+
+           family = "Arial", size =5, angle = 45, colour = "black")+
   annotate("text", label = "Cape Peninsula", x = 16.5, y = -33.9,
-             size =5, angle = 45, colour = "black")+
+             family = "Arial", size =5, angle = 45, colour = "black")+
   guides(fill = guide_colourbar()) +
@@
   theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Arial"),
-        axis.text.x = element_text(vjust = 1, family = "Arial", size = 22),
-        axis.text.y = element_text(hjust = 1, family = "Arial", size = 22),
+        axis.text.x = element_text(vjust = 1, family = "Arial", size = 12),
+        axis.text.y = element_text(hjust = 1, family = "Arial", size = 12),
         axis.title.x = element_text(vjust = 0, family = "Arial"),
@@
 
-ggsave("sa_4.tiff", units="cm", width=19.05, height=22.23, dpi=300)
+ggsave("Fig1.tiff", units="cm", width=19.05, height=22.23, dpi=600)
 ```
